### 2023-04-24 docs/manuscript.Rmd
added manuscript file and updated readme file

---
title             : "A Tutorial for Customized Power Simulation and Data Analysis using GLMM  - A Case Study for Human-AI Interaction Experiments"
shorttitle        : "Simulation Tutorial"

author: 
  - name          : "Timo K. Koch"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Florian Pargent"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "generalized linear mixed model, data simulation, sample size, power analysis"
wordcount         : "3000"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

As the integration of artificial intelligence (AI) into our daily lives continues to increase, it has become increasingly important to study how humans interact with these intelligent systems. This is particularly important as AI becomes more sophisticated and is used in decision-making processes that have significant impacts on individuals and society as a whole. Experimental psychology offers a valuable tool for studying human-AI interaction and understanding how people perceive and respond to these technologies. When conducting research in this field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable. Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.

# Methods

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Procedure

## Data analysis

We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2023-06-02 docs/manuscript.Rmd
pushed latest version of manuscript

@@
 title             : "A Tutorial for Customized Power Simulation and Data Analysis using GLMM  - A Case Study for Human-AI Interaction Experiments"
-shorttitle        : "Simulation Tutorial"
+shorttitle        : "GLMM Power Simulation Tutorial"
 
@@
   - name          : "Timo K. Koch"
-    affiliation   : "1"
+    affiliation   : "1, 2"
     corresponding : yes    # Define only one corresponding author
     address       : "Postal address"
-    email         : "my@email.com"
+    email         : "timo.koch@psy.lmu.de"
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
   - name          : "Florian Pargent"
-    affiliation   : "1,2"
+    affiliation   : "2"
+    role:
+      - "Writing - Review & Editing"
+      - "Supervision"
+  - name          : "Susanne Gaube"
+    affiliation   : "1"
     role:
@@
   - id            : "1"
-    institution   : "Wilhelm-Wundt-University"
+    institution   : "LMU Munich, Center for Leadership and People Management"
   - id            : "2"
-    institution   : "Konstanz Business School"
+    institution   : "LMU Munich, Department of Psychology"
 
@@
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
+  Human-AI Intercation experiments are becoming increasingly important.
+However, many studies are underpowered since researchers lack the tools to conduct power analyses and sample size simulations.
+In this work, we provide.
+By providing code in a case study, we equip scientists to simulate their own data and run analyses based on generalized linear mixed models (GLMM).
 
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
+keywords          : "generalized linear mixed model, data simulation, sample size, power analysis, data analysis"
 
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "generalized linear mixed model, data simulation, sample size, power analysis"
 wordcount         : "3000"
@@
 
+<!-- General intro, relevance of Human AI Interaction experiments -->
+
 As the integration of artificial intelligence (AI) into our daily lives continues to increase, it has become increasingly important to study how humans interact with these intelligent systems.
-This is particularly important as AI becomes more sophisticated and is used in decision-making processes that have significant impacts on individuals and society as a whole.
-Experimental psychology offers a valuable tool for studying human-AI interaction and understanding how people perceive and respond to these technologies.
+This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
+Therefore, thourough experiments are needed to study human-AI interaction and understand how people perceive and respond to these technologies.
 When conducting research in this field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable.
@@
 
+## LMMs and GLMMs
+
+<!-- GLMMs are often used for such research -->
+Linear Mixed Models (LMM) and Generalized Linear Mixed Models (GLMM) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
+
+Linear Mixed Models (LMM) are extensions of linear regression models that account for correlated data and hierarchical structures.
+They are used when the outcome variable is continuous and follow a normal distribution.
+LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
+
+Generalized Linear Mixed Models (GLMM) extend the LMM framework to accommodate non-normal and categorical outcome variables.
+They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
+
+Generalized Linear Mixed Models (GLMMs) are gaining increasing popularity in the field of Human-AI interaction research.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
+GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
+This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
+By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+
+<!-- Sample size planning is increasingly important! -->
+
+Power simulations play a crucial role in addressing the replication crisis and meeting the requirements set forth by journals and funding agencies in Human-AI interaction research.
+The replication crisis has underscored the problem of high false discovery rates (FDR) coupled with low-powered studies.
+Power simulations offer a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
+Moreover, many journals and funding agencies now mandate the inclusion of power simulations as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
+By incorporating power simulations into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of psychological science.
+
+## Customized power simulations are hard
+
+<!-- Power simulations are hard! (especially for GLMMs) -->
+Power simulations can be challenging, particularly when analyzing complex psychological study designs using (generalized) linear mixed models (GLMMs).
+While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of GLMMs in psychological research introduces greater complexity to power calculations.
+For GLMMs, simulations are often required, necessitating assumptions for both fixed and random effects.
+However, most tutorials on GLMMs primarily focus on the most common designs, limiting the guidance available for researchers facing more intricate scenarios.
+Moreover, existing tutorials often rely on heuristics for estimating random effect standard deviations and may incorporate meta-analytic results for both random and fixed effects, further adding to the challenge of conducting accurate power calculations for GLMMs.
+
+<!-- Research Gap & Summary -->
+When planning a Human-AI interaction experiment with complexities such as the use of Generalized Linear Mixed Models (GLMM), non-standard designs, and missing by design conditions, the process of power simulations becomes more challenging.
+Software packages designed for power simulations may not adequately address the complexities of the experiment, making it necessary to build data simulations tailored specifically to the study design.
+Unfortunately, there are no general recommendations that can be applied universally in such cases.
+Instead, researchers must acquire the necessary skills based on the specific application at hand.
+In this paper, we present a case study that serves as a practical demonstration of how to perform a bespoke data simulation for a concrete study.
+Through this case study, we aim to provide guidance and insights into the complexities involved in addressing complex design considerations and obtaining accurate power estimates.
+
 # Methods
 
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
+In this section, we describe the steps to conduct a data simulation and a power analysis. In doing so, we provide a concrete example in the form of a case study.
 
-## Procedure
+## The present case study
+<!-- The present example, describe the study idea and design  -->
+In the case study, we simulate the data for a Human-AI Interaction experiment where a health care AI is to be evaluated.
+Participants view head CT cases and evaluate if a bleeding is present.
+Also, there can be AI advice that is either correct or incorrect.
+This medical diagnosis can be correct or incorrect.
 
-## Data analysis
+In this case, it would be more challenging to recruit task expert as there is only a limited amount of such people.
+Non-experts would be easier to recruit.
+The open question is how many task experts and non-experts to recruit to achieve sufficient power.
+
+describe study design and sample size restrictions 
+- describe research question (i.e. the coefficients targeted by power simulation)
+
+<!-- basics of power simulation in GLMMs (with the concrete example in mind) -->
+
+Power simulations in generalized linear mixed models (GLMMs) are essential for estimating the statistical power of complex psychological study designs.
+The model equation for a GLMM combines fixed effects, random effects, and an appropriate link function to model the relationship between predictors and the outcome variable.
+The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
+The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial.
+Assumptions about the random effects include the assumption of normality and the covariance structure among the random effects.
+Additionally, assumptions about the error structure, such as independence or correlation, must be specified.
+Interpreting these assumptions entails understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
+
+The general model equation for a generalized linear mixed model (GLMM) can be expressed as follows:
+
+g(E(Y_ij)) = X_ij \* beta + Z_ij \* b_i + epsilon_ij
+
+In this equation:
+
+-   Y_ij represents the outcome variable for the i-th individual at the j-th level.
+-   E(Y_ij) is the expected value of Y_ij.
+-   g() is the link function that relates the linear predictor to the expected value of Y_ij.
+-   X_ij represents the design matrix of fixed effects, including the predictor variables and their corresponding coefficients (beta).
+-   Z_ij is the design matrix of random effects, representing the random effect variables and their associated random effects (b_i).
+-   epsilon_ij is the error term, assumed to follow a distribution specified by the chosen GLMM family and accounting for the variability not explained by fixed and random effects.
+
+The model equation captures the relationship between the predictors, the random effects, and the outcome variable through a combination of fixed effects (X_ij \* beta), random effects (Z_ij \* b_i), and the error term (epsilon_ij).
+The link function, g(), determines the relationship between the linear predictor and the expected value of the outcome variable based on the distributional assumptions of the GLMM.
+
+## The lme4 R package
+
+<!-- describe the lme4 package and how it works-->
+
+The lme4 R package is a state-of-the-art tool for fitting frequentist Generalized Linear Mixed Models (GLMMs).
+It provides extensive capabilities for modeling complex data structures, including models with normally distributed random intercepts and random slopes.
+The package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+
+In lme4, the parameterization used involves the concept of both the beta vector and the theta vector.
+When considering only random intercepts, the theta parameter represents the standard deviation of the random intercepts, providing information about the variability across the different levels of the random effect.
+This allows for a deeper understanding of the random intercepts' influence on the outcome variable.
+
+Regarding random slopes, the interpretation of theta becomes more complex.
+It encompasses the covariance structure between the random slopes and the random intercepts.
+The theta vector accounts for the variability and correlation between these different random effects, offering insights into the multilevel nature of the data.
+However, it is important to note that the tutorial discussed in this paper focuses solely on the simpler case of random intercepts, while acknowledging the more intricate implications of theta in models involving random slopes.
+
+## Data simulation
 
 We used `r cite_r("r-references.bib")` for all our analyses.
+This part is very well done in DeBruine et al paper!
+
+### Specify the data structure
+
+<!-- main part of the tutorial with concrete R-code displayed in the paper (Flo part!) -->
+
+- build the format of the predictor variables
+
+### Specify the parameters
+
+- find the appropriate lme4 formula (simple models first, more complex models later)
+    - discuss the specific missing by design complexity
+    - discuss different possible parametrizations of the fixed effects
+- simulate random values for the dependent variable so that testing model formulas is easier
+- determine the correct labels for model coefficients in the lme4 output
+- determine assumptions for fixed effects
+    - discuss useful strategies for determining meaningful assumptions and outline the repeated discussions with content partners
+- determine assumptions for random intercepts
+    - discuss useful strategies for determining meaningful assumptions and outline the repeated discussions with content partners
+
+### Simulate the data
+
+- simulate datasets (perhaps discuss parallelization in R)
+- test the correctness of the assumed coefficient values by comparing descriptive statistics of the simulated datasets with original assumptions
+
+### Interpreting the output
+
+### Estimate power
+
+- build function to compute the power for the coefficients/contrasts of interest
+    - estimate power for different sample sizes and number of stimuli
+    - excursus: estimate expected with of confidence intervals for different sample sizes and number of stimuli
 
@@
 
+<!-- How many participants / stimuli are needed for concrete example -->
+
+<!-- Results of power sensitivity analyses -->
+
 # Discussion
+
+<!-- Summarize key findings-->
+
+Human-AI interaction research requires meticulous planning and consideration of statistical power to ensure reliable and meaningful results.
+While heuristics and helper programs can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
+In this paper, we discuss the necessity of bespoke power simulations in human-AI interaction research, their implications for research design, and the importance of teaching these skills to Human-AI interaction researchers.
+Furthermore, we highlight the discrepancy between the perceived effort involved in performing bespoke power simulations and the actual effort required.
+Finally, we emphasize the value of bespoke power simulations as research contributions and their alignment with open science and preregistration practices.
+
+1.
+The Need for Bespoke Power Simulations in Human-AI Interaction Research: Human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or simple helper programs.
+Bespoke power simulations offer a solution by providing Human-AI interaction researchers with a tailored approach to estimating statistical power.
+These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
+
+2.
+Managing Simulations with Discrete Predictor Variables: Bespoke power simulations become more manageable when all predictor variables are discrete and fixed by the study design.
+This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need to simulate predictor values, which would introduce additional assumptions.
+By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
+
+3.
+Teaching Bespoke Power Simulation Skills to Human-AI Interaction Researchers: The ability to conduct bespoke power simulations is a valuable skill that should be taught to Human-AI interaction researchers.
+By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
+Equipping Human-AI interaction researchers with the knowledge and tools to perform bespoke power simulations empowers them to make informed decisions and enhance the rigor of their studies.
+
+4.
+Addressing the Mismatch in Effort Perception: There is often a significant disconnect between the perceived effort required to perform bespoke power simulations and the actual effort estimated by researchers and collaborators in Human-AI interaction research.
+Many content researchers and collaborators request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these simulations.
+It is crucial to raise awareness about the effort involved in bespoke power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
+
+5.
+Recognizing the Value of Bespoke Power Simulations: Bespoke power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
+They offer insights into the reliability and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
+The importance of bespoke power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports.
+
+6.
+Integration with Open Science and Preregistration Practices: Bespoke power simulations align well with the principles of open science and preregistration in Human-AI interaction research.
+When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
+By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of Human-AI interaction research.
+
+# Conclusion
+
+Bespoke power simulations play a critical role in human-AI interaction research, allowing researchers to tailor power estimation to the unique aspects of their experiments.
+The skills required to perform these simulations should be taught to Human-AI interaction researchers, fostering a deeper understanding of statistical power and enhancing research design.
+It is essential to bridge the gap between perceived and actual effort associated with bespoke power simulations and recognize their value as research contributions.
+By integrating bespoke power simulations with open science and preregistration practices, researchers can improve the robustness and transparency of their findings, advancing the field.
 

### 2023-07-17 docs/manuscript.Rmd
improved manuscript

@@
 ---
-title             : "A Tutorial for Customized Power Simulation and Data Analysis using GLMM  - A Case Study for Human-AI Interaction Experiments"
+title             : "Tutorial for Customized Power Simulations and Data Analyses for Human-AI Interaction Experiments using Generalized Linear Mixed Models"
 shorttitle        : "GLMM Power Simulation Tutorial"
@@
     institution   : "LMU Munich, Department of Psychology"
+  - id            : "3"
+    institution   : "University of St. Gallen, Institute of Behavioral Science & Technology"
+  - id            : "4"
+    institution   : "University College London"
 
@@
 abstract: |
-  Human-AI Intercation experiments are becoming increasingly important.
-However, many studies are underpowered since researchers lack the tools to conduct power analyses and sample size simulations.
-In this work, we provide.
-By providing code in a case study, we equip scientists to simulate their own data and run analyses based on generalized linear mixed models (GLMM).
+  Understanding how humans interact with articifical intelligence trough experiments becomes increasingly relevant.
+However, human-AI interaction researchers lack the appropriate tools to conduct power and data analyses for the required complex study designs.
+In this work, we provide a tutorial on how to run customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
+By providing code in a case study, we equip human-AI interaction researchers to simulate their own data and run analyses based on generalized linear mixed models (GLMM).
+We discuss the outlook.
 
-keywords          : "generalized linear mixed model, data simulation, sample size, power analysis, data analysis"
+keywords          : "generalized linear mixed model, data simulation, sample size, power analysis, human-AI interaction"
 
@@
 
-<!-- General intro, relevance of Human AI Interaction experiments -->
+<!-- General intro, relevance of Human AI Interaction experiments, problem of underpowered & inappropriately analyzed work -->
 
@@
 This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
-Therefore, thourough experiments are needed to study human-AI interaction and understand how people perceive and respond to these technologies.
+Therefore, thorough experiments are needed to study human-AI interaction and understand how people perceive and respond to these technologies.
 When conducting research in this field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable.
 Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
-
-## LMMs and GLMMs
-
-<!-- GLMMs are often used for such research -->
-Linear Mixed Models (LMM) and Generalized Linear Mixed Models (GLMM) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
-
-Linear Mixed Models (LMM) are extensions of linear regression models that account for correlated data and hierarchical structures.
-They are used when the outcome variable is continuous and follow a normal distribution.
-LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
-
-Generalized Linear Mixed Models (GLMM) extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
-
-Generalized Linear Mixed Models (GLMMs) are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
-GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
-By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
 
@@
 <!-- Power simulations are hard! (especially for GLMMs) -->
-Power simulations can be challenging, particularly when analyzing complex psychological study designs using (generalized) linear mixed models (GLMMs).
+Power simulations can be challenging, particularly when analyzing complex study designs using (generalized) linear mixed models (GLMMs).
 While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of GLMMs in psychological research introduces greater complexity to power calculations.
@@
 
+## LMMs and GLMMs
+
+<!-- GLMMs are often used for such research -->
+Linear Mixed Models (LMM) and Generalized Linear Mixed Models (GLMM) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
+
+Linear Mixed Models (LMM) are extensions of linear regression models that account for correlated data and hierarchical structures.
+They are used when the outcome variable is continuous and follow a normal distribution.
+LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
+
+Generalized Linear Mixed Models (GLMM) extend the LMM framework to accommodate non-normal and categorical outcome variables.
+They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
+
+Generalized Linear Mixed Models (GLMMs) are gaining increasing popularity in the field of Human-AI interaction research.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
+GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
+This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
+By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+
 The general model equation for a generalized linear mixed model (GLMM) can be expressed as follows:
@@
 <!-- describe the lme4 package and how it works-->
-
 The lme4 R package is a state-of-the-art tool for fitting frequentist Generalized Linear Mixed Models (GLMMs).
@@
 Furthermore, we highlight the discrepancy between the perceived effort involved in performing bespoke power simulations and the actual effort required.
-Finally, we emphasize the value of bespoke power simulations as research contributions and their alignment with open science and preregistration practices.
+Finally, we emphasize the value of power simulations as research contributions and their alignment with open science and preregistration practices.
 
 1.
-The Need for Bespoke Power Simulations in Human-AI Interaction Research: Human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or simple helper programs.
+The Need for Power Simulations in Human-AI Interaction Research: Human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or simple helper programs.
 Bespoke power simulations offer a solution by providing Human-AI interaction researchers with a tailored approach to estimating statistical power.
@@
 The skills required to perform these simulations should be taught to Human-AI interaction researchers, fostering a deeper understanding of statistical power and enhancing research design.
-It is essential to bridge the gap between perceived and actual effort associated with bespoke power simulations and recognize their value as research contributions.
+It is essential to bridge the gap between perceived and actual effort associated with power simulations and recognize their value as research contributions.
 By integrating bespoke power simulations with open science and preregistration practices, researchers can improve the robustness and transparency of their findings, advancing the field.

### 2023-07-18 r_code/powersim_minimal_example.R
add first draft of computing the powersim

@@ -1,0 +1,62 @@
+library(tidyverse)
+library(faux)
+library(lme4)
+library(multcomp)
+library(furrr)
+
+# code adapted from the vignette of the faux package:
+# https://debruine.github.io/faux/articles/sim_mixed.html
+
+# simulation function that simulates data, fits a model and computes a p-value
+sim <- function(n_subjects = 100, n_items = 50,
+  b_0 = 0, b_e = 0.2, b_a = 0.2, b_ea = 0.1, sd_u0s = 1, sd_u0i = 1, ...){
+  # simulate design
+  dat <- add_random(subject = n_subjects, item = n_items) %>%
+    add_between("subject", expert = c(1, 0), .prob = 0.2) %>%
+    add_between("subject", advice = c(1, 0)) %>%
+    
+    # add_between("subject", interpretable = c(1, 0)) %>%
+    # mutate(interpretable = replace(interpretable, advice == 1, NA)) %>%
+    
+    # add random effects
+    add_ranef("subject", u0s = sd_u0s) %>%
+    add_ranef("item", u0i = sd_u0s) %>%
+    # compute dependent variable
+    mutate(linpred = b_0 + b_e * expert + b_a * advice + 
+        b_ea * expert * advice +
+        u0i + u0s) %>%
+    mutate(y_prob = plogis(linpred)) %>%
+    mutate(y = rbinom(n = n(), size = 1, prob = y_prob))
+  # fit model
+  model <- glmer(y ~ expert * advice + (1|subject) + (1|item),
+    data = dat, family = "binomial")
+  # compute p-value for user defined hypothesis test
+  null_hypothesis <- "expert <= 0"
+  glht <- glht(model, linfct = null_hypothesis)
+  data.frame(pvalue = summary(glht)$test$pvalues)
+}
+
+# enable parallelization
+plan("multisession", workers = 4)
+# make simulation reproducible
+set.seed(2)
+
+# perform simulation study
+sim_design <- crossing(
+  rep = 1:200,
+  n_subjects = c(100, 200),
+  n_items = c(50, 100)
+) %>%
+  mutate(analysis = future_pmap(., sim, 
+    .options = furrr_options(seed = TRUE))) %>%
+  unnest(analysis)
+
+# calculate power for alpha = 0.05 and plot results
+sim_design %>%
+  group_by(n_subjects, n_items) %>% 
+  summarise(power = mean(pvalue < .05), 
[…]

### 2023-07-18 r_code/powersim_minimal_example.R
refactor functions to that data simulate function can be reused; compute power ci

@@ -2,17 +2,15 @@ library(tidyverse)
 library(faux)
 library(lme4)
 library(multcomp)
+library(binom)
 library(furrr)
 
-# code adapted from the vignette of the faux package:
-# https://debruine.github.io/faux/articles/sim_mixed.html
-
-# simulation function that simulates data, fits a model and computes a p-value
-sim <- function(n_subjects = 100, n_items = 50,
-  b_0 = 0, b_e = 0.2, b_a = 0.2, b_ea = 0.1, sd_u0s = 1, sd_u0i = 1, ...){
+# simulate a dataset based on population parameters
+simulate <- function(n_subjects = 100, n_items = 50,
+  b_0 = 0, b_e = 0.2, b_a = 0.2, sd_u0s = 1, sd_u0i = 1, ...){
   # simulate design
   dat <- add_random(subject = n_subjects, item = n_items) %>%
-    add_between("subject", expert = c(1, 0), .prob = 0.2) %>%
+    add_between("subject", expert = c(1, 0), .prob = 0.5) %>%
     add_between("subject", advice = c(1, 0)) %>%
     
     # add_between("subject", interpretable = c(1, 0)) %>%
@@ -22,18 +20,28 @@ sim <- function(n_subjects = 100, n_items = 50,
     add_ranef("subject", u0s = sd_u0s) %>%
     add_ranef("item", u0i = sd_u0s) %>%
     # compute dependent variable
-    mutate(linpred = b_0 + b_e * expert + b_a * advice + 
-        b_ea * expert * advice +
-        u0i + u0s) %>%
+    mutate(linpred = b_0 + b_e * expert + b_a * advice + u0i + u0s) %>%
     mutate(y_prob = plogis(linpred)) %>%
     mutate(y = rbinom(n = n(), size = 1, prob = y_prob))
+  dat
+}
+
+# fit a model to a data set and compute a pvalue
+analyse <- function(data, 
+  formula_chr = "y ~ 1 + expert + advice + (1|subject) + (1|item)", 
+  null_hypothesis = "expert <= 0", ...){
   # fit model
-  model <- glmer(y ~ expert * advice + (1|subject) + (1|item),
-    data = dat, family = "binomial")
+  model <- glmer(as.formula(formula_chr), data = data, family = "binomial")
   # compute p-value for user defined hypothesis test
-  null_hypothesis <- "expert <= 0"
[…]

### 2023-07-19 r_code/powersim_minimal_example.R
mini update to the model structure and data simulation

@@ -7,10 +7,10 @@ library(furrr)
 
 # simulate a dataset based on population parameters
 simulate <- function(n_subjects = 100, n_items = 50,
-  b_0 = 0, b_e = 0.2, b_a = 0.2, sd_u0s = 1, sd_u0i = 1, ...){
+  b_0 = 0, b_e = 0.2, b_a = 0.2, b_ea = 0.1, sd_u0s = 1, sd_u0i = 1, ...){
   # simulate design
   dat <- add_random(subject = n_subjects, item = n_items) %>%
-    add_between("subject", expert = c(1, 0), .prob = 0.5) %>%
+    add_between("subject", expert = c(1, 0), .prob = 0.8) %>%
     add_between("subject", advice = c(1, 0)) %>%
     
     # add_between("subject", interpretable = c(1, 0)) %>%
@@ -20,7 +20,8 @@ simulate <- function(n_subjects = 100, n_items = 50,
     add_ranef("subject", u0s = sd_u0s) %>%
     add_ranef("item", u0i = sd_u0s) %>%
     # compute dependent variable
-    mutate(linpred = b_0 + b_e * expert + b_a * advice + u0i + u0s) %>%
+    mutate(linpred = b_0 + b_e * expert + b_a * advice + b_ea * expert * advice +
+        u0i + u0s) %>%
     mutate(y_prob = plogis(linpred)) %>%
     mutate(y = rbinom(n = n(), size = 1, prob = y_prob))
   dat
@@ -28,7 +29,7 @@ simulate <- function(n_subjects = 100, n_items = 50,
 
 # fit a model to a data set and compute a pvalue
 analyse <- function(data, 
-  formula_chr = "y ~ 1 + expert + advice + (1|subject) + (1|item)", 
+  formula_chr = "y ~ 1 + expert * advice + (1|subject) + (1|item)", 
   null_hypothesis = "expert <= 0", ...){
   # fit model
   model <- glmer(as.formula(formula_chr), data = data, family = "binomial")

### 2023-07-31 docs/power_simulation_section.Rmd
add power_simulation_section.Rmd containing a first draft for the main section of the paper (including text and code); can be later transfered to manuscript.Rmd

---
title: "Power simulation section"
author: "Florian Pargent"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE)
library(tidyverse)
```

## Generalized linear mixed models

In a Generalized linear mixed model (GLMM), the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$.

$$
g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
$$
Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$.

$$
E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
$$

If the dependent variable $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability:

$$
P_{si} = P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
$$
In our case study, $P_{si}$ is the conditional probability that subject $s$ gives the correct response to item $i$.

In such a setting, we model the probability as

$$
P_{si} = inverse\_logit(\eta_{si})
$$

with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently

$$
logit(P_{si}) = \eta_{si}
$$
with the logit link $g(P_{si}) = logit(P_{si}) = ln \left( \frac{P_{si}}{1 - P_{si}} \right)$.

In our case study, the probability to give a correct response is assumed to depend on the predictors:

- $advice\_present_{si}$: whether subject $s$ was presented with algorithmic advice (1) or not (0) when asked to asses item $i$
- $advice\_correct_{si}$: whether this advice was correct (1) or not (0)
- $expert_s$: whether subject $s$ was a professional neurologist (1) or not (0)

and the random effects:

- $u_{0s}$: the deviation of suject $s$ from the average ability to solve an item with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
- $u_{0i}$: the deviation of item $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$

In total, we assume the model
$$
logit\left[P_{si}\right] = (\beta_0 + u_{0s} + u_{0i}) + \\
\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
$$

or equivalently

$$
P_{si} = inverse\_logit\left[(\beta_0 + u_{0s} + u_{0i}) + \\
\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}\right]
$$

with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.

In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).

To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions here.

## Data Simulation

The following R function simulates a full dataset structured according to our case study.
The faux package contains useful functions when simulating factorial designs including random effects.

```{r}
simulate <- function(n_subjects = 100, n_items = 50,
  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603, b_ea = 0.790, b_ec = -1.393,
  sd_u0s = 0.5, sd_u0i = 0.5, ...){
  require(dplyr)
  require(faux)
  # simulate design
  dat <- add_random(subject = n_subjects, item = n_items) %>%
    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
    mutate(advice_correct = if_else(advice_present == 1, rbinom(n(), 1, prob = 0.8), 0)) %>%
    # add random effects
    add_ranef("subject", u0s = sd_u0s) %>%
    add_ranef("item", u0i = sd_u0i) %>%
    # compute dependent variable
    mutate(linpred = b_0 + u0i + u0s +
        b_e * expert + b_a * advice_present + b_c * advice_correct +
        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
    mutate(y_prob = plogis(linpred)) %>%
    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
  dat
}
```

In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (`n_items` in total).
Thus the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
We add a between subject effect with the `add_between` command, simulating that about $25%$ of subjects are experts.
The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice and if advice is presented, the advice will be correct in about $80%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
Next we simulate one random effect each subject (`u0s`) and for each item (`u0i`).
As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with mean 0 and a standard deviation specified by the user.
With all design variables done, we are ready to simulate our model equation as outlined in equation X.
The linear predictor variable `linpred` ($\eta$ in the GLMM model equations), combines the predictor variables, random effects and model parameters as assumed by our model.
We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
In the final step, we simulate the binary dependent variable `y_bin` by -- for each trial -- drawing from a bernoulli distribution with success probability `y_prob`.

## Model fitting

In this section, we show how to fit a GLMM with lme4, interpret the model and test hypotheses derived from a research question.

We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyse this data.
```{r, message=FALSE}
set.seed(1)
dat <- simulate(n_subjects = 100, n_items = 50)
```
The lme4 package uses a special syntax for model specification.
Our assumed GLMM is represented by the formula:
```{r, message=FALSE}
library(lme4)
f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
  expert:advice_present + expert:advice_correct +
  (1|subject) + (1|item)
```
The first two lines looks similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`).
The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
The complete set of rules for the syntax are outlined in (CITE LME4 PAPER) and in the documentation of the lme4 package.

In lme4, a GLMM is fitted with the `glmer` function.
By setting `family =  "binomial"` we request a binomial GLMM, appropriate for our binary dependent variable `y_bin` (the binomial GLMM used the canonical logit link by default).
```{r}
fit <- glmer(f, data = dat, family = "binomial")
```

## Model interpretation

We can inspect the estimated model parameters with the `summary` command:
```{r}
summary(fit)
```
The output shows the estimates for all model parameters:
the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.

Unfortunately, the model parameters in a binomial GLMM are hard to interpret, because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
The most simple interpretation works by imagining a subject with average ability ($u0s = 0$) responding to an item with average difficulty ($u0i = 0$).
Then the model implied probability that such a person solves such an item is given by:
$$
P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = inverse\_logit\left[\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}\right]
$$
In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
The other extreme case would be the probability that a non-expert with average ability solves and item with average difficulty when presented without any advice:

$$
P(Y=1|expert = 0, advice\_present = 0, advice\_correct = 0, u_{0s} = 0, u_{0i} = 0) = inverse\_logit\left[\beta_0\right]
$$

Due to this complicated relationship, many experts argue not to focus too much on interpreting single model parameters when working with GLMMs.
Instead, it can be more intuitive to consider the implied predicted distribution of the dependent variable for each experimental conditions across all subjects and items.

With the marginaleffects package, we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition:

```{r, message=FALSE}
library(marginaleffects)
plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"),
  type = "response") + ylim(c(0.3, 1))
```

## Hypothesis testing

However, we need to think about the model parameters again when wanting to test hypotheses which we have theoretically derived from some research question.
Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.

The `Fixed effects` table in Figure X also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$.
However, for many research questions of interest, we are not interested in these two-sided tests referring to only one parameter.

For our case study, imagine the following combined hypothesis:
*We expect that for both experts and non-experts, correct advice leads to a higher probability to solve an item compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability to solve an item compared to no advice presented.*

This combined hypothesis leads to the following four separate null hypotheses to be tested:

$$
H_{01}: \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0 \\
H_{02}: \beta_{a} + \beta_{c} \leq 0 \\
H_{03}: \beta_{a} + \beta_{ea} \geq 0 \\
H_{04}: \beta_{a} \geq 0
$$
We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$:
The first null hypothesis states that *an expert responding to an item while presented with correct advice has lower or equal probability to solve the item compared to the same expert facing the same item without any advice.*
This implies the following inequality for each subject $s$ and item $i$

$$
inverse\_logit\left[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}\right] \leq inverse\_logit\left[(\beta_0 + u_{0s} + u_{0i}) + \beta_e\right]
$$

which simplifies to $\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0$.

We can specify and test hypotheses like these with the multcomp package as follows:
```{r, message=FALSE}
library(multcomp)
null_hypotheses <- c(
  "advice_present + advice_correct + expert:advice_present + expert:advice_correct <= 0",
  "advice_present + advice_correct <= 0",
  "-1 * (advice_present + expert:advice_present) <= 0",
  "-1 * (advice_present) <= 0")
glht <- glht(fit, linfct = null_hypotheses)
summary(glht, test = univariate())$test$pvalues
```
Because all hypotheses tested simultaneously with the `glht` function must have the same direction, we flip the sign of inequalities three and four by multiplying them with $-1$.
The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously.
However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected.
In such cases, the error probabilities do not accumulate and we would waste power when correcting for multiple testing.
Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command.
With a standard significance level of $\alpha = 0.05$, we would reject all four null hypotheses and therefore also reject the combined null hypothesis for this simulated dataset.

## Specification of plausible parameter values

When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where the numbers came from.
All parameter values have been determined in repeated exchanges with our affiliated domain experts and we here outline a few strategies on how to determine plausible parameter values.

We have already seen in our discussion of model interpretation, how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (for an average subject and an average item).

Table X shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model:

| _Experimental condition_ | _$P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$_ | _Implied equation_ |
|---|:---:|---|
| no advice, no expert | 0.70 | $logit(0.70) = \beta_0$ |
| no advice, expert | 0.90 | $logit(0.90) = \beta_0 + \beta_e$ |
| false advice, no expert | 0.40 | $logit(0.40) = \beta_0 + \beta_a$ |
| false advice, expert | 0.85 | $logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}$ |
| correct advice, no expert | 0.90 | $logit(0.90) = \beta_0 + \beta_a + \beta_c$ |
| correct advice, expert | 0.95 | $logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}$ |

This table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:

```{r}
b_0 <- qlogis(0.7)
b_e <- qlogis(0.9) - b_0
b_a <- qlogis(0.4) - b_0
b_ea <- qlogis(0.85) - b_0 - b_e - b_a
b_c <- qlogis(0.9) - b_0 - b_a
b_ec <- qlogis(0.95) - b_0 - b_e - b_a - b_c - b_ea
c(b_0 = b_0, b_e = b_e, b_a = b_a, b_c = b_c, b_ea = b_ea, b_ec = b_ec)
```

It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g.
$$P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = inverse\_logit\left[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\right]$$
```{r}
plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec)
```

Although the derivations above are straightforward, it is important not to misinterpret their implications:
In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
For example, we determined the $\beta$ parameters in a way that correspond to a desired probability, that an expert with average ability solves an item with average difficulty when presented with a correct advice.
However even if the model were true, we would not observe this probability if we estimated the probability in a group of expert responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
This implies that one must be careful when specifying parameter values based on previous studies or pilot data.

The well known inequality of conditional and marginal effects in GLMMs makes their interpretation more difficult, however, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
When parameter values have selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.

Earlier, we have already encountered one way to visualize the implied marginal distributions:
We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory.
A more sophisticated strategy is to simulate a large dataset and directly compute the averages, contrasts and distributions we are interested in.

```{r}
library(tidyverse)
library(ggdist)
dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.5, sd_u0i = 0.5)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
  ggplot(aes(x = y_prob, y = condition)) +
  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
```

Figure X shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
We can see that indeed, the average probabilities (block dots) differ from the probabilities of average subjects and items considered in the previous section.
This difference increases with the variability of the random effects.

In fact, up to this point we have not talked about plausible values for the standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$).
Plots like the one above are a useful tool to decide whether the specified standard deviations are reasonable, by comparing the ranges and overlap between conditions to domain knowledge.

In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$).
This gives us a better way to see the variability between persons.

```{r, echo=FALSE}
dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0i = 0.01)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
  ggplot(aes(x = y_prob, y = condition)) +
  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
```

In the next plot, we have set the subject standard deviation to almost zero ($\sigma_S = 0.01$.
This gives us a better way to see the variability between items.

```{r, echo=FALSE}
dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
  ggplot(aes(x = y_prob, y = condition)) +
  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
```

The final plot demonstrates, that these plots are also useful to spot standard deviations specified to high.
For example, if we set $\sigma_S = 3$ and $\sigma_I = 3$ this implies that in each experimental condition, the probabilities that a subject solves an item are always close to either 0 or 1, which is not a plausible assumption.

```{r, echo=FALSE}
dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 3, sd_u0i = 3)
dat %>% 
  mutate(condition = fct_cross(
    factor(expert), factor(advice_present), factor(advice_correct))) %>%
  mutate(condition = fct_recode(condition,
    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
  ggplot(aes(x = y_prob, y = condition)) +
  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
```

## Power simulation

With all these considerations out of the way, we are finally ready to perform a power simulation.
Wrapping the `simulate` function already introduced earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
```{r}
sim_and_analyse <- function(formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + expert:advice_present + expert:advice_correct + (1|subject) + (1|item)", 
  null_hypotheses = c(
    "advice_present + advice_correct + expert:advice_present + expert:advice_correct <= 0",
    "advice_present + advice_correct <= 0",
    "-1 * (advice_present + expert:advice_present) <= 0",
    "-1 * (advice_present) <= 0"), ...){
  require(lme4)
  require(multcomp)
  # simulate data
  dat <- simulate(...)
  # fit model
  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
  # compute p-values
  glht <- glht(model, linfct = null_hypotheses)
  pvalues <- summary(glht, test = univariate())$test$pvalues
  setNames(pvalues, paste0("p_H0", 1:length(null_hypotheses)))
}
```

Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit model for different parameter combinations or sample sizes.
Here, we use the future and furr packages to perform computations in parallel.
First, we enable parallelization and specify how many parallel cores of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.

```{r, message=FALSE}
library(future)
plan("multisession", workers = 4)
set.seed(2)
```

The next code chunk, specifies a simulation design with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
```{r}
library(furrr)
sim_design <- crossing(
  rep = 1:250,
  n_subjects = c(100, 150, 200),
  n_items = c(10, 30, 50)
) %>%
  mutate(pvalues = future_pmap(., sim_and_analyse, 
    .options = furrr_options(seed = TRUE))) %>%
  unnest_wider(pvalues)
```
The result of the computation is a data.frame that contains the p-values of all tested hypotheses for each simulated dataset.

For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations, in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code (heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>).

```{r}
library(binom)
alpha <- 0.05
power <- sim_design %>%
  group_by(n_subjects, n_items) %>% 
  summarise(power = mean(p_H01 < alpha & p_H02 < alpha & p_H03 < alpha & p_H04 < alpha), 
    n_sig = sum(p_H01 < alpha & p_H02 < alpha & p_H03 < alpha & p_H04 < alpha),
    n = n(),
    ci.lwr = binom.confint(n_sig, n, method = "wilson")$lower,
    ci.upr = binom.confint(n_sig, n, method = "wilson")$upper, 
    .groups = "drop")
power %>%
  mutate(across(c(n_subjects, n_items), factor)) %>%
  ggplot(aes(n_subjects, n_items, fill = power)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", power, ci.lwr, ci.upr)), 
    color = "white", size = 6) +
  scale_fill_viridis_c(limits = c(0, 1))
```

### 2023-08-03 r_code/powersim_minimal_example.R
delete outdated powersim_minimal_example.R

@@ -1,72 +1,0 @@
-library(tidyverse)
-library(faux)
-library(lme4)
-library(multcomp)
-library(binom)
-library(furrr)
-
-# simulate a dataset based on population parameters
-simulate <- function(n_subjects = 100, n_items = 50,
-  b_0 = 0, b_e = 0.2, b_c = 0.2, b_ec = 0.1, sd_u0s = 1, sd_u0i = 1, ...){
-  # simulate design
-  dat <- add_random(subject = n_subjects, item = n_items) %>%
-    add_between("subject", expert = c(1, 0), .prob = 0.8) %>%
-    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
-    mutate(advice_correct = if_else(advice_present == 1, rbinom(n(), 1, prob = 0.8), 0)) %>%
-    # add random effects
-    add_ranef("subject", u0s = sd_u0s) %>%
-    add_ranef("item", u0i = sd_u0s) %>%
-    # compute dependent variable
-    mutate(linpred = b_0 + b_e * expert + b_c * advice_correct + 
-        b_ec * expert * advice_correct + u0i + u0s) %>%
-    mutate(y_prob = plogis(linpred)) %>%
-    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
-  dat
-}
-
-# fit a model to a data set and compute a pvalue
-analyse <- function(data, 
-  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + expert:advice_correct + (1|subject) + (1|item)", 
-  null_hypothesis = "expert <= 0", ...){
-  # fit model
-  model <- glmer(as.formula(formula_chr), data = data, family = "binomial")
-  # compute p-value for user defined hypothesis test
-  glht <- glht(model, linfct = null_hypothesis)
-  pvalue <- summary(glht)$test$pvalues
-  as.numeric(pvalue)
-}
-
-# helper function for power simulation
-sim_helper <- function(...){
-  dat <- simulate(...)
-  analyse(dat, ...)
-}
-
-# enable parallelization
-plan("multisession", workers = 4)
-# make simulation reproducible
-set.seed(2)
-
-# perform simulation study
-sim_design <- crossing(
-  rep = 1:100,
-  n_subjects = c(100, 200),
-  n_items = c(10, 100)
-) %>%
-  mutate(pvalue = future_pmap(., sim_helper, 
[…]

### 2023-08-03 docs/manuscript.Rmd
delete power_simulation_section.Rmd, move content to manuscript; fix equations not working in pdf

@@
 In this work, we provide a tutorial on how to run customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
-By providing code in a case study, we equip human-AI interaction researchers to simulate their own data and run analyses based on generalized linear mixed models (GLMM).
+By providing code in a case study, we equip human-AI interaction researchers to simulate their own data and run analyses based on GLMMs.
 We discuss the outlook.
@@
 
-wordcount         : "3000"
+wordcount         : "5339"
 
@@
 
-floatsintext      : no
+floatsintext      : yes
 linenumbers       : yes
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output: 
+  papaja::apa6_pdf:
+    keep_tex: TRUE
 ---
@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = TRUE)
 ```
@@
 
+## LMMs, GLMMs and the lme4 package in R
+
+<!-- GLMMs are often used for such research -->
+Linear Mixed Models (LMM) and Generalized Linear Mixed Models (GLMM) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
+
+Linear Mixed Models (LMM) are extensions of linear regression models that account for correlated data and hierarchical structures.
+They are used when the outcome variable is continuous and follow a normal distribution.
+LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
+
+Generalized Linear Mixed Models (GLMM) extend the LMM framework to accommodate non-normal and categorical outcome variables.
+They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
+
+Generalized Linear Mixed Models (GLMMs) are gaining increasing popularity in the field of Human-AI interaction research.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
+GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
+This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
+By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+
+<!-- describe the lme4 package and how it works-->
+The lme4 R package is a state-of-the-art tool for fitting frequentist Generalized Linear Mixed Models (GLMMs).
+It provides extensive capabilities for modeling complex data structures, including models with normally distributed random intercepts and random slopes.
+The package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+
+In lme4, the parameterization used involves the concept of both the beta vector and the theta vector.
+When considering only random intercepts, the theta parameter represents the standard deviation of the random intercepts, providing information about the variability across the different levels of the random effect.
+This allows for a deeper understanding of the random intercepts' influence on the outcome variable.
+
+Regarding random slopes, the interpretation of theta becomes more complex.
+It encompasses the covariance structure between the random slopes and the random intercepts.
+The theta vector accounts for the variability and correlation between these different random effects, offering insights into the multilevel nature of the data.
+However, it is important to note that the tutorial discussed in this paper focuses solely on the simpler case of random intercepts, while acknowledging the more intricate implications of theta in models involving random slopes.
+
 # Methods
@@
 
+We used `r cite_r("r-references.bib")` for all our analyses.
+
 ## The present case study
+
 <!-- The present example, describe the study idea and design  -->
@@
 
-## LMMs and GLMMs
+## TODO
 
-<!-- GLMMs are often used for such research -->
-Linear Mixed Models (LMM) and Generalized Linear Mixed Models (GLMM) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
+points not discussed yet:
 
-Linear Mixed Models (LMM) are extensions of linear regression models that account for correlated data and hierarchical structures.
-They are used when the outcome variable is continuous and follow a normal distribution.
-LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
-
-Generalized Linear Mixed Models (GLMM) extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
-
-Generalized Linear Mixed Models (GLMMs) are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
-GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
-By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
-
-The general model equation for a generalized linear mixed model (GLMM) can be expressed as follows:
-
-g(E(Y_ij)) = X_ij \* beta + Z_ij \* b_i + epsilon_ij
-
-In this equation:
-
--   Y_ij represents the outcome variable for the i-th individual at the j-th level.
--   E(Y_ij) is the expected value of Y_ij.
--   g() is the link function that relates the linear predictor to the expected value of Y_ij.
--   X_ij represents the design matrix of fixed effects, including the predictor variables and their corresponding coefficients (beta).
--   Z_ij is the design matrix of random effects, representing the random effect variables and their associated random effects (b_i).
--   epsilon_ij is the error term, assumed to follow a distribution specified by the chosen GLMM family and accounting for the variability not explained by fixed and random effects.
-
-The model equation captures the relationship between the predictors, the random effects, and the outcome variable through a combination of fixed effects (X_ij \* beta), random effects (Z_ij \* b_i), and the error term (epsilon_ij).
-The link function, g(), determines the relationship between the linear predictor and the expected value of the outcome variable based on the distributional assumptions of the GLMM.
-
-## The lme4 R package
-
-<!-- describe the lme4 package and how it works-->
-The lme4 R package is a state-of-the-art tool for fitting frequentist Generalized Linear Mixed Models (GLMMs).
-It provides extensive capabilities for modeling complex data structures, including models with normally distributed random intercepts and random slopes.
-The package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
-
-In lme4, the parameterization used involves the concept of both the beta vector and the theta vector.
-When considering only random intercepts, the theta parameter represents the standard deviation of the random intercepts, providing information about the variability across the different levels of the random effect.
-This allows for a deeper understanding of the random intercepts' influence on the outcome variable.
-
-Regarding random slopes, the interpretation of theta becomes more complex.
-It encompasses the covariance structure between the random slopes and the random intercepts.
-The theta vector accounts for the variability and correlation between these different random effects, offering insights into the multilevel nature of the data.
-However, it is important to note that the tutorial discussed in this paper focuses solely on the simpler case of random intercepts, while acknowledging the more intricate implications of theta in models involving random slopes.
-
-## Data simulation
-
-We used `r cite_r("r-references.bib")` for all our analyses.
-This part is very well done in DeBruine et al paper!
-
-### Specify the data structure
-
-<!-- main part of the tutorial with concrete R-code displayed in the paper (Flo part!) -->
-
-- build the format of the predictor variables
-
-### Specify the parameters
-
-- find the appropriate lme4 formula (simple models first, more complex models later)
 - discuss the specific missing by design complexity
 - discuss different possible parametrizations of the fixed effects
-- simulate random values for the dependent variable so that testing model formulas is easier
+- find the appropriate lme4 formula (simple models first, more complex models later)
 - determine the correct labels for model coefficients in the lme4 output
-- determine assumptions for fixed effects
-    - discuss useful strategies for determining meaningful assumptions and outline the repeated discussions with content partners
-- determine assumptions for random intercepts
-    - discuss useful strategies for determining meaningful assumptions and outline the repeated discussions with content partners
-
-### Simulate the data
-
-- simulate datasets (perhaps discuss parallelization in R)
-- test the correctness of the assumed coefficient values by comparing descriptive statistics of the simulated datasets with original assumptions
-
-### Interpreting the output
-
-### Estimate power
-
-- build function to compute the power for the coefficients/contrasts of interest
-    - estimate power for different sample sizes and number of stimuli
 - excursus: estimate expected with of confidence intervals for different sample sizes and number of stimuli
 
+## Generalized linear mixed models
+
+In a Generalized linear mixed model (GLMM), the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$.
+$$
+g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
+$$
+Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$.
+$$
+E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
+$$
+If the dependent variable $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability:
+$$
+P_{si} = P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
+$$
+In our case study, $P_{si}$ is the conditional probability that subject $s$ gives the correct response to item $i$.
+
+In such a setting, we model the probability as
+$$
+P_{si} = inverse\_logit(\eta_{si})
+$$
+with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
+$$
+logit(P_{si}) = \eta_{si}
+$$
+with the logit link $g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})$.
+
+In our case study, the probability to give a correct response is assumed to depend on the predictors:
+
+- $advice\_present_{si}$: whether subject $s$ was presented with algorithmic advice (1) or not (0) when asked to asses item $i$
+- $advice\_correct_{si}$: whether this advice was correct (1) or not (0)
+- $expert_s$: whether subject $s$ was a professional neurologist (1) or not (0)
+
+and the random effects:
+
+- $u_{0s}$: the deviation of suject $s$ from the average ability to solve an item with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
+- $u_{0i}$: the deviation of item $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$
+
+In total, we assume the model
+$$
+\begin{aligned}
+logit[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
+&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
+\end{aligned}
+$$
+or equivalently
+$$
+\begin{aligned}
+P_{si} = inverse\_logit[&(\beta_0 + u_{0s} + u_{0i}) + \\
+&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+\end{aligned}
+$$
+with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
+
+In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
+
+To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions here.
+
+## Data Simulation
+
+The following R function simulates a full dataset structured according to our case study.
+The faux package contains useful functions when simulating factorial designs including random effects.
+
+```{r}
+simulate <- function(n_subjects = 100, n_items = 50,
+  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
+  b_ea = 0.790, b_ec = -1.393,
+  sd_u0s = 0.5, sd_u0i = 0.5, ...){
+  require(dplyr)
+  require(faux)
+  # simulate design
+  dat <- add_random(subject = n_subjects, item = n_items) %>%
+    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
+    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
+    mutate(advice_correct = if_else(advice_present == 1, 
+                                    rbinom(n(), 1, prob = 0.8), 0)) %>%
+    # add random effects
+    add_ranef("subject", u0s = sd_u0s) %>%
+    add_ranef("item", u0i = sd_u0i) %>%
+    # compute dependent variable
+    mutate(linpred = b_0 + u0i + u0s +
+        b_e * expert + b_a * advice_present + b_c * advice_correct +
+        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
+    mutate(y_prob = plogis(linpred)) %>%
+    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
+  dat
+}
+```
+
+In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (`n_items` in total).
+Thus the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
+We add a between subject effect with the `add_between` command, simulating that about $25%$ of subjects are experts.
+The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice and if advice is presented, the advice will be correct in about $80%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
+Next we simulate one random effect each subject (`u0s`) and for each item (`u0i`).
+As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with mean 0 and a standard deviation specified by the user.
+With all design variables done, we are ready to simulate our model equation as outlined in equation X.
+The linear predictor variable `linpred` ($\eta$ in the GLMM model equations), combines the predictor variables, random effects and model parameters as assumed by our model.
+We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
+In the final step, we simulate the binary dependent variable `y_bin` by -- for each trial -- drawing from a bernoulli distribution with success probability `y_prob`.
+
+## Model fitting
+
+In this section, we show how to fit a GLMM with lme4, interpret the model and test hypotheses derived from a research question.
+
+We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
+However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyse this data.
+
+```{r, message=FALSE}
+library(tidyverse)
+set.seed(1)
+dat <- simulate(n_subjects = 100, n_items = 50)
+```
+
+The lme4 package uses a special syntax for model specification.
+Our assumed GLMM is represented by the formula:
+
+```{r, message=FALSE}
+library(lme4)
+f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
+  expert:advice_present + expert:advice_correct +
+  (1|subject) + (1|item)
+```
+
+The first two lines looks similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`).
+The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
+The complete set of rules for the syntax are outlined in (CITE LME4 PAPER) and in the documentation of the lme4 package.
+
+In lme4, a GLMM is fitted with the `glmer` function.
+By setting `family =  "binomial"` we request a binomial GLMM, appropriate for our binary dependent variable `y_bin` (the binomial GLMM used the canonical logit link by default).
+
+```{r}
+fit <- glmer(f, data = dat, family = "binomial")
+```
+
+## Model interpretation
+
+We can inspect the estimated model parameters with the `summary` command:
+
+```{r}
+summary(fit)
+```
+
+The output shows the estimates for all model parameters:
+the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
+
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret, because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+The most simple interpretation works by imagining a subject with average ability ($u0s = 0$) responding to an item with average difficulty ($u0i = 0$).
+Then the model implied probability that such a person solves such an item is given by:
+$$
+\begin{aligned}
+P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
+= inverse\_logit[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+\end{aligned}
+$$
+
+In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
+In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
+The other extreme case would be the probability that a non-expert with average ability solves and item with average difficulty when presented without any advice:
+$$
+\begin{aligned}
+P(Y=1|expert = 0, advice\_present = 0, advice\_correct = 0, u_{0s} = 0, u_{0i} = 0) = \\
+= inverse\_logit[\beta_0]
+\end{aligned}
+$$
+
+Due to this complicated relationship, many experts argue not to focus too much on interpreting single model parameters when working with GLMMs.
+Instead, it can be more intuitive to consider the implied predicted distribution of the dependent variable for each experimental conditions across all subjects and items.
+
+With the marginaleffects package, we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition:
+
+```{r, message=FALSE}
+library(marginaleffects)
+plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"),
+  type = "response") + ylim(c(0.3, 1))
+```
+
+## Hypothesis testing
+
+However, we need to think about the model parameters again when wanting to test hypotheses which we have theoretically derived from some research question.
+Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.
+
+The `Fixed effects` table in Figure X also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$.
+However, for many research questions of interest, we are not interested in these two-sided tests referring to only one parameter.
+
+For our case study, imagine the following combined hypothesis:
+*We expect that for both experts and non-experts, correct advice leads to a higher probability to solve an item compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability to solve an item compared to no advice presented.*
+
+This combined hypothesis leads to the following four separate null hypotheses to be tested:
+$$
+\begin{aligned}
+H_{01}:\ &\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0 \\
+H_{02}:\ &\beta_{a} + \beta_{c} \leq 0 \\
+H_{03}:\ &\beta_{a} + \beta_{ea} \geq 0 \\
+H_{04}:\ &\beta_{a} \geq 0
+\end{aligned}
+$$
+
+We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$:
+The first null hypothesis states that *an expert responding to an item while presented with correct advice has lower or equal probability to solve the item compared to the same expert facing the same item without any advice.*
+This implies the following inequality for each subject $s$ and item $i$
+$$
+inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}] \leq inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e]
+$$
+which simplifies to $\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0$.
+
+We can specify and test hypotheses like these with the multcomp package as follows:
+
+```{r, message=FALSE}
+library(multcomp)
+null_hypotheses <- c(
+  "advice_present + advice_correct + expert:advice_present + 
+   expert:advice_correct <= 0",
+  "advice_present + advice_correct <= 0",
+  "-1 * (advice_present + expert:advice_present) <= 0",
+  "-1 * (advice_present) <= 0")
+glht <- glht(fit, linfct = null_hypotheses)
+summary(glht, test = univariate())$test$pvalues
+```
+
+Because all hypotheses tested simultaneously with the `glht` function must have the same direction, we flip the sign of inequalities three and four by multiplying them with $-1$.
+The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously.
+However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected.
+In such cases, the error probabilities do not accumulate and we would waste power when correcting for multiple testing.
+Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command.
+With a standard significance level of $\alpha = 0.05$, we would reject all four null hypotheses and therefore also reject the combined null hypothesis for this simulated dataset.
+
+## Specification of plausible parameter values
+
+When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where the numbers came from.
+All parameter values have been determined in repeated exchanges with our affiliated domain experts and we here outline a few strategies on how to determine plausible parameter values.
+
+We have already seen in our discussion of model interpretation, how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
+We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (for an average subject and an average item).
+
+Table X shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model:
+
+| _Experimental condition_ | _$P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$_ | _Implied equation_ |
+|---|:---:|---|
+| no advice, no expert | 0.70 | $logit(0.70) = \beta_0$ |
+| no advice, expert | 0.90 | $logit(0.90) = \beta_0 + \beta_e$ |
+| false advice, no expert | 0.40 | $logit(0.40) = \beta_0 + \beta_a$ |
+| false advice, expert | 0.85 | $logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}$ |
+| correct advice, no expert | 0.90 | $logit(0.90) = \beta_0 + \beta_a + \beta_c$ |
+| correct advice, expert | 0.95 | $logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}$ |
+
+This table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:
+
+```{r}
+b_0 <- qlogis(0.7)
+b_e <- qlogis(0.9) - b_0
+b_a <- qlogis(0.4) - b_0
+b_ea <- qlogis(0.85) - b_0 - b_e - b_a
+b_c <- qlogis(0.9) - b_0 - b_a
+b_ec <- qlogis(0.95) - b_0 - b_e - b_a - b_c - b_ea
+c(b_0 = b_0, b_e = b_e, b_a = b_a, b_c = b_c, b_ea = b_ea, b_ec = b_ec)
+```
+
+It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g.
+$$
+\begin{aligned}
+P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
+= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
+\end{aligned}
+$$
+
+```{r}
+plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec)
+```
+
+Although the derivations above are straightforward, it is important not to misinterpret their implications:
+In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
+For example, we determined the $\beta$ parameters in a way that correspond to a desired probability, that an expert with average ability solves an item with average difficulty when presented with a correct advice.
+However even if the model were true, we would not observe this probability if we estimated the probability in a group of expert responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
+This implies that one must be careful when specifying parameter values based on previous studies or pilot data.
+
+The well known inequality of conditional and marginal effects in GLMMs makes their interpretation more difficult, however, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
+When parameter values have selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
+Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
+
+Earlier, we have already encountered one way to visualize the implied marginal distributions:
+We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
+However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory.
+A more sophisticated strategy is to simulate a large dataset and directly compute the averages, contrasts and distributions we are interested in.
+
+```{r}
+library(tidyverse)
+library(ggdist)
+dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.5, sd_u0i = 0.5)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  ggplot(aes(x = y_prob, y = condition)) +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
+Figure X shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
+We can see that indeed, the average probabilities (block dots) differ from the probabilities of average subjects and items considered in the previous section.
+This difference increases with the variability of the random effects.
+
+In fact, up to this point we have not talked about plausible values for the standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$).
+Plots like the one above are a useful tool to decide whether the specified standard deviations are reasonable, by comparing the ranges and overlap between conditions to domain knowledge.
+
+In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$).
+This gives us a better way to see the variability between persons.
+
+```{r, echo=FALSE}
+dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0i = 0.01)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  ggplot(aes(x = y_prob, y = condition)) +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
+In the next plot, we have set the subject standard deviation to almost zero ($\sigma_S = 0.01$.
+This gives us a better way to see the variability between items.
+
+```{r, echo=FALSE}
+dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  ggplot(aes(x = y_prob, y = condition)) +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
+The final plot demonstrates, that these plots are also useful to spot standard deviations specified to high.
+For example, if we set $\sigma_S = 3$ and $\sigma_I = 3$ this implies that in each experimental condition, the probabilities that a subject solves an item are always close to either 0 or 1, which is not a plausible assumption.
+
+```{r, echo=FALSE}
+dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 3, sd_u0i = 3)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  ggplot(aes(x = y_prob, y = condition)) +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
 # Results
+
+## Power simulation
+
+With all these considerations out of the way, we are finally ready to perform a power simulation.
+Wrapping the `simulate` function already introduced earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
+
+```{r}
+sim_and_analyse <- function(
+  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + 
+    expert:advice_present + expert:advice_correct + (1|subject) + (1|item)", 
+  null_hypotheses = c("advice_present + advice_correct + 
+    expert:advice_present + expert:advice_correct <= 0",
+    "advice_present + advice_correct <= 0",
+    "-1 * (advice_present + expert:advice_present) <= 0",
+    "-1 * (advice_present) <= 0"), ...){
+  require(lme4)
+  require(multcomp)
+  # simulate data
+  dat <- simulate(...)
+  # fit model
+  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
+  # compute p-values
+  glht <- glht(model, linfct = null_hypotheses)
+  pvalues <- summary(glht, test = univariate())$test$pvalues
+  setNames(pvalues, paste0("p_H0", 1:length(null_hypotheses)))
+}
+```
+
+Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit model for different parameter combinations or sample sizes.
+Here, we use the future and furr packages to perform computations in parallel.
+First, we enable parallelization and specify how many parallel cores of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
+
+```{r, message=FALSE, eval=FALSE}
+library(future)
+plan("multisession", workers = 4)
+set.seed(2)
+```
+
+The next code chunk, specifies a simulation design with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
+
+```{r, eval=FALSE}
+library(furrr)
+sim_design <- crossing(
+  rep = 1:250,
+  n_subjects = c(100, 150, 200),
+  n_items = c(10, 30, 50)
+) %>%
+  mutate(pvalues = future_pmap(., sim_and_analyse, 
+    .options = furrr_options(seed = TRUE))) %>%
+  unnest_wider(pvalues)
+```
+
+The result of the computation is a data.frame that contains the p-values of all tested hypotheses for each simulated dataset.
+
+For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations, in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code (heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>).
+
+```{r, eval=FALSE}
+library(binom)
+alpha <- 0.05
+power <- sim_design %>%
+  group_by(n_subjects, n_items) %>% 
+  summarise(power = mean(p_H01 < alpha & p_H02 < alpha & 
+                           p_H03 < alpha & p_H04 < alpha), 
+    n_sig = sum(p_H01 < alpha & p_H02 < alpha & 
+                  p_H03 < alpha & p_H04 < alpha),
+    n = n(),
+    ci.lwr = binom.confint(n_sig, n, method = "wilson")$lower,
+    ci.upr = binom.confint(n_sig, n, method = "wilson")$upper, 
+    .groups = "drop")
+power %>%
+  mutate(across(c(n_subjects, n_items), factor)) %>%
+  ggplot(aes(n_subjects, n_items, fill = power)) +
+  geom_tile() +
+  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
+                                power, ci.lwr, ci.upr)), 
+    color = "white", size = 6) +
+  scale_fill_viridis_c(limits = c(0, 1))
+```
 

### 2023-08-04 docs/manuscript.Rmd
increase the simulation design

@@
 
-```{r, message=FALSE, eval=FALSE}
+```{r, message=FALSE, eval=TRUE}
 library(future)
-plan("multisession", workers = 4)
+plan("multisession", workers = 6)
 set.seed(2)
@@
 
-```{r, eval=FALSE}
+```{r, message=FALSE, eval=TRUE}
 library(furrr)
 sim_design <- crossing(
-  rep = 1:250,
-  n_subjects = c(100, 150, 200),
-  n_items = c(10, 30, 50)
+  rep = 1:300,
+  n_subjects = c(100, 150, 200, 250),
+  n_items = c(10, 30, 50, 70)
 ) %>%
@@
 
-```{r, eval=FALSE}
+```{r, eval=TRUE}
 library(binom)

### 2023-08-07 docs/manuscript.Rmd
worked on manuscript on 8/7

@@
   - name          : "Timo K. Koch"
-    affiliation   : "1, 2"
+    affiliation   : "1, 2, 3"
     corresponding : yes    # Define only one corresponding author
@@
   - name          : "Susanne Gaube"
-    affiliation   : "1"
+    affiliation   : "1, 4"
     role:
@@
 abstract: |
-  Understanding how humans interact with articifical intelligence trough experiments becomes increasingly relevant.
+  Understanding how humans interact with articifical intelligence through experiments becomes increasingly relevant.
 However, human-AI interaction researchers lack the appropriate tools to conduct power and data analyses for the required complex study designs.
 In this work, we provide a tutorial on how to run customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
-By providing code in a case study, we equip human-AI interaction researchers to simulate their own data and run analyses based on GLMMs.
-We discuss the outlook.
+By providing code in a case study, we equip human-AI interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
+We discuss implications and give an outlook.
 
@@
 
-<!-- General intro, relevance of Human AI Interaction experiments, problem of underpowered & inappropriately analyzed work -->
-
-As the integration of artificial intelligence (AI) into our daily lives continues to increase, it has become increasingly important to study how humans interact with these intelligent systems.
+<!-- General intro, relevance of Human AI Interaction experiments, problem of underpowered & inappropriately analyzed work (do not shit too much on prior research :D) -->
+As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems.
 This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
-Therefore, thorough experiments are needed to study human-AI interaction and understand how people perceive and respond to these technologies.
-When conducting research in this field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable.
+Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies.
+When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable.
 Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
@@
 <!-- Sample size planning is increasingly important! -->
-
-Power simulations play a crucial role in addressing the replication crisis and meeting the requirements set forth by journals and funding agencies in Human-AI interaction research.
+Also, power simulations play a crucial role in addressing the replication crisis and meeting the requirements set forth by journals and funding agencies in human-AI interaction research.
 The replication crisis has underscored the problem of high false discovery rates (FDR) coupled with low-powered studies.
@@
 Power simulations can be challenging, particularly when analyzing complex study designs using (generalized) linear mixed models (GLMMs).
-While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of GLMMs in psychological research introduces greater complexity to power calculations.
+While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of GLMMs in human-AI interaction research introduces greater complexity to power calculations.
 For GLMMs, simulations are often required, necessitating assumptions for both fixed and random effects.
-However, most tutorials on GLMMs primarily focus on the most common designs, limiting the guidance available for researchers facing more intricate scenarios.
+However, most tutorials on GLMMs primarily focus on the most common designs, limiting the guidance available for researchers facing more intricate scenarios [REF].
 Moreover, existing tutorials often rely on heuristics for estimating random effect standard deviations and may incorporate meta-analytic results for both random and fixed effects, further adding to the challenge of conducting accurate power calculations for GLMMs.
@@
 <!-- Research Gap & Summary -->
-When planning a Human-AI interaction experiment with complexities such as the use of Generalized Linear Mixed Models (GLMM), non-standard designs, and missing by design conditions, the process of power simulations becomes more challenging.
+When planning a human-AI interaction experiment with complexities such as the use of Generalized Linear Mixed Models (GLMM), non-standard designs, and missing by design conditions, the process of power simulations becomes more challenging.
 Software packages designed for power simulations may not adequately address the complexities of the experiment, making it necessary to build data simulations tailored specifically to the study design.
@@
 Instead, researchers must acquire the necessary skills based on the specific application at hand.
-In this paper, we present a case study that serves as a practical demonstration of how to perform a bespoke data simulation for a concrete study.
+In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study.
 Through this case study, we aim to provide guidance and insights into the complexities involved in addressing complex design considerations and obtaining accurate power estimates.
@@
 
-In this section, we describe the steps to conduct a data simulation and a power analysis. In doing so, we provide a concrete example in the form of a case study.
+In this section, we describe the steps to conduct a data simulation and a power analysis for GLMMs. In doing so, we provide a concrete example in the form of a case study.
 
@@
 <!-- The present example, describe the study idea and design  -->
-In the case study, we simulate the data for a Human-AI Interaction experiment where a health care AI is to be evaluated.
+In the case study, we simulate the data for a human-AI Interaction experiment where a health care AI is to be evaluated.
 Participants view head CT cases and evaluate if a bleeding is present.

### 2023-08-08 docs/manuscript.Rmd
work on introduction

@@
 <!-- General intro, relevance of Human AI Interaction experiments, problem of underpowered & inappropriately analyzed work (do not shit too much on prior research :D) -->
-As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems.
-This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
+As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems.This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
 Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies.
@@
 <!-- Sample size planning is increasingly important! -->
-Also, power simulations play a crucial role in addressing the replication crisis and meeting the requirements set forth by journals and funding agencies in human-AI interaction research.
+In empirical research relying on hypothesis testing, the most common strategy to determine appropriate samples size is based on statistical power.
+Statistical power is defined as the probability that a hypothesis test of interest turns out significant, when analyzing repeated samples from a population with a true effect of some pre-specified size.
+Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
+Power analysis (i.e.
+calculating the statistical power for a given true effect and sample size) plays a crucial role in addressing the replication crisis and meeting the requirements set forth by journals and funding agencies in human-AI interaction research.
 The replication crisis has underscored the problem of high false discovery rates (FDR) coupled with low-powered studies.
-Power simulations offer a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
-Moreover, many journals and funding agencies now mandate the inclusion of power simulations as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
-By incorporating power simulations into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of psychological science.
+Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
+Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
+By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of psychological science.
 
@@
 <!-- Power simulations are hard! (especially for GLMMs) -->
-Power simulations can be challenging, particularly when analyzing complex study designs using (generalized) linear mixed models (GLMMs).
-While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of GLMMs in human-AI interaction research introduces greater complexity to power calculations.
-For GLMMs, simulations are often required, necessitating assumptions for both fixed and random effects.
-However, most tutorials on GLMMs primarily focus on the most common designs, limiting the guidance available for researchers facing more intricate scenarios [REF].
-Moreover, existing tutorials often rely on heuristics for estimating random effect standard deviations and may incorporate meta-analytic results for both random and fixed effects, further adding to the challenge of conducting accurate power calculations for GLMMs.
+Power analysis can be challenging, particularly when analyzing complex study designs using (generalized) linear mixed models (GLMMs; also called multilevel models).
+While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of GLMMs in human-AI interaction research introduces greater complexity to power analysis.
+For GLMMs, analytic options to derive sample sizes are not available for general research designs.
+In contrast, simulations are often required, necessitating assumptions for all parameters in the model.
+An increasing number of tutorials describe how to perform power simulation for multilevel models.
+However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [REF].
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g.
+standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
 
 <!-- Research Gap & Summary -->
-When planning a human-AI interaction experiment with complexities such as the use of Generalized Linear Mixed Models (GLMM), non-standard designs, and missing by design conditions, the process of power simulations becomes more challenging.
-Software packages designed for power simulations may not adequately address the complexities of the experiment, making it necessary to build data simulations tailored specifically to the study design.
+When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes more challenging.
+Currently available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
 Unfortunately, there are no general recommendations that can be applied universally in such cases.
 Instead, researchers must acquire the necessary skills based on the specific application at hand.
-In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study.
+In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study for which no meta-analytic results or conclusive data from pilot studies is available.
 Through this case study, we aim to provide guidance and insights into the complexities involved in addressing complex design considerations and obtaining accurate power estimates.
@@
 <!-- GLMMs are often used for such research -->
-Linear Mixed Models (LMM) and Generalized Linear Mixed Models (GLMM) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
+LMMs and GLMMs are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
 
-Linear Mixed Models (LMM) are extensions of linear regression models that account for correlated data and hierarchical structures.
+LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
 They are used when the outcome variable is continuous and follow a normal distribution.
@@
 
-Generalized Linear Mixed Models (GLMM) extend the LMM framework to accommodate non-normal and categorical outcome variables.
+GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
 They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
@@
 
-Generalized Linear Mixed Models (GLMMs) are gaining increasing popularity in the field of Human-AI interaction research.
+GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
 As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
@@
 
-Power simulations in generalized linear mixed models (GLMMs) are essential for estimating the statistical power of complex psychological study designs.
+Power simulations in GLMMs are essential for estimating the statistical power of complex psychological study designs.
 The model equation for a GLMM combines fixed effects, random effects, and an appropriate link function to model the relationship between predictors and the outcome variable.
@@
 
-In a Generalized linear mixed model (GLMM), the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$.
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$.
 $$
@@
 
-```{r, message=FALSE, eval=TRUE}
+```{r, message=FALSE, eval=FALSE}
 library(future)
@@
 
-```{r, message=FALSE, eval=TRUE}
+```{r, message=FALSE, eval=FALSE}
 library(furrr)
@@
 
-```{r, eval=TRUE}
+```{r, eval=FALSE}
 library(binom)

### 2023-08-08 docs/manuscript.Rmd
work on lme4 section

@@
 <!-- describe the lme4 package and how it works-->
-The lme4 R package is a state-of-the-art tool for fitting frequentist Generalized Linear Mixed Models (GLMMs).
-It provides extensive capabilities for modeling complex data structures, including models with normally distributed random intercepts and random slopes.
-The package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
-
-In lme4, the parameterization used involves the concept of both the beta vector and the theta vector.
-When considering only random intercepts, the theta parameter represents the standard deviation of the random intercepts, providing information about the variability across the different levels of the random effect.
-This allows for a deeper understanding of the random intercepts' influence on the outcome variable.
-
-Regarding random slopes, the interpretation of theta becomes more complex.
-It encompasses the covariance structure between the random slopes and the random intercepts.
-The theta vector accounts for the variability and correlation between these different random effects, offering insights into the multilevel nature of the data.
-However, it is important to note that the tutorial discussed in this paper focuses solely on the simpler case of random intercepts, while acknowledging the more intricate implications of theta in models involving random slopes.
+The lme4 R package is a state-of-the-art tool for fitting frequentist GLMMs (for Bayesian GLMMs, the brms R package is currently the most prominent option).
+The lme4 package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
+Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated datasets.
 

### 2023-08-09 docs/manuscript.Rmd
add new discussion section - Expected effect size vs. smallest effect size of interest: sensitivity power analysis

@@
 
+## Expected effect size vs. smallest effect size of interest: sensitivity power analysis
+
+In our case study, we have performed power simulations based on a single set of parameter values that reflect our assumptions of an expected effect size.
+Instead of extracting this expected effect size from meta-analyses or pilot data which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge which can be considered a vague theoretical model about the data-generating process.
+When sample sizes are chosen based on the results of our power simulations, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.
+However, if the true effect is indeed smaller, power will be lower the study might not be sufficiently informative.
+A common, more conservative strategy for sample size justification is to perform power analysis for a smallest effect size of interest (SESOI). 
+An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero.
+For strategies on the even more difficult task of specifying a plausible SESOI as well as a thorough discussion of various topics concerning power analysis, see @lakensImprovingYourStatistical2022.
+When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called *sensitivity power analysis*.
+By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect, or at least get an impression how strongly power depends on the size of the true effect.
+In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g.
+analyse power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
+However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
+In this setting, sensitivity power analysis would require to manually specify additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
+Power could then be simulated for several of these scenarios (across different number of subjects and items as considered earlier).
+
 # Conclusion

### 2023-08-09 docs/manuscript.Rmd
exclude figure on item variability from main text

@@
 
-In the next plot, we have set the subject standard deviation to almost zero ($\sigma_S = 0.01$.
-This gives us a better way to see the variability between items.
+As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts.
 
-```{r, echo=FALSE}
+We could also have a closer look at variability between items, by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$, see Figure X in Appedix X).
+
+```{r, include=FALSE}
 dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)

### 2023-08-10 docs/manuscript.Rmd
extend the plausible parameter values and results sections

@@
 
-wordcount         : "5339"
+wordcount         : "6417"
 
@@
 
-As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts.
+As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts:
+With wrong advice, virtually all experts have a higher probability to make a correct diagnosis compared to non-experts when considering only items with average difficulty.
+In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
+Patterns like these should be considered carefully, discussed with the domain experts and parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
 
-We could also have a closer look at variability between items, by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$, see Figure X in Appedix X).
+We could also have a closer look at variability between items, by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$, see Figure X in Appendix X).
 
@@
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations, in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code (heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>).
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code (which is heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>).
 
@@
 <!-- How many participants / stimuli are needed for concrete example -->
+As should be the case, power estimates increase with both the number ob subjects and the number of items.
+The confidence intervals indicate how precisely power was estimated by our simulation.
+Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
+In practice, power simulations are often run multiple times with adjusted combinations of sample sizes.
+When run for the first time, it might be revealed that power is way too low (or too high) for some combinations of `n_subjects` and `n_items`.
+When narrowing down the best combination that achieves sufficient power, while at the same time strikes a good balance of how many subjects and items are practically feasible, later rounds of power simulations will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
+This will assure high precision for the final power estimates which are then used for sample size justification of the future study.
+
+Much has been written on the optimal amount of power to target in empirical research.
+The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard.
+When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other, based on practical considerations.
+For the sake of the example, let the targeted power be indeed about 0.8 to detect an effect of the expected size.
+This could be achieved by collecting data of 200 subjects (about 25% of which will be experts) each completing the same 50 items (with advice present in about 67% of cases which is correct in about 80% of cases with present advice).
+If collecting 200 subjects is not practically feasible, it would also be possible to recruit 150 subjects but increase the length of the experiment to over 70 items.
+However, 70 items might take to long to complete (especially for the expert radiologists who might have a busy schedule).
+The simulation suggests that it might also be possible to plan an even shorter experiment with only 30 items, if it is feasible to recruit an even higher number of subjects (> 250, to be determined by additional rounds of power analysis).
+Design parameters which also affect power and which could be investigated in the simulation to find a more optimal trade-off are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
 

### 2023-08-10 docs/manuscript.Rmd
revamped intro and knitted pdf

@@
 By providing code in a case study, we equip human-AI interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
-We discuss implications and give an outlook.
+We discuss implications for human-AI interaction research and give an outlook.
 
@@
 
-<!-- General intro, relevance of Human AI Interaction experiments, problem of underpowered & inappropriately analyzed work (do not shit too much on prior research :D) -->
-As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems.This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
+<!-- General intro, relevance of Human AI Interaction experiments, need for well-powered and analyzed experimental research in this field -->
+As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems. 
+This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
 Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies. 
-When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable.
-Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable, and use appropriate statistical models to analyze the data.
 
-<!-- Sample size planning is increasingly important! -->
+<!-- Some basics on statistical power, power analysis and why it is important -->
 In empirical research relying on hypothesis testing, the most common strategy to determine appropriate samples size is based on statistical power. 
@@
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-Power analysis (i.e.
-calculating the statistical power for a given true effect and sample size) plays a crucial role in addressing the replication crisis and meeting the requirements set forth by journals and funding agencies in human-AI interaction research.
-The replication crisis has underscored the problem of high false discovery rates (FDR) coupled with low-powered studies.
-Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
-Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
-By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of psychological science.
-
-## Customized power simulations are hard
+In line with this, power analysis represents the act of calculating the statistical power for a given true effect and sample size.
 
 <!-- Power simulations are hard! (especially for GLMMs) -->
-Power analysis can be challenging, particularly when analyzing complex study designs using (generalized) linear mixed models (GLMMs; also called multilevel models).
-While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of GLMMs in human-AI interaction research introduces greater complexity to power analysis.
+Hereby, for a desired statistical power and an assumed true effect, the required sample size to find such an effect can be calculated.
+Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
+While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
 For GLMMs, analytic options to derive sample sizes are not available for general research designs.
-In contrast, simulations are often required, necessitating assumptions for all parameters in the model.
 An increasing number of tutorials describe how to perform power simulation for multilevel models.
@@
 standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
+In contrast, simulations are often required, necessitating assumptions for all parameters in the model.
+Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
+By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science.
 
 <!-- Research Gap & Summary -->
-When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes more challenging.
+When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes challenging.
 Currently available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
@@
 Through this case study, we aim to provide guidance and insights into the complexities involved in addressing complex design considerations and obtaining accurate power estimates.
-
-## LMMs, GLMMs and the lme4 package in R
-
-<!-- GLMMs are often used for such research -->
-LMMs and GLMMs are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
-
-LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
-They are used when the outcome variable is continuous and follow a normal distribution.
-LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
-
-GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
-
-GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
-GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
-By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
-
-<!-- describe the lme4 package and how it works-->
-The lme4 R package is a state-of-the-art tool for fitting frequentist GLMMs (for Bayesian GLMMs, the brms R package is currently the most prominent option).
-The lme4 package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
-However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated datasets.
 
@@
 - excursus: estimate expected with of confidence intervals for different sample sizes and number of stimuli
+
+## LMMs, GLMMs and the lme4 package in R
+
+<!-- GLMMs are often used for such research -->
+LMMs and GLMMs are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
+
+LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
+They are used when the outcome variable is continuous and follow a normal distribution.
+LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
+
+GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
+They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
+
+GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
+GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
+This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
+By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+
+<!-- describe the lme4 package and how it works-->
+The lme4 R package is a state-of-the-art tool for fitting frequentist GLMMs (for Bayesian GLMMs, the brms R package is currently the most prominent option).
+The lme4 package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
+Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated datasets.
 
@@
 3.
-Teaching Bespoke Power Simulation Skills to Human-AI Interaction Researchers: The ability to conduct bespoke power simulations is a valuable skill that should be taught to Human-AI interaction researchers.
+Teaching Power Simulation Skills to Human-AI Interaction Researchers: The ability to conduct bespoke power simulations is a valuable skill that should be taught to Human-AI interaction researchers.
 By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
@@
 5.
-Recognizing the Value of Bespoke Power Simulations: Bespoke power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
+Recognizing the Value of Power Simulations: Bespoke power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
 They offer insights into the reliability and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
@@
 An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero.
-For strategies on the even more difficult task of specifying a plausible SESOI as well as a thorough discussion of various topics concerning power analysis, see @lakensImprovingYourStatistical2022.
+For strategies on the even more difficult task of specifying a plausible SESOI as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].
 When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called *sensitivity power analysis*.
@@
 
-Bespoke power simulations play a critical role in human-AI interaction research, allowing researchers to tailor power estimation to the unique aspects of their experiments.
+Power simulations play a critical role in human-AI interaction research, allowing researchers to tailor power estimation to the unique aspects of their experiments.
 The skills required to perform these simulations should be taught to Human-AI interaction researchers, fostering a deeper understanding of statistical power and enhancing research design.

### 2023-08-11 docs/manuscript.Rmd
manuscript update 8/11 by Timo

@@
 ---
-title             : "Tutorial for Customized Power Simulations and Data Analyses for Human-AI Interaction Experiments using Generalized Linear Mixed Models"
+title             : "A Tutorial for Customized Power Simulations and Data Analyses for Human-AI Interaction Experiments using Generalized Linear Mixed Models"
 shorttitle        : "GLMM Power Simulation Tutorial"
@@
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
+    #address       : "Postal address"
     email         : "timo.koch@psy.lmu.de"
@@
       - "Conceptualization"
+      - "Formal Analysis"
       - "Writing - Original Draft Preparation"
@@
     role:
+      - "Formal Analysis"
+      - "Methodology"
+      - "Visualization"
+      - "Writing - Original Draft Preparation"
       - "Writing - Review & Editing"
-      - "Supervision"
+  - name          : "Anne-Kathrin Kleine"
+    affiliation   : "1"
+    role:
+      - "Writing - Review & Editing"
   - name          : "Susanne Gaube"
@@
 
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
+# authornote: |
+#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+# 
+#   Enter author note here.
 
 abstract: |
-  Understanding how humans interact with articifical intelligence through experiments becomes increasingly relevant.
+  Understanding how humans interact with articifical intelligence (AI) through experiments becomes increasingly relevant.
 However, human-AI interaction researchers lack the appropriate tools to conduct power and data analyses for the required complex study designs.
 In this work, we provide a tutorial on how to run customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
-By providing code in a case study, we equip human-AI interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
-We discuss implications for human-AI interaction research and give an outlook.
+By providing code in a specific case study, we equip human-AI interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
+We discuss implications for human-AI interaction research and provide an outlook on the relevance of power simulations in the field.
 
@@
 
-wordcount         : "6417"
+wordcount         : "6264"
 
-bibliography      : "r-references.bib"
+bibliography      : "Simulation Tutorial Paper.bib"
 
@@
 An increasing number of tutorials describe how to perform power simulation for multilevel models.
-However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [REF].
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g.
-standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
+However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
 In contrast, simulations are often required, necessitating assumptions for all parameters in the model.
@@
 
-In this section, we describe the steps to conduct a data simulation and a power analysis for GLMMs. In doing so, we provide a concrete example in the form of a case study.
-
-We used `r cite_r("r-references.bib")` for all our analyses.
+<!-- overview of this section -->
+In this section, we describe the steps to conduct a data simulation and a power analysis for GLMMs with a specific experimental design in human-AI interaction research in mind in the form of a case study.
 
@@
 <!-- The present example, describe the study idea and design -->
-In the case study, we simulate the data for a human-AI Interaction experiment where a health care AI is to be evaluated.
-Participants view head CT cases and evaluate if a bleeding is present.
-Also, there can be AI advice that is either correct or incorrect.
-This medical diagnosis can be correct or incorrect.
+In the case study, we simulate the data for a human-AI Interaction experiment where the diagnostic performance of users of a health care AI is to be evaluated.
+The goal is to understanding how AI advice influences medical decision-making.
+Participants, radiologists (task experts) and emergency doctors (non task experts) view head CT (computertomography) scans and evaluate if an intracranial bleeding is present.
+To support their decision making, an AI provides a suggestion for a medical diagnosis.
+This AI advice can be either either correct or incorrect.
+Participants' medical diagnosis (i.e., bleeding or no bleeding) can be correct or incorrect. 
+Also, we have a control group with no AI advice present. This introduces missings by design that has an effect on our data simulations. 
+Through this experiment, we want to determine if experts are better than non-experts in reading head CT scans.
+Further, we want to research if correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
+In the present example, it would be more challenging to recruit task expert (i.e., radiologists) as there is only a limited amount of such people.
+Non-experts would be easier to recruit.
+The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
 
-In this case, it would be more challenging to recruit task expert as there is only a limited amount of such people.
-Non-experts would be easier to recruit.
-The open question is how many task experts and non-experts to recruit to achieve sufficient power.
+## The lme4 package in R
 
-describe study design and sample size restrictions 
-- describe research question (i.e. the coefficients targeted by power simulation)
+<!-- describe the lme4 package and how it works-->
+For our data simulation and data analysis, we use the lme4 R package [@batesFittingLinearMixedEffects2015].
+The lme4 package is a state-of-the-art tool for fitting frequentist GLMMs (for Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017]).
+The lme4 package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
+Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.
 
-<!-- basics of power simulation in GLMMs (with the concrete example in mind) -->
+## Generalized linear mixed models
 
-Power simulations in GLMMs are essential for estimating the statistical power of complex psychological study designs.
+<!-- GLMM intro -->
+LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects. 
+LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
+They are used when the outcome variable is continuous and follow a normal distribution.
+LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
+GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
+They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
+
+<!-- Relevance of GLMMs -->
+GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
+GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
+This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
+By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+
+<!-- basics of power simulation in GLMMs (with our specific example in mind) -->
+Power simulations in GLMMs are essential for estimating the statistical power of complex experimental study designs.
 The model equation for a GLMM combines fixed effects, random effects, and an appropriate link function to model the relationship between predictors and the outcome variable.
@@
 
-## TODO
-
-points not discussed yet:
-
-- discuss the specific missing by design complexity
-- discuss different possible parametrizations of the fixed effects
-- find the appropriate lme4 formula (simple models first, more complex models later)
-- determine the correct labels for model coefficients in the lme4 output
-- excursus: estimate expected with of confidence intervals for different sample sizes and number of stimuli
-
-## LMMs, GLMMs and the lme4 package in R
-
-<!-- GLMMs are often used for such research -->
-LMMs and GLMMs are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
-
-LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
-They are used when the outcome variable is continuous and follow a normal distribution.
-LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
-
-GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
-
-GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
-GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
-By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
-
-<!-- describe the lme4 package and how it works-->
-The lme4 R package is a state-of-the-art tool for fitting frequentist GLMMs (for Bayesian GLMMs, the brms R package is currently the most prominent option).
-The lme4 package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
-However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated datasets.
-
-## Generalized linear mixed models
-
+<!-- deep dive into our specific example -->
 In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$.
@@
 
-## Data Simulation
+## Data simulation
 
@@
 
-<!-- Results of power sensitivity analyses -->
-
 # Discussion
 
-<!-- Summarize key findings-->
-
-Human-AI interaction research requires meticulous planning and consideration of statistical power to ensure reliable and meaningful results.
+<!-- short summary of key take aways from this tutorial -->
+Human-AI interaction research requires careful planning and consideration of statistical power to ensure reliable and meaningful results.
 While heuristics and helper programs can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
-In this paper, we discuss the necessity of bespoke power simulations in human-AI interaction research, their implications for research design, and the importance of teaching these skills to Human-AI interaction researchers.
-Furthermore, we highlight the discrepancy between the perceived effort involved in performing bespoke power simulations and the actual effort required.
-Finally, we emphasize the value of power simulations as research contributions and their alignment with open science and preregistration practices.
-
-1.
-The Need for Power Simulations in Human-AI Interaction Research: Human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or simple helper programs.
-Bespoke power simulations offer a solution by providing Human-AI interaction researchers with a tailored approach to estimating statistical power.
-These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
-
-2.
-Managing Simulations with Discrete Predictor Variables: Bespoke power simulations become more manageable when all predictor variables are discrete and fixed by the study design.
-This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need to simulate predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
-
-3.
-Teaching Power Simulation Skills to Human-AI Interaction Researchers: The ability to conduct bespoke power simulations is a valuable skill that should be taught to Human-AI interaction researchers.
-By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping Human-AI interaction researchers with the knowledge and tools to perform bespoke power simulations empowers them to make informed decisions and enhance the rigor of their studies.
-
-4.
-Addressing the Mismatch in Effort Perception: There is often a significant disconnect between the perceived effort required to perform bespoke power simulations and the actual effort estimated by researchers and collaborators in Human-AI interaction research.
-Many content researchers and collaborators request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these simulations.
-It is crucial to raise awareness about the effort involved in bespoke power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
-
-5.
-Recognizing the Value of Power Simulations: Bespoke power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
-They offer insights into the reliability and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
-The importance of bespoke power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports.
-
-6.
-Integration with Open Science and Preregistration Practices: Bespoke power simulations align well with the principles of open science and preregistration in Human-AI interaction research.
-When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of Human-AI interaction research.
+The present tutorial presents a specific example on how to run power simulations for human-AI interaction experiments.
 
@@
 By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect, or at least get an impression how strongly power depends on the size of the true effect.
-In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g.
-analyse power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
+In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyse power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
 However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
@@
 
+## Outlook
+
+<!-- Outlook -->
+Beyond the specifics, we want to outline six avenues regarding the outlook on power simulations in human-AI interaction research:
+
+1.
+The need for power simulations in human-AI interaction research: As this work has clearly pointed out, human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or simple helper programs.
+Power simulations offer a solution by providing Human-AI interaction researchers with a tailored approach to estimating statistical power.
+These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
+
+2.
+Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
+This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need to simulate predictor values, which would introduce additional assumptions.
+By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
+
+3.
+Teaching power simulation skills to human-AI interaction researchers: The ability to conduct  power simulations is a valuable skill that should be taught to Human-AI interaction researchers.
+By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
+Equipping Human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
+
+4.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the perceived effort required to perform power simulations and the actual effort estimated by researchers and collaborators in Human-AI interaction research.
+Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these simulations.
+It is crucial to raise awareness about the effort involved in  power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
+
+5.
+Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
+They offer insights into the reliability and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
+The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports.
+
+6.
+Integration with Open Science and preregistration practices: Power simulations align well with the principles of open science and preregistration in Human-AI interaction research.
+When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
+By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of Human-AI interaction research.
+
 # Conclusion
@@
 Power simulations play a critical role in human-AI interaction research, allowing researchers to tailor power estimation to the unique aspects of their experiments.
-The skills required to perform these simulations should be taught to Human-AI interaction researchers, fostering a deeper understanding of statistical power and enhancing research design.
-It is essential to bridge the gap between perceived and actual effort associated with power simulations and recognize their value as research contributions.
-By integrating bespoke power simulations with open science and preregistration practices, researchers can improve the robustness and transparency of their findings, advancing the field.
+Through this tutorial, we aim to provide researchers with the necessary skills and tools to perform these simulations themselves.
+By integrating power simulations with open science and preregistration practices, researchers can improve the robustness and transparency of their findings, advancing the field.
 

### 2023-08-18 docs/manuscript.Rmd
iterate on introduction

@@
 
-bibliography      : "Simulation Tutorial Paper.bib"
+bibliography      : "../Simulation Tutorial Paper.bib"
 
@@
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-In line with this, power analysis represents the act of calculating the statistical power for a given true effect and sample size.
+Power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
+By running the calculation for a range of sample sizes, the required sample size to find an assumed true effect with desired statistical power can be determined.
 
 <!-- Power simulations are hard! (especially for GLMMs) -->
-Hereby, for a desired statistical power and an assumed true effect, the required sample size to find such an effect can be calculated.
-Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
+Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for results with high evidencial value.
 While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
@@
 However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies are available to determine plausible values for all model parameters.
 In contrast, simulations are often required, necessitating assumptions for all parameters in the model.
-Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
 Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
@@
 Instead, researchers must acquire the necessary skills based on the specific application at hand.
-In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study for which no meta-analytic results or conclusive data from pilot studies is available.
+In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study for which no meta-analytic results or conclusive data from pilot studies are available.
 Through this case study, we aim to provide guidance and insights into the complexities involved in addressing complex design considerations and obtaining accurate power estimates.
@@
 <!-- overview of this section -->
-In this section, we describe the steps to conduct a data simulation and a power analysis for GLMMs with a specific experimental design in human-AI interaction research in mind in the form of a case study.
+In this section, we describe the steps to conduct a data simulation and a power analysis for GLMMs, with a specific experimental design from human-AI interaction research, in the form of a case study.
 
@@
 In the case study, we simulate the data for a human-AI Interaction experiment where the diagnostic performance of users of a health care AI is to be evaluated.
-The goal is to understanding how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and emergency doctors (non task experts) view head CT (computertomography) scans and evaluate if an intracranial bleeding is present.
-To support their decision making, an AI provides a suggestion for a medical diagnosis.
-This AI advice can be either either correct or incorrect.
-Participants' medical diagnosis (i.e., bleeding or no bleeding) can be correct or incorrect. 
-Also, we have a control group with no AI advice present. This introduces missings by design that has an effect on our data simulations. 
+The goal is to understand how AI advice influences medical decision-making.
+Participants, radiologists (task experts) and emergency doctors (non task experts), view head CT (computertomography) scans and evaluate if an intracranial bleeding is present.
+To support their decision making, an AI provides a suggestion for a medical diagnosis. This AI advice can be either either correct or incorrect. 
+Participants then provide a final medical diagnosis (i.e., bleeding or no bleeding) that can be correct or incorrect. 
+Also, we have control trials, in which no AI advice is present. 
+This introduces some missing values by design (advice is neither correct nor incorrect, when no advice is present) which must be taken into account when simulating and analyzing the data. 
 Through this experiment, we want to determine if experts are better than non-experts in reading head CT scans.
 Further, we want to research if correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
-In the present example, it would be more challenging to recruit task expert (i.e., radiologists) as there is only a limited amount of such people.
-Non-experts would be easier to recruit.
+In the present example, it would be more challenging to recruit task expert (i.e., radiologists) as there is only a limited amount of such people. Non-experts would be easier to recruit.
 The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
@@
 <!-- describe the lme4 package and how it works-->
-For our data simulation and data analysis, we use the lme4 R package [@batesFittingLinearMixedEffects2015].
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015].
 The lme4 package is a state-of-the-art tool for fitting frequentist GLMMs (for Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017]).
-The lme4 package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+The lme4 package also includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
 However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.

### 2023-08-18 docs/manuscript.Rmd
iterate on methods and results

@@
 LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
-They are used when the outcome variable is continuous and follow a normal distribution.
+They are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables).
 LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
@@
 GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
+The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.
 
@@
 GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs.
 GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
@@
 - $advice\_correct_{si}$: whether this advice was correct (1) or not (0)
-- $expert_s$: whether subject $s$ was a professional neurologist (1) or not (0)
+- $expert_s$: whether subject $s$ was a radiologist (1) or not (0)
 
@@
 
-The following R function simulates a full dataset structured according to our case study.
+The following R function simulates a full dataset structured according to the desgin of our case study.
 The faux package contains useful functions when simulating factorial designs including random effects.
@@
 Thus the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
-We add a between subject effect with the `add_between` command, simulating that about $25%$ of subjects are experts.
-The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice and if advice is presented, the advice will be correct in about $80%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
-Next we simulate one random effect each subject (`u0s`) and for each item (`u0i`).
+We add a between subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
+The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
+Next we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
 As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with mean 0 and a standard deviation specified by the user.
@@
 The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
-The complete set of rules for the syntax are outlined in (CITE LME4 PAPER) and in the documentation of the lme4 package.
+The complete set of rules for the syntax are outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
 
 In lme4, a GLMM is fitted with the `glmer` function.
-By setting `family =  "binomial"` we request a binomial GLMM, appropriate for our binary dependent variable `y_bin` (the binomial GLMM used the canonical logit link by default).
+By setting `family =  "binomial"` we request a binomial GLMM, appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default).
 
@@
 Unfortunately, the model parameters in a binomial GLMM are hard to interpret, because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation works by imagining a subject with average ability ($u0s = 0$) responding to an item with average difficulty ($u0i = 0$).
+The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item with average difficulty ($u_{0i} = 0$).
 Then the model implied probability that such a person solves such an item is given by:
@@
 In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-The other extreme case would be the probability that a non-expert with average ability solves and item with average difficulty when presented without any advice:
+The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
 $$
@@
 
-Due to this complicated relationship, many experts argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider the implied predicted distribution of the dependent variable for each experimental conditions across all subjects and items.
+Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
+Instead, it can be more intuitive to consider the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
 
@@
 
-However, we need to think about the model parameters again when wanting to test hypotheses which we have theoretically derived from some research question.
+However, we need to think about the model parameters again when we want to test hypotheses which we have theoretically derived from some research question.
 Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.
@@
 Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command.
-With a standard significance level of $\alpha = 0.05$, we would reject all four null hypotheses and therefore also reject the combined null hypothesis for this simulated dataset.
+With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value for hypothesis $H_{03}$ is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
+Note that this decision would be wrong, because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
 
@@
 
-When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where the numbers came from.
+When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
 All parameter values have been determined in repeated exchanges with our affiliated domain experts and we here outline a few strategies on how to determine plausible parameter values.
@@
 In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
-For example, we determined the $\beta$ parameters in a way that correspond to a desired probability, that an expert with average ability solves an item with average difficulty when presented with a correct advice.
-However even if the model were true, we would not observe this probability if we estimated the probability in a group of expert responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
+For example, we determined the $\beta$ parameters in a way that correspond to a desired probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with a correct advice.
+However even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
 This implies that one must be careful when specifying parameter values based on previous studies or pilot data.
@@
 The well known inequality of conditional and marginal effects in GLMMs makes their interpretation more difficult, however, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
-When parameter values have selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
+When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
 Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
@@
 Figure X shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
-We can see that indeed, the average probabilities (block dots) differ from the probabilities of average subjects and items considered in the previous section.
+We can see that indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
 This difference increases with the variability of the random effects.
@@
 
-The final plot demonstrates, that these plots are also useful to spot standard deviations specified to high.
-For example, if we set $\sigma_S = 3$ and $\sigma_I = 3$ this implies that in each experimental condition, the probabilities that a subject solves an item are always close to either 0 or 1, which is not a plausible assumption.
+The final plot demonstrates, that these plots are also useful to spot standard deviations that are specified too high.
+For this example, we have set $\sigma_S = 3$ and $\sigma_I = 3$.
+This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
 
@@
 
-With all these considerations out of the way, we are finally ready to perform a power simulation.
+With all these considerations out of the way, we are finally ready to perform a power analysis.
 Wrapping the `simulate` function already introduced earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
@@
 
-Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit model for different parameter combinations or sample sizes.
-Here, we use the future and furr packages to perform computations in parallel.
+Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
+Thus, we use the future and furr packages to perform computations in parallel.
 First, we enable parallelization and specify how many parallel cores of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
@@
 
-The next code chunk, specifies a simulation design with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
+The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
 
@@
 In practice, power simulations are often run multiple times with adjusted combinations of sample sizes.
-When run for the first time, it might be revealed that power is way too low (or too high) for some combinations of `n_subjects` and `n_items`.
+When run for the first time, it might be revealed that power is way too low (or higher than required) for some combinations of `n_subjects` and `n_items`.
 When narrowing down the best combination that achieves sufficient power, while at the same time strikes a good balance of how many subjects and items are practically feasible, later rounds of power simulations will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
@@
 When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other, based on practical considerations.
-For the sake of the example, let the targeted power be indeed about 0.8 to detect an effect of the expected size.
+For the sake of the example, let the targeted power be indeed about 0.8 to detect an effect of the expected size implied by our data simulation.
 This could be achieved by collecting data of 200 subjects (about 25% of which will be experts) each completing the same 50 items (with advice present in about 67% of cases which is correct in about 80% of cases with present advice).
 If collecting 200 subjects is not practically feasible, it would also be possible to recruit 150 subjects but increase the length of the experiment to over 70 items.
-However, 70 items might take to long to complete (especially for the expert radiologists who might have a busy schedule).
-The simulation suggests that it might also be possible to plan an even shorter experiment with only 30 items, if it is feasible to recruit an even higher number of subjects (> 250, to be determined by additional rounds of power analysis).
+However, 70 items might take to long to complete for the doctors participating in the study, who probably all have a busy schedule.
+The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items, if it is feasible to recruit an even higher number of subjects (> 250, to be determined by additional rounds of power analysis).
 Design parameters which also affect power and which could be investigated in the simulation to find a more optimal trade-off are the ratio of experts, the frequency of whether advice is presented and whether it is correct.

### 2023-08-18 docs/manuscript.Rmd
iterate on discussion

@@
 Human-AI interaction research requires careful planning and consideration of statistical power to ensure reliable and meaningful results.
-While heuristics and helper programs can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
+While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
 The present tutorial presents a specific example on how to run power simulations for human-AI interaction experiments.
@@
 In our case study, we have performed power simulations based on a single set of parameter values that reflect our assumptions of an expected effect size.
-Instead of extracting this expected effect size from meta-analyses or pilot data which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge which can be considered a vague theoretical model about the data-generating process.
+Instead of extracting this expected effect size from meta-analyses or pilot data which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
+Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
 When sample sizes are chosen based on the results of our power simulations, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.
-However, if the true effect is indeed smaller, power will be lower the study might not be sufficiently informative.
+However, if the true effect is indeed smaller, power will be lower and the study might not be sufficiently informative.
 A common, more conservative strategy for sample size justification is to perform power analysis for a smallest effect size of interest (SESOI). 
@@
 When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called *sensitivity power analysis*.
-By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect, or at least get an impression how strongly power depends on the size of the true effect.
+By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect, or at least get an impression on how strongly power depends on the size of the true effect.
 In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyse power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
@@
 <!-- Outlook -->
-Beyond the specifics, we want to outline six avenues regarding the outlook on power simulations in human-AI interaction research:
+Beyond the specifics of our concrete case study, we want to outline six avenues regarding the outlook on power simulations in human-AI interaction research:
 
 1.
-The need for power simulations in human-AI interaction research: As this work has clearly pointed out, human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or simple helper programs.
+The need for power simulations in human-AI interaction research: As our case study has clearly pointed out, human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
 Power simulations offer a solution by providing Human-AI interaction researchers with a tailored approach to estimating statistical power.
@@
 Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
-This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need to simulate predictor values, which would introduce additional assumptions.
+This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
 By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
@@
 Equipping Human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
+The need to reason about how to simulate plausible data which is in line with the research hypothesis while not violating domain expertise on how plausible data should look like might also contribute to planning more insightful studies that can answer more precise research questions.
 
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the perceived effort required to perform power simulations and the actual effort estimated by researchers and collaborators in Human-AI interaction research.
-Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these simulations.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in Human-AI interaction research.
+Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
 It is crucial to raise awareness about the effort involved in power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
@@
 Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
-They offer insights into the reliability and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
+They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
 The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports.

### 2023-08-18 docs/manuscript.Rmd
add references

@@
 
-bibliography      : "../Simulation Tutorial Paper.bib"
+bibliography      : ["../Simulation Tutorial Paper.bib", "r-references.bib"]
 
@@
 library("papaja")
+library("tidyverse")
+library("faux")
+library("lme4")
+library("multcomp")
+library("marginaleffects")
+library("future")
+library("furrr")
+library("binom")
 r_refs("r-references.bib")
@@
 Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies. 
-When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable, and use appropriate statistical models to analyze the data.
+When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.
 
@@
 For GLMMs, analytic options to derive sample sizes are not available for general research designs.
-An increasing number of tutorials describe how to perform power simulation for multilevel models.
+An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @murayamaSummarystatisticsbasedPowerAnalysis2022; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
 However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
 Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies are available to determine plausible values for all model parameters.
-In contrast, simulations are often required, necessitating assumptions for all parameters in the model.
+In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
 Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
 Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
-By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science.
+By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science [@yarkoniGeneralizabilityCrisis2022].
 
@@
 However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.
+Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.[^simr]
+
+[^simr]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016] which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
 
@@
 The following R function simulates a full dataset structured according to the desgin of our case study.
-The faux package contains useful functions when simulating factorial designs including random effects.
+The faux package [@R-faux] contains useful functions when simulating factorial designs including random effects.
 
@@
 
-With the marginaleffects package, we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition:
+With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition:
 
@@
 
-We can specify and test hypotheses like these with the multcomp package as follows:
+We can specify and test hypotheses like these with the multcomp package [@R-multcomp] as follows:
 
@@
 Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
-Thus, we use the future and furr packages to perform computations in parallel.
+Thus, we use the future [@R-future] and furrr [@R-furrr] packages to perform computations in parallel.
 First, we enable parallelization and specify how many parallel cores of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
@@
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations, in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code (which is heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>).
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code [^faux].
+
+[^faux]: This code is heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
 
@@
 Equipping Human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
-The need to reason about how to simulate plausible data which is in line with the research hypothesis while not violating domain expertise on how plausible data should look like might also contribute to planning more insightful studies that can answer more precise research questions.
+The need to reason about how to simulate plausible data which is in line with the research hypothesis while not violating domain expertise on how plausible data should look like might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
 

### 2023-08-18 docs/manuscript.Rmd
fix small problem with future reference

@@
 Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
-Thus, we use the future [@R-future] and furrr [@R-furrr] packages to perform computations in parallel.
+Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
 First, we enable parallelization and specify how many parallel cores of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
@@
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations, in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code [^faux].
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^faux]
 

### 2023-08-18 docs/manuscript.Rmd
add more references

@@
 
-wordcount         : "6264"
+wordcount         : "6551"
 
@@
 While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
-For GLMMs, analytic options to derive sample sizes are not available for general research designs.
-An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @murayamaSummarystatisticsbasedPowerAnalysis2022; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+For GLMMs, analytic options to derive sample sizes are not available for general research designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
+An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
 However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
@@
 <!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015].
-The lme4 package is a state-of-the-art tool for fitting frequentist GLMMs (for Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017]).
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015]. The lme4 package is a state-of-the-art tool for fitting frequentist GLMMs.[^brms]
+
+[^brms]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
+
 The lme4 package also includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
@@
 <!-- GLMM intro -->
-LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects. 
+LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
 LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
@@
 GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
 GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
@@
 <!-- deep dive into our specific example -->
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$.
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021].
 $$
@@
 Because all hypotheses tested simultaneously with the `glht` function must have the same direction, we flip the sign of inequalities three and four by multiplying them with $-1$.
-The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously.
-However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected.
+The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously [@hothornSimultaneousInferenceGeneral2008].
+However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected [i.e.
+intersection-union setting\; @dmitrienkoTraditionalMultiplicityAdjustment2013].
 In such cases, the error probabilities do not accumulate and we would waste power when correcting for multiple testing.
@@
 
-The well known inequality of conditional and marginal effects in GLMMs makes their interpretation more difficult, however, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
+The well known inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult, however, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
 When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
@@
 Much has been written on the optimal amount of power to target in empirical research.
-The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard.
+The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
 When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other, based on practical considerations.
@@
 A common, more conservative strategy for sample size justification is to perform power analysis for a smallest effect size of interest (SESOI). 
-An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero.
+An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011].
 For strategies on the even more difficult task of specifying a plausible SESOI as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].

### 2023-08-18 docs/appendix.Rmd
add appendix.Rmd (only contains some fragments at the moment)

## General principles in power analysis

```{r, message=FALSE}
library(tidyverse)
set.seed(1)
dat <- simulate(n_subjects = 100, n_items = 50)
dat$expert <- factor(dat$expert)
dat$advice_present <- factor(dat$advice_present)
dat$advice_correct <- factor(dat$advice_correct)
```

```{r, message=FALSE}
library(lme4)
f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
  expert:advice_present + expert:advice_correct +
  (1|subject) + (1|item)
fit <- glmer(f, data = dat, family = "binomial")
```

```{r}
library(marginaleffects)
marginal_means(fit)
avg_comparisons(fit, variables = list(expert = c(0, 1)))
comparisons(fit,
  newdata = datagrid(expert = c("0", "1"), advice_present = c("0", "1"), advice_correct = c("0", "1"))
)
```

## Sample size justification based on precision: simulate expected confidence interval width

```{r}
simulate <- function(n_subjects = 100, n_items = 50,
  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
  b_ea = 0.790, b_ec = -1.393,
  sd_u0s = 0.5, sd_u0i = 0.5, ...){
  require(dplyr)
  require(faux)
  # simulate design
  dat <- add_random(subject = n_subjects, item = n_items) %>%
    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
    mutate(advice_correct = if_else(advice_present == 1, 
                                    rbinom(n(), 1, prob = 0.8), 0)) %>%
    # add random effects
    add_ranef("subject", u0s = sd_u0s) %>%
    add_ranef("item", u0i = sd_u0i) %>%
    # compute dependent variable
    mutate(linpred = b_0 + u0i + u0s +
        b_e * expert + b_a * advice_present + b_c * advice_correct +
        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
    mutate(y_prob = plogis(linpred)) %>%
    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
  dat
}

sim_and_analyse <- function(
  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + 
    expert:advice_present + expert:advice_correct + (1|subject) + (1|item)", 
  null_hypotheses = c("expert:advice_correct = 0"), ...){
  require(lme4)
  require(multcomp)
  # simulate data
  dat <- simulate(...)
  # fit model
  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
  # compute p-values
  glht <- glht(model, linfct = null_hypotheses)
  confint(glht, level = 0.95, calpha = univariate_calpha())$confint
}
```

```{r, message=FALSE, eval=TRUE}
library(future)
plan("multisession", workers = 6)
set.seed(2)
```

```{r, message=FALSE, eval=TURE}
library(tidyverse)
library(furrr)
sim_design <- crossing(
  rep = 1:200,
  n_subjects = c(150, 200),
  n_items = c(30, 50)
) %>%
  mutate(ci = future_pmap(., sim_and_analyse, 
    .options = furrr_options(seed = TRUE))) %>%
  unnest_wider(ci)
```

```{r, eval=FALSE}
precision <- sim_design %>%
  group_by(n_subjects, n_items) %>% 
  mutate(width = `expert:advice_correct`[,3] - `expert:advice_correct`[,2]) %>%
  summarise(precision = mean(width),
    ci.lwr = t.test(width)$conf.int[1],
    ci.upr = t.test(width)$conf.int[2], 
    .groups = "drop")
precision %>%
  mutate(across(c(n_subjects, n_items), factor)) %>%
  ggplot(aes(n_subjects, n_items, fill = precision)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
                                precision, ci.lwr, ci.upr)), 
    color = "white", size = 6) +
  scale_fill_viridis_c(limits = c(0, 2))
```

### 2023-08-22 docs/manuscript.Rmd
Update manuscript.Rmd

@@
   - id            : "4"
-    institution   : "University College London"
+    institution   : "University College London, Global Business School for Health"
 
@@
 abstract: |
-  Understanding how humans interact with articifical intelligence (AI) through experiments becomes increasingly relevant.
-However, human-AI interaction researchers lack the appropriate tools to conduct power and data analyses for the required complex study designs.
-In this work, we provide a tutorial on how to run customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
+  Understanding how humans interact with artificial intelligence (AI) through experiments becomes increasingly relevant.
+However, human-AI interaction researchers often lack the appropriate tools to conduct power and data analyses for the complex study designs required.
+In this work, we provide a comprehensive guide to performing customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
 By providing code in a specific case study, we equip human-AI interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
@@
 <!-- General intro, relevance of Human AI Interaction experiments, need for well-powered and analyzed experimental research in this field -->
-As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems. 
-This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
-Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies. 
+With the growing integration of artificial intelligence (AI) into our daily lives and routines, it has become increasingly important to study and understand how humans interact with these intelligent systems.
+This is particularly relevant as AI becomes more sophisticated and is deployed to support decision-making processes in fields such as law and medicine that have a significant impact on individuals and society at large.
+Therefore, thorough experimental research is needed to study human-AI interaction to understand how people perceive and respond to these novel technologies. 
 When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.

### 2023-08-29 docs/manuscript.Rmd
added comments and suggestions

--- 2023-08-18 docs/manuscript.Rmd
@@
 
-bibliography      : ["../Simulation Tutorial Paper.bib", "r-references.bib"]
+bibliography      : "Simulation Tutorial Paper.bib"
 
@@
 library("papaja")
-library("tidyverse")
-library("faux")
-library("lme4")
-library("multcomp")
-library("marginaleffects")
-library("future")
-library("furrr")
-library("binom")
 r_refs("r-references.bib")
@@
 Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies.
-When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.
+<!-- AK: A bit out of context - not clear to me why these experiments are needed.
+Maybe you could provide some examples what can be achieved with experiments in that context.
+-->
+When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size to ensure that the results obtained are both statistically significant and reliable, and use appropriate statistical models to analyze the data.
+<!-- AK: refs are generally missing - will they be added later?
+-->
 
 <!-- Some basics on statistical power, power analysis and why it is important -->
-In empirical research relying on hypothesis testing, the most common strategy to determine appropriate samples size is based on statistical power. 
+In empirical research relying on hypothesis testing, the most common strategy to determine appropriate sample size is based on statistical power. 
 Statistical power is defined as the probability that a hypothesis test of interest turns out significant, when analyzing repeated samples from a population with a true effect of some pre-specified size.
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-Power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
-By running the calculation for a range of sample sizes, the required sample size to find an assumed true effect with desired statistical power can be determined.
+In line with this, power analysis represents the act of calculating the statistical power for a given true effect and sample size.
 
 <!-- Power simulations are hard! (especially for GLMMs) -->
-Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for results with high evidencial value.
+For a desired statistical power and an assumed true effect, the required sample size to find such an effect can be calculated.
+Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
 While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
 For GLMMs, analytic options to derive sample sizes are not available for general research designs.
-An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @murayamaSummarystatisticsbasedPowerAnalysis2022; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+An increasing number of tutorials describe how to perform power simulation for multilevel models.
 However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies are available to determine plausible values for all model parameters.
-In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
-Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
-Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
-By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science [@yarkoniGeneralizabilityCrisis2022].
+<!-- AK: Here I would provide examples for complex study designs and what can be tested with them -->
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
+<!-- AK: Unclear to me what "data from pilot studies is available to determine plausible values for all model parameters" means -->
+In contrast<!-- AK: UWhy "in contrast". Does not fit well here -->, simulations are often required, necessitating assumptions for all parameters in the model.
+
+Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+<!-- AK: Rewrite.
+You wrote before about the tutorials.
+"Therefore" doe snot fit here.
+Moreover, the link to human-AI interaction is not clear to me.
+Up to this point it seems a bit arbitrary.
+-->
+Many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
+By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science.<!-- AK: This is a little too general.
+This last paragraph addresses the reasons for conducting power analysis.
+"Advancement of science" is a bit too vague.
+-->
 
 <!-- Research Gap & Summary -->
-When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes challenging.
+When planning a human-AI interaction experiment <!-- AK: Is it really needed to have the human-AI interaction focus?
+-->with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes challenging.<!-- AK: Readers might not know what non-standard designs are.
+Either work more with examples or elaborate more on these designs and how your paper addresses these difficulties.
+-->
 Currently available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
@@
 Instead, researchers must acquire the necessary skills based on the specific application at hand.
-In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study for which no meta-analytic results or conclusive data from pilot studies are available.
+In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study for which no meta-analytic results or conclusive data from pilot studies is available.
 Through this case study, we aim to provide guidance and insights into the complexities involved in addressing complex design considerations and obtaining accurate power estimates.
@@
 <!-- overview of this section -->
-In this section, we describe the steps to conduct a data simulation and a power analysis for GLMMs, with a specific experimental design from human-AI interaction research, in the form of a case study.
+In this section, we describe the steps to conduct a data simulation and a power analysis for GLMMs with a specific experimental design in human-AI interaction research in mind in the form of a case study.
 
@@
 In the case study, we simulate the data for a human-AI Interaction experiment where the diagnostic performance of users of a health care AI is to be evaluated.
-The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and emergency doctors (non task experts), view head CT (computertomography) scans and evaluate if an intracranial bleeding is present.
-To support their decision making, an AI provides a suggestion for a medical diagnosis. This AI advice can be either either correct or incorrect. 
-Participants then provide a final medical diagnosis (i.e., bleeding or no bleeding) that can be correct or incorrect. 
-Also, we have control trials, in which no AI advice is present. 
-This introduces some missing values by design (advice is neither correct nor incorrect, when no advice is present) which must be taken into account when simulating and analyzing the data. 
+The goal is to understanding how AI advice influences medical decision-making.
+Participants, radiologists (task experts) and emergency doctors (non task experts) view head CT (computertomography) scans and evaluate if an intracranial bleeding is present.
+To support their decision making, an AI provides a suggestion for a medical diagnosis.
+This AI advice can be either either correct or incorrect.
+Participants' medical diagnosis (i.e., bleeding or no bleeding) can be correct or incorrect. 
+Also, we have a control group with no AI advice present. This introduces missings by design that has an effect on our data simulations. 
 Through this experiment, we want to determine if experts are better than non-experts in reading head CT scans.
 Further, we want to research if correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
-In the present example, it would be more challenging to recruit task expert (i.e., radiologists) as there is only a limited amount of such people. Non-experts would be easier to recruit.
+In the present example, it would be more challenging to recruit task experts (i.e., radiologists) as there is only a limited amount of such people.
+Non-experts would be easier to recruit<!-- AK: Why is that info relevant?
+I would remove this -->.
 The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
@@
 <!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015].
+For our data simulation and data analysis, we use the lme4 R package [@batesFittingLinearMixedEffects2015].
 The lme4 package is a state-of-the-art tool for fitting frequentist GLMMs (for Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017]).
-The lme4 package also includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+The lme4 package includes a useful function called "simulate" that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
 However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.[^simr]
-
-[^simr]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016] which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
+Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.<!-- AK: I think this last 2 sentences are better for the introduction (i.e.
+what and why?) - maybe here stick to the description of the lme4 package.
+I am also not sure if you need to describe the lme4 package in detail.
+I think for the tutorial, understanding the underlying technique (GLMM) is more relevant.
+-->
 
@@
 LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
-They are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables).
+<!-- AK: This is a bit confusing.
+I would restructure this part.
+You do not explain what correlated and hierarchical mean in this context and it is difficult to follow you here.
+-->They are used when the outcome variable is continuous and follow a normal distribution.
 LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
@@
 GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
+They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma <!-- AK: I would briefly explain these distributions -->.
 GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.
+The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
 
@@
 GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
 GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
+This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance <!-- AK: Still not sure why this is particularly relevant in this context.
+These distributions also occur in other research contexts -->.
 By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
@@
 - $advice\_correct_{si}$: whether this advice was correct (1) or not (0)
-- $expert_s$: whether subject $s$ was a radiologist (1) or not (0)
+- $expert_s$: whether subject $s$ was a professional neurologist (1) or not (0)
 
@@
 
-The following R function simulates a full dataset structured according to the desgin of our case study.
-The faux package [@R-faux] contains useful functions when simulating factorial designs including random effects.
+The following R function simulates a full dataset structured according to our case study.
+The faux package contains useful functions when simulating factorial designs including random effects.
 
@@
 Thus the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
-We add a between subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
-The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
-Next we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
+We add a between subject effect with the `add_between` command, simulating that about $25%$ of subjects are experts.
+The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice and if advice is presented, the advice will be correct in about $80%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
+Next we simulate one random effect each subject (`u0s`) and for each item (`u0i`).
 As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with mean 0 and a standard deviation specified by the user.
@@
 The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
-The complete set of rules for the syntax are outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
+The complete set of rules for the syntax are outlined in (CITE LME4 PAPER) and in the documentation of the lme4 package.
 
 In lme4, a GLMM is fitted with the `glmer` function.
-By setting `family =  "binomial"` we request a binomial GLMM, appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default).
+By setting `family =  "binomial"` we request a binomial GLMM, appropriate for our binary dependent variable `y_bin` (the binomial GLMM used the canonical logit link by default).
 
@@
 Unfortunately, the model parameters in a binomial GLMM are hard to interpret, because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item with average difficulty ($u_{0i} = 0$).
+The most simple interpretation works by imagining a subject with average ability ($u0s = 0$) responding to an item with average difficulty ($u0i = 0$).
 Then the model implied probability that such a person solves such an item is given by:
@@
 In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
+The other extreme case would be the probability that a non-expert with average ability solves and item with average difficulty when presented without any advice:
 $$
@@
 
-Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
+Due to this complicated relationship, many experts argue not to focus too much on interpreting single model parameters when working with GLMMs.
+Instead, it can be more intuitive to consider the implied predicted distribution of the dependent variable for each experimental conditions across all subjects and items.
 
-With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition:
+With the marginaleffects package, we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition:
 
@@
 
-However, we need to think about the model parameters again when we want to test hypotheses which we have theoretically derived from some research question.
+However, we need to think about the model parameters again when wanting to test hypotheses which we have theoretically derived from some research question.
 Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.
@@
 
-We can specify and test hypotheses like these with the multcomp package [@R-multcomp] as follows:
+We can specify and test hypotheses like these with the multcomp package as follows:
 
@@
 Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command.
-With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value for hypothesis $H_{03}$ is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
-Note that this decision would be wrong, because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
+With a standard significance level of $\alpha = 0.05$, we would reject all four null hypotheses and therefore also reject the combined null hypothesis for this simulated dataset.
 
@@
 
-When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
+When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where the numbers came from.
 All parameter values have been determined in repeated exchanges with our affiliated domain experts and we here outline a few strategies on how to determine plausible parameter values.
@@
 In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
-For example, we determined the $\beta$ parameters in a way that correspond to a desired probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with a correct advice.
-However even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
+For example, we determined the $\beta$ parameters in a way that correspond to a desired probability, that an expert with average ability solves an item with average difficulty when presented with a correct advice.
+However even if the model were true, we would not observe this probability if we estimated the probability in a group of expert responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
 This implies that one must be careful when specifying parameter values based on previous studies or pilot data.
@@
 The well known inequality of conditional and marginal effects in GLMMs makes their interpretation more difficult, however, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
-When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
+When parameter values have selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
 Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
@@
 Figure X shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
-We can see that indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
+We can see that indeed, the average probabilities (block dots) differ from the probabilities of average subjects and items considered in the previous section.
 This difference increases with the variability of the random effects.
@@
 
-The final plot demonstrates, that these plots are also useful to spot standard deviations that are specified too high.
-For this example, we have set $\sigma_S = 3$ and $\sigma_I = 3$.
-This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
+The final plot demonstrates, that these plots are also useful to spot standard deviations specified to high.
+For example, if we set $\sigma_S = 3$ and $\sigma_I = 3$ this implies that in each experimental condition, the probabilities that a subject solves an item are always close to either 0 or 1, which is not a plausible assumption.
 
@@
 
-With all these considerations out of the way, we are finally ready to perform a power analysis.
+With all these considerations out of the way, we are finally ready to perform a power simulation.
 Wrapping the `simulate` function already introduced earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
@@
 
-Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
-Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
+Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit model for different parameter combinations or sample sizes.
+Here, we use the future and furr packages to perform computations in parallel.
 First, we enable parallelization and specify how many parallel cores of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
@@
 
-The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
+The next code chunk, specifies a simulation design with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
 
@@
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations, in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^faux]
-
-[^faux]: This code is heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code (which is heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>).
 
@@
 In practice, power simulations are often run multiple times with adjusted combinations of sample sizes.
-When run for the first time, it might be revealed that power is way too low (or higher than required) for some combinations of `n_subjects` and `n_items`.
+When run for the first time, it might be revealed that power is way too low (or too high) for some combinations of `n_subjects` and `n_items`.
 When narrowing down the best combination that achieves sufficient power, while at the same time strikes a good balance of how many subjects and items are practically feasible, later rounds of power simulations will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
@@
 When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other, based on practical considerations.
-For the sake of the example, let the targeted power be indeed about 0.8 to detect an effect of the expected size implied by our data simulation.
+For the sake of the example, let the targeted power be indeed about 0.8 to detect an effect of the expected size.
 This could be achieved by collecting data of 200 subjects (about 25% of which will be experts) each completing the same 50 items (with advice present in about 67% of cases which is correct in about 80% of cases with present advice).
 If collecting 200 subjects is not practically feasible, it would also be possible to recruit 150 subjects but increase the length of the experiment to over 70 items.
-However, 70 items might take to long to complete for the doctors participating in the study, who probably all have a busy schedule.
-The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items, if it is feasible to recruit an even higher number of subjects (> 250, to be determined by additional rounds of power analysis).
+However, 70 items might take to long to complete (especially for the expert radiologists who might have a busy schedule).
+The simulation suggests that it might also be possible to plan an even shorter experiment with only 30 items, if it is feasible to recruit an even higher number of subjects (> 250, to be determined by additional rounds of power analysis).
 Design parameters which also affect power and which could be investigated in the simulation to find a more optimal trade-off are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
@@
 Human-AI interaction research requires careful planning and consideration of statistical power to ensure reliable and meaningful results.
-While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
+While heuristics and helper programs can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
 The present tutorial presents a specific example on how to run power simulations for human-AI interaction experiments.
@@
 In our case study, we have performed power simulations based on a single set of parameter values that reflect our assumptions of an expected effect size.
-Instead of extracting this expected effect size from meta-analyses or pilot data which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
-Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
+Instead of extracting this expected effect size from meta-analyses or pilot data which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge which can be considered a vague theoretical model about the data-generating process.
 When sample sizes are chosen based on the results of our power simulations, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.
-However, if the true effect is indeed smaller, power will be lower and the study might not be sufficiently informative.
+However, if the true effect is indeed smaller, power will be lower the study might not be sufficiently informative.
 A common, more conservative strategy for sample size justification is to perform power analysis for a smallest effect size of interest (SESOI). 
@@
 When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called *sensitivity power analysis*.
-By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect, or at least get an impression on how strongly power depends on the size of the true effect.
+By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect, or at least get an impression how strongly power depends on the size of the true effect.
 In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyse power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
@@
 <!-- Outlook -->
-Beyond the specifics of our concrete case study, we want to outline six avenues regarding the outlook on power simulations in human-AI interaction research:
+Beyond the specifics, we want to outline six avenues regarding the outlook on power simulations in human-AI interaction research:
 
 1.
-The need for power simulations in human-AI interaction research: As our case study has clearly pointed out, human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
+The need for power simulations in human-AI interaction research: As this work has clearly pointed out, human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or simple helper programs.
 Power simulations offer a solution by providing Human-AI interaction researchers with a tailored approach to estimating statistical power.
@@
 Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
-This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
+This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need to simulate predictor values, which would introduce additional assumptions.
 By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
@@
 Equipping Human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
-The need to reason about how to simulate plausible data which is in line with the research hypothesis while not violating domain expertise on how plausible data should look like might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
 
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in Human-AI interaction research.
-Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the perceived effort required to perform power simulations and the actual effort estimated by researchers and collaborators in Human-AI interaction research.
+Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these simulations.
 It is crucial to raise awareness about the effort involved in  power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
@@
 Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
-They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
+They offer insights into the reliability and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
 The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports.

### 2023-08-31 docs/manuscript.Rmd
Update manuscript.Rmd
Changed intro and left comments

--- 2023-08-22 docs/manuscript.Rmd
@@
 <!-- General intro, relevance of Human AI Interaction experiments, need for well-powered and analyzed experimental research in this field -->
-With the growing integration of artificial intelligence (AI) into our daily lives and routines, it has become increasingly important to study and understand how humans interact with these intelligent systems.
-This is particularly relevant as AI becomes more sophisticated and is deployed to support decision-making processes in fields such as law and medicine that have a significant impact on individuals and society at large.
-Therefore, thorough experimental research is needed to study human-AI interaction to understand how people perceive and respond to these novel technologies. 
-When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.
+As artificial intelligence (AI) becomes more and more ubiquitous in our everyday lives and routines, it has become increasingly important to study and understand how humans interact with these intelligent systems.
+This is particularly relevant as AI models are getting more sophisticated and are deployed to support decision-making processes in areas such as law and medicine that have significant implications for individuals and society at large.
+Therefore, thorough research is needed to study the interaction between humans and AI technology to understand how people perceive, respond to and are influenced by these novel technologies. 
+<!-- SG: I think that defining a problem is missing here.
+It could be that most computer scientists developing the models are not well-trained in experimental design, which results in many papers on human-AI interaction being poorly designed and executed.
+This is problematic as it waists resources and might even be dangerous when readers rely on results from "bad" research.
+-->
+When planning experimental research in this emerging field, it is essential to determine an adequate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative.
+Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for answering their research questions and analysing the data.
+<!-- SG: I think the second point, "and use appropriate statistical models to analyze the data." needs further explanation.
+I tried to change it, but I am not sure if this is what you actually wanted to say-->
+<!-- SG: Overall, I am not entirely sure if the last two sentences are super well placed here since the benefits of performing a power analysis will be repeated below.
+Please ignore this comment, when you think otherise-->
 
 <!-- Some basics on statistical power, power analysis and why it is important -->
-In empirical research relying on hypothesis testing, the most common strategy to determine appropriate samples size is based on statistical power. 
-Statistical power is defined as the probability that a hypothesis test of interest turns out significant, when analyzing repeated samples from a population with a true effect of some pre-specified size.
+In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power. 
+Statistical power is defined as the probability that a hypothesis test of interest <!-- is the "of interest" needed?
+--> turns out significant when analyzing repeated samples from a population with a true effect of some pre-specified size.
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-Power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
-By running the calculation for a range of sample sizes, the required sample size to find an assumed true effect with desired statistical power can be determined.
+A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
+By running the calculation for a range of sample sizes, the required sample size can be determined so that researchers to find an assumed true effect with the desired statistical power.
 
 <!-- Power simulations are hard! (especially for GLMMs) -->
-Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for results with high evidencial value.
-While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
+Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
+<!-- SG: I am not 100% sure what the added benefit of this sentence is.
+It is pretty much what has been said in the previous sentence without the results with high evidential value".
+I would just add the last part to the previous sentence and then delete this one.
+-->
+While user-friendly power analysis software is available for simple statistical models like t-tests, ANOVA, and linear regression, there are few tools are available for more sophisticated statistical methods.
+However, researchers studying human-AI interaction are increasingly using (generalized) linear mixed models (GLMMs; also called multilevel models), which add complexity to power analysis.
 For GLMMs, analytic options to derive sample sizes are not available for general research designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
-An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies are available to determine plausible values for all model parameters.
+<!-- SG: I think I do not fully understand this sentence; do you mean interface-based software, or are you trying to focuse more on the flexibility of the research design/model specification?
+-->
+A number of tutorials have been published describing how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs.
+This narrow focus provides little guidance for researchers faced with more complex study designs, which are very common [@brownIntroductionLinearMixedEffects2021].
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
 In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].

### 2023-09-01 docs/manuscript.Rmd
Update manuscript.Rmd
Changes and comments

@@
 Therefore, thorough research is needed to study the interaction between humans and AI technology to understand how people perceive, respond to and are influenced by these novel technologies. 
-<!-- SG: I think that defining a problem is missing here.
+<!-- SG: Defining a problem is missing here.
 It could be that most computer scientists developing the models are not well-trained in experimental design, which results in many papers on human-AI interaction being poorly designed and executed.
@@
 I tried to change it, but I am not sure if this is what you actually wanted to say-->
-<!-- SG: Overall, I am not entirely sure if the last two sentences are super well placed here since the benefits of performing a power analysis will be repeated below.
-Please ignore this comment, when you think otherise-->
+<!-- SG: Overall, I am not sure if the last two sentences are super well placed here since the benefits of performing a power analysis will be repeated below.
+Please ignore this comment when you think otherwise -->
 
@@
 Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
+However, in the novel field of human-AI interaction, limited prior studies offer little guidance, rendering estimation a practical challenge.
+<!-- SG: I guess the content of the sentence I added was implied, but I think it would be useful to add, to highlight the challenges.
+-->
 In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
+<!-- SG: I don't totally get the "in contrast".
+Unfortunately, I don't know if I get the entire argument.
+Before, you said that other tutorials expect you to have a good assumption for all model parameters, and now you say, in contrast, simulations are needed that require assumptions for all model parameters.
+What is the difference between the two statements, and why are they contrasting?
+-->
 Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+<!-- SG: Unfortunately, again, I do not follow the argument structure.
+How do we get from it being hard to estimate parameters to why we need sample size calculations for the experimental design?
+As far as I understand, we also needed parameter estimates to run this sample size calculation.
+Also, I don't think this sentence adds any benefit since it has already been said twice before.
+-->
 Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
 By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science [@yarkoniGeneralizabilityCrisis2022].
+<!-- SG: Unclear to the reader why incorporating GLMMS is enhancing replicability and credibility.
+Either you need a section above that describes the benefits of GLMMs, or it should be removed-->
+<!-- SG: What would really help the argument is that while conducting a proper power analysis, you have to consider every aspect of your experimental design.
+This really makes you think through everything carefully and might highlight statistical or design challenges before the start of an experiment.
+This gives the researcher a great opportunity to improve the design.
+Especially in fast-moving research fields, where speed is often regarded as more important than rigour, adding a proper sample size calculation into the process can act as a safeguard for ensuring high-quality research.
+A lot of this is implied, but it doesn't come through clearly.-->
 
 <!-- Research Gap & Summary -->
-When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes challenging.
-Currently available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
+When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and "missing by design" conditions, the process of power simulations becomes challenging.
+Available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
+<!-- SG: Hmm, both sentences above are basically a repetition from the previous sections with a bit more details.
+I wonder if we could restructure the introduction to avoid repetitions.
+My suggestion would be:
+1. AI technology is becoming ubiquitous, and we need to understand how humans interact with it. 
+2. Lots of people conduct research in this field; however, since many are not properly trained in experimental design, there is a lot of low-quality research out there, which might pose a risk. 
+3. One of the problems is that a lot of these studies are underpowered with way too many conditions and way too few participants
+4. Background information on why a properly powered study is important, including definitions of statistical power 
+5.
+Introducing the concept of power analysis and sample size calculation as a solution for the issue of underpowered low-quality research and highlighting the additional benefits of doing a power analysis
+6. Describing the challenges of performing a power analysis for complex research designs 
+7. Offering this tutorial as a way to overcome the above-described challenges. 
+I think ´that strictly sticking to a storyline like this would help to avoid repetition.-->
 Unfortunately, there are no general recommendations that can be applied universally in such cases.
@@
 In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study for which no meta-analytic results or conclusive data from pilot studies are available.
-Through this case study, we aim to provide guidance and insights into the complexities involved in addressing complex design considerations and obtaining accurate power estimates.
+<!-- SG: I am curious about the latter considering that we took both of our prior studies apart to get parameter estimates.
+-->
+Through this case study, we aim to provide guidance and insights into the complexities involved in addressing sophisticated design considerations and obtaining accurate power estimates.
+<!-- SG: sample size estimates?
+This was the main goal, right?
+-->
 
@@
 <!-- overview of this section -->
-In this section, we describe the steps to conduct a data simulation and a power analysis for GLMMs, with a specific experimental design from human-AI interaction research, in the form of a case study.
+In this section, we outline the steps for performing data simulation and power analysis for GLMMs, using a case study based on a specific experimental design from human-AI interaction research.
 
 ## The present case study
+<!-- SG: I am wondering if this is the perfect place for this section.
+By the end of the end of "basics of power simulation in GLMMs", readers have probably already forgotten the experimental design of the case study-->
+<!-- The present example, describe the study idea and design -->
+In the case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
+The goal is to understand how AI advice influences medical decision-making. 
+Participants, radiologists (task experts) and students/interns (non-task experts), review 50 head CT (computer tomography) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
+To support their decision-making, in two experimental conditions, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
+This AI advice can be either either correct or incorrect.
+In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
+After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
+Each participant undergoes all experimental conditions, separated by a two-week washout period to prevent scan recall.
 
-<!-- The present example, describe the study idea and design -->
-In the case study, we simulate the data for a human-AI Interaction experiment where the diagnostic performance of users of a health care AI is to be evaluated.
-The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and emergency doctors (non task experts), view head CT (computertomography) scans and evaluate if an intracranial bleeding is present.
-To support their decision making, an AI provides a suggestion for a medical diagnosis. This AI advice can be either either correct or incorrect. 
-Participants then provide a final medical diagnosis (i.e., bleeding or no bleeding) that can be correct or incorrect. 
-Also, we have control trials, in which no AI advice is present. 
-This introduces some missing values by design (advice is neither correct nor incorrect, when no advice is present) which must be taken into account when simulating and analyzing the data. 
-Through this experiment, we want to determine if experts are better than non-experts in reading head CT scans.
-Further, we want to research if correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
-In the present example, it would be more challenging to recruit task expert (i.e., radiologists) as there is only a limited amount of such people. Non-experts would be easier to recruit.
-The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
+This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
+With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans, (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice, and (c) different presentations of AI advice influence performance.
+In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible.
+The aim of the power simulation is to ascertain the required number of task experts and non-experts for achieving adequate statistical power in our experiment.
 
 ## The lme4 package in R
+<!-- SG: I was wondering why you introduced the tool before the statistical model? -->
 
@@
 
-The lme4 package also includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
-However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.[^simr]
+The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
+<!-- SG: Do you think it is clear to readers what a population model is?
+-->
+Therefore in this tutorial, we simulate data for GLMMs from first principles <!-- SG: I do not know what "first principles" means in this context.
+This might be true for some readers as well--> to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.[^simr]
 
-[^simr]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016] which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
+[^simr]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
 
 ## Generalized linear mixed models
-
 <!-- GLMM intro -->
@@
 They are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables).
-LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
+LMMs allow for the modelling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
 Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
 GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
+They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
 GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.
+The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.
 
 <!-- Relevance of GLMMs -->
-GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
-By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+<!-- SG: Auch hier hat sich wieder viel wiederholt. Ich habe mal versucht zu kürzen -->
+As the complexity of studying human interactions with AI systems grows, researchers require advanced statistical models to effectively capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+The above-described flexibility of GLMMs is particularly suitable for investigating various aspects of human-AI interaction, such as user preferences, trust, engagement, and performance.
+By incorporating GLMMs into their analyses, researchers can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioural aspects involved.
 
@@
 <!-- deep dive into our specific example -->
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021].
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modelled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021].
 $$
@@
 $$
-Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$.
+Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$.
 $$

### 2023-09-01 docs/manuscript.Rmd
Update manuscript.Rmd

@@
 To support their decision-making, in two experimental conditions, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-This AI advice can be either either correct or incorrect.
+This AI advice can be either either correct (80% of cases) or incorrect (20%).
 In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
@@
 $$
-If the dependent variable $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability:
+If the dependent variable (i.e., performance) $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability:
 $$
@@
 $$
-In our case study, $P_{si}$ is the conditional probability that subject $s$ gives the correct response to item $i$.
+In our case study, $P_{si}$ is the conditional probability that subject (i.e., participant) $s$ gives the correct response to item (i.e., CT scan) $i$.
 
@@
 
-In our case study, the probability to give a correct response is assumed to depend on the predictors:
-
-- $advice\_present_{si}$: whether subject $s$ was presented with algorithmic advice (1) or not (0) when asked to asses item $i$
+In our case study, the probability of giving a correct response is assumed to depend on the predictors:
+<!-- SG: What happened to the different experimental conditions? AI advice had three levels: Advice with annotation, without annotation, and no advice-->
+- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to asses item $i$
 - $advice\_correct_{si}$: whether this advice was correct (1) or not (0)
-- $expert_s$: whether subject $s$ was a radiologist (1) or not (0)
+- $expert_s$: whether subject $s$ was a task expert (1) or not (0)
 
@@
 
-- $u_{0s}$: the deviation of suject $s$ from the average ability to solve an item with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
-- $u_{0i}$: the deviation of item $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$
+- $u_{0s}$: the deviation of subject $s$ from the average ability to solve an item (i.e., CT scan) with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
+- $u_{0i}$: the deviation of item (i.e., CT scan) $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$
 
@@
 
-The following R function simulates a full dataset structured according to the desgin of our case study.
-The faux package [@R-faux] contains useful functions when simulating factorial designs including random effects.
+The following R function simulates a full dataset structured according to the design of our case study.
+The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.
 
@@
 ```
-
-In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (`n_items` in total).
+<!-- SG: You basically start the explanation at line 257; I think it would be better to actually start at the beginning of the function line 251, otherwise, this might be very confusing to people.
+If you do not want to change the structure, then I would at least mention that you discuss the first section of the function below.-->
+In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan) (`n_items` in total).
 Thus the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
-We add a between subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
-The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
-Next we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
-As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with mean 0 and a standard deviation specified by the user.
+We add a between-subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
+The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
+Next, we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
+As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user.
 With all design variables done, we are ready to simulate our model equation as outlined in equation X.
-The linear predictor variable `linpred` ($\eta$ in the GLMM model equations), combines the predictor variables, random effects and model parameters as assumed by our model.
-We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
-In the final step, we simulate the binary dependent variable `y_bin` by -- for each trial -- drawing from a bernoulli distribution with success probability `y_prob`.
+The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects and model parameters as assumed by our model.
+We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
+In the final step, we simulate the binary dependent variable `y_bin` by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
 
@@
 
-The first two lines looks similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`).
+The first two lines look similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`).
 The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
-The complete set of rules for the syntax are outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
+The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
 
 In lme4, a GLMM is fitted with the `glmer` function.
-By setting `family =  "binomial"` we request a binomial GLMM, appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default).
+By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate vs.
+inaccurate diagnosis.
 
@@
 
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret, because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item with average difficulty ($u_{0i} = 0$).
-Then the model implied probability that such a person solves such an item is given by:
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modelled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
+Then the model implied probability that such a person solves such an item accurately is given by:
 $$
@@
 
-However, we need to think about the model parameters again when we want to test hypotheses which we have theoretically derived from some research question.
+However, we need to think about the model parameters again when we want to test hypotheses that we have theoretically derived from some research question.
 Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.
@@
 For our case study, imagine the following combined hypothesis:
-*We expect that for both experts and non-experts, correct advice leads to a higher probability to solve an item compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability to solve an item compared to no advice presented.*
+*We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
 
@@
 We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$:
-The first null hypothesis states that *an expert responding to an item while presented with correct advice has lower or equal probability to solve the item compared to the same expert facing the same item without any advice.*
+The first null hypothesis states that *an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.*
 This implies the following inequality for each subject $s$ and item $i$
@@
 intersection-union setting\; @dmitrienkoTraditionalMultiplicityAdjustment2013].
-In such cases, the error probabilities do not accumulate and we would waste power when correcting for multiple testing.
+In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
 Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command.
 With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value for hypothesis $H_{03}$ is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
-Note that this decision would be wrong, because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
+Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
 
@@
 When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
-All parameter values have been determined in repeated exchanges with our affiliated domain experts and we here outline a few strategies on how to determine plausible parameter values.
+All parameter values have been determined in repeated exchanges with our affiliated domain experts, and we now outline a few strategies on how to determine plausible parameter values.
 
-We have already seen in our discussion of model interpretation, how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
+We have already seen in our discussion of model interpretation how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
 We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (for an average subject and an average item).
@@
 In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
-For example, we determined the $\beta$ parameters in a way that correspond to a desired probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with a correct advice.
-However even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
-This implies that one must be careful when specifying parameter values based on previous studies or pilot data.
+For example, we determined the $\beta$ parameters in a way that corresponds to a desired probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
+c
+However, even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
+This implies that one must be careful when specifying parameter values based on previous studies or pilot data. <!-- SG: What does that mean for someone who wants to do this then?-->
 
-The well known inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult, however, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
+The inequality <!-- SG: Can you explain what this means?--> of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
+However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
 When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
@@
 Figure X shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
-We can see that indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
+We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
 This difference increases with the variability of the random effects.
 
-In fact, up to this point we have not talked about plausible values for the standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$).
-Plots like the one above are a useful tool to decide whether the specified standard deviations are reasonable, by comparing the ranges and overlap between conditions to domain knowledge.
+Up to this point, we have not talked about plausible values for the standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$).
+Plots like the one above are a useful tool to decide whether the specified standard deviations are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
 
@@
 As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts:
-With wrong advice, virtually all experts have a higher probability to make a correct diagnosis compared to non-experts when considering only items with average difficulty.
+With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
 In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
-Patterns like these should be considered carefully, discussed with the domain experts and parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
+Patterns like these should be considered carefully and discussed with the domain experts.
+Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
 
-We could also have a closer look at variability between items, by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$, see Figure X in Appendix X).
+We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$, see Figure X in Appendix X).
 
@@
 
-The final plot demonstrates, that these plots are also useful to spot standard deviations that are specified too high.
+The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
 For this example, we have set $\sigma_S = 3$ and $\sigma_I = 3$.
 This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
+<!-- SG: Maybe explain why this is not plausible.-->
 
@@
 ## Power simulation
-
-With all these considerations out of the way, we are finally ready to perform a power analysis.
+With all these considerations addressed, we are finally ready to perform a power analysis. 
 Wrapping the `simulate` function already introduced earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
@@
 
-Power analysis can quickly become computationally intensive, when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
+Power analysis can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
 Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
 First, we enable parallelization and specify how many parallel cores of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
+<!-- SG: Are workers the cores you want to use?-->
 
@@
 The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
+<!-- SG: Can you explain how you choose the repetitions?-->
 
@@
 
-For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations, in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
+For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
 Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^faux]
@@
 
-<!-- How many participants / stimuli are needed for concrete example -->
-As should be the case, power estimates increase with both the number ob subjects and the number of items.
+<!-- How many participants/stimuli are needed for concrete example -->
+As should be the case, power estimates increase with both the number of subjects and the number of items.
 The confidence intervals indicate how precisely power was estimated by our simulation.
@@
 In practice, power simulations are often run multiple times with adjusted combinations of sample sizes.
-When run for the first time, it might be revealed that power is way too low (or higher than required) for some combinations of `n_subjects` and `n_items`.
-When narrowing down the best combination that achieves sufficient power, while at the same time strikes a good balance of how many subjects and items are practically feasible, later rounds of power simulations will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
-This will assure high precision for the final power estimates which are then used for sample size justification of the future study.
+When running for the first time, it might be revealed that power is way too low (or higher than required) for some combinations of `n_subjects` and `n_items`.
+When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of power simulations will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
+This will assure high precision for the final power estimates, which are then used for sample size justification for the future study.
 
@@
 The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
-When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other, based on practical considerations.
+<!-- Maybe give an example?-->
+When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
 For the sake of the example, let the targeted power be indeed about 0.8 to detect an effect of the expected size implied by our data simulation.
-This could be achieved by collecting data of 200 subjects (about 25% of which will be experts) each completing the same 50 items (with advice present in about 67% of cases which is correct in about 80% of cases with present advice).
-If collecting 200 subjects is not practically feasible, it would also be possible to recruit 150 subjects but increase the length of the experiment to over 70 items.
-However, 70 items might take to long to complete for the doctors participating in the study, who probably all have a busy schedule.
-The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items, if it is feasible to recruit an even higher number of subjects (> 250, to be determined by additional rounds of power analysis).
-Design parameters which also affect power and which could be investigated in the simulation to find a more optimal trade-off are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
+This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases which is correct in about 80% of cases with present advice).
+If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items.
+However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule.
+The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (> 250, to be determined by additional rounds of power analysis).
+Design parameters that also affect power and which could be investigated in the simulation to find a more optimal trade-off are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
 
@@
 
-<!-- short summary of key take aways from this tutorial -->
+<!-- short summary of key takeaways from this tutorial -->
 Human-AI interaction research requires careful planning and consideration of statistical power to ensure reliable and meaningful results.
 While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
-The present tutorial presents a specific example on how to run power simulations for human-AI interaction experiments.
+The present tutorial presents a specific example of how to run power simulations for human-AI interaction experiments, which can also applied to other fields of research.
 
@@
 In our case study, we have performed power simulations based on a single set of parameter values that reflect our assumptions of an expected effect size.
-Instead of extracting this expected effect size from meta-analyses or pilot data which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
+Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
+<!-- I would try to highlight this more clearly in the methods part.
+-->
 Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
 When sample sizes are chosen based on the results of our power simulations, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.
-However, if the true effect is indeed smaller, power will be lower and the study might not be sufficiently informative.
-A common, more conservative strategy for sample size justification is to perform power analysis for a smallest effect size of interest (SESOI). 
+However, if the true effect is indeed smaller, the power will be lower, and the study might not be sufficiently informative.
+A common, more conservative strategy for sample size justification is to perform power analysis for the smallest effect size of interest (SESOI). 
 An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011].
-For strategies on the even more difficult task of specifying a plausible SESOI as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].
+For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].
 When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called *sensitivity power analysis*.
-By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect, or at least get an impression on how strongly power depends on the size of the true effect.
+By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power depends on the size of the true effect.
 In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyse power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
-However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
-In this setting, sensitivity power analysis would require to manually specify additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
-Power could then be simulated for several of these scenarios (across different number of subjects and items as considered earlier).
+However, for our case study that investigates combined hypotheses in a GLMM modelling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
+In this setting, sensitivity power analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
+Power could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
 
@@
 1.
-The need for power simulations in human-AI interaction research: As our case study has clearly pointed out, human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
-Power simulations offer a solution by providing Human-AI interaction researchers with a tailored approach to estimating statistical power.
+The need for power simulations in human-AI interaction research: Human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
+Power simulations offer a solution by providing human-AI interaction researchers with a tailored approach to estimating statistical power.
 These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
@@
 Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
-This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
+This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
 By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
@@
 3.
-Teaching power simulation skills to human-AI interaction researchers: The ability to conduct  power simulations is a valuable skill that should be taught to Human-AI interaction researchers.
+Teaching power simulation skills: The ability to conduct power simulations is a valuable skill that should be taught to human-AI interaction researchers.
 By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping Human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
-The need to reason about how to simulate plausible data which is in line with the research hypothesis while not violating domain expertise on how plausible data should look like might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
+Equipping researchers with the knowledge and tools to perform power simulations enables them to make informed decisions and enhance the rigour of their studies.
+The need to reason about how to simulate plausible data that is in line with the research hypothesis while not violating domain expertise on how plausible data should look might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
 
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in Human-AI interaction research.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in human-AI interaction research.
 Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-It is crucial to raise awareness about the effort involved in power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
+<!-- SG: Hahah Timo, do you want to tell me something =) --> It is crucial to raise awareness about the effort involved in power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
 
 5.
-Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
+Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in human-AI interaction research.
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
-The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports.
+The importance of power simulations can be reflected by allocating them to a separate publication or incorporating them as a significant component of stage 1 preregistered reports.
 
 6.
-Integration with Open Science and preregistration practices: Power simulations align well with the principles of open science and preregistration in Human-AI interaction research.
+Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in human-AI interaction research.
 When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of Human-AI interaction research.
+By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of research.
 
@@
 
-Power simulations play a critical role in human-AI interaction research, allowing researchers to tailor power estimation to the unique aspects of their experiments.
+Power simulations play a critical role in human-AI interaction research, allowing investigators to tailor power estimation to the unique aspects of their experiments.
 Through this tutorial, we aim to provide researchers with the necessary skills and tools to perform these simulations themselves.

### 2023-09-04 docs/manuscript.Rmd
made some minor changes to manuscript and updated bib file

--- 2023-08-31 docs/manuscript.Rmd
@@
     #address       : "Postal address"
-    email         : "timo.koch@psy.lmu.de"
+    email         : "timo.koch@unisg.ch"
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
   - id            : "4"
-    institution   : "University College London, Global Business School for Health"
+    institution   : "University College London"
 
@@
   Understanding how humans interact with artificial intelligence (AI) through experiments becomes increasingly relevant.
-However, human-AI interaction researchers often lack the appropriate tools to conduct power and data analyses for the complex study designs required.
-In this work, we provide a comprehensive guide to performing customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
+However, human-AI interaction researchers lack the appropriate tools to conduct power and data analyses for the required complex study designs.
+In this work, we provide a tutorial on how to run customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
 By providing code in a specific case study, we equip human-AI interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
@@
 <!-- General intro, relevance of Human AI Interaction experiments, need for well-powered and analyzed experimental research in this field -->
-As artificial intelligence (AI) becomes more and more ubiquitous in our everyday lives and routines, it has become increasingly important to study and understand how humans interact with these intelligent systems.
-This is particularly relevant as AI models are getting more sophisticated and are deployed to support decision-making processes in areas such as law and medicine that have significant implications for individuals and society at large.
-Therefore, thorough research is needed to study the interaction between humans and AI technology to understand how people perceive, respond to and are influenced by these novel technologies. 
-<!-- SG: I think that defining a problem is missing here.
-It could be that most computer scientists developing the models are not well-trained in experimental design, which results in many papers on human-AI interaction being poorly designed and executed.
-This is problematic as it waists resources and might even be dangerous when readers rely on results from "bad" research.
--->
-When planning experimental research in this emerging field, it is essential to determine an adequate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative.
-Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for answering their research questions and analysing the data.
-<!-- SG: I think the second point, "and use appropriate statistical models to analyze the data." needs further explanation.
-I tried to change it, but I am not sure if this is what you actually wanted to say-->
-<!-- SG: Overall, I am not entirely sure if the last two sentences are super well placed here since the benefits of performing a power analysis will be repeated below.
-Please ignore this comment, when you think otherise-->
+As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems. 
+This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
+Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies. 
+When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.
 
 <!-- Some basics on statistical power, power analysis and why it is important -->
-In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power. 
-Statistical power is defined as the probability that a hypothesis test of interest <!-- is the "of interest" needed?
---> turns out significant when analyzing repeated samples from a population with a true effect of some pre-specified size.
+In empirical research relying on hypothesis testing, the most common strategy to determine appropriate samples size is based on statistical power. 
+Statistical power is defined as the probability that a hypothesis test of interest turns out significant, when analyzing repeated samples from a population with a true effect of some pre-specified size.
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
-By running the calculation for a range of sample sizes, the required sample size can be determined so that researchers to find an assumed true effect with the desired statistical power.
+Power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
+By running the calculation for a range of sample sizes, the required sample size to find an assumed true effect with desired statistical power can be determined.
 
 <!-- Power simulations are hard! (especially for GLMMs) -->
-Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
-<!-- SG: I am not 100% sure what the added benefit of this sentence is.
-It is pretty much what has been said in the previous sentence without the results with high evidential value".
-I would just add the last part to the previous sentence and then delete this one.
--->
-While user-friendly power analysis software is available for simple statistical models like t-tests, ANOVA, and linear regression, there are few tools are available for more sophisticated statistical methods.
-However, researchers studying human-AI interaction are increasingly using (generalized) linear mixed models (GLMMs; also called multilevel models), which add complexity to power analysis.
+Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for results with high evidencial value.
+While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
 For GLMMs, analytic options to derive sample sizes are not available for general research designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
-<!-- SG: I think I do not fully understand this sentence; do you mean interface-based software, or are you trying to focuse more on the flexibility of the research design/model specification?
--->
-A number of tutorials have been published describing how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs.
-This narrow focus provides little guidance for researchers faced with more complex study designs, which are very common [@brownIntroductionLinearMixedEffects2021].
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
+An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies are available to determine plausible values for all model parameters.
 In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
@@
 <!-- Relevance of GLMMs -->
-GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
+GLMMs are gaining increasing popularity in the field of human-AI interaction research.
 As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
 GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance.
-By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+This versatility makes GLMMs particularly suitable for investigating various aspects of human-AI interaction, such as user preferences, trust, engagement, and performance.
+By incorporating GLMMs into their analyses, researchers in the field of human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
 
@@
 Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
-This allows Human-AI interaction researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
+This allows human-AI interaction researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
 By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
@@
 3.
-Teaching power simulation skills to human-AI interaction researchers: The ability to conduct  power simulations is a valuable skill that should be taught to Human-AI interaction researchers.
+Teaching power simulation skills to human-AI interaction researchers: The ability to conduct  power simulations is a valuable skill that should be taught to human-AI interaction researchers.
 By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping Human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
+Equipping human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
 The need to reason about how to simulate plausible data which is in line with the research hypothesis while not violating domain expertise on how plausible data should look like might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
@@
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in Human-AI interaction research.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in human-AI interaction research.
 Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
@@
 5.
-Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in Human-AI interaction research.
+Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in human-AI interaction research.
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
@@
 6.
-Integration with Open Science and preregistration practices: Power simulations align well with the principles of open science and preregistration in Human-AI interaction research.
+Integration with Open Science and preregistration practices: Power simulations align well with the principles of open science and preregistration in human-AI interaction research.
 When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.

### 2023-09-04 docs/manuscript.Rmd

@@
 This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
+
 Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies.
+<!-- AK: A bit out of context - not clear to me why these experiments are needed.
+Maybe you could provide some examples what can be achieved with experiments in that context.
+-->
 When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.
+<!-- AK: refs are generally missing - will they be added later?
+-->
 
 <!-- Some basics on statistical power, power analysis and why it is important -->
-In empirical research relying on hypothesis testing, the most common strategy to determine appropriate samples size is based on statistical power. 
+In empirical research relying on hypothesis testing, the most common strategy to determine appropriate sample size is based on statistical power. 
 Statistical power is defined as the probability that a hypothesis test of interest turns out significant, when analyzing repeated samples from a population with a true effect of some pre-specified size.
@@
 <!-- Power simulations are hard! (especially for GLMMs) -->
+
+For a desired statistical power and an assumed true effect, the required sample size to find such an effect can be calculated.
+Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
+While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
+For GLMMs, analytic options to derive sample sizes are not available for general research designs.
+An increasing number of tutorials describe how to perform power simulation for multilevel models.
+However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
+<!-- AK: Here I would provide examples for complex study designs and what can be tested with them -->
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
+<!-- AK: Unclear to me what "data from pilot studies is available to determine plausible values for all model parameters" means -->
+In contrast<!-- AK: UWhy "in contrast". Does not fit well here -->, simulations are often required, necessitating assumptions for all parameters in the model.
+
+Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+<!-- AK: Rewrite.
+You wrote before about the tutorials.
+"Therefore" doe snot fit here.
+Moreover, the link to human-AI interaction is not clear to me.
+Up to this point it seems a bit arbitrary.
+-->
+Many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
+By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science.
+
+<!-- AK: This is a little too general. This last paragraph addresses the reasons for conducting power analysis. "Advancement of science" is a bit too vague. -->
 Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for results with high evidencial value.
@@
 <!-- Research Gap & Summary -->
-When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes challenging.
+When planning a human-AI interaction experiment <!-- AK: Is it really needed to have the human-AI interaction focus?
+-->with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes challenging.<!-- AK: Readers might not know what non-standard designs are.
+Either work more with examples or elaborate more on these designs and how your paper addresses these difficulties.
+-->
 Currently available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
@@
 Further, we want to research if correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
-In the present example, it would be more challenging to recruit task expert (i.e., radiologists) as there is only a limited amount of such people. Non-experts would be easier to recruit.
+
+In the present example, it would be more challenging to recruit task experts (i.e., radiologists) as there is only a limited amount of such people.
+Non-experts would be easier to recruit<!-- AK: Why is that info relevant?
+I would remove this -->.
+The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
 The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
@@
 However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.[^simr]
+
+Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.<!-- AK: I think this last 2 sentences are better for the introduction (i.e.
+what and why?) - maybe here stick to the description of the lme4 package.
+I am also not sure if you need to describe the lme4 package in detail.
+I think for the tutorial, understanding the underlying technique (GLMM) is more relevant.
+-->
 
@@
 <!-- GLMM intro -->
+
+LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
+LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
+<!-- AK: This is a bit confusing.
+I would restructure this part.
+You do not explain what correlated and hierarchical mean in this context and it is difficult to follow you here.
+-->They are used when the outcome variable is continuous and follow a normal distribution.
+LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
+GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
+They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma <!-- AK: I would briefly explain these distributions -->.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
+
+<!-- Relevance of GLMMs -->
+GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
+GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
+This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance <!-- AK: Still not sure why this is particularly relevant in this context.
+These distributions also occur in other research contexts -->.
+By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+
 LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects [@fahrmeirRegressionModelsMethods2021].

### 2023-09-04 docs/manuscript.Rmd

--- 2023-09-01 docs/manuscript.Rmd
@@
 abstract: |
+
   Understanding how humans interact with artificial intelligence (AI) through experiments becomes increasingly relevant.
@@
 <!-- General intro, relevance of Human AI Interaction experiments, need for well-powered and analyzed experimental research in this field -->
+
 As artificial intelligence (AI) becomes more and more ubiquitous in our everyday lives and routines, it has become increasingly important to study and understand how humans interact with these intelligent systems.
@@
 --> turns out significant when analyzing repeated samples from a population with a true effect of some pre-specified size.
+
+As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems. 
+This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
+
+Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies.
+<!-- AK: A bit out of context - not clear to me why these experiments are needed.
+Maybe you could provide some examples what can be achieved with experiments in that context.
+-->
+When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.
+<!-- AK: refs are generally missing - will they be added later?
+-->
+
+<!-- Some basics on statistical power, power analysis and why it is important -->
+In empirical research relying on hypothesis testing, the most common strategy to determine appropriate sample size is based on statistical power. 
+Statistical power is defined as the probability that a hypothesis test of interest turns out significant, when analyzing repeated samples from a population with a true effect of some pre-specified size.
+
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
@@
 <!-- Power simulations are hard! (especially for GLMMs) -->
+
 Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
@@
 -->
+
+For a desired statistical power and an assumed true effect, the required sample size to find such an effect can be calculated.
+Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
+While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
+For GLMMs, analytic options to derive sample sizes are not available for general research designs.
+An increasing number of tutorials describe how to perform power simulation for multilevel models.
+However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
+<!-- AK: Here I would provide examples for complex study designs and what can be tested with them -->
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
+<!-- AK: Unclear to me what "data from pilot studies is available to determine plausible values for all model parameters" means -->
+In contrast<!-- AK: UWhy "in contrast". Does not fit well here -->, simulations are often required, necessitating assumptions for all parameters in the model.
+
+Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+<!-- AK: Rewrite.
+You wrote before about the tutorials.
+"Therefore" doe snot fit here.
+Moreover, the link to human-AI interaction is not clear to me.
+Up to this point it seems a bit arbitrary.
+-->
+Many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
+By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science.
+
+<!-- AK: This is a little too general. This last paragraph addresses the reasons for conducting power analysis. "Advancement of science" is a bit too vague. -->
+Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for results with high evidencial value.
+While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
+For GLMMs, analytic options to derive sample sizes are not available for general research designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
+An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies are available to determine plausible values for all model parameters.
+In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
+Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+
 Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
@@
 <!-- Research Gap & Summary -->
+
 When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and "missing by design" conditions, the process of power simulations becomes challenging.
@@
 I think ´that strictly sticking to a storyline like this would help to avoid repetition.-->
+
+When planning a human-AI interaction experiment <!-- AK: Is it really needed to have the human-AI interaction focus?
+-->with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes challenging.<!-- AK: Readers might not know what non-standard designs are.
+Either work more with examples or elaborate more on these designs and how your paper addresses these difficulties.
+-->
+Currently available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
+
 Unfortunately, there are no general recommendations that can be applied universally in such cases.
@@
 <!-- The present example, describe the study idea and design -->
+
 In the case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
@@
 The aim of the power simulation is to ascertain the required number of task experts and non-experts for achieving adequate statistical power in our experiment.
+
+In the case study, we simulate the data for a human-AI Interaction experiment where the diagnostic performance of users of a health care AI is to be evaluated.
+The goal is to understand how AI advice influences medical decision-making.
+Participants, radiologists (task experts) and emergency doctors (non task experts), view head CT (computertomography) scans and evaluate if an intracranial bleeding is present.
+To support their decision making, an AI provides a suggestion for a medical diagnosis. This AI advice can be either either correct or incorrect. 
+Participants then provide a final medical diagnosis (i.e., bleeding or no bleeding) that can be correct or incorrect. 
+Also, we have control trials, in which no AI advice is present. 
+This introduces some missing values by design (advice is neither correct nor incorrect, when no advice is present) which must be taken into account when simulating and analyzing the data. 
+Through this experiment, we want to determine if experts are better than non-experts in reading head CT scans.
+Further, we want to research if correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
+
+In the present example, it would be more challenging to recruit task experts (i.e., radiologists) as there is only a limited amount of such people.
+Non-experts would be easier to recruit<!-- AK: Why is that info relevant?
+I would remove this -->.
+The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
+The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
 
@@
 
+The lme4 package also includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
+
+Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.<!-- AK: I think this last 2 sentences are better for the introduction (i.e.
+what and why?) - maybe here stick to the description of the lme4 package.
+I am also not sure if you need to describe the lme4 package in detail.
+I think for the tutorial, understanding the underlying technique (GLMM) is more relevant.
+-->
+
 [^simr]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
@@
 <!-- Relevance of GLMMs -->
+
 <!-- SG: Auch hier hat sich wieder viel wiederholt. Ich habe mal versucht zu kürzen -->
@@
 By incorporating GLMMs into their analyses, researchers can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioural aspects involved.
+
+GLMMs are gaining increasing popularity in the field of human-AI interaction research.
+As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
+This versatility makes GLMMs particularly suitable for investigating various aspects of human-AI interaction, such as user preferences, trust, engagement, and performance.
+By incorporating GLMMs into their analyses, researchers in the field of human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
 
@@
 
+(there are some duplicate sentences here after merging!)
+
+2.
+Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
+This allows human-AI interaction researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
+By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
+
+3.
+Teaching power simulation skills to human-AI interaction researchers: The ability to conduct  power simulations is a valuable skill that should be taught to human-AI interaction researchers.
+By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
+Equipping human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
+The need to reason about how to simulate plausible data which is in line with the research hypothesis while not violating domain expertise on how plausible data should look like might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
+
+4.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in human-AI interaction research.
+Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
+It is crucial to raise awareness about the effort involved in power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
+
+5.
+Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in human-AI interaction research.
+They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
+The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports.
+
+6.
+Integration with Open Science and preregistration practices: Power simulations align well with the principles of open science and preregistration in human-AI interaction research.
+When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
+By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of Human-AI interaction research.
+
 # Conclusion

### 2023-09-05 docs/manuscript.Rmd
re-wrote manuscript based on feedback and points discussed in zoom call; this still needs another round of work from Timo before the next feedback round; added more lit

@@
 ---
-title             : "A Tutorial for Customized Power Simulations and Data Analyses for Human-AI Interaction Experiments using Generalized Linear Mixed Models"
+title             : "A Tutorial for Customized Power Simulations for Complex Experimental Study Designs using Generalized Linear Mixed Models"
 shorttitle        : "GLMM Power Simulation Tutorial"
@@
   Understanding how humans interact with artificial intelligence (AI) through experiments becomes increasingly relevant.
-However, human-AI interaction researchers often lack the appropriate tools to conduct power and data analyses for the complex study designs required.
+However, experimental interaction researchers often lack the appropriate tools to conduct power and data analyses for the complex study designs required.
 In this work, we provide a comprehensive guide to performing customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
-By providing code in a specific case study, we equip human-AI interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
-We discuss implications for human-AI interaction research and provide an outlook on the relevance of power simulations in the field.
+By providing code in a specific case study, we equip experimental interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
+We discuss implications for experimental interaction research and provide an outlook on the relevance of power simulations in the field.
 
-keywords          : "generalized linear mixed model, data simulation, sample size, power analysis, human-AI interaction"
+keywords          : "power analysis, data simulation, sample size, generalized linear mixed model"
 
@@
 
-<!-- General intro, relevance of Human AI Interaction experiments, need for well-powered and analyzed experimental research in this field -->
+<!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
+When planning experimental research it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data.
+However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), has illustrated many challenges surrounding the reproducibility and reliability of findings.
+As a results, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
+While accessible software solutions exists for simple statistical models and experimental designs, many experimental researcher lack the skills and tools to conduct power analyses for more complex research endeavors, such as powerful generalized linear mixed models (GLMM).
+In the present case study, we provide a tutorial comprised of a concrete example for custom power simulations using GLMMs.
 
-As artificial intelligence (AI) becomes more and more ubiquitous in our everyday lives and routines, it has become increasingly important to study and understand how humans interact with these intelligent systems.
-This is particularly relevant as AI models are getting more sophisticated and are deployed to support decision-making processes in areas such as law and medicine that have significant implications for individuals and society at large.
-Therefore, thorough research is needed to study the interaction between humans and AI technology to understand how people perceive, respond to and are influenced by these novel technologies. 
-<!-- SG: Defining a problem is missing here.
-It could be that most computer scientists developing the models are not well-trained in experimental design, which results in many papers on human-AI interaction being poorly designed and executed.
-This is problematic as it waists resources and might even be dangerous when readers rely on results from "bad" research.
--->
-When planning experimental research in this emerging field, it is essential to determine an adequate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative.
-Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for answering their research questions and analysing the data.
-<!-- SG: I think the second point, "and use appropriate statistical models to analyze the data." needs further explanation.
-I tried to change it, but I am not sure if this is what you actually wanted to say-->
-<!-- SG: Overall, I am not sure if the last two sentences are super well placed here since the benefits of performing a power analysis will be repeated below.
-Please ignore this comment when you think otherwise -->
+## Statistical power
 
-<!-- Some basics on statistical power, power analysis and why it is important -->
+<!-- Some basics on statistical power -->
 In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power.
-Statistical power is defined as the probability that a hypothesis test of interest <!-- is the "of interest" needed?
---> turns out significant when analyzing repeated samples from a population with a true effect of some pre-specified size.
+Statistical power is defined as the probability that a hypothesis test turns out significant when analyzing repeated samples from a population with a true effect of some pre-specified size.
+Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
+If the sample size (i.e., the number of participants or observations) used for data collection is insufficient to reliably detect the effects or relationships being investigated., this particular study would be considered "under-powered".
 
-As the integration of artificial intelligence (AI) into our daily lives continues to grow, it has become increasingly important to study how humans interact with these intelligent systems. 
-This is particularly important as AI becomes more sophisticated and is used in decision-making processes, for instance in medicine, that have significant impacts on individuals and society as a whole.
+<!-- what are the resulting issues with under-powered studies? -->
+Conducting under-powered research has many negative consequences.
+First, relying on under-powered experiments may yield inconclusive or misleading results hindering the accumulation of knowledge.
+Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results, thereby missing valuable discoveries and opportunities for scientific advancement.
 
-Therefore, thorough experiments are needed to study human-AI interaction to understand how people perceive and respond to these novel technologies.
-<!-- AK: A bit out of context - not clear to me why these experiments are needed.
-Maybe you could provide some examples what can be achieved with experiments in that context.
--->
-When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.
-<!-- AK: refs are generally missing - will they be added later?
--->
+## Power analysis
 
-<!-- Some basics on statistical power, power analysis and why it is important -->
-In empirical research relying on hypothesis testing, the most common strategy to determine appropriate sample size is based on statistical power. 
-Statistical power is defined as the probability that a hypothesis test of interest turns out significant, when analyzing repeated samples from a population with a true effect of some pre-specified size.
-
-Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
+<!-- define power analysis -->
 A power analysis represents the act of calculating the statistical power for a given true effect and sample size.
 By running the calculation for a range of sample sizes, the required sample size can be determined so that researchers to find an assumed true effect with the desired statistical power.
+Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value. 
+Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for answering their research questions and analysing the data.
 
-<!-- Power simulations are hard! (especially for GLMMs) -->
+<!-- advantages of an a priori power analysis -->
+For a desired statistical power and an assumed true effect, the required sample size to find such an effect can be calculated a priori.
+Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
+Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design.
+This really makes one think through everything carefully and might highlight statistical or design challenges before the start of an experiment.
+This gives the researcher a great opportunity to improve the design.
+Especially in fast-moving research fields, where speed is often regarded as more important than rigor, adding a proper sample size calculation into the process can act as a safeguard for ensuring high-quality research.
 
-Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
-<!-- SG: I am not 100% sure what the added benefit of this sentence is.
-It is pretty much what has been said in the previous sentence without the results with high evidential value".
-I would just add the last part to the previous sentence and then delete this one.
--->
-While user-friendly power analysis software is available for simple statistical models like t-tests, ANOVA, and linear regression, there are few tools are available for more sophisticated statistical methods.
-However, researchers studying human-AI interaction are increasingly using (generalized) linear mixed models (GLMMs; also called multilevel models), which add complexity to power analysis.
-For GLMMs, analytic options to derive sample sizes are not available for general research designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
-<!-- SG: I think I do not fully understand this sentence; do you mean interface-based software, or are you trying to focuse more on the flexibility of the research design/model specification?
--->
-A number of tutorials have been published describing how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs.
-This narrow focus provides little guidance for researchers faced with more complex study designs, which are very common [@brownIntroductionLinearMixedEffects2021].
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in the novel field of human-AI interaction, limited prior studies offer little guidance, rendering estimation a practical challenge.
-<!-- SG: I guess the content of the sentence I added was implied, but I think it would be useful to add, to highlight the challenges.
--->
-In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
-<!-- SG: I don't totally get the "in contrast".
-Unfortunately, I don't know if I get the entire argument.
-Before, you said that other tutorials expect you to have a good assumption for all model parameters, and now you say, in contrast, simulations are needed that require assumptions for all model parameters.
-What is the difference between the two statements, and why are they contrasting?
--->
-Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
-<!-- SG: Unfortunately, again, I do not follow the argument structure.
-How do we get from it being hard to estimate parameters to why we need sample size calculations for the experimental design?
-As far as I understand, we also needed parameter estimates to run this sample size calculation.
-Also, I don't think this sentence adds any benefit since it has already been said twice before.
--->
+<!-- external incentives to do a power analysis -->
+Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
+By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through the gained statistically-sound insights.
 
-For a desired statistical power and an assumed true effect, the required sample size to find such an effect can be calculated.
-Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
-While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
-For GLMMs, analytic options to derive sample sizes are not available for general research designs.
-An increasing number of tutorials describe how to perform power simulation for multilevel models.
-However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
-<!-- AK: Here I would provide examples for complex study designs and what can be tested with them -->
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies is available to determine plausible values for all model parameters.
-<!-- AK: Unclear to me what "data from pilot studies is available to determine plausible values for all model parameters" means -->
-In contrast<!-- AK: UWhy "in contrast". Does not fit well here -->, simulations are often required, necessitating assumptions for all parameters in the model.
+<!-- standard software is available for "simple" models and study designs -->
+For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups) user-friendly power analysis software is readily available [@faulStatisticalPowerAnalyses2009].
 
-Therefore, running sample size estimations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
-<!-- AK: Rewrite.
-You wrote before about the tutorials.
-"Therefore" doe snot fit here.
-Moreover, the link to human-AI interaction is not clear to me.
-Up to this point it seems a bit arbitrary.
--->
-Many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
-By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science.
+## Generalized linear mixed models (GLMM)
 
-<!-- AK: This is a little too general. This last paragraph addresses the reasons for conducting power analysis. "Advancement of science" is a bit too vague. -->
-Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for results with high evidencial value.
-While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
-For GLMMs, analytic options to derive sample sizes are not available for general research designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
-An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies are available to determine plausible values for all model parameters.
-In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
-Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
+<!-- Advantages of GLMMs over "simpler" models-->
+However, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+Therefore, GLMMs (generalized) linear mixed models (GLMMs; also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI.
+GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
 
-Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
-By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science [@yarkoniGeneralizabilityCrisis2022].
-<!-- SG: Unclear to the reader why incorporating GLMMS is enhancing replicability and credibility.
-Either you need a section above that describes the benefits of GLMMs, or it should be removed-->
-<!-- SG: What would really help the argument is that while conducting a proper power analysis, you have to consider every aspect of your experimental design.
-This really makes you think through everything carefully and might highlight statistical or design challenges before the start of an experiment.
-This gives the researcher a great opportunity to improve the design.
-Especially in fast-moving research fields, where speed is often regarded as more important than rigour, adding a proper sample size calculation into the process can act as a safeguard for ensuring high-quality research.
-A lot of this is implied, but it doesn't come through clearly.-->
-
-<!-- Research Gap & Summary -->
-
-When planning a human-AI interaction experiment with complexities such as the use of GLMMs, non-standard designs, and "missing by design" conditions, the process of power simulations becomes challenging.
-Available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
-<!-- SG: Hmm, both sentences above are basically a repetition from the previous sections with a bit more details.
-I wonder if we could restructure the introduction to avoid repetitions.
-My suggestion would be:
-1. AI technology is becoming ubiquitous, and we need to understand how humans interact with it. 
-2. Lots of people conduct research in this field; however, since many are not properly trained in experimental design, there is a lot of low-quality research out there, which might pose a risk. 
-3. One of the problems is that a lot of these studies are underpowered with way too many conditions and way too few participants
-4. Background information on why a properly powered study is important, including definitions of statistical power 
-5.
-Introducing the concept of power analysis and sample size calculation as a solution for the issue of underpowered low-quality research and highlighting the additional benefits of doing a power analysis
-6. Describing the challenges of performing a power analysis for complex research designs 
-7. Offering this tutorial as a way to overcome the above-described challenges. 
-I think ´that strictly sticking to a storyline like this would help to avoid repetition.-->
-
-When planning a human-AI interaction experiment <!-- AK: Is it really needed to have the human-AI interaction focus?
--->with complexities such as the use of GLMMs, non-standard designs, and missing by design conditions, the process of power simulations becomes challenging.<!-- AK: Readers might not know what non-standard designs are.
-Either work more with examples or elaborate more on these designs and how your paper addresses these difficulties.
--->
-Currently available software packages for power simulations do not include common designs in human-AI interaction research, making it necessary to build data simulations tailored specifically to the study design.
-
-Unfortunately, there are no general recommendations that can be applied universally in such cases.
-Instead, researchers must acquire the necessary skills based on the specific application at hand. 
-In this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study for which no meta-analytic results or conclusive data from pilot studies are available.
-<!-- SG: I am curious about the latter considering that we took both of our prior studies apart to get parameter estimates.
--->
-Through this case study, we aim to provide guidance and insights into the complexities involved in addressing sophisticated design considerations and obtaining accurate power estimates.
-<!-- SG: sample size estimates?
-This was the main goal, right?
--->
-
-# Methods
-
-<!-- overview of this section -->
-In this section, we outline the steps for performing data simulation and power analysis for GLMMs, using a case study based on a specific experimental design from human-AI interaction research.
-
-## The present case study
-<!-- SG: I am wondering if this is the perfect place for this section.
-By the end of the end of "basics of power simulation in GLMMs", readers have probably already forgotten the experimental design of the case study-->
-<!-- The present example, describe the study idea and design -->
-
-In the case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
-The goal is to understand how AI advice influences medical decision-making. 
-Participants, radiologists (task experts) and students/interns (non-task experts), review 50 head CT (computer tomography) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
-To support their decision-making, in two experimental conditions, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-This AI advice can be either either correct (80% of cases) or incorrect (20%).
-In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
-After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
-Each participant undergoes all experimental conditions, separated by a two-week washout period to prevent scan recall.
-
-This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans, (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice, and (c) different presentations of AI advice influence performance.
-In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible.
-The aim of the power simulation is to ascertain the required number of task experts and non-experts for achieving adequate statistical power in our experiment.
-
-In the case study, we simulate the data for a human-AI Interaction experiment where the diagnostic performance of users of a health care AI is to be evaluated.
-The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and emergency doctors (non task experts), view head CT (computertomography) scans and evaluate if an intracranial bleeding is present.
-To support their decision making, an AI provides a suggestion for a medical diagnosis. This AI advice can be either either correct or incorrect. 
-Participants then provide a final medical diagnosis (i.e., bleeding or no bleeding) that can be correct or incorrect. 
-Also, we have control trials, in which no AI advice is present. 
-This introduces some missing values by design (advice is neither correct nor incorrect, when no advice is present) which must be taken into account when simulating and analyzing the data. 
-Through this experiment, we want to determine if experts are better than non-experts in reading head CT scans.
-Further, we want to research if correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
-
-In the present example, it would be more challenging to recruit task experts (i.e., radiologists) as there is only a limited amount of such people.
-Non-experts would be easier to recruit<!-- AK: Why is that info relevant?
-I would remove this -->.
-The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
-The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
-
-## The lme4 package in R
-<!-- SG: I was wondering why you introduced the tool before the statistical model? -->
-
-<!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015]. The lme4 package is a state-of-the-art tool for fitting frequentist GLMMs.[^brms]
-
-[^brms]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
-
-The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
-However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
-<!-- SG: Do you think it is clear to readers what a population model is?
--->
-Therefore in this tutorial, we simulate data for GLMMs from first principles <!-- SG: I do not know what "first principles" means in this context.
-This might be true for some readers as well--> to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.[^simr]
-
-The lme4 package also includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
-However, the model parameterization used by the lme4 package is quite technical, making it more difficult for applied researchers to determine whether their specified population model implies plausible associations in their simulated data.
-
-Therefore in this tutorial, we simulate data for GLMMs from first principles to assist applied researchers in better understanding all model assumptions, and then use lme4 to analyze the simulated data sets.<!-- AK: I think this last 2 sentences are better for the introduction (i.e.
-what and why?) - maybe here stick to the description of the lme4 package.
-I am also not sure if you need to describe the lme4 package in detail.
-I think for the tutorial, understanding the underlying technique (GLMM) is more relevant.
--->
-
-[^simr]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
-
-## Generalized linear mixed models
-<!-- GLMM intro -->
-
-LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects.
+<!-- intro into GLMMs -->
+LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
 LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
-<!-- AK: This is a bit confusing.
-I would restructure this part.
-You do not explain what correlated and hierarchical mean in this context and it is difficult to follow you here.
--->They are used when the outcome variable is continuous and follow a normal distribution.
-LMMs allow for the modeling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
+In this context, correlated data means that the observations within a given dataset are not independent of each other.
+In other words, the values of the dependent variable for one observation may be related to the values of the dependent variable for other observations in a systematic way.
+This correlation can arise for various reasons, such as repeated measurements on the same subjects over time, observations clustered within certain groups or locations, or data collected at different levels of granularity.
+Hierarchical structures can be found in datasets that have a nested or hierarchical organization.
+For example, one might have data on students nested within schools, employees nested within departments, or patients nested within hospitals.
+In such cases, the observations within each higher-level grouping (e.g., school, department, hospital) tend to be more similar to each other than to observations in other groups.
+LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables).
+They allow for the modelling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
 Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
 GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family, such as binomial, Poisson, or gamma <!-- AK: I would briefly explain these distributions -->.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a way that is appropriate for the specific distribution family of the response variable.
-
-<!-- Relevance of GLMMs -->
-GLMMs are gaining increasing popularity in the field of Human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures within the data.
-GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of Human-AI interaction, such as user preferences, trust, engagement, and performance <!-- AK: Still not sure why this is particularly relevant in this context.
-These distributions also occur in other research contexts -->.
-By incorporating GLMMs into their analyses, researchers in the field of Human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
-
-LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
-LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
-They are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables).
-LMMs allow for the modelling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
-GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family, such as binomial, Poisson, or gamma.
+They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family (e.g., binomial).
 GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
@@
 
-<!-- Relevance of GLMMs -->
+## Power simulations for GLMMs
 
-<!-- SG: Auch hier hat sich wieder viel wiederholt. Ich habe mal versucht zu kürzen -->
-As the complexity of studying human interactions with AI systems grows, researchers require advanced statistical models to effectively capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-The above-described flexibility of GLMMs is particularly suitable for investigating various aspects of human-AI interaction, such as user preferences, trust, engagement, and performance.
-By incorporating GLMMs into their analyses, researchers can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioural aspects involved.
+<!-- what are the existing option for a priori power analysis for GLMMs / where are existing tutorials limited -->
+Due to the previously mentioned advantages of GLMMs, researchers are increasingly using those for their statistical analyses. 
+However, available software packages for power simulations do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
+As a consequence, custom power simulations are needed.
 
-GLMMs are gaining increasing popularity in the field of human-AI interaction research.
-As the complexity of studying human interactions with artificial intelligence systems grows, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
-This versatility makes GLMMs particularly suitable for investigating various aspects of human-AI interaction, such as user preferences, trust, engagement, and performance.
-By incorporating GLMMs into their analyses, researchers in the field of human-AI interaction can obtain more robust and comprehensive insights into the intricate dynamics between humans and AI systems, leading to a deeper understanding of the psychological and behavioral aspects involved.
+A number of tutorials have been published describing how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE).
+This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common [@brownIntroductionLinearMixedEffects2021].
+Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
+However, in practice, knowledge about those parameters from prior studies is often limited, rendering estimation a practical challenge [@maxwellSampleSizePlanning2008].
 
-<!-- basics of power simulation in GLMMs (with our specific example in mind) -->
+<!-- basics of power simulation for GLMMs  -->
 Power simulations in GLMMs are essential for estimating the statistical power of complex experimental study designs.
@@
 Interpreting these assumptions entails understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
+When planning an experiment with complexities such as the use of GLMMs, non-standard designs (e.g., hierarchical structure), and "missing by design" conditions, the process of power simulations becomes challenging.
+Unfortunately, there are no general recommendations that can be applied universally in such cases. Instead, researchers must acquire the necessary skills based on the specific application at hand.
+By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work, [@yarkoniGeneralizabilityCrisis2022].
+
+<!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short -->
+Based on need for sufficiently powered experimental research and the lack of tools to run custom power analyses for complex study designs using GLMMs, in this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study.
+Through this case study, we aim to provide guidance and insights into the complexities involved in addressing sophisticated design considerations and obtaining accurate required sample size estimates.
+
+# Method
+
+<!-- overview of this section -->
+
+In this section, we outline the steps for performing data simulation and power analysis for GLMMs, using a case study based on a specific experimental study design from the area of experimental interaction research.
+
+## Experimental study design
+
+<!-- describe the study idea and design -->
+In the case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
+The goal is to understand how AI advice influences medical decision-making.
+Participants, radiologists (task experts) and students/interns (non-task experts), review head CT (computer tomography) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
+To support their decision-making, in two experimental conditions, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
+This AI advice can be either either correct (80% of cases) or incorrect (20%).
+In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
+After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
+Each participant undergoes all experimental conditions, separated by a two-week washout period to prevent scan recall.
+This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
+With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans, (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice, and (c) different presentations of AI advice influence performance.
+In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
+The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
+
+## The lme4 package in R
+
+<!-- describe the lme4 package and how it works-->
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015]. The lme4 package is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
+
+[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
+
+The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+
+However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
+Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., starting from the most basic and fundamental concepts and building up from there to create synthetic data) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.[^2]
+
+[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
+
+## Our specific GLMM
 
@@
 g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$
-Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$.
-$$
+$$ Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-$$
-If the dependent variable (i.e., performance) $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability:
-$$
+$$ If the dependent variable (i.e., performance) $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability: $$
 P_{si} = P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
-$$
-In our case study, $P_{si}$ is the conditional probability that subject (i.e., participant) $s$ gives the correct response to item (i.e., CT scan) $i$.
+$$ In our case study, $P_{si}$ is the conditional probability that subject (i.e., participant) $s$ gives the correct response to item (i.e., CT scan) $i$.
 
-In such a setting, we model the probability as
-$$
+In such a setting, we model the probability as $$
 P_{si} = inverse\_logit(\eta_{si})
-$$
-with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
-$$
+$$ with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently $$
 logit(P_{si}) = \eta_{si}
-$$
-with the logit link $g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})$.
+$$ with the logit link $g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})$.
 
-In our case study, the probability of giving a correct response is assumed to depend on the predictors:
-<!-- SG: What happened to the different experimental conditions? AI advice had three levels: Advice with annotation, without annotation, and no advice-->
-- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to asses item $i$
-- $advice\_correct_{si}$: whether this advice was correct (1) or not (0)
-- $expert_s$: whether subject $s$ was a task expert (1) or not (0)
+In our case study, the probability of giving a correct response is assumed to depend on the predictors: <!-- SG: What happened to the different experimental conditions?
+AI advice had three levels: Advice with annotation, without annotation, and no advice--> - $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to asses item $i$ - $advice\_correct_{si}$: whether this advice was correct (1) or not (0) - $expert_s$: whether subject $s$ was a task expert (1) or not (0)
 
@@
 
-In total, we assume the model
-$$
+In total, we assume the model $$
 \begin{aligned}
@@
 \end{aligned}
-$$
-or equivalently
-$$
+$$ or equivalently $$
 \begin{aligned}
@@
 \end{aligned}
-$$
-with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
+$$ with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
 
@@
 ```
+
 <!-- SG: You basically start the explanation at line 257; I think it would be better to actually start at the beginning of the function line 251, otherwise, this might be very confusing to people.
 If you do not want to change the structure, then I would at least mention that you discuss the first section of the function below.-->
+
 In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan) (`n_items` in total).
@@
 
-The lme4 package uses a special syntax for model specification.
-Our assumed GLMM is represented by the formula:
+The lme4 package uses a special syntax for model specification. Our assumed GLMM is represented by the formula:
 
@@
 
-The output shows the estimates for all model parameters:
-the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
+The output shows the estimates for all model parameters: the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
 
@@
 The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
-Then the model implied probability that such a person solves such an item accurately is given by:
-$$
+Then the model implied probability that such a person solves such an item accurately is given by: $$
 \begin{aligned}
@@
 In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
-$$
+The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented: $$
 \begin{aligned}
@@
 
-For our case study, imagine the following combined hypothesis:
-*We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
+For our case study, imagine the following combined hypothesis: *We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
 
-This combined hypothesis leads to the following four separate null hypotheses to be tested:
-$$
+This combined hypothesis leads to the following four separate null hypotheses to be tested: $$
 \begin{aligned}
@@
 
-We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$:
-The first null hypothesis states that *an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.*
-This implies the following inequality for each subject $s$ and item $i$
-$$
+We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$: The first null hypothesis states that *an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.* This implies the following inequality for each subject $s$ and item $i$ $$
 inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}] \leq inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e]
-$$
-which simplifies to $\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0$.
+$$ which simplifies to $\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0$.
 
@@
 However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected [i.e.
-intersection-union setting\; @dmitrienkoTraditionalMultiplicityAdjustment2013].
+intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
 In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
@@
 
+Often no meta-analytic results or conclusive data from pilot studies are available. Then domain knowledge is essential.
+
 Table X shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model:
 
-| _Experimental condition_ | _$P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$_ | _Implied equation_ |
-|---|:---:|---|
+| *Experimental condition*  | $P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$ | *Implied equation*                                                              |
+|--------------------|:------------------------------:|--------------------|
 | no advice, no expert      |                      0.70                      | $logit(0.70) = \beta_0$                                                         |
@@
 
-It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g.
-$$
+It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g. $$
 \begin{aligned}
@@
 
-Although the derivations above are straightforward, it is important not to misinterpret their implications:
-In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
+Although the derivations above are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
 For example, we determined the $\beta$ parameters in a way that corresponds to a desired probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
-c
-However, even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
-This implies that one must be careful when specifying parameter values based on previous studies or pilot data. <!-- SG: What does that mean for someone who wants to do this then?-->
+c However, even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
+This implies that one must be careful when specifying parameter values based on previous studies or pilot data.
+<!-- SG: What does that mean for someone who wants to do this then?-->
 
 The inequality <!-- SG: Can you explain what this means?--> of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
-However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values:
-When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
+However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
 Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
 
-Earlier, we have already encountered one way to visualize the implied marginal distributions:
-We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
+Earlier, we have already encountered one way to visualize the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
 However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory.
@@
 
-In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$).
-This gives us a better way to see the variability between persons.
+In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$). This gives us a better way to see the variability between persons.
 
@@
 
-As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts:
-With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
+As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
 In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
@@
 
-## Power simulation
 With all these considerations addressed, we are finally ready to perform a power analysis.
@@
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^faux]
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^3]
 
-[^faux]: This code is heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
+[^3]: This code is heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
 
@@
 <!-- How many participants/stimuli are needed for concrete example -->
+
 As should be the case, power estimates increase with both the number of subjects and the number of items.
@@
 The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
-<!-- Maybe give an example?-->
-When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
+<!-- Maybe give an example?--> When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
 For the sake of the example, let the targeted power be indeed about 0.8 to detect an effect of the expected size implied by our data simulation.
@@
 However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule.
-The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (> 250, to be determined by additional rounds of power analysis).
+The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
 Design parameters that also affect power and which could be investigated in the simulation to find a more optimal trade-off are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
@@
 <!-- short summary of key takeaways from this tutorial -->
-Human-AI interaction research requires careful planning and consideration of statistical power to ensure reliable and meaningful results.
+
+Experimental research requires careful planning and consideration of statistical power to ensure reliable and meaningful results.
 While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
-The present tutorial presents a specific example of how to run power simulations for human-AI interaction experiments, which can also applied to other fields of research.
+The present tutorial presents a specific example of how to run power simulations for experimental interaction experiments, which can also applied to other fields of research.
 
@@
 <!-- I would try to highlight this more clearly in the methods part.
--->
-Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
+--> Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
 When sample sizes are chosen based on the results of our power simulations, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.
@@
 <!-- Outlook -->
-Beyond the specifics of our concrete case study, we want to outline six avenues regarding the outlook on power simulations in human-AI interaction research:
+
+Beyond the specifics of our concrete case study, we want to outline six avenues regarding the outlook on power simulations in experimental research:
 
 1.
-The need for power simulations in human-AI interaction research: Human-AI interaction research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
-Power simulations offer a solution by providing human-AI interaction researchers with a tailored approach to estimating statistical power.
+The need for power simulations in experimental research: Experimental research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
+Power simulations offer a solution by providing experimental interaction researchers with a tailored approach to estimating statistical power.
 These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
@@
 3.
-Teaching power simulation skills: The ability to conduct power simulations is a valuable skill that should be taught to human-AI interaction researchers.
+Teaching power simulation skills: The ability to conduct power simulations is a valuable skill that should be taught to experimental interaction researchers.
 By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
@@
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in human-AI interaction research.
-Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-<!-- SG: Hahah Timo, do you want to tell me something =) --> It is crucial to raise awareness about the effort involved in power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
-
-5.
-Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in human-AI interaction research.
-They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
-The importance of power simulations can be reflected by allocating them to a separate publication or incorporating them as a significant component of stage 1 preregistered reports.
-
-6.
-Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in human-AI interaction research.
-When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of research.
-
-(there are some duplicate sentences here after merging!)
-
-2.
-Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
-This allows human-AI interaction researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
-
-3.
-Teaching power simulation skills to human-AI interaction researchers: The ability to conduct  power simulations is a valuable skill that should be taught to human-AI interaction researchers.
-By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping human-AI interaction researchers with the knowledge and tools to perform power simulations empowers them to make informed decisions and enhance the rigor of their studies.
-The need to reason about how to simulate plausible data which is in line with the research hypothesis while not violating domain expertise on how plausible data should look like might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
-
-4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and  the perceived effort estimated by researchers and collaborators in human-AI interaction research.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and the perceived effort estimated by researchers and collaborators in experimental interaction research.
 Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
@@
 5.
-Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in human-AI interaction research.
+Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in experimental interaction research.
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
@@
 6.
-Integration with Open Science and preregistration practices: Power simulations align well with the principles of open science and preregistration in human-AI interaction research.
+Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in experimental interaction research.
 When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of Human-AI interaction research.
+By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of research.
 
@@
 
-Power simulations play a critical role in human-AI interaction research, allowing investigators to tailor power estimation to the unique aspects of their experiments.
+Power simulations play a critical role in experimental research, allowing investigators to tailor power estimation to the unique aspects of their experiments.
 Through this tutorial, we aim to provide researchers with the necessary skills and tools to perform these simulations themselves.

### 2023-09-06 docs/manuscript.Rmd
further improved manuscript
went through manuscript again and made final edits before handing over to Flo

@@
 ---
-title             : "A Tutorial for Customized Power Simulations for Complex Experimental Study Designs using Generalized Linear Mixed Models"
+title             : "A Tutorial on Tailored Power Simulations for Experimental Study Designs with Generalized Linear Mixed Models"
 shorttitle        : "GLMM Power Simulation Tutorial"
@@
   - name          : "Timo K. Koch"
-    affiliation   : "1, 2, 3"
+    affiliation   : "1, 2"
     corresponding : yes    # Define only one corresponding author
@@
   - name          : "Florian Pargent"
-    affiliation   : "2"
+    affiliation   : "1"
     role:
@@
     role:
+      - "Validation"
       - "Writing - Review & Editing"
+  - name          : "Eva Lermer"
+    affiliation   : "1,3"
+    role:
+      - "Resources"
   - name          : "Susanne Gaube"
-    affiliation   : "1, 4"
+    affiliation   : "1,4"
     role:
@@
   - id            : "1"
-    institution   : "LMU Munich, Center for Leadership and People Management"
+    institution   : "LMU Munich, Department of Psychology"
   - id            : "2"
-    institution   : "LMU Munich, Department of Psychology"
+    institution   : "University of St. Gallen, Institute of Behavioral Science & Technology"
   - id            : "3"
-    institution   : "University of St. Gallen, Institute of Behavioral Science & Technology"
+    institution   : "Technical University of Applied Sciences Augsburg, Department of Business Psychology"
   - id            : "4"
@@
 
-  Understanding how humans interact with artificial intelligence (AI) through experiments becomes increasingly relevant.
+  [THIS WILL BE UPDATED BASED ON GOOGLE DOC] Understanding how humans interact with artificial intelligence (AI) through experiments becomes increasingly relevant.
 However, experimental interaction researchers often lack the appropriate tools to conduct power and data analyses for the complex study designs required.
@@
 
-wordcount         : "6551"
+wordcount         : "7059"
 
@@
 As a results, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
-While accessible software solutions exists for simple statistical models and experimental designs, many experimental researcher lack the skills and tools to conduct power analyses for more complex research endeavors, such as powerful generalized linear mixed models (GLMM).
-In the present case study, we provide a tutorial comprised of a concrete example for custom power simulations using GLMMs.
+While accessible software solutions exists for simple statistical models and experimental designs, many experimental researcher lack the skills and tools to conduct power analyses for more complex research designs using powerful generalized linear mixed models (GLMM).
+In the present case study, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
 
@@
 
-## Power analysis
+## A priori power analysis
 
@@
 A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
-By running the calculation for a range of sample sizes, the required sample size can be determined so that researchers to find an assumed true effect with the desired statistical power.
-Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value. 
-Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for answering their research questions and analysing the data.
+By running the calculation for a range of sample sizes "a priori" (i.e., beforehand), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
+Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
 
 <!-- advantages of an a priori power analysis -->
-For a desired statistical power and an assumed true effect, the required sample size to find such an effect can be calculated a priori.
-Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for reliable results.
+Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for answering their research questions and analyzing the data.
 Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design.
-This really makes one think through everything carefully and might highlight statistical or design challenges before the start of an experiment.
-This gives the researcher a great opportunity to improve the design.
-Especially in fast-moving research fields, where speed is often regarded as more important than rigor, adding a proper sample size calculation into the process can act as a safeguard for ensuring high-quality research.
+This makes one think through everything carefully and might highlight statistical or design challenges before the start of an experiment.
+This gives researchers a great opportunity to improve the design.
+Especially in fast-moving research fields, where speed is often regarded as more important than rigor, adding a solid sample size calculation into the process can act as a safeguard for ensuring high-quality research.
 
@@
 
-<!-- Advantages of GLMMs over "simpler" models-->
-However, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-Therefore, GLMMs (generalized) linear mixed models (GLMMs; also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI.
-GLMMs offer a flexible framework for analyzing data with non-normal and categorical outcomes, accounting for both fixed and random effects.
+<!-- Advantages of GLMMs over "simpler" models -->
+As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+As a result, GLMMs (generalized) linear mixed models (GLMMs; also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
 
 <!-- intro into GLMMs -->
-LMMs (Linear Mixed Models) and GLMMs (Generalized Linear Mixed Models) are powerful statistical frameworks that handle complex data structures by incorporating both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
-LMMs are extensions of linear regression models that account for correlated data and hierarchical structures.
-In this context, correlated data means that the observations within a given dataset are not independent of each other.
+GLMMs are an extension of LMMs (Linear Mixed Models) which are in turn extensions of classic linear regression models that account for correlated data and hierarchical structures [@fahrmeirRegressionModelsMethods2021].
+In this context, correlated data means that observations within a given dataset are not independent of each other.
 In other words, the values of the dependent variable for one observation may be related to the values of the dependent variable for other observations in a systematic way.
@@
 Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
-GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
+As mentioned, GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
 They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family (e.g., binomial).
@@
 <!-- what are the existing option for a priori power analysis for GLMMs / where are existing tutorials limited -->
-Due to the previously mentioned advantages of GLMMs, researchers are increasingly using those for their statistical analyses. 
-However, available software packages for power simulations do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
-As a consequence, custom power simulations are needed.
-
-A number of tutorials have been published describing how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+Currently available software packages for power simulations do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
+A number of tutorials have been published describing how to perform such power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
 However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE).
 This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common [@brownIntroductionLinearMixedEffects2021].
-Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, rendering estimation a practical challenge [@maxwellSampleSizePlanning2008].
-
-<!-- basics of power simulation for GLMMs  -->
-Power simulations in GLMMs are essential for estimating the statistical power of complex experimental study designs.
-The model equation for a GLMM combines fixed effects, random effects, and an appropriate link function to model the relationship between predictors and the outcome variable.
 The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
@@
 Interpreting these assumptions entails understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
-When planning an experiment with complexities such as the use of GLMMs, non-standard designs (e.g., hierarchical structure), and "missing by design" conditions, the process of power simulations becomes challenging.
-Unfortunately, there are no general recommendations that can be applied universally in such cases. Instead, researchers must acquire the necessary skills based on the specific application at hand.
-By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work, [@yarkoniGeneralizabilityCrisis2022].
+Here, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
+However, in practice, knowledge about those parameters from prior studies is often limited, rendering estimation a practical challenge [@maxwellSampleSizePlanning2008].
+<!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
+Based on need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a priori power simulations of such study designs.
+Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes with GLMMs.
 
-<!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short -->
-Based on need for sufficiently powered experimental research and the lack of tools to run custom power analyses for complex study designs using GLMMs, in this paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a concrete study.
-Through this case study, we aim to provide guidance and insights into the complexities involved in addressing sophisticated design considerations and obtaining accurate required sample size estimates.
-
-# Method
+# The present case study
 
@@
 
-In this section, we outline the steps for performing data simulation and power analysis for GLMMs, using a case study based on a specific experimental study design from the area of experimental interaction research.
+In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs, using a case study based on a specific experimental study design from the area of human-AI interaction research.
 
@@
 <!-- describe the study idea and design -->
-In the case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
+In the present case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
 The goal is to understand how AI advice influences medical decision-making.
@@
 After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
-Each participant undergoes all experimental conditions, separated by a two-week washout period to prevent scan recall.
+Each participant undergoes all experimental conditions.
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
@@
 
-In our case study, the probability of giving a correct response is assumed to depend on the predictors: <!-- SG: What happened to the different experimental conditions?
-AI advice had three levels: Advice with annotation, without annotation, and no advice--> - $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to asses item $i$ - $advice\_correct_{si}$: whether this advice was correct (1) or not (0) - $expert_s$: whether subject $s$ was a task expert (1) or not (0)
+In our case study, the probability of giving a correct response is assumed to depend on the predictors:
+
+- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to asses item $i$ 
+- $advice\_correct_{si}$: whether this advice was correct (1) or not (0) 
+- $expert_s$: whether subject $s$ was a task expert (1) or not (0)
 
@@
 ```
-
-<!-- SG: You basically start the explanation at line 257; I think it would be better to actually start at the beginning of the function line 251, otherwise, this might be very confusing to people.
-If you do not want to change the structure, then I would at least mention that you discuss the first section of the function below.-->
 
@@
 When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
-All parameter values have been determined in repeated exchanges with our affiliated domain experts, and we now outline a few strategies on how to determine plausible parameter values.
 
+Often no meta-analytic results or conclusive data from pilot studies are available.
+All parameter values in our present case study have been determined based on results from prior own work and from the literature.
+Additionally, we had repeated exchanges with our affiliated domain experts from radiology.
+
+We now outline a few strategies on how to determine plausible parameter values.
 We have already seen in our discussion of model interpretation how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
 We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (for an average subject and an average item).
-
-Often no meta-analytic results or conclusive data from pilot studies are available. Then domain knowledge is essential.
 
@@
 For example, we determined the $\beta$ parameters in a way that corresponds to a desired probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
-c However, even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
+However, even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
 This implies that one must be careful when specifying parameter values based on previous studies or pilot data.
@@
 
-The inequality <!-- SG: Can you explain what this means?--> of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
+The inequality of conditional and marginal effects in GLMMs (i.e., how predictor variables affect the response within groups versus their overall impact across all groups, which can differ due to random effects) [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
 However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
@@
 Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
-First, we enable parallelization and specify how many parallel cores of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
-<!-- SG: Are workers the cores you want to use?-->
+First, we enable parallelization and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
 
@@
 library(future)
-plan("multisession", workers = 6)
+plan("multisession", workers = 4)
 set.seed(2)
@@
 The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
-<!-- SG: Can you explain how you choose the repetitions?-->
+We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
 
@@
 While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
-The present tutorial presents a specific example of how to run power simulations for experimental interaction experiments, which can also applied to other fields of research.
+The present tutorial presents a specific example of how to run power simulations for experimental designs with GLMMs.
 
@@
 The need for power simulations in experimental research: Experimental research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
-Power simulations offer a solution by providing experimental interaction researchers with a tailored approach to estimating statistical power.
+Power simulations offer a solution by providing experimental researchers with a tailored approach to estimating statistical power.
 These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
@@
 3.
-Teaching power simulation skills: The ability to conduct power simulations is a valuable skill that should be taught to experimental interaction researchers.
+Teaching power simulation skills: The ability to conduct power simulations is a valuable skill that should be taught to experimental researchers.
 By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
@@
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and the perceived effort estimated by researchers and collaborators in experimental interaction research.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and the perceived effort estimated by researchers and collaborators in experimental research.
 Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
@@
 5.
-Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in experimental interaction research.
+Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
@@
 6.
-Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in experimental interaction research.
+Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in experimental research.
 When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
@@
 Power simulations play a critical role in experimental research, allowing investigators to tailor power estimation to the unique aspects of their experiments.
-Through this tutorial, we aim to provide researchers with the necessary skills and tools to perform these simulations themselves.
-By integrating power simulations with open science and preregistration practices, researchers can improve the robustness and transparency of their findings, advancing the field.
+Through this tutorial, we aim to provide researchers with the necessary skills and tools to perform these simulations for GLMMs themselves.
+By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work [@yarkoniGeneralizabilityCrisis2022].
 

### 2023-09-07 docs/manuscript.Rmd
improve the manuscript and add some comments

@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = TRUE)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = FALSE)
 ```
@@
 However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), has illustrated many challenges surrounding the reproducibility and reliability of findings.
-As a results, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
-While accessible software solutions exists for simple statistical models and experimental designs, many experimental researcher lack the skills and tools to conduct power analyses for more complex research designs using powerful generalized linear mixed models (GLMM).
+<!-- FP: do we need a reference here?
+--> As a results, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
+While accessible software solutions exists for simple statistical models and experimental designs, many experimental researchers lack the skills and tools to conduct power analyses for more complex research designs using powerful <!-- FP: that can be confused with statistical power.
+Alternative: ...
+using the flexible GLMM framework --> generalized linear mixed models (GLMM).
 In the present case study, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
@@
 <!-- Some basics on statistical power -->
-In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power.
-Statistical power is defined as the probability that a hypothesis test turns out significant when analyzing repeated samples from a population with a true effect of some pre-specified size.
+In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power <!-- FP:  cite lakens again here?
+-->.
+Statistical power is defined as the probability that a hypothesis test turns out significant <!-- FP: ...
+has a significant p-value?
+to make sure that we mean the statistical/technical meaning of significant  --> when analyzing repeated samples from a population with a true effect of some pre-specified size.
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-If the sample size (i.e., the number of participants or observations) used for data collection is insufficient to reliably detect the effects or relationships being investigated., this particular study would be considered "under-powered".
+If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to reliably detect <!-- FP: alternative to reliably detect: detect the effects with high probability --> the effects or relationships being investigated, this particular study would be considered "under-powered".
 
@@
 Conducting under-powered research has many negative consequences.
-First, relying on under-powered experiments may yield inconclusive or misleading results hindering the accumulation of knowledge.
-Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results, thereby missing valuable discoveries and opportunities for scientific advancement.
+First, relying on under-powered experiments may yield inconclusive (if low power is discussed by the researchers) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
+Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results, thereby missing valuable discoveries and opportunities for scientific advancement <!-- FP: I would cut everything after funding  -->.
 
@@
 A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
-By running the calculation for a range of sample sizes "a priori" (i.e., beforehand), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
-Thereby, power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
+By running the calculation for a range of sample sizes "a priori" (i.e., before data collection), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
+Thereby, power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
 
 <!-- advantages of an a priori power analysis -->
-Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for answering their research questions and analyzing the data.
-Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design.
-This makes one think through everything carefully and might highlight statistical or design challenges before the start of an experiment.
+Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
+Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design and will possibly notice statistical or design challenges before the start of an experiment.
 This gives researchers a great opportunity to improve the design.
-Especially in fast-moving research fields, where speed is often regarded as more important than rigor, adding a solid sample size calculation into the process can act as a safeguard for ensuring high-quality research.
+Especially in fast-moving research fields, adding a solid sample size calculation into the process can act as a safeguard for ensuring high-quality research.
 
 <!-- external incentives to do a power analysis -->
-Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
-By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through the gained statistically-sound insights.
+Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights.
+<!-- FP: this last sentence seems like a repetition  -->
 
 <!-- standard software is available for "simple" models and study designs -->
-For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups) user-friendly power analysis software is readily available [@faulStatisticalPowerAnalyses2009].
+For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups) user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009].
 
@@
 As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-As a result, GLMMs (generalized) linear mixed models (GLMMs; also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
+GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
 
@@
 <!-- what are the existing option for a priori power analysis for GLMMs / where are existing tutorials limited -->
-Currently available software packages for power simulations do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
+Currently available software packages for power simulations <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation  --> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
 A number of tutorials have been published describing how to perform such power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
@@
 Here, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, rendering estimation a practical challenge [@maxwellSampleSizePlanning2008].
+However, in practice, knowledge about those parameters from prior studies is often limited, rendering estimation <!-- FP: why estimation?
+more like specifying assumptions?
+--> a practical challenge [@maxwellSampleSizePlanning2008].
 <!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
-Based on need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation for a priori power simulations of such study designs.
-Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes with GLMMs.
+Based on need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design.
+Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.
 
@@
 
-In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs, using a case study based on a specific experimental study design from the area of human-AI interaction research.
+In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs, using a case study based on a specific experimental study design from the area of human-AI <!-- FP: this is the first time AI appears; probably have to define the AI abbreviation --> interaction research.
 
@@
 The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts), review head CT (computer tomography) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
+Participants, radiologists (task experts) and students/interns (non-task experts) <!-- FP: during our work on the paper, I have heard two different stories of who the non experts are; either students/interns or full doctors that where not radiologists.
+Although this is only a data simulation example, it is nonetheless important here because the assumed effect sizes will probably differ!
+So who did you imagine when choosing the effect sizes?
+-->, review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
 To support their decision-making, in two experimental conditions, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-This AI advice can be either either correct (80% of cases) or incorrect (20%).
+This AI advice can be either correct (80% of cases) or incorrect (20%).
 In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
@@
 Each participant undergoes all experimental conditions.
+<!-- FP: I would remove the phrase experimental condition from this section. I think this was newly introduced here and does not match well with the experiment and the model as described later  -->
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
@@
 In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
-The goal of the power simulation is to determine how many task experts and non-experts to recruit to achieve sufficient statistical power in our experiment.
+The goal of the power simulation is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
 
@@
 <!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015]. The lme4 package is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015] which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
 
@@
 However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., starting from the most basic and fundamental concepts and building up from there to create synthetic data) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.[^2]
+Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., starting from the most basic and fundamental concepts and building up from there to create synthetic data <!-- FP: ALTERNATIVE: creating synthetic data step by step instead of using black box functions -->) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.[^2]
 
@@
 <!-- deep dive into our specific example -->
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modelled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021].
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021].
 $$
 g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$ Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
+$$ Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-$$ If the dependent variable (i.e., performance) $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability: $$
-P_{si} = P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
-$$ In our case study, $P_{si}$ is the conditional probability that subject (i.e., participant) $s$ gives the correct response to item (i.e., CT scan) $i$.
+$$ If the dependent variable (i.e., performance <!-- FP: is performance really the best option to describe our DV?
+-->) $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability: $$
+P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
+$$ In our case study, $P_{si}$ is the conditional probability that subject (i.e., participant <!-- FP: can't we just use subject instead of participant everywhere in the manuscript, i.e.
+also in the intro and when describing the experiment?
+I thought we had already agreed on this when deciding on how to write the equations and I do not want to change all the s subscripts now  -->) $s$ gives the correct response to item (i.e., CT scan) $i$.
 
-In such a setting, we model the probability as $$
+In such a setting, we model this probability as $$
 P_{si} = inverse\_logit(\eta_{si})
@@
 
-In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan) (`n_items` in total).
+In the first six lines of the function definition, we set some default parameter values (which we will explain in a later section) and load the packages we use to manipulate and simulate data. 
+In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).
 Thus the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
@@
 We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
-In the final step, we simulate the binary dependent variable `y_bin` by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
+In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject chooses the correct diagnosis for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
 
@@
 
-The lme4 package uses a special syntax for model specification. Our assumed GLMM is represented by the formula:
+The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:
 
@@
 In lme4, a GLMM is fitted with the `glmer` function.
-By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate vs.
-inaccurate diagnosis.
+By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs.
+inaccurate (0) diagnosis.
 
@@
 
-We can inspect the estimated model parameters with the `summary` command:
+We can inspect the estimates for all model parameters with the `summary` command:
 
@@
 
-The output shows the estimates for all model parameters: the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
+In the output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
 
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modelled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
 The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
@@
 Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
+Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
 
@@
 The `Fixed effects` table in Figure X also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$.
-However, for many research questions of interest, we are not interested in these two-sided tests referring to only one parameter.
+However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.
 
@@
 
-It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g. $$
+It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g. 
+$$
 \begin{aligned}
@@
 Although the derivations above are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
-For example, we determined the $\beta$ parameters in a way that corresponds to a desired probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
-However, even if the model were true, we would not observe this probability if we estimated the probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
-This implies that one must be careful when specifying parameter values based on previous studies or pilot data.
-<!-- SG: What does that mean for someone who wants to do this then?-->
+The first perspective implies a so-called marginal interpretation while the second one implies a conditional interpretation.
+For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
+However, even if the model were true, we would not observe this probability value if we estimated the marginal probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
 
-The inequality of conditional and marginal effects in GLMMs (i.e., how predictor variables affect the response within groups versus their overall impact across all groups, which can differ due to random effects) [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
-However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
+The inequality of conditional and marginal effects in GLMMs (i.e., how predictor variables affect the response within groups versus their overall impact across all groups, which can differ due to random effects) <!-- FP: I have tried to introduce contidional and marginal in the above section, so I would suggest to remove the bracket here --> [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
+One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
+<!-- SG: What does that mean for someone who wants to do this then?
+FP: is it more clear now?
+--> However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
 Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
@@
 This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-<!-- SG: Maybe explain why this is not plausible.-->
+For example, we would expect experts with low ability to solve a difficult item with a probability substantially larger than zero, when presented with wrong advice.
+<!-- SG: Maybe explain why this is not plausible.
+FP: like this?-->
 
@@
 With all these considerations addressed, we are finally ready to perform a power analysis.
-Wrapping the `simulate` function already introduced earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
+Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
 
@@
 library(future)
-plan("multisession", workers = 4)
+plan("multisession", workers = 6)
 set.seed(2)
@@
 We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
+With the current settings, this simulation takes about X minutes on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
+If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.
 
@@
 The result of the computation is a data.frame that contains the p-values of all tested hypotheses for each simulated dataset.
+In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
+We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
 
@@
 
-[^3]: This code is heavily inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
+[^3]: This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
 
@@
                                 power, ci.lwr, ci.upr)), 
-    color = "white", size = 6) +
-  scale_fill_viridis_c(limits = c(0, 1))
+    color = "white", size = 4.5) +
+  scale_fill_viridis_c(limits = c(0, 1)) +
+  xlab("number of subjects") + ylab("number of items")
 ```
@@
 In practice, power simulations are often run multiple times with adjusted combinations of sample sizes.
-When running for the first time, it might be revealed that power is way too low (or higher than required) for some combinations of `n_subjects` and `n_items`.
-When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of power simulations will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
-This will assure high precision for the final power estimates, which are then used for sample size justification for the future study.
+When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of `n_subjects` and `n_items`.
+When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of power simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
+This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.
 
@@
 <!-- Maybe give an example?--> When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
-For the sake of the example, let the targeted power be indeed about 0.8 to detect an effect of the expected size implied by our data simulation.
+For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation.
 This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases which is correct in about 80% of cases with present advice).
@@
 The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
-Design parameters that also affect power and which could be investigated in the simulation to find a more optimal trade-off are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
+Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
 
@@
 
-Experimental research requires careful planning and consideration of statistical power to ensure reliable and meaningful results.
+Experimental research requires careful planning and consideration of statistical power to ensure robust and meaningful results.
 While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
@@
 
-Beyond the specifics of our concrete case study, we want to outline six avenues regarding the outlook on power simulations in experimental research:
+Beyond the specifics of our concrete case study, we want to outline six avenues regarding the outlook on power simulations <!-- FP: ALTERNATIVE: regarding the future of power simulations --> in experimental research:
 
@@
 This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
+By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy <!-- FP: is accuracy the right term here?
+compromising realistic assumptions about the data generating process implied by the study design -->.
 
@@
 By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping researchers with the knowledge and tools to perform power simulations enables them to make informed decisions and enhance the rigour of their studies. 
-The need to reason about how to simulate plausible data that is in line with the research hypothesis while not violating domain expertise on how plausible data should look might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
+Equipping researchers with the knowledge and tools to perform power simulations enables them to make informed decisions and enhance the rigor of their studies. 
+The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look like, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
 
@@
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
-The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports.
+The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports <!-- FP: cite the Chambers RR  paper on what a preregistered report is?
+-->.
 
@@
 Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in experimental research.
-When researchers have access to simulated data based on their prespecified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their power simulations, researchers enhance transparency and accountability in their experimental procedures, contributing to the credibility and reproducibility of research.
+When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
+By preregistering their power simulations, researchers enhance transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
 

### 2023-09-07 docs/manuscript.Rmd
add/work on more comments

@@
 Statistical power is defined as the probability that a hypothesis test turns out significant <!-- FP: ...
-has a significant p-value?
+probability that a hypothesis test has a significant p-value?
 to make sure that we mean the statistical/technical meaning of significant  --> when analyzing repeated samples from a population with a true effect of some pre-specified size.
@@
 Conducting under-powered research has many negative consequences.
-First, relying on under-powered experiments may yield inconclusive (if low power is discussed by the researchers) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
+First, relying on under-powered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
 Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results, thereby missing valuable discoveries and opportunities for scientific advancement <!-- FP: I would cut everything after funding  -->.
@@
 Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
-Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design and will possibly notice statistical or design challenges before the start of an experiment.
+Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design and will possibly notice statistical or design challenges before starting with data collection.
 This gives researchers a great opportunity to improve the design.
-Especially in fast-moving research fields, adding a solid sample size calculation into the process can act as a safeguard for ensuring high-quality research.
+Especially in fast-moving research fields, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.
 
 <!-- external incentives to do a power analysis -->
-Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+Moreover, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
 By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights.
-<!-- FP: this last sentence seems like a repetition  -->
+<!-- FP: this last sentence seems like a repetition.
+wouldn't some sentence better fit here stating that proficiency with power analysis improves your career chances in academia by increasing the probability to get papers published and grants funded (which is a new argument we haven't already made elsewhere)  -->
 
@@
 For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups) user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009].
+<!-- FP: we could reintroduce the old section her on: with more complex models and designs, power analysis is more complicated because no analytic solution is possible, power has to be estimated by simulation, and there is no user-friendly software that includes all relevant designs.
+-->
 
@@
 
-The inequality of conditional and marginal effects in GLMMs (i.e., how predictor variables affect the response within groups versus their overall impact across all groups, which can differ due to random effects) <!-- FP: I have tried to introduce contidional and marginal in the above section, so I would suggest to remove the bracket here --> [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
+The inequality of conditional and marginal effects in GLMMs (i.e., how predictor variables affect the response within groups versus their overall impact across all groups, which can differ due to random effects) <!-- FP: I have tried to introduce conditional and marginal in the above section, so I would suggest to remove the bracket here --> [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
 One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).

### 2023-09-07 docs/manuscript.Rmd
TK: made some minor changes and replied to some of FP's comments & implemented them

@@
 author: 
-  - name          : "Timo K. Koch"
+  - name          : "Timo K. Koch*"
     affiliation   : "1, 2"
@@
       - "Writing - Review & Editing"
-  - name          : "Florian Pargent"
+  - name          : "Florian Pargent*"
     affiliation   : "1"
@@
     role:
-      - "Validation"
       - "Writing - Review & Editing"
@@
 
-# authornote: |
-#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-# 
-#   Enter author note here.
+authornote: |
+  \* The authors have contributed equally and share first authorship of this work.
 
@@
 
-  [THIS WILL BE UPDATED BASED ON GOOGLE DOC] Understanding how humans interact with artificial intelligence (AI) through experiments becomes increasingly relevant.
-However, experimental interaction researchers often lack the appropriate tools to conduct power and data analyses for the complex study designs required.
-In this work, we provide a comprehensive guide to performing customized power analyses using data simulation based on Generalized Linear Mixed Models (GLMMs).
-By providing code in a specific case study, we equip experimental interaction researchers with the tools to simulate their own data and run analyses based on GLMMs.
-We discuss implications for experimental interaction research and provide an outlook on the relevance of power simulations in the field.
+  [THIS WILL BE UPDATED BASED ON GOOGLE DOC] Researchers often lack the tools to conduct a priori power analysis for complex experimental designs deployed in human-computer interaction (HCI) studies.
+Generalized Linear Mixed Models (GLMMs) are well-suited to analyze experiments in which subjects respond to a set of stimuli.
+However, existing software packages can only be used to simulate power for simple designs, which are often not useful for more sophisticated investigations.
+This tutorial addresses this gap by providing a comprehensive tutorial to perform tailored power analyses using data simulation.
+Our tutorial provides code for an HCI case study, empowering researchers to simulate their own data and ascertain appropriate sample sizes for GLMMs.
+This resource should help to enhance the precision of study designs and to mitigate the risk of underpowered studies.
+Moreover, we discuss further implications for experimental research and provide an outlook on the increasing relevance of power simulations in the field.
 
@@
 When planning experimental research it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data.
-However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), has illustrated many challenges surrounding the reproducibility and reliability of findings.
-<!-- FP: do we need a reference here?
---> As a results, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
-While accessible software solutions exists for simple statistical models and experimental designs, many experimental researchers lack the skills and tools to conduct power analyses for more complex research designs using powerful <!-- FP: that can be confused with statistical power.
-Alternative: ...
-using the flexible GLMM framework --> generalized linear mixed models (GLMM).
+However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), has illustrated many challenges surrounding the reproducibility and reliability of findings [@robertsonImprovingStatisticalPractice2016].
+As a results, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
+While accessible software solutions exists for simple statistical models and experimental designs, many experimental researchers lack the skills and tools to conduct power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework.
 In the present case study, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
@@
 <!-- Some basics on statistical power -->
-In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power <!-- FP:  cite lakens again here?
--->.
-Statistical power is defined as the probability that a hypothesis test turns out significant <!-- FP: ...
-probability that a hypothesis test has a significant p-value?
-to make sure that we mean the statistical/technical meaning of significant  --> when analyzing repeated samples from a population with a true effect of some pre-specified size.
+In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022].
+Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to reliably detect <!-- FP: alternative to reliably detect: detect the effects with high probability --> the effects or relationships being investigated, this particular study would be considered "under-powered".
+If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated, this particular study would be considered "under-powered".
 
@@
 First, relying on under-powered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
-Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results, thereby missing valuable discoveries and opportunities for scientific advancement <!-- FP: I would cut everything after funding  -->.
+Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
 
@@
 <!-- FP: this last sentence seems like a repetition.
-wouldn't some sentence better fit here stating that proficiency with power analysis improves your career chances in academia by increasing the probability to get papers published and grants funded (which is a new argument we haven't already made elsewhere)  -->
+wouldn't some sentence better fit here stating that proficiency with power analysis improves your career chances in academia by increasing the probability to get papers published and grants funded (which is a new argument we haven't already made elsewhere); TK: I like this argument.
+Might be a little idealistic though since bc flashy results get you ahead mostly :D   -->
 
@@
 For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups) user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009].
-<!-- FP: we could reintroduce the old section her on: with more complex models and designs, power analysis is more complicated because no analytic solution is possible, power has to be estimated by simulation, and there is no user-friendly software that includes all relevant designs.
+<!-- FP: we could reintroduce the old section her on: with more complex models and designs, power analysis is more complicated because no analytic solution is possible, power has to be estimated by simulation, and there is no user-friendly software that includes all relevant designs.; TK: yes let's do this.
+I am not entirely sure what you mean with "analytic" solution, i think we should define this.
 -->
@@
 <!-- what are the existing option for a priori power analysis for GLMMs / where are existing tutorials limited -->
-Currently available software packages for power simulations <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation  --> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
+Currently available software packages for power simulations <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.
+--> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
 A number of tutorials have been published describing how to perform such power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
@@
 Here, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, rendering estimation <!-- FP: why estimation?
-more like specifying assumptions?
---> a practical challenge [@maxwellSampleSizePlanning2008].
+However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions a practical challenge [@maxwellSampleSizePlanning2008].
+
 <!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
@@
 
-In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs, using a case study based on a specific experimental study design from the area of human-AI <!-- FP: this is the first time AI appears; probably have to define the AI abbreviation --> interaction research.
+In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs, using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
 
@@
 Although this is only a data simulation example, it is nonetheless important here because the assumed effect sizes will probably differ!
-So who did you imagine when choosing the effect sizes?
+So who did you imagine when choosing the effect sizes?; SG can you clarify here please?
 -->, review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
@@
 After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
-Each participant undergoes all experimental conditions.
-<!-- FP: I would remove the phrase experimental condition from this section. I think this was newly introduced here and does not match well with the experiment and the model as described later  -->
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
@@
 However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., starting from the most basic and fundamental concepts and building up from there to create synthetic data <!-- FP: ALTERNATIVE: creating synthetic data step by step instead of using black box functions -->) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.[^2]
+Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.[^2]
 
@@
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-$$ If the dependent variable (i.e., performance <!-- FP: is performance really the best option to describe our DV?
--->) $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability: $$
+$$ If the dependent variable (i.e., diagnostic accuracy) $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability: $$
 P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
-$$ In our case study, $P_{si}$ is the conditional probability that subject (i.e., participant <!-- FP: can't we just use subject instead of participant everywhere in the manuscript, i.e.
-also in the intro and when describing the experiment?
-I thought we had already agreed on this when deciding on how to write the equations and I do not want to change all the s subscripts now  -->) $s$ gives the correct response to item (i.e., CT scan) $i$.
+$$ In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.
 
@@
 
-The inequality of conditional and marginal effects in GLMMs (i.e., how predictor variables affect the response within groups versus their overall impact across all groups, which can differ due to random effects) <!-- FP: I have tried to introduce conditional and marginal in the above section, so I would suggest to remove the bracket here --> [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
+The inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
 One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
@@
 
-Beyond the specifics of our concrete case study, we want to outline six avenues regarding the outlook on power simulations <!-- FP: ALTERNATIVE: regarding the future of power simulations --> in experimental research:
+Beyond the specifics of our concrete case study, we want to outline six avenues regarding the future of power simulations in experimental research:
 

### 2023-09-07 docs/manuscript.Rmd
integrated shared first authorship info using papaja

@@
 author: 
-  - name          : "Timo K. Koch*"
+  - name          : "Timo K. Koch"
     affiliation   : "1, 2"
@@
     email         : "timo.koch@unisg.ch"
+    equal_contrib: yes
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
       - "Writing - Review & Editing"
-  - name          : "Florian Pargent*"
+  - name          : "Florian Pargent"
     affiliation   : "1"
+    equal_contrib: yes
     role:
@@
 
-authornote: |
-  \* The authors have contributed equally and share first authorship of this work.
+#authornote: |
+  # author note here
 

### 2023-09-08 docs/manuscript.Rmd
Update manuscript.Rmd

@@
 
+<!-- SG: In the intro power analysis, a priori power analysis and sample size calculation is now used interchangeably, should we use one term consistently?-->
+
 <!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
-When planning experimental research it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data.
-However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), has illustrated many challenges surrounding the reproducibility and reliability of findings [@robertsonImprovingStatisticalPractice2016].
-As a results, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
-While accessible software solutions exists for simple statistical models and experimental designs, many experimental researchers lack the skills and tools to conduct power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework.
-In the present case study, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
+When planning experimental research, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data.
+However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), <!-- SG: is the HCI needed here?
+Also, can we claim a general replication crisis in experimental research?
+For instance, have no idea if experimental physics also has a problem.
+Maybe "in several disciplines --> has illustrated many challenges surrounding the reproducibility and reliability of findings [@robertsonImprovingStatisticalPractice2016].
+As a result, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
+While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework.
+In the present work, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
 
@@
 Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated, this particular study would be considered "under-powered".
+If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated <!-- SG: I think there is something wrong with the logic of this sentence.
+Should it be like this?
+If the sample size used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered"-->, this particular study would be considered "underpowered".
 
-<!-- what are the resulting issues with under-powered studies? -->
-Conducting under-powered research has many negative consequences.
-First, relying on under-powered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
-Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
+<!-- what are the resulting issues with underpowered studies? -->
+Conducting underpowered research has many negative consequences.
+First, relying on underpowered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
+Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
 
@@
 <!-- standard software is available for "simple" models and study designs -->
-For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups) user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009].
+For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009].
 <!-- FP: we could reintroduce the old section her on: with more complex models and designs, power analysis is more complicated because no analytic solution is possible, power has to be estimated by simulation, and there is no user-friendly software that includes all relevant designs.; TK: yes let's do this.
 I am not entirely sure what you mean with "analytic" solution, i think we should define this.
--->
+SG: I am not sure what the plan is here.
+This sentence alone makes little sense.
+The argument that these tools cannot be used or adjusted for more complex models is missing.
+If you want to link it to the next section, it would argue that these simple statistical methods are often not sufficient for researchers to answer their research questions and that more complex designs are required.
+I would not go in detail here, that no tools for GLMM exist, since this argument is already mentioned below-->
 
@@
 <!-- intro into GLMMs -->
-GLMMs are an extension of LMMs (Linear Mixed Models) which are in turn extensions of classic linear regression models that account for correlated data and hierarchical structures [@fahrmeirRegressionModelsMethods2021].
+GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of classic linear regression models that account for correlated data and hierarchical structures [@fahrmeirRegressionModelsMethods2021].
 In this context, correlated data means that observations within a given dataset are not independent of each other.
@@
 As mentioned, GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution, but instead belongs to a different distribution family (e.g., binomial).
+They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family (e.g., binomial) <!-- SG: do you think examples for the different types of variables are needed-->.
 GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
@@
 
-<!-- what are the existing option for a priori power analysis for GLMMs / where are existing tutorials limited -->
+<!-- what are the existing options for a priori power analysis for GLMMs / where are existing tutorials limited -->
 Currently available software packages for power simulations <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.
@@
 <!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
-Based on need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design.
+Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design.
 Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.
@@
 
-In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs, using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
+In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
 
@@
 The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts) <!-- FP: during our work on the paper, I have heard two different stories of who the non experts are; either students/interns or full doctors that where not radiologists.
+Participants, radiologists (task experts) and students/interns (non-task experts)<!-- FP: during our work on the paper, I have heard two different stories of who the non experts are; either students/interns or full doctors that where not radiologists.
 Although this is only a data simulation example, it is nonetheless important here because the assumed effect sizes will probably differ!
 So who did you imagine when choosing the effect sizes?; SG can you clarify here please?
--->, review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
-To support their decision-making, in two experimental conditions, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
+SG: Sure, when planning the study, our non-expert group was going to be students and interns.
+In a previous study, the non-expert groups were physicians with little task expertise.
+In the end, we did base some assumptions on these previous studies since we did not have anything else to work with -->, review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
+To support their decision-making, in two experimental conditions <!--SG: I still think this will be confusing to readers.
+Why two experimental conditions?-->, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
 This AI advice can be either correct (80% of cases) or incorrect (20%).
@@
 With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans, (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice, and (c) different presentations of AI advice influence performance.
+<!--SG: c) has to be removed, if it is not in the analysis, right?
+-->
 In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
@@
 $$
-g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$ Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
+g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta <!-- SG: Sorry for maybe a stupid question, but what does E stand for? and why are model parameters β mentioned but not in the equation?-->
+$$ Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-$$ If the dependent variable (i.e., diagnostic accuracy) $Y$ is a binary variable with values $0$ or $1$, the conditional expected value is equivalent to the probability: $$
+$$ If the dependent variable (i.e., diagnostic decision)<!-- SG: not sure if diagnostic accuracy as a term is useful as a binary dependent variable since it is usually calculated as a percentage over many cases.)  --> $Y$ is a binary variable with values $0$ (i.e.
+inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
 P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
@@
 In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).
-Thus the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
+Thus, the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
 We add a between-subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
-The next two lines simulate that in about $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
+The next two lines simulate that in $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
 Next, we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
@@
 We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
-In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject chooses the correct diagnosis for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
+In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject chooses the accurate diagnosis for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
 
@@
 
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modelled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
 The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
@@
 However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.
+<!-- Figure is not included in the pdf -->
 
@@
 Table X shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model:
-
+<!-- SG Table number, title and note are missing. -->
 | *Experimental condition*  | $P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$ | *Implied equation*                                                              |
@@
 Although the derivations above are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
-The first perspective implies a so-called marginal interpretation while the second one implies a conditional interpretation.
+The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
 For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
@@
 FP: is it more clear now?
+SG: Definitely better; the question might still be what your solution to this issue is.
 --> However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
-Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
+Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
 
-Earlier, we have already encountered one way to visualize the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
+Earlier, we have already encountered one way to visualize <!-- SG: so far, no figures in the pdf --> the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
 However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory.
@@
 This difference increases with the variability of the random effects.
+<!-- SG: Figure is not included in the PDF-->
 
@@
 
-The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
+The final plot demonstrates <!-- SG: no figures in the pdf --> that these plots are also useful for spotting standard deviations that are specified too high.
 For this example, we have set $\sigma_S = 3$ and $\sigma_I = 3$.
 This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-For example, we would expect experts with low ability to solve a difficult item with a probability substantially larger than zero, when presented with wrong advice.
+For example, we would expect even experts with low ability to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
 <!-- SG: Maybe explain why this is not plausible.
-FP: like this?-->
+FP: like this?
+SG: Better, but maybe still include a reason for this (years of training to do exactly this).
+Do you think it should be mentioned that already when just guessing, their probability would be .5-->
 
@@
 In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
-We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
+<!-- SG: Maybe explain what this means--> We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
 
@@
 For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation.
-This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases which is correct in about 80% of cases with present advice).
+This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice).
 If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items.
@@
 This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy <!-- FP: is accuracy the right term here?
+By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy<!-- FP: is accuracy the right term here?
 compromising realistic assumptions about the data generating process implied by the study design -->.

### 2023-09-11 docs/manuscript.Rmd
anne's comments

@@
 For instance, have no idea if experimental physics also has a problem.
-Maybe "in several disciplines --> has illustrated many challenges surrounding the reproducibility and reliability of findings [@robertsonImprovingStatisticalPractice2016].
+Maybe "in several disciplines.
+AK: I agree with Susanne.
+I am not sure whether the HCI is needed.
+Seems a bit arbitrary.
+Why does this apply to HCI in particular?
+Either work with an example or keep more generel --> has illustrated many challenges surrounding the reproducibility and reliability of findings [@robertsonImprovingStatisticalPractice2016].
 As a result, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
@@
 In the present work, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
+<!-- AK: Great, I like the Intro.
+I take my previous comment back - I think this can stay as it is now, without examples.
+I feel that it is not tailored towards HCI, but I also don't think that's a big issue.-->
 
@@
 Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
-Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated <!-- SG: I think there is something wrong with the logic of this sentence.
+Less formally, power is described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.<!-- AK: Just a very minor thing: I learned that power is the probability that a statistical test correctly rejects the null hypothesis when the alternative hypothesis is true.
+That is basically the same as what you wrote, but maybe a bit clearer because you don't have H1 and H1 right after another.
+But up to you!
+--> If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated <!-- SG: I think there is something wrong with the logic of this sentence.
 Should it be like this?
-If the sample size used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered"-->, this particular study would be considered "underpowered".
+If the sample size used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered"-->, this particular study would be considered "underpowered".<!--AK: I agree with Susanne.
+This is not a sentence anymore.
+-->
 
@@
 Conducting underpowered research has many negative consequences.
-First, relying on underpowered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
+First, relying on underpowered experiments may yield inconclusive (if low power is taken into account<!--AK: Unclear what this means in this context-  how is it "taken into account"?
+-->) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
 Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
@@
 By running the calculation for a range of sample sizes "a priori" (i.e., before data collection), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
-Thereby, power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
+<!--AK: I am not sure, but isn't this the definition of power simulation more or less?
+Maybe you could introduce that here already.
+-->
+Thereby, power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.<!--AK: This is already an advantage.
+I would combine this with the part below.-->
 
@@
 Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
-Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design and will possibly notice statistical or design challenges before starting with data collection.
-This gives researchers a great opportunity to improve the design.
-Especially in fast-moving research fields, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.
+Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
+Especially in fast-moving research fields<!--AK: I don't understand how this applies to fast-moving research fields (btw, what is that??) in particular.
+I would delere this.-->, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.
 
@@
 wouldn't some sentence better fit here stating that proficiency with power analysis improves your career chances in academia by increasing the probability to get papers published and grants funded (which is a new argument we haven't already made elsewhere); TK: I like this argument.
-Might be a little idealistic though since bc flashy results get you ahead mostly :D   -->
+Might be a little idealistic though since bc flashy results get you ahead mostly :D  AK: I agree, the first part is repetitive.
+-->
 
@@
 If you want to link it to the next section, it would argue that these simple statistical methods are often not sufficient for researchers to answer their research questions and that more complex designs are required.
-I would not go in detail here, that no tools for GLMM exist, since this argument is already mentioned below-->
+I would not go in detail here, that no tools for GLMM exist, since this argument is already mentioned below; AK: I agree with Susanne.-->
 
@@
 <!-- Advantages of GLMMs over "simpler" models -->
-As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
+As study designs become more complex<!--AK: I am not even sure whether this is true.
+Do study designs become more complex or do we become more aware of the pitfalls of underpowered studies?
+I would say it's the latter.-->, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI<!--AK: This "such as in HCI" always sounds super weird.
+It becomes apparant that you need to mention this for the submission.
+I would eitehr remove this entirely or explain this more in context of HCI research (using an example from that field)-->, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
 
@@
 GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.
+The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.<!--AK: And here I would add one research example.
+Something like, "Imagine an experiment where experts are repeatedly asked for...
+In this case, the data is not independent because ...
+I would add an example that explains both aspects you talk about before: 1) The dependency, 2) the outcome distribution (e.g., binomial).
+I think this would make it much easier for the reader to follow you here.-->
 
@@
 However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE).
-This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common [@brownIntroductionLinearMixedEffects2021].
+This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common <!--AK: Are very common how?
+Unclear sentence-->[@brownIntroductionLinearMixedEffects2021].
 The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
 The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial.
-Assumptions about the random effects include the assumption of normality and the covariance structure among the random effects.
+Assumptions about the random effects include the assumption of normality<!--AK: Normality of what?--> and the covariance structure among the random effects<!--AK: Unclear what this is-->.
 Additionally, assumptions about the error structure, such as independence or correlation, must be specified.
-Interpreting these assumptions entails understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
-Here, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions a practical challenge [@maxwellSampleSizePlanning2008].
+Interpreting these assumptions entails understanding the underlying assumptions <!--AK: Repeated "assumptions".
+Find other word.-->of the model and ensuring they align with the characteristics of the data being analyzed.
+Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
+However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions <!--AK: Again "assumptions".-->a practical challenge [@maxwellSampleSizePlanning2008].
 
 <!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
-Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design.
+Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design<!--AK: I would just write GLMM.-->.
 Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.
@@
 <!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015] which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015] which is a state-of-the-art tool for fitting frequentist GLMMs<!--AK: I think you need a ref for this statement.-->.[^1]
 
@@
 ## Our specific GLMM
-
 <!-- deep dive into our specific example -->
@@
 $$
-g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta <!-- SG: Sorry for maybe a stupid question, but what does E stand for? and why are model parameters β mentioned but not in the equation?-->
+g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta <!-- SG: Sorry for maybe a stupid question, but what does E stand for?
+and why are model parameters beta mentioned but not in the equation?-->
 $$ Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
@@
     mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
-    mutate(advice_correct = if_else(advice_present == 1, 
-                                    rbinom(n(), 1, prob = 0.8), 0)) %>%
+    mutate(advice_correct = if_else(advice_present == 1L, 
+                                    rbinom(n(), 1L, prob = 0.8), 0L)) %>%
     # add random effects

### 2023-09-11 docs/manuscript.tex
anne's comments

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{power analysis, data simulation, sample size, generalized linear mixed model\newline\indent Word count: 7059}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={A Tutorial on Tailored Power Simulations for Experimental Study Designs with Generalized Linear Mixed Models},
  pdfauthor={Timo K. Koch,1, 2, Florian Pargent,1, Anne-Kathrin Kleine1, Eva Lermer1,3, \& Susanne Gaube1,4},
  pdflang={en-EN},
  pdfkeywords={power analysis, data simulation, sample size, generalized linear mixed model},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{A Tutorial on Tailored Power Simulations for Experimental Study Designs with Generalized Linear Mixed Models}
\author{Timo K. Koch\textsuperscript{$\dagger{}$,1, 2}, Florian Pargent\textsuperscript{$\dagger{}$,1}, Anne-Kathrin Kleine\textsuperscript{1}, Eva Lermer\textsuperscript{1,3}, \& Susanne Gaube\textsuperscript{1,4}}
\date{}

\shorttitle{GLMM Power Simulation Tutorial}

\authornote{

The authors made the following contributions. Timo K. Koch: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Florian Pargent: Formal Analysis, Methodology, Visualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Anne-Kathrin Kleine: Writing - Review \& Editing; Eva Lermer: Resources; Susanne Gaube: Writing - Review \& Editing, Supervision. \textsuperscript{$\dagger{}$} Timo K. Koch and Florian Pargent contributed equally to this work.

Correspondence concerning this article should be addressed to Timo K. Koch. E-mail: \href{mailto:timo.koch@unisg.ch}{\nolinkurl{timo.koch@unisg.ch}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} LMU Munich, Department of Psychology\\\textsuperscript{2} University of St.~Gallen, Institute of Behavioral Science \& Technology\\\textsuperscript{3} Technical University of Applied Sciences Augsburg, Department of Business Psychology\\\textsuperscript{4} University College London, Global Business School for Health}

\abstract{%
{[}THIS WILL BE UPDATED BASED ON GOOGLE DOC{]} Researchers often lack the tools to conduct a priori power analysis for complex experimental designs deployed in human-computer interaction (HCI) studies. Generalized Linear Mixed Models (GLMMs) are well-suited to analyze experiments in which subjects respond to a set of stimuli. However, existing software packages can only be used to simulate power for simple designs, which are often not useful for more sophisticated investigations. This tutorial addresses this gap by providing a comprehensive tutorial to perform tailored power analyses using data simulation. Our tutorial provides code for an HCI case study, empowering researchers to simulate their own data and ascertain appropriate sample sizes for GLMMs. This resource should help to enhance the precision of study designs and to mitigate the risk of underpowered studies. Moreover, we discuss further implications for experimental research and provide an outlook on the increasing relevance of power simulations in the field.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

When planning experimental research, it is essential to determine an appropriate sample size (Lakens, 2022b) to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data. However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), has illustrated many challenges surrounding the reproducibility and reliability of findings (Robertson \& Kaptein, 2016). As a result, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs. While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework. In the present work, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.

\hypertarget{statistical-power}{%
\subsection{Statistical power}\label{statistical-power}}

In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power (Lakens, 2022b). Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size. Less formally, power is described as the probability that a hypothesis test accepts the alternative hypothesis \(H_1\) if \(H_1\) is indeed true. If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated , this particular study would be considered ``underpowered''.

Conducting underpowered research has many negative consequences. First, relying on underpowered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge. Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.

\hypertarget{a-priori-power-analysis}{%
\subsection{A priori power analysis}\label{a-priori-power-analysis}}

A power analysis represents the act of calculating the statistical power for a given true effect and sample size.
By running the calculation for a range of sample sizes ``a priori'' (i.e., before data collection), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
Thereby, power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.

Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
Especially in fast-moving research fields, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.

Moreover, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings. By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights.

For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for power analysis is readily available (Faul, Erdfelder, Buchner, \& Lang, 2009).

\hypertarget{generalized-linear-mixed-models-glmm}{%
\subsection{Generalized linear mixed models (GLMM)}\label{generalized-linear-mixed-models-glmm}}

As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs (Yarkoni, 2022). GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects (Fahrmeir, Kneib, Lang, \& Marx, 2021; Kaptein, 2016).

GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of classic linear regression models that account for correlated data and hierarchical structures (Fahrmeir et al., 2021). In this context, correlated data means that observations within a given dataset are not independent of each other. In other words, the values of the dependent variable for one observation may be related to the values of the dependent variable for other observations in a systematic way. This correlation can arise for various reasons, such as repeated measurements on the same subjects over time, observations clustered within certain groups or locations, or data collected at different levels of granularity. Hierarchical structures can be found in datasets that have a nested or hierarchical organization. For example, one might have data on students nested within schools, employees nested within departments, or patients nested within hospitals. In such cases, the observations within each higher-level grouping (e.g., school, department, hospital) tend to be more similar to each other than to observations in other groups.
LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables). They allow for the modelling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects. Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
As mentioned, GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables. They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family (e.g., binomial) . GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable. The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.

\hypertarget{power-simulations-for-glmms}{%
\subsection{Power simulations for GLMMs}\label{power-simulations-for-glmms}}

Currently available software packages for power simulations do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design (Murayama, Usami, \& Sakaki, 2022; Westfall, Kenny, \& Judd, 2014).
A number of tutorials have been published describing how to perform such power simulation for multilevel models (Arend \& Schäfer, 2019; Brysbaert \& Stevens, 2018; DeBruine \& Barr, 2021; Kumle, Võ, \& Draschkow, 2021; Lafit et al., 2021; Zimmer, Henninger, \& Debelak, 2022).
However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE). This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common (Brown, 2021).
The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure. The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial. Assumptions about the random effects include the assumption of normality and the covariance structure among the random effects. Additionally, assumptions about the error structure, such as independence or correlation, must be specified. Interpreting these assumptions entails understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters. However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions a practical challenge (Maxwell, Kelley, \& Rausch, 2008).

Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design. Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.

\hypertarget{the-present-case-study}{%
\section{The present case study}\label{the-present-case-study}}

In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.

\hypertarget{experimental-study-design}{%
\subsection{Experimental study design}\label{experimental-study-design}}

In the present case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated.
The goal is to understand how AI advice influences medical decision-making. Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
To support their decision-making, in two experimental conditions , an AI model provides initial diagnostic advice, which can be used as guidance by the participants. This AI advice can be either correct (80\% of cases) or incorrect (20\%). In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support. After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data. With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans, (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice, and (c) different presentations of AI advice influence performance.
In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible.
The goal of the power simulation is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.

\hypertarget{the-lme4-package-in-r}{%
\subsection{The lme4 package in R}\label{the-lme4-package-in-r}}

In our case study, we use the lme4 R package (Bates, Mächler, Bolker, \& Walker, 2015) which is a state-of-the-art tool for fitting frequentist GLMMs.\footnote{For Bayesian GLMMs, the brms R package is currently the most prominent option (Bürkner, 2017).}

The lme4 package includes a function called \texttt{simulate} that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.

However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data. Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.\footnote{A less flexible alternative would be to use the simr package (Green \& MacLeod, 2016), which can be used to both simulate data and perform power analysis for models supported by the lme4 package.}

\hypertarget{our-specific-glmm}{%
\subsection{Our specific GLMM}\label{our-specific-glmm}}

In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables \(\mathbf{X}\) and random effects \(\mathbf{U}\), transformed by a link function \(g()\) is modeled as a linear combination \(\eta\) of the predictor variables \(\mathbf{X}\), the random effects \(\mathbf{U}\) and the model parameters \(\mathbf{\beta}\) (Fahrmeir et al., 2021). \[
g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta <!-- SG: Sorry for maybe a stupid question, but what does E stand for? and why are model parameters beta mentioned but not in the equation?-->
\] Equivalently, the conditional expected value is modelled as the linear combination \(\eta\), transformed by the inverse link function \(g^{-1}()\). \[
E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
\] If the dependent variable (i.e., diagnostic decision) \(Y\) is a binary variable with values \(0\) (i.e.~inaccurate), or \(1\) (i.e., accurate), the conditional expected value is equivalent to the probability: \[
P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
\] In our case study, \(P_{si}\) is the conditional probability that a subject \(s\) gives the correct response to item (i.e., CT scan) \(i\).

In such a setting, we model this probability as \[
P_{si} = inverse\_logit(\eta_{si})
\] with the inverse-logit link \(g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}\) or equivalently \[
logit(P_{si}) = \eta_{si}
\] with the logit link \(g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})\).

In our case study, the probability of giving a correct response is assumed to depend on the predictors:

\begin{itemize}
\tightlist
\item
  \(advice\_present_{si}\): whether subject \(s\) was presented with AI advice (1) or not (0) when asked to asses item \(i\)
\item
  \(advice\_correct_{si}\): whether this advice was correct (1) or not (0)
\item
  \(expert_s\): whether subject \(s\) was a task expert (1) or not (0)
\end{itemize}

and the random effects:

\begin{itemize}
\tightlist
\item
  \(u_{0s}\): the deviation of subject \(s\) from the average ability to solve an item (i.e., CT scan) with average difficulty; assumed to be distributed as \(u_{0s} \sim N(0, \sigma_S^2)\)
\item
  \(u_{0i}\): the deviation of item (i.e., CT scan) \(i\) from the average difficulty to be solved by a person with average ability; assumed to be distributed as \(u_{0i} \sim N(0, \sigma_I^2)\)
\end{itemize}

In total, we assume the model \[
\begin{aligned}
logit[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
\end{aligned}
\] or equivalently \[
\begin{aligned}
P_{si} = inverse\_logit[&(\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
\] with model parameters \(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\), \(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\).

In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors (\(advice\_present\), \(advice\_correct\)), one level-2 predictor (\(expert\)) and two cross-level interactions (\(expert \cdot advice\_present\), \(expert \cdot advice\_correct\)).

To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions here.

\hypertarget{data-simulation}{%
\subsection{Data simulation}\label{data-simulation}}

The following R function simulates a full dataset structured according to the design of our case study. The faux package (DeBruine, 2023) contains useful functions when simulating factorial designs, including random effects.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{simulate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{100}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{50}\NormalTok{,}
  \AttributeTok{b\_0 =} \FloatTok{0.847}\NormalTok{, }\AttributeTok{b\_e =} \FloatTok{1.350}\NormalTok{, }\AttributeTok{b\_a =} \SpecialCharTok{{-}}\FloatTok{1.253}\NormalTok{, }\AttributeTok{b\_c =} \FloatTok{2.603}\NormalTok{,}
  \AttributeTok{b\_ea =} \FloatTok{0.790}\NormalTok{, }\AttributeTok{b\_ec =} \SpecialCharTok{{-}}\FloatTok{1.393}\NormalTok{,}
  \AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{, ...)\{}
  \FunctionTok{require}\NormalTok{(dplyr)}
  \FunctionTok{require}\NormalTok{(faux)}
  \CommentTok{\# simulate design}
\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{add\_random}\NormalTok{(}\AttributeTok{subject =}\NormalTok{ n\_subjects, }\AttributeTok{item =}\NormalTok{ n\_items) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{add\_between}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{expert =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{.prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.25}\NormalTok{, }\FloatTok{0.75}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_present =} \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{, }\AttributeTok{prob =} \DecValTok{2}\SpecialCharTok{/}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_correct =} \FunctionTok{if\_else}\NormalTok{(advice\_present }\SpecialCharTok{==}\NormalTok{ 1L, }
                                    \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), 1L, }\AttributeTok{prob =} \FloatTok{0.8}\NormalTok{), 0L)) }\SpecialCharTok{\%\textgreater{}\%}
    \CommentTok{\# add random effects}
    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{u0s =}\NormalTok{ sd\_u0s) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"item"}\NormalTok{, }\AttributeTok{u0i =}\NormalTok{ sd\_u0i) }\SpecialCharTok{\%\textgreater{}\%}
    \CommentTok{\# compute dependent variable}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{linpred =}\NormalTok{ b\_0 }\SpecialCharTok{+}\NormalTok{ u0i }\SpecialCharTok{+}\NormalTok{ u0s }\SpecialCharTok{+}
\NormalTok{        b\_e }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{*}\NormalTok{ advice\_correct }\SpecialCharTok{+}
\NormalTok{        b\_ea }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_ec }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_correct) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_prob =} \FunctionTok{plogis}\NormalTok{(linpred)) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_bin =} \FunctionTok{rbinom}\NormalTok{(}\AttributeTok{n =} \FunctionTok{n}\NormalTok{(), }\AttributeTok{size =} \DecValTok{1}\NormalTok{, }\AttributeTok{prob =}\NormalTok{ y\_prob))}
\NormalTok{  dat}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

In the first six lines of the function definition, we set some default parameter values (which we will explain in a later section) and load the packages we use to manipulate and simulate data.
In our case study, each subject (\texttt{n\_subjects} in total) is assumed to respond to each item (i.e., CT scan; \texttt{n\_items} in total). Thus, the \texttt{add\_random} command creates a fully-crossed \texttt{data.frame} with \texttt{n\_subjects} \(\times\) \texttt{n\_items} rows. We add a between-subject effect with the \texttt{add\_between} command, simulating that about \(25\%\) of subjects are experts. The next two lines simulate that in \(\frac{2}{3}\) of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about \(80\%\) of cases (the variable \texttt{advice\_correct} is always 0 when no advice is presented). Next, we simulate one random effect for each subject (\texttt{u0s}) and for each item (\texttt{u0i}). As assumed by standard GLMMs, the \texttt{add\_ranef} function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user. With all design variables done, we are ready to simulate our model equation as outlined in equation X. The linear predictor variable \texttt{linpred} (\(\eta\) in the GLMM model equations) combines the predictor variables, random effects and model parameters as assumed by our model. We then transform the linear predictor with the inverse-link function to compute \texttt{y\_prob}, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with \texttt{plogis} and the logit link with \texttt{qlogis}). In the final step, we simulate the binary dependent variable \texttt{y\_bin} (i.e., whether the subject chooses the accurate diagnosis for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability \texttt{y\_prob}.

\hypertarget{model-fitting}{%
\subsection{Model fitting}\label{model-fitting}}

In this section, we show how to fit a GLMM with lme4, interpret the model and test hypotheses derived from a research question.

We simulate data according to our model, in which 100 subjects respond to 50 items (we use \texttt{set.seed} to make the simulation reproducible). However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyse this data.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{100}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{50}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(lme4)}
\NormalTok{f }\OtherTok{\textless{}{-}}\NormalTok{ y\_bin }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ expert }\SpecialCharTok{+}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ advice\_correct }\SpecialCharTok{+} 
\NormalTok{  expert}\SpecialCharTok{:}\NormalTok{advice\_present }\SpecialCharTok{+}\NormalTok{ expert}\SpecialCharTok{:}\NormalTok{advice\_correct }\SpecialCharTok{+}
\NormalTok{  (}\DecValTok{1}\SpecialCharTok{|}\NormalTok{subject) }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{|}\NormalTok{item)}
\end{Highlighting}
\end{Shaded}

The first two lines look similar to any linear model in R (general intercept indicated by \texttt{1}; main effects indicated by variable names in the dataset; interactions indicated by \texttt{variable1:variable2}). The third line specifies a random intercept for each subject \texttt{(1\textbar{}subject)} and for each item \texttt{(1\textbar{}item)}. The complete set of rules for the syntax is outlined in Bates et al. (2015) and in the documentation of the lme4 package.

In lme4, a GLMM is fitted with the \texttt{glmer} function. By setting \texttt{family\ =\ \ "binomial"}, we request a binomial GLMM appropriate for our binary dependent variable \texttt{y\_bin} (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs.~inaccurate (0) diagnosis.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(f, }\AttributeTok{data =}\NormalTok{ dat, }\AttributeTok{family =} \StringTok{"binomial"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{model-interpretation}{%
\subsection{Model interpretation}\label{model-interpretation}}

We can inspect the estimates for all model parameters with the \texttt{summary} command:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{summary}\NormalTok{(fit)}
\end{Highlighting}
\end{Shaded}

In the output, the \texttt{Estimate} column in the \texttt{Fixed\ effects} table contains the estimates for the \(\beta\) parameters, while the \texttt{Std.Dev.} column in the \texttt{Random\ effects} table contains the estimates for \(\sigma_S\) and \(\sigma_I\).

Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the \(\beta\) parameters are connected to the modelled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects. The most simple interpretation works by imagining a subject with average ability (\(u_{0s} = 0\)) responding to an item (i.e., CT scan) with average difficulty (\(u_{0i} = 0\)). Then the model implied probability that such a person solves such an item accurately is given by: \[
\begin{aligned}
P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
= inverse\_logit[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
\]

In fact, we would only need the full equation if the subject is an expert and correct advice is presented. In all other experimental conditions, some terms drop from the equation because they are multiplied by \(0\). The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented: \[
\begin{aligned}
P(Y=1|expert = 0, advice\_present = 0, advice\_correct = 0, u_{0s} = 0, u_{0i} = 0) = \\
= inverse\_logit[\beta_0]
\end{aligned}
\]

Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs. Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.

With the marginaleffects package (Arel-Bundock, 2023), we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed \textbf{and} random effects), and plot the average probability with confidence intervals for each experimental condition:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(marginaleffects)}
\FunctionTok{plot\_predictions}\NormalTok{(fit, }\AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"advice\_present"}\NormalTok{, }\StringTok{"advice\_correct"}\NormalTok{, }\StringTok{"expert"}\NormalTok{),}
  \AttributeTok{type =} \StringTok{"response"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{ylim}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FloatTok{0.3}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{hypothesis-testing}{%
\subsection{Hypothesis testing}\label{hypothesis-testing}}

However, we need to think about the model parameters again when we want to test hypotheses that we have theoretically derived from some research question. Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.

The \texttt{Fixed\ effects} table in Figure X also includes p-values for hypothesis tests with null hypotheses of the style \(H_0: \beta = 0\). However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.

For our case study, imagine the following combined hypothesis: \emph{We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.}

This combined hypothesis leads to the following four separate null hypotheses to be tested: \[
\begin{aligned}
H_{01}:\ &\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0 \\
H_{02}:\ &\beta_{a} + \beta_{c} \leq 0 \\
H_{03}:\ &\beta_{a} + \beta_{ea} \geq 0 \\
H_{04}:\ &\beta_{a} \geq 0
\end{aligned}
\]

We arrive at these inequalities based on the following logic, exemplified here only for \(H_{01}\): The first null hypothesis states that \emph{an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.} This implies the following inequality for each subject \(s\) and item \(i\) \[
inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}] \leq inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e]
\] which simplifies to \(\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0\).

We can specify and test hypotheses like these with the multcomp package (Hothorn, Bretz, \& Westfall, 2008a) as follows:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(multcomp)}
\NormalTok{null\_hypotheses }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"advice\_present + advice\_correct + expert:advice\_present + }
\StringTok{   expert:advice\_correct \textless{}= 0"}\NormalTok{,}
  \StringTok{"advice\_present + advice\_correct \textless{}= 0"}\NormalTok{,}
  \StringTok{"{-}1 * (advice\_present + expert:advice\_present) \textless{}= 0"}\NormalTok{,}
  \StringTok{"{-}1 * (advice\_present) \textless{}= 0"}\NormalTok{)}
\NormalTok{glht }\OtherTok{\textless{}{-}} \FunctionTok{glht}\NormalTok{(fit, }\AttributeTok{linfct =}\NormalTok{ null\_hypotheses)}
\FunctionTok{summary}\NormalTok{(glht, }\AttributeTok{test =} \FunctionTok{univariate}\NormalTok{())}\SpecialCharTok{$}\NormalTok{test}\SpecialCharTok{$}\NormalTok{pvalues}
\end{Highlighting}
\end{Shaded}

Because all hypotheses tested simultaneously with the \texttt{glht} function must have the same direction, we flip the sign of inequalities three and four by multiplying them with \(-1\). The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously (Hothorn, Bretz, \& Westfall, 2008b). However, the combined null hypothesis in our exemplary research question should only be rejected if \textbf{all} individual null hypotheses are rejected {[}i.e.~intersection-union setting; Dmitrienko and D'Agostino (2013){]}. In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests. Thus, we request unadjusted p-values by setting \texttt{test\ =\ univariate()} in the \texttt{summary} command. With a standard significance level of \(\alpha = 0.05\), we would not reject all four null hypotheses (the p-value for hypothesis \(H_{03}\) is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset. Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.

\hypertarget{specification-of-plausible-parameter-values}{%
\subsection{Specification of plausible parameter values}\label{specification-of-plausible-parameter-values}}

When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters (\(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\), \(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\)), but have not talked about where these numbers came from.

Often no meta-analytic results or conclusive data from pilot studies are available. All parameter values in our present case study have been determined based on results from prior own work and from the literature. Additionally, we had repeated exchanges with our affiliated domain experts from radiology.

We now outline a few strategies on how to determine plausible parameter values. We have already seen in our discussion of model interpretation how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty. We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (for an average subject and an average item).

Table X shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model:
\textbar{} \emph{Experimental condition} \textbar{} \(P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})\) \textbar{} \emph{Implied equation} \textbar{}
\textbar--------------------\textbar:------------------------------:\textbar--------------------\textbar{}
\textbar{} no advice, no expert \textbar{} 0.70 \textbar{} \(logit(0.70) = \beta_0\) \textbar{}
\textbar{} no advice, expert \textbar{} 0.90 \textbar{} \(logit(0.90) = \beta_0 + \beta_e\) \textbar{}
\textbar{} false advice, no expert \textbar{} 0.40 \textbar{} \(logit(0.40) = \beta_0 + \beta_a\) \textbar{}
\textbar{} false advice, expert \textbar{} 0.85 \textbar{} \(logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \textbar{}
\textbar{} correct advice, no expert \textbar{} 0.90 \textbar{} \(logit(0.90) = \beta_0 + \beta_a + \beta_c\) \textbar{}
\textbar{} correct advice, expert \textbar{} 0.95 \textbar{} \(logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\) \textbar{}

This table can be used to compute the implied values for the \(\beta\) parameters, starting with the first equation and reinserting the computed \(\beta\) values in all following equations:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{b\_0 }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.7}\NormalTok{)}
\NormalTok{b\_e }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
\NormalTok{b\_a }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.4}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
\NormalTok{b\_ea }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.85}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a}
\NormalTok{b\_c }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_a}
\NormalTok{b\_ec }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.95}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a }\SpecialCharTok{{-}}\NormalTok{ b\_c }\SpecialCharTok{{-}}\NormalTok{ b\_ea}
\FunctionTok{c}\NormalTok{(}\AttributeTok{b\_0 =}\NormalTok{ b\_0, }\AttributeTok{b\_e =}\NormalTok{ b\_e, }\AttributeTok{b\_a =}\NormalTok{ b\_a, }\AttributeTok{b\_c =}\NormalTok{ b\_c, }\AttributeTok{b\_ea =}\NormalTok{ b\_ea, }\AttributeTok{b\_ec =}\NormalTok{ b\_ec)}
\end{Highlighting}
\end{Shaded}

It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g.~
\[
\begin{aligned}
P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
\end{aligned}
\]

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plogis}\NormalTok{(b\_0 }\SpecialCharTok{+}\NormalTok{ b\_e }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{+}\NormalTok{ b\_ea }\SpecialCharTok{+}\NormalTok{ b\_ec)}
\end{Highlighting}
\end{Shaded}

Although the derivations above are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is \textbf{not} equal to the probability that a person with average ability solves an item with average difficulty. The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation. For example, we determined the \(\beta\) parameters in a way that corresponds to a desired conditional probability of \(0.95\), that an expert with average ability solves an item with average difficulty when presented with correct advice. However, even if the model were true, we would not observe this probability value if we estimated the marginal probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.

The inequality of conditional and marginal effects in GLMMs (Fahrmeir et al., 2021) makes their interpretation more difficult. One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans). However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge. Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.

Earlier, we have already encountered one way to visualize the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest. However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory. A more sophisticated strategy is to simulate a large dataset and directly compute the averages, contrasts and distributions we are interested in.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(ggdist)}
\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{2000}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{2000}\NormalTok{, }\AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{)}
\NormalTok{dat }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
    \StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
    \StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
    \StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ y\_prob, }\AttributeTok{y =}\NormalTok{ condition)) }\SpecialCharTok{+}
  \FunctionTok{stat\_histinterval}\NormalTok{(}\AttributeTok{point\_interval =} \StringTok{"mean\_qi"}\NormalTok{, }\AttributeTok{slab\_color =} \StringTok{"gray45"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{breaks =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FloatTok{0.1}\NormalTok{), }\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

Figure X shows the model implied marginal distributions, including the mean, 66\% and 95\% intervals. We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section. This difference increases with the variability of the random effects.

Up to this point, we have not talked about plausible values for the standard deviations of the subject and item random intercepts (\(\sigma_S\) and \(\sigma_I\)). Plots like the one above are a useful tool to decide whether the specified standard deviations are reasonable by comparing the ranges and overlap between conditions to domain knowledge.

In the next plot, we have set the item standard deviation to almost zero (\(\sigma_I = 0.01\)). This gives us a better way to see the variability between persons.

As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty. In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice. Patterns like these should be considered carefully and discussed with the domain experts. Parameter values (\(\beta\) parameters, and \(\sigma_S\)) should be adjusted if the implications do not seem reasonable.

We could also have a closer look at variability between items by setting the subject standard deviation to almost zero (\(\sigma_S = 0.01\), see Figure X in Appendix X).

The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high. For this example, we have set \(\sigma_S = 3\) and \(\sigma_I = 3\). This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption. For example, we would expect even experts with low ability to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.

\hypertarget{results}{%
\section{Results}\label{results}}

With all these considerations addressed, we are finally ready to perform a power analysis. Wrapping the \texttt{simulate} function already constructed earlier, the helper function \texttt{sim\_and\_analyse} performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sim\_and\_analyse }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}
  \AttributeTok{formula\_chr =} \StringTok{"y\_bin \textasciitilde{} 1 + expert + advice\_present + advice\_correct + }
\StringTok{    expert:advice\_present + expert:advice\_correct + (1|subject) + (1|item)"}\NormalTok{, }
  \AttributeTok{null\_hypotheses =} \FunctionTok{c}\NormalTok{(}\StringTok{"advice\_present + advice\_correct + }
\StringTok{    expert:advice\_present + expert:advice\_correct \textless{}= 0"}\NormalTok{,}
    \StringTok{"advice\_present + advice\_correct \textless{}= 0"}\NormalTok{,}
    \StringTok{"{-}1 * (advice\_present + expert:advice\_present) \textless{}= 0"}\NormalTok{,}
    \StringTok{"{-}1 * (advice\_present) \textless{}= 0"}\NormalTok{), ...)\{}
  \FunctionTok{require}\NormalTok{(lme4)}
  \FunctionTok{require}\NormalTok{(multcomp)}
  \CommentTok{\# simulate data}
\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(...)}
  \CommentTok{\# fit model}
\NormalTok{  model }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(}\FunctionTok{as.formula}\NormalTok{(formula\_chr), }\AttributeTok{data =}\NormalTok{ dat, }\AttributeTok{family =} \StringTok{"binomial"}\NormalTok{)}
  \CommentTok{\# compute p{-}values}
\NormalTok{  glht }\OtherTok{\textless{}{-}} \FunctionTok{glht}\NormalTok{(model, }\AttributeTok{linfct =}\NormalTok{ null\_hypotheses)}
\NormalTok{  pvalues }\OtherTok{\textless{}{-}} \FunctionTok{summary}\NormalTok{(glht, }\AttributeTok{test =} \FunctionTok{univariate}\NormalTok{())}\SpecialCharTok{$}\NormalTok{test}\SpecialCharTok{$}\NormalTok{pvalues}
  \FunctionTok{setNames}\NormalTok{(pvalues, }\FunctionTok{paste0}\NormalTok{(}\StringTok{"p\_H0"}\NormalTok{, }\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(null\_hypotheses)))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Power analysis can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes. Thus, we use the future (Bengtsson, 2021) and furrr (Vaughan \& Dancho, 2022) packages to perform computations in parallel. First, we enable parallelization and specify how many parallel cores (``workers'') of our computer to use (users can find out the maximum number of cores on their computer with the command \texttt{parallel::detectCores()}), and set a seed to make the simulation reproducible.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(future)}
\FunctionTok{plan}\NormalTok{(}\StringTok{"multisession"}\NormalTok{, }\AttributeTok{workers =} \DecValTok{6}\NormalTok{)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The next code chunk specifies a simulation grid with different settings for both the number of subjects (\texttt{n\_subjects}) and the number of items (\texttt{n\_items}), each combination being repeated \texttt{rep} times. We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
With the current settings, this simulation takes about X minutes on a MacBook Pro from 2020 with M1 chip and 16 GB working memory. If you want to quickly experiment with the code yourself, a setting with \texttt{workers\ =\ 4} and \texttt{rep\ =\ 5} should finish in less than 5 minutes, even on smaller machines.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(furrr)}
\NormalTok{sim\_design }\OtherTok{\textless{}{-}} \FunctionTok{crossing}\NormalTok{(}
  \AttributeTok{rep =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{300}\NormalTok{,}
  \AttributeTok{n\_subjects =} \FunctionTok{c}\NormalTok{(}\DecValTok{100}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{250}\NormalTok{),}
  \AttributeTok{n\_items =} \FunctionTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{50}\NormalTok{, }\DecValTok{70}\NormalTok{)}
\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{pvalues =} \FunctionTok{future\_pmap}\NormalTok{(., sim\_and\_analyse, }
    \AttributeTok{.options =} \FunctionTok{furrr\_options}\NormalTok{(}\AttributeTok{seed =} \ConstantTok{TRUE}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{unnest\_wider}\NormalTok{(pvalues)}
\end{Highlighting}
\end{Shaded}

The result of the computation is a data.frame that contains the p-values of all tested hypotheses for each simulated dataset.
In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package. We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.

For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the \(\alpha = 0.05\) level. Based on our simulation outcomes, we compute a power estimate for each combination of \texttt{n\_subjects} \(\times\) \texttt{n\_items} (including 95\% confidence intervals) and visualize the results with the following code.\footnote{This code was inspired by the ``Mixed Design Simulation'' vignette of the faux package at \url{https://debruine.github.io/faux/articles/sim_mixed.html}.}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(binom)}
\NormalTok{alpha }\OtherTok{\textless{}{-}} \FloatTok{0.05}
\NormalTok{power }\OtherTok{\textless{}{-}}\NormalTok{ sim\_design }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{power =} \FunctionTok{mean}\NormalTok{(p\_H01 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}\NormalTok{ p\_H02 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
\NormalTok{                           p\_H03 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}\NormalTok{ p\_H04 }\SpecialCharTok{\textless{}}\NormalTok{ alpha), }
    \AttributeTok{n\_sig =} \FunctionTok{sum}\NormalTok{(p\_H01 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}\NormalTok{ p\_H02 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
\NormalTok{                  p\_H03 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}\NormalTok{ p\_H04 }\SpecialCharTok{\textless{}}\NormalTok{ alpha),}
    \AttributeTok{n =} \FunctionTok{n}\NormalTok{(),}
    \AttributeTok{ci.lwr =} \FunctionTok{binom.confint}\NormalTok{(n\_sig, n, }\AttributeTok{method =} \StringTok{"wilson"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{lower,}
    \AttributeTok{ci.upr =} \FunctionTok{binom.confint}\NormalTok{(n\_sig, n, }\AttributeTok{method =} \StringTok{"wilson"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{upper, }
    \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{)}
\NormalTok{power }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(n\_subjects, n\_items, }\AttributeTok{fill =}\NormalTok{ power)) }\SpecialCharTok{+}
  \FunctionTok{geom\_tile}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{label =} \FunctionTok{sprintf}\NormalTok{(}\StringTok{"\%.2f }\SpecialCharTok{\textbackslash{}n}\StringTok{ [\%.2f; \%.2f]"}\NormalTok{, }
\NormalTok{                                power, ci.lwr, ci.upr)), }
    \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \FloatTok{4.5}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{xlab}\NormalTok{(}\StringTok{"number of subjects"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{ylab}\NormalTok{(}\StringTok{"number of items"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

As should be the case, power estimates increase with both the number of subjects and the number of items. The confidence intervals indicate how precisely power was estimated by our simulation. Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (\texttt{rep}) in the simulation. In practice, power simulations are often run multiple times with adjusted combinations of sample sizes. When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of \texttt{n\_subjects} and \texttt{n\_items}. When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of power simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions. This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.

Much has been written on the optimal amount of power to target in empirical research. The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of \(\alpha = 0.05\)), but depending on the research goals of the study, there are often good reasons to move away from this standard (Lakens, 2022b; Lakens et al., 2018). When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations. For the sake of the example, let the targeted power be indeed about 0.8, using an \(\alpha\) of 0.05 to detect an effect of the expected size implied by our data simulation. This could be achieved by collecting data from 200 subjects (about 25\% of which will be experts), each completing the same 50 items (with advice present in about 67\% of cases, which is correct in about 80\% of cases with present advice). If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items. However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule. The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\textgreater{} 250, to be determined by additional rounds of power analysis). Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Experimental research requires careful planning and consideration of statistical power to ensure robust and meaningful results. While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required. The present tutorial presents a specific example of how to run power simulations for experimental designs with GLMMs.

\hypertarget{expected-effect-size-vs.-smallest-effect-size-of-interest-sensitivity-power-analysis}{%
\subsection{Expected effect size vs.~smallest effect size of interest: sensitivity power analysis}\label{expected-effect-size-vs.-smallest-effect-size-of-interest-sensitivity-power-analysis}}

In our case study, we have performed power simulations based on a single set of parameter values that reflect our assumptions of an expected effect size. Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge. Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data. When sample sizes are chosen based on the results of our power simulations, a future study will be informative to reject the null hypothesis if an effect of our \emph{expected size} is present. However, if the true effect is indeed smaller, the power will be lower, and the study might not be sufficiently informative. A common, more conservative strategy for sample size justification is to perform power analysis for the smallest effect size of interest (SESOI). An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero (King, 2011). For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see (Lakens, 2022a). When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called \emph{sensitivity power analysis}. By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power depends on the size of the true effect. In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyse power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8). However, for our case study that investigates combined hypotheses in a GLMM modelling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions. In this setting, sensitivity power analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question. Power could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).

\hypertarget{outlook}{%
\subsection{Outlook}\label{outlook}}

Beyond the specifics of our concrete case study, we want to outline six avenues regarding the future of power simulations in experimental research:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  The need for power simulations in experimental research: Experimental research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software. Power simulations offer a solution by providing experimental researchers with a tailored approach to estimating statistical power. These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
\item
  Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
  This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions. By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
\item
  Teaching power simulation skills: The ability to conduct power simulations is a valuable skill that should be taught to experimental researchers. By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
  Equipping researchers with the knowledge and tools to perform power simulations enables them to make informed decisions and enhance the rigor of their studies.
  The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look like, might also contribute to planning more insightful studies that can answer more precise research questions (Yarkoni, 2022).
\item
  Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and the perceived effort estimated by researchers and collaborators in experimental research. Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations. It is crucial to raise awareness about the effort involved in power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
\item
  Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research. They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power. The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports .
\item
  Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in experimental research. When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous. By preregistering their power simulations, researchers enhance transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
\end{enumerate}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Power simulations play a critical role in experimental research, allowing investigators to tailor power estimation to the unique aspects of their experiments. Through this tutorial, we aim to provide researchers with the necessary skills and tools to perform these simulations for GLMMs themselves. By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work (Yarkoni, 2022).

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-R-marginaleffects}{}}%
Arel-Bundock, V. (2023). \emph{Marginaleffects: Predictions, comparisons, slopes, marginal means, and hypothesis tests}. Retrieved from \url{https://CRAN.R-project.org/package=marginaleffects}

\leavevmode\vadjust pre{\hypertarget{ref-arendStatisticalPowerTwolevel2019}{}}%
Arend, M. G., \& Schäfer, T. (2019). Statistical power in two-level models: {A} tutorial based on {Monte Carlo} simulation. \emph{Psychological Methods}, \emph{24}(1), 1--19. \url{https://doi.org/10.1037/met0000195}

\leavevmode\vadjust pre{\hypertarget{ref-batesFittingLinearMixedEffects2015}{}}%
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting {Linear Mixed-Effects Models Using} {\textbf{Lme4}}. \emph{Journal of Statistical Software}, \emph{67}(1). \url{https://doi.org/10.18637/jss.v067.i01}

\leavevmode\vadjust pre{\hypertarget{ref-R-RJ-2021-048}{}}%
Bengtsson, H. (2021). A unifying framework for parallel and distributed processing in r using futures. \emph{The R Journal}, \emph{13}(2), 208--227. \url{https://doi.org/10.32614/RJ-2021-048}

\leavevmode\vadjust pre{\hypertarget{ref-brownIntroductionLinearMixedEffects2021}{}}%
Brown, V. A. (2021). An {Introduction} to {Linear Mixed-Effects Modeling} in {R}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{4}(1), 2515245920960351. \url{https://doi.org/10.1177/2515245920960351}

\leavevmode\vadjust pre{\hypertarget{ref-brysbaertPowerAnalysisEffect2018}{}}%
Brysbaert, M., \& Stevens, M. (2018). Power {Analysis} and {Effect Size} in {Mixed Effects Models}: {A Tutorial}. \emph{Journal of Cognition}, \emph{1}(1), 9. \url{https://doi.org/10.5334/joc.10}

\leavevmode\vadjust pre{\hypertarget{ref-burknerBrmsPackageBayesian2017}{}}%
Bürkner, P.-C. (2017). Brms: {An R Package} for {Bayesian Multilevel Models Using Stan}. \emph{Journal of Statistical Software}, \emph{80}, 1--28. \url{https://doi.org/10.18637/jss.v080.i01}

\leavevmode\vadjust pre{\hypertarget{ref-R-faux}{}}%
DeBruine, L. (2023). \emph{Faux: Simulation for factorial designs}. Zenodo. \url{https://doi.org/10.5281/zenodo.2669586}

\leavevmode\vadjust pre{\hypertarget{ref-debruineUnderstandingMixedEffectsModels2021}{}}%
DeBruine, L., \& Barr, D. J. (2021). Understanding {Mixed-Effects Models Through Data Simulation}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{4}(1), 2515245920965119. \url{https://doi.org/10.1177/2515245920965119}

\leavevmode\vadjust pre{\hypertarget{ref-dmitrienkoTraditionalMultiplicityAdjustment2013}{}}%
Dmitrienko, A., \& D'Agostino, R. (2013). Traditional multiplicity adjustment methods in clinical trials. \emph{Statistics in Medicine}, \emph{32}(29), 5172--5218. \url{https://doi.org/10.1002/sim.5990}

\leavevmode\vadjust pre{\hypertarget{ref-fahrmeirRegressionModelsMethods2021}{}}%
Fahrmeir, L., Kneib, T., Lang, S., \& Marx, B. D. (2021). \emph{Regression: {Models}, {Methods} and {Applications}}. {Berlin, Heidelberg}: {Springer Berlin Heidelberg}. \url{https://doi.org/10.1007/978-3-662-63882-8}

\leavevmode\vadjust pre{\hypertarget{ref-faulStatisticalPowerAnalyses2009}{}}%
Faul, F., Erdfelder, E., Buchner, A., \& Lang, A.-G. (2009). Statistical power analyses using {G}*{Power} 3.1: {Tests} for correlation and regression analyses. \emph{Behavior Research Methods}, \emph{41}(4), 1149--1160. \url{https://doi.org/10.3758/BRM.41.4.1149}

\leavevmode\vadjust pre{\hypertarget{ref-greenSIMRPackagePower2016}{}}%
Green, P., \& MacLeod, C. J. (2016). {SIMR}: An {R} package for power analysis of generalized linear mixed models by simulation. \emph{Methods in Ecology and Evolution}, \emph{7}(4), 493--498. \url{https://doi.org/10.1111/2041-210X.12504}

\leavevmode\vadjust pre{\hypertarget{ref-hothornSimultaneousInferenceGeneral2008}{}}%
Hothorn, T., Bretz, F., \& Westfall, P. (2008b). Simultaneous {Inference} in {General Parametric Models}. \emph{Biometrical Journal}, \emph{50}(3), 346--363. \url{https://doi.org/10.1002/bimj.200810425}

\leavevmode\vadjust pre{\hypertarget{ref-R-multcomp}{}}%
Hothorn, T., Bretz, F., \& Westfall, P. (2008a). Simultaneous inference in general parametric models. \emph{Biometrical Journal}, \emph{50}(3), 346--363.

\leavevmode\vadjust pre{\hypertarget{ref-kapteinUsingGeneralizedLinear2016}{}}%
Kaptein, M. (2016). Using {Generalized Linear} ({Mixed}) {Models} in {HCI}. In J. Robertson \& M. Kaptein (Eds.), \emph{Modern {Statistical Methods} for {HCI}} (pp. 251--274). {Cham}: {Springer International Publishing}. \url{https://doi.org/10.1007/978-3-319-26633-6_11}

\leavevmode\vadjust pre{\hypertarget{ref-kingPointMinimalImportant2011}{}}%
King, M. T. (2011). A point of minimal important difference ({MID}): A critique of terminology and methods. \emph{Expert Review of Pharmacoeconomics \& Outcomes Research}, \emph{11}(2), 171--184. \url{https://doi.org/10.1586/erp.11.9}

\leavevmode\vadjust pre{\hypertarget{ref-kumleEstimatingPowerGeneralized2021}{}}%
Kumle, L., Võ, M. L.-H., \& Draschkow, D. (2021). Estimating power in (generalized) linear mixed models: {An} open introduction and tutorial in {R}. \emph{Behavior Research Methods}, \emph{53}(6), 2528--2543. \url{https://doi.org/10.3758/s13428-021-01546-0}

\leavevmode\vadjust pre{\hypertarget{ref-lafitSelectionNumberParticipants2021}{}}%
Lafit, G., Adolf, J. K., Dejonckheere, E., Myin-Germeys, I., Viechtbauer, W., \& Ceulemans, E. (2021). Selection of the {Number} of {Participants} in {Intensive Longitudinal Studies}: {A User-Friendly Shiny App} and {Tutorial} for {Performing Power Analysis} in {Multilevel Regression Models That Account} for {Temporal Dependencies}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{4}(1), 251524592097873. \url{https://doi.org/10.1177/2515245920978738}

\leavevmode\vadjust pre{\hypertarget{ref-lakensImprovingYourStatistical2022}{}}%
Lakens, D. (2022a). \emph{Improving {Your Statistical Inferences}}. Zenodo. \url{https://doi.org/10.5281/ZENODO.6409077}

\leavevmode\vadjust pre{\hypertarget{ref-lakensSampleSizeJustification2022}{}}%
Lakens, D. (2022b). Sample {Size Justification}. \emph{Collabra: Psychology}, \emph{8}(1), 33267. \url{https://doi.org/10.1525/collabra.33267}

\leavevmode\vadjust pre{\hypertarget{ref-lakensJustifyYourAlpha2018}{}}%
Lakens, D., Adolfi, F. G., Albers, C. J., Anvari, F., Apps, M. A. J., Argamon, S. E., \ldots{} Zwaan, R. A. (2018). Justify your alpha. \emph{Nature Human Behaviour}, \emph{2}(3), 168--171. \url{https://doi.org/10.1038/s41562-018-0311-x}

\leavevmode\vadjust pre{\hypertarget{ref-maxwellSampleSizePlanning2008}{}}%
Maxwell, S. E., Kelley, K., \& Rausch, J. R. (2008). Sample {Size Planning} for {Statistical Power} and {Accuracy} in {Parameter Estimation}. \emph{Annual Review of Psychology}, \emph{59}(1), 537--563. \url{https://doi.org/10.1146/annurev.psych.59.103006.093735}

\leavevmode\vadjust pre{\hypertarget{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{}}%
Murayama, K., Usami, S., \& Sakaki, M. (2022). Summary-statistics-based power analysis: {A} new and practical method to determine sample size for mixed-effects modeling. \emph{Psychological Methods}. \url{https://doi.org/10.1037/met0000330}

\leavevmode\vadjust pre{\hypertarget{ref-robertsonImprovingStatisticalPractice2016}{}}%
Robertson, J., \& Kaptein, M. (2016). Improving {Statistical Practice} in {HCI}. In J. Robertson \& M. Kaptein (Eds.), \emph{Modern {Statistical Methods} for {HCI}} (pp. 331--348). {Cham}: {Springer International Publishing}. \url{https://doi.org/10.1007/978-3-319-26633-6_14}

\leavevmode\vadjust pre{\hypertarget{ref-R-furrr}{}}%
Vaughan, D., \& Dancho, M. (2022). \emph{Furrr: Apply mapping functions in parallel using futures}. Retrieved from \url{https://CRAN.R-project.org/package=furrr}

\leavevmode\vadjust pre{\hypertarget{ref-westfallStatisticalPowerOptimal2014}{}}%
Westfall, J., Kenny, D. A., \& Judd, C. M. (2014). Statistical power and optimal design in experiments in which samples of participants respond to samples of stimuli. \emph{Journal of Experimental Psychology: General}, \emph{143}, 2020--2045. \url{https://doi.org/10.1037/xge0000014}

\leavevmode\vadjust pre{\hypertarget{ref-yarkoniGeneralizabilityCrisis2022}{}}%
Yarkoni, T. (2022). The generalizability crisis. \emph{Behavioral and Brain Sciences}, \emph{45}, e1. \url{https://doi.org/10.1017/S0140525X20001685}

\leavevmode\vadjust pre{\hypertarget{ref-zimmerSampleSizePlanning2022}{}}%
Zimmer, F., Henninger, M., \& Debelak, R. (2022). \emph{Sample {Size Planning} for {Complex Study Designs}: {A Tutorial} for the mlpwr {Package}}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/r9w6t}

\end{CSLReferences}

\end{document}

### 2023-09-12 docs/manuscript.Rmd
Timo's edits based on Monday meeting and comments

@@
 ---
-title             : "A Tutorial on Tailored Power Simulations for Experimental Study Designs with Generalized Linear Mixed Models"
-shorttitle        : "GLMM Power Simulation Tutorial"
+title             : "A Tutorial on Tailored Simulation-Based Power Analysis for Experimental Designs with Generalized Linear Mixed Models"
+shorttitle        : "Simulation-Based Power Analysis for GLMMs"
 
@@
 
-wordcount         : "7059"
+wordcount         : "7037"
 
@@
 <!-- SG: In the intro power analysis, a priori power analysis and sample size calculation is now used interchangeably, should we use one term consistently?-->
+<!-- TK: I have repharsed this - should be clearer now -->
 
 <!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
-When planning experimental research, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data.
-However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), <!-- SG: is the HCI needed here?
-Also, can we claim a general replication crisis in experimental research?
-For instance, have no idea if experimental physics also has a problem.
-Maybe "in several disciplines.
-AK: I agree with Susanne.
-I am not sure whether the HCI is needed.
-Seems a bit arbitrary.
-Why does this apply to HCI in particular?
-Either work with an example or keep more generel --> has illustrated many challenges surrounding the reproducibility and reliability of findings [@robertsonImprovingStatisticalPractice2016].
+When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data [@lakensSampleSizeJustification2022].
+However, the recent replication crisis in Human-Computer Interaction (HCI) and several other disciplines grounded on empirical research has illustrated many challenges surrounding the reproducibility and reliability of findings [@cockburnThreatsReplicationCrisis2020; @robertsonImprovingStatisticalPractice2016].
 As a result, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
-While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework.
-In the present work, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
-<!-- AK: Great, I like the Intro.
-I take my previous comment back - I think this can stay as it is now, without examples.
-I feel that it is not tailored towards HCI, but I also don't think that's a big issue.-->
+While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
+In the present work, we provide a tutorial comprised of a concrete example for tailored a priori power analyses using data simulations based on GLMMs.
 
@@
 But up to you!
---> If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated <!-- SG: I think there is something wrong with the logic of this sentence.
-Should it be like this?
-If the sample size used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered"-->, this particular study would be considered "underpowered".<!--AK: I agree with Susanne.
-This is not a sentence anymore.
--->
+TK: FP what do you like better?
+Please insert the one you think fits better.
+--> If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, this particular study would be considered "underpowered".
 
@@
 Conducting underpowered research has many negative consequences.
-First, relying on underpowered experiments may yield inconclusive (if low power is taken into account<!--AK: Unclear what this means in this context-  how is it "taken into account"?
+First, relying on underpowered experiments may yield inconclusive (if low power is taken into account<!--AK: Unclear what this means in this context-  how is it "taken into account"?, TK: FP can you clarify what is meant here?
 -->) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
@@
 A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
-By running the calculation for a range of sample sizes "a priori" (i.e., before data collection), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
-<!--AK: I am not sure, but isn't this the definition of power simulation more or less?
-Maybe you could introduce that here already.
--->
-Thereby, power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.<!--AK: This is already an advantage.
-I would combine this with the part below.-->
+When running a power analysis before data collection (i.e., "a priori"), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
 
 <!-- advantages of an a priori power analysis -->
-Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
+Thereby, a priori power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
+Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
 Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
-Especially in fast-moving research fields<!--AK: I don't understand how this applies to fast-moving research fields (btw, what is that??) in particular.
-I would delere this.-->, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.
-
-<!-- external incentives to do a power analysis -->
-Moreover, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights.
-<!-- FP: this last sentence seems like a repetition.
-wouldn't some sentence better fit here stating that proficiency with power analysis improves your career chances in academia by increasing the probability to get papers published and grants funded (which is a new argument we haven't already made elsewhere); TK: I like this argument.
-Might be a little idealistic though since bc flashy results get you ahead mostly :D  AK: I agree, the first part is repetitive.
--->
+Here, adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
+Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+Therefore, proficiency with power analysis can aid one's academic career chances.
 
 <!-- standard software is available for "simple" models and study designs -->
-For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009].
-<!-- FP: we could reintroduce the old section her on: with more complex models and designs, power analysis is more complicated because no analytic solution is possible, power has to be estimated by simulation, and there is no user-friendly software that includes all relevant designs.; TK: yes let's do this.
-I am not entirely sure what you mean with "analytic" solution, i think we should define this.
-SG: I am not sure what the plan is here.
-This sentence alone makes little sense.
-The argument that these tools cannot be used or adjusted for more complex models is missing.
-If you want to link it to the next section, it would argue that these simple statistical methods are often not sufficient for researchers to answer their research questions and that more complex designs are required.
-I would not go in detail here, that no tools for GLMM exist, since this argument is already mentioned below; AK: I agree with Susanne.-->
+For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori power analysis is readily available [@faulStatisticalPowerAnalyses2009].
+Yet, these simple statistical methods are often not sufficient for research questions requiring more complex designs and methods.
+However, software-based power analysis tools cannot be used or adjusted for those.
+<!-- TK: FP do you like the previous section how it is now? Please adjust, if required -->
 
-## Generalized linear mixed models (GLMM)
+## Generalized linear mixed models (GLMMs)
 
@@
 I would say it's the latter.-->, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI<!--AK: This "such as in HCI" always sounds super weird.
-It becomes apparant that you need to mention this for the submission.
-I would eitehr remove this entirely or explain this more in context of HCI research (using an example from that field)-->, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
+GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data in HCI and other empirical disciplines because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
 
@@
 As mentioned, GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family (e.g., binomial) <!-- SG: do you think examples for the different types of variables are needed-->.
+They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family (e.g., binomial) <!-- SG: do you think examples for the different types of variables are needed; TK: i don't think so -->.
 GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.<!--AK: And here I would add one research example.
+The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.
+As an example, think of an experiment on different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
+Here, participants' interaction is measured repeatedly (e.g., over several sessions).
+Also, the interactions of participants in one session are likely to be correlated with their previous sessions.
+Finally, the outcome variable is binary (click/no click) for each interaction, which follows a binomial distribution.
+
+<!--AK: And here I would add one research example.
 Something like, "Imagine an experiment where experts are repeatedly asked for...
@@
 I would add an example that explains both aspects you talk about before: 1) The dependency, 2) the outcome distribution (e.g., binomial).
-I think this would make it much easier for the reader to follow you here.-->
+I think this would make it much easier for the reader to follow you here.
+TK: FP how do you like the new example? Feel free to adjust -->
 
-## Power simulations for GLMMs
+## Power analysis for GLMMs
 
@@
 A number of tutorials have been published describing how to perform such power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE).
-This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common <!--AK: Are very common how?
-Unclear sentence-->[@brownIntroductionLinearMixedEffects2021].
+However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->).
+This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->) [@brownIntroductionLinearMixedEffects2021].
 The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
@@
 Additionally, assumptions about the error structure, such as independence or correlation, must be specified.
-Interpreting these assumptions entails understanding the underlying assumptions <!--AK: Repeated "assumptions".
-Find other word.-->of the model and ensuring they align with the characteristics of the data being analyzed.
+Interpreting these assumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
 Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions <!--AK: Again "assumptions".-->a practical challenge [@maxwellSampleSizePlanning2008].
+However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions <!--AK: Again "assumptions".--> a practical challenge [@maxwellSampleSizePlanning2008].
 
 <!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
-Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design<!--AK: I would just write GLMM.-->.
+Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power simulation with GLMMs.
 Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.
@@
 The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts)<!-- FP: during our work on the paper, I have heard two different stories of who the non experts are; either students/interns or full doctors that where not radiologists.
-Although this is only a data simulation example, it is nonetheless important here because the assumed effect sizes will probably differ!
-So who did you imagine when choosing the effect sizes?; SG can you clarify here please?
-SG: Sure, when planning the study, our non-expert group was going to be students and interns.
-In a previous study, the non-expert groups were physicians with little task expertise.
-In the end, we did base some assumptions on these previous studies since we did not have anything else to work with -->, review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
-To support their decision-making, in two experimental conditions <!--SG: I still think this will be confusing to readers.
-Why two experimental conditions?-->, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
+Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
+To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
 This AI advice can be either correct (80% of cases) or incorrect (20%).
@@
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans, (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice, and (c) different presentations of AI advice influence performance.
-<!--SG: c) has to be removed, if it is not in the analysis, right?
--->
+With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
+
 In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
@@
 <!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015] which is a state-of-the-art tool for fitting frequentist GLMMs<!--AK: I think you need a ref for this statement.-->.[^1]
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015] which is a state-of-the-art tool for fitting frequentist GLMMs<!--AK: I think you need a ref for this statement.
+TK: not sure if we can find a ref for this, but i think it's fine-->.[^1]
 
@@
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-$$ If the dependent variable (i.e., diagnostic decision)<!-- SG: not sure if diagnostic accuracy as a term is useful as a binary dependent variable since it is usually calculated as a percentage over many cases.)  --> $Y$ is a binary variable with values $0$ (i.e.
+$$ If the dependent variable (i.e., diagnostic decision)<!-- SG: not sure if diagnostic accuracy as a term is useful as a binary dependent variable since it is usually calculated as a percentage over many cases.).
+inserted "diagnostic decision".
+TK: i like it  --> $Y$ is a binary variable with values $0$ (i.e.
 inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
@@
 
-Often no meta-analytic results or conclusive data from pilot studies are available.
+Ideally, one would rely here on meta-analytic results or conclusive data from pilot studies.
+However, these are sometimes not readily available.
 All parameter values in our present case study have been determined based on results from prior own work and from the literature.
@@
 Table X shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model:
-<!-- SG Table number, title and note are missing. -->
+<!-- SG Table number, title and note are missing. TK: FP an you please insert this in final manuscript? -->
 | *Experimental condition*  | $P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$ | *Implied equation*                                                              |
@@
 This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-For example, we would expect even experts with low ability to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
+However, these high standard deviations do not account for the inherent variability and complexity of human performance.
+For example, we would expect that a participant with low ability compared to other task-experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
 <!-- SG: Maybe explain why this is not plausible.
@@
 SG: Better, but maybe still include a reason for this (years of training to do exactly this).
-Do you think it should be mentioned that already when just guessing, their probability would be .5-->
+Do you think it should be mentioned that already when just guessing, their probability would be .5
+TK: I have adjusted/ extended on this. how do you like it FP? -->
 
@@
 In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
-<!-- SG: Maybe explain what this means--> We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
+When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process.
+We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
 
@@
 Much has been written on the optimal amount of power to target in empirical research.
-The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
-<!-- Maybe give an example?--> When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
+The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints  [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
+When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
 For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation.
@@
 <!-- I would try to highlight this more clearly in the methods part.
+TK: is it clear enough now?
 --> Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
@@
 This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy<!-- FP: is accuracy the right term here?
-compromising realistic assumptions about the data generating process implied by the study design -->.
+By simplifying the simulation process, researchers can obtain reliable power estimates without compromising realistic assumptions about the data generating process implied by the study design.
 
@@
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
-The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports <!-- FP: cite the Chambers RR  paper on what a preregistered report is?
--->.
+The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
 

### 2023-09-12 docs/manuscript.Rmd
changed some terminology

@@
 <!-- what are the existing options for a priori power analysis for GLMMs / where are existing tutorials limited -->
-Currently available software packages for power simulations <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.
+Currently available software packages for power analysis <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.
 --> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
-A number of tutorials have been published describing how to perform such power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
 However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->).
@@
 <!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
-Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power simulation with GLMMs.
+Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs.
 Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.

### 2023-09-12 docs/manuscript.Rmd
include table and figure captions; work on some comments

@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = FALSE)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = TRUE)
 ```
@@
 Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
-Less formally, power is described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.<!-- AK: Just a very minor thing: I learned that power is the probability that a statistical test correctly rejects the null hypothesis when the alternative hypothesis is true.
-That is basically the same as what you wrote, but maybe a bit clearer because you don't have H1 and H1 right after another.
-But up to you!
-TK: FP what do you like better?
-Please insert the one you think fits better.
---> If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, this particular study would be considered "underpowered".
+Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
+If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, this particular study would be considered "underpowered".
 
@@
 Conducting underpowered research has many negative consequences.
-First, relying on underpowered experiments may yield inconclusive (if low power is taken into account<!--AK: Unclear what this means in this context-  how is it "taken into account"?, TK: FP can you clarify what is meant here?
--->) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
+First, relying on underpowered experiments may yield inconclusive (if researchers acknowledge the small evidential value of an underpowered study in the limitations section) <!--AK: Unclear what this means in this context-  how is it "taken into account"?, TK: FP can you clarify what is meant here?
+FP: better?
+-->) or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
 Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
@@
 Yet, these simple statistical methods are often not sufficient for research questions requiring more complex designs and methods.
-However, software-based power analysis tools cannot be used or adjusted for those.
-<!-- TK: FP do you like the previous section how it is now? Please adjust, if required -->
+However, software-based power analysis tools are often not flexible enough to perform power analysis for complex designs.
+<!-- TK: FP do you like the previous section how it is now? Please adjust, if required? FP: I made some changes -->
 
@@
 Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions <!--AK: Again "assumptions".--> a practical challenge [@maxwellSampleSizePlanning2008].
+However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions <!--AK: Again "assumptions".--> a practical challenge [@maxwellSampleSizePlanning2008].
 
@@
 Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs.
-Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.
+Thereby, we aim to equip researchers with the tools needed to simulate data and determine appropriate sample sizes for their own research.
 
@@
 <!-- describe the study idea and design -->
-In the present case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
+In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
 The goal is to understand how AI advice influences medical decision-making.
@@
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-$$ If the dependent variable (i.e., diagnostic decision)<!-- SG: not sure if diagnostic accuracy as a term is useful as a binary dependent variable since it is usually calculated as a percentage over many cases.).
-inserted "diagnostic decision".
-TK: i like it  --> $Y$ is a binary variable with values $0$ (i.e.
+$$ If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e.
 inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
@@
 
-In our case study, the probability of giving a correct response is assumed to depend on the predictors:
+In our case study, the probability of making an accurate diagnostic decision is assumed to depend on the predictors:
 
@@
 As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user.
-With all design variables done, we are ready to simulate our model equation as outlined in equation X.
-The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects and model parameters as assumed by our model.
+With all design variables done, we are ready to simulate our model equation outlined in the last section.
+The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects, and model parameters as assumed by our model.
 We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
-In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject chooses the accurate diagnosis for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
+In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject makes an accurate diagnostic decision for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
 
@@
 
-In this section, we show how to fit a GLMM with lme4, interpret the model and test hypotheses derived from a research question.
+In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
 
@@
 
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modelled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
 The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
@@
 
-With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition:
+With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects):
 
-```{r, message=FALSE}
+(ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package.
+
+```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"}
 library(marginaleffects)
@@
 
-The `Fixed effects` table in Figure X also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$.
+The `Fixed effects` table in the lme4 summary output also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$.
 However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.
-<!-- Figure is not included in the pdf -->
 
@@
 
-Table X shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model:
-<!-- SG Table number, title and note are missing. TK: FP an you please insert this in final manuscript? -->
-| *Experimental condition*  | $P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$ | *Implied equation*                                                              |
-|--------------------|:------------------------------:|--------------------|
-| no advice, no expert      |                      0.70                      | $logit(0.70) = \beta_0$                                                         |
-| no advice, expert         |                      0.90                      | $logit(0.90) = \beta_0 + \beta_e$                                               |
-| false advice, no expert   |                      0.40                      | $logit(0.40) = \beta_0 + \beta_a$                                               |
-| false advice, expert      |                      0.85                      | $logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}$                      |
-| correct advice, no expert |                      0.90                      | $logit(0.90) = \beta_0 + \beta_a + \beta_c$                                     |
-| correct advice, expert    |                      0.95                      | $logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}$ |
+```{r probtable, echo=FALSE}
+probtable <- data.frame(
+  col1 = 
+    c("no advice, no expert", "no advice, expert", "false advice, no expert", 
+      "false advice, expert", "correct advice, no expert", "correct advice, expert"),
+  col2 =
+    c("0.70", "0.90", "0.40", "0.85", "0.90", "0.95"),
+  col3 = 
+    c("$logit(0.70) = \\beta_0$",
+      "$logit(0.90) = \\beta_0 + \\beta_e$", 
+      "$logit(0.40) = \\beta_0 + \\beta_a$",
+      "$logit(0.85) = \\beta_0 + \\beta_e + \\beta_{a} + \\beta_{ea}$",
+      "$logit(0.90) = \\beta_0 + \\beta_a + \\beta_c$",
+      "$logit(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
+apa_table(
+  probtable
+  , col.names = c("Experimental condition",
+    "$P(Y=1|\\mathbf{X=x}, \\mathbf{U} = \\mathbf{0})$",
+    "Implied equation")
+  , caption = "Assumed probabilities that an average subject solves and average item in each experimental condition."
+  , note = "Implied equations are derived based on the model equations and setting all random intercept terms to 0."
+  , escape = FALSE
+)
+```
 
-This table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:
+Table \@ref(tab:probtable) shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
+The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:
 
@@
 
-```{r}
+(ref:margdist1-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions.
+
+```{r margdist1, fig.cap="(ref:margdist1-caption)"}
 library(tidyverse)
@@
 
-Figure X shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
+Figure \@ref(fig:margdist1) shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
 We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
 This difference increases with the variability of the random effects.
-<!-- SG: Figure is not included in the PDF-->
 
@@
 
-```{r, echo=FALSE}
+(ref:margdist2-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01.
+
+```{r margdist2, echo=FALSE, fig.cap="(ref:margdist2-caption)"}
 dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0i = 0.01)
@@
 
-As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
+As an example, Figure \@ref(fig:margdist2) reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
 In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
@@
 
-We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$, see Figure X in Appendix X).
+We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
 
@@
 The final plot demonstrates <!-- SG: no figures in the pdf --> that these plots are also useful for spotting standard deviations that are specified too high.
-For this example, we have set $\sigma_S = 3$ and $\sigma_I = 3$.
+For Figure \@ref(fig:margdist3), we have set $\sigma_S = 3$ and $\sigma_I = 3$.
 This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
@@
 
-```{r, echo=FALSE}
+(ref:margdist3-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3.
+
+```{r margdist3, echo=FALSE, fig.cap="(ref:margdist3-caption)"}
 dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 3, sd_u0i = 3)
@@
 We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
-With the current settings, this simulation takes about X minutes on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
+With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
 If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.
@@
 
-```{r, eval=FALSE}
+(ref:finalpwr-caption) Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05.
+
+```{r finalpwr, eval=FALSE}
 library(binom)
@@
 
-As should be the case, power estimates increase with both the number of subjects and the number of items.
+As should be the case, power estimates in Figure \@ref(fig:finalpwr) increase with both the number of subjects and the number of items.
 The confidence intervals indicate how precisely power was estimated by our simulation.

### 2023-09-12 docs/manuscript.Rmd
adjust manuscript.Rmd so that it loads a results.rds file (to avoid long evaluations when preparing the final manuscript)

@@
 
-The final plot demonstrates <!-- SG: no figures in the pdf --> that these plots are also useful for spotting standard deviations that are specified too high.
+The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
 For Figure \@ref(fig:margdist3), we have set $\sigma_S = 3$ and $\sigma_I = 3$.
@@
 
+```{r, message=FALSE, echo=FALSE, eval=FALSE}
+library(future)
+library(furrr)
+plan("multisession", workers = 6)
+set.seed(2)
+sim_design <- crossing(
+  rep = 1:300,
+  n_subjects = c(100, 150, 200, 250),
+  n_items = c(10, 30, 50, 70)
+) %>%
+  mutate(pvalues = future_pmap(., sim_and_analyse, 
+    .options = furrr_options(seed = TRUE))) %>%
+  unnest_wider(pvalues)
+saveRDS(sim_design, file = "results.rds")
+```
+
 ```{r, message=FALSE, eval=FALSE}
@@
 
-```{r finalpwr, eval=FALSE}
+```{r, echo=FALSE}
+sim_design <- readRDS(file = "results.rds")
+```
+
+```{r finalpwr, fig.cap="(ref:finalpwr-caption)"}
 library(binom)

### 2023-09-13 docs/manuscript.Rmd
work on comments

@@
 Do study designs become more complex or do we become more aware of the pitfalls of underpowered studies?
-I would say it's the latter.-->, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+I would say it's the latter.
+-->, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
 GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data in HCI and other empirical disciplines because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
@@
 Currently available software packages for power analysis <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.
---> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
+--> often do not include GLMMs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
 A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->).
-This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->) [@brownIntroductionLinearMixedEffects2021].
+However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
+This narrow focus provides limited guidance for researchers faced with more complex study designs especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
+<!-- FP: es war nicht leicht das hier zu lösen: ich möchte ungern debruineUnderstandingMixedEffectsModels2021 als negativ beispiel hier nennen, das hat zwar nur lmms im hauptteil aber glmms im appendix und ist einfach ein super paper.
+außerdem hatte ich beim nochmal lesen den eindruck, dass wir kumleEstimatingPowerGeneralized2021 in unserem ganzen paper nicht genug credit geben.
+daher hab ich es jetzt so umgestellt dass wir nicht die negativbeispiele sondern das positivbeispiel aufzählen -->
 The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
@@
 Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions <!--AK: Again "assumptions".--> a practical challenge [@maxwellSampleSizePlanning2008].
+However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions <!--AK: Again "assumptions".--> a practical challenge [see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021].
 

### 2023-09-13 docs/manuscript.Rmd
remove comment that broke equation

@@
 $$
-g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta <!-- SG: Sorry for maybe a stupid question, but what does E stand for?
-and why are model parameters beta mentioned but not in the equation?-->
+g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
 $$ Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$

### 2023-09-13 docs/manuscript.Rmd
made final changes before sending out pdf to co-authors

@@
 
-  [THIS WILL BE UPDATED BASED ON GOOGLE DOC] Researchers often lack the tools to conduct a priori power analysis for complex experimental designs deployed in human-computer interaction (HCI) studies.
-Generalized Linear Mixed Models (GLMMs) are well-suited to analyze experiments in which subjects respond to a set of stimuli.
-However, existing software packages can only be used to simulate power for simple designs, which are often not useful for more sophisticated investigations.
-This tutorial addresses this gap by providing a comprehensive tutorial to perform tailored power analyses using data simulation.
-Our tutorial provides code for an HCI case study, empowering researchers to simulate their own data and ascertain appropriate sample sizes for GLMMs.
-This resource should help to enhance the precision of study designs and to mitigate the risk of underpowered studies.
-Moreover, we discuss further implications for experimental research and provide an outlook on the increasing relevance of power simulations in the field.
+  When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
+However, the recent replication crisis in Human-Computer Interaction (HCI) and other empirical research fields underlines the need for more rigorous statistical methodology and well-powered designs.
+Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
+Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs necessitating data simulation-based approaches.
+Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs.
+Finally, we give an outlook on the increasing importance of simulation-based power analysis in experimental research.
 
@@
 
-wordcount         : "7037"
+wordcount         : "7461"
 
@@
 When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data [@lakensSampleSizeJustification2022].
-However, the recent replication crisis in Human-Computer Interaction (HCI) and several other disciplines grounded on empirical research has illustrated many challenges surrounding the reproducibility and reliability of findings [@cockburnThreatsReplicationCrisis2020; @robertsonImprovingStatisticalPractice2016].
-As a result, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
+However, the recent replication crisis in Human-Computer Interaction (HCI) and several other disciplines grounded on empirical research has illustrated many challenges surrounding the reproducibility and reliability of findings [@cockburnThreatsReplicationCrisis2020; @robertsonImprovingStatisticalPractice2016; @yarkoniGeneralizabilityCrisis2022].
+As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
 While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
@@
 <!-- standard software is available for "simple" models and study designs -->
-For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori power analysis is readily available [@faulStatisticalPowerAnalyses2009].
+For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori power analysis is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018].
 Yet, these simple statistical methods are often not sufficient for research questions requiring more complex designs and methods.
@@
 Finally, the outcome variable is binary (click/no click) for each interaction, which follows a binomial distribution.
-
-<!--AK: And here I would add one research example.
-Something like, "Imagine an experiment where experts are repeatedly asked for...
-In this case, the data is not independent because ...
-I would add an example that explains both aspects you talk about before: 1) The dependency, 2) the outcome distribution (e.g., binomial).
-I think this would make it much easier for the reader to follow you here.
-TK: FP how do you like the new example? Feel free to adjust -->
+<!-- TK: I've added an example here -->
 
@@
 außerdem hatte ich beim nochmal lesen den eindruck, dass wir kumleEstimatingPowerGeneralized2021 in unserem ganzen paper nicht genug credit geben.
-daher hab ich es jetzt so umgestellt dass wir nicht die negativbeispiele sondern das positivbeispiel aufzählen -->
-The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
+daher hab ich es jetzt so umgestellt dass wir nicht die negativbeispiele sondern das positivbeispiel aufzählen.
+TK: stimme dir voll und ganz zu!
+-->
+The necessary assumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
 The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial.
-Assumptions about the random effects include the assumption of normality<!--AK: Normality of what?--> and the covariance structure among the random effects<!--AK: Unclear what this is-->.
+Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e.,  how they are correlated).
 Additionally, assumptions about the error structure, such as independence or correlation, must be specified.
@@
 With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
-
 In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
-The goal of the power simulation is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
+The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
 
@@
 
-The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
 
@@
 Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
-In practice, power simulations are often run multiple times with adjusted combinations of sample sizes.
+In practice, data simulations are often run multiple times with adjusted combinations of sample sizes.
 When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of `n_subjects` and `n_items`.
-When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of power simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
+When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
 This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.
@@
 <!-- short summary of key takeaways from this tutorial -->
-
 Experimental research requires careful planning and consideration of statistical power to ensure robust and meaningful results.
-While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
-The present tutorial presents a specific example of how to run power simulations for experimental designs with GLMMs.
+While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations with GLMMs are required.
+The present tutorial presents a specific case study with corresponding code of how to conduct a simulation-based power analysis for experimental designs with GLMMs.
 
@@
 
-In our case study, we have performed power simulations based on a single set of parameter values that reflect our assumptions of an expected effect size.
+In our case study, we have performed simulation-based power analysis from a single set of parameter values that reflect our assumptions of an expected effect size.
 Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
@@
 --> Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
-When sample sizes are chosen based on the results of our power simulations, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.
+When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.
 However, if the true effect is indeed smaller, the power will be lower, and the study might not be sufficiently informative.
@@
 <!-- Outlook -->
-
-Beyond the specifics of our concrete case study, we want to outline six avenues regarding the future of power simulations in experimental research:
+Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based power analysis in experimental research:
 
 1.
-The need for power simulations in experimental research: Experimental research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
-Power simulations offer a solution by providing experimental researchers with a tailored approach to estimating statistical power.
-These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
+The growing need for simulation-based power analyses in experimental research: In order to conduct well-powered research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
+Therefore, simulation-based power analysis are becoming increasingly needed since they provide experimental researchers with a tailored approach to estimating required sample sizes before data collection.
 
 2.
-Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
+Managing data simulations more easily with discrete predictor variables: Simulation-based power analysis becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
 This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
@@
 3.
-Teaching power simulation skills: The ability to conduct power simulations is a valuable skill that should be taught to experimental researchers.
-By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping researchers with the knowledge and tools to perform power simulations enables them to make informed decisions and enhance the rigor of their studies. 
+Teaching data simulation skills: The ability to conduct simulation-based power analysis is a valuable skill that should be taught to experimental researchers.
+By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
+Equipping researchers with the knowledge and tools to perform simulation-based power analyes enables them to make informed decisions and enhance the rigor of their studies. 
 The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look like, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
@@
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and the perceived effort estimated by researchers and collaborators in experimental research.
-Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-It is crucial to raise awareness about the effort involved in power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform simulation-based a priori power analysis and the perceived effort estimated by researchers and collaborators in experimental research.
+Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
+It is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.
 
 5.
-Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
+Recognizing the value of simulation-based power analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
-The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
+Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
 
 6.
-Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in experimental research.
+Integration with Open Science and preregistration practices: Simulation-based powers analysis aligns well with the principles of Open Science and preregistration in experimental research.
 When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their power simulations, researchers enhance transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
+By preregistering their simulation-based power analysis, researchers enhance transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
 
@@
 
-Power simulations play a critical role in experimental research, allowing investigators to tailor power estimation to the unique aspects of their experiments.
-Through this tutorial, we aim to provide researchers with the necessary skills and tools to perform these simulations for GLMMs themselves.
-By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work [@yarkoniGeneralizabilityCrisis2022].
+In the wake of the replication crisis and myriad of under-powered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
+Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs necessitating data simulation-based approaches.
+Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based power analysis with GLMMs themselves.
+By incorporating GLMMs and  a priori power analysis into their work, researchers can enhance the replicability and credibility of their experiments [@yarkoniGeneralizabilityCrisis2022].
 

### 2023-09-14 docs/manuscript.Rmd
added final feedback from google doc

@@
     corresponding : yes    # Define only one corresponding author
-    #address       : "Postal address"
+    address       : "Torstrasse 25, 9000 St. Gallen, Switzerland"
     email         : "timo.koch@unisg.ch"
@@
 
-#authornote: |
-  # author note here
+authornote: |
+  This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525).
 
@@
 Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs necessitating data simulation-based approaches.
+Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs, necessitating data simulation approaches.
 Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs.
@@
 output: 
-  papaja::apa6_pdf:
+  papaja::apa6_docx:
     keep_tex: TRUE
@@
 While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
-In the present work, we provide a tutorial comprised of a concrete example for tailored a priori power analyses using data simulations based on GLMMs.
+In the present work, we provide a tutorial consisting of a concrete example for tailored a priori power analyses using data simulations based on GLMMs.
 
@@
 Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, this particular study would be considered "underpowered".
+If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered".
 
@@
 FP: better?
--->) or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
+--> or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
 Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
@@
 A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
-When running a power analysis before data collection (i.e., "a priori"), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
+When running a power analysis before data collection, the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
 
@@
 Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
-Here, adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
+Adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
 Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-Therefore, proficiency with power analysis can aid one's academic career chances.
 
@@
 For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori power analysis is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018].
-Yet, these simple statistical methods are often not sufficient for research questions requiring more complex designs and methods.
-However, software-based power analysis tools are often not flexible enough to perform power analysis for complex designs.
+However, these software packages are often not flexible enough to perform power analysis for complex designs.
 <!-- TK: FP do you like the previous section how it is now? Please adjust, if required? FP: I made some changes -->
@@
 I would say it's the latter.
--->, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data in HCI and other empirical disciplines because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
+-->, researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data in HCI and other empirical disciplines because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
 
 <!-- intro into GLMMs -->
-GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of classic linear regression models that account for correlated data and hierarchical structures  [@fahrmeirRegressionModelsMethods2021].
-In this context, correlated data means that observations within a given dataset are not independent of each other.
-In other words, the values of the dependent variable for one observation may be related to the values of the dependent variable for other observations in a systematic way.
-This correlation can arise for various reasons, such as repeated measurements on the same subjects over time, observations clustered within certain groups or locations, or data collected at different levels of granularity.
-Hierarchical structures can be found in datasets that have a nested or hierarchical organization.
-For example, one might have data on students nested within schools, employees nested within departments, or patients nested within hospitals.
-In such cases, the observations within each higher-level grouping (e.g., school, department, hospital) tend to be more similar to each other than to observations in other groups.
-LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables).
-They allow for the modelling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
-As mentioned, GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family (e.g., binomial) <!-- SG: do you think examples for the different types of variables are needed; TK: i don't think so -->.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.
-As an example, think of an experiment on different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
-Here, participants' interaction is measured repeatedly (e.g., over several sessions).
-Also, the interactions of participants in one session are likely to be correlated with their previous sessions.
+GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data including hierarchical structures  [@fahrmeirRegressionModelsMethods2021].
+In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value for another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
+This correlation can arise for various reasons: Responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), both participants come from the same neighborhood (clustering) or both participants responded to the same stimulus (stimulus effects).
+Thus, modeling such correlations is especially important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
+LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on all fixed and random effects).
+They allow for the modeling of fixed effects, which capture the relationships between our usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects exemplified above.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
+
+As mentioned, GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
+As an example, think of an experiment with different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
+Here, participants’ behavior is measured repeatedly (e.g., over several sessions).
+The click patterns of participants in one session are likely to be correlated with their previous sessions.
 Finally, the outcome variable is binary (click/no click) for each interaction, which follows a binomial distribution.
-<!-- TK: I've added an example here -->
 
@@
 <!-- what are the existing options for a priori power analysis for GLMMs / where are existing tutorials limited -->
-Currently available software packages for power analysis <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.
---> often do not include GLMMs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
+Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
+Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e.
+how often the hypothesis test is significant for the simulated data).
+Currently available formula-based software packages for power analysis often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
 A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
 However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
-This narrow focus provides limited guidance for researchers faced with more complex study designs especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
+This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
 <!-- FP: es war nicht leicht das hier zu lösen: ich möchte ungern debruineUnderstandingMixedEffectsModels2021 als negativ beispiel hier nennen, das hat zwar nur lmms im hauptteil aber glmms im appendix und ist einfach ein super paper.
@@
 -->
-The necessary assumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
-The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial.
-Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e.,  how they are correlated).
-Additionally, assumptions about the error structure, such as independence or correlation, must be specified.
-Interpreting these assumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
+The necessary presumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
+The distributional assumption specifies the distributional family for the outcome variable (when conditioned on all fixed and random effects).
+Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e.,  if and how they are correlated).
+Interpreting these presumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
 Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions <!--AK: Again "assumptions".--> a practical challenge [see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021].
+However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge [see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021].
 
@@
 <!-- describe the study idea and design -->
-In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated. 
+In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
 The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
+Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding.
 To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
@@
 <!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015] which is a state-of-the-art tool for fitting frequentist GLMMs<!--AK: I think you need a ref for this statement.
-TK: not sure if we can find a ref for this, but i think it's fine-->.[^1]
-
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
 [^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
-
 The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
@@
 g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$ Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
+$$ Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
@@
 
-To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions here.
+To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.
 
@@
 We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
-However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyse this data.
+However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.
 
@@
 
-Ideally, one would rely here on meta-analytic results or conclusive data from pilot studies.
+Ideally, one would rely on meta-analytic results or conclusive data from pilot studies.
 However, these are sometimes not readily available.
-All parameter values in our present case study have been determined based on results from prior own work and from the literature.
-Additionally, we had repeated exchanges with our affiliated domain experts from radiology.
+All parameter values in our present case study have been determined based on results from related prior work.
+Additionally, we had repeated discussions with our affiliated domain experts in radiology to check our assumptions.
 
@@
     "Implied equation")
-  , caption = "Assumed probabilities that an average subject solves and average item in each experimental condition."
+  , caption = "Assumed probabilities that an average subject solves an average item in each experimental condition."
   , note = "Implied equations are derived based on the model equations and setting all random intercept terms to 0."
@@
 
-Earlier, we have already encountered one way to visualize <!-- SG: so far, no figures in the pdf --> the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
+Earlier, we have already encountered one way to visualize the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
 However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory.
@@
 However, these high standard deviations do not account for the inherent variability and complexity of human performance.
-For example, we would expect that a participant with low ability compared to other task-experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
+For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
 <!-- SG: Maybe explain why this is not plausible.
@@
 
-The result of the computation is a data.frame that contains the p-values of all tested hypotheses for each simulated dataset.
+The result of the computation is a data frame that contains the p-values of all tested hypotheses for each simulated dataset.
 In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
@@
 By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power depends on the size of the true effect.
-In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyse power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
-However, for our case study that investigates combined hypotheses in a GLMM modelling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
+In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyze power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
+However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
 In this setting, sensitivity power analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
@@
 The growing need for simulation-based power analyses in experimental research: In order to conduct well-powered research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
-Therefore, simulation-based power analysis are becoming increasingly needed since they provide experimental researchers with a tailored approach to estimating required sample sizes before data collection.
+Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
 
@@
 This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising realistic assumptions about the data generating process implied by the study design.
+By simplifying the simulation process, researchers can obtain reliable power estimates without compromising realistic assumptions about the data-generating process implied by the study design.
 
@@
 By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping researchers with the knowledge and tools to perform simulation-based power analyes enables them to make informed decisions and enhance the rigor of their studies. 
-The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look like, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
+Equipping researchers with the knowledge and tools to perform simulation-based power analyses enables them to make informed decisions and enhance the rigor of their studies. 
+The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
 
@@
 When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their simulation-based power analysis, researchers enhance transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
+By preregistering their simulation-based power analysis, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
 
@@
 
-In the wake of the replication crisis and myriad of under-powered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
+In the wake of the replication crisis and myriad of underpowered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
 Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs necessitating data simulation-based approaches.

### 2023-09-14 docs/manuscript.Rmd
last mini-edits

@@
 
-wordcount         : "7461"
+wordcount         : "7892"
 
@@
 Conducting underpowered research has many negative consequences.
-First, relying on underpowered experiments may yield inconclusive (if researchers acknowledge the small evidential value of an underpowered study in the limitations section) <!--AK: Unclear what this means in this context-  how is it "taken into account"?, TK: FP can you clarify what is meant here?
-FP: better?
---> or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
+First, relying on underpowered experiments may yield inconclusive (if researchers acknowledge the small evidential value of an underpowered study in the limitations section) or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
 Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
@@
 Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e.
-how often the hypothesis test is significant for the simulated data).
+Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
 Currently available formula-based software packages for power analysis often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
@@
 In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
-[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
 The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
 
+[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
+
 However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.[^2]
+Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]
 
@@
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-$$ If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e.
-inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
+$$ If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
 P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
@@
 
-- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to asses item $i$ 
+- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to assess item $i$ 
 - $advice\_correct_{si}$: whether this advice was correct (1) or not (0) 
@@
 The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously [@hothornSimultaneousInferenceGeneral2008].
-However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected [i.e.
-intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
+However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
 In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
@@
 
-(ref:finalpwr-caption) Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05.
-
 ```{r, echo=FALSE}
@@
 
-```{r finalpwr, fig.cap="(ref:finalpwr-caption)"}
+<!-- (ref:finalpwr-caption) Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05.
+TK: Caption did not work, included it manually for quick fix -->
+
+```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05"}
 library(binom)

### 2023-09-14 r_code/archive/samplesizesimulation.R
cleanup repository

@@ -1,270 +1,0 @@
-#### PREPARATION ####
-
-# inspiration for this code came from..
-
-# online ideas: https://datascienceplus.com/how-to-create-a-loop-to-run-multiple-regression-models/
-# this looks really (!!) good: https://rstudio-pubs-static.s3.amazonaws.com/11703_21d1c073558845e1b56ec921c6e0931e.html
-# source for theta parameters: https://stats.stackexchange.com/questions/155424/how-does-lme4-optimize-the-variance-parameters-of-random-effects-theta-vector
-
-# Load packages
-
-library(lme4) 
-library(ggplot2)
-set.seed(101)
-
-#### INSERT SAMPLE SIZE ASSUMPTIONS ####
-
-nSubjects <- 250 # number of participants (experts to non experts with 1:5 ratio)
-nExp <- 3 # number of experiments
-nObs <- 50 # number of observations per experiment
-
-N = nSubjects*nExp*nObs # number of total observations
-
-#### SET UM EXPERIMENT DATA FRAME ####
-
-# Set up data frame
-
-# generate experimental variables 
-# interpretability is only relevant when AI advice is given!
-
-## Define parameter/experimental design values:
-
-subject <- factor(rep(seq(nSubjects)))
-experiment <- factor(rep(seq(nExp)))
-observation <- factor(rep(seq(nObs)))
-
-# set up experimental conditions (see https://docs.google.com/document/d/1-3MCjl_JsVmqOevUTN-LPg3bJJk1YDGn/edit)
-
-exp_conditions <- data.frame(group = rep(c(1,2,3,4,5,6), times = 1, each = 3),
-                             experiment = rep(c(1,2,3), times = 6, each = 1),
-                             advice = rep(c(1,1,0,1,0,1,0,1,1), times = 2, each =1),
-                             interpretable = c(1,0,0,1,0,0,0,1,0,
-                                               0,1,0,0,0,1,0,0,1))
-
-# randomly assign subjects to experimental conditions
-
-subject_cond <- data.frame(subject = c(1:250), expert = sample(x = 0:1, size = 250, replace = TRUE, prob = c(0.8, 0.2)), group = sample(1:6, size = 250, replace = TRUE))
-
-# we aim to recruit 250 participant with a 5:1 ratio of non experts to experts
-
-# create df for experiment data
-
-experiment_data <- expand.grid(observation, experiment, subject)
-
[…]

### 2023-09-14 docs/manuscript.Rmd
made some final edits i figured out while preparing the chi submission in word

@@
 
+csl               : "acm-sigchi-proceedings.csl" #added this for chi submission
+
 classoption       : "man"
@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = TRUE)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = TRUE, fig.keep = "high", dev = "jpeg")
 ```
@@
 In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
+The corresponding code in this manuscript will be made available in a public repository upon publication.
 
@@
 However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]
+Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]
 
@@
 <!-- deep dive into our specific example -->
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021].
-$$
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]: $$
 g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$ Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
+$$ Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: $$
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
@@
 In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
-
 To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.
@@
 In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
-
 We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
@@
 
-We can specify and test hypotheses like these with the multcomp package [@R-multcomp] as follows:
+We can specify and test hypotheses like these with the multcomp package [@hothornSimultaneousInferenceGeneral2008] as follows:
 
@@
 
-```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05"}
+```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."}
 library(binom)
@@
                                 power, ci.lwr, ci.upr)), 
-    color = "white", size = 4.5) +
+    color = "white", size = 4) +
   scale_fill_viridis_c(limits = c(0, 1)) +

### 2023-09-15 docs/appendix.Rmd

@@
+# Appendix:
+
+**Note**: At the moment, this file contains only a few fragments of code but is not mentioned in the manuscript.
+Some of these pieces might be used in a later stage of the probject (e.g.
+after the first round of peer review).
+
+## A different model equation with random slopes
+
+$$
+logit(P_{si}) = (\beta_0 + u_{0p} + u_{0i})  + \beta_e \cdot expert_p + \beta_a \cdot advice\_present_{si} + (\beta_c + u_{1p}) \cdot advice\_correct_{si} + \beta_{ea} \cdot expert_{p} \cdot advice\_present_{si}
+$$
+
 ## General principles in power analysis

### 2023-09-18 docs/manuscript.Rmd
updated repo for preprint publication

@@
 
+note: "\\clearpage" #added this to move entire author note to second page
 authornote: |
   This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525).
+  All R code in this manuscript is available in the project's repository on the Open Science Framework (OSF): https://osf.io/dhwf4/
 
@@
 
-wordcount         : "7892"
+wordcount         : "7598"
 
@@
 
-csl               : "acm-sigchi-proceedings.csl" #added this for chi submission
+#csl               : "acm-sigchi-proceedings.csl" #added this for chi submission, not needed for preprint
 
@@
 output: 
-  papaja::apa6_docx:
-    keep_tex: TRUE
+  papaja::apa6_pdf:
 ---
@@
 # Introduction
-
-<!-- SG: In the intro power analysis, a priori power analysis and sample size calculation is now used interchangeably, should we use one term consistently?-->
-<!-- TK: I have repharsed this - should be clearer now -->
 
@@
 However, these software packages are often not flexible enough to perform power analysis for complex designs.
-<!-- TK: FP do you like the previous section how it is now? Please adjust, if required? FP: I made some changes -->
 
@@
 <!-- Advantages of GLMMs over "simpler" models -->
-As study designs become more complex<!--AK: I am not even sure whether this is true.
-Do study designs become more complex or do we become more aware of the pitfalls of underpowered studies?
-I would say it's the latter.
--->, researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
 GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data in HCI and other empirical disciplines because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
@@
 This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
-<!-- FP: es war nicht leicht das hier zu lösen: ich möchte ungern debruineUnderstandingMixedEffectsModels2021 als negativ beispiel hier nennen, das hat zwar nur lmms im hauptteil aber glmms im appendix und ist einfach ein super paper.
-außerdem hatte ich beim nochmal lesen den eindruck, dass wir kumleEstimatingPowerGeneralized2021 in unserem ganzen paper nicht genug credit geben.
-daher hab ich es jetzt so umgestellt dass wir nicht die negativbeispiele sondern das positivbeispiel aufzählen.
-TK: stimme dir voll und ganz zu!
--->
 The necessary presumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
@@
 In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
-The corresponding code in this manuscript will be made available in a public repository upon publication.
+The corresponding code in this manuscript is available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
 
@@
 ## Our specific GLMM
+
 <!-- deep dive into our specific example -->
@@
 One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
-<!-- SG: What does that mean for someone who wants to do this then?
-FP: is it more clear now?
-SG: Definitely better; the question might still be what your solution to this issue is.
---> However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
+However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
 Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
@@
 For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
-<!-- SG: Maybe explain why this is not plausible.
-FP: like this?
-SG: Better, but maybe still include a reason for this (years of training to do exactly this).
-Do you think it should be mentioned that already when just guessing, their probability would be .5
-TK: I have adjusted/ extended on this. how do you like it FP? -->
 
@@
 
-<!-- (ref:finalpwr-caption) Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05.
-TK: Caption did not work, included it manually for quick fix -->
-
 ```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."}
@@
 Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
-<!-- I would try to highlight this more clearly in the methods part.
-TK: is it clear enough now?
---> Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
+Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
 When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.

### 2023-09-18 docs/manuscript.tex
updated repo for preprint publication

@@
 {}{\PatchFailed}
-\keywords{power analysis, data simulation, sample size, generalized linear mixed model\newline\indent Word count: 7059}
+\keywords{power analysis, data simulation, sample size, generalized linear mixed model\newline\indent Word count: 7598}
 \usepackage{lineno}
@@
 \hypersetup{
-  pdftitle={A Tutorial on Tailored Power Simulations for Experimental Study Designs with Generalized Linear Mixed Models},
+  pdftitle={A Tutorial on Tailored Simulation-Based Power Analysis for Experimental Designs with Generalized Linear Mixed Models},
   pdfauthor={Timo K. Koch,1, 2, Florian Pargent,1, Anne-Kathrin Kleine1, Eva Lermer1,3, \& Susanne Gaube1,4},
@@
 
-\title{A Tutorial on Tailored Power Simulations for Experimental Study Designs with Generalized Linear Mixed Models}
+\title{A Tutorial on Tailored Simulation-Based Power Analysis for Experimental Designs with Generalized Linear Mixed Models}
 \author{Timo K.
@@
 
-\shorttitle{GLMM Power Simulation Tutorial}
+\shorttitle{Simulation-Based Power Analysis for GLMMs}
 
 \authornote{
+
+This research was funded by a grant from the Volkswagen Foundation (Grant No.~98525).
+All R code in this manuscript is available in the project's repository on the Open Science Framework (OSF): \url{https://osf.io/dhwf4/}
 
@@
 
-Correspondence concerning this article should be addressed to Timo K. Koch. E-mail: \href{mailto:timo.koch@unisg.ch}{\nolinkurl{timo.koch@unisg.ch}}
+Correspondence concerning this article should be addressed to Timo K. Koch, Torstrasse 25, 9000 St.~Gallen, Switzerland. E-mail: \href{mailto:timo.koch@unisg.ch}{\nolinkurl{timo.koch@unisg.ch}}
 
@@
 
+\note{\clearpage}
+
 \abstract{%
-{[}THIS WILL BE UPDATED BASED ON GOOGLE DOC{]} Researchers often lack the tools to conduct a priori power analysis for complex experimental designs deployed in human-computer interaction (HCI) studies.
-Generalized Linear Mixed Models (GLMMs) are well-suited to analyze experiments in which subjects respond to a set of stimuli.
-However, existing software packages can only be used to simulate power for simple designs, which are often not useful for more sophisticated investigations.
-This tutorial addresses this gap by providing a comprehensive tutorial to perform tailored power analyses using data simulation.
-Our tutorial provides code for an HCI case study, empowering researchers to simulate their own data and ascertain appropriate sample sizes for GLMMs.
-This resource should help to enhance the precision of study designs and to mitigate the risk of underpowered studies.
-Moreover, we discuss further implications for experimental research and provide an outlook on the increasing relevance of power simulations in the field.
+When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
+However, the recent replication crisis in Human-Computer Interaction (HCI) and other empirical research fields underlines the need for more rigorous statistical methodology and well-powered designs.
+Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
+Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs, necessitating data simulation approaches.
+Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs.
+Finally, we give an outlook on the increasing importance of simulation-based power analysis in experimental research.
 }
@@
 
-When planning experimental research, it is essential to determine an appropriate sample size (Lakens, 2022b) to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data.
-However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), has illustrated many challenges surrounding the reproducibility and reliability of findings (Robertson \& Kaptein, 2016).
-As a result, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs.
-While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework.
-In the present work, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
+When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data (Lakens, 2022b).
+However, the recent replication crisis in Human-Computer Interaction (HCI) and several other disciplines grounded on empirical research has illustrated many challenges surrounding the reproducibility and reliability of findings (Cockburn, Dragicevic, Besançon, \& Gutwin, 2020; Robertson \& Kaptein, 2016; Yarkoni, 2022).
+As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
+While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct ``a priori'' (i.e., before data collection) power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
+In the present work, we provide a tutorial consisting of a concrete example for tailored a priori power analyses using data simulations based on GLMMs.
 
@@
 Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
-Less formally, power is described as the probability that a hypothesis test accepts the alternative hypothesis \(H_1\) if \(H_1\) is indeed true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated , this particular study would be considered ``underpowered''.
+Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
+If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered ``underpowered''.
 
 Conducting underpowered research has many negative consequences.
-First, relying on underpowered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge.
+First, relying on underpowered experiments may yield inconclusive (if researchers acknowledge the small evidential value of an underpowered study in the limitations section) or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
 Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
@@
 A power analysis represents the act of calculating the statistical power for a given true effect and sample size.
-By running the calculation for a range of sample sizes ``a priori'' (i.e., before data collection), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
-Thereby, power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
-
-Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
+When running a power analysis before data collection, the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
+
+Thereby, a priori power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
+Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
 Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
-Especially in fast-moving research fields, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.
-
-Moreover, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights.
-
-For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for power analysis is readily available (Faul, Erdfelder, Buchner, \& Lang, 2009).
-
-\hypertarget{generalized-linear-mixed-models-glmm}{%
-\subsection{Generalized linear mixed models (GLMM)}\label{generalized-linear-mixed-models-glmm}}
-
-As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs (Yarkoni, 2022).
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects (Fahrmeir, Kneib, Lang, \& Marx, 2021; Kaptein, 2016).
-
-GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of classic linear regression models that account for correlated data and hierarchical structures (Fahrmeir et al., 2021).
-In this context, correlated data means that observations within a given dataset are not independent of each other.
-In other words, the values of the dependent variable for one observation may be related to the values of the dependent variable for other observations in a systematic way.
-This correlation can arise for various reasons, such as repeated measurements on the same subjects over time, observations clustered within certain groups or locations, or data collected at different levels of granularity.
-Hierarchical structures can be found in datasets that have a nested or hierarchical organization.
-For example, one might have data on students nested within schools, employees nested within departments, or patients nested within hospitals.
-In such cases, the observations within each higher-level grouping (e.g., school, department, hospital) tend to be more similar to each other than to observations in other groups.
-LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables).
-They allow for the modelling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects.
-As mentioned, GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables.
-They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family (e.g., binomial) .
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable.
-The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.
-
-\hypertarget{power-simulations-for-glmms}{%
-\subsection{Power simulations for GLMMs}\label{power-simulations-for-glmms}}
-
-Currently available software packages for power simulations do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design (Murayama, Usami, \& Sakaki, 2022; Westfall, Kenny, \& Judd, 2014).
-A number of tutorials have been published describing how to perform such power simulation for multilevel models (Arend \& Schäfer, 2019; Brysbaert \& Stevens, 2018; DeBruine \& Barr, 2021; Kumle, Võ, \& Draschkow, 2021; Lafit et al., 2021; Zimmer, Henninger, \& Debelak, 2022).
-However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE).
-This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common (Brown, 2021).
-The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure.
-The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial.
-Assumptions about the random effects include the assumption of normality and the covariance structure among the random effects.
-Additionally, assumptions about the error structure, such as independence or correlation, must be specified.
-Interpreting these assumptions entails understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
+Adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
+Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+
+For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori power analysis is readily available (Champely et al., 2018; Faul, Erdfelder, Buchner, \& Lang, 2009).
+However, these software packages are often not flexible enough to perform power analysis for complex designs.
+
+\hypertarget{generalized-linear-mixed-models-glmms}{%
+\subsection{Generalized linear mixed models (GLMMs)}\label{generalized-linear-mixed-models-glmms}}
+
+As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs (Yarkoni, 2022).
+GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data in HCI and other empirical disciplines because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects (Fahrmeir, Kneib, Lang, \& Marx, 2021; Kaptein, 2016).
+
+GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data including hierarchical structures (Fahrmeir et al., 2021).
+In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value for another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
+This correlation can arise for various reasons: Responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), both participants come from the same neighborhood (clustering) or both participants responded to the same stimulus (stimulus effects).
+Thus, modeling such correlations is especially important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
+LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on all fixed and random effects).
+They allow for the modeling of fixed effects, which capture the relationships between our usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects exemplified above.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
+
+As mentioned, GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
+As an example, think of an experiment with different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
+Here, participants' behavior is measured repeatedly (e.g., over several sessions).
+The click patterns of participants in one session are likely to be correlated with their previous sessions.
+Finally, the outcome variable is binary (click/no click) for each interaction, which follows a binomial distribution.
+
+\hypertarget{power-analysis-for-glmms}{%
+\subsection{Power analysis for GLMMs}\label{power-analysis-for-glmms}}
+
+Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods (Murayama, Usami, \& Sakaki, 2022).
+Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
+Currently available formula-based software packages for power analysis often do not include GLMMs or are limited to very simple designs (Murayama et al., 2022; Westfall, Kenny, \& Judd, 2014), making it necessary to build data simulations tailored specifically to the study design.
+A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models (Arend \& Schäfer, 2019; Brysbaert \& Stevens, 2018; DeBruine \& Barr, 2021; Kumle, Võ, \& Draschkow, 2021; Lafit et al., 2021; Zimmer, Henninger, \& Debelak, 2022).
+However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs (but see Kumle et al., 2021 for a tutorial that also covers more advanced settings).
+This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available (see the discussion in Kumle et al., 2021).
+The necessary presumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
+The distributional assumption specifies the distributional family for the outcome variable (when conditioned on all fixed and random effects).
+Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e., if and how they are correlated).
+Interpreting these presumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
 Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions a practical challenge (Maxwell, Kelley, \& Rausch, 2008).
-
-Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design.
-Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.
+However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge Kumle et al.
+(2021).
+
+Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs.
+Thereby, we aim to equip researchers with the tools needed to simulate data and determine appropriate sample sizes for their own research.
 
@@
 In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
+The corresponding code in this manuscript is available in the project's repository on the Open Science Framework (\url{https://osf.io/dhwf4/}).
 
@@
 
-In the present case study, we simulate the data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system is to be evaluated.
+In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated.
 The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of bleeding, more specifically, intracranial haemorrhage (ICH).
-To support their decision-making, in two experimental conditions , an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
+Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding.
+To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
 This AI advice can be either correct (80\% of cases) or incorrect (20\%).
@@
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans, (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice, and (c) different presentations of AI advice influence performance.
+With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
 In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible.
-The goal of the power simulation is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
+The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
 
@@
 
-In our case study, we use the lme4 R package (Bates, Mächler, Bolker, \& Walker, 2015) which is a state-of-the-art tool for fitting frequentist GLMMs.\footnote{For Bayesian GLMMs, the brms R package is currently the most prominent option (Bürkner, 2017).}
-
-The lme4 package includes a function called \texttt{simulate} that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling power simulations and other related analyses.
+In our case study, we use the lme4 R package (Bates, Mächler, Bolker, \& Walker, 2015), which is a state-of-the-art tool for fitting frequentist GLMMs.\footnote{For Bayesian GLMMs, the brms R package is currently the most prominent option (Bürkner, 2017).}
+The lme4 package includes a function called \texttt{simulate} that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
 
 However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated data sets.\footnote{A less flexible alternative would be to use the simr package (Green \& MacLeod, 2016), which can be used to both simulate data and perform power analysis for models supported by the lme4 package.}
+Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.\footnote{A less flexible alternative would be to use the simr package (Green \& MacLeod, 2016), which can be used to both simulate data and perform power analysis for models supported by the lme4 package.}
 
@@
 
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables \(\mathbf{X}\) and random effects \(\mathbf{U}\), transformed by a link function \(g()\) is modeled as a linear combination \(\eta\) of the predictor variables \(\mathbf{X}\), the random effects \(\mathbf{U}\) and the model parameters \(\mathbf{\beta}\) (Fahrmeir et al., 2021).
-\[
-g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta <!-- SG: Sorry for maybe a stupid question, but what does E stand for?
-and why are model parameters beta mentioned but not in the equation?-->
-\] Equivalently, the conditional expected value is modelled as the linear combination \(\eta\), transformed by the inverse link function \(g^{-1}()\). \[
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables \(\mathbf{X}\) and random effects \(\mathbf{U}\), transformed by a link function \(g()\) is modeled as a linear combination \(\eta\) of the predictor variables \(\mathbf{X}\), the random effects \(\mathbf{U}\) and the model parameters \(\mathbf{\beta}\) (Fahrmeir et al., 2021): \[
+g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
+\] Equivalently, the conditional expected value is modeled as the linear combination \(\eta\), transformed by the inverse link function \(g^{-1}()\): \[
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-\] If the dependent variable (i.e., diagnostic decision) \(Y\) is a binary variable with values \(0\) (i.e.~inaccurate), or \(1\) (i.e., accurate), the conditional expected value is equivalent to the probability: \[
+\] If the dependent variable (i.e., diagnostic decision) \(Y\) is a binary variable with values \(0\) (i.e., inaccurate), or \(1\) (i.e., accurate), the conditional expected value is equivalent to the probability: \[
 P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
@@
 \item
-  \(advice\_present_{si}\): whether subject \(s\) was presented with AI advice (1) or not (0) when asked to asses item \(i\)
+  \(advice\_present_{si}\): whether subject \(s\) was presented with AI advice (1) or not (0) when asked to assess item \(i\)
 \item
@@
 In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors (\(advice\_present\), \(advice\_correct\)), one level-2 predictor (\(expert\)) and two cross-level interactions (\(expert \cdot advice\_present\), \(expert \cdot advice\_correct\)).
-
-To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions here.
+To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.
 
@@
 As assumed by standard GLMMs, the \texttt{add\_ranef} function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user.
-With all design variables done, we are ready to simulate our model equation as outlined in equation X.
-The linear predictor variable \texttt{linpred} (\(\eta\) in the GLMM model equations) combines the predictor variables, random effects and model parameters as assumed by our model.
+With all design variables done, we are ready to simulate our model equation outlined in the last section.
+The linear predictor variable \texttt{linpred} (\(\eta\) in the GLMM model equations) combines the predictor variables, random effects, and model parameters as assumed by our model.
 We then transform the linear predictor with the inverse-link function to compute \texttt{y\_prob}, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with \texttt{plogis} and the logit link with \texttt{qlogis}).
-In the final step, we simulate the binary dependent variable \texttt{y\_bin} (i.e., whether the subject chooses the accurate diagnosis for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability \texttt{y\_prob}.
+In the final step, we simulate the binary dependent variable \texttt{y\_bin} (i.e., whether the subject makes an accurate diagnostic decision for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability \texttt{y\_prob}.
 
@@
 
-In this section, we show how to fit a GLMM with lme4, interpret the model and test hypotheses derived from a research question.
-
+In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
 We simulate data according to our model, in which 100 subjects respond to 50 items (we use \texttt{set.seed} to make the simulation reproducible).
-However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyse this data.
+However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.
 
@@
 
+\begin{verbatim}
+## Generalized linear mixed model fit by maximum likelihood (Laplace
+##   Approximation) [glmerMod]
+##  Family: binomial  ( logit )
+## Formula: 
+## y_bin ~ 1 + expert + advice_present + advice_correct + expert:advice_present +  
+##     expert:advice_correct + (1 | subject) + (1 | item)
+##    Data: dat
+## 
+##      AIC      BIC   logLik deviance df.resid 
+##   4149.4   4201.6  -2066.7   4133.4     4992 
+## 
+## Scaled residuals: 
+##     Min      1Q  Median      3Q     Max 
+## -5.7669  0.2125  0.3046  0.4317  2.1056 
+## 
+## Random effects:
+##  Groups  Name        Variance Std.Dev.
+##  subject (Intercept) 0.3148   0.5611  
+##  item    (Intercept) 0.1624   0.4029  
+## Number of obs: 5000, groups:  subject, 100; item, 50
+## 
+## Fixed effects:
+##                       Estimate Std. Error z value Pr(>|z|)    
+## (Intercept)             1.0339     0.1103   9.374  < 2e-16 ***
+## expert                  1.1849     0.2096   5.654 1.57e-08 ***
+## advice_present         -1.3436     0.1206 -11.143  < 2e-16 ***
+## advice_correct          2.6154     0.1273  20.540  < 2e-16 ***
+## expert:advice_present   1.0589     0.2940   3.601 0.000317 ***
+## expert:advice_correct  -1.8104     0.2915  -6.211 5.27e-10 ***
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##             (Intr) expert advc_p advc_c exprt:dvc_p
+## expert      -0.377                                 
+## advic_prsnt -0.349  0.176                          
+## advic_crrct  0.023  0.001 -0.668                   
+## exprt:dvc_p  0.143 -0.448 -0.412  0.276            
+## exprt:dvc_c -0.008  0.004  0.292 -0.435 -0.686
+\end{verbatim}
+
 In the output, the \texttt{Estimate} column in the \texttt{Fixed\ effects} table contains the estimates for the \(\beta\) parameters, while the \texttt{Std.Dev.} column in the \texttt{Random\ effects} table contains the estimates for \(\sigma_S\) and \(\sigma_I\).
 
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the \(\beta\) parameters are connected to the modelled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the \(\beta\) parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
 The most simple interpretation works by imagining a subject with average ability (\(u_{0s} = 0\)) responding to an item (i.e., CT scan) with average difficulty (\(u_{0i} = 0\)).
@@
 
-With the marginaleffects package (Arel-Bundock, 2023), we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed \textbf{and} random effects), and plot the average probability with confidence intervals for each experimental condition:
+With the marginaleffects package (Arel-Bundock, 2023), we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed \textbf{and} random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \ref{fig:marginaleffects}:
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/marginaleffects-1.jpeg}
+\caption{\label{fig:marginaleffects}Marginal distributions including means and 95\% confidence intervals for all experimental conditions computed with the marginaleffects package.}
+\end{figure}
+
 \hypertarget{hypothesis-testing}{%
@@
 
-The \texttt{Fixed\ effects} table in Figure X also includes p-values for hypothesis tests with null hypotheses of the style \(H_0: \beta = 0\).
+The \texttt{Fixed\ effects} table in the lme4 summary output also includes p-values for hypothesis tests with null hypotheses of the style \(H_0: \beta = 0\).
 However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.
@@
 
-We can specify and test hypotheses like these with the multcomp package (Hothorn, Bretz, \& Westfall, 2008a) as follows:
+We can specify and test hypotheses like these with the multcomp package (Hothorn, Bretz, \& Westfall, 2008) as follows:
 
@@
 
+\begin{verbatim}
+## advice_present + advice_correct + expert:advice_present + expert:advice_correct 
+##                                                                     0.006407391 
+##                                                 advice_present + advice_correct 
+##                                                                     0.000000000 
+##                                   -1 * (advice_present + expert:advice_present) 
+##                                                                     0.143963670 
+##                                                           -1 * (advice_present) 
+##                                                                     0.000000000
+\end{verbatim}
+
 Because all hypotheses tested simultaneously with the \texttt{glht} function must have the same direction, we flip the sign of inequalities three and four by multiplying them with \(-1\).
-The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously (Hothorn, Bretz, \& Westfall, 2008b).
-However, the combined null hypothesis in our exemplary research question should only be rejected if \textbf{all} individual null hypotheses are rejected {[}i.e.~intersection-union setting; Dmitrienko and D'Agostino (2013){]}.
+The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously (Hothorn et al., 2008).
+However, the combined null hypothesis in our exemplary research question should only be rejected if \textbf{all} individual null hypotheses are rejected {[}i.e., intersection-union setting; Dmitrienko and D'Agostino (2013){]}.
 In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
@@
 
-Often no meta-analytic results or conclusive data from pilot studies are available.
-All parameter values in our present case study have been determined based on results from prior own work and from the literature.
-Additionally, we had repeated exchanges with our affiliated domain experts from radiology.
+Ideally, one would rely on meta-analytic results or conclusive data from pilot studies.
+However, these are sometimes not readily available.
+All parameter values in our present case study have been determined based on results from related prior work.
+Additionally, we had repeated discussions with our affiliated domain experts in radiology to check our assumptions.
 
@@
 
-Table X shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model:
-\textbar{} \emph{Experimental condition} \textbar{} \(P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})\) \textbar{} \emph{Implied equation} \textbar{}
-\textbar--------------------\textbar:------------------------------:\textbar--------------------\textbar{}
-\textbar{} no advice, no expert \textbar{} 0.70 \textbar{} \(logit(0.70) = \beta_0\) \textbar{}
-\textbar{} no advice, expert \textbar{} 0.90 \textbar{} \(logit(0.90) = \beta_0 + \beta_e\) \textbar{}
-\textbar{} false advice, no expert \textbar{} 0.40 \textbar{} \(logit(0.40) = \beta_0 + \beta_a\) \textbar{}
-\textbar{} false advice, expert \textbar{} 0.85 \textbar{} \(logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \textbar{}
-\textbar{} correct advice, no expert \textbar{} 0.90 \textbar{} \(logit(0.90) = \beta_0 + \beta_a + \beta_c\) \textbar{}
-\textbar{} correct advice, expert \textbar{} 0.95 \textbar{} \(logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\) \textbar{}
-
-This table can be used to compute the implied values for the \(\beta\) parameters, starting with the first equation and reinserting the computed \(\beta\) values in all following equations:
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:probtable}Assumed probabilities that an average subject solves an average item in each experimental condition.}
+
+\begin{tabular}{lll}
+\toprule
+Experimental condition & $P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$ & Implied equation\\
+\midrule
+no advice, no expert & 0.70 & $logit(0.70) = \beta_0$\\
+no advice, expert & 0.90 & $logit(0.90) = \beta_0 + \beta_e$\\
+false advice, no expert & 0.40 & $logit(0.40) = \beta_0 + \beta_a$\\
+false advice, expert & 0.85 & $logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}$\\
+correct advice, no expert & 0.90 & $logit(0.90) = \beta_0 + \beta_a + \beta_c$\\
+correct advice, expert & 0.95 & $logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}$\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Implied equations are derived based on the model equations and setting all random intercept terms to 0.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+Table \ref{tab:probtable} shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
+The table can be used to compute the implied values for the \(\beta\) parameters, starting with the first equation and reinserting the computed \(\beta\) values in all following equations:
 
@@
 
+\begin{verbatim}
+##        b_0        b_e        b_a        b_c       b_ea       b_ec 
+##  0.8472979  1.3499267 -1.2527630  2.6026897  0.7901394 -1.3928518
+\end{verbatim}
+
 It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g.~
@@
 \end{Shaded}
+
+\begin{verbatim}
+## [1] 0.95
+\end{verbatim}
 
@@
 
-Figure X shows the model implied marginal distributions, including the mean, 66\% and 95\% intervals.
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/margdist1-1.jpeg}
+\caption{\label{fig:margdist1}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions.}
+\end{figure}
+
+Figure \ref{fig:margdist1} shows the model implied marginal distributions, including the mean, 66\% and 95\% intervals.
 We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
@@
 
-As an example, this presentation reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/margdist2-1.jpeg}
+\caption{\label{fig:margdist2}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01.}
+\end{figure}
+
+As an example, Figure \ref{fig:margdist2} reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
 In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
@@
 
-We could also have a closer look at variability between items by setting the subject standard deviation to almost zero (\(\sigma_S = 0.01\), see Figure X in Appendix X).
+We could also have a closer look at variability between items by setting the subject standard deviation to almost zero (\(\sigma_S = 0.01\)).
 
 The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
-For this example, we have set \(\sigma_S = 3\) and \(\sigma_I = 3\).
+For Figure \ref{fig:margdist3}, we have set \(\sigma_S = 3\) and \(\sigma_I = 3\).
 This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-For example, we would expect even experts with low ability to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
+However, these high standard deviations do not account for the inherent variability and complexity of human performance.
+For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/margdist3-1.jpeg}
+\caption{\label{fig:margdist3}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3.}
+\end{figure}
 
@@
 We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
-With the current settings, this simulation takes about X minutes on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
+With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
 If you want to quickly experiment with the code yourself, a setting with \texttt{workers\ =\ 4} and \texttt{rep\ =\ 5} should finish in less than 5 minutes, even on smaller machines.
@@
 
-The result of the computation is a data.frame that contains the p-values of all tested hypotheses for each simulated dataset.
+The result of the computation is a data frame that contains the p-values of all tested hypotheses for each simulated dataset.
 In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
+When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process.
 We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
@@
 \NormalTok{                                power, ci.lwr, ci.upr)), }
-    \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \FloatTok{4.5}\NormalTok{) }\SpecialCharTok{+}
+    \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \DecValTok{4}\NormalTok{) }\SpecialCharTok{+}
   \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)) }\SpecialCharTok{+}
@@
 
-As should be the case, power estimates increase with both the number of subjects and the number of items.
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/finalpwr-1.jpeg}
+\caption{\label{fig:finalpwr}Simulation-based power estimates including 95\% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05.}
+\end{figure}
+
+As should be the case, power estimates in Figure \ref{fig:finalpwr} increase with both the number of subjects and the number of items.
 The confidence intervals indicate how precisely power was estimated by our simulation.
 Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (\texttt{rep}) in the simulation.
-In practice, power simulations are often run multiple times with adjusted combinations of sample sizes.
+In practice, data simulations are often run multiple times with adjusted combinations of sample sizes.
 When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of \texttt{n\_subjects} and \texttt{n\_items}.
-When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of power simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
+When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
 This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.
@@
 Much has been written on the optimal amount of power to target in empirical research.
-The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of \(\alpha = 0.05\)), but depending on the research goals of the study, there are often good reasons to move away from this standard (Lakens, 2022b; Lakens et al., 2018).
+The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of \(\alpha = 0.05\)), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints (Lakens, 2022b; Lakens et al., 2018).
 When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
@@
 Experimental research requires careful planning and consideration of statistical power to ensure robust and meaningful results.
-While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations are required.
-The present tutorial presents a specific example of how to run power simulations for experimental designs with GLMMs.
+While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations with GLMMs are required.
+The present tutorial presents a specific case study with corresponding code of how to conduct a simulation-based power analysis for experimental designs with GLMMs.
 
@@
 
-In our case study, we have performed power simulations based on a single set of parameter values that reflect our assumptions of an expected effect size.
+In our case study, we have performed simulation-based power analysis from a single set of parameter values that reflect our assumptions of an expected effect size.
 Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
 Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
-When sample sizes are chosen based on the results of our power simulations, a future study will be informative to reject the null hypothesis if an effect of our \emph{expected size} is present.
+When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our \emph{expected size} is present.
 However, if the true effect is indeed smaller, the power will be lower, and the study might not be sufficiently informative.
@@
 
-Beyond the specifics of our concrete case study, we want to outline six avenues regarding the future of power simulations in experimental research:
+Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based power analysis in experimental research:
 
@@
 \item
-  The need for power simulations in experimental research: Experimental research often involves intricate designs and complex models that cannot be adequately addressed by heuristics or user-friendly software.
-Power simulations offer a solution by providing experimental researchers with a tailored approach to estimating statistical power.
-These simulations take into account the specific study design, account for the underlying assumptions, and offer more accurate power estimates.
+  The growing need for simulation-based power analyses in experimental research: In order to conduct well-powered research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
+Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
 \item
-  Managing simulations with discrete predictor variables: Power simulations become more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
+  Managing data simulations more easily with discrete predictor variables: Simulation-based power analysis becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
   This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising accuracy.
+By simplifying the simulation process, researchers can obtain reliable power estimates without compromising realistic assumptions about the data-generating process implied by the study design.
 \item
-  Teaching power simulation skills: The ability to conduct power simulations is a valuable skill that should be taught to experimental researchers.
-By incorporating this training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-  Equipping researchers with the knowledge and tools to perform power simulations enables them to make informed decisions and enhance the rigor of their studies.
-  The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look like, might also contribute to planning more insightful studies that can answer more precise research questions (Yarkoni, 2022).
+  Teaching data simulation skills: The ability to conduct simulation-based power analysis is a valuable skill that should be taught to experimental researchers.
+By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
+  Equipping researchers with the knowledge and tools to perform simulation-based power analyses enables them to make informed decisions and enhance the rigor of their studies.
+  The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions (Yarkoni, 2022).
 \item
-  Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform power analysis and the perceived effort estimated by researchers and collaborators in experimental research.
-Many researchers request power simulations from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-It is crucial to raise awareness about the effort involved in power simulations to ensure realistic expectations and effective collaboration between researchers and methodological experts.
+  Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform simulation-based a priori power analysis and the perceived effort estimated by researchers and collaborators in experimental research.
+Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
+It is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.
 \item
-  Recognizing the value of power simulations: Power simulations are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
+  Recognizing the value of simulation-based power analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
-The importance of power simulations can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports .
+Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports (Chambers \& Tzavella, 2022).
 \item
-  Integration with Open Science and preregistration practices: Power simulations align well with the principles of Open Science and preregistration in experimental research.
+  Integration with Open Science and preregistration practices: Simulation-based powers analysis aligns well with the principles of Open Science and preregistration in experimental research.
 When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their power simulations, researchers enhance transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
+By preregistering their simulation-based power analysis, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
 \end{enumerate}
@@
 
-Power simulations play a critical role in experimental research, allowing investigators to tailor power estimation to the unique aspects of their experiments.
-Through this tutorial, we aim to provide researchers with the necessary skills and tools to perform these simulations for GLMMs themselves.
-By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work (Yarkoni, 2022).
+In the wake of the replication crisis and myriad of underpowered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
+Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs necessitating data simulation-based approaches.
+Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based power analysis with GLMMs themselves.
+By incorporating GLMMs and a priori power analysis into their work, researchers can enhance the replicability and credibility of their experiments (Yarkoni, 2022).
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-brownIntroductionLinearMixedEffects2021}{}}%
-Brown, V.
-A.
-(2021).
-An {Introduction} to {Linear Mixed-Effects Modeling} in {R}.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{4}(1), 2515245920960351.
-\url{https://doi.org/10.1177/2515245920960351}
-
 \leavevmode\vadjust pre{\hypertarget{ref-brysbaertPowerAnalysisEffect2018}{}}%
@@
 Bürkner, P.-C. (2017). Brms: {An R Package} for {Bayesian Multilevel Models Using Stan}. \emph{Journal of Statistical Software}, \emph{80}, 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
+
+\leavevmode\vadjust pre{\hypertarget{ref-chambersPresentFutureRegistered2022}{}}%
+Chambers, C. D., \& Tzavella, L. (2022). The past, present and future of {Registered Reports}. \emph{Nature Human Behaviour}, \emph{6}(1), 29--42. \url{https://doi.org/10.1038/s41562-021-01193-7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-champelyPackagePwr2018}{}}%
+Champely, S., Ekstrom, C., Dalgaard, P., Gill, J., Weibelzahl, S., Anandkumar, A., \ldots{} De Rosario, M. H. (2018). Package {``pwr.''} \emph{R Package Version}, \emph{1}(2).
+
+\leavevmode\vadjust pre{\hypertarget{ref-cockburnThreatsReplicationCrisis2020}{}}%
+Cockburn, A., Dragicevic, P., Besançon, L., \& Gutwin, C.
+(2020).
+Threats of a replication crisis in empirical computer science.
+\emph{Communications of the ACM}, \emph{63}(8), 70--79.
+\url{https://doi.org/10.1145/3360311}
 
@@
 Hothorn, T., Bretz, F., \& Westfall, P.
-(2008b).
+(2008).
 Simultaneous {Inference} in {General Parametric Models}.
@@
 \url{https://doi.org/10.1002/bimj.200810425}
-
-\leavevmode\vadjust pre{\hypertarget{ref-R-multcomp}{}}%
-Hothorn, T., Bretz, F., \& Westfall, P. (2008a). Simultaneous inference in general parametric models. \emph{Biometrical Journal}, \emph{50}(3), 346--363.
 

### 2023-09-18 docs/manuscript.Rmd
added osf info to manuscript

@@
   This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525).
-  All R code in this manuscript is available in the project's repository on the Open Science Framework (OSF): https://osf.io/dhwf4/
+  All R code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (OSF): https://osf.io/dhwf4/
 
@@
 In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
-The corresponding code in this manuscript is available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
+All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
 

### 2023-09-18 docs/manuscript.tex
last edits for preprint publication

@@
 {}{\PatchFailed}
-\keywords{power analysis, data simulation, sample size, generalized linear mixed model\newline\indent Word count: 7598}
+\keywords{power analysis, data simulation, sample size, generalized linear mixed model\newline\indent Word count: 7600}
 \usepackage{lineno}
@@
 This research was funded by a grant from the Volkswagen Foundation (Grant No.~98525).
-All R code in this manuscript is available in the project's repository on the Open Science Framework (OSF): \url{https://osf.io/dhwf4/}
+All materials (reproducible manuscript incl.~R code, simulation results) are available in the project's repository on the Open Science Framework (OSF): \url{https://osf.io/dhwf4/}
 
@@
 In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
-The corresponding code in this manuscript is available in the project's repository on the Open Science Framework (\url{https://osf.io/dhwf4/}).
+All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (\url{https://osf.io/dhwf4/}).
 

### 2023-09-18 docs/manuscript.Rmd
last edits for preprint publication

@@
   This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525).
-  All R code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (OSF): https://osf.io/dhwf4/
+  All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF): https://osf.io/dhwf4/
 
@@
 
-wordcount         : "7598"
+wordcount         : "7600"
 
@@
 floatsintext      : yes
-linenumbers       : yes
+linenumbers       : no
 draft             : no

### 2023-10-19 docs/manuscript.Rmd
removed hci references from manuscript

@@
   When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
-However, the recent replication crisis in Human-Computer Interaction (HCI) and other empirical research fields underlines the need for more rigorous statistical methodology and well-powered designs.
+However, the recent replication crisis underlines the need for more rigorous statistical methodology and well-powered designs.
 Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
@@
 When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data [@lakensSampleSizeJustification2022].
-However, the recent replication crisis in Human-Computer Interaction (HCI) and several other disciplines grounded on empirical research has illustrated many challenges surrounding the reproducibility and reliability of findings [@cockburnThreatsReplicationCrisis2020; @robertsonImprovingStatisticalPractice2016; @yarkoniGeneralizabilityCrisis2022].
+However, the recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of findings [@yarkoniGeneralizabilityCrisis2022].
 As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
@@
 As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data in HCI and other empirical disciplines because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].
+GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
 

### 2023-10-19 docs/manuscript.Rmd
removed general outlook and discussion recap

@@
 
-# Discussion
-
-<!-- short summary of key takeaways from this tutorial -->
-Experimental research requires careful planning and consideration of statistical power to ensure robust and meaningful results.
-While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations with GLMMs are required.
-The present tutorial presents a specific case study with corresponding code of how to conduct a simulation-based power analysis for experimental designs with GLMMs.
-
 ## Expected effect size vs. smallest effect size of interest: sensitivity power analysis
@@
 
-## Outlook
-
-<!-- Outlook -->
-Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based power analysis in experimental research:
-
-1.
-The growing need for simulation-based power analyses in experimental research: In order to conduct well-powered research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
-Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
-
-2.
-Managing data simulations more easily with discrete predictor variables: Simulation-based power analysis becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
-This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising realistic assumptions about the data-generating process implied by the study design.
-
-3.
-Teaching data simulation skills: The ability to conduct simulation-based power analysis is a valuable skill that should be taught to experimental researchers.
-By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping researchers with the knowledge and tools to perform simulation-based power analyses enables them to make informed decisions and enhance the rigor of their studies. 
-The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
-
-4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform simulation-based a priori power analysis and the perceived effort estimated by researchers and collaborators in experimental research.
-Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-It is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.
-
-5.
-Recognizing the value of simulation-based power analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
-They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
-Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
-
-6.
-Integration with Open Science and preregistration practices: Simulation-based powers analysis aligns well with the principles of Open Science and preregistration in experimental research.
-When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their simulation-based power analysis, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
-
 # Conclusion

### 2023-10-19 docs/manuscript.Rmd
added more Credit roles in line with journal authorship requirements

@@
     role:
+      - "Formal Analysis"      
       - "Writing - Review & Editing"
@@
     role:
-      - "Resources"
+      - "Resources" # this is not sufficient as co-authors for AMPPS...
   - name          : "Susanne Gaube"
@@
     role:
+      - "Conceptualization"
       - "Writing - Review & Editing"

### 2023-10-23 docs/manuscript.Rmd
made final changes for submission

--- 2023-10-19 docs/manuscript.Rmd
@@
     role:
+      - "Conceptualization"
       - "Resources"
+      - "Writing - Review & Editing"
   - name          : "Susanne Gaube"
@@
     role:
+      - "Conceptualization"
+      - "Supervision"
       - "Writing - Review & Editing"
-      - "Supervision"
 
@@
 Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs, necessitating data simulation approaches.
+Yet, available methods and software for a priori power analyses for GLMMs are often limited to specific designs, while data simulation approaches offer more flexibility.
 Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs.
@@
 
-wordcount         : "7600"
+wordcount         : "7504"
 
@@
 footnotelist      : no
-
-#csl               : "acm-sigchi-proceedings.csl" #added this for chi submission, not needed for preprint
 

### 2024-03-12 docs/manuscript.Rmd
start rewrite of intro structure (work in progress)

@@
 ---
-title             : "A Tutorial on Tailored Simulation-Based Power Analysis for Experimental Designs with Generalized Linear Mixed Models"
-shorttitle        : "Simulation-Based Power Analysis for GLMMs"
+title             : "A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models"
+shorttitle        : "Simulation-Based Sample Size Planning for GLMMs"
+author:
 
-author: 
   - name          : "Timo K. Koch"
@@
       - "Writing - Review & Editing"
-
 affiliation:
@@
     institution   : "University College London, Global Business School for Health"
-
 note: "\\clearpage" #added this to move entire author note to second page
@@
   All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF): https://osf.io/dhwf4/
-
 abstract: |
-
   When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
@@
 keywords          : "power analysis, data simulation, sample size, generalized linear mixed model"
-
 wordcount         : "7504"
-
 bibliography      : ["../Simulation Tutorial Paper.bib", "r-references.bib"]
-
 floatsintext      : yes
@@
 mask              : no
-
 figurelist        : no
@@
 footnotelist      : no
-
 classoption       : "man"
@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = TRUE, fig.keep = "high", dev = "jpeg")
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = FALSE, fig.keep = "high", dev = "jpeg")
 ```
+
+# New Intro Structure
+
+# Introduction
+
+<!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
+When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data [@lakensSampleSizeJustification2022].
+However, the recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of findings [@yarkoniGeneralizabilityCrisis2022].
+As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
+While easy-to-use software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) design analysis for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
+In the present work, we provide a tutorial on performing tailored simulation based sample size planning for GLMMs.
+After introducing general concepts, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
+Then we describe general steps and decisions that are necessary whenever performing tailored data simulation.
+Because the details of these steps vary considerably for each specific study, we finish with a case study from the field of human-AI (artificial intelligence) interaction research.
+
+## Prior Knowledge
+
+Simulation-based sample size planning for generalized linear models is a complex topic that touches on a variety of statistical concepts and programming skills.
+Although we give some short introduction to clarify our terminology, to profit the most from our tutorial, our readers require prior knowledge in the following topics (or refer to our collection of background literature):
+
+Readers should understand basic statistical concepts like hypothesis tests and their statistical power but also confidence intervals and their precision.
+Some background knowledge in empirical estimands, statistical estimates, and causal inference is very useful but not absolutely necessary.
+
+In addition, readers should have basic programming skills in R on how to simulate data and repeat these simulations in parallel to speed up computation.
+While there are a myriad options for data simulation in R beyond the basic functionality in the base and stats packages, in our tutorial we use functions from the tidyverse and the faux package.
+
+Finally, readers should be familiar with regression modeling in general and GLMMs in particular.
+In our tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process.
+While it is not necessary to understand the technical details of how GLMMs are estimated, it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts).
+
+Literature:
+
+- Lakens ebook
+- R for data science
+- Bruine article, faux documentation
+- draschkow power tutorial
+- what is your estimand? paper
+
+## General concepts
+
+### Planning for statistical power or precision
+
+The majority of empirical studies in psychology applies hypothesis testing.
+As a consequence, the dominant approach for sample size planning is based on the power analysis (i.e., planing for power).
+However, sample size planning can also be based on the precision of parameter estimates (planning for precision).
+
+#### Planning for power
+
+<!-- Some basics on statistical power -->
+In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022].
+Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
+Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
+If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered".
+
+<!-- what are the resulting issues with underpowered studies? -->
+Conducting underpowered research has many negative consequences.
+First, relying on underpowered experiments may yield inconclusive (if researchers acknowledge the small evidential value of an underpowered study in the limitations section) or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
+Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
+
+#### Planning for precision
+
+### Generalized linear mixed models (GLMMs)
+
+## Simulation based sample size planning with GLMMs
+
+- Insert table on existing R packages for power analysis with GLMMs
+
+## When do we profit from tailored data simulation?
+
+Performing tailored simulation based design analysis is more complicated and time consuming than using existing software tools for sample size planning.
+Thus, it is worth discussing under which circumstances we most profit from or are forced to perform tailored simulation based design analysis, because existing software solutions do not meet our requirements.
+
+### Complex study designs
+
+Real studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
+
+#### Missing data
+
+Missing data is a frequent issue in applied data analysis.
+
+- Missing data processes; data missing by design; dropout
+
+#### Data generating processes beyond standard GLMMs
+
+While GLMMs are a very flexible model class that can handle a large variety of outcome variables, psychological researchers are becoming increasingly aware that many psychological datasets might profit from even more sophisticated models.
+Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects.
+
+    - possibility to extend beyond standard GLMMs (e.g. zero-inflated outcomes, censoring, nonlinear effects, …)
+
+### Complex statistical hypotheses
+
+The most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is some coefficient in a regression model.
+However, many research questions in psychology actually require testing more complex statistical hypotheses.
+
+#### Directed hypotheses and testing against values other than 0
+
+In the new era of preregistration and registered reports, most research questions are tested with directed hypotheses, because good theories should at least postulate whether some psychological effect of interest is positive or negative.
+Even better theories should be able to specify a smallest effect sizes of interest (SESOI) that must be exceeded if the psychological effect has some practical relevance in a specific application.
+In combination, this might require a test like $H_0: \beta \leq 0.1$.
+
+#### Combined hypotheses
+
+Elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example $H_0: \beta_0 + \beta_1 \leq 0$.
+If the research question consists only of a single hypothesis of this sort, it is often possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
+However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypotheses.
+For example, a combined null hypothesis $H_0$ might consist of two single null hypotheses $H_{01}: \beta_1 \leq 0$ and $H_{02}: \beta_0 + \beta_1 \leq 0$.
+For some research questions, the combined null hypothesis $H_0$ would be rejected if both $H_{01}$ *AND* $H_{02}$ are rejected.
+For other research questions, the combined null hypothesis $H_0$ would be rejected if $H_{01}$ *OR* $H_{02}$ *OR* both are rejected.
+If the global hypothesis $H_0$ is combined with *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis.
+However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global hypothesis test.
+
+None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here.
+In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation based design analysis.
+
+### Planning for precision
+
+- Planning for expected width of confidence intervals
+    - The most interesting contrasts are usually on the scale of the outcome instead of the scale of the coefficients
+
+### No prior studies or pilot data
+
+- No prior studies or pilot data available
+    - Detailed data simulation is necessary to identify plausible scenarios for effects sizes
+
+## General steps in simulation-based sample size planning
+
+Each tailored simulation based design analysis requires a series of steps and decisions.
+
+### Simulate the data generating process
+
+### Estimate the statistical model
+
+### Specify the estimand
+
+- Specify estimand (conditional vs. marginal perspective)
+
+### Compute the estimate
+
+- Compute hypothesis test or confidence interval
+
+### Perform repeated simulations
+
+- Write code to perform repeated simulations
+
+<!-- ### Sensitivity analysis -->
+
+## Strategies on specifying the population parameters
+
+As discussed earlier, tailored simulation based design analysis is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
+In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
+Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
+In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available.
+Our strategies to specify population parameters will require access to domain knowledge from domain experts.
+Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based design analysis.
+This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available domain knowledge of domain experts.
+
+Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
+In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions, that can be adapted for our purposes.
+
+### Insightful descriptive statistics
+
+For most GLMMs, the interpretation of model parameters if quite complicated which makes it difficult to specify plausible values for a specific application.
+However, the mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
+The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
+For example, domain experts might not be able to specify plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
+However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.
+which descriptive statistics they expect to observe (on average), when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
+The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
+Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
+
+### Insightful model based quantities
+
+Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
+This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
+An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
+For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
+Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
+For example, the domain expert might deem it unplausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
+
+### Iterative process with domain experts
+
+In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
+In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
+As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
+For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
+Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
+Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
+During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
+
+# Case study
+
+# Conclusion
+
+----
 
@@
 $$ Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: $$
-E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
+E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
 $$ If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
@@
 We can specify and test hypotheses like these with the multcomp package [@hothornSimultaneousInferenceGeneral2008] as follows:
+
+```{r}
+library(marginaleffects)
+hypotheses(fit, hypothesis = c(
+  "`advice_present` + `advice_correct` + `expert:advice_present` +
+  `expert:advice_correct` = 0",
+  "`advice_present` + `advice_correct` = 0",
+  "`advice_present` + `expert:advice_present` = 0",
+  "`advice_present` = 0"
+), equivalence = 0)
+
+avg_predictions(fit, variables = list(expert = 0:1, advice_present = 0:1, advice_correct = 0:1)) %>%
+  hypotheses(hypothesis = c(
+    "b8 - b2 = 0",
+    "b7 - b1 = 0",
+    "b2 - b4 = 0",
+    "b1 - b3 = 0"
+    )) %>%
+  inferences(method = "delta")
+
+comparisons(fit)
+avg_comparisons(fit, hypothesis = c(
+  "`advice_present` + `advice_correct` + `expert:advice_present` +
+  `expert:advice_correct` = 0",
+  "`advice_present` + `advice_correct` = 0",
+  "`advice_present` + `expert:advice_present` = 0",
+  "`advice_present` = 0"
+))
+```
 

### 2024-03-12 docs/manuscript.Rmd
further work on new intro (work in progress)

@@
 
-In addition, readers should have basic programming skills in R on how to simulate data and repeat these simulations in parallel to speed up computation.
-While there are a myriad options for data simulation in R beyond the basic functionality in the base and stats packages, in our tutorial we use functions from the tidyverse and the faux package.
+In addition, readers should have basic programming skills in R on how to simulate data and how to repeat these simulations in parallel to speed up computation.
+While there are a myriad of options for data simulation in R beyond the basic functionality in the base and stats packages, in our tutorial we use functions from the tidyverse and the faux package.
 
@@
 In our tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process.
-While it is not necessary to understand the technical details of how GLMMs are estimated, it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts).
+While it is not necessary to understand the technical details of how GLMMs are estimated, it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the values in the dependent variable are determined by the predictor variables and the random effects.
 
@@
 Missing data is a frequent issue in applied data analysis.
-
-- Missing data processes; data missing by design; dropout
+There can be various reasons for why data is missing.
+For example data can be missing completely at random.
+Alternatively, subjects might drop out or produce missing data as a result of some covariate also measured in the dataset.
+A a different note, many experimental designs contain some conditions in which some values of the predictor variables are missing by design.
+Missing by design conditions can make data analysis more complicated because predictors have to be coded in highly specific ways that prevent the estimated GLMM to become unidentified.
+Whether missing data has an effect on the result of design analysis depends on the characteristics of the missing data process.
+However, except for very simple scenarios, deciding whether missing data can be safely ignored in the design analysis is often difficult to decide on a theoretical basis.
+A better strategy is to explicitly include the (assumed) process of missing data into the data simulation process.
 
@@
 
-- Planning for expected width of confidence intervals
-    - The most interesting contrasts are usually on the scale of the outcome instead of the scale of the coefficients
+As discussed earlier, the two major frameworks for sample size planning are planning for power and planning for precision.
+Although planning for precision has been increasingly discussed as a more suitable strategy for exploratory research, all available software packages for sample size planning with GLMMs are based on power analysis.
+
+As we will show in our case study, tailored simulation based design analysis can also be based on planning for precision.
+The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset and the expected width is estimated across repetitions.
+
+As long as one were interested only in confidence intervals of single model parameters, it might be possible to adapt existing software packages to plan for precision.
+However, we demonstrate in our case study that in realistic research projects with GLMMs, the contrasts of interests and their respective confidence intervals are usually situated on the scale of the (expected) outcomea in contrast to the scale of the model parameters.
+Computing these contrasts and estimating the expected width of these confidence intervals usually requires tailored data analysis that is not possible with the existing software packages for sample size planning.
 
@@
 
-- No prior studies or pilot data available
-    - Detailed data simulation is necessary to identify plausible scenarios for effects sizes
-
-## General steps in simulation-based sample size planning
-
-Each tailored simulation based design analysis requires a series of steps and decisions.
+All frameworks for sample size planning require the user to make assumptions about the expected effect size.
+Assuming the true effect is of this size (or greater), one can compute the (minimum) power of a hypothesis test or the (maximum) expected width of a confidence interval.
+Existing software packages for sample size planning for GLMMs usually require to provide the assumed effect in the unit of some standardized measure of effect size.
+When the researcher has access to similar studies or pilot data, providing such standardized effect sizes is feasible.
+However, this requirement can be an almost impossible challenge when no prior studies of pilot data are available.
+This problem is further exacerbated by the fact that GLMMs are so flexible, that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
+In our experience, using domain knowledge to construct a tailored data simulation is the only solution to determine plausible effect sizes in the absence of prior evidence.
+It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
+However, we would argue that when tailored data simulations are necessary anyways, performing the whole desing analysis in a customized way is preferred over using the existing software packages.
+
+## General steps in tailored simulation-based sample size planning
+
+Although the details differ depending on the specific study characteristics, each tailored simulation based design analysis requires a series of steps and decisions that we will outline in this section.
 
@@
 
+The first step is to write code that simulates the (assumed) data generating process of the planned study.
+This requires specifying a generative process for all predictor variables used in the final data analysis.
+While such assumptions can be quite challenging for observational studies and/or continuous predictor variables, this is less of a problem for experimental studies with categorical predictor variables.
+When all predictor variables have been simulated, one can use the structure of a suitable GLMM to simulate the dependent variable.
+To simulate the GLMM, one requires plausible values for all model parameters.
+We will discuss strategies for how these values can be obtained later.
+Because one has full control over the data generating process in a tailored simulation, it is simple to also model highly specific aspects of the planned study, like data missing by design or assuming that subjects dropout (i.e.
+deleting their data from the simulated dataset) based on other variables in the simulated dataset.
+The quality of the results of the design analysis crucially depends on the plausibility of the simulated data generating process.
+For this reasons, deciding on a specific set of assumptions can easily feel overwelming.
+However, we would argue that even a strongly simplified data generating process (e.g.
+including only a small number of interaction effect; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current state of psychological research where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.
+
 ### Estimate the statistical model
 
+At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset that will later be collected in the planned study.
+The next step is to specify how the statistical model shall be estimated in the actual study collected later.
+This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
+Note that this does not always mean that one will specify the same GLMM that was used when specifying the data generating process.
+On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
+On the other hand, using a more complex model (that perhaps even goes beyond a standard GLMM) for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
+In such more complicated scenarios, the final model specification might also depend on the next decision on specifying the theoretical estimand.
+
 ### Specify the estimand
 
-- Specify estimand (conditional vs. marginal perspective)
+This step requires a clear definition of which theoretical quantity is necessary to answer the specific research question of the study under planning.
+In an ideal world, this should be the first step in every research process.
+However in our experience, many psychological researchers are still not well trained in this task.
+We think that simulating data from a generic GLMM that is suited to describe (at least a rough approximation of) the assumed data generating process, can help to better reason about what information is actually needed to answer the specific research question, and how this information is actually represented in the context of the statistical model representing the data generating process of the planned study.
 

### 2024-03-13 docs/manuscript.Rmd
further work on new intro; delete old intro

@@
 
-# New Intro Structure
-
 # Introduction
@@
 As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
-While easy-to-use software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) design analysis for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
+While easy-to-use software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) sample size planning for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
 In the present work, we provide a tutorial on performing tailored simulation based sample size planning for GLMMs.
@@
 Because the details of these steps vary considerably for each specific study, we finish with a case study from the field of human-AI (artificial intelligence) interaction research.
+
+<!-- <!-- Old Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! --> -->
+<!-- Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs.
+Thereby, we aim to equip researchers with the tools needed to simulate data and determine appropriate sample sizes for their own research.
+-->
 
@@
 
+Conducting research with too small sample sizes can have many negative consequences.
+First, experiments may yield inconclusive or misleading results, hindering the accumulation of knowledge.
+Second, studies that are doomed to never finding a postulated effect due to small sample size waste resources by consuming time, effort, and funding without delivering meaningful results.
+For such reasons, many journals and funding agencies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+Although many pragmatic justifications are possible [@lakensSampleSizeJustification2022], the ideal is often considered to determine a suitable sample size a priori (i.e., before the study is conducted) based on some meaningful computation to assure that the study will be able to fulfill its purpose.
+In this way, adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
+
+<!-- <!-- advantages of an a priori power analysis --> -->
+<!-- Thereby, a priori power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
+-->
+<!-- Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
+-->
+<!-- Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
+-->
+<!-- Adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research. -->
+<!-- Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+-->
+
 The majority of empirical studies in psychology applies hypothesis testing.
-As a consequence, the dominant approach for sample size planning is based on the power analysis (i.e., planing for power).
-However, sample size planning can also be based on the precision of parameter estimates (planning for precision).
+As a consequence, the dominant approach for sample size planning is based on power analysis (i.e., planing for power).
+However, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision).
 
@@
 
-<!-- what are the resulting issues with underpowered studies? -->
-Conducting underpowered research has many negative consequences.
-First, relying on underpowered experiments may yield inconclusive (if researchers acknowledge the small evidential value of an underpowered study in the limitations section) or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
-Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
+When planning for power, a target is set for the statistical power of a hypothesis test of interest (a very common heuristic is 0.8).
+Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that (on average) would guarantee reaching this target.
 
@@
 
+Not all research questions are best answered using hypothesis testing.
+Especially for exploratory studies when little previous research has been conducted, only estimating effects of interest is often more useful.
+When no hypothesis tests are conducted when analysing the planned study, power analysis is not not available for sample size planning.
+Nonetheless, sample size still has a crucial effect on how informative the planned study will be to learn about the effect of interest.
+In this case, the expected width of a confidence interval is used as a target for sample size planning.
+The values inside a confidence interval are commonly interpreted as plausible values for the quantity of interest the confidence interval is supposed to estimate.
+More formally, a 95%-confidence interval provides the smallest interval with the property that upon repeated sampling, 95% of individual confidence intervals would include the true quantity of interest.
+Thus, a narrow confidence interval (assuming the confidence level is kept at a high value) is more informative about the size of the true effect than a wide interval.
+Apart from the desired confidence level, the width of a confidence interval depends strongly on the sample size.
+Because bigger samples carry more information, they lead to smaller confidence intervals (assuming everything else stays the same).
+When planning for precision, a target is set for the expected width of a confidence interval of interest.
+Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that (on average) would guarantee reaching this target.
+
 ### Generalized linear mixed models (GLMMs)
 
+<!-- Advantages of GLMMs over "simpler" models -->
+As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
+GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
+
+<!-- intro into GLMMs -->
+GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data including hierarchical structures  [@fahrmeirRegressionModelsMethods2021].
+In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
+This correlation can arise for various reasons: Responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), both participants come from the same neighborhood (clustering) or both participants responded to the same stimulus (stimulus effects).
+Thus, modeling such correlations is especially important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
+LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on all fixed and random effects).
+They allow for the modeling of fixed effects, which capture the relationships between our usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects exemplified above.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
+
+As mentioned, GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
+GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
+The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
+As an example, think of an experiment with different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
+Here, participants’ behavior is measured repeatedly (e.g., over several sessions).
+The click patterns of participants in one session are likely to be correlated with their previous sessions.
+Finally, the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
+
 ## Simulation based sample size planning with GLMMs
 
+For simpler statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori sample size planning is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018; @kelleyObtainingPowerObtaining2003; @kelleySampleSizePlanning2006].
+However, these software packages are often not flexible enough to perform sample size planning for complex designs.
+
+Existing approaches for sample size planning with GLMMs have almost exlusively focused on planning for power. 
+Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
+Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
+Currently available formula-based software packages for power analysis with multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
+A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
+However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
+This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
+The necessary presumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
+The distributional assumption specifies the distributional family for the outcome variable (when conditioned on all fixed and random effects).
+Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e.,  if and how they are correlated).
+Interpreting these presumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
+Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
+However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge [see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021].
+
 - Insert table on existing R packages for power analysis with GLMMs
@@
 
-Performing tailored simulation based design analysis is more complicated and time consuming than using existing software tools for sample size planning.
-Thus, it is worth discussing under which circumstances we most profit from or are forced to perform tailored simulation based design analysis, because existing software solutions do not meet our requirements.
+Performing tailored simulation based sample size planning is more complicated and time consuming than using existing software tools for sample size planning.
+Thus, it is worth discussing under which circumstances we most profit from or are forced to perform tailored simulation based sample size planning, because existing software solutions do not meet our requirements.
 
@@
 Missing by design conditions can make data analysis more complicated because predictors have to be coded in highly specific ways that prevent the estimated GLMM to become unidentified.
-Whether missing data has an effect on the result of design analysis depends on the characteristics of the missing data process.
-However, except for very simple scenarios, deciding whether missing data can be safely ignored in the design analysis is often difficult to decide on a theoretical basis.
+Whether missing data has an effect on the result of sample size planning depends on the characteristics of the missing data process.
+However, except for very simple scenarios, deciding whether missing data can be safely ignored in the sample size planning is often difficult to decide on a theoretical basis.
 A better strategy is to explicitly include the (assumed) process of missing data into the data simulation process.
@@
 None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here.
-In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation based design analysis.
+In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation based sample size planning.
 
@@
 
-As we will show in our case study, tailored simulation based design analysis can also be based on planning for precision.
-The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset and the expected width is estimated across repetitions.
-
-As long as one were interested only in confidence intervals of single model parameters, it might be possible to adapt existing software packages to plan for precision.
-However, we demonstrate in our case study that in realistic research projects with GLMMs, the contrasts of interests and their respective confidence intervals are usually situated on the scale of the (expected) outcomea in contrast to the scale of the model parameters.
-Computing these contrasts and estimating the expected width of these confidence intervals usually requires tailored data analysis that is not possible with the existing software packages for sample size planning.
+As long as one were interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
+However, we demonstrate in our case study that in realistic research projects with GLMMs, the contrasts of interests and their respective confidence intervals are usually situated on the scale of the (expected) outcome in contrast to the scale of the model parameters.
+We will show that tailored simulation based sample size planning can easily handle the planning for precision scenario.
+The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset and the expected width is estimated.
 
@@
 
-Although the details differ depending on the specific study characteristics, each tailored simulation based design analysis requires a series of steps and decisions that we will outline in this section.
+Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions that we will outline in this section.
 
@@
 deleting their data from the simulated dataset) based on other variables in the simulated dataset.
-The quality of the results of the design analysis crucially depends on the plausibility of the simulated data generating process.
+The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
 For this reasons, deciding on a specific set of assumptions can easily feel overwelming.
@@
 
-- Compute hypothesis test or confidence interval
+When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
+This decision is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests with the marginaleffects R package, using the delta method).
 
@@
 
-- Write code to perform repeated simulations
+Conducting the previous steps enable the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
+The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
+On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter.
+When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.
+When planning for precision, the function should instead return the width of the confidence interval(s) of interest.
+Secondly, we must run this function repeatedly with the same sample and population parameters.
+Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computations, that is running simulations on multiple cores of the computer at the same time to reduce total computation time.
+Thirdly, the results of the repeated simulation must be collected and aggregated.
+When planning for power, we compute the relative frequency of a significant p-value across repeated simulations.
+When planning for precision, we compute the average width of the confidence interval.
+Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted statistical power or the expected width of the confidence interval.
 
@@
 
-As discussed earlier, tailored simulation based design analysis is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
+As discussed earlier, tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
 In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
@@
 Our strategies to specify population parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based design analysis.
+Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based sample size planning.
 This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available domain knowledge of domain experts.
@@
 # Case study
-
-# Conclusion
-
-----
-
-# Introduction
-
-<!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
-When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data [@lakensSampleSizeJustification2022].
-However, the recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of findings [@yarkoniGeneralizabilityCrisis2022].
-As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
-While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
-In the present work, we provide a tutorial consisting of a concrete example for tailored a priori power analyses using data simulations based on GLMMs.
-
-## Statistical power
-
-<!-- Some basics on statistical power -->
-In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022].
-Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
-Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered".
-
-<!-- what are the resulting issues with underpowered studies? -->
-Conducting underpowered research has many negative consequences.
-First, relying on underpowered experiments may yield inconclusive (if researchers acknowledge the small evidential value of an underpowered study in the limitations section) or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
-Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
-
-## A priori power analysis
-
-<!-- define power analysis -->
-A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
-When running a power analysis before data collection, the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
-
-<!-- advantages of an a priori power analysis -->
-Thereby, a priori power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
-Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
-Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
-Adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
-Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-
-<!-- standard software is available for "simple" models and study designs -->
-For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori power analysis is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018].
-However, these software packages are often not flexible enough to perform power analysis for complex designs.
-
-## Generalized linear mixed models (GLMMs)
-
-<!-- Advantages of GLMMs over "simpler" models -->
-As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
-
-<!-- intro into GLMMs -->
-GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data including hierarchical structures  [@fahrmeirRegressionModelsMethods2021].
-In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value for another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
-This correlation can arise for various reasons: Responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), both participants come from the same neighborhood (clustering) or both participants responded to the same stimulus (stimulus effects).
-Thus, modeling such correlations is especially important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
-LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on all fixed and random effects).
-They allow for the modeling of fixed effects, which capture the relationships between our usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects exemplified above.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
-
-As mentioned, GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
-As an example, think of an experiment with different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
-Here, participants’ behavior is measured repeatedly (e.g., over several sessions).
-The click patterns of participants in one session are likely to be correlated with their previous sessions.
-Finally, the outcome variable is binary (click/no click) for each interaction, which follows a binomial distribution.
-
-## Power analysis for GLMMs
-
-<!-- what are the existing options for a priori power analysis for GLMMs / where are existing tutorials limited -->
-Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
-Currently available formula-based software packages for power analysis often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
-A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
-This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
-The necessary presumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
-The distributional assumption specifies the distributional family for the outcome variable (when conditioned on all fixed and random effects).
-Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e.,  if and how they are correlated).
-Interpreting these presumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
-Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge [see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021].
-
-<!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
-Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs.
-Thereby, we aim to equip researchers with the tools needed to simulate data and determine appropriate sample sizes for their own research.
-
-# The present case study
 

### 2024-03-13 docs/manuscript.Rmd
adjust outlook and conclusion to not only speak about power analysis

@@
 
-# Results
+## Results
 
@@
 
-## Expected effect size vs. smallest effect size of interest: sensitivity power analysis
+## Expected effect size vs. smallest effect size of interest: sensitivity analysis
 
@@
 
-## Outlook
+# Outlook
 
 <!-- Outlook -->
-Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based power analysis in experimental research:
+Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based sample size planning in experimental research:
 
 1.
-The growing need for simulation-based power analyses in experimental research: In order to conduct well-powered research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
+The growing need for simulation-based sample size planning in experimental research: In order to conduct informative research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
 Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
@@
 2.
-Managing data simulations more easily with discrete predictor variables: Simulation-based power analysis becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
+Managing data simulations more easily with discrete predictor variables: Simulation-based sample size planning becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
 This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising realistic assumptions about the data-generating process implied by the study design.
+By simplifying the simulation process, researchers can obtain reliable estimates for power or precision without compromising realistic assumptions about the data-generating process implied by the study design.
 
 3.
-Teaching data simulation skills: The ability to conduct simulation-based power analysis is a valuable skill that should be taught to experimental researchers.
+Teaching data simulation skills: The ability to conduct simulation-based sample size planning is a valuable skill that should be taught to experimental researchers.
 By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-Equipping researchers with the knowledge and tools to perform simulation-based power analyses enables them to make informed decisions and enhance the rigor of their studies. 
+Equipping researchers with the knowledge and tools to perform simulation-based sample size planning enables them to make informed decisions and enhance the rigor of their studies. 
 The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
@@
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform simulation-based a priori power analysis and the perceived effort estimated by researchers and collaborators in experimental research.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform simulation-based a priori sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
 Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
@@
 5.
-Recognizing the value of simulation-based power analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
+Recognizing the value of simulation-based design analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
@@
 6.
-Integration with Open Science and preregistration practices: Simulation-based powers analysis aligns well with the principles of Open Science and preregistration in experimental research.
+Integration with Open Science and preregistration practices: Simulation-based sample size planning aligns well with the principles of Open Science and preregistration in experimental research.
 When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their simulation-based power analysis, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
+By preregistering their simulation-based sample size plan, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
 
@@
 In the wake of the replication crisis and myriad of underpowered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs necessitating data simulation-based approaches.
-Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based power analysis with GLMMs themselves.
-By incorporating GLMMs and  a priori power analysis into their work, researchers can enhance the replicability and credibility of their experiments [@yarkoniGeneralizabilityCrisis2022].
+Yet, analytic methods and software cannot be applied to conduct a priori sample size planning for GLMMs necessitating data simulation-based approaches.
+Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based sample size planning with GLMMs themselves.
+By incorporating GLMMs and a priori sample size planning into their work, researchers can enhance the replicability and credibility of their experiments [@yarkoniGeneralizabilityCrisis2022].
 

### 2024-03-13 docs/manuscript.Rmd
move the lme4 package section
minimal work on new marginaleffects code for hypotheses testing

@@
 
-## The lme4 package in R
-
-<!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
-The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
-
-[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
-
-However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]
-
-[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
-
 ## Our specific GLMM
@@
 
+<!-- ## The lme4 package in R -->
+
+<!-- describe the lme4 package and how it works-->
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
+The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
+
+[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
+
+However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
+Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]
+
+[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
+
 In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
@@
 We can specify and test hypotheses like these with the multcomp package [@hothornSimultaneousInferenceGeneral2008] as follows:
-
-```{r}
-library(marginaleffects)
-hypotheses(fit, hypothesis = c(
-  "`advice_present` + `advice_correct` + `expert:advice_present` +
-  `expert:advice_correct` = 0",
-  "`advice_present` + `advice_correct` = 0",
-  "`advice_present` + `expert:advice_present` = 0",
-  "`advice_present` = 0"
-), equivalence = 0)
-
-avg_predictions(fit, variables = list(expert = 0:1, advice_present = 0:1, advice_correct = 0:1)) %>%
-  hypotheses(hypothesis = c(
-    "b8 - b2 = 0",
-    "b7 - b1 = 0",
-    "b2 - b4 = 0",
-    "b1 - b3 = 0"
-    )) %>%
-  inferences(method = "delta")
-
-comparisons(fit)
-avg_comparisons(fit, hypothesis = c(
-  "`advice_present` + `advice_correct` + `expert:advice_present` +
-  `expert:advice_correct` = 0",
-  "`advice_present` + `advice_correct` = 0",
-  "`advice_present` + `expert:advice_present` = 0",
-  "`advice_present` = 0"
-))
-```
 
@@
 Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
+
+We can specify and test hypothesis like these with the marginaleffects [@R-marginaleffects] package.
+
+```{r}
+library(marginaleffects)
+
+hypotheses(fit, hypothesis = c(
+  "`advice_present` + `advice_correct` + `expert:advice_present` +
+  `expert:advice_correct` = 0",
+  "`advice_present` + `advice_correct` = 0",
+  "`advice_present` + `expert:advice_present` = 0",
+  "`advice_present` = 0"
+), equivalence = 0)
+
+hypotheses(fit, hypothesis = c(
+  "`advice_present` + `advice_correct` + `expert:advice_present` +
+  `expert:advice_correct` = 0",
+  "`advice_present` + `advice_correct` = 0",
+  "-1 * (`advice_present` + `expert:advice_present`) = 0",
+  "-1 * (`advice_present`) = 0"
+), equivalence = 0)
+```
+
+## Interval estimation
+
+STILL MISSING
 

### 2024-03-25 docs/manuscript.Rmd
add marginaleffects code on interval estimation
initial work on estimand intro section

@@
 
-- Lakens ebook
-- R for data science
-- Bruine article, faux documentation
-- draschkow power tutorial
-- what is your estimand? paper
+- Lakens ebook evtl. Andy Field für die Basics?
+- R for data science, Bruine article, faux documentation für die R skills
+- kumle power tutorial und irgendwelche paper (am besten aus AMPPS) zu GLMMs
+- what is your estimand? paper für das estimand topic
 
@@
 
+### Define the estimand
+
+#### THEORY
+
+The first step in every research process should always be a clear definition of which theoretical quantity is necessary to answer the specific research question of the study under planning.
+Unfortunately, many psychological researchers are still not well trained in this task.
+The theoretical estimand is a unit specific quantity defined outside of any statistical model.
+Depending on the research and whether the research question is causal, this quantity might be counterfactual and not directly observable.
+Part of the theoretical estimand is also a clear definition of the target population, for which the quantity is of interest.
+
+When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reexpresses the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
+For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantities that can be estimated with a regression model.
+Note however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression.
+
+When the empirical estimand has been specified, the last step is to clearly define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
+When a regression model is used for estimation, this includes decisions like which statistical framework and package is used to fit the model, and how to compute hypothesis tests or confidence intervals.
+
+@lundbergWhatYourEstimand2021
+
+#### PRACTICE
+
+In the present case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) in an AI-enabled diagnostic decision support system.
+The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography.
+In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans require to make the final diagnostic decision.
+The research goal is to validate the effectiveness of the AI-enabled advice.
+We consider the AI-enable as effective, if the following pattern holds, which we will here describe verbally:
+
+*We expect that for BOTH expert radiologists and students under training, correct AI-advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI-advice presented, AND, we expect that for BOTH experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
+
+I becomes clear that our theoretical estimand is be a causal quantity, based on individual quantities define outside of any statistical model.
+To explain why, we focus on only on part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists.
+For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice were presented and the probability that a correct diagnosis is made if no AI-advice were presented.
+Mathematically, we could describe this unit-specific quantity with some pseudo-equation like this (where $s$ stands for a specific expert radiologist and $i$ stands for a specific CT scan):
+
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{correct advice, a specific expert, a specific scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, a specific expert, a specific scan})
+\end{aligned}
+$$
+However, our verbal description is not precise enough yet, because the above quantity exists and may differ for each combination of expert and scan.
+We must specify how we want to aggregate the quantity across different experts and scans.
+If we would compute the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts.
+Although often debated whether this is a good idea, psychological research often focuses on a different quantity.
+Instead of averaging over experts and scans, we focus in a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
+With this definition, the theoretical estimand in our above example can be expressed as:
+
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{correct advice, an average expert, an average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan})
+\end{aligned}
+$$
+In total, our theoretical estimand consists of four expressions like the one above, the remaining three are:
+
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{wrong advice, an average expert, an average scan})
+\end{aligned}
+$$
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{correct advice, an average student, an average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, an average student, an average scan})
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{no advice, an average student, an average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{wrong advice, an average student, an average scan})
+\end{aligned}
+$$
+To complete our definition of the theoretical estimand, we have to clearly define out target population that consists of both persons and scans:
+With respect to persons, we are only interested in expert radiologists and medical students at German universities.
+With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
+Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system under study.
+
+While a theoretical estimand is defined outside of any statistical model and be impossible to observe directly, an empirical estimand is a quantity that can be estimated based on observed data.
+For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of brain scans, but the kind of AI-advice given for each brain scan is randomly assigned within participants.
+This random intervention allows to estimate our theoretical estimand, although in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each brain scan.
+
+For our example, each probability expression in our theoretical estimand can be modeled with a concrete GLMM.
+Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability and these estimates can be used to compute an estimate for each of the four probability contrasts.
+
 ### Simulate the data generating process
 
-The first step is to write code that simulates the (assumed) data generating process of the planned study.
+#### THEORY
+
+When the estimand has been defined, the next step in the research process is to write code that simulates the (assumed) data generating process of the planned study.
 This requires specifying a generative process for all predictor variables used in the final data analysis.
-While such assumptions can be quite challenging for observational studies and/or continuous predictor variables, this is less of a problem for experimental studies with categorical predictor variables.
+While such assumptions can be quite challenging for observational studies and/or continuous predictor variables, this is less of a problem for experimental studies with only categorical predictor variables.
 When all predictor variables have been simulated, one can use the structure of a suitable GLMM to simulate the dependent variable.
@@
 The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
-For this reasons, deciding on a specific set of assumptions can easily feel overwelming.
+For this reason, deciding on a specific set of assumptions can easily feel overwelming.
 However, we would argue that even a strongly simplified data generating process (e.g.
-including only a small number of interaction effect; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current state of psychological research where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.
+including only a small number of interaction effect; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current status quo in psychological research, where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.
+
+#### PRACTICE
+
+We simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
+The goal is to understand how AI advice influences medical decision-making.
+Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding. 
+To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
+This AI advice can be either correct (80% of cases) or incorrect (20%).
+In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
+After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
+This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
+With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
+In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
+The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
+
+#### Strategies on specifying the population parameters
+
+As discussed earlier, tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
+In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
+Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
+In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available.
+Our strategies to specify population parameters will require access to domain knowledge from domain experts.
+Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based sample size planning.
+This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available domain knowledge of domain experts.
+
+Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
+In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions, that can be adapted for our purposes.
+
+#### Insightful descriptive statistics
+
+For most GLMMs, the interpretation of model parameters if quite complicated which makes it difficult to specify plausible values for a specific application.
+However, the mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
+The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
+For example, domain experts might not be able to specify plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
+However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.
+which descriptive statistics they expect to observe (on average), when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
+The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
+Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
+
+#### Insightful model based quantities
+
+Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
+This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
+An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
+For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
+Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
+For example, the domain expert might deem it unplausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
+
+#### Iterative process with domain experts
+
+In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
+In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
+As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
+For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
+Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
+Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
+During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
 
@@
 In such more complicated scenarios, the final model specification might also depend on the next decision on specifying the theoretical estimand.
-
-### Specify the estimand
-
-This step requires a clear definition of which theoretical quantity is necessary to answer the specific research question of the study under planning.
-In an ideal world, this should be the first step in every research process.
-However in our experience, many psychological researchers are still not well trained in this task.
-We think that simulating data from a generic GLMM that is suited to describe (at least a rough approximation of) the assumed data generating process, can help to better reason about what information is actually needed to answer the specific research question, and how this information is actually represented in the context of the statistical model representing the data generating process of the planned study.
 
@@
 <!-- ### Sensitivity analysis -->
-
-## Strategies on specifying the population parameters
-
-As discussed earlier, tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
-In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
-Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
-In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available.
-Our strategies to specify population parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based sample size planning.
-This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available domain knowledge of domain experts.
-
-Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
-In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions, that can be adapted for our purposes.
-
-### Insightful descriptive statistics
-
-For most GLMMs, the interpretation of model parameters if quite complicated which makes it difficult to specify plausible values for a specific application.
-However, the mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
-The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
-For example, domain experts might not be able to specify plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
-However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.
-which descriptive statistics they expect to observe (on average), when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
-The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
-Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
-
-### Insightful model based quantities
-
-Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
-This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
-An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
-For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
-Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
-For example, the domain expert might deem it unplausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
-
-### Iterative process with domain experts
-
-In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
-In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
-As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
-For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
-Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
-Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
-During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
 
@@
 
+```{r}
+probs <- avg_predictions(fit, variables = list(
+  expert = 0:1, advice_present = 0:1, advice_correct = 0:1), 
+  re.form = NA, type = "response")
+
+probs
+
+probs %>% 
+  hypotheses(hypothesis = c(
+    "b8 - b2 = 0",
+    "b7 - b1 = 0",
+    "b2 - b4 = 0",
+    "b1 - b3 = 0"))
+```
+
 STILL MISSING

### 2024-03-27 docs/manuscript.Rmd
continue work on restructuring the manuscript

@@
 
-## Prior Knowledge
+# Prior Knowledge
 
@@
 
-## General concepts
-
-### Planning for statistical power or precision
+# General concepts
+
+## Planning for statistical power or precision
 
@@
 
-#### Planning for power
+### Planning for power
 
@@
 
-#### Planning for precision
+### Planning for precision
 
@@
 
-### Generalized linear mixed models (GLMMs)
+## Generalized linear mixed models (GLMMs)
 
@@
 
-## Simulation based sample size planning with GLMMs
+# Simulation based sample size planning with GLMMs
 
@@
 
-## When do we profit from tailored data simulation?
+# When do we profit from tailored data simulation?
 
@@
 
-### Complex study designs
+## Complex study designs
 
@@
 
-#### Missing data
+### Missing data
 
@@
 
-#### Data generating processes beyond standard GLMMs
+### Data generating processes beyond standard GLMMs
 
@@
 
-### Complex statistical hypotheses
+## Complex statistical hypotheses
 
@@
 
-#### Directed hypotheses and testing against values other than 0
+### Directed hypotheses and testing against values other than 0
 
@@
 
-#### Combined hypotheses
+### Combined hypotheses
 
@@
 
-### Planning for precision
+## Planning for precision
 
@@
 
-### No prior studies or pilot data
+## No prior studies or pilot data
 
@@
 
-## General steps in tailored simulation-based sample size planning
+# General steps in tailored simulation-based sample size planning
 
@@
 
-### Define the estimand
-
-#### THEORY
+## Define the estimand
+
+### THEORY
 
@@
 
-#### PRACTICE
+### PRACTICE
 
@@
 
-### Simulate the data generating process
+## Simulate the data generating process
+
+### Data simulation
 
@@
 We simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
-The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding. 
-To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-This AI advice can be either correct (80% of cases) or incorrect (20%).
-In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
-After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
+<!-- The goal is to understand how AI advice influences medical decision-making. -->
+Participants, radiologists (task experts; 1/4 of participants) and students/interns (non-task experts; 3/4), review a series of head computer tomography (CT) scans to assess the presence of a bleeding.
+To support their decision-making, an AI model provides initial diagnostic advice (2/3 of cases), which can be used as guidance by the participants.
+In the control condition (1/3), no AI advice is presented, meaning that the participants have to read the CT scan without any support.
+When AI advice is given, this advice can be either correct (4/5 of cases) or incorrect (1/5).
+The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across brain scans.
+After reviewing a CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
+<!-- With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
+-->
 In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
-The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
-
-#### Strategies on specifying the population parameters
-
-As discussed earlier, tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
-In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
-Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
-In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available.
-Our strategies to specify population parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based sample size planning.
-This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available domain knowledge of domain experts.
-
-Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
-In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions, that can be adapted for our purposes.
-
-#### Insightful descriptive statistics
-
-For most GLMMs, the interpretation of model parameters if quite complicated which makes it difficult to specify plausible values for a specific application.
-However, the mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
-The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
-For example, domain experts might not be able to specify plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
-However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.
-which descriptive statistics they expect to observe (on average), when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
-The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
-Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
-
-#### Insightful model based quantities
-
-Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
-This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
-An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
-For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
-Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
-For example, the domain expert might deem it unplausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
-
-#### Iterative process with domain experts
-
-In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
-In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
-As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
-For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
-Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
-Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
-During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
-
-### Estimate the statistical model
-
-At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset that will later be collected in the planned study.
-The next step is to specify how the statistical model shall be estimated in the actual study collected later.
-This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
-Note that this does not always mean that one will specify the same GLMM that was used when specifying the data generating process.
-On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
-On the other hand, using a more complex model (that perhaps even goes beyond a standard GLMM) for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
-In such more complicated scenarios, the final model specification might also depend on the next decision on specifying the theoretical estimand.
-
-### Compute the estimate
-
-When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
-This decision is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests with the marginaleffects R package, using the delta method).
-
-### Perform repeated simulations
-
-Conducting the previous steps enable the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
-The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
-On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter.
-When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.
-When planning for precision, the function should instead return the width of the confidence interval(s) of interest.
-Secondly, we must run this function repeatedly with the same sample and population parameters.
-Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computations, that is running simulations on multiple cores of the computer at the same time to reduce total computation time.
-Thirdly, the results of the repeated simulation must be collected and aggregated.
-When planning for power, we compute the relative frequency of a significant p-value across repeated simulations.
-When planning for precision, we compute the average width of the confidence interval.
-Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted statistical power or the expected width of the confidence interval.
-
-<!-- ### Sensitivity analysis -->
-
-# Case study
-
-<!-- overview of this section -->
-
-In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
-All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
-
-## Experimental study design
-
-<!-- describe the study idea and design -->
-In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
-The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding.
-To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-This AI advice can be either correct (80% of cases) or incorrect (20%).
-In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
-After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
-This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
-In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
-The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
-
-## Our specific GLMM
-
-<!-- deep dive into our specific example -->
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]: $$
+The goal of simulation-based sample size planning is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power or precision in the planned experiment.
+
+##### Our specific GLMM
+
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
+$$
 g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$ Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: $$
+$$
+Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: 
+$$
 E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
-$$ If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
+$$
+If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability:
+$$
 P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
-$$ In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.
-
-In such a setting, we model this probability as $$
-P_{si} = inverse\_logit(\eta_{si})
-$$ with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently $$
-logit(P_{si}) = \eta_{si}
-$$ with the logit link $g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})$.
+$$
+In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.
+
+In such a setting, we model this probability as
+$$
+P_{si} = \text{inverse_logit}(\eta_{si})
+$$
+with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
+$$
+\text{logit}(P_{si}) = \eta_{si}
+$$
+with the logit link $g(P_{si}) = \text{logit}(P_{si}) = \text{ln} (\frac{P_{si}}{1 - P_{si}})$.
 
@@
 
-In total, we assume the model $$
+In total, we assume the model
+$$
 \begin{aligned}
-logit[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
+\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
 &\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
@@
 \end{aligned}
-$$ or equivalently $$
+$$
+or equivalently
+$$
 \begin{aligned}
-P_{si} = inverse\_logit[&(\beta_0 + u_{0s} + u_{0i}) + \\
+P_{si} = \text{inverse_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
 &\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
@@
 \end{aligned}
-$$ with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
+$$
+with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
 
@@
 
-## Data simulation
+##### Simulation function in R
 
@@
 
+In the first six lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data. 
+In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).
+Thus, the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
+We add a between-subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
+The next two lines simulate that in $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
+Next, we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
+As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user.
+With all design variables done, we are ready to simulate our model equation outlined in the last section.
+The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects, and model parameters as assumed by our model.
+We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
+In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject makes an accurate diagnostic decision for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
+
+### Strategies on specifying the population parameters
+
+As discussed earlier, tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
+In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
+Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
+In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available.
+Our strategies to specify population parameters will require access to domain knowledge from domain experts.
+Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based sample size planning.
+This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available domain knowledge of domain experts.
+
+Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
+In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions, that can be adapted for our purposes.
+
+##### Insightful descriptive statistics
+
+For most GLMMs, the interpretation of model parameters if quite complicated which makes it difficult to specify plausible values for a specific application.
+However, the mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
+The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
+For example, domain experts might not be able to specify plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
+However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.
+which descriptive statistics they expect to observe (on average), when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
+The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
+Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
+
+##### Insightful model based quantities
+
+Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
+This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
+An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
+For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
+Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
+For example, the domain expert might deem it unplausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
+
+#### Iterative process with domain experts
+
+In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
+In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
+As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
+For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
+Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
+Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
+During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
+
+## Estimate the statistical model
+
+### THEORY
+
+At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset that will later be collected in the planned study.
+The next step is to specify how the statistical model shall be estimated in the actual study collected later.
+This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
+Note that this does not always mean that one will specify the same GLMM that was used when specifying the data generating process.
+On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
+On the other hand, using a more complex model (that perhaps even goes beyond a standard GLMM) for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
+In such more complicated scenarios, the final model specification might also depend on the next decision on specifying the theoretical estimand.
+
+### PRACTICE
+
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
+The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
+
+[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
+
+However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
+Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]
+
+[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
+
+In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
+We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
+However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.
+
+```{r, message=FALSE}
+library(tidyverse)
+set.seed(1)
+dat <- simulate(n_subjects = 100, n_items = 50)
+```
+
+The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:
+
+```{r, message=FALSE}
+library(lme4)
+f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
+  expert:advice_present + expert:advice_correct +
+  (1|subject) + (1|item)
+```
+
+The first two lines look similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`).
+The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
+The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
+
+In lme4, a GLMM is fitted with the `glmer` function.
+By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs.
+inaccurate (0) diagnosis.
+
+```{r}
+fit <- glmer(f, data = dat, family = "binomial")
+```
+
+## Compute the estimate
+
+### THEORY
+
+When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
+This decision is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests with the marginaleffects R package, using the delta method).
+
+### PRACTICE
+
+#### Model interpretation
+
+We can inspect the estimates for all model parameters with the `summary` command:
+
+```{r}
+summary(fit)
+```
+
+In the output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
+
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
+Then the model implied probability that such a person solves such an item accurately is given by:
+$$
+\begin{aligned}
+P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
+= inverse\_logit[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+\end{aligned}
+$$
+
+In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
+In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
+The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
+$$
+\begin{aligned}
+P(Y=1|expert = 0, advice\_present = 0, advice\_correct = 0, u_{0s} = 0, u_{0i} = 0) = \\
+= inverse\_logit[\beta_0]
+\end{aligned}
+$$
+
+Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
+Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
+
+With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects):
+
+(ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package.
+
+```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"}
+library(marginaleffects)
+plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"),
+  type = "response") + ylim(c(0.3, 1))
+```
+
+#### Hypothesis testing
+
+However, we need to think about the model parameters again when we want to test hypotheses that we have theoretically derived from some research question.
+Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.
+
+The `Fixed effects` table in the lme4 summary output also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$.
+However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.
+
+For our case study, imagine the following combined hypothesis: *We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
+
+This combined hypothesis leads to the following four separate null hypotheses to be tested: $$
+\begin{aligned}
+H_{01}:\ &\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0 \\
+H_{02}:\ &\beta_{a} + \beta_{c} \leq 0 \\
+H_{03}:\ &\beta_{a} + \beta_{ea} \geq 0 \\
+H_{04}:\ &\beta_{a} \geq 0
+\end{aligned}
+$$
+
+We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$: The first null hypothesis states that *an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.* This implies the following inequality for each subject $s$ and item $i$ $$
+inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}] \leq inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e]
+$$ which simplifies to $\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0$.
+
+We can specify and test hypotheses like these with the multcomp package [@hothornSimultaneousInferenceGeneral2008] as follows:
+
+```{r, message=FALSE}
+library(multcomp)
+null_hypotheses <- c(
+  "advice_present + advice_correct + expert:advice_present + 
+   expert:advice_correct <= 0",
+  "advice_present + advice_correct <= 0",
+  "-1 * (advice_present + expert:advice_present) <= 0",
+  "-1 * (advice_present) <= 0")
+glht <- glht(fit, linfct = null_hypotheses)
+summary(glht, test = univariate())$test$pvalues
+```
+
+Because all hypotheses tested simultaneously with the `glht` function must have the same direction, we flip the sign of inequalities three and four by multiplying them with $-1$.
+The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously [@hothornSimultaneousInferenceGeneral2008].
+However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
+In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
+Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command.
+With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value for hypothesis $H_{03}$ is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
+Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
+
+We can specify and test hypothesis like these with the marginaleffects [@R-marginaleffects] package.
+
+```{r}
+library(marginaleffects)
+
+hypotheses(fit, hypothesis = c(
+  "`advice_present` + `advice_correct` + `expert:advice_present` +
+  `expert:advice_correct` = 0",
+  "`advice_present` + `advice_correct` = 0",
+  "`advice_present` + `expert:advice_present` = 0",
+  "`advice_present` = 0"
+), equivalence = 0)
+
+hypotheses(fit, hypothesis = c(
+  "`advice_present` + `advice_correct` + `expert:advice_present` +
+  `expert:advice_correct` = 0",
+  "`advice_present` + `advice_correct` = 0",
+  "-1 * (`advice_present` + `expert:advice_present`) = 0",
+  "-1 * (`advice_present`) = 0"
+), equivalence = 0)
+```
+
+#### Interval estimation
+
+```{r}
+probs <- avg_predictions(fit, variables = list(
+  expert = 0:1, advice_present = 0:1, advice_correct = 0:1), 
+  re.form = NA, type = "response")
+
+probs
+
+probs %>% 
+  hypotheses(hypothesis = c(
+    "b8 - b2 = 0",
+    "b7 - b1 = 0",
+    "b2 - b4 = 0",
+    "b1 - b3 = 0"))
+```
+
+## Perform repeated simulations
+
+### THEORY
+
+Conducting the previous steps enable the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
+The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
+On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter.
+When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.
+When planning for precision, the function should instead return the width of the confidence interval(s) of interest.
+Secondly, we must run this function repeatedly with the same sample and population parameters.
+Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computations, that is running simulations on multiple cores of the computer at the same time to reduce total computation time.
+Thirdly, the results of the repeated simulation must be collected and aggregated.
+When planning for power, we compute the relative frequency of a significant p-value across repeated simulations.
+When planning for precision, we compute the average width of the confidence interval.
+Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted statistical power or the expected width of the confidence interval.
+
+### PRACTICE
+
+With all these considerations addressed, we are finally ready to perform a power analysis.
+Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
+
+```{r}
+sim_and_analyse <- function(
+  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + 
+    expert:advice_present + expert:advice_correct + (1|subject) + (1|item)", 
+  null_hypotheses = c("advice_present + advice_correct + 
+    expert:advice_present + expert:advice_correct <= 0",
+    "advice_present + advice_correct <= 0",
+    "-1 * (advice_present + expert:advice_present) <= 0",
+    "-1 * (advice_present) <= 0"), ...){
+  require(lme4)
+  require(multcomp)
+  # simulate data
+  dat <- simulate(...)
+  # fit model
+  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
+  # compute p-values
+  glht <- glht(model, linfct = null_hypotheses)
+  pvalues <- summary(glht, test = univariate())$test$pvalues
+  setNames(pvalues, paste0("p_H0", 1:length(null_hypotheses)))
+}
+```
+
+Power analysis can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
+Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
+First, we enable parallelization and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
+
+```{r, message=FALSE, eval=FALSE}
+library(future)
+plan("multisession", workers = 6)
+set.seed(2)
+```
+
+The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
+We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
+With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
+If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.
+
+```{r, message=FALSE, echo=FALSE, eval=FALSE}
+library(future)
+library(furrr)
+plan("multisession", workers = 6)
+set.seed(2)
+sim_design <- crossing(
+  rep = 1:300,
+  n_subjects = c(100, 150, 200, 250),
+  n_items = c(10, 30, 50, 70)
+) %>%
+  mutate(pvalues = future_pmap(., sim_and_analyse, 
+    .options = furrr_options(seed = TRUE))) %>%
+  unnest_wider(pvalues)
+saveRDS(sim_design, file = "results.rds")
+```
+
+```{r, message=FALSE, eval=FALSE}
+library(furrr)
+sim_design <- crossing(
+  rep = 1:300,
+  n_subjects = c(100, 150, 200, 250),
+  n_items = c(10, 30, 50, 70)
+) %>%
+  mutate(pvalues = future_pmap(., sim_and_analyse, 
+    .options = furrr_options(seed = TRUE))) %>%
+  unnest_wider(pvalues)
+```
+
+The result of the computation is a data frame that contains the p-values of all tested hypotheses for each simulated dataset.
+In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
+When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process.
+We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
+
+For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^3]
+
+[^3]: This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
+
+```{r, echo=FALSE}
+sim_design <- readRDS(file = "results.rds")
+```
+
+```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."}
+library(binom)
+alpha <- 0.05
+power <- sim_design %>%
+  group_by(n_subjects, n_items) %>% 
+  summarise(power = mean(p_H01 < alpha & p_H02 < alpha & 
+                           p_H03 < alpha & p_H04 < alpha), 
+    n_sig = sum(p_H01 < alpha & p_H02 < alpha & 
+                  p_H03 < alpha & p_H04 < alpha),
+    n = n(),
+    ci.lwr = binom.confint(n_sig, n, method = "wilson")$lower,
+    ci.upr = binom.confint(n_sig, n, method = "wilson")$upper, 
+    .groups = "drop")
+power %>%
+  mutate(across(c(n_subjects, n_items), factor)) %>%
+  ggplot(aes(n_subjects, n_items, fill = power)) +
+  geom_tile() +
+  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
+                                power, ci.lwr, ci.upr)), 
+    color = "white", size = 4) +
+  scale_fill_viridis_c(limits = c(0, 1)) +
+  xlab("number of subjects") + ylab("number of items")
+```
+
+<!-- How many participants/stimuli are needed for concrete example -->
+
+As should be the case, power estimates in Figure \@ref(fig:finalpwr) increase with both the number of subjects and the number of items.
+The confidence intervals indicate how precisely power was estimated by our simulation.
+Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
+In practice, data simulations are often run multiple times with adjusted combinations of sample sizes.
+When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of `n_subjects` and `n_items`.
+When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
+This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.
+
+Much has been written on the optimal amount of power to target in empirical research.
+The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints  [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
+When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
+For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation.
+This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice).
+If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items.
+However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule.
+The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
+Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
+
+<!-- ## Sensitivity analysis -->
+
+# THIS WAS THE OLD CASE STUDY CHAPTER
+
+<!-- overview of this section -->
+
+In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
+All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
+
+## Experimental study design
+
+<!-- describe the study idea and design -->
+In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
+The goal is to understand how AI advice influences medical decision-making.
+Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding.
+To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
+This AI advice can be either correct (80% of cases) or incorrect (20%).
+In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
+After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
+This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
+With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
+In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
+The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
+
+## Our specific GLMM
+
+<!-- deep dive into our specific example -->
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
+$$
+g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
+$$
+Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: 
+$$
+E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
+$$
+If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability:
+$$
+P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
+$$
+In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.
+
+In such a setting, we model this probability as
+$$
+P_{si} = inverse\_logit(\eta_{si})
+$$
+with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
+$$
+logit(P_{si}) = \eta_{si}
+$$
+with the logit link $g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})$.
+
+In our case study, the probability of making an accurate diagnostic decision is assumed to depend on the predictors:
+
+- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to assess item $i$ 
+- $advice\_correct_{si}$: whether this advice was correct (1) or not (0) 
+- $expert_s$: whether subject $s$ was a task expert (1) or not (0)
+
+and the random effects:
+
+-   $u_{0s}$: the deviation of subject $s$ from the average ability to solve an item (i.e., CT scan) with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
+-   $u_{0i}$: the deviation of item (i.e., CT scan) $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$
+
+In total, we assume the model
+$$
+\begin{aligned}
+logit[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
+&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
+\end{aligned}
+$$
+or equivalently
+$$
+\begin{aligned}
+P_{si} = inverse\_logit[&(\beta_0 + u_{0s} + u_{0i}) + \\
+&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+\end{aligned}
+$$
+with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
+
+In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
+To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.
+
+## Data simulation
+
+The following R function simulates a full dataset structured according to the design of our case study.
+The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.
+
+```{r}
+simulate <- function(n_subjects = 100, n_items = 50,
+  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
+  b_ea = 0.790, b_ec = -1.393,
+  sd_u0s = 0.5, sd_u0i = 0.5, ...){
+  require(dplyr)
+  require(faux)
+  # simulate design
+  dat <- add_random(subject = n_subjects, item = n_items) %>%
+    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
+    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
+    mutate(advice_correct = if_else(advice_present == 1L, 
+                                    rbinom(n(), 1L, prob = 0.8), 0L)) %>%
+    # add random effects
+    add_ranef("subject", u0s = sd_u0s) %>%
+    add_ranef("item", u0i = sd_u0i) %>%
+    # compute dependent variable
+    mutate(linpred = b_0 + u0i + u0s +
+        b_e * expert + b_a * advice_present + b_c * advice_correct +
+        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
+    mutate(y_prob = plogis(linpred)) %>%
+    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
+  dat
+}
+```
+
 In the first six lines of the function definition, we set some default parameter values (which we will explain in a later section) and load the packages we use to manipulate and simulate data. 
@@
 The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
-Then the model implied probability that such a person solves such an item accurately is given by: $$
+Then the model implied probability that such a person solves such an item accurately is given by:
+$$
 \begin{aligned}
@@
 In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented: $$
+The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
+$$
 \begin{aligned}
@@
 Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
-
-We can specify and test hypothesis like these with the marginaleffects [@R-marginaleffects] package.
-
-```{r}
-library(marginaleffects)
-
-hypotheses(fit, hypothesis = c(
-  "`advice_present` + `advice_correct` + `expert:advice_present` +
-  `expert:advice_correct` = 0",
-  "`advice_present` + `advice_correct` = 0",
-  "`advice_present` + `expert:advice_present` = 0",
-  "`advice_present` = 0"
-), equivalence = 0)
-
-hypotheses(fit, hypothesis = c(
-  "`advice_present` + `advice_correct` + `expert:advice_present` +
-  `expert:advice_correct` = 0",
-  "`advice_present` + `advice_correct` = 0",
-  "-1 * (`advice_present` + `expert:advice_present`) = 0",
-  "-1 * (`advice_present`) = 0"
-), equivalence = 0)
-```
-
-## Interval estimation
-
-```{r}
-probs <- avg_predictions(fit, variables = list(
-  expert = 0:1, advice_present = 0:1, advice_correct = 0:1), 
-  re.form = NA, type = "response")
-
-probs
-
-probs %>% 
-  hypotheses(hypothesis = c(
-    "b8 - b2 = 0",
-    "b7 - b1 = 0",
-    "b2 - b4 = 0",
-    "b1 - b3 = 0"))
-```
-
-STILL MISSING
 

### 2024-03-27 docs/manuscript.Rmd
add susannes paragraph on the iterative exchange with domain experts

@@
 
+##### PRACTICE
+
+All parameter values in our present case study have been determined from results from related prior work and repeated discussions with our affiliated domain experts in Radiology to validate our assumptions.
+Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for ICH.
+Existing studies indicate that radiologists typically demonstrate high accuracies, often exceeding or hovering around 90%, while interns have been shown to perform below 80%, and medical students fall even shorter.
+For simplicity, we assumed plausible probability values of .90 for experts and .70 for non-experts, respectively.
+Our experts confirmed that these values are realistic baselines for reviewing diverse head CT images without AI assistance.
+Subsequently, we consulted several published papers investigating the effect of correct and incorrect advice on decision-making performance in other settings.
+From their findings, we inferred that both experts and non-experts would benefit from correct and suffer losses from incorrect advice.
+However, the magnitude of these effects would be substantially greater for non-experts, given their demonstrated reliance on advice compared to experts.
+We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
+For our simulation, we used the probabilities of average participants to solve an average case, as shown in Table 1.
+
 ## Estimate the statistical model
@@
 library(ggdist)
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.5, sd_u0i = 0.5)
+dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
 dat %>% 
@@
   scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
+```{r}
+library(tidyverse)
+library(ggdist)
+dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>%
+  group_by(condition) %>%
+  summarize(freq_y_1 = sum(y_bin) / n())
 ```

### 2024-03-27 docs/manuscript.Rmd
work on the "compute the estimate" section

@@
 
-For our example, each probability expression in our theoretical estimand can be modeled with a concrete GLMM.
+For our example, each probability expression in our theoretical estimand can be modeled with a GLMM.
 Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability and these estimates can be used to compute an estimate for each of the four probability contrasts.
@@
 We will discuss strategies for how these values can be obtained later.
-Because one has full control over the data generating process in a tailored simulation, it is simple to also model highly specific aspects of the planned study, like data missing by design or assuming that subjects dropout (i.e.
+Because one has full control over the data generating process in a tailored simulation, it is simple to also model highly specific aspects of the planned study, like data missing by design or assuming that subjects drop out (i.e.
 deleting their data from the simulated dataset) based on other variables in the simulated dataset.
@@
 
-We simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
+In our case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
 <!-- The goal is to understand how AI advice influences medical decision-making. -->
@@
 When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
-This decision is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests with the marginaleffects R package, using the delta method).
+This decision is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests and confidence intervals with the marginaleffects R package using the delta method).
 
@@
 
-#### Model interpretation
-
-We can inspect the estimates for all model parameters with the `summary` command:
+For our case study, we investigate the following research question:
+*We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
+
+In the earlier section on our theoretical estimand we translated this verbal description into four probability statements that are specified outside of any specific statistical model.
+The next translation step is to specify how these probabilites can be estimated within the framework of our concrete GLMM.
+
+The first of the four statements was:
+
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{correct advice, an average expert, an average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan})
+\end{aligned}
+$$
+In the context of our GLMM, the corresponding model based quantity is:
+
+$$
+\begin{aligned}
+& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
+& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
+\end{aligned}
+$$
+Plugging in the model equation of the GLMM produces an equation on how to compute this contrast when all model parameters are known.
+
+<!-- $$ -->
+<!-- \begin{aligned} -->
+<!-- & \text{inverse_logit}[\beta_0 + \beta_a \cdot 1 + \beta_c \cdot 1 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 1 + \beta_{ec} \cdot 1 \cdot 1]\\ -->
+<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_a \cdot 0 + \beta_c \cdot 0 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 0 + \beta_{ec} \cdot 1 \cdot 0] = \\ -->
+<!-- & = \text{inverse_logit}[\beta_0 + \beta_a  + \beta_c + \beta_e + \beta_{ea} + \beta_{ec} ]\\ -->
+<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_e ] \\ -->
+<!-- \end{aligned} -->
+<!-- $$ -->
+
+When model parameters are not known, the corresponding parameter estimates are plugged into the equation to compute an estimate of the contrast instead.
+
+In general, we can inspect the estimates for all model parameters with the `summary` command:
 
@@
 
-In the output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
-
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
-Then the model implied probability that such a person solves such an item accurately is given by:
+In the lme4 output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
+
+Unfortunately, there is no single model parameter that corresponds with our contrast of interest.
+the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+When we are interested in a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$), the model implied probability that such a person solves such an item accurately is given by:
 $$
@@
 P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
-= inverse\_logit[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+= \text{inverse_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
 &\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
@@
 $$
-
 In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
@@
 \begin{aligned}
-P(Y=1|expert = 0, advice\_present = 0, advice\_correct = 0, u_{0s} = 0, u_{0i} = 0) = \\
-= inverse\_logit[\beta_0]
-\end{aligned}
-$$
-
-Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
-
-With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects):
-
-(ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package.
-
-```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"}
-library(marginaleffects)
-plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"),
-  type = "response") + ylim(c(0.3, 1))
-```
-
-#### Hypothesis testing
-
-However, we need to think about the model parameters again when we want to test hypotheses that we have theoretically derived from some research question.
-Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.
-
-The `Fixed effects` table in the lme4 summary output also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$.
-However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.
-
-For our case study, imagine the following combined hypothesis: *We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
-
-This combined hypothesis leads to the following four separate null hypotheses to be tested: $$
-\begin{aligned}
-H_{01}:\ &\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0 \\
-H_{02}:\ &\beta_{a} + \beta_{c} \leq 0 \\
-H_{03}:\ &\beta_{a} + \beta_{ea} \geq 0 \\
-H_{04}:\ &\beta_{a} \geq 0
-\end{aligned}
-$$
-
-We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$: The first null hypothesis states that *an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.* This implies the following inequality for each subject $s$ and item $i$ $$
-inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}] \leq inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e]
-$$ which simplifies to $\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0$.
-
-We can specify and test hypotheses like these with the multcomp package [@hothornSimultaneousInferenceGeneral2008] as follows:
-
-```{r, message=FALSE}
-library(multcomp)
-null_hypotheses <- c(
-  "advice_present + advice_correct + expert:advice_present + 
-   expert:advice_correct <= 0",
-  "advice_present + advice_correct <= 0",
-  "-1 * (advice_present + expert:advice_present) <= 0",
-  "-1 * (advice_present) <= 0")
-glht <- glht(fit, linfct = null_hypotheses)
-summary(glht, test = univariate())$test$pvalues
-```
-
-Because all hypotheses tested simultaneously with the `glht` function must have the same direction, we flip the sign of inequalities three and four by multiplying them with $-1$.
-The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously [@hothornSimultaneousInferenceGeneral2008].
-However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
-In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
-Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command.
-With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value for hypothesis $H_{03}$ is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
-Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
-
-We can specify and test hypothesis like these with the marginaleffects [@R-marginaleffects] package.
+P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
+= \text{inverse_logit}[\beta_0]
+\end{aligned}
+$$
+One could use such reasoning, to determine the exact formula needed to compute our contrast of interest.
+However, this would be very tedious and we can use R to compute this contrast for us without doing the math.
+
+We can use the predict function of the lme4 package to compute the predicted probability for a correct diagnosis based on our model, plugin the two sets of predictor values, and compute the difference between the two estimated probabilities.
 
 ```{r}
-library(marginaleffects)
-
-hypotheses(fit, hypothesis = c(
-  "`advice_present` + `advice_correct` + `expert:advice_present` +
-  `expert:advice_correct` = 0",
-  "`advice_present` + `advice_correct` = 0",
-  "`advice_present` + `expert:advice_present` = 0",
-  "`advice_present` = 0"
-), equivalence = 0)
-
-hypotheses(fit, hypothesis = c(
-  "`advice_present` + `advice_correct` + `expert:advice_present` +
-  `expert:advice_correct` = 0",
-  "`advice_present` + `advice_correct` = 0",
-  "-1 * (`advice_present` + `expert:advice_present`) = 0",
-  "-1 * (`advice_present`) = 0"
-), equivalence = 0)
-```
-
-#### Interval estimation
+pred_df <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), expert = c(1, 1))
+pred_df
+preds <- predict(fit, newdata = pred_df, type = "response", re.form = NA)
+preds
+preds[1] - preds[2]
+```
+
+The argument `type = "response"` specifies that predictions are made on the probability scale, while `re.form = NA` sets all random effects to 0.
+
+We could use this method to compute point estimates for all four contrast that are part of our estimand.
+However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute hypothesis tests or confidence interval for our point estimates.
+
+The marginaleffects package [@R-marginaleffects] is a very flexible package to compute hypothesis tests and confidence intervals for contrasts with a variety of statistical models, including GLMM estimated with lme4.
+
+First, we compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
 
 ```{r}
-probs <- avg_predictions(fit, variables = list(
-  expert = 0:1, advice_present = 0:1, advice_correct = 0:1), 
-  re.form = NA, type = "response")
-
+pred_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
+probs <- predictions(fit, newdata = pred_df, type = "response", re.form = NA)
 probs
-
+```
+
+Note that the result also contains the two missing by design conditions that are never observed in the actual study ($advice\_present = 0, advice\_correct = 1, expert = 1$ and $advice\_present = 0, advice\_correct = 1, expert = 0$), and we should never interpret those.
+
+Next, we use these predictions to compute the four specific contrasts that are part of our estimand.
+For this we must specify which rows in `probs` have to be subtracted from each other.
+We will use the `hypotheses` function to compute our four contrasts of interest together with hypothesis tests and confidence intervals.
+We use the default inference options of the marginaleffects package that compute hypothesis tests and confidence intervals based on the approximate delta method.
+
+```{r}
 probs %>% 
   hypotheses(hypothesis = c(
-    "b8 - b2 = 0",
-    "b7 - b1 = 0",
-    "b2 - b4 = 0",
-    "b1 - b3 = 0"))
-```
+    "b8 = b2",
+    "b2 = b6",
+    "b7 = b1",
+    "b1 = b5"),
+    equivalence = c(0, 0))
+```
+
+The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2.
+The argument `equivalence = c(0, 0)` can be used to compute one-sided p-values testing whether the constrast in the population is smaller than 0 (`p (NonSub)` column) or greater than zero (`p (NonInf)` column).
+We computed the contrasts in a way that we expect positive values for all four of them.
+
+#### Hypothesis testing
+
+If we are interested in hypothesis testing in our case study, we test a combined null hypothesis that consists of four separate null hypotheses:
+
+$$
+\begin{aligned}
+H_{01}:\ & P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
+& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
+H_{02}:\ &P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
+& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
+H_{03}:\ &P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
+& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
+H_{04}:\ & P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
+& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
+\end{aligned}
+$$
+The combined null hypothesis should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
+In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
+
+With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value in the `p (NonInf)` column for the third hypothesis is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
+Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
+
+#### Interval estimation
+
+If we are interested in parameter estimation in our case study, we focus on the two-sided confidence intervals of the four contrasts of interest.
+With a standard confidence level of $1 - \alpha = 0.95$, plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
+Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts.
+The smallest width is observed for the first contrast where both underlying probabilities are close to 1.
+The largest width is observed for the fourth contrast, where both underlying probabilities are closer to 0.5.
+
+<!-- Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
+Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
+-->
+
+<!-- With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects): -->
+
+<!-- (ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package. -->
+
+<!-- ```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"} -->
+<!-- library(marginaleffects) -->
+<!-- plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"), -->
+<!--   type = "response") + ylim(c(0.3, 1)) -->
+<!-- ``` -->
 

### 2024-04-08 docs/manuscript.Rmd
work on power and precision results

@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, eval = FALSE, fig.keep = "high", dev = "jpeg")
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, fig.keep = "high", dev = "jpeg")
 ```
@@
 The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.
+
+```{r, echo = FALSE}
+b_0 = 0.847
+b_e = 1.350
+b_a = -1.253
+b_c = 2.603
+b_ea = 0.790
+b_ec = -1.393
+```
 
@@
 Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts.
-The smallest width is observed for the first contrast where both underlying probabilities are close to 1.
-The largest width is observed for the fourth contrast, where both underlying probabilities are closer to 0.5.
+The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice) where both underlying probabilities are close to 1.
+The largest width is observed for the fourth contrast (non-expert with wrong advice vs. non-expert without advice), where both underlying probabilities are closer to 0.5.
 
@@
 
-With all these considerations addressed, we are finally ready to perform a power analysis.
-Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
+With all these considerations addressed, we are finally ready to run our simulation-based sample size planning analyses to plan for power and precision.
+Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and confidence intervals) in a single command.
 
@@
     expert:advice_present + expert:advice_correct + (1|subject) + (1|item)",
-  null_hypotheses = c("advice_present + advice_correct + 
-    expert:advice_present + expert:advice_correct <= 0",
-    "advice_present + advice_correct <= 0",
-    "-1 * (advice_present + expert:advice_present) <= 0",
-    "-1 * (advice_present) <= 0"), ...){
+  contrasts = c("b8 = b2", "b2 = b6", "b7 = b1", "b1 = b5"), ...){
   require(lme4)
-  require(multcomp)
+  require(marginaleffects)
+  require(tidyr)
   # simulate data
@@
   model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
-  # compute p-values
-  glht <- glht(model, linfct = null_hypotheses)
-  pvalues <- summary(glht, test = univariate())$test$pvalues
-  setNames(pvalues, paste0("p_H0", 1:length(null_hypotheses)))
+  # compute contrasts
+  contr_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
+  predictions(model, newdata = contr_df, type = "response", re.form = NA) %>%
+    hypotheses(hypothesis = contrasts, equivalence = c(0, 0))
 }
@@
 
-Power analysis can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
+Simulation-based smaple size planning can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
 Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
@@
 set.seed(2)
-sim_design <- crossing(
-  rep = 1:300,
+sim_result <- crossing(
+  rep = 1:100,
   n_subjects = c(100, 150, 200, 250),
@@
 ) %>%
-  mutate(pvalues = future_pmap(., sim_and_analyse, 
+  mutate(res = future_pmap(., sim_and_analyse, 
     .options = furrr_options(seed = TRUE))) %>%
-  unnest_wider(pvalues)
-saveRDS(sim_design, file = "results.rds")
+  unnest(col = res)
+saveRDS(sim_result, file = "results.rds")
 ```
@@
 library(furrr)
-sim_design <- crossing(
-  rep = 1:300,
+sim_result <- crossing(
+  rep = 1:5,
   n_subjects = c(100, 150, 200, 250),
@@
 ) %>%
-  mutate(pvalues = future_pmap(., sim_and_analyse, 
+  mutate(res = future_pmap(., sim_and_analyse, 
     .options = furrr_options(seed = TRUE))) %>%
-  unnest_wider(pvalues)
-```
-
-The result of the computation is a data frame that contains the p-values of all tested hypotheses for each simulated dataset.
+  unnest(col = res)
+```
+
+The result of the computation is a data frame that contains the p-values and confidence intervals of all specified contrasts for each simulated dataset.
 In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
@@
 
+#### Power results
+
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
@@
 ```{r, echo=FALSE}
-sim_design <- readRDS(file = "results.rds")
+sim_result <- readRDS(file = "results.rds")
 ```
@@
 alpha <- 0.05
-power <- sim_design %>%
+power <- sim_result %>%
+  pivot_wider(names_from = term, names_sep = "_", 
+    values_from = estimate:p.value.equiv) %>%
   group_by(n_subjects, n_items) %>% 
-  summarise(power = mean(p_H01 < alpha & p_H02 < alpha & 
-                           p_H03 < alpha & p_H04 < alpha), 
-    n_sig = sum(p_H01 < alpha & p_H02 < alpha & 
-                  p_H03 < alpha & p_H04 < alpha),
+  summarise(
+    power = mean(`p.value.noninf_b1=b5` < alpha & `p.value.noninf_b8=b2` < alpha &
+      `p.value.noninf_b2=b6` < alpha & `p.value.noninf_b7=b1` < alpha), 
+    n_sig = sum(`p.value.noninf_b1=b5` < alpha & `p.value.noninf_b8=b2` < alpha &
+      `p.value.noninf_b2=b6` < alpha & `p.value.noninf_b7=b1` < alpha),
     n = n(),
@@
 Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
+
+#### Precision results
+
+When planning for precision, one could monitor the width of all four confidence intervals at the same time.
+However, because the confidence intervals of the four contrasts strongly differ in width, it is not trivial to decide which width one should target when deciding on the appropriate sample size.
+In contrast to planning for power, there are no common standards on how to specify the targeted precision.
+For our example, we use a simple heuristic but we strongly encourage readers to think about better alternatives that are appropriate in their own applications.
+
+Our simulations show that the smallest confidence interval can be expected for the first contrast (expert with correct advice vs. expert without advice).
+The true contrast in probability for an average expert and an average item in this condition is `plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec) - plogis(b_0 + b_e) =` $`r round(plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec) - plogis(b_0 + b_e), 3)`$.
+We want the width of this confidence interval to be smaller than 0.1.
+This would mean that if the point estimate happens to be close to the true value, the plausible values inside of a 95% confidence interval would all be positive.
+
+Thus in our example, precision is defined as the (long-run) average width of a 95% confidence interval for the probability contrast between experts with correct advice and experts without advice.
+Of course, lower width implies better precision.
+Based on our simulation outcomes, we compute the precision estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.
+
+```{r finalprecision, fig.cap="Simulation-based precision estimates (expected width of confidence intervals) including 95% confidence interval of the case study for different numbers of subjects and items, based on a confidence level of 0.95."}
+precision <- sim_result %>%
+  pivot_wider(names_from = term, names_sep = "_", 
+    values_from = estimate:p.value.equiv) %>%
+  group_by(n_subjects, n_items) %>%
+  mutate(width = `conf.high_b8=b2` - `conf.low_b8=b2`) %>%
+  summarise(precision = mean(width),
+    ci.lwr = t.test(width)$conf.int[1],
+    ci.upr = t.test(width)$conf.int[2], 
+    .groups = "drop")
+precision %>%
+  mutate(across(c(n_subjects, n_items), factor)) %>%
+  ggplot(aes(n_subjects, n_items, fill = precision)) +
+  geom_tile() +
+  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
+                                precision, ci.lwr, ci.upr)), 
+    color = "white", size = 6) +
+  scale_fill_viridis_c(limits = c(0, 0.3), direction = -1) + guides(fill = guide_legend(reverse=FALSE))
+```
+
+As should be the case, precision estimates in Figure \@ref(fig:finalprecision) increase  (i.e., average width of confidence interval decreases) with both the number of subjects and the number of items.
+Applying our simple heuristic of targeting an expected width smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.
+We could either choose 100 subjects and 30 items or 200 subjects and 10 items.
+
+Note that our simple heuristic for determining sample size in the planning for precision scenario was quite liberal.
+This is reflected by the result that we would need a smaller sample size than in the planning for power scenario.
+With a more conservative precision target, the result is generally the opposite: As a rule, precise parameter estimates usually require bigger samples than null hypothesis testing.
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, eval=FALSE}
 sim_design <- readRDS(file = "results.rds")
@@
 
-```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."}
+```{r finpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05.", eval=FALSE}
 library(binom)
@@
 
-As should be the case, power estimates in Figure \@ref(fig:finalpwr) increase with both the number of subjects and the number of items.
+As should be the case, power estimates in Figure \@ref(fig:finpwr) increase with both the number of subjects and the number of items.
 The confidence intervals indicate how precisely power was estimated by our simulation.

### 2024-04-08 docs/manuscript.Rmd
fix escape _ in \text

@@
 $$
-P_{si} = \text{inverse_logit}(\eta_{si})
+P_{si} = \text{inverse\_logit}(\eta_{si})
 $$
@@
 \begin{aligned}
-P_{si} = \text{inverse_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
+P_{si} = \text{inverse\_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
 &\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
@@
 <!-- \begin{aligned} -->
-<!-- & \text{inverse_logit}[\beta_0 + \beta_a \cdot 1 + \beta_c \cdot 1 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 1 + \beta_{ec} \cdot 1 \cdot 1]\\ -->
+<!-- & \text{omv}[\beta_0 + \beta_a \cdot 1 + \beta_c \cdot 1 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 1 + \beta_{ec} \cdot 1 \cdot 1]\\ -->
 <!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_a \cdot 0 + \beta_c \cdot 0 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 0 + \beta_{ec} \cdot 1 \cdot 0] = \\ -->
@@
 P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
-= \text{inverse_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
 &\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
@@
 P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
-= \text{inverse_logit}[\beta_0]
+= \text{inverse\_logit}[\beta_0]
 \end{aligned}

### 2024-04-08 docs/manuscript.Rmd
work on new practice sections on how to specify population parameters

@@
 
-### Strategies on specifying the population parameters
+## Strategies on specifying the population parameters
+
+### THEORY
 
@@
 
-##### Insightful descriptive statistics
-
-For most GLMMs, the interpretation of model parameters if quite complicated which makes it difficult to specify plausible values for a specific application.
+### PRACTICE
+
+When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
+
+Ideally, one would rely on meta-analytic results or conclusive data from pilot studies.
+However, these are sometimes not readily available.
+All parameter values in our present case study have been determined based on results from related study designs from the literature.
+Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions seems plausible.
+
+We now outline the strategies we used to determine plausible parameter values:
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret in isolation because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
+Then the model implied probability that such a person solves such an item accurately is given by:
+
+$$
+\begin{aligned}
+P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
+= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+\end{aligned}
+$$
+In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
+In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
+The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
+$$
+\begin{aligned}
+P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
+= \text{inverse\_logit}[\beta_0]
+\end{aligned}
+$$
+We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (that a subject with average ability solves an item with average difficulty) for each experimental condition.
+
+```{r probtable, echo=FALSE}
+probtable <- data.frame(
+  col1 = 
+    c("no advice, no expert", "no advice, expert", "false advice, no expert", 
+      "false advice, expert", "correct advice, no expert", "correct advice, expert"),
+  col2 =
+    c("0.70", "0.90", "0.40", "0.85", "0.90", "0.95"),
+  col3 = 
+    c("$logit(0.70) = \\beta_0$",
+      "$logit(0.90) = \\beta_0 + \\beta_e$", 
+      "$logit(0.40) = \\beta_0 + \\beta_a$",
+      "$logit(0.85) = \\beta_0 + \\beta_e + \\beta_{a} + \\beta_{ea}$",
+      "$logit(0.90) = \\beta_0 + \\beta_a + \\beta_c$",
+      "$logit(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
+apa_table(
+  probtable
+  , col.names = c("Experimental condition",
+    "$P(Y=1|\\mathbf{X=x}, \\mathbf{U} = \\mathbf{0})$",
+    "Implied equation")
+  , caption = "Assumed probabilities that an average subject solves an average item in each experimental condition."
+  , note = "Implied equations are derived based on the model equations and setting all random intercept terms to 0."
+  , escape = FALSE
+)
+```
+
+Table \@ref(tab:probtable) shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
+The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:
+
+```{r}
+b_0 <- qlogis(0.7)
+b_e <- qlogis(0.9) - b_0
+b_a <- qlogis(0.4) - b_0
+b_ea <- qlogis(0.85) - b_0 - b_e - b_a
+b_c <- qlogis(0.9) - b_0 - b_a
+b_ec <- qlogis(0.95) - b_0 - b_e - b_a - b_c - b_ea
+c(b_0 = b_0, b_e = b_e, b_a = b_a, b_c = b_c, b_ea = b_ea, b_ec = b_ec)
+```
+
+It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g. 
+$$
+\begin{aligned}
+P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
+= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
+\end{aligned}
+$$
+
+```{r}
+plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec)
+```
+
+### Insightful descriptive statistics
+
+#### THEORY
+
+For most GLMMs, the interpretation of model parameters are quite complicated which makes it difficult to specify plausible values for a specific application.
 However, the mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
@@
 
-##### Insightful model based quantities
+#### PRACTICE
+
+In the last section, we showed how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
+Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
+The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
+For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
+
+However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
+How much this estimated probability differs from the model implied probability of an average expert and an average items depends on the standard deviation of the random subject and random item intercepts (the two quantities are only equal if both standard deviations would be zero).
+We can use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
+
+We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model implied true values) and simply compute the relative frequency of correct diagnoses for each experimental conditions.
+
+```{r}
+library(tidyverse)
+set.seed(1)
+dat <- simulate(n_subjects = 3000, n_items = 3000,
+  sd_u0s = 0.5, sd_u0i = 0.5)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  group_by(condition) %>%
+  summarize(relative_frequency = sum(y_bin) / n())
+```
+
+We can use these descriptive statistics to judge whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
+
+However, in our example they are not informative enough to determine a final set of plausible parameter values.
+For this reason, we will additionally look at insightful model based quantities.
+
+### Insightful model based quantities
+
+#### THEORY
 
@@
 
-#### Iterative process with domain experts
+#### PRACTICE
+
+The inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
+One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
+However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
+Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
+
+In the last section, we simulated a large simulated dataset and computed a descriptive statistic, the relative frequencies of correct diagnoses, for each experimental condition.
+We will now use the model implied probability of each simulated data entry (y_prob) to visualize the whole model based implied marginal distribution of correct diagnoses for each condition.
+
+(ref:margdist1-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions.
+
+```{r margdist1, fig.cap="(ref:margdist1-caption)"}
+library(tidyverse)
+library(ggdist)
+dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  ggplot(aes(x = y_prob, y = condition)) +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
+```{r}
+library(tidyverse)
+library(ggdist)
+dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>%
+  group_by(condition) %>%
+  summarize(freq_y_1 = sum(y_bin) / n())
+```
+
+Figure \@ref(fig:margdist1) shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
+We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
+This difference increases with the variability of the random effects.
+
+We can use plots like the one above as a useful tool to decide whether the specified standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
+
+In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$). This gives us a better way to see the variability between persons.
+
+(ref:margdist2-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01.
+
+```{r margdist2, echo=FALSE, fig.cap="(ref:margdist2-caption)"}
+dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0i = 0.01)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  ggplot(aes(x = y_prob, y = condition)) +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
+As an example, Figure \@ref(fig:margdist2) reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
+In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
+Patterns like these should be considered carefully and discussed with the domain experts.
+Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
+
+We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
+
+```{r, include=FALSE}
+dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  ggplot(aes(x = y_prob, y = condition)) +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
+The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
+For Figure \@ref(fig:margdist3), we have set $\sigma_S = 3$ and $\sigma_I = 3$.
+This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
+However, these high standard deviations do not account for the inherent variability and complexity of human performance.
+For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
+
+(ref:margdist3-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3.
+
+```{r margdist3, echo=FALSE, fig.cap="(ref:margdist3-caption)"}
+dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 3, sd_u0i = 3)
+dat %>% 
+  mutate(condition = fct_cross(
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+  mutate(condition = fct_recode(condition,
+    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  ggplot(aes(x = y_prob, y = condition)) +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
+```
+
+### Iterative process with domain experts
 
@@
 
-```{r probtable, echo=FALSE}
+```{r, echo=FALSE}
 probtable <- data.frame(
@@
 
-```{r margdist1, fig.cap="(ref:margdist1-caption)"}
+```{r}
 library(tidyverse)
@@
 
-```{r margdist2, echo=FALSE, fig.cap="(ref:margdist2-caption)"}
+```{r}
 dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0i = 0.01)
@@
 
-```{r margdist3, echo=FALSE, fig.cap="(ref:margdist3-caption)"}
+```{r}
 dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 3, sd_u0i = 3)

### 2024-04-09 docs/manuscript.Rmd
iterate on first part of case study sections

@@
 
-Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions that we will outline in this section.
+Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions that we will outline in the remainder of our tutorial.
+Will introduce each step in a theoretical section, followed by the practical application based on a case study.
 
@@
 Unfortunately, many psychological researchers are still not well trained in this task.
-The theoretical estimand is a unit specific quantity defined outside of any statistical model.
+The theoretical estimand is a unit specific quantity defined outside of any statistical model [@lundbergWhatYourEstimand2021].
 Depending on the research and whether the research question is causal, this quantity might be counterfactual and not directly observable.
@@
 
-When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reexpresses the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
-For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantities that can be estimated with a regression model.
-Note however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression.
+When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reformulate the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
+For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantities within the scope that can be estimated with a regression model.
+Note however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression (CITATION).
 
 When the empirical estimand has been specified, the last step is to clearly define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
-When a regression model is used for estimation, this includes decisions like which statistical framework and package is used to fit the model, and how to compute hypothesis tests or confidence intervals.
-
-@lundbergWhatYourEstimand2021
+When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique, and software package is used to fit the model, and how to compute hypothesis tests or confidence intervals.
 
@@
 The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography.
-In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans require to make the final diagnostic decision.
+In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans that are required to make the final diagnostic decision.
 The research goal is to validate the effectiveness of the AI-enabled advice.
-We consider the AI-enable as effective, if the following pattern holds, which we will here describe verbally:
+We consider the AI-enabled advice as effective, if the following pattern holds, which we will first describe verbally:
 
@@
 
-I becomes clear that our theoretical estimand is be a causal quantity, based on individual quantities define outside of any statistical model.
-To explain why, we focus on only on part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists.
+I becomes clear that our theoretical estimand is be a causal quantity, based on several individual quantities that are defined outside of any statistical model [@lundbergWhatYourEstimand2021].
+To elaborate further, we focus on only one part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists.
 For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice were presented and the probability that a correct diagnosis is made if no AI-advice were presented.
@@
 \begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, a specific expert, a specific scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, a specific expert, a specific scan})
-\end{aligned}
-$$
-However, our verbal description is not precise enough yet, because the above quantity exists and may differ for each combination of expert and scan.
+& P(\text{correct diagnosis} | \text{correct advice, specific expert, specific scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, specific expert, specific scan})
+\end{aligned}
+$$
+However, our verbal description is not precise enough yet, because the above quantity may differ for each combination of expert and scan.
 We must specify how we want to aggregate the quantity across different experts and scans.
-If we would compute the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts.
-Although often debated whether this is a good idea, psychological research often focuses on a different quantity.
+If we would compute the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts (CITATION).
+Although often debated whether this is a good idea, psychological research often focuses on a different quantity:
 Instead of averaging over experts and scans, we focus in a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
@@
 \begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, an average expert, an average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan})
-\end{aligned}
-$$
-In total, our theoretical estimand consists of four expressions like the one above, the remaining three are:
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, an average expert, an average scan})
-\end{aligned}
-$$
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, an average student, an average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, an average student, an average scan})
-\end{aligned}
-$$
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{no advice, an average student, an average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, an average student, an average scan})
+& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
+\end{aligned}
+$$
+In total, our theoretical estimand consists of four such expressions, the remaining three are:
+
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan})
+\end{aligned}
+$$
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{correct advice, average student, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, average student, average scan})
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan})
 \end{aligned}
@@
 
-While a theoretical estimand is defined outside of any statistical model and be impossible to observe directly, an empirical estimand is a quantity that can be estimated based on observed data.
+While a theoretical estimand is defined outside of any statistical model and can be impossible to be observe directly, an empirical estimand is a quantity that can be estimated based on observed data.
 For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of brain scans, but the kind of AI-advice given for each brain scan is randomly assigned within participants.
@@
 To simulate the GLMM, one requires plausible values for all model parameters.
-We will discuss strategies for how these values can be obtained later.
+We will discuss strategies on how these values can be obtained later.
 Because one has full control over the data generating process in a tailored simulation, it is simple to also model highly specific aspects of the planned study, like data missing by design or assuming that subjects drop out (i.e.
@@
 However, we would argue that even a strongly simplified data generating process (e.g.
-including only a small number of interaction effect; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current status quo in psychological research, where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.
+including only a small number of interaction effects; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current status quo in psychological research, where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.
 
@@
 <!-- The goal is to understand how AI advice influences medical decision-making. -->
-Participants, radiologists (task experts; 1/4 of participants) and students/interns (non-task experts; 3/4), review a series of head computer tomography (CT) scans to assess the presence of a bleeding.
-To support their decision-making, an AI model provides initial diagnostic advice (2/3 of cases), which can be used as guidance by the participants.
-In the control condition (1/3), no AI advice is presented, meaning that the participants have to read the CT scan without any support.
-When AI advice is given, this advice can be either correct (4/5 of cases) or incorrect (1/5).
+Participants, radiologists (task experts) and students/interns (non experts), review a series of head computer tomography (CT) scans to assess the presence of a bleeding. 
+To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
+In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
+When AI advice is given, this advice can be either correct or incorrect.
 The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across brain scans.
@@
 $$
-In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.
+In our case study, $P_{si}$ is the conditional probability that subject $s$ gives the correct response to item (i.e., CT scan) $i$.
 
@@
 Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
-In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available.
+In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available for GLMMs.
 Our strategies to specify population parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferred over standardized effect sizes for tailored simulation based sample size planning.
-This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available domain knowledge of domain experts.
+Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation based sample size planning.
+This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
 
 Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
-In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions, that can be adapted for our purposes.
+In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions [@gelmanBayesianWorkflow2020], that can be adapted for our purposes.
 
@@
 
-When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
+When introducing the simulation function for our case study, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
 
@@
 However, these are sometimes not readily available.
-All parameter values in our present case study have been determined based on results from related study designs from the literature.
-Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions seems plausible.
-
-We now outline the strategies we used to determine plausible parameter values:
+All parameter values in our present case study have been determined based on results from distantly related study designs in the literature.
+Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions seem plausible.
+
+We now outline our main strategy to determine plausible parameter values for the fixed effects ($\beta$ parameters):
 Unfortunately, the model parameters in a binomial GLMM are hard to interpret in isolation because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
+The most simple interpretation, that allows us to ignore the random effects for now, works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
 Then the model implied probability that such a person solves such an item accurately is given by:
@@
 $$
-We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (that a subject with average ability solves an item with average difficulty) for each experimental condition.
+We can revert this perspective by choosing plausible probability values based on domain knowledge and deriving the parameter values implied by these probabilities for each experimental condition.
 
@@
 
+which we compute in R as:
+
 ```{r}
@@
 
+This leaves us with the question on how to determine plausible values for the two remaining model parameters ($\sigma_S$, and $\sigma_I$) that are the standard deviations for the random intercepts.
+For this, we introduce two more strategies.
+
 ### Insightful descriptive statistics
@@
 
-For most GLMMs, the interpretation of model parameters are quite complicated which makes it difficult to specify plausible values for a specific application.
-However, the mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
+<!-- For most GLMMs, the interpretation of model parameters are quite complicated which makes it difficult to specify plausible values for a specific application. -->
+The mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
 The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
-For example, domain experts might not be able to specify plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
+For example, domain experts might not be able to directly choose plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
 However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.
-which descriptive statistics they expect to observe (on average), when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
+which descriptive statistics they expect to observe (on average) when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
 The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
@@
 In the last section, we showed how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
-Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
+Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty [@fahrmeirRegressionModelsMethods2021].
 The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
+
 For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
-
 However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
-How much this estimated probability differs from the model implied probability of an average expert and an average items depends on the standard deviation of the random subject and random item intercepts (the two quantities are only equal if both standard deviations would be zero).
-We can use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
-
-We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model implied true values) and simply compute the relative frequency of correct diagnoses for each experimental conditions.
+How much this estimated probability differs from the model implied probability of an average expert and an average items depends on the standard deviations of the random intercepts (the two quantities are only equal if both standard deviations would be zero).
+We want to use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
+
+We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model implied true values) and simply compute the relative frequency of correct diagnoses for each experimental condition.
 
@@
 
-We can use these descriptive statistics to judge whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
-
+We tried using these descriptive statistics to judge together with our domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
 However, in our example they are not informative enough to determine a final set of plausible parameter values.
@@
 
-The inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
+The already discussed inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
 One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
@@
 
-In the last section, we simulated a large simulated dataset and computed a descriptive statistic, the relative frequencies of correct diagnoses, for each experimental condition.
-We will now use the model implied probability of each simulated data entry (y_prob) to visualize the whole model based implied marginal distribution of correct diagnoses for each condition.
+In the last section, we simulated a large simulated dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
+We will now use the model implied probability of each simulated data point (stored in the variable `y_prob`) to visualize the whole model based implied marginal distribution of correct diagnoses for each experimental condition.
 
@@
 ```{r margdist1, fig.cap="(ref:margdist1-caption)"}
-library(tidyverse)
-library(ggdist)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
 dat %>% 
@@
 
-```{r}
-library(tidyverse)
-library(ggdist)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
-dat %>% 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>%
-  group_by(condition) %>%
-  summarize(freq_y_1 = sum(y_bin) / n())
-```
-
 Figure \@ref(fig:margdist1) shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
@@
 ```{r margdist2, echo=FALSE, fig.cap="(ref:margdist2-caption)"}
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0i = 0.01)
+set.seed(1)
+dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0i = 0.01)
 dat %>% 
@@
 ```{r, include=FALSE}
+set.seed(1)
 dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
@@
 ```{r margdist3, echo=FALSE, fig.cap="(ref:margdist3-caption)"}
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 3, sd_u0i = 3)
+set.seed(1)
+dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 3, sd_u0i = 3)
 dat %>% 
@@
 
+#### THEORY
+
 In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
@@
 
-All parameter values in our present case study have been determined from results from related prior work and repeated discussions with our affiliated domain experts in Radiology to validate our assumptions.
-Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for ICH.
+All parameter values in our present case study have been determined based on repeated discussions with our affiliated domain experts in Radiology to validate our assumptions.
+Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for intracranial hemorrhage.
 Existing studies indicate that radiologists typically demonstrate high accuracies, often exceeding or hovering around 90%, while interns have been shown to perform below 80%, and medical students fall even shorter.
@@
 Subsequently, we consulted several published papers investigating the effect of correct and incorrect advice on decision-making performance in other settings.
-From their findings, we inferred that both experts and non-experts would benefit from correct and suffer losses from incorrect advice.
-However, the magnitude of these effects would be substantially greater for non-experts, given their demonstrated reliance on advice compared to experts.
+From their findings, we inferred that both experts and non-experts should benefit from correct and suffer losses from incorrect advice.
+However, the magnitude of these effects should be substantially greater for non-experts, given their demonstrated reliance on advice compared to experts.
 We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
@@
 On the other hand, using a more complex model (that perhaps even goes beyond a standard GLMM) for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
-In such more complicated scenarios, the final model specification might also depend on the next decision on specifying the theoretical estimand.
 
@@
 However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]
+Therefore, for our case study, we have written our own function to simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions).
+We think this choice assists applied researchers in better understanding all model assumptions before using lme4 to analyze the simulated datasets.[^2]
 
@@
 
-In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
+In this section, we show how to fit a GLMM with lme4.
 We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
@@
 ```
+
+We can inspect the estimates for all model parameters with the `summary` command:
+
+```{r}
+summary(fit)
+```
+
+In the lme4 output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
 
@@
 When model parameters are not known, the corresponding parameter estimates are plugged into the equation to compute an estimate of the contrast instead.
-
-In general, we can inspect the estimates for all model parameters with the `summary` command:
-
-```{r}
-summary(fit)
-```
-
-In the lme4 output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
 

### 2024-04-09 docs/manuscript.Rmd
iterate on remainder of case study sections

@@
 Will introduce each step in a theoretical section, followed by the practical application based on a case study.
+All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
 
@@
 This random intervention allows to estimate our theoretical estimand, although in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each brain scan.
-
-For our example, each probability expression in our theoretical estimand can be modeled with a GLMM.
+For our example, each probability expression in our theoretical estimand can be modeled with the same GLMM.
 Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability and these estimates can be used to compute an estimate for each of the four probability contrasts.
+For pedagogical reasons, we will skip the concrete definition of our empirical estimand until we have discussed how to simulate data based on a concrete GLMM in the next section.
 
 ## Simulate the data generating process
-
-### Data simulation
 
@@
 
-As discussed earlier, tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with a similar design or any pilot data.
+Tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with the same design or pilot data.
 In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
@@
 Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation based sample size planning.
-This requires new strategies on how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
+The basic idea of all strategies is how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
 
 Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
-In this literature, there have been developed some strategies to monitor the plausibility of (prior) model assumptions [@gelmanBayesianWorkflow2020], that can be adapted for our purposes.
+In this literature, strategies have been developed to monitor the plausibility of (prior) model assumptions [@gelmanBayesianWorkflow2020], that can be adapted for our purposes.
 
@@
 
-For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
-However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
-How much this estimated probability differs from the model implied probability of an average expert and an average items depends on the standard deviations of the random intercepts (the two quantities are only equal if both standard deviations would be zero).
+For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice (the conditional perspective).
+However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty (the marginal perspective).
+How much the two probabilities differ depend on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations would be zero).
 We want to use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
@@
 We tried using these descriptive statistics to judge together with our domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
-However, in our example they are not informative enough to determine a final set of plausible parameter values.
+However although they result was deemed plausible, in our example these statistics are not informative enough to determine a final set of plausible parameter values.
 For this reason, we will additionally look at insightful model based quantities.
@@
 
-The already discussed inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
+The discussed inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
 One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
@@
 In the last section, we simulated a large simulated dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
-We will now use the model implied probability of each simulated data point (stored in the variable `y_prob`) to visualize the whole model based implied marginal distribution of correct diagnoses for each experimental condition.
+We will now use the model implied probability of each simulated data point (stored in the variable `y_prob`) to visualize the whole model implied marginal distribution of correct diagnoses for each experimental condition.
 
@@
 When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
-This decision is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests and confidence intervals with the marginaleffects R package using the delta method).
+While the majority of psychological research uses hypothesis testing, there are repeated claims that a focus on estimation can also be a good choice in many circumstances (CITATION).
+This decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests and confidence intervals with the marginaleffects R package using the delta method).
 
@@
 In the earlier section on our theoretical estimand we translated this verbal description into four probability statements that are specified outside of any specific statistical model.
-The next translation step is to specify how these probabilites can be estimated within the framework of our concrete GLMM.
-
-The first of the four statements was:
+The next translation step is to specify the empirical estimand, that is how these probabilites can be estimated within the scope of our concrete GLMM.
+
+The first of the four statements of the theoretical estimand was:
 
@@
 \begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, an average expert, an average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, an average expert, an average scan})
+& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
 \end{aligned}
 $$
-In the context of our GLMM, the corresponding model based quantity is:
+In the context of our GLMM, the corresponding probability statement is:
 
@@
 Plugging in the model equation of the GLMM produces an equation on how to compute this contrast when all model parameters are known.
+We have already discussed how to compute the individual probabilities in the section on specifying population parameters.
+When we want to estimate the above contrast based on observed data, the only difference is that model parameters are not known and we instead plug the corresponding parameter estimates into the equation.
 
@@
 
-When model parameters are not known, the corresponding parameter estimates are plugged into the equation to compute an estimate of the contrast instead.
-
-Unfortunately, there is no single model parameter that corresponds with our contrast of interest.
-the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-When we are interested in a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$), the model implied probability that such a person solves such an item accurately is given by:
-$$
-\begin{aligned}
-P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
-= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
-\end{aligned}
-$$
-In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
-In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
-$$
-\begin{aligned}
-P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
-= \text{inverse\_logit}[\beta_0]
-\end{aligned}
-$$
-One could use such reasoning, to determine the exact formula needed to compute our contrast of interest.
-However, this would be very tedious and we can use R to compute this contrast for us without doing the math.
-
-We can use the predict function of the lme4 package to compute the predicted probability for a correct diagnosis based on our model, plugin the two sets of predictor values, and compute the difference between the two estimated probabilities.
+We could use our knowledge of the structure of our GLMM to determine the exact formula needed to compute our contrast of interest and then plug in the parameter estimates manually from the `summary(fit)` output.
+However, this would be very tedious and we can use R to compute this contrast without doing the math.
+
+We could use the `predict` function of the lme4 package to compute the predicted probability for a correct diagnosis based on our model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
 
@@
 
-The argument `type = "response"` specifies that predictions are made on the probability scale, while `re.form = NA` sets all random effects to 0.
-
-We could use this method to compute point estimates for all four contrast that are part of our estimand.
+The argument `type = "response"` specifies that predictions are made on the probability scale (instead of the log-odds scale of the $\beta$ parameters), while `re.form = NA` sets all random effects to 0.
+
+We could use this method to compute point estimates for all four contrast that are part of our empirical estimand.
 However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute hypothesis tests or confidence interval for our point estimates.
 
-The marginaleffects package [@R-marginaleffects] is a very flexible package to compute hypothesis tests and confidence intervals for contrasts with a variety of statistical models, including GLMM estimated with lme4.
-
-First, we compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
+The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute hypothesis tests and confidence intervals for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
+
+First, we specify a grid of all combinations of predictor variable and then compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
 
@@
 
-Note that the result also contains the two missing by design conditions that are never observed in the actual study ($advice\_present = 0, advice\_correct = 1, expert = 1$ and $advice\_present = 0, advice\_correct = 1, expert = 0$), and we should never interpret those.
-
-Next, we use these predictions to compute the four specific contrasts that are part of our estimand.
+The point estimates for all experimental conditions are reported in the `Estimate` column.
+Note that the output also contains the two missing by design conditions that will never be observed in the actual study ($advice\_present = 0, advice\_correct = 1, expert = 1$ and $advice\_present = 0, advice\_correct = 1, expert = 0$).
+This is no problem as long as we never interpret those estimates.
+
+Next, we use the estimated probabilities to compute the four specific contrasts that are part of our empirical estimand.
 For this we must specify which rows in `probs` have to be subtracted from each other.
@@
 The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2.
-The argument `equivalence = c(0, 0)` can be used to compute one-sided p-values testing whether the constrast in the population is smaller than 0 (`p (NonSub)` column) or greater than zero (`p (NonInf)` column).
-We computed the contrasts in a way that we expect positive values for all four of them.
+The argument `equivalence = c(0, 0)` can be used to compute one-sided p-values, testing whether the contrast in the population is smaller than 0 (`p (NonSub)` column) or greater than 0 (`p (NonInf)` column).
+The point estimates for four contrasts are reported in the `Estimate` column.
+Note that to facilitate interpretation, we arranged the contrasts in a way that we theoretically expect positive values for all four of them.
 
@@
 
-If we are interested in hypothesis testing in our case study, we test a combined null hypothesis that consists of four separate null hypotheses:
+If we chose hypothesis testing for our case study, we would test a combined null hypothesis $H_0$ that consists of four separate null hypotheses:
 
@@
 $$
-The combined null hypothesis should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
+The combined null hypothesis $H_0$ should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting\; @dmitrienkoTraditionalMultiplicityAdjustment2013].
 In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
 
-With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value in the `p (NonInf)` column for the third hypothesis is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
-Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
+With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value in the `p (NonInf)` column for the third hypothesis is not significant) and therefore also not reject the combined null hypothesis for this particular (simulated) dataset.
+Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis $H_1$ is actually true in the population.
 
@@
 
-If we are interested in parameter estimation in our case study, we focus on the two-sided confidence intervals of the four contrasts of interest.
+If we chose parameter estimation for our case study, we would focus on the two-sided confidence intervals of the four contrasts of interest.
 With a standard confidence level of $1 - \alpha = 0.95$, plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
-Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts.
+Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts (which the expected behavior for binomial GLMMs).
 The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice) where both underlying probabilities are close to 1.
@@
 
-Conducting the previous steps enable the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
+Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
 The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
-On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter.
+On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter values.
 When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.
-When planning for precision, the function should instead return the width of the confidence interval(s) of interest.
-Secondly, we must run this function repeatedly with the same sample and population parameters.
-Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computations, that is running simulations on multiple cores of the computer at the same time to reduce total computation time.
+When planning for precision, the function should return the width of the confidence interval(s) of interest.
+Secondly, we must run this function repeatedly with the same sample size and population parameters.
+Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computing, that is running simulations on multiple cores of the computer at the same time to reduce total run time.
 Thirdly, the results of the repeated simulation must be collected and aggregated.
-When planning for power, we compute the relative frequency of a significant p-value across repeated simulations.
-When planning for precision, we compute the average width of the confidence interval.
-Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted statistical power or the expected width of the confidence interval.
+When planning for power, we compute the relative frequency of a significant p-value(s) across repeated simulations.
+When planning for precision, we compute the average width of the confidence interval(s).
+Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted power or precision.
 
@@
 
-With all these considerations addressed, we are finally ready to run our simulation-based sample size planning analyses to plan for power and precision.
+We are finally ready to run our simulation-based sample size planning analyses to plan for power and for precision.
 Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and confidence intervals) in a single command.
@@
 
-Simulation-based smaple size planning can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
+Simulation-based sample size planning can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
 Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
-First, we enable parallelization and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
+First, we enable parallelization with the `plan` function and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
 
@@
 
-The result of the computation is a data frame that contains the p-values and confidence intervals of all specified contrasts for each simulated dataset.
+The result of this computation is a data frame that contains the p-values and confidence intervals of all specified contrasts for each simulated dataset.
 In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
@@
 
-For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
+For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the $\alpha = 0.05$ level.
 Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^3]
@@
 As should be the case, power estimates in Figure \@ref(fig:finalpwr) increase with both the number of subjects and the number of items.
-The confidence intervals indicate how precisely power was estimated by our simulation.
+The confidence intervals reported here indicate how precisely power was estimated by our simulation.
 Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
@@
 As should be the case, precision estimates in Figure \@ref(fig:finalprecision) increase  (i.e., average width of confidence interval decreases) with both the number of subjects and the number of items.
+The confidence intervals reported here indicate how precisely the expected width of the confidence interval for our focal contrast was estimated by our simulation.
 Applying our simple heuristic of targeting an expected width smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.

### 2024-04-10 docs/manuscript.Rmd
delete the old case study chapter that has now been fully integrated into the new structure

@@
 
-# THIS WAS THE OLD CASE STUDY CHAPTER
-
-<!-- overview of this section -->
-
-In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
-All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
-
-## Experimental study design
-
-<!-- describe the study idea and design -->
-In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
-The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding.
-To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-This AI advice can be either correct (80% of cases) or incorrect (20%).
-In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
-After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
-This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
-In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
-The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
-
-## Our specific GLMM
-
-<!-- deep dive into our specific example -->
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
-$$
-g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$
-Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: 
-$$
-E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
-$$
-If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability:
-$$
-P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
-$$
-In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.
-
-In such a setting, we model this probability as
-$$
-P_{si} = inverse\_logit(\eta_{si})
-$$
-with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
-$$
-logit(P_{si}) = \eta_{si}
-$$
-with the logit link $g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})$.
-
-In our case study, the probability of making an accurate diagnostic decision is assumed to depend on the predictors:
-
-- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to assess item $i$ 
-- $advice\_correct_{si}$: whether this advice was correct (1) or not (0) 
-- $expert_s$: whether subject $s$ was a task expert (1) or not (0)
-
-and the random effects:
-
--   $u_{0s}$: the deviation of subject $s$ from the average ability to solve an item (i.e., CT scan) with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
--   $u_{0i}$: the deviation of item (i.e., CT scan) $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$
-
-In total, we assume the model
-$$
-\begin{aligned}
-logit[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
-&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
-\end{aligned}
-$$
-or equivalently
-$$
-\begin{aligned}
-P_{si} = inverse\_logit[&(\beta_0 + u_{0s} + u_{0i}) + \\
-&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
-\end{aligned}
-$$
-with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
-
-In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
-To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.
-
-## Data simulation
-
-The following R function simulates a full dataset structured according to the design of our case study.
-The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.
-
-```{r}
-simulate <- function(n_subjects = 100, n_items = 50,
-  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
-  b_ea = 0.790, b_ec = -1.393,
-  sd_u0s = 0.5, sd_u0i = 0.5, ...){
-  require(dplyr)
-  require(faux)
-  # simulate design
-  dat <- add_random(subject = n_subjects, item = n_items) %>%
-    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
-    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
-    mutate(advice_correct = if_else(advice_present == 1L, 
-                                    rbinom(n(), 1L, prob = 0.8), 0L)) %>%
-    # add random effects
-    add_ranef("subject", u0s = sd_u0s) %>%
-    add_ranef("item", u0i = sd_u0i) %>%
-    # compute dependent variable
-    mutate(linpred = b_0 + u0i + u0s +
-        b_e * expert + b_a * advice_present + b_c * advice_correct +
-        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
-    mutate(y_prob = plogis(linpred)) %>%
-    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
-  dat
-}
-```
-
-In the first six lines of the function definition, we set some default parameter values (which we will explain in a later section) and load the packages we use to manipulate and simulate data. 
-In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).
-Thus, the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
-We add a between-subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
-The next two lines simulate that in $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
-Next, we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
-As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user.
-With all design variables done, we are ready to simulate our model equation outlined in the last section.
-The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects, and model parameters as assumed by our model.
-We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
-In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject makes an accurate diagnostic decision for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
-
-## Model fitting
-
-<!-- ## The lme4 package in R -->
-
-<!-- describe the lme4 package and how it works-->
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
-The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
-
-[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
-
-However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.[^2]
-
-[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
-
-In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
-We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
-However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.
-
-```{r, message=FALSE}
-library(tidyverse)
-set.seed(1)
-dat <- simulate(n_subjects = 100, n_items = 50)
-```
-
-The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:
-
-```{r, message=FALSE}
-library(lme4)
-f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
-  expert:advice_present + expert:advice_correct +
-  (1|subject) + (1|item)
-```
-
-The first two lines look similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`).
-The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
-The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
-
-In lme4, a GLMM is fitted with the `glmer` function.
-By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs.
-inaccurate (0) diagnosis.
-
-```{r}
-fit <- glmer(f, data = dat, family = "binomial")
-```
-
-## Model interpretation
-
-We can inspect the estimates for all model parameters with the `summary` command:
-
-```{r}
-summary(fit)
-```
-
-In the output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
-
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
-Then the model implied probability that such a person solves such an item accurately is given by:
-$$
-\begin{aligned}
-P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
-= inverse\_logit[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
-\end{aligned}
-$$
-
-In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
-In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
-$$
-\begin{aligned}
-P(Y=1|expert = 0, advice\_present = 0, advice\_correct = 0, u_{0s} = 0, u_{0i} = 0) = \\
-= inverse\_logit[\beta_0]
-\end{aligned}
-$$
-
-Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
-
-With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects):
-
-(ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package.
-
-```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"}
-library(marginaleffects)
-plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"),
-  type = "response") + ylim(c(0.3, 1))
-```
-
-## Hypothesis testing
-
-However, we need to think about the model parameters again when we want to test hypotheses that we have theoretically derived from some research question.
-Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.
-
-The `Fixed effects` table in the lme4 summary output also includes p-values for hypothesis tests with null hypotheses of the style $H_0: \beta = 0$.
-However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.
-
-For our case study, imagine the following combined hypothesis: *We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
-
-This combined hypothesis leads to the following four separate null hypotheses to be tested: $$
-\begin{aligned}
-H_{01}:\ &\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0 \\
-H_{02}:\ &\beta_{a} + \beta_{c} \leq 0 \\
-H_{03}:\ &\beta_{a} + \beta_{ea} \geq 0 \\
-H_{04}:\ &\beta_{a} \geq 0
-\end{aligned}
-$$
-
-We arrive at these inequalities based on the following logic, exemplified here only for $H_{01}$: The first null hypothesis states that *an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.* This implies the following inequality for each subject $s$ and item $i$ $$
-inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}] \leq inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e]
-$$ which simplifies to $\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0$.
-
-We can specify and test hypotheses like these with the multcomp package [@hothornSimultaneousInferenceGeneral2008] as follows:
-
-```{r, message=FALSE}
-library(multcomp)
-null_hypotheses <- c(
-  "advice_present + advice_correct + expert:advice_present + 
-   expert:advice_correct <= 0",
-  "advice_present + advice_correct <= 0",
-  "-1 * (advice_present + expert:advice_present) <= 0",
-  "-1 * (advice_present) <= 0")
-glht <- glht(fit, linfct = null_hypotheses)
-summary(glht, test = univariate())$test$pvalues
-```
-
-Because all hypotheses tested simultaneously with the `glht` function must have the same direction, we flip the sign of inequalities three and four by multiplying them with $-1$.
-The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously [@hothornSimultaneousInferenceGeneral2008].
-However, the combined null hypothesis in our exemplary research question should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting; @dmitrienkoTraditionalMultiplicityAdjustment2013].
-In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
-Thus, we request unadjusted p-values by setting `test = univariate()` in the `summary` command.
-With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value for hypothesis $H_{03}$ is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
-Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
-
-## Specification of plausible parameter values
-
-When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
-
-Ideally, one would rely on meta-analytic results or conclusive data from pilot studies.
-However, these are sometimes not readily available.
-All parameter values in our present case study have been determined based on results from related prior work.
-Additionally, we had repeated discussions with our affiliated domain experts in radiology to check our assumptions.
-
-We now outline a few strategies on how to determine plausible parameter values.
-We have already seen in our discussion of model interpretation how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
-We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (for an average subject and an average item).
-
-```{r, echo=FALSE}
-probtable <- data.frame(
-  col1 = 
-    c("no advice, no expert", "no advice, expert", "false advice, no expert", 
-      "false advice, expert", "correct advice, no expert", "correct advice, expert"),
-  col2 =
-    c("0.70", "0.90", "0.40", "0.85", "0.90", "0.95"),
-  col3 = 
-    c("$logit(0.70) = \\beta_0$",
-      "$logit(0.90) = \\beta_0 + \\beta_e$", 
-      "$logit(0.40) = \\beta_0 + \\beta_a$",
-      "$logit(0.85) = \\beta_0 + \\beta_e + \\beta_{a} + \\beta_{ea}$",
-      "$logit(0.90) = \\beta_0 + \\beta_a + \\beta_c$",
-      "$logit(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
-apa_table(
-  probtable
-  , col.names = c("Experimental condition",
-    "$P(Y=1|\\mathbf{X=x}, \\mathbf{U} = \\mathbf{0})$",
-    "Implied equation")
-  , caption = "Assumed probabilities that an average subject solves an average item in each experimental condition."
-  , note = "Implied equations are derived based on the model equations and setting all random intercept terms to 0."
-  , escape = FALSE
-)
-```
-
-Table \@ref(tab:probtable) shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
-The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:
-
-```{r}
-b_0 <- qlogis(0.7)
-b_e <- qlogis(0.9) - b_0
-b_a <- qlogis(0.4) - b_0
-b_ea <- qlogis(0.85) - b_0 - b_e - b_a
-b_c <- qlogis(0.9) - b_0 - b_a
-b_ec <- qlogis(0.95) - b_0 - b_e - b_a - b_c - b_ea
-c(b_0 = b_0, b_e = b_e, b_a = b_a, b_c = b_c, b_ea = b_ea, b_ec = b_ec)
-```
-
-It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g. 
-$$
-\begin{aligned}
-P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
-= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
-\end{aligned}
-$$
-
-```{r}
-plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec)
-```
-
-Although the derivations above are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty.
-The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
-For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice.
-However, even if the model were true, we would not observe this probability value if we estimated the marginal probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
-
-The inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
-One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
-However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
-Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
-
-Earlier, we have already encountered one way to visualize the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
-However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory.
-A more sophisticated strategy is to simulate a large dataset and directly compute the averages, contrasts and distributions we are interested in.
-
-(ref:margdist1-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions.
-
-```{r}
-library(tidyverse)
-library(ggdist)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
-dat %>% 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
-```{r}
-library(tidyverse)
-library(ggdist)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.5, sd_u0i = 0.5)
-dat %>% 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>%
-  group_by(condition) %>%
-  summarize(freq_y_1 = sum(y_bin) / n())
-```
-
-Figure \@ref(fig:margdist1) shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
-We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
-This difference increases with the variability of the random effects.
-
-Up to this point, we have not talked about plausible values for the standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$).
-Plots like the one above are a useful tool to decide whether the specified standard deviations are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
-
-In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$). This gives us a better way to see the variability between persons.
-
-(ref:margdist2-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01.
-
-```{r}
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0i = 0.01)
-dat %>% 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
-As an example, Figure \@ref(fig:margdist2) reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
-In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
-Patterns like these should be considered carefully and discussed with the domain experts.
-Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
-
-We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
-
-```{r, include=FALSE}
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
-dat %>% 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
-The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
-For Figure \@ref(fig:margdist3), we have set $\sigma_S = 3$ and $\sigma_I = 3$.
-This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-However, these high standard deviations do not account for the inherent variability and complexity of human performance.
-For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
-
-(ref:margdist3-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3.
-
-```{r}
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 3, sd_u0i = 3)
-dat %>% 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
-## Results
-
-With all these considerations addressed, we are finally ready to perform a power analysis.
-Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
-
-```{r}
-sim_and_analyse <- function(
-  formula_chr = "y_bin ~ 1 + expert + advice_present + advice_correct + 
-    expert:advice_present + expert:advice_correct + (1|subject) + (1|item)", 
-  null_hypotheses = c("advice_present + advice_correct + 
-    expert:advice_present + expert:advice_correct <= 0",
-    "advice_present + advice_correct <= 0",
-    "-1 * (advice_present + expert:advice_present) <= 0",
-    "-1 * (advice_present) <= 0"), ...){
-  require(lme4)
-  require(multcomp)
-  # simulate data
-  dat <- simulate(...)
-  # fit model
-  model <- glmer(as.formula(formula_chr), data = dat, family = "binomial")
-  # compute p-values
-  glht <- glht(model, linfct = null_hypotheses)
-  pvalues <- summary(glht, test = univariate())$test$pvalues
-  setNames(pvalues, paste0("p_H0", 1:length(null_hypotheses)))
-}
-```
-
-Power analysis can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
-Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
-First, we enable parallelization and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
-
-```{r, message=FALSE, eval=FALSE}
-library(future)
-plan("multisession", workers = 6)
-set.seed(2)
-```
-
-The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
-We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
-With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
-If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.
-
-```{r, message=FALSE, echo=FALSE, eval=FALSE}
-library(future)
-library(furrr)
-plan("multisession", workers = 6)
-set.seed(2)
-sim_design <- crossing(
-  rep = 1:300,
-  n_subjects = c(100, 150, 200, 250),
-  n_items = c(10, 30, 50, 70)
-) %>%
-  mutate(pvalues = future_pmap(., sim_and_analyse, 
-    .options = furrr_options(seed = TRUE))) %>%
-  unnest_wider(pvalues)
-saveRDS(sim_design, file = "results.rds")
-```
-
-```{r, message=FALSE, eval=FALSE}
-library(furrr)
-sim_design <- crossing(
-  rep = 1:300,
-  n_subjects = c(100, 150, 200, 250),
-  n_items = c(10, 30, 50, 70)
-) %>%
-  mutate(pvalues = future_pmap(., sim_and_analyse, 
-    .options = furrr_options(seed = TRUE))) %>%
-  unnest_wider(pvalues)
-```
-
-The result of the computation is a data frame that contains the p-values of all tested hypotheses for each simulated dataset.
-In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
-When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process.
-We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
-
-For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^3]
-
-[^3]: This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
-
-```{r, echo=FALSE, eval=FALSE}
-sim_design <- readRDS(file = "results.rds")
-```
-
-```{r finpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05.", eval=FALSE}
-library(binom)
-alpha <- 0.05
-power <- sim_design %>%
-  group_by(n_subjects, n_items) %>% 
-  summarise(power = mean(p_H01 < alpha & p_H02 < alpha & 
-                           p_H03 < alpha & p_H04 < alpha), 
-    n_sig = sum(p_H01 < alpha & p_H02 < alpha & 
-                  p_H03 < alpha & p_H04 < alpha),
-    n = n(),
-    ci.lwr = binom.confint(n_sig, n, method = "wilson")$lower,
-    ci.upr = binom.confint(n_sig, n, method = "wilson")$upper, 
-    .groups = "drop")
-power %>%
-  mutate(across(c(n_subjects, n_items), factor)) %>%
-  ggplot(aes(n_subjects, n_items, fill = power)) +
-  geom_tile() +
-  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
-                                power, ci.lwr, ci.upr)), 
-    color = "white", size = 4) +
-  scale_fill_viridis_c(limits = c(0, 1)) +
-  xlab("number of subjects") + ylab("number of items")
-```
-
-<!-- How many participants/stimuli are needed for concrete example -->
-
-As should be the case, power estimates in Figure \@ref(fig:finpwr) increase with both the number of subjects and the number of items.
-The confidence intervals indicate how precisely power was estimated by our simulation.
-Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
-In practice, data simulations are often run multiple times with adjusted combinations of sample sizes.
-When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of `n_subjects` and `n_items`.
-When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
-This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.
-
-Much has been written on the optimal amount of power to target in empirical research.
-The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints  [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
-When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
-For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation.
-This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice).
-If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items.
-However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule.
-The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
-Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
-
-## Expected effect size vs. smallest effect size of interest: sensitivity analysis
-
-In our case study, we have performed simulation-based power analysis from a single set of parameter values that reflect our assumptions of an expected effect size.
+## Sensitivity analysis
+
+In our case study, we have performed simulation-based sample size planning from a single set of parameter values that reflect our assumptions of an expected effect size.
 Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
 Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
-When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present.
-However, if the true effect is indeed smaller, the power will be lower, and the study might not be sufficiently informative.
-A common, more conservative strategy for sample size justification is to perform power analysis for the smallest effect size of interest (SESOI).
+When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present (or estimate the effect with satisfying precision).
+However, if the true effect is indeed smaller, power (or precision) will be lower, and the study might not be sufficiently informative.
+A common, more conservative strategy for sample size justification is to perform sample size planning for the smallest effect size of interest (SESOI).
 An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011].
 For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].
-When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called *sensitivity power analysis*.
-By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power depends on the size of the true effect.
-In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyze power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
+When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power or precision for different effect sizes in what is called *sensitivity power analysis*.
+By simulating power (or precision) for different effect sizes (in addition to the different number of subjects and items), one can make sure that power (or precision) would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power (or precision) depends on the size of the true effect.
+In simple study designs, it is possible to perform sensitivity analysis based on a single standardized effect size (e.g., analyze power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
 However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
-In this setting, sensitivity power analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
-Power could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
+In this setting, sensitivity analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
+Power (or precision) could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
 
@@
 1.
-The growing need for simulation-based sample size planning in experimental research: In order to conduct informative research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
+The growing need for simulation-based sample size planning in experimental research: In order to conduct informative research using varying complex experimental designs with GLMMs, formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
 Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
@@
 Teaching data simulation skills: The ability to conduct simulation-based sample size planning is a valuable skill that should be taught to experimental researchers.
-By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
+By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power or precision, and improve the quality of their experimental designs.
 Equipping researchers with the knowledge and tools to perform simulation-based sample size planning enables them to make informed decisions and enhance the rigor of their studies. 
@@
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform simulation-based a priori sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform a priori simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
 Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
@@
 Recognizing the value of simulation-based design analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
-They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
+They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power or precision.
 Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
@@
 In the wake of the replication crisis and myriad of underpowered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, analytic methods and software cannot be applied to conduct a priori sample size planning for GLMMs necessitating data simulation-based approaches.
+Yet, analytic methods and software cannot be applied to conduct a priori sample size planning for GLMMs, necessitating data simulation-based approaches.
 Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based sample size planning with GLMMs themselves.

### 2024-04-10 docs/manuscript.Rmd
small formatting fixes

@@
 ```{r margdist1, fig.cap="(ref:margdist1-caption)"}
+library(ggdist)
 dat %>% 
@@
 ```{r}
-pred_df <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), expert = c(1, 1))
+pred_df <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), 
+  expert = c(1, 1))
 pred_df
@@
 pred_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
+pred_df
 probs <- predictions(fit, newdata = pred_df, type = "response", re.form = NA)
@@
   # compute contrasts
-  contr_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
+  contr_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1,
+    expert = 0:1)
   predictions(model, newdata = contr_df, type = "response", re.form = NA) %>%
-    hypotheses(hypothesis = contrasts, equivalence = c(0, 0))
+    hypotheses(hypothesis = contrasts, equivalence = c(0, 0)) %>%
+    data.frame()
 }
@@
 sim_result <- crossing(
-  rep = 1:100,
+  rep = 1:300,
   n_subjects = c(100, 150, 200, 250),
@@
 sim_result <- crossing(
-  rep = 1:5,
+  rep = 1:300,
   n_subjects = c(100, 150, 200, 250),
@@
   summarise(
-    power = mean(`p.value.noninf_b1=b5` < alpha & `p.value.noninf_b8=b2` < alpha &
-      `p.value.noninf_b2=b6` < alpha & `p.value.noninf_b7=b1` < alpha), 
-    n_sig = sum(`p.value.noninf_b1=b5` < alpha & `p.value.noninf_b8=b2` < alpha &
-      `p.value.noninf_b2=b6` < alpha & `p.value.noninf_b7=b1` < alpha),
+    power = mean(`p.value.noninf_b1=b5` < alpha & 
+        `p.value.noninf_b8=b2` < alpha & `p.value.noninf_b2=b6` < alpha & 
+        `p.value.noninf_b7=b1` < alpha), 
+    n_sig = sum(`p.value.noninf_b1=b5` < alpha & 
+        `p.value.noninf_b8=b2` < alpha & `p.value.noninf_b2=b6` < alpha & 
+        `p.value.noninf_b7=b1` < alpha),
     n = n(),
@@
                                 precision, ci.lwr, ci.upr)), 
-    color = "white", size = 6) +
-  scale_fill_viridis_c(limits = c(0, 0.3), direction = -1) + guides(fill = guide_legend(reverse=FALSE))
+    color = "white", size = 4) +
+  scale_fill_viridis_c(limits = c(0, 0.3), direction = -1) +
+  guides(fill = guide_legend(reverse=FALSE))
 ```

### 2024-04-10 docs/manuscript.Rmd
new pdf for team review; update packages, update results file, update bib file

@@
 
+TODO: fertig ausformulieren
+
 Simulation-based sample size planning for generalized linear models is a complex topic that touches on a variety of statistical concepts and programming skills.
@@
 
-- Lakens ebook evtl. Andy Field für die Basics?
-- R for data science, Bruine article, faux documentation für die R skills
-- kumle power tutorial und irgendwelche paper (am besten aus AMPPS) zu GLMMs
-- what is your estimand? paper für das estimand topic
+- Lakens ebook [@lakensImprovingYourStatistical2022] evtl. Andy Field für die Basics?
+- R for data science [@wickhamDataScienceImport2023], Bruine article [@debruineUnderstandingMixedEffectsModels2021], faux documentation [@R-faux] für die R skills
+- kumle power tutorial [@kumleEstimatingPowerGeneralized2021] und irgendwelche paper (am besten aus AMPPS) zu GLMMs
+- what is your estimand? paper [@lundbergWhatYourEstimand2021] für das estimand topic
 
@@
 
+TODO: fertig ausformulieren
+
 Real studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
@@
 ### Missing data
+
+TODO: fertig ausformulieren
 
@@
 
+TODO: fertig ausformulieren
+
 While GLMMs are a very flexible model class that can handle a large variety of outcome variables, psychological researchers are becoming increasingly aware that many psychological datasets might profit from even more sophisticated models.
@@
 
+TODO: fertig ausformulieren
+
 The most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is some coefficient in a regression model.
@@
 ### Directed hypotheses and testing against values other than 0
+
+TODO: fertig ausformulieren
 
@@
 ### Combined hypotheses
+
+TODO: fertig ausformulieren
 
@@
 
+TODO: fertig ausformulieren
+
 As discussed earlier, the two major frameworks for sample size planning are planning for power and planning for precision.
@@
 ## No prior studies or pilot data
+
+TODO: fertig ausformulieren
 

### 2024-04-10 docs/manuscript.tex
last edits for preprint publication
new pdf for team review; update packages, update results file, update bib file

@@
 
-% \usepackage{etoolbox}
 \makeatletter
-\patchcmd{\HyOrg@maketitle}
+\usepackage{etoolbox}
+\patchcmd{\maketitle}
   {\section{\normalfont\normalsize\abstractname}}
@@
   {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
+\patchcmd{\maketitle}
   {\section{\protect\normalfont{\@title}}}
@@
 {}{\PatchFailed}
-\keywords{power analysis, data simulation, sample size, generalized linear mixed model\newline\indent Word count: 7600}
-\usepackage{lineno}
-
-\linenumbers
+\keywords{power analysis, data simulation, sample size, generalized linear mixed model\newline\indent Word count: 7504}
 \usepackage{csquotes}
@@
 \hypersetup{
-  pdftitle={A Tutorial on Tailored Simulation-Based Power Analysis for Experimental Designs with Generalized Linear Mixed Models},
+  pdftitle={A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models},
   pdfauthor={Timo K. Koch,1, 2, Florian Pargent,1, Anne-Kathrin Kleine1, Eva Lermer1,3, \& Susanne Gaube1,4},
@@
 
-\title{A Tutorial on Tailored Simulation-Based Power Analysis for Experimental Designs with Generalized Linear Mixed Models}
+\title{A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models}
 \author{Timo K.
@@
 
-\shorttitle{Simulation-Based Power Analysis for GLMMs}
+\shorttitle{Simulation-Based Sample Size Planning for GLMMs}
 
@@
 Timo K.
-Koch: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Florian Pargent: Formal Analysis, Methodology, Visualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Anne-Kathrin Kleine: Writing - Review \& Editing; Eva Lermer: Resources; Susanne Gaube: Writing - Review \& Editing, Supervision.
+Koch: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Florian Pargent: Formal Analysis, Methodology, Visualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Anne-Kathrin Kleine: Formal Analysis, Writing - Review \& Editing; Eva Lermer: Conceptualization, Resources, Writing - Review \& Editing; Susanne Gaube: Conceptualization, Supervision, Writing - Review \& Editing.
 \textsuperscript{$\dagger{}$} Timo K.
@@
 When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
-However, the recent replication crisis in Human-Computer Interaction (HCI) and other empirical research fields underlines the need for more rigorous statistical methodology and well-powered designs.
+However, the recent replication crisis underlines the need for more rigorous statistical methodology and well-powered designs.
 Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs, necessitating data simulation approaches.
+Yet, available methods and software for a priori power analyses for GLMMs are often limited to specific designs, while data simulation approaches offer more flexibility.
 Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs.
@@
 When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data (Lakens, 2022b).
-However, the recent replication crisis in Human-Computer Interaction (HCI) and several other disciplines grounded on empirical research has illustrated many challenges surrounding the reproducibility and reliability of findings (Cockburn, Dragicevic, Besançon, \& Gutwin, 2020; Robertson \& Kaptein, 2016; Yarkoni, 2022).
+However, the recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of findings (Yarkoni, 2022).
 As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
-While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct ``a priori'' (i.e., before data collection) power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
-In the present work, we provide a tutorial consisting of a concrete example for tailored a priori power analyses using data simulations based on GLMMs.
-
-\hypertarget{statistical-power}{%
-\subsection{Statistical power}\label{statistical-power}}
+While easy-to-use software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct ``a priori'' (i.e., before data collection) sample size planning for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
+In the present work, we provide a tutorial on performing tailored simulation based sample size planning for GLMMs.
+After introducing general concepts, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
+Then we describe general steps and decisions that are necessary whenever performing tailored data simulation.
+Because the details of these steps vary considerably for each specific study, we finish with a case study from the field of human-AI (artificial intelligence) interaction research.
+
+--\textgreater{}
+
+\hypertarget{prior-knowledge}{%
+\section{Prior Knowledge}\label{prior-knowledge}}
+
+TODO: fertig ausformulieren
+
+Simulation-based sample size planning for generalized linear models is a complex topic that touches on a variety of statistical concepts and programming skills.
+Although we give some short introduction to clarify our terminology, to profit the most from our tutorial, our readers require prior knowledge in the following topics (or refer to our collection of background literature):
+
+Readers should understand basic statistical concepts like hypothesis tests and their statistical power but also confidence intervals and their precision.
+Some background knowledge in empirical estimands, statistical estimates, and causal inference is very useful but not absolutely necessary.
+
+In addition, readers should have basic programming skills in R on how to simulate data and how to repeat these simulations in parallel to speed up computation.
+While there are a myriad of options for data simulation in R beyond the basic functionality in the base and stats packages, in our tutorial we use functions from the tidyverse and the faux package.
+
+Finally, readers should be familiar with regression modeling in general and GLMMs in particular.
+In our tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process.
+While it is not necessary to understand the technical details of how GLMMs are estimated, it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the values in the dependent variable are determined by the predictor variables and the random effects.
+
+Literature:
+
+\begin{itemize}
+\tightlist
+\item
+  Lakens ebook (Lakens, 2022a) evtl. Andy Field für die Basics?
+\item
+  R for data science (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023), Bruine article (DeBruine \& Barr, 2021), faux documentation (DeBruine, 2023) für die R skills
+\item
+  kumle power tutorial (Kumle, Võ, \& Draschkow, 2021) und irgendwelche paper (am besten aus AMPPS) zu GLMMs
+\item
+  what is your estimand? paper (Lundberg, Johnson, \& Stewart, 2021) für das estimand topic
+\end{itemize}
+
+\hypertarget{general-concepts}{%
+\section{General concepts}\label{general-concepts}}
+
+\hypertarget{planning-for-statistical-power-or-precision}{%
+\subsection{Planning for statistical power or precision}\label{planning-for-statistical-power-or-precision}}
+
+Conducting research with too small sample sizes can have many negative consequences.
+First, experiments may yield inconclusive or misleading results, hindering the accumulation of knowledge.
+Second, studies that are doomed to never finding a postulated effect due to small sample size waste resources by consuming time, effort, and funding without delivering meaningful results.
+For such reasons, many journals and funding agencies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+Although many pragmatic justifications are possible (Lakens, 2022b), the ideal is often considered to determine a suitable sample size a priori (i.e., before the study is conducted) based on some meaningful computation to assure that the study will be able to fulfill its purpose.
+In this way, adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
+
+--\textgreater{}
+
+The majority of empirical studies in psychology applies hypothesis testing.
+As a consequence, the dominant approach for sample size planning is based on power analysis (i.e., planing for power).
+However, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision).
+
+\hypertarget{planning-for-power}{%
+\subsubsection{Planning for power}\label{planning-for-power}}
 
@@
 
-Conducting underpowered research has many negative consequences.
-First, relying on underpowered experiments may yield inconclusive (if researchers acknowledge the small evidential value of an underpowered study in the limitations section) or misleading (if low power is ignored by the researchers) results, hindering the accumulation of knowledge.
-Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.
-
-\hypertarget{a-priori-power-analysis}{%
-\subsection{A priori power analysis}\label{a-priori-power-analysis}}
-
-A power analysis represents the act of calculating the statistical power for a given true effect and sample size.
-When running a power analysis before data collection, the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power.
-
-Thereby, a priori power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
-Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
-Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
-Adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
-Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-
-For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori power analysis is readily available (Champely et al., 2018; Faul, Erdfelder, Buchner, \& Lang, 2009).
-However, these software packages are often not flexible enough to perform power analysis for complex designs.
+When planning for power, a target is set for the statistical power of a hypothesis test of interest (a very common heuristic is 0.8).
+Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that (on average) would guarantee reaching this target.
+
+\hypertarget{planning-for-precision}{%
+\subsubsection{Planning for precision}\label{planning-for-precision}}
+
+Not all research questions are best answered using hypothesis testing.
+Especially for exploratory studies when little previous research has been conducted, only estimating effects of interest is often more useful.
+When no hypothesis tests are conducted when analysing the planned study, power analysis is not not available for sample size planning.
+Nonetheless, sample size still has a crucial effect on how informative the planned study will be to learn about the effect of interest.
+In this case, the expected width of a confidence interval is used as a target for sample size planning.
+The values inside a confidence interval are commonly interpreted as plausible values for the quantity of interest the confidence interval is supposed to estimate.
+More formally, a 95\%-confidence interval provides the smallest interval with the property that upon repeated sampling, 95\% of individual confidence intervals would include the true quantity of interest.
+Thus, a narrow confidence interval (assuming the confidence level is kept at a high value) is more informative about the size of the true effect than a wide interval.
+Apart from the desired confidence level, the width of a confidence interval depends strongly on the sample size.
+Because bigger samples carry more information, they lead to smaller confidence intervals (assuming everything else stays the same).
+When planning for precision, a target is set for the expected width of a confidence interval of interest.
+Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that (on average) would guarantee reaching this target.
 
@@
 
-As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs (Yarkoni, 2022).
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data in HCI and other empirical disciplines because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects (Fahrmeir, Kneib, Lang, \& Marx, 2021; Kaptein, 2016).
+As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs (Yarkoni, 2022).
+GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects (Fahrmeir, Kneib, Lang, \& Marx, 2021).
 
 GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data including hierarchical structures (Fahrmeir et al., 2021).
-In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value for another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
+In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
 This correlation can arise for various reasons: Responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), both participants come from the same neighborhood (clustering) or both participants responded to the same stimulus (stimulus effects).
@@
 The click patterns of participants in one session are likely to be correlated with their previous sessions.
-Finally, the outcome variable is binary (click/no click) for each interaction, which follows a binomial distribution.
-
-\hypertarget{power-analysis-for-glmms}{%
-\subsection{Power analysis for GLMMs}\label{power-analysis-for-glmms}}
-
+Finally, the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
+
+\hypertarget{simulation-based-sample-size-planning-with-glmms}{%
+\section{Simulation based sample size planning with GLMMs}\label{simulation-based-sample-size-planning-with-glmms}}
+
+For simpler statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori sample size planning is readily available (Champely et al., 2018; Faul, Erdfelder, Buchner, \& Lang, 2009; Kelley, Maxwell, \& Rausch, 2003; Kelley \& Rausch, 2006).
+However, these software packages are often not flexible enough to perform sample size planning for complex designs.
+
+Existing approaches for sample size planning with GLMMs have almost exlusively focused on planning for power.
 Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods (Murayama, Usami, \& Sakaki, 2022).
 Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
-Currently available formula-based software packages for power analysis often do not include GLMMs or are limited to very simple designs (Murayama et al., 2022; Westfall, Kenny, \& Judd, 2014), making it necessary to build data simulations tailored specifically to the study design.
-A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models (Arend \& Schäfer, 2019; Brysbaert \& Stevens, 2018; DeBruine \& Barr, 2021; Kumle, Võ, \& Draschkow, 2021; Lafit et al., 2021; Zimmer, Henninger, \& Debelak, 2022).
+Currently available formula-based software packages for power analysis with multilevel models often do not include GLMMs or are limited to very simple designs (Murayama et al., 2022; Westfall, Kenny, \& Judd, 2014), making it necessary to build data simulations tailored specifically to the study design.
+A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models (Arend \& Schäfer, 2019; Brysbaert \& Stevens, 2018; DeBruine \& Barr, 2021; Kumle et al., 2021; Lafit et al., 2021; Zimmer, Henninger, \& Debelak, 2022).
 However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs (but see Kumle et al., 2021 for a tutorial that also covers more advanced settings).
@@
 
-Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs.
-Thereby, we aim to equip researchers with the tools needed to simulate data and determine appropriate sample sizes for their own research.
-
-\hypertarget{the-present-case-study}{%
-\section{The present case study}\label{the-present-case-study}}
-
-In this section, we outline the steps for performing data simulation and a priori power analysis for GLMMs using a case study based on a specific experimental study design from the area of human-AI (artificial intelligence) interaction research.
+\begin{itemize}
+\tightlist
+\item
+  Insert table on existing R packages for power analysis with GLMMs
+\end{itemize}
+
+\hypertarget{when-do-we-profit-from-tailored-data-simulation}{%
+\section{When do we profit from tailored data simulation?}\label{when-do-we-profit-from-tailored-data-simulation}}
+
+Performing tailored simulation based sample size planning is more complicated and time consuming than using existing software tools for sample size planning.
+Thus, it is worth discussing under which circumstances we most profit from or are forced to perform tailored simulation based sample size planning, because existing software solutions do not meet our requirements.
+
+\hypertarget{complex-study-designs}{%
+\subsection{Complex study designs}\label{complex-study-designs}}
+
+TODO: fertig ausformulieren
+
+Real studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
+
+\hypertarget{missing-data}{%
+\subsubsection{Missing data}\label{missing-data}}
+
+TODO: fertig ausformulieren
+
+Missing data is a frequent issue in applied data analysis.
+There can be various reasons for why data is missing.
+For example data can be missing completely at random.
+Alternatively, subjects might drop out or produce missing data as a result of some covariate also measured in the dataset.
+A a different note, many experimental designs contain some conditions in which some values of the predictor variables are missing by design.
+Missing by design conditions can make data analysis more complicated because predictors have to be coded in highly specific ways that prevent the estimated GLMM to become unidentified.
+Whether missing data has an effect on the result of sample size planning depends on the characteristics of the missing data process.
+However, except for very simple scenarios, deciding whether missing data can be safely ignored in the sample size planning is often difficult to decide on a theoretical basis.
+A better strategy is to explicitly include the (assumed) process of missing data into the data simulation process.
+
+\hypertarget{data-generating-processes-beyond-standard-glmms}{%
+\subsubsection{Data generating processes beyond standard GLMMs}\label{data-generating-processes-beyond-standard-glmms}}
+
+TODO: fertig ausformulieren
+
+While GLMMs are a very flexible model class that can handle a large variety of outcome variables, psychological researchers are becoming increasingly aware that many psychological datasets might profit from even more sophisticated models.
+Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects.
+
+\begin{verbatim}
+- possibility to extend beyond standard GLMMs (e.g. zero-inflated outcomes, censoring, nonlinear effects, …)
+\end{verbatim}
+
+\hypertarget{complex-statistical-hypotheses}{%
+\subsection{Complex statistical hypotheses}\label{complex-statistical-hypotheses}}
+
+TODO: fertig ausformulieren
+
+The most common hypotheses tested in psychological research are of the type \(H_0: \beta = 0\), where \(\beta\) is some coefficient in a regression model.
+However, many research questions in psychology actually require testing more complex statistical hypotheses.
+
+\hypertarget{directed-hypotheses-and-testing-against-values-other-than-0}{%
+\subsubsection{Directed hypotheses and testing against values other than 0}\label{directed-hypotheses-and-testing-against-values-other-than-0}}
+
+TODO: fertig ausformulieren
+
+In the new era of preregistration and registered reports, most research questions are tested with directed hypotheses, because good theories should at least postulate whether some psychological effect of interest is positive or negative.
+Even better theories should be able to specify a smallest effect sizes of interest (SESOI) that must be exceeded if the psychological effect has some practical relevance in a specific application.
+In combination, this might require a test like \(H_0: \beta \leq 0.1\).
+
+\hypertarget{combined-hypotheses}{%
+\subsubsection{Combined hypotheses}\label{combined-hypotheses}}
+
+TODO: fertig ausformulieren
+
+Elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example \(H_0: \beta_0 + \beta_1 \leq 0\).
+If the research question consists only of a single hypothesis of this sort, it is often possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
+However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypotheses.
+For example, a combined null hypothesis \(H_0\) might consist of two single null hypotheses \(H_{01}: \beta_1 \leq 0\) and \(H_{02}: \beta_0 + \beta_1 \leq 0\).
+For some research questions, the combined null hypothesis \(H_0\) would be rejected if both \(H_{01}\) \emph{AND} \(H_{02}\) are rejected.
+For other research questions, the combined null hypothesis \(H_0\) would be rejected if \(H_{01}\) \emph{OR} \(H_{02}\) \emph{OR} both are rejected.
+If the global hypothesis \(H_0\) is combined with \emph{OR}, the p-values of the single hypotheses must be corrected for multiple testing to avoid \(\alpha\)-inflation for the global hypothesis.
+However, if the global hypothesis \(H_0\) is combined with \emph{AND}, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global hypothesis test.
+
+None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here.
+In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation based sample size planning.
+
+\hypertarget{planning-for-precision-1}{%
+\subsection{Planning for precision}\label{planning-for-precision-1}}
+
+TODO: fertig ausformulieren
+
+As discussed earlier, the two major frameworks for sample size planning are planning for power and planning for precision.
+Although planning for precision has been increasingly discussed as a more suitable strategy for exploratory research, all available software packages for sample size planning with GLMMs are based on power analysis.
+
+As long as one were interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
+However, we demonstrate in our case study that in realistic research projects with GLMMs, the contrasts of interests and their respective confidence intervals are usually situated on the scale of the (expected) outcome in contrast to the scale of the model parameters.
+We will show that tailored simulation based sample size planning can easily handle the planning for precision scenario.
+The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset and the expected width is estimated.
+
+\hypertarget{no-prior-studies-or-pilot-data}{%
+\subsection{No prior studies or pilot data}\label{no-prior-studies-or-pilot-data}}
+
+TODO: fertig ausformulieren
+
+All frameworks for sample size planning require the user to make assumptions about the expected effect size.
+Assuming the true effect is of this size (or greater), one can compute the (minimum) power of a hypothesis test or the (maximum) expected width of a confidence interval.
+Existing software packages for sample size planning for GLMMs usually require to provide the assumed effect in the unit of some standardized measure of effect size.
+When the researcher has access to similar studies or pilot data, providing such standardized effect sizes is feasible.
+However, this requirement can be an almost impossible challenge when no prior studies of pilot data are available.
+This problem is further exacerbated by the fact that GLMMs are so flexible, that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
+In our experience, using domain knowledge to construct a tailored data simulation is the only solution to determine plausible effect sizes in the absence of prior evidence.
+It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
+However, we would argue that when tailored data simulations are necessary anyways, performing the whole desing analysis in a customized way is preferred over using the existing software packages.
+
+\hypertarget{general-steps-in-tailored-simulation-based-sample-size-planning}{%
+\section{General steps in tailored simulation-based sample size planning}\label{general-steps-in-tailored-simulation-based-sample-size-planning}}
+
+Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions that we will outline in the remainder of our tutorial.
+Will introduce each step in a theoretical section, followed by the practical application based on a case study.
 All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (\url{https://osf.io/dhwf4/}).
 
-\hypertarget{experimental-study-design}{%
-\subsection{Experimental study design}\label{experimental-study-design}}
-
-In the present case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated.
-The goal is to understand how AI advice influences medical decision-making.
-Participants, radiologists (task experts) and students/interns (non-task experts), review head computer tomography (CT) scans to assess the presence of a bleeding.
+\hypertarget{define-the-estimand}{%
+\subsection{Define the estimand}\label{define-the-estimand}}
+
+\hypertarget{theory}{%
+\subsubsection{THEORY}\label{theory}}
+
+The first step in every research process should always be a clear definition of which theoretical quantity is necessary to answer the specific research question of the study under planning.
+Unfortunately, many psychological researchers are still not well trained in this task.
+The theoretical estimand is a unit specific quantity defined outside of any statistical model (Lundberg et al., 2021).
+Depending on the research and whether the research question is causal, this quantity might be counterfactual and not directly observable.
+Part of the theoretical estimand is also a clear definition of the target population, for which the quantity is of interest.
+
+When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reformulate the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
+For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantities within the scope that can be estimated with a regression model.
+Note however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression (CITATION).
+
+When the empirical estimand has been specified, the last step is to clearly define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
+When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique, and software package is used to fit the model, and how to compute hypothesis tests or confidence intervals.
+
+\hypertarget{practice}{%
+\subsubsection{PRACTICE}\label{practice}}
+
+In the present case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) in an AI-enabled diagnostic decision support system.
+The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography.
+In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans that are required to make the final diagnostic decision.
+The research goal is to validate the effectiveness of the AI-enabled advice.
+We consider the AI-enabled advice as effective, if the following pattern holds, which we will first describe verbally:
+
+\emph{We expect that for BOTH expert radiologists and students under training, correct AI-advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI-advice presented, AND, we expect that for BOTH experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.}
+
+I becomes clear that our theoretical estimand is be a causal quantity, based on several individual quantities that are defined outside of any statistical model (Lundberg et al., 2021).
+To elaborate further, we focus on only one part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists.
+For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice were presented and the probability that a correct diagnosis is made if no AI-advice were presented.
+Mathematically, we could describe this unit-specific quantity with some pseudo-equation like this (where \(s\) stands for a specific expert radiologist and \(i\) stands for a specific CT scan):
+
+\[
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{correct advice, specific expert, specific scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, specific expert, specific scan})
+\end{aligned}
+\]
+However, our verbal description is not precise enough yet, because the above quantity may differ for each combination of expert and scan.
+We must specify how we want to aggregate the quantity across different experts and scans.
+If we would compute the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts (CITATION).
+Although often debated whether this is a good idea, psychological research often focuses on a different quantity:
+Instead of averaging over experts and scans, we focus in a hypothetical \emph{typical} expert and a \emph{typical} scan, where \emph{typical} is usually defined as an average score on all attributes of the expert or scan.
+With this definition, the theoretical estimand in our above example can be expressed as:
+
+\[
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
+\end{aligned}
+\]
+In total, our theoretical estimand consists of four such expressions, the remaining three are:
+
+\[
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan})
+\end{aligned}
+\]
+\[
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{correct advice, average student, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, average student, average scan})
+\end{aligned}
+\]
+
+\[
+\begin{aligned}
+& P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan})
+\end{aligned}
+\]
+To complete our definition of the theoretical estimand, we have to clearly define out target population that consists of both persons and scans:
+With respect to persons, we are only interested in expert radiologists and medical students at German universities.
+With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
+Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system under study.
+
+While a theoretical estimand is defined outside of any statistical model and can be impossible to be observe directly, an empirical estimand is a quantity that can be estimated based on observed data.
+For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of brain scans, but the kind of AI-advice given for each brain scan is randomly assigned within participants.
+This random intervention allows to estimate our theoretical estimand, although in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each brain scan.
+For our example, each probability expression in our theoretical estimand can be modeled with the same GLMM.
+Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability and these estimates can be used to compute an estimate for each of the four probability contrasts.
+For pedagogical reasons, we will skip the concrete definition of our empirical estimand until we have discussed how to simulate data based on a concrete GLMM in the next section.
+
+\hypertarget{simulate-the-data-generating-process}{%
+\subsection{Simulate the data generating process}\label{simulate-the-data-generating-process}}
+
+\hypertarget{theory-1}{%
+\paragraph{THEORY}\label{theory-1}}
+
+When the estimand has been defined, the next step in the research process is to write code that simulates the (assumed) data generating process of the planned study.
+This requires specifying a generative process for all predictor variables used in the final data analysis.
+While such assumptions can be quite challenging for observational studies and/or continuous predictor variables, this is less of a problem for experimental studies with only categorical predictor variables.
+When all predictor variables have been simulated, one can use the structure of a suitable GLMM to simulate the dependent variable.
+To simulate the GLMM, one requires plausible values for all model parameters.
+We will discuss strategies on how these values can be obtained later.
+Because one has full control over the data generating process in a tailored simulation, it is simple to also model highly specific aspects of the planned study, like data missing by design or assuming that subjects drop out (i.e.~deleting their data from the simulated dataset) based on other variables in the simulated dataset.
+The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
+For this reason, deciding on a specific set of assumptions can easily feel overwelming.
+However, we would argue that even a strongly simplified data generating process (e.g.~including only a small number of interaction effects; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current status quo in psychological research, where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.
+
+\hypertarget{practice-1}{%
+\paragraph{PRACTICE}\label{practice-1}}
+
+In our case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated.
+Participants, radiologists (task experts) and students/interns (non experts), review a series of head computer tomography (CT) scans to assess the presence of a bleeding.
 To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-This AI advice can be either correct (80\% of cases) or incorrect (20\%).
 In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
-After reviewing the CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
+When AI advice is given, this advice can be either correct or incorrect.
+The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across brain scans.
+After reviewing a CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
 In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible.
-The goal of the present simulation-based power analysis is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power in the planned experiment.
-
-\hypertarget{the-lme4-package-in-r}{%
-\subsection{The lme4 package in R}\label{the-lme4-package-in-r}}
-
-In our case study, we use the lme4 R package (Bates, Mächler, Bolker, \& Walker, 2015), which is a state-of-the-art tool for fitting frequentist GLMMs.\footnote{For Bayesian GLMMs, the brms R package is currently the most prominent option (Bürkner, 2017).}
-The lme4 package includes a function called \texttt{simulate} that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
-
-However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore, in this tutorial, we simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions) to assist applied researchers in better understanding all model assumptions and then use lme4 to analyze the simulated datasets.\footnote{A less flexible alternative would be to use the simr package (Green \& MacLeod, 2016), which can be used to both simulate data and perform power analysis for models supported by the lme4 package.}
+The goal of simulation-based sample size planning is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power or precision in the planned experiment.
 
 \hypertarget{our-specific-glmm}{%
-\subsection{Our specific GLMM}\label{our-specific-glmm}}
-
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables \(\mathbf{X}\) and random effects \(\mathbf{U}\), transformed by a link function \(g()\) is modeled as a linear combination \(\eta\) of the predictor variables \(\mathbf{X}\), the random effects \(\mathbf{U}\) and the model parameters \(\mathbf{\beta}\) (Fahrmeir et al., 2021): \[
+\subparagraph{Our specific GLMM}\label{our-specific-glmm}}
+
+In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables \(\mathbf{X}\) and random effects \(\mathbf{U}\), transformed by a link function \(g()\) is modeled as a linear combination \(\eta\) of the predictor variables \(\mathbf{X}\), the random effects \(\mathbf{U}\) and the model parameters \(\mathbf{\beta}\) (Fahrmeir et al., 2021):
+\[
 g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-\] Equivalently, the conditional expected value is modeled as the linear combination \(\eta\), transformed by the inverse link function \(g^{-1}()\): \[
-E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
-\] If the dependent variable (i.e., diagnostic decision) \(Y\) is a binary variable with values \(0\) (i.e., inaccurate), or \(1\) (i.e., accurate), the conditional expected value is equivalent to the probability: \[
+\]
+Equivalently, the conditional expected value is modeled as the linear combination \(\eta\), transformed by the inverse link function \(g^{-1}()\):
+\[
+E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
+\]
+If the dependent variable (i.e., diagnostic decision) \(Y\) is a binary variable with values \(0\) (i.e., inaccurate), or \(1\) (i.e., accurate), the conditional expected value is equivalent to the probability:
+\[
 P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
-\] In our case study, \(P_{si}\) is the conditional probability that a subject \(s\) gives the correct response to item (i.e., CT scan) \(i\).
-
-In such a setting, we model this probability as \[
-P_{si} = inverse\_logit(\eta_{si})
-\] with the inverse-logit link \(g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}\) or equivalently \[
-logit(P_{si}) = \eta_{si}
-\] with the logit link \(g(P_{si}) = logit(P_{si}) = ln (\frac{P_{si}}{1 - P_{si}})\).
+\]
+In our case study, \(P_{si}\) is the conditional probability that subject \(s\) gives the correct response to item (i.e., CT scan) \(i\).
+
+In such a setting, we model this probability as
+\[
+P_{si} = \text{inverse\_logit}(\eta_{si})
+\]
+with the inverse-logit link \(g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}\) or equivalently
+\[
+\text{logit}(P_{si}) = \eta_{si}
+\]
+with the logit link \(g(P_{si}) = \text{logit}(P_{si}) = \text{ln} (\frac{P_{si}}{1 - P_{si}})\).
 
@@
 
-In total, we assume the model \[
+In total, we assume the model
+\[
 \begin{aligned}
-logit[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
+\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
 &\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
@@
 \end{aligned}
-\] or equivalently \[
+\]
+or equivalently
+\[
 \begin{aligned}
-P_{si} = inverse\_logit[&(\beta_0 + u_{0s} + u_{0i}) + \\
+P_{si} = \text{inverse\_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
 &\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
@@
 \end{aligned}
-\] with model parameters \(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\), \(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\).
+\]
+with model parameters \(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\), \(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\).
 
@@
 
-\hypertarget{data-simulation}{%
-\subsection{Data simulation}\label{data-simulation}}
+\hypertarget{simulation-function-in-r}{%
+\subparagraph{Simulation function in R}\label{simulation-function-in-r}}
 
@@
 
-In the first six lines of the function definition, we set some default parameter values (which we will explain in a later section) and load the packages we use to manipulate and simulate data.
+In the first six lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data.
 In our case study, each subject (\texttt{n\_subjects} in total) is assumed to respond to each item (i.e., CT scan; \texttt{n\_items} in total).
@@
 
-\hypertarget{model-fitting}{%
-\subsection{Model fitting}\label{model-fitting}}
-
-In this section, we show how to fit a GLMM with lme4, interpret the model, and test hypotheses derived from a research question.
+\hypertarget{strategies-on-specifying-the-population-parameters}{%
+\subsection{Strategies on specifying the population parameters}\label{strategies-on-specifying-the-population-parameters}}
+
+\hypertarget{theory-2}{%
+\subsubsection{THEORY}\label{theory-2}}
+
+Tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with the same design or pilot data.
+In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
+Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
+In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available for GLMMs.
+Our strategies to specify population parameters will require access to domain knowledge from domain experts.
+Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation based sample size planning.
+The basic idea of all strategies is how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
+
+Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
+In this literature, strategies have been developed to monitor the plausibility of (prior) model assumptions (Gelman et al., 2020), that can be adapted for our purposes.
+
+\hypertarget{practice-2}{%
+\subsubsection{PRACTICE}\label{practice-2}}
+
+When introducing the simulation function for our case study, we have used theoretically plausible values as defaults for all model parameters (\(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\), \(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\)), but have not talked about where these numbers came from.
+
+Ideally, one would rely on meta-analytic results or conclusive data from pilot studies.
+However, these are sometimes not readily available.
+All parameter values in our present case study have been determined based on results from distantly related study designs in the literature.
+Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions seem plausible.
+
+We now outline our main strategy to determine plausible parameter values for the fixed effects (\(\beta\) parameters):
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret in isolation because 1) the \(\beta\) parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
+The most simple interpretation, that allows us to ignore the random effects for now, works by imagining a subject with average ability (\(u_{0s} = 0\)) responding to an item (i.e., CT scan) with average difficulty (\(u_{0i} = 0\)).
+Then the model implied probability that such a person solves such an item accurately is given by:
+
+\[
+\begin{aligned}
+P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
+= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
+&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+\end{aligned}
+\]
+In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
+In all other experimental conditions, some terms drop from the equation because they are multiplied by \(0\).
+The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
+\[
+\begin{aligned}
+P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
+= \text{inverse\_logit}[\beta_0]
+\end{aligned}
+\]
+We can revert this perspective by choosing plausible probability values based on domain knowledge and deriving the parameter values implied by these probabilities for each experimental condition.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:probtable}Assumed probabilities that an average subject solves an average item in each experimental condition.}
+
+\begin{tabular}{lll}
+\toprule
+Experimental condition & $P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$ & Implied equation\\
+\midrule
+no advice, no expert & 0.70 & $logit(0.70) = \beta_0$\\
+no advice, expert & 0.90 & $logit(0.90) = \beta_0 + \beta_e$\\
+false advice, no expert & 0.40 & $logit(0.40) = \beta_0 + \beta_a$\\
+false advice, expert & 0.85 & $logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}$\\
+correct advice, no expert & 0.90 & $logit(0.90) = \beta_0 + \beta_a + \beta_c$\\
+correct advice, expert & 0.95 & $logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}$\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Implied equations are derived based on the model equations and setting all random intercept terms to 0.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+Table \ref{tab:probtable} shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
+The table can be used to compute the implied values for the \(\beta\) parameters, starting with the first equation and reinserting the computed \(\beta\) values in all following equations:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{b\_0 }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.7}\NormalTok{)}
+\NormalTok{b\_e }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
+\NormalTok{b\_a }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.4}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
+\NormalTok{b\_ea }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.85}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a}
+\NormalTok{b\_c }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_a}
+\NormalTok{b\_ec }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.95}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a }\SpecialCharTok{{-}}\NormalTok{ b\_c }\SpecialCharTok{{-}}\NormalTok{ b\_ea}
+\FunctionTok{c}\NormalTok{(}\AttributeTok{b\_0 =}\NormalTok{ b\_0, }\AttributeTok{b\_e =}\NormalTok{ b\_e, }\AttributeTok{b\_a =}\NormalTok{ b\_a, }\AttributeTok{b\_c =}\NormalTok{ b\_c, }\AttributeTok{b\_ea =}\NormalTok{ b\_ea, }\AttributeTok{b\_ec =}\NormalTok{ b\_ec)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##        b_0        b_e        b_a        b_c       b_ea       b_ec 
+##  0.8472979  1.3499267 -1.2527630  2.6026897  0.7901394 -1.3928518
+\end{verbatim}
+
+It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g.~
+\[
+\begin{aligned}
+P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
+= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
+\end{aligned}
+\]
+
+which we compute in R as:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plogis}\NormalTok{(b\_0 }\SpecialCharTok{+}\NormalTok{ b\_e }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{+}\NormalTok{ b\_ea }\SpecialCharTok{+}\NormalTok{ b\_ec)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 0.95
+\end{verbatim}
+
+This leaves us with the question on how to determine plausible values for the two remaining model parameters (\(\sigma_S\), and \(\sigma_I\)) that are the standard deviations for the random intercepts.
+For this, we introduce two more strategies.
+
+\hypertarget{insightful-descriptive-statistics}{%
+\subsubsection{Insightful descriptive statistics}\label{insightful-descriptive-statistics}}
+
+\hypertarget{theory-3}{%
+\paragraph{THEORY}\label{theory-3}}
+
+The mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
+The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
+For example, domain experts might not be able to directly choose plausible values for the \(\beta\) coefficients in a logistic regression model (which are measured on the log-odds scale).
+However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.~which descriptive statistics they expect to observe (on average) when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
+The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
+Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
+
+\hypertarget{practice-3}{%
+\paragraph{PRACTICE}\label{practice-3}}
+
+In the last section, we showed how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
+Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is \textbf{not} equal to the probability that a person with average ability solves an item with average difficulty (Fahrmeir et al., 2021).
+The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
+
+For example, we determined the \(\beta\) parameters in a way that corresponds to a desired conditional probability of \(0.95\), that an expert with average ability solves an item with average difficulty when presented with correct advice (the conditional perspective).
+However, even if the model were true, we would not observe that 95\% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty (the marginal perspective).
+How much the two probabilities differ depend on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations would be zero).
+We want to use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
+
+We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model implied true values) and simply compute the relative frequency of correct diagnoses for each experimental condition.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(tidyverse)}
+\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
+\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{3000}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{3000}\NormalTok{,}
+  \AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{)}
+\NormalTok{dat }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
+    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
+    \StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
+    \StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
+    \StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(condition) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{summarize}\NormalTok{(}\AttributeTok{relative\_frequency =} \FunctionTok{sum}\NormalTok{(y\_bin) }\SpecialCharTok{/} \FunctionTok{n}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 6 x 2
+##   condition                 relative_frequency
+##   <fct>                                  <dbl>
+## 1 no expert, no advice                   0.683
+## 2 expert, no advice                      0.881
+## 3 no expert, wrong advice                0.409
+## 4 expert, wrong advice                   0.828
+## 5 no expert, correct advice              0.883
+## 6 expert, correct advice                 0.938
+\end{verbatim}
+
+We tried using these descriptive statistics to judge together with our domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
+However although they result was deemed plausible, in our example these statistics are not informative enough to determine a final set of plausible parameter values.
+For this reason, we will additionally look at insightful model based quantities.
+
+\hypertarget{insightful-model-based-quantities}{%
+\subsubsection{Insightful model based quantities}\label{insightful-model-based-quantities}}
+
+\hypertarget{theory-4}{%
+\paragraph{THEORY}\label{theory-4}}
+
+Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
+This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
+An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
+For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
+Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
+For example, the domain expert might deem it unplausible that the 5\% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
+
+\hypertarget{practice-4}{%
+\paragraph{PRACTICE}\label{practice-4}}
+
+The discussed inequality of conditional and marginal effects in GLMMs (Fahrmeir et al., 2021) makes their interpretation more difficult.
+One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
+However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
+Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
+
+In the last section, we simulated a large simulated dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
+We will now use the model implied probability of each simulated data point (stored in the variable \texttt{y\_prob}) to visualize the whole model implied marginal distribution of correct diagnoses for each experimental condition.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(ggdist)}
+\NormalTok{dat }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
+    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
+    \StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
+    \StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
+    \StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ y\_prob, }\AttributeTok{y =}\NormalTok{ condition)) }\SpecialCharTok{+}
+  \FunctionTok{stat\_histinterval}\NormalTok{(}\AttributeTok{point\_interval =} \StringTok{"mean\_qi"}\NormalTok{, }\AttributeTok{slab\_color =} \StringTok{"gray45"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{breaks =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FloatTok{0.1}\NormalTok{), }\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/margdist1-1.jpeg}
+\caption{\label{fig:margdist1}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions.}
+\end{figure}
+
+Figure \ref{fig:margdist1} shows the model implied marginal distributions, including the mean, 66\% and 95\% intervals.
+We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
+This difference increases with the variability of the random effects.
+
+We can use plots like the one above as a useful tool to decide whether the specified standard deviations of the subject and item random intercepts (\(\sigma_S\) and \(\sigma_I\)) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
+
+In the next plot, we have set the item standard deviation to almost zero (\(\sigma_I = 0.01\)). This gives us a better way to see the variability between persons.
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/margdist2-1.jpeg}
+\caption{\label{fig:margdist2}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01.}
+\end{figure}
+
+As an example, Figure \ref{fig:margdist2} reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
+In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
+Patterns like these should be considered carefully and discussed with the domain experts.
+Parameter values (\(\beta\) parameters, and \(\sigma_S\)) should be adjusted if the implications do not seem reasonable.
+
+We could also have a closer look at variability between items by setting the subject standard deviation to almost zero (\(\sigma_S = 0.01\)).
+
+The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
+For Figure \ref{fig:margdist3}, we have set \(\sigma_S = 3\) and \(\sigma_I = 3\).
+This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
+However, these high standard deviations do not account for the inherent variability and complexity of human performance.
+For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/margdist3-1.jpeg}
+\caption{\label{fig:margdist3}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3.}
+\end{figure}
+
+\hypertarget{iterative-process-with-domain-experts}{%
+\subsubsection{Iterative process with domain experts}\label{iterative-process-with-domain-experts}}
+
+\hypertarget{theory-5}{%
+\paragraph{THEORY}\label{theory-5}}
+
+In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
+In a first step, domain experts can be interviewed to ``elicit'' their domain knowledge about how the future data of the planned study is expected to look like.
+As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
+For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
+Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
+Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
+During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
+
+\hypertarget{practice-5}{%
+\subparagraph{PRACTICE}\label{practice-5}}
+
+All parameter values in our present case study have been determined based on repeated discussions with our affiliated domain experts in Radiology to validate our assumptions.
+Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for intracranial hemorrhage.
+Existing studies indicate that radiologists typically demonstrate high accuracies, often exceeding or hovering around 90\%, while interns have been shown to perform below 80\%, and medical students fall even shorter.
+For simplicity, we assumed plausible probability values of .90 for experts and .70 for non-experts, respectively.
+Our experts confirmed that these values are realistic baselines for reviewing diverse head CT images without AI assistance.
+Subsequently, we consulted several published papers investigating the effect of correct and incorrect advice on decision-making performance in other settings.
+From their findings, we inferred that both experts and non-experts should benefit from correct and suffer losses from incorrect advice.
+However, the magnitude of these effects should be substantially greater for non-experts, given their demonstrated reliance on advice compared to experts.
+We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
+For our simulation, we used the probabilities of average participants to solve an average case, as shown in Table 1.
+
+\hypertarget{estimate-the-statistical-model}{%
+\subsection{Estimate the statistical model}\label{estimate-the-statistical-model}}
+
+\hypertarget{theory-6}{%
+\subsubsection{THEORY}\label{theory-6}}
+
+At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset that will later be collected in the planned study.
+The next step is to specify how the statistical model shall be estimated in the actual study collected later.
+This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer ``bobyqa''), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.\\
+Note that this does not always mean that one will specify the same GLMM that was used when specifying the data generating process.
+On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
+On the other hand, using a more complex model (that perhaps even goes beyond a standard GLMM) for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
+
+\hypertarget{practice-6}{%
+\subsubsection{PRACTICE}\label{practice-6}}
+
+In our case study, we use the lme4 R package (Bates, Mächler, Bolker, \& Walker, 2015), which is a state-of-the-art tool for fitting frequentist GLMMs.\footnote{For Bayesian GLMMs, the brms R package is currently the most prominent option (Bürkner, 2017).}
+The lme4 package includes a function called \texttt{simulate} that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
+
+However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
+Therefore, for our case study, we have written our own function to simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions).
+We think this choice assists applied researchers in better understanding all model assumptions before using lme4 to analyze the simulated datasets.\footnote{A less flexible alternative would be to use the simr package (Green \& MacLeod, 2016), which can be used to both simulate data and perform power analysis for models supported by the lme4 package.}
+
+In this section, we show how to fit a GLMM with lme4.
 We simulate data according to our model, in which 100 subjects respond to 50 items (we use \texttt{set.seed} to make the simulation reproducible).
@@
 \end{Shaded}
-
-\hypertarget{model-interpretation}{%
-\subsection{Model interpretation}\label{model-interpretation}}
 
@@
 ## expert:advice_present   1.0589     0.2940   3.601 0.000317 ***
-## expert:advice_correct  -1.8104     0.2915  -6.211 5.27e-10 ***
+## expert:advice_correct  -1.8104     0.2915  -6.210 5.28e-10 ***
 ## ---
@@
 
-In the output, the \texttt{Estimate} column in the \texttt{Fixed\ effects} table contains the estimates for the \(\beta\) parameters, while the \texttt{Std.Dev.} column in the \texttt{Random\ effects} table contains the estimates for \(\sigma_S\) and \(\sigma_I\).
-
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret because 1) the \(\beta\) parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation works by imagining a subject with average ability (\(u_{0s} = 0\)) responding to an item (i.e., CT scan) with average difficulty (\(u_{0i} = 0\)).
-Then the model implied probability that such a person solves such an item accurately is given by: \[
+In the lme4 output, the \texttt{Estimate} column in the \texttt{Fixed\ effects} table contains the estimates for the \(\beta\) parameters, while the \texttt{Std.Dev.} column in the \texttt{Random\ effects} table contains the estimates for \(\sigma_S\) and \(\sigma_I\).
+
+\hypertarget{compute-the-estimate}{%
+\subsection{Compute the estimate}\label{compute-the-estimate}}
+
+\hypertarget{theory-7}{%
+\subsubsection{THEORY}\label{theory-7}}
+
+When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
+While the majority of psychological research uses hypothesis testing, there are repeated claims that a focus on estimation can also be a good choice in many circumstances (CITATION).
+This decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests and confidence intervals with the marginaleffects R package using the delta method).
+
+\hypertarget{practice-7}{%
+\subsubsection{PRACTICE}\label{practice-7}}
+
+For our case study, we investigate the following research question:
+\emph{We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.}
+
+In the earlier section on our theoretical estimand we translated this verbal description into four probability statements that are specified outside of any specific statistical model.
+The next translation step is to specify the empirical estimand, that is how these probabilites can be estimated within the scope of our concrete GLMM.
+
+The first of the four statements of the theoretical estimand was:
+
+\[
 \begin{aligned}
-P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
-= inverse\_logit[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
+& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
 \end{aligned}
 \]
-
-In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
-In all other experimental conditions, some terms drop from the equation because they are multiplied by \(0\).
-The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented: \[
+In the context of our GLMM, the corresponding probability statement is:
+
+\[
 \begin{aligned}
-P(Y=1|expert = 0, advice\_present = 0, advice\_correct = 0, u_{0s} = 0, u_{0i} = 0) = \\
-= inverse\_logit[\beta_0]
+& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
+& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
 \end{aligned}
 \]
-
-Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
-
-With the marginaleffects package (Arel-Bundock, 2023), we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed \textbf{and} random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \ref{fig:marginaleffects}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(marginaleffects)}
-\FunctionTok{plot\_predictions}\NormalTok{(fit, }\AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"advice\_present"}\NormalTok{, }\StringTok{"advice\_correct"}\NormalTok{, }\StringTok{"expert"}\NormalTok{),}
-  \AttributeTok{type =} \StringTok{"response"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{ylim}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FloatTok{0.3}\NormalTok{, }\DecValTok{1}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/marginaleffects-1.jpeg}
-\caption{\label{fig:marginaleffects}Marginal distributions including means and 95\% confidence intervals for all experimental conditions computed with the marginaleffects package.}
-\end{figure}
+Plugging in the model equation of the GLMM produces an equation on how to compute this contrast when all model parameters are known.
+We have already discussed how to compute the individual probabilities in the section on specifying population parameters.
+When we want to estimate the above contrast based on observed data, the only difference is that model parameters are not known and we instead plug the corresponding parameter estimates into the equation.
+
+We could use our knowledge of the structure of our GLMM to determine the exact formula needed to compute our contrast of interest and then plug in the parameter estimates manually from the \texttt{summary(fit)} output.
+However, this would be very tedious and we can use R to compute this contrast without doing the math.
+
+We could use the \texttt{predict} function of the lme4 package to compute the predicted probability for a correct diagnosis based on our model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pred\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{advice\_present =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{advice\_correct =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }
+  \AttributeTok{expert =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\NormalTok{pred\_df}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##   advice_present advice_correct expert
+## 1              1              1      1
+## 2              0              0      1
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{preds }\OtherTok{\textless{}{-}} \FunctionTok{predict}\NormalTok{(fit, }\AttributeTok{newdata =}\NormalTok{ pred\_df, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{)}
+\NormalTok{preds}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##        1        2 
+## 0.939292 0.901923
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{preds[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ preds[}\DecValTok{2}\NormalTok{]}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##          1 
+## 0.03736901
+\end{verbatim}
+
+The argument \texttt{type\ =\ "response"} specifies that predictions are made on the probability scale (instead of the log-odds scale of the \(\beta\) parameters), while \texttt{re.form\ =\ NA} sets all random effects to 0.
+
+We could use this method to compute point estimates for all four contrast that are part of our empirical estimand.
+However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute hypothesis tests or confidence interval for our point estimates.
+
+The marginaleffects package (Arel-Bundock, 2023) is a very flexible, increasingly popular package to compute hypothesis tests and confidence intervals for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
+
+First, we specify a grid of all combinations of predictor variable and then compute estimated probabilities for all experimental conditions in our experiment with the \texttt{predictions} function:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pred\_df }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}\AttributeTok{advice\_present =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }\AttributeTok{advice\_correct =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }\AttributeTok{expert =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{)}
+\NormalTok{pred\_df}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 8 x 3
+##   advice_present advice_correct expert
+##            <int>          <int>  <int>
+## 1              0              0      0
+## 2              0              0      1
+## 3              0              1      0
+## 4              0              1      1
+## 5              1              0      0
+## 6              1              0      1
+## 7              1              1      0
+## 8              1              1      1
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{probs }\OtherTok{\textless{}{-}} \FunctionTok{predictions}\NormalTok{(fit, }\AttributeTok{newdata =}\NormalTok{ pred\_df, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{)}
+\NormalTok{probs}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## 
+##  Estimate Std. Error     z Pr(>|z|)     S 2.5 % 97.5 % advice_present
+##     0.738    0.02134  34.6   <0.001 867.2 0.696  0.779              0
+##     0.902    0.01739  51.9   <0.001   Inf 0.868  0.936              0
+##     0.975    0.00421 231.6   <0.001   Inf 0.966  0.983              0
+##     0.954    0.01454  65.6   <0.001   Inf 0.925  0.982              0
+##     0.423    0.03221  13.1   <0.001 128.6 0.360  0.486              1
+##     0.874    0.02793  31.3   <0.001 711.4 0.819  0.928              1
+##     0.909    0.00967  94.1   <0.001   Inf 0.890  0.928              1
+##     0.939    0.01091  86.1   <0.001   Inf 0.918  0.961              1
+##  advice_correct expert
+##               0      0
+##               0      1
+##               1      0
+##               1      1
+##               0      0
+##               0      1
+##               1      0
+##               1      1
+## 
+## Columns: rowid, estimate, std.error, statistic, p.value, s.value, conf.low, conf.high, advice_present, advice_correct, expert, y_bin 
+## Type:  response
+\end{verbatim}
+
+The point estimates for all experimental conditions are reported in the \texttt{Estimate} column.
+Note that the output also contains the two missing by design conditions that will never be observed in the actual study (\(advice\_present = 0, advice\_correct = 1, expert = 1\) and \(advice\_present = 0, advice\_correct = 1, expert = 0\)).
+This is no problem as long as we never interpret those estimates.
+
+Next, we use the estimated probabilities to compute the four specific contrasts that are part of our empirical estimand.
+For this we must specify which rows in \texttt{probs} have to be subtracted from each other.
+We will use the \texttt{hypotheses} function to compute our four contrasts of interest together with hypothesis tests and confidence intervals.
+We use the default inference options of the marginaleffects package that compute hypothesis tests and confidence intervals based on the approximate delta method.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{probs }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =} \FunctionTok{c}\NormalTok{(}
+    \StringTok{"b8 = b2"}\NormalTok{,}
+    \StringTok{"b2 = b6"}\NormalTok{,}
+    \StringTok{"b7 = b1"}\NormalTok{,}
+    \StringTok{"b1 = b5"}\NormalTok{),}
+    \AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## 
+##   Term Estimate Std. Error     z Pr(>|z|)    S    2.5 % 97.5 % p (NonSup)
+##  b8=b2   0.0374     0.0162  2.31    0.021  5.6  0.00563 0.0691      0.989
+##  b2=b6   0.0282     0.0279  1.01    0.312  1.7 -0.02653 0.0830      0.844
+##  b7=b1   0.1717     0.0173  9.93   <0.001 74.8  0.13780 0.2056      1.000
+##  b1=b5   0.3145     0.0280 11.24   <0.001 95.0  0.25965 0.3693      1.000
+##  p (NonInf) p (Equiv)
+##      0.0105     0.989
+##      0.1562     0.844
+##      <0.001     1.000
+##      <0.001     1.000
+## 
+## Columns: term, estimate, std.error, statistic, p.value, s.value, conf.low, conf.high, statistic.noninf, statistic.nonsup, p.value.noninf, p.value.nonsup, p.value.equiv 
+## Type:  response
+\end{verbatim}
+
+The expression \texttt{"b8\ =\ b2"} is special syntax to subtract the estimate in row number 8 from the estimate in row number 2.
+The argument \texttt{equivalence\ =\ c(0,\ 0)} can be used to compute one-sided p-values, testing whether the contrast in the population is smaller than 0 (\texttt{p\ (NonSub)} column) or greater than 0 (\texttt{p\ (NonInf)} column).
+The point estimates for four contrasts are reported in the \texttt{Estimate} column.
+Note that to facilitate interpretation, we arranged the contrasts in a way that we theoretically expect positive values for all four of them.
 
 \hypertarget{hypothesis-testing}{%
-\subsection{Hypothesis testing}\label{hypothesis-testing}}
-
-However, we need to think about the model parameters again when we want to test hypotheses that we have theoretically derived from some research question.
-Because the inverse-logit link is still a continuously increasing function, positive parameter values always correspond to increases in probability and vice versa.
-
-The \texttt{Fixed\ effects} table in the lme4 summary output also includes p-values for hypothesis tests with null hypotheses of the style \(H_0: \beta = 0\).
-However, for many research questions of interest, we are not interested in these two-sided tests that refer to only a single parameter.
-
-For our case study, imagine the following combined hypothesis: \emph{We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.}
-
-This combined hypothesis leads to the following four separate null hypotheses to be tested: \[
+\paragraph{Hypothesis testing}\label{hypothesis-testing}}
+
+If we chose hypothesis testing for our case study, we would test a combined null hypothesis \(H_0\) that consists of four separate null hypotheses:
+
+\[
 \begin{aligned}
-H_{01}:\ &\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0 \\
-H_{02}:\ &\beta_{a} + \beta_{c} \leq 0 \\
-H_{03}:\ &\beta_{a} + \beta_{ea} \geq 0 \\
-H_{04}:\ &\beta_{a} \geq 0
+H_{01}:\ & P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
+& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
+H_{02}:\ &P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
+& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
+H_{03}:\ &P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
+& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
+H_{04}:\ & P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
+& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
 \end{aligned}
 \]
-
-We arrive at these inequalities based on the following logic, exemplified here only for \(H_{01}\): The first null hypothesis states that \emph{an expert responding to an item while presented with correct advice has a lower or equal probability of solving the item compared to the same expert facing the same item without any advice.} This implies the following inequality for each subject \(s\) and item \(i\) \[
-inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e + \beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec}] \leq inverse\_logit[(\beta_0 + u_{0s} + u_{0i}) + \beta_e]
-\] which simplifies to \(\beta_{a} + \beta_{c} + \beta_{ea} + \beta_{ec} \leq 0\).
-
-We can specify and test hypotheses like these with the multcomp package (Hothorn, Bretz, \& Westfall, 2008) as follows:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(multcomp)}
-\NormalTok{null\_hypotheses }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"advice\_present + advice\_correct + expert:advice\_present + }
-\StringTok{   expert:advice\_correct \textless{}= 0"}\NormalTok{,}
-  \StringTok{"advice\_present + advice\_correct \textless{}= 0"}\NormalTok{,}
-  \StringTok{"{-}1 * (advice\_present + expert:advice\_present) \textless{}= 0"}\NormalTok{,}
-  \StringTok{"{-}1 * (advice\_present) \textless{}= 0"}\NormalTok{)}
-\NormalTok{glht }\OtherTok{\textless{}{-}} \FunctionTok{glht}\NormalTok{(fit, }\AttributeTok{linfct =}\NormalTok{ null\_hypotheses)}
-\FunctionTok{summary}\NormalTok{(glht, }\AttributeTok{test =} \FunctionTok{univariate}\NormalTok{())}\SpecialCharTok{$}\NormalTok{test}\SpecialCharTok{$}\NormalTok{pvalues}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## advice_present + advice_correct + expert:advice_present + expert:advice_correct 
-##                                                                     0.006407391 
-##                                                 advice_present + advice_correct 
-##                                                                     0.000000000 
-##                                   -1 * (advice_present + expert:advice_present) 
-##                                                                     0.143963670 
-##                                                           -1 * (advice_present) 
-##                                                                     0.000000000
-\end{verbatim}
-
-Because all hypotheses tested simultaneously with the \texttt{glht} function must have the same direction, we flip the sign of inequalities three and four by multiplying them with \(-1\).
-The multcomp package automatically adjusts p-values when multiple hypotheses are tested simultaneously (Hothorn et al., 2008).
-However, the combined null hypothesis in our exemplary research question should only be rejected if \textbf{all} individual null hypotheses are rejected {[}i.e., intersection-union setting; Dmitrienko and D'Agostino (2013){]}.
+The combined null hypothesis \(H_0\) should only be rejected if \textbf{all} individual null hypotheses are rejected (i.e., intersection-union setting; Dmitrienko \& D'Agostino, 2013).
 In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
-Thus, we request unadjusted p-values by setting \texttt{test\ =\ univariate()} in the \texttt{summary} command.
-With a standard significance level of \(\alpha = 0.05\), we would not reject all four null hypotheses (the p-value for hypothesis \(H_{03}\) is not significant) and therefore also not reject the combined null hypothesis for this simulated dataset.
-Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis is actually true in the population.
-
-\hypertarget{specification-of-plausible-parameter-values}{%
-\subsection{Specification of plausible parameter values}\label{specification-of-plausible-parameter-values}}
-
-When introducing our simulation function and simulating data for the above example, we have used theoretically plausible values as defaults for all model parameters (\(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\), \(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\)), but have not talked about where these numbers came from.
-
-Ideally, one would rely on meta-analytic results or conclusive data from pilot studies.
-However, these are sometimes not readily available.
-All parameter values in our present case study have been determined based on results from related prior work.
-Additionally, we had repeated discussions with our affiliated domain experts in radiology to check our assumptions.
-
-We now outline a few strategies on how to determine plausible parameter values.
-We have already seen in our discussion of model interpretation how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
-We can revert this perspective by choosing plausible probability values and deriving the parameter values implied by these probabilities (for an average subject and an average item).
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:probtable}Assumed probabilities that an average subject solves an average item in each experimental condition.}
-
-\begin{tabular}{lll}
-\toprule
-Experimental condition & $P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})$ & Implied equation\\
-\midrule
-no advice, no expert & 0.70 & $logit(0.70) = \beta_0$\\
-no advice, expert & 0.90 & $logit(0.90) = \beta_0 + \beta_e$\\
-false advice, no expert & 0.40 & $logit(0.40) = \beta_0 + \beta_a$\\
-false advice, expert & 0.85 & $logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}$\\
-correct advice, no expert & 0.90 & $logit(0.90) = \beta_0 + \beta_a + \beta_c$\\
-correct advice, expert & 0.95 & $logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}$\\
-\bottomrule
-\addlinespace
-\end{tabular}
-
-\begin{tablenotes}[para]
-\normalsize{\textit{Note.} Implied equations are derived based on the model equations and setting all random intercept terms to 0.}
-\end{tablenotes}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-Table \ref{tab:probtable} shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
-The table can be used to compute the implied values for the \(\beta\) parameters, starting with the first equation and reinserting the computed \(\beta\) values in all following equations:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{b\_0 }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.7}\NormalTok{)}
-\NormalTok{b\_e }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
-\NormalTok{b\_a }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.4}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
-\NormalTok{b\_ea }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.85}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a}
-\NormalTok{b\_c }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_a}
-\NormalTok{b\_ec }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.95}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a }\SpecialCharTok{{-}}\NormalTok{ b\_c }\SpecialCharTok{{-}}\NormalTok{ b\_ea}
-\FunctionTok{c}\NormalTok{(}\AttributeTok{b\_0 =}\NormalTok{ b\_0, }\AttributeTok{b\_e =}\NormalTok{ b\_e, }\AttributeTok{b\_a =}\NormalTok{ b\_a, }\AttributeTok{b\_c =}\NormalTok{ b\_c, }\AttributeTok{b\_ea =}\NormalTok{ b\_ea, }\AttributeTok{b\_ec =}\NormalTok{ b\_ec)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##        b_0        b_e        b_a        b_c       b_ea       b_ec 
-##  0.8472979  1.3499267 -1.2527630  2.6026897  0.7901394 -1.3928518
-\end{verbatim}
-
-It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g.~
-\[
-\begin{aligned}
-P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
-= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
-\end{aligned}
-\]
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{plogis}\NormalTok{(b\_0 }\SpecialCharTok{+}\NormalTok{ b\_e }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{+}\NormalTok{ b\_ea }\SpecialCharTok{+}\NormalTok{ b\_ec)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] 0.95
-\end{verbatim}
-
-Although the derivations above are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is \textbf{not} equal to the probability that a person with average ability solves an item with average difficulty.
-The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
-For example, we determined the \(\beta\) parameters in a way that corresponds to a desired conditional probability of \(0.95\), that an expert with average ability solves an item with average difficulty when presented with correct advice.
-However, even if the model were true, we would not observe this probability value if we estimated the marginal probability in a group of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty.
-
-The inequality of conditional and marginal effects in GLMMs (Fahrmeir et al., 2021) makes their interpretation more difficult.
-One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
-However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
-Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
-
-Earlier, we have already encountered one way to visualize the implied marginal distributions: We can fit our model to a simulated dataset and use the convenience functions from the marginaleffects package to compute averaged predictions that correspond to our quantities of interest.
-However, the model predictions will only be close to the true distribution if the simulated dataset is very large, but then the model fitting consumes a lot of time and memory.
-A more sophisticated strategy is to simulate a large dataset and directly compute the averages, contrasts and distributions we are interested in.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(tidyverse)}
-\FunctionTok{library}\NormalTok{(ggdist)}
-\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{2000}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{2000}\NormalTok{, }\AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{)}
-\NormalTok{dat }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
-    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
-    \StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
-    \StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
-    \StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ y\_prob, }\AttributeTok{y =}\NormalTok{ condition)) }\SpecialCharTok{+}
-  \FunctionTok{stat\_histinterval}\NormalTok{(}\AttributeTok{point\_interval =} \StringTok{"mean\_qi"}\NormalTok{, }\AttributeTok{slab\_color =} \StringTok{"gray45"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{breaks =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FloatTok{0.1}\NormalTok{), }\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/margdist1-1.jpeg}
-\caption{\label{fig:margdist1}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions.}
-\end{figure}
-
-Figure \ref{fig:margdist1} shows the model implied marginal distributions, including the mean, 66\% and 95\% intervals.
-We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
-This difference increases with the variability of the random effects.
-
-Up to this point, we have not talked about plausible values for the standard deviations of the subject and item random intercepts (\(\sigma_S\) and \(\sigma_I\)).
-Plots like the one above are a useful tool to decide whether the specified standard deviations are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
-
-In the next plot, we have set the item standard deviation to almost zero (\(\sigma_I = 0.01\)). This gives us a better way to see the variability between persons.
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/margdist2-1.jpeg}
-\caption{\label{fig:margdist2}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01.}
-\end{figure}
-
-As an example, Figure \ref{fig:margdist2} reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
-In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
-Patterns like these should be considered carefully and discussed with the domain experts.
-Parameter values (\(\beta\) parameters, and \(\sigma_S\)) should be adjusted if the implications do not seem reasonable.
-
-We could also have a closer look at variability between items by setting the subject standard deviation to almost zero (\(\sigma_S = 0.01\)).
-
-The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
-For Figure \ref{fig:margdist3}, we have set \(\sigma_S = 3\) and \(\sigma_I = 3\).
-This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-However, these high standard deviations do not account for the inherent variability and complexity of human performance.
-For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/margdist3-1.jpeg}
-\caption{\label{fig:margdist3}Marginal distributions including means, 66\% and 95\% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3.}
-\end{figure}
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-With all these considerations addressed, we are finally ready to perform a power analysis.
-Wrapping the \texttt{simulate} function already constructed earlier, the helper function \texttt{sim\_and\_analyse} performs all previous steps (simulate a dataset, fit a GLMM, compute p-values) in a single command.
+
+With a standard significance level of \(\alpha = 0.05\), we would not reject all four null hypotheses (the p-value in the \texttt{p\ (NonInf)} column for the third hypothesis is not significant) and therefore also not reject the combined null hypothesis for this particular (simulated) dataset.
+Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis \(H_1\) is actually true in the population.
+
+\hypertarget{interval-estimation}{%
+\paragraph{Interval estimation}\label{interval-estimation}}
+
+If we chose parameter estimation for our case study, we would focus on the two-sided confidence intervals of the four contrasts of interest.
+With a standard confidence level of \(1 - \alpha = 0.95\), plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
+Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts (which the expected behavior for binomial GLMMs).
+The smallest width is observed for the first contrast (expert with correct advice vs.~expert without advice) where both underlying probabilities are close to 1.
+The largest width is observed for the fourth contrast (non-expert with wrong advice vs.~non-expert without advice), where both underlying probabilities are closer to 0.5.
+
+\hypertarget{perform-repeated-simulations}{%
+\subsection{Perform repeated simulations}\label{perform-repeated-simulations}}
+
+\hypertarget{theory-8}{%
+\subsubsection{THEORY}\label{theory-8}}
+
+Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
+The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
+On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter values.
+When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.
+When planning for precision, the function should return the width of the confidence interval(s) of interest.
+Secondly, we must run this function repeatedly with the same sample size and population parameters.
+Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computing, that is running simulations on multiple cores of the computer at the same time to reduce total run time.
+Thirdly, the results of the repeated simulation must be collected and aggregated.
+When planning for power, we compute the relative frequency of a significant p-value(s) across repeated simulations.
+When planning for precision, we compute the average width of the confidence interval(s).
+Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted power or precision.
+
+\hypertarget{practice-8}{%
+\subsubsection{PRACTICE}\label{practice-8}}
+
+We are finally ready to run our simulation-based sample size planning analyses to plan for power and for precision.
+Wrapping the \texttt{simulate} function already constructed earlier, the helper function \texttt{sim\_and\_analyse} performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and confidence intervals) in a single command.
 
@@
   \AttributeTok{formula\_chr =} \StringTok{"y\_bin \textasciitilde{} 1 + expert + advice\_present + advice\_correct + }
-\StringTok{    expert:advice\_present + expert:advice\_correct + (1|subject) + (1|item)"}\NormalTok{, }
-  \AttributeTok{null\_hypotheses =} \FunctionTok{c}\NormalTok{(}\StringTok{"advice\_present + advice\_correct + }
-\StringTok{    expert:advice\_present + expert:advice\_correct \textless{}= 0"}\NormalTok{,}
-    \StringTok{"advice\_present + advice\_correct \textless{}= 0"}\NormalTok{,}
-    \StringTok{"{-}1 * (advice\_present + expert:advice\_present) \textless{}= 0"}\NormalTok{,}
-    \StringTok{"{-}1 * (advice\_present) \textless{}= 0"}\NormalTok{), ...)\{}
+\StringTok{    expert:advice\_present + expert:advice\_correct + (1|subject) + (1|item)"}\NormalTok{,}
+  \AttributeTok{contrasts =} \FunctionTok{c}\NormalTok{(}\StringTok{"b8 = b2"}\NormalTok{, }\StringTok{"b2 = b6"}\NormalTok{, }\StringTok{"b7 = b1"}\NormalTok{, }\StringTok{"b1 = b5"}\NormalTok{), ...)\{}
   \FunctionTok{require}\NormalTok{(lme4)}
-  \FunctionTok{require}\NormalTok{(multcomp)}
+  \FunctionTok{require}\NormalTok{(marginaleffects)}
+  \FunctionTok{require}\NormalTok{(tidyr)}
   \CommentTok{\# simulate data}
@@
 \NormalTok{  model }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(}\FunctionTok{as.formula}\NormalTok{(formula\_chr), }\AttributeTok{data =}\NormalTok{ dat, }\AttributeTok{family =} \StringTok{"binomial"}\NormalTok{)}
-  \CommentTok{\# compute p{-}values}
-\NormalTok{  glht }\OtherTok{\textless{}{-}} \FunctionTok{glht}\NormalTok{(model, }\AttributeTok{linfct =}\NormalTok{ null\_hypotheses)}
-\NormalTok{  pvalues }\OtherTok{\textless{}{-}} \FunctionTok{summary}\NormalTok{(glht, }\AttributeTok{test =} \FunctionTok{univariate}\NormalTok{())}\SpecialCharTok{$}\NormalTok{test}\SpecialCharTok{$}\NormalTok{pvalues}
-  \FunctionTok{setNames}\NormalTok{(pvalues, }\FunctionTok{paste0}\NormalTok{(}\StringTok{"p\_H0"}\NormalTok{, }\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(null\_hypotheses)))}
+  \CommentTok{\# compute contrasts}
+\NormalTok{  contr\_df }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}\AttributeTok{advice\_present =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }\AttributeTok{advice\_correct =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{,}
+    \AttributeTok{expert =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{)}
+  \FunctionTok{predictions}\NormalTok{(model, }\AttributeTok{newdata =}\NormalTok{ contr\_df, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =}\NormalTok{ contrasts, }\AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{data.frame}\NormalTok{()}
 \NormalTok{\}}
@@
 
-Power analysis can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
+Simulation-based sample size planning can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
 Thus, we use the future (Bengtsson, 2021) and furrr (Vaughan \& Dancho, 2022) packages to perform computations in parallel.
-First, we enable parallelization and specify how many parallel cores (``workers'') of our computer to use (users can find out the maximum number of cores on their computer with the command \texttt{parallel::detectCores()}), and set a seed to make the simulation reproducible.
+First, we enable parallelization with the \texttt{plan} function and specify how many parallel cores (``workers'') of our computer to use (users can find out the maximum number of cores on their computer with the command \texttt{parallel::detectCores()}), and set a seed to make the simulation reproducible.
 
@@
 \FunctionTok{library}\NormalTok{(furrr)}
-\NormalTok{sim\_design }\OtherTok{\textless{}{-}} \FunctionTok{crossing}\NormalTok{(}
+\NormalTok{sim\_result }\OtherTok{\textless{}{-}} \FunctionTok{crossing}\NormalTok{(}
   \AttributeTok{rep =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{300}\NormalTok{,}
@@
 \NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{pvalues =} \FunctionTok{future\_pmap}\NormalTok{(., sim\_and\_analyse, }
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{res =} \FunctionTok{future\_pmap}\NormalTok{(., sim\_and\_analyse, }
     \AttributeTok{.options =} \FunctionTok{furrr\_options}\NormalTok{(}\AttributeTok{seed =} \ConstantTok{TRUE}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{unnest\_wider}\NormalTok{(pvalues)}
-\end{Highlighting}
-\end{Shaded}
-
-The result of the computation is a data frame that contains the p-values of all tested hypotheses for each simulated dataset.
+  \FunctionTok{unnest}\NormalTok{(}\AttributeTok{col =}\NormalTok{ res)}
+\end{Highlighting}
+\end{Shaded}
+
+The result of this computation is a data frame that contains the p-values and confidence intervals of all specified contrasts for each simulated dataset.
 In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
@@
 
-For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our component hypotheses are significant at the \(\alpha = 0.05\) level.
+\hypertarget{power-results}{%
+\paragraph{Power results}\label{power-results}}
+
+For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the \(\alpha = 0.05\) level.
 Based on our simulation outcomes, we compute a power estimate for each combination of \texttt{n\_subjects} \(\times\) \texttt{n\_items} (including 95\% confidence intervals) and visualize the results with the following code.\footnote{This code was inspired by the ``Mixed Design Simulation'' vignette of the faux package at \url{https://debruine.github.io/faux/articles/sim_mixed.html}.}
@@
 \NormalTok{alpha }\OtherTok{\textless{}{-}} \FloatTok{0.05}
-\NormalTok{power }\OtherTok{\textless{}{-}}\NormalTok{ sim\_design }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{power }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ term, }\AttributeTok{names\_sep =} \StringTok{"\_"}\NormalTok{, }
+    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{power =} \FunctionTok{mean}\NormalTok{(p\_H01 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}\NormalTok{ p\_H02 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}
-\NormalTok{                           p\_H03 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}\NormalTok{ p\_H04 }\SpecialCharTok{\textless{}}\NormalTok{ alpha), }
-    \AttributeTok{n\_sig =} \FunctionTok{sum}\NormalTok{(p\_H01 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}\NormalTok{ p\_H02 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}
-\NormalTok{                  p\_H03 }\SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}\NormalTok{ p\_H04 }\SpecialCharTok{\textless{}}\NormalTok{ alpha),}
+  \FunctionTok{summarise}\NormalTok{(}
+    \AttributeTok{power =} \FunctionTok{mean}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b1=b5}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}
+        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b2=b6}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}
+        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b7=b1}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha), }
+    \AttributeTok{n\_sig =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b1=b5}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}
+        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b2=b6}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&}
+        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b7=b1}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha),}
     \AttributeTok{n =} \FunctionTok{n}\NormalTok{(),}
@@
 As should be the case, power estimates in Figure \ref{fig:finalpwr} increase with both the number of subjects and the number of items.
-The confidence intervals indicate how precisely power was estimated by our simulation.
+The confidence intervals reported here indicate how precisely power was estimated by our simulation.
 Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (\texttt{rep}) in the simulation.
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Experimental research requires careful planning and consideration of statistical power to ensure robust and meaningful results.
-While heuristics and user-friendly software can be useful for simple designs and models, they often fall short when more complex and customized simulations with GLMMs are required.
-The present tutorial presents a specific case study with corresponding code of how to conduct a simulation-based power analysis for experimental designs with GLMMs.
-
-\hypertarget{expected-effect-size-vs.-smallest-effect-size-of-interest-sensitivity-power-analysis}{%
-\subsection{Expected effect size vs.~smallest effect size of interest: sensitivity power analysis}\label{expected-effect-size-vs.-smallest-effect-size-of-interest-sensitivity-power-analysis}}
-
-In our case study, we have performed simulation-based power analysis from a single set of parameter values that reflect our assumptions of an expected effect size.
+\hypertarget{precision-results}{%
+\paragraph{Precision results}\label{precision-results}}
+
+When planning for precision, one could monitor the width of all four confidence intervals at the same time.
+However, because the confidence intervals of the four contrasts strongly differ in width, it is not trivial to decide which width one should target when deciding on the appropriate sample size.
+In contrast to planning for power, there are no common standards on how to specify the targeted precision.
+For our example, we use a simple heuristic but we strongly encourage readers to think about better alternatives that are appropriate in their own applications.
+
+Our simulations show that the smallest confidence interval can be expected for the first contrast (expert with correct advice vs.~expert without advice).
+The true contrast in probability for an average expert and an average item in this condition is \texttt{plogis(b\_0\ +\ b\_e\ +\ b\_a\ +\ b\_c\ +\ b\_ea\ +\ b\_ec)\ -\ plogis(b\_0\ +\ b\_e)\ =} \(0.05\).
+We want the width of this confidence interval to be smaller than 0.1.
+This would mean that if the point estimate happens to be close to the true value, the plausible values inside of a 95\% confidence interval would all be positive.
+
+Thus in our example, precision is defined as the (long-run) average width of a 95\% confidence interval for the probability contrast between experts with correct advice and experts without advice.
+Of course, lower width implies better precision.
+Based on our simulation outcomes, we compute the precision estimate for each combination of \texttt{n\_subjects} \(\times\) \texttt{n\_items} (including 95\% confidence intervals) and visualize the results with the following code.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{precision }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ term, }\AttributeTok{names\_sep =} \StringTok{"\_"}\NormalTok{, }
+    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{width =} \StringTok{\textasciigrave{}}\AttributeTok{conf.high\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{{-}} \StringTok{\textasciigrave{}}\AttributeTok{conf.low\_b8=b2}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{precision =} \FunctionTok{mean}\NormalTok{(width),}
+    \AttributeTok{ci.lwr =} \FunctionTok{t.test}\NormalTok{(width)}\SpecialCharTok{$}\NormalTok{conf.int[}\DecValTok{1}\NormalTok{],}
+    \AttributeTok{ci.upr =} \FunctionTok{t.test}\NormalTok{(width)}\SpecialCharTok{$}\NormalTok{conf.int[}\DecValTok{2}\NormalTok{], }
+    \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{)}
+\NormalTok{precision }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(n\_subjects, n\_items, }\AttributeTok{fill =}\NormalTok{ precision)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_tile}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{label =} \FunctionTok{sprintf}\NormalTok{(}\StringTok{"\%.2f }\SpecialCharTok{\textbackslash{}n}\StringTok{ [\%.2f; \%.2f]"}\NormalTok{, }
+\NormalTok{                                precision, ci.lwr, ci.upr)), }
+    \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \DecValTok{4}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\AttributeTok{direction =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{fill =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{reverse=}\ConstantTok{FALSE}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/finalprecision-1.jpeg}
+\caption{\label{fig:finalprecision}Simulation-based precision estimates (expected width of confidence intervals) including 95\% confidence interval of the case study for different numbers of subjects and items, based on a confidence level of 0.95.}
+\end{figure}
+
+As should be the case, precision estimates in Figure \ref{fig:finalprecision} increase (i.e., average width of confidence interval decreases) with both the number of subjects and the number of items.
+The confidence intervals reported here indicate how precisely the expected width of the confidence interval for our focal contrast was estimated by our simulation.
+Applying our simple heuristic of targeting an expected width smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.
+We could either choose 100 subjects and 30 items or 200 subjects and 10 items.
+
+Note that our simple heuristic for determining sample size in the planning for precision scenario was quite liberal.
+This is reflected by the result that we would need a smaller sample size than in the planning for power scenario.
+With a more conservative precision target, the result is generally the opposite: As a rule, precise parameter estimates usually require bigger samples than null hypothesis testing.
+
+\hypertarget{sensitivity-analysis}{%
+\subsection{Sensitivity analysis}\label{sensitivity-analysis}}
+
+In our case study, we have performed simulation-based sample size planning from a single set of parameter values that reflect our assumptions of an expected effect size.
 Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
 Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
-When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our \emph{expected size} is present.
-However, if the true effect is indeed smaller, the power will be lower, and the study might not be sufficiently informative.
-A common, more conservative strategy for sample size justification is to perform power analysis for the smallest effect size of interest (SESOI).
+When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our \emph{expected size} is present (or estimate the effect with satisfying precision).
+However, if the true effect is indeed smaller, power (or precision) will be lower, and the study might not be sufficiently informative.
+A common, more conservative strategy for sample size justification is to perform sample size planning for the smallest effect size of interest (SESOI).
 An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero (King, 2011).
 For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see (Lakens, 2022a).
-When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power for different effect sizes in what is called \emph{sensitivity power analysis}.
-By simulating power for different effect sizes (in addition to the different number of subjects and items), one can make sure that power would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power depends on the size of the true effect.
-In simple study designs, it is possible to perform sensitivity power analysis based on a single standardized effect size (e.g., analyze power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
+When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power or precision for different effect sizes in what is called \emph{sensitivity power analysis}.
+By simulating power (or precision) for different effect sizes (in addition to the different number of subjects and items), one can make sure that power (or precision) would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power (or precision) depends on the size of the true effect.
+In simple study designs, it is possible to perform sensitivity analysis based on a single standardized effect size (e.g., analyze power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
 However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
-In this setting, sensitivity power analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
-Power could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
+In this setting, sensitivity analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
+Power (or precision) could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
 
 \hypertarget{outlook}{%
-\subsection{Outlook}\label{outlook}}
-
-Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based power analysis in experimental research:
+\section{Outlook}\label{outlook}}
+
+Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based sample size planning in experimental research:
 
@@
 \item
-  The growing need for simulation-based power analyses in experimental research: In order to conduct well-powered research using varying complex experimental designs with GLMMs formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
+  The growing need for simulation-based sample size planning in experimental research: In order to conduct informative research using varying complex experimental designs with GLMMs, formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
 Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
 \item
-  Managing data simulations more easily with discrete predictor variables: Simulation-based power analysis becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
+  Managing data simulations more easily with discrete predictor variables: Simulation-based sample size planning becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
   This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable power estimates without compromising realistic assumptions about the data-generating process implied by the study design.
+By simplifying the simulation process, researchers can obtain reliable estimates for power or precision without compromising realistic assumptions about the data-generating process implied by the study design.
 \item
-  Teaching data simulation skills: The ability to conduct simulation-based power analysis is a valuable skill that should be taught to experimental researchers.
-By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power and improve the quality of their experimental designs.
-  Equipping researchers with the knowledge and tools to perform simulation-based power analyses enables them to make informed decisions and enhance the rigor of their studies.
+  Teaching data simulation skills: The ability to conduct simulation-based sample size planning is a valuable skill that should be taught to experimental researchers.
+By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power or precision, and improve the quality of their experimental designs.
+  Equipping researchers with the knowledge and tools to perform simulation-based sample size planning enables them to make informed decisions and enhance the rigor of their studies.
   The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions (Yarkoni, 2022).
 \item
-  Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform simulation-based a priori power analysis and the perceived effort estimated by researchers and collaborators in experimental research.
+  Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform a priori simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
 Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
@@
 \item
-  Recognizing the value of simulation-based power analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
-They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power.
+  Recognizing the value of simulation-based design analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
+They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power or precision.
 Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports (Chambers \& Tzavella, 2022).
 \item
-  Integration with Open Science and preregistration practices: Simulation-based powers analysis aligns well with the principles of Open Science and preregistration in experimental research.
+  Integration with Open Science and preregistration practices: Simulation-based sample size planning aligns well with the principles of Open Science and preregistration in experimental research.
 When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their simulation-based power analysis, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
+By preregistering their simulation-based sample size plan, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
 \end{enumerate}
@@
 In the wake of the replication crisis and myriad of underpowered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, analytic methods and software cannot be applied to conduct a priori power analyses for GLMMs necessitating data simulation-based approaches.
-Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based power analysis with GLMMs themselves.
-By incorporating GLMMs and a priori power analysis into their work, researchers can enhance the replicability and credibility of their experiments (Yarkoni, 2022).
+Yet, analytic methods and software cannot be applied to conduct a priori sample size planning for GLMMs, necessitating data simulation-based approaches.
+Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based sample size planning with GLMMs themselves.
+By incorporating GLMMs and a priori sample size planning into their work, researchers can enhance the replicability and credibility of their experiments (Yarkoni, 2022).
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cockburnThreatsReplicationCrisis2020}{}}%
-Cockburn, A., Dragicevic, P., Besançon, L., \& Gutwin, C.
-(2020).
-Threats of a replication crisis in empirical computer science.
-\emph{Communications of the ACM}, \emph{63}(8), 70--79.
-\url{https://doi.org/10.1145/3360311}
-
 \leavevmode\vadjust pre{\hypertarget{ref-R-faux}{}}%
@@
 \emph{Regression: {Models}, {Methods} and {Applications}}.
-{Berlin, Heidelberg}: {Springer Berlin Heidelberg}.
+Berlin, Heidelberg: Springer Berlin Heidelberg.
 \url{https://doi.org/10.1007/978-3-662-63882-8}
@@
 \url{https://doi.org/10.3758/BRM.41.4.1149}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gelmanBayesianWorkflow2020}{}}%
+Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B., Yao, Y., \ldots{} Modrák, M. (2020). \emph{Bayesian {Workflow}}. arXiv. Retrieved from \url{https://arxiv.org/abs/2011.01808}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hothornSimultaneousInferenceGeneral2008}{}}%
-Hothorn, T., Bretz, F., \& Westfall, P.
-(2008).
-Simultaneous {Inference} in {General Parametric Models}.
-\emph{Biometrical Journal}, \emph{50}(3), 346--363.
-\url{https://doi.org/10.1002/bimj.200810425}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kapteinUsingGeneralizedLinear2016}{}}%
-Kaptein, M.
-(2016).
-Using {Generalized Linear} ({Mixed}) {Models} in {HCI}.
-In J.
-Robertson \& M.
-Kaptein (Eds.), \emph{Modern {Statistical Methods} for {HCI}} (pp.
-251--274).
-{Cham}: {Springer International Publishing}.
-\url{https://doi.org/10.1007/978-3-319-26633-6_11}
+\leavevmode\vadjust pre{\hypertarget{ref-kelleyObtainingPowerObtaining2003}{}}%
+Kelley, K., Maxwell, S.
+E., \& Rausch, J.
+R.
+(2003).
+Obtaining {Power} or {Obtaining Precision}: {Delineating Methods} of {Sample-Size Planning}.
+\emph{Evaluation \& the Health Professions}, \emph{26}(3), 258--287.
+\url{https://doi.org/10.1177/0163278703255242}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kelleySampleSizePlanning2006}{}}%
+Kelley, K., \& Rausch, J.
+R.
+(2006).
+Sample size planning for the standardized mean difference: {Accuracy} in parameter estimation via narrow confidence intervals.
+\emph{Psychological Methods}, \emph{11}(4), 363--385.
+\url{https://doi.org/10.1037/1082-989X.11.4.363}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lundbergWhatYourEstimand2021}{}}%
+Lundberg, I., Johnson, R., \& Stewart, B.
+M.
+(2021).
+What {Is Your Estimand}?
+{Defining} the {Target Quantity Connects Statistical Evidence} to {Theory}.
+\emph{American Sociological Review}, \emph{86}(3), 532--565.
+\url{https://doi.org/10.1177/00031224211004187}
+
 \leavevmode\vadjust pre{\hypertarget{ref-maxwellSampleSizePlanning2008}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-robertsonImprovingStatisticalPractice2016}{}}%
-Robertson, J., \& Kaptein, M.
-(2016).
-Improving {Statistical Practice} in {HCI}.
-In J.
-Robertson \& M.
-Kaptein (Eds.), \emph{Modern {Statistical Methods} for {HCI}} (pp.
-331--348).
-{Cham}: {Springer International Publishing}.
-\url{https://doi.org/10.1007/978-3-319-26633-6_14}
-
 \leavevmode\vadjust pre{\hypertarget{ref-R-furrr}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-wickhamDataScienceImport2023}{}}%
+Wickham, H., Çetinkaya-Rundel, M., \& Grolemund, G.
+(2023).
+\emph{R for data science: Import, tidy, transform, visualize, and model data} (2nd edition).
+Beijing Boston Farnham Sebastopol Tokyo: O'Reilly.
+
 \leavevmode\vadjust pre{\hypertarget{ref-yarkoniGeneralizabilityCrisis2022}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-zimmerSampleSizePlanning2022}{}}%
-Zimmer, F., Henninger, M., \& Debelak, R.
-(2022).
-\emph{Sample {Size Planning} for {Complex Study Designs}: {A Tutorial} for the mlpwr {Package}}.
-{PsyArXiv}.
-\url{https://doi.org/10.31234/osf.io/r9w6t}
+Zimmer, F., Henninger, M., \& Debelak, R. (2022). \emph{Sample {Size Planning} for {Complex Study Designs}: {A Tutorial} for the mlpwr {Package}}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/r9w6t}
 

### 2024-04-14 manuscript.qmd
switch to quarto: - add manuscript.qmd and _quarto.yml - add apaquarto extension and apa.csl - update .gitignore and move results.rds to main project directory

--- 2024-04-10 docs/manuscript.Rmd
@@
 ---
-title             : "A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models"
-shorttitle        : "Simulation-Based Sample Size Planning for GLMMs"
+title: "A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models"
+shorttitle: "Simulation-Based Sample Size Planning for GLMMs"
 author:
-
-  - name          : "Timo K. Koch"
-    affiliation   : "1, 2"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Torstrasse 25, 9000 St. Gallen, Switzerland"
-    email         : "timo.koch@unisg.ch"
-    equal_contrib: yes
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - "Conceptualization"
-      - "Formal Analysis"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
-  - name          : "Florian Pargent"
-    affiliation   : "1"
-    equal_contrib: yes
-    role:
-      - "Formal Analysis"
-      - "Methodology"
-      - "Visualization"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
-  - name          : "Anne-Kathrin Kleine"
-    affiliation   : "1"
-    role:
-      - "Formal Analysis"      
-      - "Writing - Review & Editing"
-  - name          : "Eva Lermer"
-    affiliation   : "1,3"
-    role:
-      - "Conceptualization"
-      - "Resources"
-      - "Writing - Review & Editing"
-  - name          : "Susanne Gaube"
-    affiliation   : "1,4"
-    role:
-      - "Conceptualization"
-      - "Supervision"
-      - "Writing - Review & Editing"
-affiliation:
-  - id            : "1"
-    institution   : "LMU Munich, Department of Psychology"
-  - id            : "2"
-    institution   : "University of St. Gallen, Institute of Behavioral Science & Technology"
-  - id            : "3"
-    institution   : "Technical University of Applied Sciences Augsburg, Department of Business Psychology"
-  - id            : "4"
-    institution   : "University College London, Global Business School for Health"
-note: "\\clearpage" #added this to move entire author note to second page
-authornote: |
-  This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525).
-  All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF): https://osf.io/dhwf4/
-abstract: |
-  When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
+  - name: Timo K. Koch
+    corresponding: true
+    orcid: 0000-0001-6728-2063
+    email: timo.koch@unisg.ch
+    roles:
+      - conceptualization
+      - formal analysis
+      - writing
+      - editing
+    affiliations:
+      - name: University of St. Gallen 
+        department: Institute of Behavioral Science & Technology
+        address: Torstrasse 25
+        city: St. Gallen
+        country: Switzerland
+        postal-code: 9000
+      - id: "lmu"
+        name: LMU Munich
+        department: Department of Psychology
+  - name: Florian Pargent
+    orcid: 0000-0002-2388-553X
+    roles:
+      - formal analysis
+      - methodology
+      - visualization
+      - writing
+      - editing
+    affiliations:
+      - ref: lmu
+  - name: Anne-Kathrin Kleine
+    roles:
+      - formal analysis
+      - editing
+    affiliations:
+      - ref: lmu
+  - name: Eva Lermer
+    roles:
+      - conceptualization
+      - funding aquisition
+      - editing
+    affiliations:
+      - ref: lmu
+      - name: Technical University of Applied Sciences Augsburg
+        department: Department of Business Psychology
+  - name: Susanne Gaube
+    roles:
+      - conceptualization
+      - supervision
+      - editing
+    affiliations:
+      - ref: lmu
+      - name: University College London
+        department: Global Business School for Health
+author-note:
+  disclosures:
+    data-sharing: "All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF) at https://osf.io/dhwf4/"
+    conflict-of-interest: "None."
+    financial-support: "This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525)."
+    gratitude: ~
+    authorship-agreements: "Timo K. Koch and Florian Pargent contributed equally to this work."
+abstract: "When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
 However, the recent replication crisis underlines the need for more rigorous statistical methodology and well-powered designs.
@@
 Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs.
-Finally, we give an outlook on the increasing importance of simulation-based power analysis in experimental research.
-
-keywords          : "power analysis, data simulation, sample size, generalized linear mixed model"
-wordcount         : "7504"
-bibliography      : ["../Simulation Tutorial Paper.bib", "r-references.bib"]
-floatsintext      : yes
-linenumbers       : no
-draft             : no
-mask              : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-classoption       : "man"
-output: 
-  papaja::apa6_pdf:
+Finally, we give an outlook on the increasing importance of simulation-based power analysis in experimental research."
+keywords: [power analysis, data simulation, sample size, generalized linear mixed model]
+bibliography: [Simulation Tutorial Paper.bib, r-references.bib]
+floatsintext: true
+numbered-lines: false
+csl: apa.csl
 ---
 
-```{r setup, include = FALSE}
+```{r}
+#| label: setup
+#| include: false
 library("papaja")
@@
 r_refs("r-references.bib")
-```
-
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = TRUE, fig.keep = "high", dev = "jpeg")
 ```
@@
 
-```{r, echo = FALSE}
+```{r}
+#| echo: false
 b_0 = 0.847
@@
 
-```{r probtable, echo=FALSE}
+```{r}
+#| label: probtable
+#| echo: false
 probtable <- data.frame(
@@
 
-(ref:margdist1-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions.
-
-```{r margdist1, fig.cap="(ref:margdist1-caption)"}
+```{r}
+#| label: fig-margdist1
+#| fig-cap: "Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions."
+
 library(ggdist)
@@
 
-Figure \@ref(fig:margdist1) shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
+@fig-margdist1 shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
 We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
@@
 
-(ref:margdist2-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01.
-
-```{r margdist2, echo=FALSE, fig.cap="(ref:margdist2-caption)"}
+```{r}
+#| label: fig-margdist2
+#| echo: false
+#| fig-cap: "Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01."
 set.seed(1)
@@
 
-As an example, Figure \@ref(fig:margdist2) reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
+As an example, @fig-margdist2 reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
 In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
@@
 
-```{r, include=FALSE}
+```{r}
+#| include: false
 set.seed(1)
@@
 The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
-For Figure \@ref(fig:margdist3), we have set $\sigma_S = 3$ and $\sigma_I = 3$.
+For @fig-margdist3 we have set $\sigma_S = 3$ and $\sigma_I = 3$.
 This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
@@
 
-(ref:margdist3-caption) Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3.
-
-```{r margdist3, echo=FALSE, fig.cap="(ref:margdist3-caption)"}
+```{r}
+#| label: fig-margdist3
+#| echo: false
+#| fig-cap: "Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3."
 set.seed(1)
@@
 
-```{r, message=FALSE}
+```{r}
+#| message: false
 library(tidyverse)
@@
 
-```{r, message=FALSE}
+```{r}
+#| message: false
 library(lme4)
@@
 
-```{r, message=FALSE, eval=FALSE}
+```{r}
+#| message: false
+#| eval: false
 library(future)
@@
 
-```{r, message=FALSE, echo=FALSE, eval=FALSE}
+```{r}
+#| echo: false
+#| eval: false
 library(future)
@@
 
-```{r, message=FALSE, eval=FALSE}
+```{r}
+#| message: false
+#| eval: false
 library(furrr)
@@
 
-```{r, echo=FALSE}
+```{r}
+#| echo: false
 sim_result <- readRDS(file = "results.rds")
@@
 
-```{r finalpwr, fig.cap="Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."}
+```{r}
+#| label: fig-finalpwr
+#| fig-cap: "Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."
 library(binom)
@@
 
-As should be the case, power estimates in Figure \@ref(fig:finalpwr) increase with both the number of subjects and the number of items.
+As should be the case, power estimates in @fig-finalpwr increase with both the number of subjects and the number of items.
 The confidence intervals reported here indicate how precisely power was estimated by our simulation.
@@
 
-```{r finalprecision, fig.cap="Simulation-based precision estimates (expected width of confidence intervals) including 95% confidence interval of the case study for different numbers of subjects and items, based on a confidence level of 0.95."}
+```{r}
+#| label: fig-finalprecision
+#| fig-cap: "Simulation-based precision estimates (expected width of confidence intervals) including 95% confidence interval of the case study for different numbers of subjects and items, based on a confidence level of 0.95."
 precision <- sim_result %>%
@@
 
-As should be the case, precision estimates in Figure \@ref(fig:finalprecision) increase  (i.e., average width of confidence interval decreases) with both the number of subjects and the number of items.
+As should be the case, precision estimates in @fig-finalprecision increase  (i.e., average width of confidence interval decreases) with both the number of subjects and the number of items.
 The confidence intervals reported here indicate how precisely the expected width of the confidence interval for our focal contrast was estimated by our simulation.
@@
 
-\newpage
-
 # References
 
-::: {#refs custom-style="Bibliography"}
+::: {#refs}
 :::

### 2024-04-14 manuscript.qmd
add rcode.qmd notebook and experiment which equal-contributor yaml option (seems currently bugged)

@@
     corresponding: true
+    equal-contributor: true
     orcid: 0000-0001-6728-2063
@@
   - name: Florian Pargent
+    equal-contributor: true
     orcid: 0000-0002-2388-553X
@@
     data-sharing: "All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF) at https://osf.io/dhwf4/"
-    conflict-of-interest: "None."
+    conflict-of-interest: "The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article."
     financial-support: "This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525)."
@@
 ```{r}
-#| label: probtable
+#| label: tab-probtable
 #| echo: false
@@
 
-Table \@ref(tab:probtable) shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
+@tab-probtable shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
 The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:
@@
 ```{r}
+library(tidyverse)
+library(marginaleffects)
 pred_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)

### 2024-04-14 rcode.qmd
add rcode.qmd notebook and experiment which equal-contributor yaml option (seems currently bugged)

--- 2024-04-14 manuscript.qmd
@@
 ---
-title: "A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models"
-shorttitle: "Simulation-Based Sample Size Planning for GLMMs"
-author:
-  - name: Timo K. Koch
-    corresponding: true
-    equal-contributor: true
-    orcid: 0000-0001-6728-2063
-    email: timo.koch@unisg.ch
-    roles:
-      - conceptualization
-      - formal analysis
-      - writing
-      - editing
-    affiliations:
-      - name: University of St. Gallen 
-        department: Institute of Behavioral Science & Technology
-        address: Torstrasse 25
-        city: St. Gallen
-        country: Switzerland
-        postal-code: 9000
-      - id: "lmu"
-        name: LMU Munich
-        department: Department of Psychology
-  - name: Florian Pargent
-    equal-contributor: true
-    orcid: 0000-0002-2388-553X
-    roles:
-      - formal analysis
-      - methodology
-      - visualization
-      - writing
-      - editing
-    affiliations:
-      - ref: lmu
-  - name: Anne-Kathrin Kleine
-    roles:
-      - formal analysis
-      - editing
-    affiliations:
-      - ref: lmu
-  - name: Eva Lermer
-    roles:
-      - conceptualization
-      - funding aquisition
-      - editing
-    affiliations:
-      - ref: lmu
-      - name: Technical University of Applied Sciences Augsburg
-        department: Department of Business Psychology
-  - name: Susanne Gaube
-    roles:
-      - conceptualization
-      - supervision
-      - editing
-    affiliations:
-      - ref: lmu
-      - name: University College London
-        department: Global Business School for Health
-author-note:
-  disclosures:
-    data-sharing: "All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF) at https://osf.io/dhwf4/"
-    conflict-of-interest: "The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article."
-    financial-support: "This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525)."
-    gratitude: ~
-    authorship-agreements: "Timo K. Koch and Florian Pargent contributed equally to this work."
-abstract: "When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
-However, the recent replication crisis underlines the need for more rigorous statistical methodology and well-powered designs.
-Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, available methods and software for a priori power analyses for GLMMs are often limited to specific designs, while data simulation approaches offer more flexibility.
-Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs.
-Finally, we give an outlook on the increasing importance of simulation-based power analysis in experimental research."
-keywords: [power analysis, data simulation, sample size, generalized linear mixed model]
-bibliography: [Simulation Tutorial Paper.bib, r-references.bib]
-floatsintext: true
-numbered-lines: false
-csl: apa.csl
+title: "R code from the practice sessions"
 ---
 
-```{r}
-#| label: setup
-#| include: false
-library("papaja")
-library("tidyverse")
-library("faux")
-library("lme4")
-library("multcomp")
-library("marginaleffects")
-library("future")
-library("furrr")
-library("binom")
-r_refs("r-references.bib")
-```
-
-# Introduction
-
-<!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
-When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data [@lakensSampleSizeJustification2022].
-However, the recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of findings [@yarkoniGeneralizabilityCrisis2022].
-As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
-While easy-to-use software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) sample size planning for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
-In the present work, we provide a tutorial on performing tailored simulation based sample size planning for GLMMs.
-After introducing general concepts, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
-Then we describe general steps and decisions that are necessary whenever performing tailored data simulation.
-Because the details of these steps vary considerably for each specific study, we finish with a case study from the field of human-AI (artificial intelligence) interaction research.
-
-<!-- <!-- Old Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! --> -->
-<!-- Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs.
-Thereby, we aim to equip researchers with the tools needed to simulate data and determine appropriate sample sizes for their own research.
--->
-
-# Prior Knowledge
-
-TODO: fertig ausformulieren
-
-Simulation-based sample size planning for generalized linear models is a complex topic that touches on a variety of statistical concepts and programming skills.
-Although we give some short introduction to clarify our terminology, to profit the most from our tutorial, our readers require prior knowledge in the following topics (or refer to our collection of background literature):
-
-Readers should understand basic statistical concepts like hypothesis tests and their statistical power but also confidence intervals and their precision.
-Some background knowledge in empirical estimands, statistical estimates, and causal inference is very useful but not absolutely necessary.
-
-In addition, readers should have basic programming skills in R on how to simulate data and how to repeat these simulations in parallel to speed up computation.
-While there are a myriad of options for data simulation in R beyond the basic functionality in the base and stats packages, in our tutorial we use functions from the tidyverse and the faux package.
-
-Finally, readers should be familiar with regression modeling in general and GLMMs in particular.
-In our tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process.
-While it is not necessary to understand the technical details of how GLMMs are estimated, it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the values in the dependent variable are determined by the predictor variables and the random effects.
-
-Literature:
-
-- Lakens ebook [@lakensImprovingYourStatistical2022] evtl. Andy Field für die Basics?
-- R for data science [@wickhamDataScienceImport2023], Bruine article [@debruineUnderstandingMixedEffectsModels2021], faux documentation [@R-faux] für die R skills
-- kumle power tutorial [@kumleEstimatingPowerGeneralized2021] und irgendwelche paper (am besten aus AMPPS) zu GLMMs
-- what is your estimand? paper [@lundbergWhatYourEstimand2021] für das estimand topic
-
-# General concepts
-
-## Planning for statistical power or precision
-
-Conducting research with too small sample sizes can have many negative consequences.
-First, experiments may yield inconclusive or misleading results, hindering the accumulation of knowledge.
-Second, studies that are doomed to never finding a postulated effect due to small sample size waste resources by consuming time, effort, and funding without delivering meaningful results.
-For such reasons, many journals and funding agencies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-Although many pragmatic justifications are possible [@lakensSampleSizeJustification2022], the ideal is often considered to determine a suitable sample size a priori (i.e., before the study is conducted) based on some meaningful computation to assure that the study will be able to fulfill its purpose.
-In this way, adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
-
-<!-- <!-- advantages of an a priori power analysis --> -->
-<!-- Thereby, a priori power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
--->
-<!-- Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
--->
-<!-- Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
--->
-<!-- Adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research. -->
-<!-- Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
--->
-
-The majority of empirical studies in psychology applies hypothesis testing.
-As a consequence, the dominant approach for sample size planning is based on power analysis (i.e., planing for power).
-However, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision).
-
-### Planning for power
-
-<!-- Some basics on statistical power -->
-In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022].
-Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
-Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered".
-
-When planning for power, a target is set for the statistical power of a hypothesis test of interest (a very common heuristic is 0.8).
-Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that (on average) would guarantee reaching this target.
-
-### Planning for precision
-
-Not all research questions are best answered using hypothesis testing.
-Especially for exploratory studies when little previous research has been conducted, only estimating effects of interest is often more useful.
-When no hypothesis tests are conducted when analysing the planned study, power analysis is not not available for sample size planning.
-Nonetheless, sample size still has a crucial effect on how informative the planned study will be to learn about the effect of interest.
-In this case, the expected width of a confidence interval is used as a target for sample size planning.
-The values inside a confidence interval are commonly interpreted as plausible values for the quantity of interest the confidence interval is supposed to estimate.
-More formally, a 95%-confidence interval provides the smallest interval with the property that upon repeated sampling, 95% of individual confidence intervals would include the true quantity of interest.
-Thus, a narrow confidence interval (assuming the confidence level is kept at a high value) is more informative about the size of the true effect than a wide interval.
-Apart from the desired confidence level, the width of a confidence interval depends strongly on the sample size.
-Because bigger samples carry more information, they lead to smaller confidence intervals (assuming everything else stays the same).
-When planning for precision, a target is set for the expected width of a confidence interval of interest.
-Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that (on average) would guarantee reaching this target.
-
-## Generalized linear mixed models (GLMMs)
-
-<!-- Advantages of GLMMs over "simpler" models -->
-As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
-
-<!-- intro into GLMMs -->
-GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data including hierarchical structures  [@fahrmeirRegressionModelsMethods2021].
-In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
-This correlation can arise for various reasons: Responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), both participants come from the same neighborhood (clustering) or both participants responded to the same stimulus (stimulus effects).
-Thus, modeling such correlations is especially important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
-LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on all fixed and random effects).
-They allow for the modeling of fixed effects, which capture the relationships between our usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects exemplified above.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
-
-As mentioned, GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
-As an example, think of an experiment with different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
-Here, participants’ behavior is measured repeatedly (e.g., over several sessions).
-The click patterns of participants in one session are likely to be correlated with their previous sessions.
-Finally, the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
-
-# Simulation based sample size planning with GLMMs
-
-For simpler statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori sample size planning is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018; @kelleyObtainingPowerObtaining2003; @kelleySampleSizePlanning2006].
-However, these software packages are often not flexible enough to perform sample size planning for complex designs.
-
-Existing approaches for sample size planning with GLMMs have almost exlusively focused on planning for power. 
-Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
-Currently available formula-based software packages for power analysis with multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
-A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
-This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
-The necessary presumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
-The distributional assumption specifies the distributional family for the outcome variable (when conditioned on all fixed and random effects).
-Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e.,  if and how they are correlated).
-Interpreting these presumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
-Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge [see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021].
-
-- Insert table on existing R packages for power analysis with GLMMs
-
-# When do we profit from tailored data simulation?
-
-Performing tailored simulation based sample size planning is more complicated and time consuming than using existing software tools for sample size planning.
-Thus, it is worth discussing under which circumstances we most profit from or are forced to perform tailored simulation based sample size planning, because existing software solutions do not meet our requirements.
-
-## Complex study designs
-
-TODO: fertig ausformulieren
-
-Real studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
-
-### Missing data
-
-TODO: fertig ausformulieren
-
-Missing data is a frequent issue in applied data analysis.
-There can be various reasons for why data is missing.
-For example data can be missing completely at random.
-Alternatively, subjects might drop out or produce missing data as a result of some covariate also measured in the dataset.
-A a different note, many experimental designs contain some conditions in which some values of the predictor variables are missing by design.
-Missing by design conditions can make data analysis more complicated because predictors have to be coded in highly specific ways that prevent the estimated GLMM to become unidentified.
-Whether missing data has an effect on the result of sample size planning depends on the characteristics of the missing data process.
-However, except for very simple scenarios, deciding whether missing data can be safely ignored in the sample size planning is often difficult to decide on a theoretical basis.
-A better strategy is to explicitly include the (assumed) process of missing data into the data simulation process.
-
-### Data generating processes beyond standard GLMMs
-
-TODO: fertig ausformulieren
-
-While GLMMs are a very flexible model class that can handle a large variety of outcome variables, psychological researchers are becoming increasingly aware that many psychological datasets might profit from even more sophisticated models.
-Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects.
-
-- possibility to extend beyond standard GLMMs (e.g. zero-inflated outcomes, censoring, nonlinear effects, …)
-
-## Complex statistical hypotheses
-
-TODO: fertig ausformulieren
-
-The most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is some coefficient in a regression model.
-However, many research questions in psychology actually require testing more complex statistical hypotheses.
-
-### Directed hypotheses and testing against values other than 0
-
-TODO: fertig ausformulieren
-
-In the new era of preregistration and registered reports, most research questions are tested with directed hypotheses, because good theories should at least postulate whether some psychological effect of interest is positive or negative.
-Even better theories should be able to specify a smallest effect sizes of interest (SESOI) that must be exceeded if the psychological effect has some practical relevance in a specific application.
-In combination, this might require a test like $H_0: \beta \leq 0.1$.
-
-### Combined hypotheses
-
-TODO: fertig ausformulieren
-
-Elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example $H_0: \beta_0 + \beta_1 \leq 0$.
-If the research question consists only of a single hypothesis of this sort, it is often possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
-However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypotheses.
-For example, a combined null hypothesis $H_0$ might consist of two single null hypotheses $H_{01}: \beta_1 \leq 0$ and $H_{02}: \beta_0 + \beta_1 \leq 0$.
-For some research questions, the combined null hypothesis $H_0$ would be rejected if both $H_{01}$ *AND* $H_{02}$ are rejected.
-For other research questions, the combined null hypothesis $H_0$ would be rejected if $H_{01}$ *OR* $H_{02}$ *OR* both are rejected.
-If the global hypothesis $H_0$ is combined with *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis.
-However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global hypothesis test.
-
-None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here.
-In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation based sample size planning.
-
-## Planning for precision
-
-TODO: fertig ausformulieren
-
-As discussed earlier, the two major frameworks for sample size planning are planning for power and planning for precision.
-Although planning for precision has been increasingly discussed as a more suitable strategy for exploratory research, all available software packages for sample size planning with GLMMs are based on power analysis.
-
-As long as one were interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
-However, we demonstrate in our case study that in realistic research projects with GLMMs, the contrasts of interests and their respective confidence intervals are usually situated on the scale of the (expected) outcome in contrast to the scale of the model parameters.
-We will show that tailored simulation based sample size planning can easily handle the planning for precision scenario.
-The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset and the expected width is estimated.
-
-## No prior studies or pilot data
-
-TODO: fertig ausformulieren
-
-All frameworks for sample size planning require the user to make assumptions about the expected effect size.
-Assuming the true effect is of this size (or greater), one can compute the (minimum) power of a hypothesis test or the (maximum) expected width of a confidence interval.
-Existing software packages for sample size planning for GLMMs usually require to provide the assumed effect in the unit of some standardized measure of effect size.
-When the researcher has access to similar studies or pilot data, providing such standardized effect sizes is feasible.
-However, this requirement can be an almost impossible challenge when no prior studies of pilot data are available.
-This problem is further exacerbated by the fact that GLMMs are so flexible, that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
-In our experience, using domain knowledge to construct a tailored data simulation is the only solution to determine plausible effect sizes in the absence of prior evidence.
-It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
-However, we would argue that when tailored data simulations are necessary anyways, performing the whole desing analysis in a customized way is preferred over using the existing software packages.
-
-# General steps in tailored simulation-based sample size planning
-
-Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions that we will outline in the remainder of our tutorial.
-Will introduce each step in a theoretical section, followed by the practical application based on a case study.
-All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
-
-## Define the estimand
-
-### THEORY
-
-The first step in every research process should always be a clear definition of which theoretical quantity is necessary to answer the specific research question of the study under planning.
-Unfortunately, many psychological researchers are still not well trained in this task.
-The theoretical estimand is a unit specific quantity defined outside of any statistical model [@lundbergWhatYourEstimand2021].
-Depending on the research and whether the research question is causal, this quantity might be counterfactual and not directly observable.
-Part of the theoretical estimand is also a clear definition of the target population, for which the quantity is of interest.
-
-When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reformulate the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
-For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantities within the scope that can be estimated with a regression model.
-Note however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression (CITATION).
-
-When the empirical estimand has been specified, the last step is to clearly define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
-When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique, and software package is used to fit the model, and how to compute hypothesis tests or confidence intervals.
-
-### PRACTICE
-
-In the present case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) in an AI-enabled diagnostic decision support system.
-The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography.
-In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans that are required to make the final diagnostic decision.
-The research goal is to validate the effectiveness of the AI-enabled advice.
-We consider the AI-enabled advice as effective, if the following pattern holds, which we will first describe verbally:
-
-*We expect that for BOTH expert radiologists and students under training, correct AI-advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI-advice presented, AND, we expect that for BOTH experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
-
-I becomes clear that our theoretical estimand is be a causal quantity, based on several individual quantities that are defined outside of any statistical model [@lundbergWhatYourEstimand2021].
-To elaborate further, we focus on only one part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists.
-For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice were presented and the probability that a correct diagnosis is made if no AI-advice were presented.
-Mathematically, we could describe this unit-specific quantity with some pseudo-equation like this (where $s$ stands for a specific expert radiologist and $i$ stands for a specific CT scan):
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, specific expert, specific scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, specific expert, specific scan})
-\end{aligned}
-$$
-However, our verbal description is not precise enough yet, because the above quantity may differ for each combination of expert and scan.
-We must specify how we want to aggregate the quantity across different experts and scans.
-If we would compute the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts (CITATION).
-Although often debated whether this is a good idea, psychological research often focuses on a different quantity:
-Instead of averaging over experts and scans, we focus in a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
-With this definition, the theoretical estimand in our above example can be expressed as:
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
-\end{aligned}
-$$
-In total, our theoretical estimand consists of four such expressions, the remaining three are:
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan})
-\end{aligned}
-$$
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, average student, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, average student, average scan})
-\end{aligned}
-$$
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan})
-\end{aligned}
-$$
-To complete our definition of the theoretical estimand, we have to clearly define out target population that consists of both persons and scans:
-With respect to persons, we are only interested in expert radiologists and medical students at German universities.
-With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
-Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system under study.
-
-While a theoretical estimand is defined outside of any statistical model and can be impossible to be observe directly, an empirical estimand is a quantity that can be estimated based on observed data.
-For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of brain scans, but the kind of AI-advice given for each brain scan is randomly assigned within participants.
-This random intervention allows to estimate our theoretical estimand, although in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each brain scan.
-For our example, each probability expression in our theoretical estimand can be modeled with the same GLMM.
-Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability and these estimates can be used to compute an estimate for each of the four probability contrasts.
-For pedagogical reasons, we will skip the concrete definition of our empirical estimand until we have discussed how to simulate data based on a concrete GLMM in the next section.
+::: {.callout-note}
+This Notebook contains only the R code from the practice sessions of the manuscript.
+The number of samples and repetitions for all simulations have been greatly reduced to ensure that the code runs quickly and with limited computational resources.
+:::
 
 ## Simulate the data generating process
-
-#### THEORY
-
-When the estimand has been defined, the next step in the research process is to write code that simulates the (assumed) data generating process of the planned study.
-This requires specifying a generative process for all predictor variables used in the final data analysis.
-While such assumptions can be quite challenging for observational studies and/or continuous predictor variables, this is less of a problem for experimental studies with only categorical predictor variables.
-When all predictor variables have been simulated, one can use the structure of a suitable GLMM to simulate the dependent variable.
-To simulate the GLMM, one requires plausible values for all model parameters.
-We will discuss strategies on how these values can be obtained later.
-Because one has full control over the data generating process in a tailored simulation, it is simple to also model highly specific aspects of the planned study, like data missing by design or assuming that subjects drop out (i.e.
-deleting their data from the simulated dataset) based on other variables in the simulated dataset.
-The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
-For this reason, deciding on a specific set of assumptions can easily feel overwelming.
-However, we would argue that even a strongly simplified data generating process (e.g.
-including only a small number of interaction effects; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current status quo in psychological research, where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.
-
-#### PRACTICE
-
-In our case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated. 
-<!-- The goal is to understand how AI advice influences medical decision-making. -->
-Participants, radiologists (task experts) and students/interns (non experts), review a series of head computer tomography (CT) scans to assess the presence of a bleeding. 
-To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
-When AI advice is given, this advice can be either correct or incorrect.
-The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across brain scans.
-After reviewing a CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
-This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-<!-- With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
--->
-In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible. 
-The goal of simulation-based sample size planning is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power or precision in the planned experiment.
-
-##### Our specific GLMM
-
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
-$$
-g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$
-Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: 
-$$
-E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
-$$
-If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability:
-$$
-P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
-$$
-In our case study, $P_{si}$ is the conditional probability that subject $s$ gives the correct response to item (i.e., CT scan) $i$.
-
-In such a setting, we model this probability as
-$$
-P_{si} = \text{inverse\_logit}(\eta_{si})
-$$
-with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
-$$
-\text{logit}(P_{si}) = \eta_{si}
-$$
-with the logit link $g(P_{si}) = \text{logit}(P_{si}) = \text{ln} (\frac{P_{si}}{1 - P_{si}})$.
-
-In our case study, the probability of making an accurate diagnostic decision is assumed to depend on the predictors:
-
-- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to assess item $i$ 
-- $advice\_correct_{si}$: whether this advice was correct (1) or not (0) 
-- $expert_s$: whether subject $s$ was a task expert (1) or not (0)
-
-and the random effects:
-
--   $u_{0s}$: the deviation of subject $s$ from the average ability to solve an item (i.e., CT scan) with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
--   $u_{0i}$: the deviation of item (i.e., CT scan) $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$
-
-In total, we assume the model
-$$
-\begin{aligned}
-\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
-&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
-\end{aligned}
-$$
-or equivalently
-$$
-\begin{aligned}
-P_{si} = \text{inverse\_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
-&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
-\end{aligned}
-$$
-with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
-
-In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
-To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.
-
-##### Simulation function in R
-
-The following R function simulates a full dataset structured according to the design of our case study.
-The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.
-
-```{r}
-#| echo: false
-b_0 = 0.847
-b_e = 1.350
-b_a = -1.253
-b_c = 2.603
-b_ea = 0.790
-b_ec = -1.393
-```
 
@@
 
-In the first six lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data. 
-In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).
-Thus, the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
-We add a between-subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
-The next two lines simulate that in $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
-Next, we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
-As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user.
-With all design variables done, we are ready to simulate our model equation outlined in the last section.
-The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects, and model parameters as assumed by our model.
-We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
-In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject makes an accurate diagnostic decision for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
-
-## Strategies on specifying the population parameters
-
-### THEORY
-
-Tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with the same design or pilot data.
-In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
-Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
-In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available for GLMMs.
-Our strategies to specify population parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation based sample size planning.
-The basic idea of all strategies is how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
-
-Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
-In this literature, strategies have been developed to monitor the plausibility of (prior) model assumptions [@gelmanBayesianWorkflow2020], that can be adapted for our purposes.
-
-### PRACTICE
-
-When introducing the simulation function for our case study, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
-
-Ideally, one would rely on meta-analytic results or conclusive data from pilot studies.
-However, these are sometimes not readily available.
-All parameter values in our present case study have been determined based on results from distantly related study designs in the literature.
-Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions seem plausible.
-
-We now outline our main strategy to determine plausible parameter values for the fixed effects ($\beta$ parameters):
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret in isolation because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation, that allows us to ignore the random effects for now, works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
-Then the model implied probability that such a person solves such an item accurately is given by:
-
-$$
-\begin{aligned}
-P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
-= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
-\end{aligned}
-$$
-In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
-In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
-$$
-\begin{aligned}
-P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
-= \text{inverse\_logit}[\beta_0]
-\end{aligned}
-$$
-We can revert this perspective by choosing plausible probability values based on domain knowledge and deriving the parameter values implied by these probabilities for each experimental condition.
-
-```{r}
-#| label: tab-probtable
-#| echo: false
-probtable <- data.frame(
-  col1 = 
-    c("no advice, no expert", "no advice, expert", "false advice, no expert", 
-      "false advice, expert", "correct advice, no expert", "correct advice, expert"),
-  col2 =
-    c("0.70", "0.90", "0.40", "0.85", "0.90", "0.95"),
-  col3 = 
-    c("$logit(0.70) = \\beta_0$",
-      "$logit(0.90) = \\beta_0 + \\beta_e$", 
-      "$logit(0.40) = \\beta_0 + \\beta_a$",
-      "$logit(0.85) = \\beta_0 + \\beta_e + \\beta_{a} + \\beta_{ea}$",
-      "$logit(0.90) = \\beta_0 + \\beta_a + \\beta_c$",
-      "$logit(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
-apa_table(
-  probtable
-  , col.names = c("Experimental condition",
-    "$P(Y=1|\\mathbf{X=x}, \\mathbf{U} = \\mathbf{0})$",
-    "Implied equation")
-  , caption = "Assumed probabilities that an average subject solves an average item in each experimental condition."
-  , note = "Implied equations are derived based on the model equations and setting all random intercept terms to 0."
-  , escape = FALSE
-)
-```
-
-@tab-probtable shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
-The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:
-
 ```{r}
@@
 
-It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g. 
-$$
-\begin{aligned}
-P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
-= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
-\end{aligned}
-$$
-
-which we compute in R as:
-
 ```{r}
@@
 
-This leaves us with the question on how to determine plausible values for the two remaining model parameters ($\sigma_S$, and $\sigma_I$) that are the standard deviations for the random intercepts.
-For this, we introduce two more strategies.
-
 ### Insightful descriptive statistics
-
-#### THEORY
-
-<!-- For most GLMMs, the interpretation of model parameters are quite complicated which makes it difficult to specify plausible values for a specific application. -->
-The mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
-The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
-For example, domain experts might not be able to directly choose plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
-However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.
-which descriptive statistics they expect to observe (on average) when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
-The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
-Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
-
-#### PRACTICE
-
-In the last section, we showed how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
-Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty [@fahrmeirRegressionModelsMethods2021].
-The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
-
-For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice (the conditional perspective).
-However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty (the marginal perspective).
-How much the two probabilities differ depend on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations would be zero).
-We want to use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
-
-We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model implied true values) and simply compute the relative frequency of correct diagnoses for each experimental condition.
 
@@
 set.seed(1)
-dat <- simulate(n_subjects = 3000, n_items = 3000,
+dat <- simulate(n_subjects = 500, n_items = 500,
   sd_u0s = 0.5, sd_u0i = 0.5)
@@
 
-We tried using these descriptive statistics to judge together with our domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
-However although they result was deemed plausible, in our example these statistics are not informative enough to determine a final set of plausible parameter values.
-For this reason, we will additionally look at insightful model based quantities.
-
 ### Insightful model based quantities
-
-#### THEORY
-
-Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
-This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
-An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
-For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
-Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
-For example, the domain expert might deem it unplausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
-
-#### PRACTICE
-
-The discussed inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
-One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
-However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
-Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
-
-In the last section, we simulated a large simulated dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
-We will now use the model implied probability of each simulated data point (stored in the variable `y_prob`) to visualize the whole model implied marginal distribution of correct diagnoses for each experimental condition.
 
@@
 
-@fig-margdist1 shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
-We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
-This difference increases with the variability of the random effects.
-
-We can use plots like the one above as a useful tool to decide whether the specified standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
-
-In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$). This gives us a better way to see the variability between persons.
-
 ```{r}
 #| label: fig-margdist2
-#| echo: false
 #| fig-cap: "Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01."
 set.seed(1)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0i = 0.01)
+dat <- simulate(n_subjects = 500, n_items = 500, sd_u0i = 0.01)
 dat %>% 
@@
 
-As an example, @fig-margdist2 reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
-In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
-Patterns like these should be considered carefully and discussed with the domain experts.
-Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
-
-We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
-
 ```{r}
@@
 set.seed(1)
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
+dat <- simulate(n_subjects = 500, n_items = 500, sd_u0s = 0.01)
 dat %>% 
@@
 
-The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
-For @fig-margdist3 we have set $\sigma_S = 3$ and $\sigma_I = 3$.
-This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-However, these high standard deviations do not account for the inherent variability and complexity of human performance.
-For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
-
 ```{r}
@@
 set.seed(1)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 3, sd_u0i = 3)
+dat <- simulate(n_subjects = 500, n_items = 500, sd_u0s = 3, sd_u0i = 3)
 dat %>% 
@@
 
-### Iterative process with domain experts
-
-#### THEORY
-
-In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
-In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
-As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
-For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
-Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
-Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
-During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
-
-##### PRACTICE
-
-All parameter values in our present case study have been determined based on repeated discussions with our affiliated domain experts in Radiology to validate our assumptions.
-Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for intracranial hemorrhage.
-Existing studies indicate that radiologists typically demonstrate high accuracies, often exceeding or hovering around 90%, while interns have been shown to perform below 80%, and medical students fall even shorter.
-For simplicity, we assumed plausible probability values of .90 for experts and .70 for non-experts, respectively.
-Our experts confirmed that these values are realistic baselines for reviewing diverse head CT images without AI assistance.
-Subsequently, we consulted several published papers investigating the effect of correct and incorrect advice on decision-making performance in other settings.
-From their findings, we inferred that both experts and non-experts should benefit from correct and suffer losses from incorrect advice.
-However, the magnitude of these effects should be substantially greater for non-experts, given their demonstrated reliance on advice compared to experts.
-We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
-For our simulation, we used the probabilities of average participants to solve an average case, as shown in Table 1.
-
 ## Estimate the statistical model
-
-### THEORY
-
-At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset that will later be collected in the planned study.
-The next step is to specify how the statistical model shall be estimated in the actual study collected later.
-This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
-Note that this does not always mean that one will specify the same GLMM that was used when specifying the data generating process.
-On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
-On the other hand, using a more complex model (that perhaps even goes beyond a standard GLMM) for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
-
-### PRACTICE
-
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
-The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
-
-[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
-
-However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore, for our case study, we have written our own function to simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions).
-We think this choice assists applied researchers in better understanding all model assumptions before using lme4 to analyze the simulated datasets.[^2]
-
-[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
-
-In this section, we show how to fit a GLMM with lme4.
-We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
-However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.
 
@@
 ```
-
-The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:
 
@@
 
-The first two lines look similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`).
-The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
-The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
-
-In lme4, a GLMM is fitted with the `glmer` function.
-By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs.
-inaccurate (0) diagnosis.
-
 ```{r}
@@
 ```
-
-We can inspect the estimates for all model parameters with the `summary` command:
 
@@
 
-In the lme4 output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
-
 ## Compute the estimate
-
-### THEORY
-
-When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
-While the majority of psychological research uses hypothesis testing, there are repeated claims that a focus on estimation can also be a good choice in many circumstances (CITATION).
-This decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests and confidence intervals with the marginaleffects R package using the delta method).
-
-### PRACTICE
-
-For our case study, we investigate the following research question:
-*We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
-
-In the earlier section on our theoretical estimand we translated this verbal description into four probability statements that are specified outside of any specific statistical model.
-The next translation step is to specify the empirical estimand, that is how these probabilites can be estimated within the scope of our concrete GLMM.
-
-The first of the four statements of the theoretical estimand was:
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
-\end{aligned}
-$$
-In the context of our GLMM, the corresponding probability statement is:
-
-$$
-\begin{aligned}
-& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
-& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
-\end{aligned}
-$$
-Plugging in the model equation of the GLMM produces an equation on how to compute this contrast when all model parameters are known.
-We have already discussed how to compute the individual probabilities in the section on specifying population parameters.
-When we want to estimate the above contrast based on observed data, the only difference is that model parameters are not known and we instead plug the corresponding parameter estimates into the equation.
-
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- & \text{omv}[\beta_0 + \beta_a \cdot 1 + \beta_c \cdot 1 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 1 + \beta_{ec} \cdot 1 \cdot 1]\\ -->
-<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_a \cdot 0 + \beta_c \cdot 0 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 0 + \beta_{ec} \cdot 1 \cdot 0] = \\ -->
-<!-- & = \text{inverse_logit}[\beta_0 + \beta_a  + \beta_c + \beta_e + \beta_{ea} + \beta_{ec} ]\\ -->
-<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_e ] \\ -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-
-We could use our knowledge of the structure of our GLMM to determine the exact formula needed to compute our contrast of interest and then plug in the parameter estimates manually from the `summary(fit)` output.
-However, this would be very tedious and we can use R to compute this contrast without doing the math.
-
-We could use the `predict` function of the lme4 package to compute the predicted probability for a correct diagnosis based on our model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
 
@@
 
-The argument `type = "response"` specifies that predictions are made on the probability scale (instead of the log-odds scale of the $\beta$ parameters), while `re.form = NA` sets all random effects to 0.
-
-We could use this method to compute point estimates for all four contrast that are part of our empirical estimand.
-However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute hypothesis tests or confidence interval for our point estimates.
-
-The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute hypothesis tests and confidence intervals for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
-
-First, we specify a grid of all combinations of predictor variable and then compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
-
 ```{r}
@@
 ```
-
-The point estimates for all experimental conditions are reported in the `Estimate` column.
-Note that the output also contains the two missing by design conditions that will never be observed in the actual study ($advice\_present = 0, advice\_correct = 1, expert = 1$ and $advice\_present = 0, advice\_correct = 1, expert = 0$).
-This is no problem as long as we never interpret those estimates.
-
-Next, we use the estimated probabilities to compute the four specific contrasts that are part of our empirical estimand.
-For this we must specify which rows in `probs` have to be subtracted from each other.
-We will use the `hypotheses` function to compute our four contrasts of interest together with hypothesis tests and confidence intervals.
-We use the default inference options of the marginaleffects package that compute hypothesis tests and confidence intervals based on the approximate delta method.
 
@@
 
-The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2.
-The argument `equivalence = c(0, 0)` can be used to compute one-sided p-values, testing whether the contrast in the population is smaller than 0 (`p (NonSub)` column) or greater than 0 (`p (NonInf)` column).
-The point estimates for four contrasts are reported in the `Estimate` column.
-Note that to facilitate interpretation, we arranged the contrasts in a way that we theoretically expect positive values for all four of them.
-
-#### Hypothesis testing
-
-If we chose hypothesis testing for our case study, we would test a combined null hypothesis $H_0$ that consists of four separate null hypotheses:
-
-$$
-\begin{aligned}
-H_{01}:\ & P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
-& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
-H_{02}:\ &P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
-& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
-H_{03}:\ &P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
-& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
-H_{04}:\ & P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
-& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
-\end{aligned}
-$$
-The combined null hypothesis $H_0$ should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting\; @dmitrienkoTraditionalMultiplicityAdjustment2013].
-In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
-
-With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value in the `p (NonInf)` column for the third hypothesis is not significant) and therefore also not reject the combined null hypothesis for this particular (simulated) dataset.
-Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis $H_1$ is actually true in the population.
-
-#### Interval estimation
-
-If we chose parameter estimation for our case study, we would focus on the two-sided confidence intervals of the four contrasts of interest.
-With a standard confidence level of $1 - \alpha = 0.95$, plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
-Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts (which the expected behavior for binomial GLMMs).
-The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice) where both underlying probabilities are close to 1.
-The largest width is observed for the fourth contrast (non-expert with wrong advice vs. non-expert without advice), where both underlying probabilities are closer to 0.5.
-
-<!-- Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
--->
-
-<!-- With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects): -->
-
-<!-- (ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package. -->
-
-<!-- ```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"} -->
-<!-- library(marginaleffects) -->
-<!-- plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"), -->
-<!--   type = "response") + ylim(c(0.3, 1)) -->
-<!-- ``` -->
-
 ## Perform repeated simulations
-
-### THEORY
-
-Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
-The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
-On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter values.
-When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.
-When planning for precision, the function should return the width of the confidence interval(s) of interest.
-Secondly, we must run this function repeatedly with the same sample size and population parameters.
-Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computing, that is running simulations on multiple cores of the computer at the same time to reduce total run time.
-Thirdly, the results of the repeated simulation must be collected and aggregated.
-When planning for power, we compute the relative frequency of a significant p-value(s) across repeated simulations.
-When planning for precision, we compute the average width of the confidence interval(s).
-Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted power or precision.
-
-### PRACTICE
-
-We are finally ready to run our simulation-based sample size planning analyses to plan for power and for precision.
-Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and confidence intervals) in a single command.
 
@@
 
-Simulation-based sample size planning can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
-Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
-First, we enable parallelization with the `plan` function and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
-
 ```{r}
 #| message: false
-#| eval: false
 library(future)
-plan("multisession", workers = 6)
+plan("multisession", workers = 4)
 set.seed(2)
-```
-
-The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
-We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
-With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
-If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.
-
-```{r}
-#| echo: false
-#| eval: false
-library(future)
-library(furrr)
-plan("multisession", workers = 6)
-set.seed(2)
-sim_result <- crossing(
-  rep = 1:300,
-  n_subjects = c(100, 150, 200, 250),
-  n_items = c(10, 30, 50, 70)
-) %>%
-  mutate(res = future_pmap(., sim_and_analyse, 
-    .options = furrr_options(seed = TRUE))) %>%
-  unnest(col = res)
-saveRDS(sim_result, file = "results.rds")
 ```
@@
 #| message: false
-#| eval: false
 library(furrr)
 sim_result <- crossing(
-  rep = 1:300,
+  rep = 1:5,
   n_subjects = c(100, 150, 200, 250),
@@
 
-The result of this computation is a data frame that contains the p-values and confidence intervals of all specified contrasts for each simulated dataset.
-In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
-When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process.
-We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
-
 #### Power results
-
-For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^3]
-
-[^3]: This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
-
-```{r}
-#| echo: false
-sim_result <- readRDS(file = "results.rds")
-```
 
@@
 
-<!-- How many participants/stimuli are needed for concrete example -->
-
-As should be the case, power estimates in @fig-finalpwr increase with both the number of subjects and the number of items.
-The confidence intervals reported here indicate how precisely power was estimated by our simulation.
-Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
-In practice, data simulations are often run multiple times with adjusted combinations of sample sizes.
-When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of `n_subjects` and `n_items`.
-When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
-This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.
-
-Much has been written on the optimal amount of power to target in empirical research.
-The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints  [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
-When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
-For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation.
-This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice).
-If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items.
-However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule.
-The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
-Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
-
 #### Precision results
-
-When planning for precision, one could monitor the width of all four confidence intervals at the same time.
-However, because the confidence intervals of the four contrasts strongly differ in width, it is not trivial to decide which width one should target when deciding on the appropriate sample size.
-In contrast to planning for power, there are no common standards on how to specify the targeted precision.
-For our example, we use a simple heuristic but we strongly encourage readers to think about better alternatives that are appropriate in their own applications.
-
-Our simulations show that the smallest confidence interval can be expected for the first contrast (expert with correct advice vs. expert without advice).
-The true contrast in probability for an average expert and an average item in this condition is `plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec) - plogis(b_0 + b_e) =` $`r round(plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec) - plogis(b_0 + b_e), 3)`$.
-We want the width of this confidence interval to be smaller than 0.1.
-This would mean that if the point estimate happens to be close to the true value, the plausible values inside of a 95% confidence interval would all be positive.
-
-Thus in our example, precision is defined as the (long-run) average width of a 95% confidence interval for the probability contrast between experts with correct advice and experts without advice.
-Of course, lower width implies better precision.
-Based on our simulation outcomes, we compute the precision estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.
 
@@
 ```
-
-As should be the case, precision estimates in @fig-finalprecision increase  (i.e., average width of confidence interval decreases) with both the number of subjects and the number of items.
-The confidence intervals reported here indicate how precisely the expected width of the confidence interval for our focal contrast was estimated by our simulation.
-Applying our simple heuristic of targeting an expected width smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.
-We could either choose 100 subjects and 30 items or 200 subjects and 10 items.
-
-Note that our simple heuristic for determining sample size in the planning for precision scenario was quite liberal.
-This is reflected by the result that we would need a smaller sample size than in the planning for power scenario.
-With a more conservative precision target, the result is generally the opposite: As a rule, precise parameter estimates usually require bigger samples than null hypothesis testing.
-
-<!-- ## Sensitivity analysis -->
-
-## Sensitivity analysis
-
-In our case study, we have performed simulation-based sample size planning from a single set of parameter values that reflect our assumptions of an expected effect size.
-Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
-Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
-When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present (or estimate the effect with satisfying precision).
-However, if the true effect is indeed smaller, power (or precision) will be lower, and the study might not be sufficiently informative.
-A common, more conservative strategy for sample size justification is to perform sample size planning for the smallest effect size of interest (SESOI).
-An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011].
-For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].
-When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power or precision for different effect sizes in what is called *sensitivity power analysis*.
-By simulating power (or precision) for different effect sizes (in addition to the different number of subjects and items), one can make sure that power (or precision) would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power (or precision) depends on the size of the true effect.
-In simple study designs, it is possible to perform sensitivity analysis based on a single standardized effect size (e.g., analyze power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
-However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
-In this setting, sensitivity analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
-Power (or precision) could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
-
-# Outlook
-
-<!-- Outlook -->
-Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based sample size planning in experimental research:
-
-1.
-The growing need for simulation-based sample size planning in experimental research: In order to conduct informative research using varying complex experimental designs with GLMMs, formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
-Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
-
-2.
-Managing data simulations more easily with discrete predictor variables: Simulation-based sample size planning becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
-This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable estimates for power or precision without compromising realistic assumptions about the data-generating process implied by the study design.
-
-3.
-Teaching data simulation skills: The ability to conduct simulation-based sample size planning is a valuable skill that should be taught to experimental researchers.
-By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power or precision, and improve the quality of their experimental designs.
-Equipping researchers with the knowledge and tools to perform simulation-based sample size planning enables them to make informed decisions and enhance the rigor of their studies. 
-The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
-
-4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform a priori simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
-Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-It is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.
-
-5.
-Recognizing the value of simulation-based design analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
-They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power or precision.
-Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
-
-6.
-Integration with Open Science and preregistration practices: Simulation-based sample size planning aligns well with the principles of Open Science and preregistration in experimental research.
-When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their simulation-based sample size plan, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
-
-# Conclusion
-
-In the wake of the replication crisis and myriad of underpowered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, analytic methods and software cannot be applied to conduct a priori sample size planning for GLMMs, necessitating data simulation-based approaches.
-Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based sample size planning with GLMMs themselves.
-By incorporating GLMMs and a priori sample size planning into their work, researchers can enhance the replicability and credibility of their experiments [@yarkoniGeneralizabilityCrisis2022].
-
-# References
-
-::: {#refs}
-:::

### 2024-05-21 manuscript.qmd
revise introduction

@@
 abstract: "When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
-However, the recent replication crisis underlines the need for more rigorous statistical methodology and well-powered designs.
+The recent replication crisis underlines the need for more rigorous statistical methodology and adequately powered designs.
 Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, available methods and software for a priori power analyses for GLMMs are often limited to specific designs, while data simulation approaches offer more flexibility.
-Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori power analyses to determine appropriate sample sizes with GLMMs.
+However, available methods and software for a priori sample size planning for GLMMs are often limited to specific designs.
+Data simulation approaches as an alternative offer much more flexibility.
+Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori sample size planning to determine appropriate sample sizes with GLMMs.
 Finally, we give an outlook on the increasing importance of simulation-based power analysis in experimental research."
-keywords: [power analysis, data simulation, sample size, generalized linear mixed model]
+keywords: [power analysis, precision, data simulation, sample size planning, generalized linear mixed model]
 bibliography: [Simulation Tutorial Paper.bib, r-references.bib]
@@
 
-<!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
-When planning experimental research, it is essential to determine an appropriate sample size to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data [@lakensSampleSizeJustification2022].
-However, the recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of findings [@yarkoniGeneralizabilityCrisis2022].
-As a result, there is a growing need for more rigorous statistical methodology and the adoption of well-powered experimental designs.
-While easy-to-use software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) sample size planning for more complex research designs using the flexible generalized linear mixed models (GLMM) framework in order to determine the required sample size in their experiments.
-In the present work, we provide a tutorial on performing tailored simulation based sample size planning for GLMMs.
-After introducing general concepts, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
-Then we describe general steps and decisions that are necessary whenever performing tailored data simulation.
-Because the details of these steps vary considerably for each specific study, we finish with a case study from the field of human-AI (artificial intelligence) interaction research.
-
-<!-- <!-- Old Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! --> -->
-<!-- Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs.
-Thereby, we aim to equip researchers with the tools needed to simulate data and determine appropriate sample sizes for their own research.
--->
-
-# Prior Knowledge
-
-TODO: fertig ausformulieren
-
-Simulation-based sample size planning for generalized linear models is a complex topic that touches on a variety of statistical concepts and programming skills.
-Although we give some short introduction to clarify our terminology, to profit the most from our tutorial, our readers require prior knowledge in the following topics (or refer to our collection of background literature):
-
-Readers should understand basic statistical concepts like hypothesis tests and their statistical power but also confidence intervals and their precision.
-Some background knowledge in empirical estimands, statistical estimates, and causal inference is very useful but not absolutely necessary.
-
-In addition, readers should have basic programming skills in R on how to simulate data and how to repeat these simulations in parallel to speed up computation.
-While there are a myriad of options for data simulation in R beyond the basic functionality in the base and stats packages, in our tutorial we use functions from the tidyverse and the faux package.
-
+When planning experimental research, it is essential to determine an appropriate sample size and to use appropriate statistical models to analyze the data to ensure that the results obtained are both robust and informative [@lakensSampleSizeJustification2022].
+The recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of study findings [@yarkoniGeneralizabilityCrisis2022].
+As a result, there is a growing need for more rigorous statistical methodology and the adoption of adequately powered experimental designs.
+Multiple easy-to-use software solutions exist for simple statistical models and experimental designs.
+However, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) sample size planning for more complex research designs such as flexible generalized linear mixed models (GLMM) framework.
+In the present work, we provide a tutorial on how to determine adequate sample sizes by performing tailored simulation-based sample size planning for GLMMs.
+After introducing some theoretical background on sample size planning, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
+Then we describe general steps and decisions involved in tailored data simulation.
+To illustrate the details of these steps, we finish with a hypothetical case study from the field of human-AI (artificial intelligence) interaction research.
+
+<!-- # Recommended background knowledge -->
+
+To benefit most of this tutorial paper, we recommend readers to familiarise themselves with basic statistical concepts like hypothesis tests and their statistical power as well as confidence intervals and their precision [@lakensImprovingYourStatistical2022; @riesthuisSimulationBasedPowerAnalyses2024; @kumleEstimatingPowerGeneralized2021].
+Some knowledge of causal inference is useful but not necessary [@lundbergWhatYourEstimand2021].
+In addition, readers should have a rough understanding of R [@wickhamDataScienceImport2023] and how to simulate data.
+For data simulation, we use functions from the tidyverse [@wickhamWelcomeTidyverse2019] and the faux package [@R-faux].
 Finally, readers should be familiar with regression modeling in general and GLMMs in particular.
-In our tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process.
-While it is not necessary to understand the technical details of how GLMMs are estimated, it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the values in the dependent variable are determined by the predictor variables and the random effects.
-
-Literature:
-
-- Lakens ebook [@lakensImprovingYourStatistical2022] evtl. Andy Field für die Basics?
-- R for data science [@wickhamDataScienceImport2023], Bruine article [@debruineUnderstandingMixedEffectsModels2021], faux documentation [@R-faux] für die R skills
-- kumle power tutorial [@kumleEstimatingPowerGeneralized2021] und irgendwelche paper (am besten aus AMPPS) zu GLMMs
-- what is your estimand? paper [@lundbergWhatYourEstimand2021] für das estimand topic
-
-# General concepts
+In this tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process [@debruineUnderstandingMixedEffectsModels2021].
+It is not necessary to understand the technical details of how GLMMs are estimated.
+However it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the values in the dependent variable are determined by the predictor variables and the random effects.
+
+# Theoretical background
 
@@
 
-Conducting research with too small sample sizes can have many negative consequences.
+Conducting research with insufficiently large sample sizes can have many negative consequences.
 First, experiments may yield inconclusive or misleading results, hindering the accumulation of knowledge.
-Second, studies that are doomed to never finding a postulated effect due to small sample size waste resources by consuming time, effort, and funding without delivering meaningful results.
-For such reasons, many journals and funding agencies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-Although many pragmatic justifications are possible [@lakensSampleSizeJustification2022], the ideal is often considered to determine a suitable sample size a priori (i.e., before the study is conducted) based on some meaningful computation to assure that the study will be able to fulfill its purpose.
-In this way, adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research.
-
-<!-- <!-- advantages of an a priori power analysis --> -->
-<!-- Thereby, a priori power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.
--->
-<!-- Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
--->
-<!-- Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection.
--->
-<!-- Adding a solid sample size calculation to the research process can act as a safeguard for ensuring high-quality research. -->
-<!-- Finally, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
--->
-
-The majority of empirical studies in psychology applies hypothesis testing.
-As a consequence, the dominant approach for sample size planning is based on power analysis (i.e., planing for power).
+Second, studies that are doomed to never finding a postulated effect waste resources by consuming time, effort, and funding without delivering meaningful results.
+For these reasons, many journals and funding bodies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+Although many pragmatic justifications are possible [@lakensSampleSizeJustification2022], ideally a suitable sample size should be determined a priori (i.e., before the study is conducted) based on some meaningful computation to ensure that the study will be able to fulfill its purpose.
+
+The majority of empirical studies in psychology apply hypothesis testing.
+As a consequence, the dominant approach for sample size planning is based on power analysis (i.e., planning for power).
 However, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision).
+We will now discuss both concepts.
 
@@
 
-<!-- Some basics on statistical power -->
 In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022].
@@
 Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered".
-
+If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study is considered "underpowered".
 When planning for power, a target is set for the statistical power of a hypothesis test of interest (a very common heuristic is 0.8).
-Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that (on average) would guarantee reaching this target.
+Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that,  (on average,) would guarantee reaching this target.
 
@@
 Not all research questions are best answered using hypothesis testing.
-Especially for exploratory studies when little previous research has been conducted, only estimating effects of interest is often more useful.
-When no hypothesis tests are conducted when analysing the planned study, power analysis is not not available for sample size planning.
-Nonetheless, sample size still has a crucial effect on how informative the planned study will be to learn about the effect of interest.
+Especially for exploratory studies, when little previous research has been conducted, only estimating the effects of interest is often more useful.
+When no hypothesis tests are conducted for the planned study, power analysis is not available for sample size planning.
+Nonetheless, the sample size still has a crucial effect on how informative the planned study will be to learn about the effect of interest.
 In this case, the expected width of a confidence interval is used as a target for sample size planning.
@@
 More formally, a 95%-confidence interval provides the smallest interval with the property that upon repeated sampling, 95% of individual confidence intervals would include the true quantity of interest.
-Thus, a narrow confidence interval (assuming the confidence level is kept at a high value) is more informative about the size of the true effect than a wide interval.
+Thus, a narrow confidence interval is more informative about the size of the true effect than a wide interval.
 Apart from the desired confidence level, the width of a confidence interval depends strongly on the sample size.
-Because bigger samples carry more information, they lead to smaller confidence intervals (assuming everything else stays the same).
+Because bigger samples carry more information, they lead to smaller confidence intervals.
 When planning for precision, a target is set for the expected width of a confidence interval of interest.
-Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that (on average) would guarantee reaching this target.
+Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
 
@@
 
-<!-- Advantages of GLMMs over "simpler" models -->
-As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
-
-<!-- intro into GLMMs -->
-GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data including hierarchical structures  [@fahrmeirRegressionModelsMethods2021].
-In this context, correlated data means that the value in the outcome variable for one observation may be related (i.e., more similar or less similar) to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
-This correlation can arise for various reasons: Responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), both participants come from the same neighborhood (clustering) or both participants responded to the same stimulus (stimulus effects).
-Thus, modeling such correlations is especially important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
-LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on all fixed and random effects).
-They allow for the modeling of fixed effects, which capture the relationships between our usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects exemplified above.
+As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by them [@yarkoniGeneralizabilityCrisis2022].
+GLMMs (also called multilevel models) are gaining increasing popularity because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
+GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data, including hierarchical structures [@fahrmeirRegressionModelsMethods2021].
+In this context, correlated data means that the value in the outcome variable for one observation may be related to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
+This correlation can arise for various reasons: For instance, responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), participants come from the same neighborhood (clustering), or participants responded to the same stimulus (stimulus effects).
+Thus, modeling such correlations is important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
+LMMs are used when the outcome variable is continuous and follows a normal distribution.
+They allow for the modeling of fixed effects, which capture the relationships between the usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects.
 Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
-
 As mentioned, GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
@@
 The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
-As an example, think of an experiment with different design factors (e.g., picture, headline) impacting the likelihood of users clicking on an online advertisement.
-Here, participants’ behavior is measured repeatedly (e.g., over several sessions).
-The click patterns of participants in one session are likely to be correlated with their previous sessions.
-Finally, the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
+For example, think of an experiment with different design factors (e.g., picture variations, headline texts) impacting the likelihood of users clicking on an online advertisement.
+The participants’ behavior is measured repeatedly over several sessions.
+The click patterns of participants in one session are likely to be correlated with their previous sessions and the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
 

### 2024-05-21 manuscript.qmd
revise more

@@
 
-For simpler statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori sample size planning is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018; @kelleyObtainingPowerObtaining2003; @kelleySampleSizePlanning2006].
-However, these software packages are often not flexible enough to perform sample size planning for complex designs.
-
-Existing approaches for sample size planning with GLMMs have almost exlusively focused on planning for power. 
+<!-- For less complex statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori sample size planning is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018; @kelleyObtainingPowerObtaining2003; @kelleySampleSizePlanning2006].
+-->
+<!-- However, these software packages are often not flexible enough to perform sample size planning for complex designs. -->
+
+Existing approaches for sample size planning for GLMMs have exclusively focused on planning for power.
 Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on often complicated formulas that can be used to directly calculate power while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
-Currently available formula-based software packages for power analysis with multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
+Formula-based methods rely on formulas to calculate power directly while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
+Available formula-based software packages for multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
 A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials focus on linear mixed models (LMMs) and the most common designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
+However, most of these tutorials focus on linear mixed models (LMMs) and the common study designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
 This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
-The necessary presumptions for simulation-based power analysis with GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
-The distributional assumption specifies the distributional family for the outcome variable (when conditioned on all fixed and random effects).
+The prerequisites for simulation-based power analysis with GLMMs include making assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
+The distributional assumption specifies the distributional family for the outcome variable.
 Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e., if and how they are correlated).
-Interpreting these presumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
+Making these decisions requires understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
 Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge [see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021].
+However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge (see the discussion in [@maxwellSampleSizePlanning2008] and [@kumleEstimatingPowerGeneralized2021]).
 
@@
 
-In the first six lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data. 
+In the first four lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data. 
 In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).

### 2024-05-24 manuscript.qmd
work on general background revision

@@
 
-The majority of empirical studies in psychology apply hypothesis testing.
-As a consequence, the dominant approach for sample size planning is based on power analysis (i.e., planning for power).
-However, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision).
-We will now discuss both concepts.
-
-### Planning for power
-
-In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022].
-Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size.
+The majority of empirical studies in psychology and other social sciences apply hypothesis testing.
+As a consequence, the dominant approach for determining an adequate sample siz is based on power analysis (i.e., planning for power) [@lakensSampleSizeJustification2022].
+Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size (REFERENCES).
 Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
@@
 
-### Planning for precision
-
+In contrast to power analysis, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision).
 Not all research questions are best answered using hypothesis testing.
-Especially for exploratory studies, when little previous research has been conducted, only estimating the effects of interest is often more useful.
-When no hypothesis tests are conducted for the planned study, power analysis is not available for sample size planning.
+Especially for exploratory or pilot studies, when little previous research has been conducted, only estimating the effects of interest is considered more useful (REFERENCES).
+When no hypothesis tests are conducted for the planned study, power analysis is not relevant for sample size planning.
 Nonetheless, the sample size still has a crucial effect on how informative the planned study will be to learn about the effect of interest.
-In this case, the expected width of a confidence interval is used as a target for sample size planning.
-The values inside a confidence interval are commonly interpreted as plausible values for the quantity of interest the confidence interval is supposed to estimate.
+In this case, the target quantity used for sample size planning is the expected width of a confidence interval.
+The values inside a confidence interval are commonly interpreted as plausible values for the quantity of interest is is supposed to estimate.
 More formally, a 95%-confidence interval provides the smallest interval with the property that upon repeated sampling, 95% of individual confidence intervals would include the true quantity of interest.
@@
 As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by them [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity because they offer a flexible framework for handling data with outcome variables that are not normally distributed (e.g., categorical outcomes) while accounting for both fixed and random effects [@fahrmeirRegressionModelsMethods2021].
+GLMMs (also called multilevel models) are gaining increasing popularity because they offer great flexibility [@fahrmeirRegressionModelsMethods2021].
 GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data, including hierarchical structures [@fahrmeirRegressionModelsMethods2021].
@@
 Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
-As mentioned, GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
-GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
+GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
+GLMMs involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
 The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
@@
 
-Existing approaches for sample size planning for GLMMs have exclusively focused on planning for power.
+Existing approaches for sample size planning for GLMMs have exclusively focused on planning for power (REFERENCE).
 Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
@@
 A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials focus on linear mixed models (LMMs) and the common study designs [but see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
+However, most of these tutorials focus on linear mixed models (LMMs) and the common study designs [but see @kumleEstimatingPowerGeneralized2021, for a tutorial that also covers more advanced settings].
 This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
@@
 
-# When do we profit from tailored data simulation?
-
-Performing tailored simulation based sample size planning is more complicated and time consuming than using existing software tools for sample size planning.
-Thus, it is worth discussing under which circumstances we most profit from or are forced to perform tailored simulation based sample size planning, because existing software solutions do not meet our requirements.
+# When to use tailored data simulation?
+
+Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in TABLE XXX.
+Thus, it is worth discussing under which circumstances tailored simulation-based sample size planning is needed.
 
@@
 
-TODO: fertig ausformulieren
-
-Real studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
-
-### Missing data
-
-TODO: fertig ausformulieren
-
-Missing data is a frequent issue in applied data analysis.
-There can be various reasons for why data is missing.
-For example data can be missing completely at random.
-Alternatively, subjects might drop out or produce missing data as a result of some covariate also measured in the dataset.
-A a different note, many experimental designs contain some conditions in which some values of the predictor variables are missing by design.
-Missing by design conditions can make data analysis more complicated because predictors have to be coded in highly specific ways that prevent the estimated GLMM to become unidentified.
-Whether missing data has an effect on the result of sample size planning depends on the characteristics of the missing data process.
-However, except for very simple scenarios, deciding whether missing data can be safely ignored in the sample size planning is often difficult to decide on a theoretical basis.
-A better strategy is to explicitly include the (assumed) process of missing data into the data simulation process.
-
-### Data generating processes beyond standard GLMMs
-
-TODO: fertig ausformulieren
-
-While GLMMs are a very flexible model class that can handle a large variety of outcome variables, psychological researchers are becoming increasingly aware that many psychological datasets might profit from even more sophisticated models.
-Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects.
-
-- possibility to extend beyond standard GLMMs (e.g. zero-inflated outcomes, censoring, nonlinear effects, …)
+<!-- ### Missing data -->
+
+Requirements of real-world studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
+One frequent issue in applied data analysis is missing data, and there can be various reasons for this (REFERENCES).
+For example, data can be missing completely at random (e.g., because an electronic measurement device randomly failed for some technical reasons).
+Alternatively, subjects might systematically drop out or produce missing data, but this dropout can be explained by some attributes also measured in the dataset (e.g., older subjects have a higher probability to refuse answering a question on income).
+In a more complicated scenario, missing data in some variable is caused by the measured attribute itself (e.g., wealthy people are more likely to refuse reporting their income).
+Moreover, many experimental designs contain conditions in which values of the predictor variables are missing by design.
+This can make data analysis more complicated because predictors have to be coded in specific ways that prevent the estimated GLMM from becoming unidentified.
+Whether missing data has an effect on the sample size planning depends on our theoretical assumptions on how the missingness is caused.
+However, it is difficult to make the decision whether missing data can be safely ignored in the sample size planning based on a merely theoretical approach.
+In case there is missing data, tailored simulation-based approaches offer the possibility to include the assumed process how data become missing in the data simulation and to test whether the intended data analysis provides the expected results.
+
+<!-- ### Data generating processes beyond standard GLMMs -->
+
+Although GLMMs can handle a large variety of outcome variables, researchers are becoming increasingly aware that many datasets might profit from even more sophisticated models.
+Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects that can be modeled with the R packages glmmTMB [@brooksGlmmTMBBalancesSpeed2017] or brms [@burknerAdvancedBayesianMultilevel2018].
+Tailored simulation-based approaches do not share the same limitation than the existing software solutions for power analysis that focus exclusively on GLMMs.
+As long as there is a software package available to estimate the model of interest, it is possible to perform tailored simulation-based sample size planning.
 
 ## Complex statistical hypotheses
-
-TODO: fertig ausformulieren
 
@@
 
-### Directed hypotheses and testing against values other than 0
-
-TODO: fertig ausformulieren
-
-In the new era of preregistration and registered reports, most research questions are tested with directed hypotheses, because good theories should at least postulate whether some psychological effect of interest is positive or negative.
-Even better theories should be able to specify a smallest effect sizes of interest (SESOI) that must be exceeded if the psychological effect has some practical relevance in a specific application.
+<!-- ### Directed hypotheses and testing against values other than 0 -->
+
+In the new era of preregistration and registered reports, most research questions are tested with directed hypotheses because good theories should postulate whether some psychological effect of interest is positive or negative.
+Even better theories should be able to specify the smallest effect sizes of interest (SESOI) that must be exceeded if the effect has any practical relevance.
 In combination, this might require a test like $H_0: \beta \leq 0.1$.
 
-### Combined hypotheses
-
-TODO: fertig ausformulieren
-
-Elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example $H_0: \beta_0 + \beta_1 \leq 0$.
-If the research question consists only of a single hypothesis of this sort, it is often possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
-However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypotheses.
+<!-- ### Combined hypotheses  -->
+
+Elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example, $H_0: \beta_0 + \beta_1 \leq 0$.
+If the research question consists only of a single hypothesis of this sort, it might be possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
+However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypothesis.
 For example, a combined null hypothesis $H_0$ might consist of two single null hypotheses $H_{01}: \beta_1 \leq 0$ and $H_{02}: \beta_0 + \beta_1 \leq 0$.
@@
 
-None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here.
-In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation based sample size planning.
+None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here and only some can handle directed hypotheses.
+In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation-based sample size planning.
 
@@
 
-TODO: fertig ausformulieren
-
-As discussed earlier, the two major frameworks for sample size planning are planning for power and planning for precision.
 Although planning for precision has been increasingly discussed as a more suitable strategy for exploratory research, all available software packages for sample size planning with GLMMs are based on power analysis.
-
-As long as one were interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
+Therefore, researchers that want to want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE XXX.
+<!-- As long as one is interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
 However, we demonstrate in our case study that in realistic research projects with GLMMs, the contrasts of interests and their respective confidence intervals are usually situated on the scale of the (expected) outcome in contrast to the scale of the model parameters.
-We will show that tailored simulation based sample size planning can easily handle the planning for precision scenario.
-The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset and the expected width is estimated.
+-->
+However, tailored simulation-based sample size planning can easily handle the planning for precision strategy.
+The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset, and the expected width is estimated.
 
 ## No prior studies or pilot data
-
-TODO: fertig ausformulieren
 
@@
 It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
-However, we would argue that when tailored data simulations are necessary anyways, performing the whole desing analysis in a customized way is preferred over using the existing software packages.
+However, we would argue that when tailored data simulations are necessary anyways, performing the whole design analysis in a customized way is preferred over using the existing software packages.
 
@@
 
-Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions that we will outline in the remainder of our tutorial.
-Will introduce each step in a theoretical section, followed by the practical application based on a case study.
-All code in this manuscript and simulation results are available in the project's repository on the Open Science Framework (https://osf.io/dhwf4/).
+Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions.
+We will introduce each step in a theoretical section, followed by the practical application based on a hypothetical case study.
+All code in this manuscript and simulation results are available in the project’s repository on the Open Science Framework (https://osf.io/dhwf4/).
 

### 2024-05-24 manuscript.qmd
work on case study part

@@
 
-## Define the estimand
+## Step 1: Define the estimand
 
@@
 
-The first step in every research process should always be a clear definition of which theoretical quantity is necessary to answer the specific research question of the study under planning.
-Unfortunately, many psychological researchers are still not well trained in this task.
+The first step in every research process should be a clear definition of which theoretical quantity is necessary to answer the specific research question.
 The theoretical estimand is a unit specific quantity defined outside of any statistical model [@lundbergWhatYourEstimand2021].
@@
 
-When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reformulate the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
-For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantities within the scope that can be estimated with a regression model.
-Note however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression (CITATION).
-
-When the empirical estimand has been specified, the last step is to clearly define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
-When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique, and software package is used to fit the model, and how to compute hypothesis tests or confidence intervals.
+When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reformulates the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
+For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantity within the scope that can be estimated with a regression model.
+Note, however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression (CITATION).
+
+When the empirical estimand has been specified, the last step is to define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
+When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique and software package are used to fit the model and how to compute hypothesis tests or confidence intervals.
 
@@
 
-In the present case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) in an AI-enabled diagnostic decision support system.
-The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography.
-In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans that are required to make the final diagnostic decision.
+In the present hypothetical case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) embedded in a diagnostic decision support system.
+The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography (CT).
+In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans who are required to make the final diagnostic decision.
 The research goal is to validate the effectiveness of the AI-enabled advice.
@@
 
-*We expect that for BOTH expert radiologists and students under training, correct AI-advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI-advice presented, AND, we expect that for BOTH experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
-
-I becomes clear that our theoretical estimand is be a causal quantity, based on several individual quantities that are defined outside of any statistical model [@lundbergWhatYourEstimand2021].
+*We expect that for BOTH expert radiologists and medical students, correct AI-advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI-advice presented, AND, we expect that for BOTH experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
+
+It becomes clear that our theoretical estimand is a causal quantity, based on several individual quantities that are defined outside of any statistical model [@lundbergWhatYourEstimand2021].
 To elaborate further, we focus on only one part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists.
-For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice were presented and the probability that a correct diagnosis is made if no AI-advice were presented.
+For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice is presented and the probability that a correct diagnosis is made if no AI-advice is presented.
 Mathematically, we could describe this unit-specific quantity with some pseudo-equation like this (where $s$ stands for a specific expert radiologist and $i$ stands for a specific CT scan):
@@
 We must specify how we want to aggregate the quantity across different experts and scans.
-If we would compute the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts (CITATION).
-Although often debated whether this is a good idea, psychological research often focuses on a different quantity:
-Instead of averaging over experts and scans, we focus in a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
+If we computed the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts (CITATION).
+Although often debated whether this is a good idea, psychological research often focuses on a different quantity: Instead of averaging over experts and scans, we focus on a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
 With this definition, the theoretical estimand in our above example can be expressed as:
@@
 $$
-To complete our definition of the theoretical estimand, we have to clearly define out target population that consists of both persons and scans:
-With respect to persons, we are only interested in expert radiologists and medical students at German universities.
+To complete our definition of the theoretical estimand, we have to define ourt target population that consists of both persons and scans: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
 With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
-Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system under study.
-
-While a theoretical estimand is defined outside of any statistical model and can be impossible to be observe directly, an empirical estimand is a quantity that can be estimated based on observed data.
-For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of brain scans, but the kind of AI-advice given for each brain scan is randomly assigned within participants.
-This random intervention allows to estimate our theoretical estimand, although in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each brain scan.
+Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system.
+
+While a theoretical estimand is defined outside of any statistical model, an empirical estimand is a quantity that can be estimated based on observed data.
+For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of head CT scans, but the kind of AI-advice given for each scan is randomly assigned within participants.
+This random intervention allows us to estimate our theoretical estimand, although, in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each scan.
 For our example, each probability expression in our theoretical estimand can be modeled with the same GLMM.
-Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability and these estimates can be used to compute an estimate for each of the four probability contrasts.
+Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability, and these estimates can be used to compute an estimate for each of the four probability contrasts.
 For pedagogical reasons, we will skip the concrete definition of our empirical estimand until we have discussed how to simulate data based on a concrete GLMM in the next section.
 
-## Simulate the data generating process
+## Step 2: Simulate the data generating process
 
@@
 This requires specifying a generative process for all predictor variables used in the final data analysis.
-While such assumptions can be quite challenging for observational studies and/or continuous predictor variables, this is less of a problem for experimental studies with only categorical predictor variables.
+While such assumptions can be quite challenging for observational studies or continuous predictor variables, this is less of a problem for experimental studies with only categorical predictor variables.
 When all predictor variables have been simulated, one can use the structure of a suitable GLMM to simulate the dependent variable.
@@
 We will discuss strategies on how these values can be obtained later.
-Because one has full control over the data generating process in a tailored simulation, it is simple to also model highly specific aspects of the planned study, like data missing by design or assuming that subjects drop out (i.e.
+Because we have full control over the data generating process in a tailored simulation, it is possible to model specific aspects of the planned study, like data missing by design or assuming that subjects drop out (i.e.
 deleting their data from the simulated dataset) based on other variables in the simulated dataset.
 The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
-For this reason, deciding on a specific set of assumptions can easily feel overwelming.
+For this reason, deciding on a specific set of assumptions can easily feel overwhelming.
 However, we would argue that even a strongly simplified data generating process (e.g.
-including only a small number of interaction effects; including only random intercepts but no random slopes; making rather implausible assumptions for missing data) can yield informative result and should be strongly preferred to the current status quo in psychological research, where most studies do not justify their sample size or use highly questionable heuristics that are independent of their specific research question.
+only a small number of interaction effects; only random intercepts and no random slopes; making rather implausible assumptions for missing data) can yield informative results.
 
@@
 In our case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated.
-<!-- The goal is to understand how AI advice influences medical decision-making. -->
-Participants, radiologists (task experts) and students/interns (non experts), review a series of head computer tomography (CT) scans to assess the presence of a bleeding. 
-To support their decision-making, an AI model provides initial diagnostic advice, which can be used as guidance by the participants.
-In the control condition, no AI advice is presented, meaning that the participants have to read the CT scan without any support.
+Radiologists (task experts) and students/interns (non experts) review a series of head computer tomography (CT) scans to assess the presence of a bleeding.
+An AI model provides initial diagnostic advice to assist their decision-making.
+In the control condition, no AI advice is presented.
 When AI advice is given, this advice can be either correct or incorrect.
@@
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-<!-- With this experiment, we want to determine if (a) experts are better than non-experts in reading head CT scans and if (b) correct AI advice leads to better diagnostic accuracy than incorrect AI advice.
--->
 In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible.
@@
 
-## Strategies on specifying the population parameters
+## Step 3: Specify the population parameters
 
@@
 
-Although we focus on frequentist statistics in our tutorial, selecting plausible values for data simulation shares many similarities with the specification of (weakly) informative prior distribution in applied Bayesian statistics.
-In this literature, strategies have been developed to monitor the plausibility of (prior) model assumptions [@gelmanBayesianWorkflow2020], that can be adapted for our purposes.
+Although we use frequentist model estimation in our tutorial, many strategies described in this chapter are inspired by research on how to monitor the plausibility of model assumptions in applied Bayesian statistics [@gelmanBayesianWorkflow2020].
 
@@
 
-Ideally, one would rely on meta-analytic results or conclusive data from pilot studies.
-However, these are sometimes not readily available.
 All parameter values in our present case study have been determined based on results from distantly related study designs in the literature.
@@
 
-<!-- For most GLMMs, the interpretation of model parameters are quite complicated which makes it difficult to specify plausible values for a specific application. -->
 The mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
 The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
-For example, domain experts might not be able to directly choose plausible values for the $\beta$ coefficients in a logistic regression model (which are measured on the log-odds scale).
-However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental condition, i.e.
-which descriptive statistics they expect to observe (on average) when they will analyse the actual dataset produced by the study under planning (if everything turns out as postulated by the investigated theory).
+For example, domain experts might not be able to directly choose plausible values for the  coefficients in a logistic regression model (which are measured on the log-odds scale).
+However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental conditions, i.e.
+which descriptive statistics they expect to observe (on average).
 The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
-Although insightful descriptive statistics often depend on the model parameters in complicated ways, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
+Although insightful descriptive statistics often depend on the model parameters in a non-linear way, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
 
@@
 
-In the last section, we showed how we can derive the model implied probability for each experimental condition, that a subject with average ability solves an item with average difficulty.
+In the last section, we showed how we can derive the model implied probability that a subject with average ability solves an item with average difficulty  for each experimental condition.
 Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty [@fahrmeirRegressionModelsMethods2021].
@@
 However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty (the marginal perspective).
-How much the two probabilities differ depend on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations would be zero).
+How much the two probabilities differ depends on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations would be zero).
 We want to use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
@@
 
-We tried using these descriptive statistics to judge together with our domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
-However although they result was deemed plausible, in our example these statistics are not informative enough to determine a final set of plausible parameter values.
+We tried using these descriptive statistics to judge together with domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
+However, although the result was deemed plausible, in our example these statistics are not informative enough to determine a final set of plausible parameter values.
 For this reason, we will additionally look at insightful model based quantities.
@@
 Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
-This is especially true for the standard deviation of random effects, that have complicated (and often unexpected) effects on the model-implied results.
+This is especially true for the standard deviation of random effects that have complicated (and often unexpected) effects on the model-implied results.
 An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
 For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
-Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out unplausible boundary conditions.
-For example, the domain expert might deem it unplausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
+Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out implausible boundary conditions.
+For example, the domain expert might deem it implausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
 
@@
 
-In the last section, we simulated a large simulated dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
+In the last section, we simulated a large dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
 We will now use the model implied probability of each simulated data point (stored in the variable `y_prob`) to visualize the whole model implied marginal distribution of correct diagnoses for each experimental condition.
@@
 As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
-For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst which is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
+For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst who is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
 Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
@@
 
-## Estimate the statistical model
+## Step 4: Estimate the statistical model
 
@@
 
-## Compute the estimate
+## Step 5: Compute the estimate
 
@@
 In the earlier section on our theoretical estimand we translated this verbal description into four probability statements that are specified outside of any specific statistical model.
-The next translation step is to specify the empirical estimand, that is how these probabilites can be estimated within the scope of our concrete GLMM.
+The next translation step is to specify the empirical estimand, that is how these probabilities can be estimated within the scope of our concrete GLMM.
 
@@
 With a standard confidence level of $1 - \alpha = 0.95$, plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
-Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts (which the expected behavior for binomial GLMMs).
+Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts (which is the expected behavior for binomial GLMMs).
 The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice) where both underlying probabilities are close to 1.
@@
 
-## Perform repeated simulations
+## Step 6: Perform repeated simulations
 
@@
 The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
-On a conceptual level, we first require a function, that takes as input the sample size and the full set of population parameter values.
+On a conceptual level, we first require a function that takes as input the sample size and the full set of population parameter values.
 When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.

### 2024-06-04 manuscript.qmd
extend the planning for power or precision section

@@
 
-To benefit most of this tutorial paper, we recommend readers to familiarise themselves with basic statistical concepts like hypothesis tests and their statistical power as well as confidence intervals and their precision [@lakensImprovingYourStatistical2022; @riesthuisSimulationBasedPowerAnalyses2024; @kumleEstimatingPowerGeneralized2021].
+To benefit most of this tutorial paper, we recommend readers to familiarise themselves with basic statistical concepts like hypothesis tests (HTs) and their statistical power as well as confidence intervals (CIs) and their precision [@lakensImprovingYourStatistical2022; @riesthuisSimulationBasedPowerAnalyses2024; @kumleEstimatingPowerGeneralized2021].
 Some knowledge of causal inference is useful but not necessary [@lundbergWhatYourEstimand2021].
@@
 For these reasons, many journals and funding bodies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-Although many pragmatic justifications are possible [@lakensSampleSizeJustification2022], ideally a suitable sample size should be determined a priori (i.e., before the study is conducted) based on some meaningful computation to ensure that the study will be able to fulfill its purpose.
+Although scientists often do not justify sample size or use general heuristics from the literature, resource constraints often play an important role [@lakensSampleSizeJustification2022].
+But ideally, a suitable sample size should be determined a priori (i.e., before the study is conducted) based on some meaningful computation to ensure that the study will be able to fulfill its purpose.
 
 The majority of empirical studies in psychology and other social sciences apply hypothesis testing.
-As a consequence, the dominant approach for determining an adequate sample siz is based on power analysis (i.e., planning for power) [@lakensSampleSizeJustification2022].
-Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size (REFERENCES).
-Less formally, power is described as the probability that a hypothesis test correctly rejects the null hypothesis when the alternative hypothesis is true.
+As a consequence, the dominant approach for determining an adequate sample size is based on power analysis (i.e., planning for power) [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
+Statistical power is defined as the probability that a HT has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size [@cohenPowerPrimer1992].
+Less formally, power is described as the probability that a HT correctly rejects the null hypothesis when the alternative hypothesis is true.
 If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study is considered "underpowered".
-When planning for power, a target is set for the statistical power of a hypothesis test of interest (a very common heuristic is 0.8).
-Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that,  (on average,) would guarantee reaching this target.
-
-In contrast to power analysis, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision).
+When planning for power, a target is set for the statistical power of a HT of interest.
+Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
+Although it is recommended to justify the desired significance level and power [@lakensJustifyYourAlpha2018], most empirical studies adopt the heuristic of $\alpha = 0.05$ and $1 - \beta = 0.80$.
+
+In contrast to power analysis, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision or planning for accuracy) [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
 Not all research questions are best answered using hypothesis testing.
-Especially for exploratory or pilot studies, when little previous research has been conducted, only estimating the effects of interest is considered more useful (REFERENCES).
-When no hypothesis tests are conducted for the planned study, power analysis is not relevant for sample size planning.
-Nonetheless, the sample size still has a crucial effect on how informative the planned study will be to learn about the effect of interest.
-In this case, the target quantity used for sample size planning is the expected width of a confidence interval.
-The values inside a confidence interval are commonly interpreted as plausible values for the quantity of interest is is supposed to estimate.
-More formally, a 95%-confidence interval provides the smallest interval with the property that upon repeated sampling, 95% of individual confidence intervals would include the true quantity of interest.
-Thus, a narrow confidence interval is more informative about the size of the true effect than a wide interval.
-Apart from the desired confidence level, the width of a confidence interval depends strongly on the sample size.
-Because bigger samples carry more information, they lead to smaller confidence intervals.
-When planning for precision, a target is set for the expected width of a confidence interval of interest.
+It has been argued that basic research rarely requires making discrete decisions on whether some effect has been "discovered" and should thus shift from a hypothesis testing towards an estimation framework [@mcelreathStatisticalRethinkingBayesian2020; @cummingNewStatisticsWhy2014; @kruschkeBayesianNewStatistics2018].
+Although this view is not without critique [@uyguntuncEpistemicPragmaticFunction2023], at least for exploratory or pilot studies where little previous research has been conducted, more scientists seem to agree that simply estimating the effects of interest and making the estimation uncertainty transparent by reporting CIs is more useful.
+Assuming that no HTs are conducted for a planned study, power analysis is not relevant for sample size planning.
+Nonetheless, the sample size still has a crucial effect on how informative the planned study will be, because an effect of interest is estimated more precisely with bigger samples.
+In the precision framework, the target quantity commonly used for sample size planning is the expected width of a CI [@kelleySampleSizePlanning2006; @lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
+The values inside a CI are often interpreted as plausible values for the quantity of interest it is supposed to estimate.
+More formally, a CI with confidence level 0.95 provides the smallest interval with the property that upon repeated sampling, 95% of individual CIs would include the true quantity of interest.
+Thus, a narrow CI is more informative about the size of the true effect than a wide interval.
+Apart from the desired confidence level, the width of a CI depends strongly on the sample size.
+Because bigger samples carry more information, they lead to smaller CIs.
+When planning for precision, a target can be set for the expected width of a CI of interest.
 Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
+Because planning for precision is still rare, there are no common heuristics on how to choose the desired width of the CI [@lakensSampleSizeJustification2022].
 
@@
 
-# Simulation based sample size planning with GLMMs
+# Simulation-based sample size planning with GLMMs
 
@@
 Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on formulas to calculate power directly while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the hypothesis test is significant for the simulated data).
+Formula-based methods rely on formulas to calculate power directly while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the HT is significant for the simulated data).
 Available formula-based software packages for multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
@@
 If the global hypothesis $H_0$ is combined with *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis.
-However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global hypothesis test.
+However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
 
@@
 However, tailored simulation-based sample size planning can easily handle the planning for precision strategy.
-The only change in procedure is that instead of computing hypothesis tests for each simulated dataset and estimating statistical power across repetitions, confidence intervals are computed for each simulated dataset, and the expected width is estimated.
+The only change in procedure is that instead of computing HTs for each simulated dataset and estimating statistical power across repetitions, CIs are computed for each simulated dataset, and the expected width is estimated.
 
@@
 All frameworks for sample size planning require the user to make assumptions about the expected effect size.
-Assuming the true effect is of this size (or greater), one can compute the (minimum) power of a hypothesis test or the (maximum) expected width of a confidence interval.
+Assuming the true effect is of this size (or greater), one can compute the (minimum) power of a HT or the (maximum) expected width of a CI.
 Existing software packages for sample size planning for GLMMs usually require to provide the assumed effect in the unit of some standardized measure of effect size.
@@
 
-Although the details differ depending on the specific study characteristics, each tailored simulation based sample size planning requires a series of steps and decisions.
+Although the details differ depending on the specific study characteristics, each tailored simulation-based sample size planning requires a series of steps and decisions.
 We will introduce each step in a theoretical section, followed by the practical application based on a hypothetical case study.
@@
 When the empirical estimand has been specified, the last step is to define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
-When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique and software package are used to fit the model and how to compute hypothesis tests or confidence intervals.
+When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique and software package are used to fit the model and how to compute HTs or CIs.
 
@@
 
-Tailored simulation based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with the same design or pilot data.
+Tailored simulation-based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with the same design or pilot data.
 In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
@@
 Our strategies to specify population parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation based sample size planning.
+Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
 The basic idea of all strategies is how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
@@
 While the majority of psychological research uses hypothesis testing, there are repeated claims that a focus on estimation can also be a good choice in many circumstances (CITATION).
-This decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the hypothesis test(s) or confidence interval(s) (e.g., compute hypothesis tests and confidence intervals with the marginaleffects R package using the delta method).
+This decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with the marginaleffects R package using the delta method).
 
@@
 We could use this method to compute point estimates for all four contrast that are part of our empirical estimand.
-However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute hypothesis tests or confidence interval for our point estimates.
-
-The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute hypothesis tests and confidence intervals for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
+However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute HTs or CI for our point estimates.
+
+The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute HTs and CIs for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
 
@@
 For this we must specify which rows in `probs` have to be subtracted from each other.
-We will use the `hypotheses` function to compute our four contrasts of interest together with hypothesis tests and confidence intervals.
-We use the default inference options of the marginaleffects package that compute hypothesis tests and confidence intervals based on the approximate delta method.
+We will use the `hypotheses` function to compute our four contrasts of interest together with HTs and CIs.
+We use the default inference options of the marginaleffects package that compute HTs and CIs based on the approximate delta method.
 
@@
 
-If we chose parameter estimation for our case study, we would focus on the two-sided confidence intervals of the four contrasts of interest.
+If we chose parameter estimation for our case study, we would focus on the two-sided CIs of the four contrasts of interest.
 With a standard confidence level of $1 - \alpha = 0.95$, plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
-Note that due to the constrained range of the probability scale, the width of the confidence interval differs between the four contrasts (which is the expected behavior for binomial GLMMs).
+Note that due to the constrained range of the probability scale, the width of the CI differs between the four contrasts (which is the expected behavior for binomial GLMMs).
 The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice) where both underlying probabilities are close to 1.
@@
 
-Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute hypothesis tests or confidence intervals for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
+Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute HTs or CIs for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
 The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
 On a conceptual level, we first require a function that takes as input the sample size and the full set of population parameter values.
-When planning for power, the function should return the p-value(s) of the hypothesis test(s) of interest when conducted on the simulated dataset.
-When planning for precision, the function should return the width of the confidence interval(s) of interest.
+When planning for power, the function should return the p-value(s) of the HT(s) of interest when conducted on the simulated dataset.
+When planning for precision, the function should return the width of the CI(s) of interest.
 Secondly, we must run this function repeatedly with the same sample size and population parameters.
@@
 When planning for power, we compute the relative frequency of a significant p-value(s) across repeated simulations.
-When planning for precision, we compute the average width of the confidence interval(s).
+When planning for precision, we compute the average width of the CI(s).
 Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted power or precision.
@@
 We are finally ready to run our simulation-based sample size planning analyses to plan for power and for precision.
-Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and confidence intervals) in a single command.
+Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and CIs) in a single command.
 
@@
 
-The result of this computation is a data frame that contains the p-values and confidence intervals of all specified contrasts for each simulated dataset.
+The result of this computation is a data frame that contains the p-values and CIs of all specified contrasts for each simulated dataset.
 In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
@@
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.[^3]
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.[^3]
 
@@
 As should be the case, power estimates in @fig-finalpwr increase with both the number of subjects and the number of items.
-The confidence intervals reported here indicate how precisely power was estimated by our simulation.
-Higher precision (which would be reflected in narrower confidence intervals) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
+The CIs reported here indicate how precisely power was estimated by our simulation.
+Higher precision (which would be reflected in narrower CIs) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
 In practice, data simulations are often run multiple times with adjusted combinations of sample sizes.
@@
 
-When planning for precision, one could monitor the width of all four confidence intervals at the same time.
-However, because the confidence intervals of the four contrasts strongly differ in width, it is not trivial to decide which width one should target when deciding on the appropriate sample size.
+When planning for precision, one could monitor the width of all four CIs at the same time.
+However, because the CIs of the four contrasts strongly differ in width, it is not trivial to decide which width one should target when deciding on the appropriate sample size.
 In contrast to planning for power, there are no common standards on how to specify the targeted precision.
@@
 
-Our simulations show that the smallest confidence interval can be expected for the first contrast (expert with correct advice vs. expert without advice).
+Our simulations show that the smallest CI can be expected for the first contrast (expert with correct advice vs. expert without advice).
 The true contrast in probability for an average expert and an average item in this condition is `plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec) - plogis(b_0 + b_e) =` $`r round(plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec) - plogis(b_0 + b_e), 3)`$.
-We want the width of this confidence interval to be smaller than 0.1.
-This would mean that if the point estimate happens to be close to the true value, the plausible values inside of a 95% confidence interval would all be positive.
-
-Thus in our example, precision is defined as the (long-run) average width of a 95% confidence interval for the probability contrast between experts with correct advice and experts without advice.
+We want the width of this CI to be smaller than 0.1.
+This would mean that if the point estimate happens to be close to the true value, the plausible values inside of a 95% CI would all be positive.
+
+Thus in our example, precision is defined as the (long-run) average width of a 95% CI for the probability contrast between experts with correct advice and experts without advice.
 Of course, lower width implies better precision.
-Based on our simulation outcomes, we compute the precision estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% confidence intervals) and visualize the results with the following code.
+Based on our simulation outcomes, we compute the precision estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.
 
@@
 
-As should be the case, precision estimates in @fig-finalprecision increase  (i.e., average width of confidence interval decreases) with both the number of subjects and the number of items.
-The confidence intervals reported here indicate how precisely the expected width of the confidence interval for our focal contrast was estimated by our simulation.
+As should be the case, precision estimates in @fig-finalprecision increase  (i.e., average width of CI decreases) with both the number of subjects and the number of items.
+The CIs reported here indicate how precisely the expected width of the CI for our focal contrast was estimated by our simulation.
 Applying our simple heuristic of targeting an expected width smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.

### 2024-06-04 manuscript.qmd
update glmm intro sections

@@
 This correlation can arise for various reasons: For instance, responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), participants come from the same neighborhood (clustering), or participants responded to the same stimulus (stimulus effects).
-Thus, modeling such correlations is important whenever the data has a clear structure, while the grouping variables can be hierarchically organized (e.g., students nested in schools, schools nested in districts) or not (e.g., students solve math exercises, but neither student sees all exercises).
+Thus, modeling such correlations is important whenever the data has a clear structure, while the grouping variables can be hierarchically nested (e.g., grouping variables students and schools: each student belongs to exactly one school) or cross-classified (e.g., grouping variables students and math exercises: each student is presented with several math exercises).
 LMMs are used when the outcome variable is continuous and follows a normal distribution.
@@
 The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
-For example, think of an experiment with different design factors (e.g., picture variations, headline texts) impacting the likelihood of users clicking on an online advertisement.
+For example, think of an experiment with different design factors (e.g., picture positions, headline aesthetics) impacting the likelihood of users clicking on an online advertisement.
 The participants’ behavior is measured repeatedly over several sessions.
@@
 
-Existing approaches for sample size planning for GLMMs have exclusively focused on planning for power (REFERENCE).
+To our knowledge, existing approaches for sample size planning for GLMMs have exclusively focused on planning for power.
 Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on formulas to calculate power directly while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., how often the HT is significant for the simulated data).
+Formula-based methods rely on formulas to calculate power directly while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., what percentage of simulated datasets produces a significant p-value).
 Available formula-based software packages for multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
@@
 This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
-The prerequisites for simulation-based power analysis with GLMMs include making assumptions about the distributional form of the outcome variable, the random effects, and the correlation structure.
-The distributional assumption specifies the distributional family for the outcome variable.
+Simulation-based power analysis with GLMMs requires making a range of assumptions: The (conditional) distribution assumption specifies the distributional family for the outcome variable.
 Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e., if and how they are correlated).

### 2024-06-05 manuscript.qmd
update when to use tailored data simulation section

@@
 Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in TABLE XXX.
-Thus, it is worth discussing under which circumstances tailored simulation-based sample size planning is needed.
-
-## Complex study designs
-
-<!-- ### Missing data -->
+The most important circumstances under which tailored simulation-based sample size planning is necessary are 1) complex study designs, 2) complex statistical hypotheses, 3) planning for precision, 4) no available prior studies or pilot data.
+
+<!-- ## Complex study designs -->
 
 Requirements of real-world studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
-One frequent issue in applied data analysis is missing data, and there can be various reasons for this (REFERENCES).
+One frequent issue in applied data analysis is missing data, and there can be various reasons for this [@littleStatisticalAnalysisMissing2014].
 For example, data can be missing completely at random (e.g., because an electronic measurement device randomly failed for some technical reasons).
@@
 Whether missing data has an effect on the sample size planning depends on our theoretical assumptions on how the missingness is caused.
-However, it is difficult to make the decision whether missing data can be safely ignored in the sample size planning based on a merely theoretical approach.
-In case there is missing data, tailored simulation-based approaches offer the possibility to include the assumed process how data become missing in the data simulation and to test whether the intended data analysis provides the expected results.
-
-<!-- ### Data generating processes beyond standard GLMMs -->
-
+However, it is often challenging to decide whether missing data can be safely ignored in the data analysis and sample size planning process based on a merely theoretical approach [@gomilaMissingDataExperiments2022].
+Tailored simulation-based approaches offer the possibility to include the assumed process of how data become missing in the data simulation, thereby determining the required sample based on simulated datasets that contain missing values [for an example, see @lanePowerStrugglesEstimating2018].
+As a byproduct, the simulated datasets can also be used to test whether the intended data analysis provides the expected (unbiased) results, despite the missing data.
 Although GLMMs can handle a large variety of outcome variables, researchers are becoming increasingly aware that many datasets might profit from even more sophisticated models.
@@
 Tailored simulation-based approaches do not share the same limitation than the existing software solutions for power analysis that focus exclusively on GLMMs.
-As long as there is a software package available to estimate the model of interest, it is possible to perform tailored simulation-based sample size planning.
-
-## Complex statistical hypotheses
-
-The most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is some coefficient in a regression model.
+As long as there is a software package available to estimate the model of interest, it is always possible to perform tailored simulation-based sample size planning.
+
+<!-- ## Complex statistical hypotheses -->
+
+The most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is a slope or intercept of a regression model.
 However, many research questions in psychology actually require testing more complex statistical hypotheses.
-
-<!-- ### Directed hypotheses and testing against values other than 0 -->
-
-In the new era of preregistration and registered reports, most research questions are tested with directed hypotheses because good theories should postulate whether some psychological effect of interest is positive or negative.
-Even better theories should be able to specify the smallest effect sizes of interest (SESOI) that must be exceeded if the effect has any practical relevance.
-In combination, this might require a test like $H_0: \beta \leq 0.1$.
-
-<!-- ### Combined hypotheses  -->
-
-Elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example, $H_0: \beta_0 + \beta_1 \leq 0$.
+In the new era of preregistration and registered reports [@chambersPresentFutureRegistered2022], most research questions should be tested with directed hypotheses because good theories at least postulate whether some psychological effect of interest is positive or negative.
+Even better theories should be able to specify the smallest effect sizes of interest (SESOI) that must be exceeded if the effect has any practical relevance [@lakensEquivalenceTestingPsychological2018].
+In combination, this might require a test such as $H_0: \beta \leq 0.1$.
+More elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example testing simple slopes [@preacherComputationalToolsProbing2006] with a hypothesis such as $H_0: \beta_0 + \beta_1 \leq 0$.
 If the research question consists only of a single hypothesis of this sort, it might be possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
-However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypothesis.
+However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypothesis [for a tutorial on contrast analysis in GLMMs, see @schadHowCapitalizePriori2020].
 For example, a combined null hypothesis $H_0$ might consist of two single null hypotheses $H_{01}: \beta_1 \leq 0$ and $H_{02}: \beta_0 + \beta_1 \leq 0$.
@@
 For other research questions, the combined null hypothesis $H_0$ would be rejected if $H_{01}$ *OR* $H_{02}$ *OR* both are rejected.
-If the global hypothesis $H_0$ is combined with *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis.
+If the global hypothesis $H_0$ is combined with *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis [@dmitrienkoTraditionalMultiplicityAdjustment2013].
 However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
-
 None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here and only some can handle directed hypotheses.
@@
 
-## Planning for precision
-
-Although planning for precision has been increasingly discussed as a more suitable strategy for exploratory research, all available software packages for sample size planning with GLMMs are based on power analysis.
+<!-- ## Planning for precision -->
+
+Although planning for precision [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008; @cummingNewStatisticsWhy2014] has been increasingly discussed as a useful strategy for empirical research, all available software packages for sample size planning with GLMMs are based on power analysis.
 Therefore, researchers that want to want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE XXX.
@@
 -->
-However, tailored simulation-based sample size planning can easily handle the planning for precision strategy.
+However, tailored simulation-based sample size planning can easily handle the planning for precision approach [@maxwellSampleSizePlanning2008].
 The only change in procedure is that instead of computing HTs for each simulated dataset and estimating statistical power across repetitions, CIs are computed for each simulated dataset, and the expected width is estimated.
 
-## No prior studies or pilot data
+<!-- ## No prior studies or pilot data -->
 
@@
 When the researcher has access to similar studies or pilot data, providing such standardized effect sizes is feasible.
-However, this requirement can be an almost impossible challenge when no prior studies of pilot data are available.
+However, note that choosing effect sizes based on small pilot studies is generally not recommended, as those estimates can be heavily biased [@albersWhenPowerAnalyses2018a; @lakensSampleSizeJustification2022].
+Providing an informed standardized effect size can be an almost impossible challenge when no prior studies of pilot data are available.
 This problem is further exacerbated by the fact that GLMMs are so flexible, that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
@@
 It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
-However, we would argue that when tailored data simulations are necessary anyways, performing the whole design analysis in a customized way is preferred over using the existing software packages.
+However, we would argue that when tailored data simulations are necessary anyways, performing the whole sample size planning in a customized way is preferred over using the existing software packages.
 

### 2024-06-05 manuscript.qmd
update estimand, data simulation and outlook sections

@@
 
-The first step in every research process should be a clear definition of which theoretical quantity is necessary to answer the specific research question.
-The theoretical estimand is a unit specific quantity defined outside of any statistical model [@lundbergWhatYourEstimand2021].
-Depending on the research and whether the research question is causal, this quantity might be counterfactual and not directly observable.
-Part of the theoretical estimand is also a clear definition of the target population, for which the quantity is of interest.
-
-When the theoretical estimand has been specified, the next step is to define the empirical estimand, which reformulates the theoretical estimand in a way that can be estimated based on the observed data that will be collected in the planned study.
-For many common research questions in psychology, the empirical estimand can be expressed as a statistical quantity within the scope that can be estimated with a regression model.
-Note, however, that this is not always possible and that the literature on causal inference includes many estimation strategies beyond regression (CITATION).
-
-When the empirical estimand has been specified, the last step is to define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
+The first step in every research process should be a clear definition of the *estimand* [@lundbergWhatYourEstimand2021], i.e.
+the theoretical quantity which is necessary to answer the specific research question.
+The estimand consists of a quantity, that can be described for each unit under investigation and a clear definition of the target population, for which the quantity is of interest.
+
+The estimand should always be defined outside of any statistical model, because there are usually a range of statistical methods that can be used to estimate the same estimand based on the observed data that will be collected in the planned study.
+For many common research questions in psychology, the estimand can be expressed as a statistical quantity that can be estimated with a regression model.
+However, this is not always possible all causal estimands, which is why the literature discusses many estimation strategies beyond regression [@lundbergWhatYourEstimand2021].
+
+<!-- When the empirical estimand has been specified, the last step is to define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
 When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique and software package are used to fit the model and how to compute HTs or CIs.
+-->
 
@@
 
-It becomes clear that our theoretical estimand is a causal quantity, based on several individual quantities that are defined outside of any statistical model [@lundbergWhatYourEstimand2021].
-To elaborate further, we focus on only one part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists.
-For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice is presented and the probability that a correct diagnosis is made if no AI-advice is presented.
-Mathematically, we could describe this unit-specific quantity with some pseudo-equation like this (where $s$ stands for a specific expert radiologist and $i$ stands for a specific CT scan):
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, specific expert, specific scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, specific expert, specific scan})
-\end{aligned}
-$$
-However, our verbal description is not precise enough yet, because the above quantity may differ for each combination of expert and scan.
-We must specify how we want to aggregate the quantity across different experts and scans.
-If we computed the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts (CITATION).
-Although often debated whether this is a good idea, psychological research often focuses on a different quantity: Instead of averaging over experts and scans, we focus on a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
-With this definition, the theoretical estimand in our above example can be expressed as:
+It becomes clear that our estimand consists of four comparisons between experimental conditions [@lundbergWhatYourEstimand2021].
+However, the verbal descriptions is not precise enough yet, which is why we try to give a unique mathematical expression:
 
@@
 $$
-In total, our theoretical estimand consists of four such expressions, the remaining three are:
 
@@
 $$
-To complete our definition of the theoretical estimand, we have to define ourt target population that consists of both persons and scans: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
+For example, the first expression is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice is presented and the probability that a correct diagnosis is made if no AI-advice is presented.
+This contrast is quantified for a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.^[Note that a different estimand would be the so-called average treatment effect (ATE).
+For the ATE, the probability contrast would be computed for each expert and scan, and then these contrasts are averaged across all experts and scan in the target population [@lundbergWhatYourEstimand2021]].
+To complete our definition of the estimand, we have to define our target population that consists of both persons and scans: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
 With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
@@
 
-While a theoretical estimand is defined outside of any statistical model, an empirical estimand is a quantity that can be estimated based on observed data.
+Although the estimand is initially defined outside of any statistical model, it is only useful if we find a way to estimate it based on observed data.
+For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of head CT scans, but the kind of AI-advice given for each scan is randomly assigned within participants.
+This random intervention allows us to produce an empirical estimate of our estimand, although, in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each scan.
+We will see soon that each of the probability expressions in our estimand can be modeled with the same GLMM.
+Estimating this GLMM based on the data observed in our planned study will produce an estimate for each probability, and these estimates can then be combined to compute an estimate for each of the four probability contrasts.
+For pedagogical reasons, we will skip the concrete definition of our estimate until we have discussed how to simulate data based on a concrete GLMM in the next section.
+
+<!-- It becomes clear that our estimand consists of four comparisons between experimental conditions [@lundbergWhatYourEstimand2021].
+To elaborate further, we focus on one part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists first.
+For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice is presented and the probability that a correct diagnosis is made if no AI-advice is presented.
+Mathematically, we could describe this unit-specific quantity with some pseudo-equation like this (where $s$ stands for a specific expert radiologist and $i$ stands for a specific CT scan): -->
+
+<!-- $$ -->
+<!-- \begin{aligned} -->
+<!-- & P(\text{correct diagnosis} | \text{correct advice, specific expert, specific scan}) \\ -->
+<!-- & \quad - P(\text{correct diagnosis} | \text{no advice, specific expert, specific scan}) -->
+<!-- \end{aligned} -->
+<!-- $$ -->
+<!-- However, our verbal description is not precise enough yet, because the above quantity may differ for each combination of expert and scan.
+We must specify how we want to aggregate the quantity across different experts and scans.
+If we computed the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts (CITATION).
+Although often debated whether this is a good idea, psychological research often focuses on a different quantity: Instead of averaging over experts and scans, we focus on a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
+With this definition, the theoretical estimand in our above example can be expressed as: -->
+
+<!-- $$ -->
+<!-- \begin{aligned} -->
+<!-- & P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\ -->
+<!-- & \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) -->
+<!-- \end{aligned} -->
+<!-- $$ -->
+<!-- In total, our theoretical estimand consists of four such expressions, the remaining three are: -->
+
+<!-- $$ -->
+<!-- \begin{aligned} -->
+<!-- & P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\ -->
+<!-- & \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan}) -->
+<!-- \end{aligned} -->
+<!-- $$ -->
+<!-- $$ -->
+<!-- \begin{aligned} -->
+<!-- & P(\text{correct diagnosis} | \text{correct advice, average student, average scan}) \\ -->
+<!-- & \quad - P(\text{correct diagnosis} | \text{no advice, average student, average scan}) -->
+<!-- \end{aligned} -->
+<!-- $$ -->
+
+<!-- $$ -->
+<!-- \begin{aligned} -->
+<!-- & P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\ -->
+<!-- & \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan}) -->
+<!-- \end{aligned} -->
+<!-- $$ -->
+<!-- To complete our definition of the theoretical estimand, we have to define our target population that consists of both persons and scans: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
+With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
+Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system.
+-->
+
+<!-- While a theoretical estimand is defined outside of any statistical model, an empirical estimand is a quantity that can be estimated based on observed data.
 For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of head CT scans, but the kind of AI-advice given for each scan is randomly assigned within participants.
@@
 For pedagogical reasons, we will skip the concrete definition of our empirical estimand until we have discussed how to simulate data based on a concrete GLMM in the next section.
+-->
 
@@
 However, we would argue that even a strongly simplified data generating process (e.g.
-only a small number of interaction effects; only random intercepts and no random slopes; making rather implausible assumptions for missing data) can yield informative results.
+only a small number of interaction effects; only random intercepts and no random slopes; assuming that data is missing completely at random) can yield informative results.
 
@@
 
-In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
+In a GLMM, the expected value of the dependent variable $Y$ conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
 $$
@@
 In our case study, we have performed simulation-based sample size planning from a single set of parameter values that reflect our assumptions of an expected effect size.
-Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials, we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
+Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials [e.g., @kumleEstimatingPowerGeneralized2021], we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
 Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
@@
 A common, more conservative strategy for sample size justification is to perform sample size planning for the smallest effect size of interest (SESOI).
-An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011].
+An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011; @lakensEquivalenceTestingPsychological2018].
 For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].
@@
 By simulating power (or precision) for different effect sizes (in addition to the different number of subjects and items), one can make sure that power (or precision) would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power (or precision) depends on the size of the true effect.
-In simple study designs, it is possible to perform sensitivity analysis based on a single standardized effect size (e.g., analyze power in a two-sample t-test for a standardized mean difference varying between 0.1 and 0.8).
-However, for our case study that investigates combined hypotheses in a GLMM modeling framework, the effect size is implicitly represented by the complex distribution of probabilities within and between experimental conditions.
-In this setting, sensitivity analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
+For our case study that investigates combined hypotheses in a GLMM modeling framework, sensitivity analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
 Power (or precision) could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
 
-# Outlook
-
-<!-- Outlook -->
+# Conclusion and outlook
+
+The goal of this tutorial was to teach researchers how to perform tailored simulation-based sample size planning for GLMMs.
 Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based sample size planning in experimental research:
@@
 1.
-The growing need for simulation-based sample size planning in experimental research: In order to conduct informative research using varying complex experimental designs with GLMMs, formula-based heuristics and user-friendly software tools for a priori power analysis are often not suitable.
+In light of the ongoing replication crisis and an existing literature full of underpowered studies, there is a growing need for simulation-based sample size planning in experimental research: In order to conduct informative research, GLMMs offer a flexible statistical framework to analyze complex experimental study designs.
+However, existing formula-based heuristics and user-friendly software tools for a priori power analysis are often not sufficient.
 Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
@@
 4.
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform a priori simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
+Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform tailored simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
 Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
@@
 5.
-Recognizing the value of simulation-based design analysis: Simulation-based power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
+Recognizing the value of simulation-based design analysis: Tailored data simulations and power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
 They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power or precision.
@@
 
-# Conclusion
-
-In the wake of the replication crisis and myriad of underpowered experimental work, generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-Yet, analytic methods and software cannot be applied to conduct a priori sample size planning for GLMMs, necessitating data simulation-based approaches.
-Through this applied tutorial, we aim to provide researchers with the necessary skills and tools to perform simulation-based sample size planning with GLMMs themselves.
-By incorporating GLMMs and a priori sample size planning into their work, researchers can enhance the replicability and credibility of their experiments [@yarkoniGeneralizabilityCrisis2022].
-
 # References

### 2024-06-07 manuscript.qmd
extend estimand section

@@
 To benefit most of this tutorial paper, we recommend readers to familiarise themselves with basic statistical concepts like hypothesis tests (HTs) and their statistical power as well as confidence intervals (CIs) and their precision [@lakensImprovingYourStatistical2022; @riesthuisSimulationBasedPowerAnalyses2024; @kumleEstimatingPowerGeneralized2021].
-Some knowledge of causal inference is useful but not necessary [@lundbergWhatYourEstimand2021].
+Some knowledge of causal inference is useful but not necessary [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
 In addition, readers should have a rough understanding of R [@wickhamDataScienceImport2023] and how to simulate data.
@@
 Providing an informed standardized effect size can be an almost impossible challenge when no prior studies of pilot data are available.
-This problem is further exacerbated by the fact that GLMMs are so flexible, that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
+This problem is further exacerbated by the fact that GLMMs are so flexible that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
 In our experience, using domain knowledge to construct a tailored data simulation is the only solution to determine plausible effect sizes in the absence of prior evidence.
 It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
-However, we would argue that when tailored data simulations are necessary anyways, performing the whole sample size planning in a customized way is preferred over using the existing software packages.
+However, we would argue that when tailored data simulations necessary to determine effect sizes anyway, performing the whole sample size planning in a customized way is preferred over using the existing software packages.
 
@@
 
-The first step in every research process should be a clear definition of the *estimand* [@lundbergWhatYourEstimand2021], i.e.
-the theoretical quantity which is necessary to answer the specific research question.
+The first step in every research process should be a clear definition of the theoretical *estimand* [@lundbergWhatYourEstimand2021], i.e.
+the theoretical quantity which is necessary to answer a specific research question.
 The estimand consists of a quantity, that can be described for each unit under investigation and a clear definition of the target population, for which the quantity is of interest.
-
-The estimand should always be defined outside of any statistical model, because there are usually a range of statistical methods that can be used to estimate the same estimand based on the observed data that will be collected in the planned study.
-For many common research questions in psychology, the estimand can be expressed as a statistical quantity that can be estimated with a regression model.
-However, this is not always possible all causal estimands, which is why the literature discusses many estimation strategies beyond regression [@lundbergWhatYourEstimand2021].
+For example, an estimand might be the probability that a clinical psychologists makes the correct diagnosis for a psychiatric patient with major depression, averaged across all clinical psychologists and depressed patients in psychiatric institutions in Germany.
+
+A theoretical estimand is a quantifiable measure in statistical analysis that is designed to estimate the underlying parameter of interest within a theoretical model or framework.
+Essentially, it represents a defined characteristic or feature of the theoretical model.
+This estimand serves as a target for estimation and provides insight into the behavior and properties of the phenomena being studied within the theoretical construct.
+
+The estimand should always be defined outside of any statistical model, because there are usually a range of statistical methods that can be used to estimate the same estimand, depending on the study design (e.g., a randomized experiment) that will produce the observed data in the planned study.
+For many common research questions in psychology, the estimand can be expressed as a statistical quantity that can be estimated with a regression model, for example a single $\beta$ coefficient.
+However, this is not possible for all estimands, which is why the literature discusses many estimation strategies beyond regression [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
 
@@
 It becomes clear that our estimand consists of four comparisons between experimental conditions [@lundbergWhatYourEstimand2021].
-However, the verbal descriptions is not precise enough yet, which is why we try to give a unique mathematical expression:
+However, the verbal description is still somewhat vague, which is why we try to give a more precise expression for each comparison:
 
@@
 This contrast is quantified for a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.^[Note that a different estimand would be the so-called average treatment effect (ATE).
-For the ATE, the probability contrast would be computed for each expert and scan, and then these contrasts are averaged across all experts and scan in the target population [@lundbergWhatYourEstimand2021]].
-To complete our definition of the estimand, we have to define our target population that consists of both persons and scans: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
+For the ATE, the probability contrast is defined for each combination of expert and scan, and then these contrasts are averaged across all experts and scans from the target population [@lundbergWhatYourEstimand2021]].
+To complete our definition of the estimand, we have to define our target population that consists of persons, stimuli, and AI-advice: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
 With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
-Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system.
+Lastly, we are only interested in AI-advice given by a specific AI-enabled diagnostic decision support system.
 
@@
 This random intervention allows us to produce an empirical estimate of our estimand, although, in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each scan.
-We will see soon that each of the probability expressions in our estimand can be modeled with the same GLMM.
+We will see later how each of the probability expressions in our estimand can be modeled with the same GLMM.
 Estimating this GLMM based on the data observed in our planned study will produce an estimate for each probability, and these estimates can then be combined to compute an estimate for each of the four probability contrasts.
 For pedagogical reasons, we will skip the concrete definition of our estimate until we have discussed how to simulate data based on a concrete GLMM in the next section.
-
-<!-- It becomes clear that our estimand consists of four comparisons between experimental conditions [@lundbergWhatYourEstimand2021].
-To elaborate further, we focus on one part of the research question concerning the comparison between correct AI-advice and no AI-advice for expert radiologists first.
-For a specific expert radiologist and a specific CT scan, our question is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice is presented and the probability that a correct diagnosis is made if no AI-advice is presented.
-Mathematically, we could describe this unit-specific quantity with some pseudo-equation like this (where $s$ stands for a specific expert radiologist and $i$ stands for a specific CT scan): -->
-
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- & P(\text{correct diagnosis} | \text{correct advice, specific expert, specific scan}) \\ -->
-<!-- & \quad - P(\text{correct diagnosis} | \text{no advice, specific expert, specific scan}) -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-<!-- However, our verbal description is not precise enough yet, because the above quantity may differ for each combination of expert and scan.
-We must specify how we want to aggregate the quantity across different experts and scans.
-If we computed the average of our individual quantity across all experts and scans in the target population, the result would be called the average treatment effect (ATE) in the population of experts (CITATION).
-Although often debated whether this is a good idea, psychological research often focuses on a different quantity: Instead of averaging over experts and scans, we focus on a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.
-With this definition, the theoretical estimand in our above example can be expressed as: -->
-
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- & P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\ -->
-<!-- & \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-<!-- In total, our theoretical estimand consists of four such expressions, the remaining three are: -->
-
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- & P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\ -->
-<!-- & \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan}) -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- & P(\text{correct diagnosis} | \text{correct advice, average student, average scan}) \\ -->
-<!-- & \quad - P(\text{correct diagnosis} | \text{no advice, average student, average scan}) -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- & P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\ -->
-<!-- & \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan}) -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-<!-- To complete our definition of the theoretical estimand, we have to define our target population that consists of both persons and scans: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
-With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
-Similarly, we are only interested in AI-advice given by the specific AI-enabled diagnostic decision support system.
--->
-
-<!-- While a theoretical estimand is defined outside of any statistical model, an empirical estimand is a quantity that can be estimated based on observed data.
-For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of head CT scans, but the kind of AI-advice given for each scan is randomly assigned within participants.
-This random intervention allows us to estimate our theoretical estimand, although, in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each scan.
-For our example, each probability expression in our theoretical estimand can be modeled with the same GLMM.
-Estimating the GLMM based on the data observed in our planned study produces an estimate for each probability, and these estimates can be used to compute an estimate for each of the four probability contrasts.
-For pedagogical reasons, we will skip the concrete definition of our empirical estimand until we have discussed how to simulate data based on a concrete GLMM in the next section.
--->
 
@@
 The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
-Although insightful descriptive statistics often depend on the model parameters in a non-linear way, it is usually possible to manually adjust the population parameters (if only with the help of trial and error) until the model implied quantities produce the desired result.
+Although insightful descriptive statistics often depend on the model parameters in a non-linear way, it is not necessary to solve the exact relationship mathematically.
+Instead, one can simply adjust the population parameters by trial and error until the model implied quantities produce the desired result.
 

### 2024-06-09 manuscript.qmd
work on case study part

@@
 Available formula-based software packages for multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
-A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
-However, most of these tutorials focus on linear mixed models (LMMs) and the common study designs [but see @kumleEstimatingPowerGeneralized2021, for a tutorial that also covers more advanced settings].
+A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a].
+However, many of these tutorials focus on linear mixed models (LMMs) and the common study designs [but see @kumleEstimatingPowerGeneralized2021, for a tutorial that also covers more advanced settings].
 This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
@@
 
-- Insert table on existing R packages for power analysis with GLMMs
+In Table XXX, we give a short review of existing R packages that can be used for power analysis for GLMMs.
 
@@
 
-A theoretical estimand is a quantifiable measure in statistical analysis that is designed to estimate the underlying parameter of interest within a theoretical model or framework.
-Essentially, it represents a defined characteristic or feature of the theoretical model.
-This estimand serves as a target for estimation and provides insight into the behavior and properties of the phenomena being studied within the theoretical construct.
-
 The estimand should always be defined outside of any statistical model, because there are usually a range of statistical methods that can be used to estimate the same estimand, depending on the study design (e.g., a randomized experiment) that will produce the observed data in the planned study.
@@
 We will discuss strategies on how these values can be obtained later.
-Because we have full control over the data generating process in a tailored simulation, it is possible to model specific aspects of the planned study, like data missing by design or assuming that subjects drop out (i.e.
-deleting their data from the simulated dataset) based on other variables in the simulated dataset.
+Because we have full control over the data generating process in a tailored simulation, it is possible to model specific aspects of the planned study, like data missing by design or assuming that some subjects might drop out.
 The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
-For this reason, deciding on a specific set of assumptions can easily feel overwhelming.
 However, we would argue that even a strongly simplified data generating process (e.g.
@@
 The following R function simulates a full dataset structured according to the design of our case study.
-The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.
 
@@
 
-In the first four lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data. 
+In the first four lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data.^[The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.]
 In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).
@@
 
-Tailored simulation-based sample size planning is especially relevant when studies have to be planned in the absence of previous studies with the same design or pilot data.
-In this scenario, researchers require strategies on how to specify the population parameters used in their data simulation.
+In the absence of previous studies with the same design or pilot data, researchers require strategies on how to specify the population parameters used in their data simulation.
 Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
@@
 Our strategies to specify population parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge of domain experts can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
+Because most domain knowledge can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
 The basic idea of all strategies is how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
@@
 
-All parameter values in our present case study have been determined based on results from distantly related study designs in the literature.
+The starting point for all parameter values in our present case study were based on results from distantly related study designs in the literature.
 Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions seem plausible.
@@
 In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-The other extreme case would be the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented:
+For example, the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented, only requires the intercept:
 $$
@@
 @tab-probtable shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
-The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations:
+The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations (`b_0` stands for the intercept $\beta_0$, `b_e` for the slope $\beta_e$, and so on):
 
@@
 However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental conditions, i.e.
-which descriptive statistics they expect to observe (on average).
+which relative frequency they expect to observe (on average).
 The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
-Although insightful descriptive statistics often depend on the model parameters in a non-linear way, it is not necessary to solve the exact relationship mathematically.
+Although insightful descriptive statistics usually depend on the model parameters in a non-linear way, it is not necessary to solve the exact relationship mathematically.
 Instead, one can simply adjust the population parameters by trial and error until the model implied quantities produce the desired result.
@@
 For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice (the conditional perspective).
-However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty (the marginal perspective).
+However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability, and items drawn from their natural distribution of difficulty (the marginal perspective).
 How much the two probabilities differ depends on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations would be zero).
@@
 We tried using these descriptive statistics to judge together with domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
-However, although the result was deemed plausible, in our example these statistics are not informative enough to determine a final set of plausible parameter values.
+However, although the result was deemed plausible, these statistics were not informative enough to determine a final set of plausible parameter values.
 For this reason, we will additionally look at insightful model based quantities.
@@
 Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
-This is especially true for the standard deviation of random effects that have complicated (and often unexpected) effects on the model-implied results.
+This is especially true for the standard deviation of random effects that have non-linear (and often unexpected) effects on the model-implied results.
 An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
@@
 The discussed inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
-One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans).
+One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often radiologists make an accurate diagnosis based on brain scans).
 However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
@@
 Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
-
 We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
@@
 This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-However, these high standard deviations do not account for the inherent variability and complexity of human performance.
+These high standard deviations do not account for the inherent variability and complexity of human performance.
 For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.

### 2024-06-10 manuscript.qmd
work on steps 4 and 5

@@
 On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
-On the other hand, using a more complex model (that perhaps even goes beyond a standard GLMM) for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
+On the other hand, using a more complex model for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
 
@@
 
+<!-- The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
+-->
+
+<!-- [^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017]. -->
+
+<!-- However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
+Therefore, for our case study, we have written our own function to simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions).
+-->
+<!-- We think this choice assists applied researchers in better understanding all model assumptions before using lme4 to analyze the simulated datasets.[^2] -->
+
+<!-- [^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
+-->
+
 In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
-The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
-
-[^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].
-
-However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore, for our case study, we have written our own function to simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions).
-We think this choice assists applied researchers in better understanding all model assumptions before using lme4 to analyze the simulated datasets.[^2]
-
-[^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
-
-In this section, we show how to fit a GLMM with lme4.
-We simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
+For the current example, we simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
 However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.
@@
 
-When the theoretical and empirical estimands have been identified in the modeling framework used to estimate the GLMM, one has to decide on whether hypothesis testing or interval estimation should be used to answer the research question.
-While the majority of psychological research uses hypothesis testing, there are repeated claims that a focus on estimation can also be a good choice in many circumstances (CITATION).
-This decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with the marginaleffects R package using the delta method).
+In previous steps, we have defined the theoretical estimand, written a data simulation function and specified how to estimate a GLMM using simulated data.
+The next step is to specify how to compute a concrete point estimate of the theoretical estimand within the framework of the fitted GLMM.
+For some research question, the estimate corresponds with a single regression coefficient. In more complicated scenarios, the estimate is computed from a combination of coefficients.
+Beyond computing the point estimate, we have already discussed that both hypothesis testing and interval estimation can be used to answer the research question.
+The decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with the marginaleffects R package using the delta method).
 
@@
 
-For our case study, we investigate the following research question:
-*We expect that for both experts and non-experts, correct advice leads to a higher probability of accurately diagnosing a CT scan compared to no advice presented, AND, we expect that for both experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
-
-In the earlier section on our theoretical estimand we translated this verbal description into four probability statements that are specified outside of any specific statistical model.
-The next translation step is to specify the empirical estimand, that is how these probabilities can be estimated within the scope of our concrete GLMM.
-
-The first of the four statements of the theoretical estimand was:
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
-\end{aligned}
-$$
-In the context of our GLMM, the corresponding probability statement is:
+In the estimand section, we have translated a verbal description of our research question into four probability statements that are specified outside of any specific statistical model.
+For a concrete estimate within the context of our specified GLMM, we must compute the following probability contrasts:
 
@@
 $$
-Plugging in the model equation of the GLMM produces an equation on how to compute this contrast when all model parameters are known.
-We have already discussed how to compute the individual probabilities in the section on specifying population parameters.
-When we want to estimate the above contrast based on observed data, the only difference is that model parameters are not known and we instead plug the corresponding parameter estimates into the equation.
+$$
+\begin{aligned}
+& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
+& \quad - P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
+& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
+& \quad - P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
+\end{aligned}
+$$
+We have already discussed how to compute the involved probabilities in the section on specifying population parameters.
+Plugging in the model equation of the GLMM produces an equation on how to compute each contrast if all model parameters were known.
+When we want to estimate the above contrasts based on *observed* data, the only difference is that model parameters are not known and we instead use the corresponding parameter *estimates*.
 
@@
 
-We could use our knowledge of the structure of our GLMM to determine the exact formula needed to compute our contrast of interest and then plug in the parameter estimates manually from the `summary(fit)` output.
-However, this would be very tedious and we can use R to compute this contrast without doing the math.
-
-We could use the `predict` function of the lme4 package to compute the predicted probability for a correct diagnosis based on our model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
-
-```{r}
+We could use our knowledge of the structure of the GLMM to determine the exact formula needed to compute the contrasts of interest and then plug in the parameter estimates manually from the `summary(fit)` output.
+However, this would be tedious and we can use R to compute this contrast without doing the math.
+
+Using the first contrast as our example, we could apply the `predict` function of the lme4 package to compute the predicted probability for a correct diagnosis based on our fitted model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
+
+```{r}
+#| echo: true
 pred_df <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), 
@@
 The argument `type = "response"` specifies that predictions are made on the probability scale (instead of the log-odds scale of the $\beta$ parameters), while `re.form = NA` sets all random effects to 0.
-
-We could use this method to compute point estimates for all four contrast that are part of our empirical estimand.
-However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute HTs or CI for our point estimates.
+We could use this method to compute point estimates for all four contrasts that are part of our estimand.
+However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute HTs or CI.
 
 The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute HTs and CIs for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
-
 First, we specify a grid of all combinations of predictor variable and then compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
@@
 ```{r}
+#| echo: true
 library(tidyverse)
@@
 This is no problem as long as we never interpret those estimates.
-
-Next, we use the estimated probabilities to compute the four specific contrasts that are part of our empirical estimand.
+Next, we use the estimated probabilities to compute the four specific contrasts that are part of our estimand.
 For this we must specify which rows in `probs` have to be subtracted from each other.
@@
 
-The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2.
+The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2 in the `probs`-output.
 The argument `equivalence = c(0, 0)` can be used to compute one-sided p-values, testing whether the contrast in the population is smaller than 0 (`p (NonSub)` column) or greater than 0 (`p (NonInf)` column).
@@
 Thirdly, the results of the repeated simulation must be collected and aggregated.
-When planning for power, we compute the relative frequency of a significant p-value(s) across repeated simulations.
+When planning for power, we compute the relative frequency of (a) significant p-value(s) across repeated simulations.
 When planning for precision, we compute the average width of the CI(s).

### 2024-06-10 manuscript.qmd
some formatting

@@
 author:
+  - name: Florian Pargent
+    corresponding: true
+    equal-contributor: true
+    orcid: 0000-0002-2388-553X
+    roles:
+      - formal analysis
+      - methodology
+      - visualization
+      - writing
+      - editing
+    affiliations:
+      - id: "lmu"
+        name: LMU Munich
+        department: Department of Psychology
   - name: Timo K. Koch
@@
         postal-code: 9000
-      - id: "lmu"
-        name: LMU Munich
-        department: Department of Psychology
-  - name: Florian Pargent
-    equal-contributor: true
-    orcid: 0000-0002-2388-553X
-    roles:
-      - formal analysis
-      - methodology
-      - visualization
-      - writing
-      - editing
-    affiliations:
       - ref: lmu
@@
     gratitude: ~
-    authorship-agreements: "Timo K. Koch and Florian Pargent contributed equally to this work."
+    authorship-agreements: "Florian Pargent and Timo K. Koch contributed equally to this work."
 abstract: "When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
@@
 However, available methods and software for a priori sample size planning for GLMMs are often limited to specific designs.
-Data simulation approaches as an alternative offer much more flexibility.
-Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori sample size planning to determine appropriate sample sizes with GLMMs.
-Finally, we give an outlook on the increasing importance of simulation-based power analysis in experimental research."
+Tailored data simulation approaches are a more flexible alternative.
+Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori sample size planning with GLMMs.
+We not only focus on power analysis but also explain how to use the precision of parameter estimates to determine appropriate sample sizes.
+We conclude with an outlook on the increasing importance of simulation-based sample size planning."
 keywords: [power analysis, precision, data simulation, sample size planning, generalized linear mixed model]
@@
 ```{r}
+#| echo: true
 simulate <- function(n_subjects = 100, n_items = 50,
@@
   # simulate design
-  dat <- add_random(subject = n_subjects, item = n_items) %>%
-    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
-    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
+  dat <- add_random(subject = n_subjects, item = n_items) |>
+    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) |>
+    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) |>
     mutate(advice_correct = if_else(advice_present == 1L, 
-                                    rbinom(n(), 1L, prob = 0.8), 0L)) %>%
+                                    rbinom(n(), 1L, prob = 0.8), 0L)) |>
     # add random effects
-    add_ranef("subject", u0s = sd_u0s) %>%
-    add_ranef("item", u0i = sd_u0i) %>%
+    add_ranef("subject", u0s = sd_u0s) |>
+    add_ranef("item", u0i = sd_u0i) |>
     # compute dependent variable
@@
         b_e * expert + b_a * advice_present + b_c * advice_correct +
-        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
-    mutate(y_prob = plogis(linpred)) %>%
+        b_ea * expert * advice_present + b_ec * expert * advice_correct) |>
+    mutate(y_prob = plogis(linpred)) |>
     mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
@@
 ```{r}
+#| echo: true
 b_0 <- qlogis(0.7)
@@
 ```{r}
+#| echo: true
 plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec)
@@
   sd_u0s = 0.5, sd_u0i = 0.5)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
     "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
-  group_by(condition) %>%
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
+  group_by(condition) |>
   summarize(relative_frequency = sum(y_bin) / n())
@@
 #| fig-cap: "Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions."
-
+#| echo: true
 library(ggdist)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
     "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
   ggplot(aes(x = y_prob, y = condition)) +
@@
 dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0i = 0.01)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
 dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
 dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 3, sd_u0i = 3)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
 #| message: false
+#| echo: true
 library(tidyverse)
@@
 #| message: false
+#| echo: true
 library(lme4)
@@
 ```{r}
+#| echo: true
 fit <- glmer(f, data = dat, family = "binomial")
@@
 ```{r}
+#| echo: true
 summary(fit)
@@
 ```{r}
-probs %>% 
+#| echo: true
+probs |> 
   hypotheses(hypothesis = c(
@@
 ```{r}
+#| echo: true
 sim_and_analyse <- function(
@@
     expert = 0:1)
-  predictions(model, newdata = contr_df, type = "response", re.form = NA) %>%
-    hypotheses(hypothesis = contrasts, equivalence = c(0, 0)) %>%
+  predictions(model, newdata = contr_df, type = "response", re.form = NA) |>
+    hypotheses(hypothesis = contrasts, equivalence = c(0, 0)) |>
     data.frame()
@@
 #| eval: false
+#| echo: true
 library(future)
@@
   n_items = c(10, 30, 50, 70)
-) %>%
+) |>
   mutate(res = future_pmap(., sim_and_analyse, 
-    .options = furrr_options(seed = TRUE))) %>%
+    .options = furrr_options(seed = TRUE))) |>
   unnest(col = res)
@@
 #| eval: false
+#| echo: true
 library(furrr)
@@
 #| fig-cap: "Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."
+#| echo: true
 library(binom)
 alpha <- 0.05
-power <- sim_result %>%
+power <- sim_result |>
   pivot_wider(names_from = term, names_sep = "_", 
-    values_from = estimate:p.value.equiv) %>%
-  group_by(n_subjects, n_items) %>% 
+    values_from = estimate:p.value.equiv) |>
+  group_by(n_subjects, n_items) |> 
   summarise(
@@
     .groups = "drop")
-power %>%
-  mutate(across(c(n_subjects, n_items), factor)) %>%
+power |>
+  mutate(across(c(n_subjects, n_items), factor)) |>
   ggplot(aes(n_subjects, n_items, fill = power)) +
@@
 #| fig-cap: "Simulation-based precision estimates (expected width of confidence intervals) including 95% confidence interval of the case study for different numbers of subjects and items, based on a confidence level of 0.95."
-precision <- sim_result %>%
+#| echo: true
+precision <- sim_result |>
   pivot_wider(names_from = term, names_sep = "_", 
-    values_from = estimate:p.value.equiv) %>%
-  group_by(n_subjects, n_items) %>%
-  mutate(width = `conf.high_b8=b2` - `conf.low_b8=b2`) %>%
+    values_from = estimate:p.value.equiv) |>
+  group_by(n_subjects, n_items) |>
+  mutate(width = `conf.high_b8=b2` - `conf.low_b8=b2`) |>
   summarise(precision = mean(width),
@@
     .groups = "drop")
-precision %>%
-  mutate(across(c(n_subjects, n_items), factor)) %>%
+precision |>
+  mutate(across(c(n_subjects, n_items), factor)) |>
   ggplot(aes(n_subjects, n_items, fill = precision)) +
@@
 
-1.
 In light of the ongoing replication crisis and an existing literature full of underpowered studies, there is a growing need for simulation-based sample size planning in experimental research: In order to conduct informative research, GLMMs offer a flexible statistical framework to analyze complex experimental study designs.
@@
 
-2.
 Managing data simulations more easily with discrete predictor variables: Simulation-based sample size planning becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
@@
 
-3.
 Teaching data simulation skills: The ability to conduct simulation-based sample size planning is a valuable skill that should be taught to experimental researchers.
@@
 
-4.
 Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform tailored simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
@@
 
-5.
 Recognizing the value of simulation-based design analysis: Tailored data simulations and power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
@@
 
-6.
 Integration with Open Science and preregistration practices: Simulation-based sample size planning aligns well with the principles of Open Science and preregistration in experimental research.

### 2024-06-11 full_simulation.qmd
update bibfile, add full_simulation.qmd notebook, change notebook-links option

--- 2024-06-10 manuscript.qmd
@@
 ---
-title: "A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models"
-shorttitle: "Simulation-Based Sample Size Planning for GLMMs"
-author:
-  - name: Florian Pargent
-    corresponding: true
-    equal-contributor: true
-    orcid: 0000-0002-2388-553X
-    roles:
-      - formal analysis
-      - methodology
-      - visualization
-      - writing
-      - editing
-    affiliations:
-      - id: "lmu"
-        name: LMU Munich
-        department: Department of Psychology
-  - name: Timo K. Koch
-    corresponding: true
-    equal-contributor: true
-    orcid: 0000-0001-6728-2063
-    email: timo.koch@unisg.ch
-    roles:
-      - conceptualization
-      - formal analysis
-      - writing
-      - editing
-    affiliations:
-      - name: University of St. Gallen 
-        department: Institute of Behavioral Science & Technology
-        address: Torstrasse 25
-        city: St. Gallen
-        country: Switzerland
-        postal-code: 9000
-      - ref: lmu
-  - name: Anne-Kathrin Kleine
-    roles:
-      - formal analysis
-      - editing
-    affiliations:
-      - ref: lmu
-  - name: Eva Lermer
-    roles:
-      - conceptualization
-      - funding aquisition
-      - editing
-    affiliations:
-      - ref: lmu
-      - name: Technical University of Applied Sciences Augsburg
-        department: Department of Business Psychology
-  - name: Susanne Gaube
-    roles:
-      - conceptualization
-      - supervision
-      - editing
-    affiliations:
-      - ref: lmu
-      - name: University College London
-        department: Global Business School for Health
-author-note:
-  disclosures:
-    data-sharing: "All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF) at https://osf.io/dhwf4/"
-    conflict-of-interest: "The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article."
-    financial-support: "This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525)."
-    gratitude: ~
-    authorship-agreements: "Florian Pargent and Timo K. Koch contributed equally to this work."
-abstract: "When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
-The recent replication crisis underlines the need for more rigorous statistical methodology and adequately powered designs.
-Generalized linear mixed models (GLMMs) offer a flexible statistical framework to analyze experimental data with complex (e.g., dependent and hierarchical) data structures.
-However, available methods and software for a priori sample size planning for GLMMs are often limited to specific designs.
-Tailored data simulation approaches are a more flexible alternative.
-Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori sample size planning with GLMMs.
-We not only focus on power analysis but also explain how to use the precision of parameter estimates to determine appropriate sample sizes.
-We conclude with an outlook on the increasing importance of simulation-based sample size planning."
-keywords: [power analysis, precision, data simulation, sample size planning, generalized linear mixed model]
-bibliography: [Simulation Tutorial Paper.bib, r-references.bib]
-floatsintext: true
-numbered-lines: false
-csl: apa.csl
+title: "Full simulation"
 ---
 
-```{r}
-#| label: setup
-#| include: false
-library("papaja")
-library("tidyverse")
-library("faux")
-library("lme4")
-library("multcomp")
-library("marginaleffects")
-library("future")
-library("furrr")
-library("binom")
-r_refs("r-references.bib")
-```
+::: {.callout-note}
+This Notebook is used to run the full simulation for the case study discussed in the manuscript. The simulation takes several hours to run on a Mac Book Pro (13-inch, M1, 2020) with 16 GB memory.
+:::
 
-# Introduction
-
-When planning experimental research, it is essential to determine an appropriate sample size and to use appropriate statistical models to analyze the data to ensure that the results obtained are both robust and informative [@lakensSampleSizeJustification2022].
-The recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of study findings [@yarkoniGeneralizabilityCrisis2022].
-As a result, there is a growing need for more rigorous statistical methodology and the adoption of adequately powered experimental designs.
-Multiple easy-to-use software solutions exist for simple statistical models and experimental designs.
-However, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) sample size planning for more complex research designs such as flexible generalized linear mixed models (GLMM) framework.
-In the present work, we provide a tutorial on how to determine adequate sample sizes by performing tailored simulation-based sample size planning for GLMMs.
-After introducing some theoretical background on sample size planning, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
-Then we describe general steps and decisions involved in tailored data simulation.
-To illustrate the details of these steps, we finish with a hypothetical case study from the field of human-AI (artificial intelligence) interaction research.
-
-<!-- # Recommended background knowledge -->
-
-To benefit most of this tutorial paper, we recommend readers to familiarise themselves with basic statistical concepts like hypothesis tests (HTs) and their statistical power as well as confidence intervals (CIs) and their precision [@lakensImprovingYourStatistical2022; @riesthuisSimulationBasedPowerAnalyses2024; @kumleEstimatingPowerGeneralized2021].
-Some knowledge of causal inference is useful but not necessary [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
-In addition, readers should have a rough understanding of R [@wickhamDataScienceImport2023] and how to simulate data.
-For data simulation, we use functions from the tidyverse [@wickhamWelcomeTidyverse2019] and the faux package [@R-faux].
-Finally, readers should be familiar with regression modeling in general and GLMMs in particular.
-In this tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process [@debruineUnderstandingMixedEffectsModels2021].
-It is not necessary to understand the technical details of how GLMMs are estimated.
-However it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the values in the dependent variable are determined by the predictor variables and the random effects.
-
-# Theoretical background
-
-## Planning for statistical power or precision
-
-Conducting research with insufficiently large sample sizes can have many negative consequences.
-First, experiments may yield inconclusive or misleading results, hindering the accumulation of knowledge.
-Second, studies that are doomed to never finding a postulated effect waste resources by consuming time, effort, and funding without delivering meaningful results.
-For these reasons, many journals and funding bodies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-Although scientists often do not justify sample size or use general heuristics from the literature, resource constraints often play an important role [@lakensSampleSizeJustification2022].
-But ideally, a suitable sample size should be determined a priori (i.e., before the study is conducted) based on some meaningful computation to ensure that the study will be able to fulfill its purpose.
-
-The majority of empirical studies in psychology and other social sciences apply hypothesis testing.
-As a consequence, the dominant approach for determining an adequate sample size is based on power analysis (i.e., planning for power) [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
-Statistical power is defined as the probability that a HT has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size [@cohenPowerPrimer1992].
-Less formally, power is described as the probability that a HT correctly rejects the null hypothesis when the alternative hypothesis is true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study is considered "underpowered".
-When planning for power, a target is set for the statistical power of a HT of interest.
-Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
-Although it is recommended to justify the desired significance level and power [@lakensJustifyYourAlpha2018], most empirical studies adopt the heuristic of $\alpha = 0.05$ and $1 - \beta = 0.80$.
-
-In contrast to power analysis, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision or planning for accuracy) [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
-Not all research questions are best answered using hypothesis testing.
-It has been argued that basic research rarely requires making discrete decisions on whether some effect has been "discovered" and should thus shift from a hypothesis testing towards an estimation framework [@mcelreathStatisticalRethinkingBayesian2020; @cummingNewStatisticsWhy2014; @kruschkeBayesianNewStatistics2018].
-Although this view is not without critique [@uyguntuncEpistemicPragmaticFunction2023], at least for exploratory or pilot studies where little previous research has been conducted, more scientists seem to agree that simply estimating the effects of interest and making the estimation uncertainty transparent by reporting CIs is more useful.
-Assuming that no HTs are conducted for a planned study, power analysis is not relevant for sample size planning.
-Nonetheless, the sample size still has a crucial effect on how informative the planned study will be, because an effect of interest is estimated more precisely with bigger samples.
-In the precision framework, the target quantity commonly used for sample size planning is the expected width of a CI [@kelleySampleSizePlanning2006; @lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
-The values inside a CI are often interpreted as plausible values for the quantity of interest it is supposed to estimate.
-More formally, a CI with confidence level 0.95 provides the smallest interval with the property that upon repeated sampling, 95% of individual CIs would include the true quantity of interest.
-Thus, a narrow CI is more informative about the size of the true effect than a wide interval.
-Apart from the desired confidence level, the width of a CI depends strongly on the sample size.
-Because bigger samples carry more information, they lead to smaller CIs.
-When planning for precision, a target can be set for the expected width of a CI of interest.
-Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
-Because planning for precision is still rare, there are no common heuristics on how to choose the desired width of the CI [@lakensSampleSizeJustification2022].
-
-## Generalized linear mixed models (GLMMs)
-
-As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by them [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity because they offer great flexibility [@fahrmeirRegressionModelsMethods2021].
-GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data, including hierarchical structures [@fahrmeirRegressionModelsMethods2021].
-In this context, correlated data means that the value in the outcome variable for one observation may be related to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
-This correlation can arise for various reasons: For instance, responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), participants come from the same neighborhood (clustering), or participants responded to the same stimulus (stimulus effects).
-Thus, modeling such correlations is important whenever the data has a clear structure, while the grouping variables can be hierarchically nested (e.g., grouping variables students and schools: each student belongs to exactly one school) or cross-classified (e.g., grouping variables students and math exercises: each student is presented with several math exercises).
-LMMs are used when the outcome variable is continuous and follows a normal distribution.
-They allow for the modeling of fixed effects, which capture the relationships between the usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
-GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
-GLMMs involve a link function that connects the linear combination of predictor variables to the expected value of the outcome variable.
-The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
-For example, think of an experiment with different design factors (e.g., picture positions, headline aesthetics) impacting the likelihood of users clicking on an online advertisement.
-The participants’ behavior is measured repeatedly over several sessions.
-The click patterns of participants in one session are likely to be correlated with their previous sessions and the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
-
-# Simulation-based sample size planning with GLMMs
-
-<!-- For less complex statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for a priori sample size planning is readily available [@faulStatisticalPowerAnalyses2009; @champelyPackagePwr2018; @kelleyObtainingPowerObtaining2003; @kelleySampleSizePlanning2006].
--->
-<!-- However, these software packages are often not flexible enough to perform sample size planning for complex designs. -->
-
-To our knowledge, existing approaches for sample size planning for GLMMs have exclusively focused on planning for power.
-Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on formulas to calculate power directly while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., what percentage of simulated datasets produces a significant p-value).
-Available formula-based software packages for multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
-A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a].
-However, many of these tutorials focus on linear mixed models (LMMs) and the common study designs [but see @kumleEstimatingPowerGeneralized2021, for a tutorial that also covers more advanced settings].
-This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
-Simulation-based power analysis with GLMMs requires making a range of assumptions: The (conditional) distribution assumption specifies the distributional family for the outcome variable.
-Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e., if and how they are correlated).
-Making these decisions requires understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
-Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge (see the discussion in [@maxwellSampleSizePlanning2008] and [@kumleEstimatingPowerGeneralized2021]).
-
-In Table XXX, we give a short review of existing R packages that can be used for power analysis for GLMMs.
-
-# When to use tailored data simulation?
-
-Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in TABLE XXX.
-The most important circumstances under which tailored simulation-based sample size planning is necessary are 1) complex study designs, 2) complex statistical hypotheses, 3) planning for precision, 4) no available prior studies or pilot data.
-
-<!-- ## Complex study designs -->
-
-Requirements of real-world studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
-One frequent issue in applied data analysis is missing data, and there can be various reasons for this [@littleStatisticalAnalysisMissing2014].
-For example, data can be missing completely at random (e.g., because an electronic measurement device randomly failed for some technical reasons).
-Alternatively, subjects might systematically drop out or produce missing data, but this dropout can be explained by some attributes also measured in the dataset (e.g., older subjects have a higher probability to refuse answering a question on income).
-In a more complicated scenario, missing data in some variable is caused by the measured attribute itself (e.g., wealthy people are more likely to refuse reporting their income).
-Moreover, many experimental designs contain conditions in which values of the predictor variables are missing by design.
-This can make data analysis more complicated because predictors have to be coded in specific ways that prevent the estimated GLMM from becoming unidentified.
-Whether missing data has an effect on the sample size planning depends on our theoretical assumptions on how the missingness is caused.
-However, it is often challenging to decide whether missing data can be safely ignored in the data analysis and sample size planning process based on a merely theoretical approach [@gomilaMissingDataExperiments2022].
-Tailored simulation-based approaches offer the possibility to include the assumed process of how data become missing in the data simulation, thereby determining the required sample based on simulated datasets that contain missing values [for an example, see @lanePowerStrugglesEstimating2018].
-As a byproduct, the simulated datasets can also be used to test whether the intended data analysis provides the expected (unbiased) results, despite the missing data.
-Although GLMMs can handle a large variety of outcome variables, researchers are becoming increasingly aware that many datasets might profit from even more sophisticated models.
-Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects that can be modeled with the R packages glmmTMB [@brooksGlmmTMBBalancesSpeed2017] or brms [@burknerAdvancedBayesianMultilevel2018].
-Tailored simulation-based approaches do not share the same limitation than the existing software solutions for power analysis that focus exclusively on GLMMs.
-As long as there is a software package available to estimate the model of interest, it is always possible to perform tailored simulation-based sample size planning.
-
-<!-- ## Complex statistical hypotheses -->
-
-The most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is a slope or intercept of a regression model.
-However, many research questions in psychology actually require testing more complex statistical hypotheses.
-In the new era of preregistration and registered reports [@chambersPresentFutureRegistered2022], most research questions should be tested with directed hypotheses because good theories at least postulate whether some psychological effect of interest is positive or negative.
-Even better theories should be able to specify the smallest effect sizes of interest (SESOI) that must be exceeded if the effect has any practical relevance [@lakensEquivalenceTestingPsychological2018].
-In combination, this might require a test such as $H_0: \beta \leq 0.1$.
-More elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example testing simple slopes [@preacherComputationalToolsProbing2006] with a hypothesis such as $H_0: \beta_0 + \beta_1 \leq 0$.
-If the research question consists only of a single hypothesis of this sort, it might be possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
-However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypothesis [for a tutorial on contrast analysis in GLMMs, see @schadHowCapitalizePriori2020].
-For example, a combined null hypothesis $H_0$ might consist of two single null hypotheses $H_{01}: \beta_1 \leq 0$ and $H_{02}: \beta_0 + \beta_1 \leq 0$.
-For some research questions, the combined null hypothesis $H_0$ would be rejected if both $H_{01}$ *AND* $H_{02}$ are rejected.
-For other research questions, the combined null hypothesis $H_0$ would be rejected if $H_{01}$ *OR* $H_{02}$ *OR* both are rejected.
-If the global hypothesis $H_0$ is combined with *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis [@dmitrienkoTraditionalMultiplicityAdjustment2013].
-However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
-None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here and only some can handle directed hypotheses.
-In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation-based sample size planning.
-
-<!-- ## Planning for precision -->
-
-Although planning for precision [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008; @cummingNewStatisticsWhy2014] has been increasingly discussed as a useful strategy for empirical research, all available software packages for sample size planning with GLMMs are based on power analysis.
-Therefore, researchers that want to want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE XXX.
-<!-- As long as one is interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
-However, we demonstrate in our case study that in realistic research projects with GLMMs, the contrasts of interests and their respective confidence intervals are usually situated on the scale of the (expected) outcome in contrast to the scale of the model parameters.
--->
-However, tailored simulation-based sample size planning can easily handle the planning for precision approach [@maxwellSampleSizePlanning2008].
-The only change in procedure is that instead of computing HTs for each simulated dataset and estimating statistical power across repetitions, CIs are computed for each simulated dataset, and the expected width is estimated.
-
-<!-- ## No prior studies or pilot data -->
-
-All frameworks for sample size planning require the user to make assumptions about the expected effect size.
-Assuming the true effect is of this size (or greater), one can compute the (minimum) power of a HT or the (maximum) expected width of a CI.
-Existing software packages for sample size planning for GLMMs usually require to provide the assumed effect in the unit of some standardized measure of effect size.
-When the researcher has access to similar studies or pilot data, providing such standardized effect sizes is feasible.
-However, note that choosing effect sizes based on small pilot studies is generally not recommended, as those estimates can be heavily biased [@albersWhenPowerAnalyses2018a; @lakensSampleSizeJustification2022].
-Providing an informed standardized effect size can be an almost impossible challenge when no prior studies of pilot data are available.
-This problem is further exacerbated by the fact that GLMMs are so flexible that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
-In our experience, using domain knowledge to construct a tailored data simulation is the only solution to determine plausible effect sizes in the absence of prior evidence.
-It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
-However, we would argue that when tailored data simulations necessary to determine effect sizes anyway, performing the whole sample size planning in a customized way is preferred over using the existing software packages.
-
-# General steps in tailored simulation-based sample size planning
-
-Although the details differ depending on the specific study characteristics, each tailored simulation-based sample size planning requires a series of steps and decisions.
-We will introduce each step in a theoretical section, followed by the practical application based on a hypothetical case study.
-All code in this manuscript and simulation results are available in the project’s repository on the Open Science Framework (https://osf.io/dhwf4/).
-
-## Step 1: Define the estimand
-
-### THEORY
-
-The first step in every research process should be a clear definition of the theoretical *estimand* [@lundbergWhatYourEstimand2021], i.e.
-the theoretical quantity which is necessary to answer a specific research question.
-The estimand consists of a quantity, that can be described for each unit under investigation and a clear definition of the target population, for which the quantity is of interest.
-For example, an estimand might be the probability that a clinical psychologists makes the correct diagnosis for a psychiatric patient with major depression, averaged across all clinical psychologists and depressed patients in psychiatric institutions in Germany.
-
-The estimand should always be defined outside of any statistical model, because there are usually a range of statistical methods that can be used to estimate the same estimand, depending on the study design (e.g., a randomized experiment) that will produce the observed data in the planned study.
-For many common research questions in psychology, the estimand can be expressed as a statistical quantity that can be estimated with a regression model, for example a single $\beta$ coefficient.
-However, this is not possible for all estimands, which is why the literature discusses many estimation strategies beyond regression [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
-
-<!-- When the empirical estimand has been specified, the last step is to define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
-When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique and software package are used to fit the model and how to compute HTs or CIs.
--->
-
-### PRACTICE
-
-In the present hypothetical case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) embedded in a diagnostic decision support system.
-The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography (CT).
-In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans who are required to make the final diagnostic decision.
-The research goal is to validate the effectiveness of the AI-enabled advice.
-We consider the AI-enabled advice as effective, if the following pattern holds, which we will first describe verbally:
-
-*We expect that for BOTH expert radiologists and medical students, correct AI-advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI-advice presented, AND, we expect that for BOTH experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
-
-It becomes clear that our estimand consists of four comparisons between experimental conditions [@lundbergWhatYourEstimand2021].
-However, the verbal description is still somewhat vague, which is why we try to give a more precise expression for each comparison:
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
-\end{aligned}
-$$
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan})
-\end{aligned}
-$$
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{correct advice, average student, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{no advice, average student, average scan})
-\end{aligned}
-$$
-
-$$
-\begin{aligned}
-& P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan})
-\end{aligned}
-$$
-For example, the first expression is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice is presented and the probability that a correct diagnosis is made if no AI-advice is presented.
-This contrast is quantified for a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.^[Note that a different estimand would be the so-called average treatment effect (ATE).
-For the ATE, the probability contrast is defined for each combination of expert and scan, and then these contrasts are averaged across all experts and scans from the target population [@lundbergWhatYourEstimand2021]].
-To complete our definition of the estimand, we have to define our target population that consists of persons, stimuli, and AI-advice: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
-With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
-Lastly, we are only interested in AI-advice given by a specific AI-enabled diagnostic decision support system.
-
-Although the estimand is initially defined outside of any statistical model, it is only useful if we find a way to estimate it based on observed data.
-For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of head CT scans, but the kind of AI-advice given for each scan is randomly assigned within participants.
-This random intervention allows us to produce an empirical estimate of our estimand, although, in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each scan.
-We will see later how each of the probability expressions in our estimand can be modeled with the same GLMM.
-Estimating this GLMM based on the data observed in our planned study will produce an estimate for each probability, and these estimates can then be combined to compute an estimate for each of the four probability contrasts.
-For pedagogical reasons, we will skip the concrete definition of our estimate until we have discussed how to simulate data based on a concrete GLMM in the next section.
-
-## Step 2: Simulate the data generating process
-
-#### THEORY
-
-When the estimand has been defined, the next step in the research process is to write code that simulates the (assumed) data generating process of the planned study.
-This requires specifying a generative process for all predictor variables used in the final data analysis.
-While such assumptions can be quite challenging for observational studies or continuous predictor variables, this is less of a problem for experimental studies with only categorical predictor variables.
-When all predictor variables have been simulated, one can use the structure of a suitable GLMM to simulate the dependent variable.
-To simulate the GLMM, one requires plausible values for all model parameters.
-We will discuss strategies on how these values can be obtained later.
-Because we have full control over the data generating process in a tailored simulation, it is possible to model specific aspects of the planned study, like data missing by design or assuming that some subjects might drop out.
-The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
-However, we would argue that even a strongly simplified data generating process (e.g.
-only a small number of interaction effects; only random intercepts and no random slopes; assuming that data is missing completely at random) can yield informative results.
-
-#### PRACTICE
-
-In our case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated.
-Radiologists (task experts) and students/interns (non experts) review a series of head computer tomography (CT) scans to assess the presence of a bleeding.
-An AI model provides initial diagnostic advice to assist their decision-making.
-In the control condition, no AI advice is presented.
-When AI advice is given, this advice can be either correct or incorrect.
-The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across brain scans.
-After reviewing a CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
-This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible.
-The goal of simulation-based sample size planning is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power or precision in the planned experiment.
-
-##### Our specific GLMM
-
-In a GLMM, the expected value of the dependent variable $Y$ conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
-$$
-g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
-$$
-Equivalently, the conditional expected value is modeled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$: 
-$$
-E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
-$$
-If the dependent variable (i.e., diagnostic decision) $Y$ is a binary variable with values $0$ (i.e., inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability:
-$$
-P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
-$$
-In our case study, $P_{si}$ is the conditional probability that subject $s$ gives the correct response to item (i.e., CT scan) $i$.
-
-In such a setting, we model this probability as
-$$
-P_{si} = \text{inverse\_logit}(\eta_{si})
-$$
-with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
-$$
-\text{logit}(P_{si}) = \eta_{si}
-$$
-with the logit link $g(P_{si}) = \text{logit}(P_{si}) = \text{ln} (\frac{P_{si}}{1 - P_{si}})$.
-
-In our case study, the probability of making an accurate diagnostic decision is assumed to depend on the predictors:
-
-- $advice\_present_{si}$: whether subject $s$ was presented with AI advice (1) or not (0) when asked to assess item $i$ 
-- $advice\_correct_{si}$: whether this advice was correct (1) or not (0) 
-- $expert_s$: whether subject $s$ was a task expert (1) or not (0)
-
-and the random effects:
-
--   $u_{0s}$: the deviation of subject $s$ from the average ability to solve an item (i.e., CT scan) with average difficulty; assumed to be distributed as $u_{0s} \sim N(0, \sigma_S^2)$
--   $u_{0i}$: the deviation of item (i.e., CT scan) $i$ from the average difficulty to be solved by a person with average ability; assumed to be distributed as $u_{0i} \sim N(0, \sigma_I^2)$
-
-In total, we assume the model
-$$
-\begin{aligned}
-\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
-&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
-\end{aligned}
-$$
-or equivalently
-$$
-\begin{aligned}
-P_{si} = \text{inverse\_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
-&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
-\end{aligned}
-$$
-with model parameters $\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$.
-
-In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
-To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.
-
-##### Simulation function in R
-
-The following R function simulates a full dataset structured according to the design of our case study.
-
-```{r}
-#| echo: false
-b_0 = 0.847
-b_e = 1.350
-b_a = -1.253
-b_c = 2.603
-b_ea = 0.790
-b_ec = -1.393
-```
+## Simulation functions
 
@@
 
-In the first four lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data.^[The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.]
-In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).
-Thus, the `add_random` command creates a fully-crossed `data.frame` with `n_subjects` $\times$ `n_items` rows.
-We add a between-subject effect with the `add_between` command, simulating that about $25\%$ of subjects are experts.
-The next two lines simulate that in $\frac{2}{3}$ of trials, subjects will be presented with AI advice, and if advice is presented, the advice will be correct in about $80\%$ of cases (the variable `advice_correct` is always 0 when no advice is presented).
-Next, we simulate one random effect for each subject (`u0s`) and for each item (`u0i`).
-As assumed by standard GLMMs, the `add_ranef` function draws the random effects from a normal distribution with a mean 0 and a standard deviation specified by the user.
-With all design variables done, we are ready to simulate our model equation outlined in the last section.
-The linear predictor variable `linpred` ($\eta$ in the GLMM model equations) combines the predictor variables, random effects, and model parameters as assumed by our model.
-We then transform the linear predictor with the inverse-link function to compute `y_prob`, the probability that the subject correctly solved the item (in R, the inverse-logit link is computed with `plogis` and the logit link with `qlogis`).
-In the final step, we simulate the binary dependent variable `y_bin` (i.e., whether the subject makes an accurate diagnostic decision for the CT scan) by -- for each trial -- drawing from a Bernoulli distribution with success probability `y_prob`.
-
-## Step 3: Specify the population parameters
-
-### THEORY
-
-In the absence of previous studies with the same design or pilot data, researchers require strategies on how to specify the population parameters used in their data simulation.
-Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
-In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available for GLMMs.
-Our strategies to specify population parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
-The basic idea of all strategies is how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
-
-Although we use frequentist model estimation in our tutorial, many strategies described in this chapter are inspired by research on how to monitor the plausibility of model assumptions in applied Bayesian statistics [@gelmanBayesianWorkflow2020].
-
-### PRACTICE
-
-When introducing the simulation function for our case study, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
-
-The starting point for all parameter values in our present case study were based on results from distantly related study designs in the literature.
-Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions seem plausible.
-
-We now outline our main strategy to determine plausible parameter values for the fixed effects ($\beta$ parameters):
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret in isolation because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation, that allows us to ignore the random effects for now, works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
-Then the model implied probability that such a person solves such an item accurately is given by:
-
-$$
-\begin{aligned}
-P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
-= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
-\end{aligned}
-$$
-In fact, we would only need the full equation if the subject is an expert and correct advice is presented.
-In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-For example, the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented, only requires the intercept:
-$$
-\begin{aligned}
-P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
-= \text{inverse\_logit}[\beta_0]
-\end{aligned}
-$$
-We can revert this perspective by choosing plausible probability values based on domain knowledge and deriving the parameter values implied by these probabilities for each experimental condition.
-
-```{r}
-#| label: tab-probtable
-#| echo: false
-probtable <- data.frame(
-  col1 = 
-    c("no advice, no expert", "no advice, expert", "false advice, no expert", 
-      "false advice, expert", "correct advice, no expert", "correct advice, expert"),
-  col2 =
-    c("0.70", "0.90", "0.40", "0.85", "0.90", "0.95"),
-  col3 = 
-    c("$logit(0.70) = \\beta_0$",
-      "$logit(0.90) = \\beta_0 + \\beta_e$", 
-      "$logit(0.40) = \\beta_0 + \\beta_a$",
-      "$logit(0.85) = \\beta_0 + \\beta_e + \\beta_{a} + \\beta_{ea}$",
-      "$logit(0.90) = \\beta_0 + \\beta_a + \\beta_c$",
-      "$logit(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
-apa_table(
-  probtable
-  , col.names = c("Experimental condition",
-    "$P(Y=1|\\mathbf{X=x}, \\mathbf{U} = \\mathbf{0})$",
-    "Implied equation")
-  , caption = "Assumed probabilities that an average subject solves an average item in each experimental condition."
-  , note = "Implied equations are derived based on the model equations and setting all random intercept terms to 0."
-  , escape = FALSE
-)
-```
-
-@tab-probtable shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
-The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations (`b_0` stands for the intercept $\beta_0$, `b_e` for the slope $\beta_e$, and so on):
-
-```{r}
-#| echo: true
-b_0 <- qlogis(0.7)
-b_e <- qlogis(0.9) - b_0
-b_a <- qlogis(0.4) - b_0
-b_ea <- qlogis(0.85) - b_0 - b_e - b_a
-b_c <- qlogis(0.9) - b_0 - b_a
-b_ec <- qlogis(0.95) - b_0 - b_e - b_a - b_c - b_ea
-c(b_0 = b_0, b_e = b_e, b_a = b_a, b_c = b_c, b_ea = b_ea, b_ec = b_ec)
-```
-
-It is always possible to double-check these computations by transforming the parameter values back to probabilities, e.g. 
-$$
-\begin{aligned}
-P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
-= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
-\end{aligned}
-$$
-
-which we compute in R as:
-
-```{r}
-#| echo: true
-plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec)
-```
-
-This leaves us with the question on how to determine plausible values for the two remaining model parameters ($\sigma_S$, and $\sigma_I$) that are the standard deviations for the random intercepts.
-For this, we introduce two more strategies.
-
-### Insightful descriptive statistics
-
-#### THEORY
-
-The mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
-The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
-For example, domain experts might not be able to directly choose plausible values for the  coefficients in a logistic regression model (which are measured on the log-odds scale).
-However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental conditions, i.e.
-which relative frequency they expect to observe (on average).
-The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
-Although insightful descriptive statistics usually depend on the model parameters in a non-linear way, it is not necessary to solve the exact relationship mathematically.
-Instead, one can simply adjust the population parameters by trial and error until the model implied quantities produce the desired result.
-
-#### PRACTICE
-
-In the last section, we showed how we can derive the model implied probability that a subject with average ability solves an item with average difficulty for each experimental condition.
-Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty [@fahrmeirRegressionModelsMethods2021].
-The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
-
-For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice (the conditional perspective).
-However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability, and items drawn from their natural distribution of difficulty (the marginal perspective).
-How much the two probabilities differ depends on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations would be zero).
-We want to use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
-
-We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model implied true values) and simply compute the relative frequency of correct diagnoses for each experimental condition.
-
-```{r}
-#| echo: true
-library(tidyverse)
-set.seed(1)
-dat <- simulate(n_subjects = 3000, n_items = 3000,
-  sd_u0s = 0.5, sd_u0i = 0.5)
-dat |> 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
-  group_by(condition) |>
-  summarize(relative_frequency = sum(y_bin) / n())
-```
-
-We tried using these descriptive statistics to judge together with domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
-However, although the result was deemed plausible, these statistics were not informative enough to determine a final set of plausible parameter values.
-For this reason, we will additionally look at insightful model based quantities.
-
-### Insightful model based quantities
-
-#### THEORY
-
-Because GLMMs are complicated models, descriptive statistics alone are usually not enough to specify plausible values for all model parameters.
-This is especially true for the standard deviation of random effects that have non-linear (and often unexpected) effects on the model-implied results.
-An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
-For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
-Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out implausible boundary conditions.
-For example, the domain expert might deem it implausible that the 5% most able participants have a probability of more than 0.99 to solve the difficult cognitive task.
-
-#### PRACTICE
-
-The discussed inequality of conditional and marginal effects in GLMMs [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult.
-One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often radiologists make an accurate diagnosis based on brain scans).
-However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge.
-Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
-
-In the last section, we simulated a large dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
-We will now use the model implied probability of each simulated data point (stored in the variable `y_prob`) to visualize the whole model implied marginal distribution of correct diagnoses for each experimental condition.
-
-```{r}
-#| label: fig-margdist1
-#| fig-cap: "Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions."
-#| echo: true
-library(ggdist)
-dat |> 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
-@fig-margdist1 shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
-We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
-This difference increases with the variability of the random effects.
-
-We can use plots like the one above as a useful tool to decide whether the specified standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
-
-In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$). This gives us a better way to see the variability between persons.
-
-```{r}
-#| label: fig-margdist2
-#| echo: false
-#| fig-cap: "Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of item random intercepts to 0.01."
-set.seed(1)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0i = 0.01)
-dat |> 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
-As an example, @fig-margdist2 reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
-In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
-Patterns like these should be considered carefully and discussed with the domain experts.
-Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
-We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
-
-```{r}
-#| include: false
-set.seed(1)
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
-dat |> 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
-The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
-For @fig-margdist3 we have set $\sigma_S = 3$ and $\sigma_I = 3$.
-This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
-These high standard deviations do not account for the inherent variability and complexity of human performance.
-For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
-
-```{r}
-#| label: fig-margdist3
-#| echo: false
-#| fig-cap: "Marginal distributions including means, 66% and 95% confidence intervals for all experimental conditions while setting the standard deviation of subject and item random intercepts to 3."
-set.seed(1)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 3, sd_u0i = 3)
-dat |> 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
-  mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
-### Iterative process with domain experts
-
-#### THEORY
-
-In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
-In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
-As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
-For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst who is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
-Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
-Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
-During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
-
-##### PRACTICE
-
-All parameter values in our present case study have been determined based on repeated discussions with our affiliated domain experts in Radiology to validate our assumptions.
-Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for intracranial hemorrhage.
-Existing studies indicate that radiologists typically demonstrate high accuracies, often exceeding or hovering around 90%, while interns have been shown to perform below 80%, and medical students fall even shorter.
-For simplicity, we assumed plausible probability values of .90 for experts and .70 for non-experts, respectively.
-Our experts confirmed that these values are realistic baselines for reviewing diverse head CT images without AI assistance.
-Subsequently, we consulted several published papers investigating the effect of correct and incorrect advice on decision-making performance in other settings.
-From their findings, we inferred that both experts and non-experts should benefit from correct and suffer losses from incorrect advice.
-However, the magnitude of these effects should be substantially greater for non-experts, given their demonstrated reliance on advice compared to experts.
-We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
-For our simulation, we used the probabilities of average participants to solve an average case, as shown in Table 1.
-
-## Step 4: Estimate the statistical model
-
-### THEORY
-
-At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset that will later be collected in the planned study.
-The next step is to specify how the statistical model shall be estimated in the actual study collected later.
-This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
-Note that this does not always mean that one will specify the same GLMM that was used when specifying the data generating process.
-On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
-On the other hand, using a more complex model for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
-
-### PRACTICE
-
-<!-- The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
--->
-
-<!-- [^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017]. -->
-
-<!-- However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore, for our case study, we have written our own function to simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions).
--->
-<!-- We think this choice assists applied researchers in better understanding all model assumptions before using lme4 to analyze the simulated datasets.[^2] -->
-
-<!-- [^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
--->
-
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
-For the current example, we simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
-However, for the sake of the exercise, we can imagine that this would be real data resulting from our future experiment and think about how we would analyze this data.
-
-```{r}
-#| message: false
-#| echo: true
-library(tidyverse)
-set.seed(1)
-dat <- simulate(n_subjects = 100, n_items = 50)
-```
-
-The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:
-
-```{r}
-#| message: false
-#| echo: true
-library(lme4)
-f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
-  expert:advice_present + expert:advice_correct +
-  (1|subject) + (1|item)
-```
-
-The first two lines look similar to any linear model in R (general intercept indicated by `1`; main effects indicated by variable names in the dataset; interactions indicated by `variable1:variable2`).
-The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
-The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
-
-In lme4, a GLMM is fitted with the `glmer` function.
-By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs.
-inaccurate (0) diagnosis.
-
-```{r}
-#| echo: true
-fit <- glmer(f, data = dat, family = "binomial")
-```
-
-We can inspect the estimates for all model parameters with the `summary` command:
-
-```{r}
-#| echo: true
-summary(fit)
-```
-
-In the lme4 output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
-
-## Step 5: Compute the estimate
-
-### THEORY
-
-In previous steps, we have defined the theoretical estimand, written a data simulation function and specified how to estimate a GLMM using simulated data.
-The next step is to specify how to compute a concrete point estimate of the theoretical estimand within the framework of the fitted GLMM.
-For some research question, the estimate corresponds with a single regression coefficient. In more complicated scenarios, the estimate is computed from a combination of coefficients.
-Beyond computing the point estimate, we have already discussed that both hypothesis testing and interval estimation can be used to answer the research question.
-The decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with the marginaleffects R package using the delta method).
-
-### PRACTICE
-
-In the estimand section, we have translated a verbal description of our research question into four probability statements that are specified outside of any specific statistical model.
-For a concrete estimate within the context of our specified GLMM, we must compute the following probability contrasts:
-
-$$
-\begin{aligned}
-& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
-& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
-\end{aligned}
-$$
-$$
-\begin{aligned}
-& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
-& \quad - P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
-\end{aligned}
-$$
-
-$$
-\begin{aligned}
-& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
-& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
-\end{aligned}
-$$
-
-$$
-\begin{aligned}
-& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
-& \quad - P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
-\end{aligned}
-$$
-We have already discussed how to compute the involved probabilities in the section on specifying population parameters.
-Plugging in the model equation of the GLMM produces an equation on how to compute each contrast if all model parameters were known.
-When we want to estimate the above contrasts based on *observed* data, the only difference is that model parameters are not known and we instead use the corresponding parameter *estimates*.
-
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- & \text{omv}[\beta_0 + \beta_a \cdot 1 + \beta_c \cdot 1 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 1 + \beta_{ec} \cdot 1 \cdot 1]\\ -->
-<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_a \cdot 0 + \beta_c \cdot 0 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 0 + \beta_{ec} \cdot 1 \cdot 0] = \\ -->
-<!-- & = \text{inverse_logit}[\beta_0 + \beta_a  + \beta_c + \beta_e + \beta_{ea} + \beta_{ec} ]\\ -->
-<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_e ] \\ -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-
-We could use our knowledge of the structure of the GLMM to determine the exact formula needed to compute the contrasts of interest and then plug in the parameter estimates manually from the `summary(fit)` output.
-However, this would be tedious and we can use R to compute this contrast without doing the math.
-
-Using the first contrast as our example, we could apply the `predict` function of the lme4 package to compute the predicted probability for a correct diagnosis based on our fitted model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
-
-```{r}
-#| echo: true
-pred_df <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), 
-  expert = c(1, 1))
-pred_df
-preds <- predict(fit, newdata = pred_df, type = "response", re.form = NA)
-preds
-preds[1] - preds[2]
-```
-
-The argument `type = "response"` specifies that predictions are made on the probability scale (instead of the log-odds scale of the $\beta$ parameters), while `re.form = NA` sets all random effects to 0.
-We could use this method to compute point estimates for all four contrasts that are part of our estimand.
-However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute HTs or CI.
-
-The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute HTs and CIs for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
-First, we specify a grid of all combinations of predictor variable and then compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
-
-```{r}
-#| echo: true
-library(tidyverse)
-library(marginaleffects)
-pred_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
-pred_df
-probs <- predictions(fit, newdata = pred_df, type = "response", re.form = NA)
-probs
-```
-
-The point estimates for all experimental conditions are reported in the `Estimate` column.
-Note that the output also contains the two missing by design conditions that will never be observed in the actual study ($advice\_present = 0, advice\_correct = 1, expert = 1$ and $advice\_present = 0, advice\_correct = 1, expert = 0$).
-This is no problem as long as we never interpret those estimates.
-Next, we use the estimated probabilities to compute the four specific contrasts that are part of our estimand.
-For this we must specify which rows in `probs` have to be subtracted from each other.
-We will use the `hypotheses` function to compute our four contrasts of interest together with HTs and CIs.
-We use the default inference options of the marginaleffects package that compute HTs and CIs based on the approximate delta method.
-
-```{r}
-#| echo: true
-probs |> 
-  hypotheses(hypothesis = c(
-    "b8 = b2",
-    "b2 = b6",
-    "b7 = b1",
-    "b1 = b5"),
-    equivalence = c(0, 0))
-```
-
-The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2 in the `probs`-output.
-The argument `equivalence = c(0, 0)` can be used to compute one-sided p-values, testing whether the contrast in the population is smaller than 0 (`p (NonSub)` column) or greater than 0 (`p (NonInf)` column).
-The point estimates for four contrasts are reported in the `Estimate` column.
-Note that to facilitate interpretation, we arranged the contrasts in a way that we theoretically expect positive values for all four of them.
-
-#### Hypothesis testing
-
-If we chose hypothesis testing for our case study, we would test a combined null hypothesis $H_0$ that consists of four separate null hypotheses:
-
-$$
-\begin{aligned}
-H_{01}:\ & P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
-& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
-H_{02}:\ &P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
-& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
-H_{03}:\ &P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
-& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
-H_{04}:\ & P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
-& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
-\end{aligned}
-$$
-The combined null hypothesis $H_0$ should only be rejected if **all** individual null hypotheses are rejected [i.e., intersection-union setting\; @dmitrienkoTraditionalMultiplicityAdjustment2013].
-In such cases, the error probabilities do not accumulate, and we would waste power when correcting for multiple tests.
-
-With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value in the `p (NonInf)` column for the third hypothesis is not significant) and therefore also not reject the combined null hypothesis for this particular (simulated) dataset.
-Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis $H_1$ is actually true in the population.
-
-#### Interval estimation
-
-If we chose parameter estimation for our case study, we would focus on the two-sided CIs of the four contrasts of interest.
-With a standard confidence level of $1 - \alpha = 0.95$, plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
-Note that due to the constrained range of the probability scale, the width of the CI differs between the four contrasts (which is the expected behavior for binomial GLMMs).
-The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice) where both underlying probabilities are close to 1.
-The largest width is observed for the fourth contrast (non-expert with wrong advice vs. non-expert without advice), where both underlying probabilities are closer to 0.5.
-
-<!-- Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
--->
-
-<!-- With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects): -->
-
-<!-- (ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package. -->
-
-<!-- ```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"} -->
-<!-- library(marginaleffects) -->
-<!-- plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"), -->
-<!--   type = "response") + ylim(c(0.3, 1)) -->
-<!-- ``` -->
-
-## Step 6: Perform repeated simulations
-
-### THEORY
-
-Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute HTs or CIs for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
-The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
-On a conceptual level, we first require a function that takes as input the sample size and the full set of population parameter values.
-When planning for power, the function should return the p-value(s) of the HT(s) of interest when conducted on the simulated dataset.
-When planning for precision, the function should return the width of the CI(s) of interest.
-Secondly, we must run this function repeatedly with the same sample size and population parameters.
-Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computing, that is running simulations on multiple cores of the computer at the same time to reduce total run time.
-Thirdly, the results of the repeated simulation must be collected and aggregated.
-When planning for power, we compute the relative frequency of (a) significant p-value(s) across repeated simulations.
-When planning for precision, we compute the average width of the CI(s).
-Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted power or precision.
-
-### PRACTICE
-
-We are finally ready to run our simulation-based sample size planning analyses to plan for power and for precision.
-Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and CIs) in a single command.
-
 ```{r}
@@
 
-Simulation-based sample size planning can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
-Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
-First, we enable parallelization with the `plan` function and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
+## Run simulation
 
 ```{r}
-#| message: false
-#| eval: false
 #| echo: true
+library(tidyverse)
 library(future)
+library(furrr)
+
 plan("multisession", workers = 6)
 set.seed(2)
-```
 
-The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
-We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
-With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
-If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.
-
-```{r}
-#| echo: false
-#| eval: false
-library(future)
-library(furrr)
-plan("multisession", workers = 6)
-set.seed(2)
-sim_result <- crossing(
-  rep = 1:300,
-  n_subjects = c(100, 150, 200, 250),
-  n_items = c(10, 30, 50, 70)
-) |>
-  mutate(res = future_pmap(., sim_and_analyse, 
-    .options = furrr_options(seed = TRUE))) |>
-  unnest(col = res)
-saveRDS(sim_result, file = "results.rds")
-```
-
-```{r}
-#| message: false
-#| eval: false
-#| echo: true
-library(furrr)
 sim_result <- crossing(
@@
 
-The result of this computation is a data frame that contains the p-values and CIs of all specified contrasts for each simulated dataset.
-In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
-When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process.
-We do not remove these results because non-convergence can also happen when analyzing the real data we plan to collect, thus, we want to factor in this possibility to keep our simulation more realistic.
-
-#### Power results
-
-For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.[^3]
-
-[^3]: This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
+## Save results file to be used in the manuscript
 
 ```{r}
-#| echo: false
-sim_result <- readRDS(file = "results.rds")
+saveRDS(sim_result, file = "results.rds")
 ```
-
-```{r}
-#| label: fig-finalpwr
-#| fig-cap: "Simulation-based power estimates including 95% confidence interval of the case study for different numbers of subjects and items, based on a significance level of 0.05."
-#| echo: true
-library(binom)
-alpha <- 0.05
-power <- sim_result |>
-  pivot_wider(names_from = term, names_sep = "_", 
-    values_from = estimate:p.value.equiv) |>
-  group_by(n_subjects, n_items) |> 
-  summarise(
-    power = mean(`p.value.noninf_b1=b5` < alpha & 
-        `p.value.noninf_b8=b2` < alpha & `p.value.noninf_b2=b6` < alpha & 
-        `p.value.noninf_b7=b1` < alpha), 
-    n_sig = sum(`p.value.noninf_b1=b5` < alpha & 
-        `p.value.noninf_b8=b2` < alpha & `p.value.noninf_b2=b6` < alpha & 
-        `p.value.noninf_b7=b1` < alpha),
-    n = n(),
-    ci.lwr = binom.confint(n_sig, n, method = "wilson")$lower,
-    ci.upr = binom.confint(n_sig, n, method = "wilson")$upper, 
-    .groups = "drop")
-power |>
-  mutate(across(c(n_subjects, n_items), factor)) |>
-  ggplot(aes(n_subjects, n_items, fill = power)) +
-  geom_tile() +
-  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
-                                power, ci.lwr, ci.upr)), 
-    color = "white", size = 4) +
-  scale_fill_viridis_c(limits = c(0, 1)) +
-  xlab("number of subjects") + ylab("number of items")
-```
-
-<!-- How many participants/stimuli are needed for concrete example -->
-
-As should be the case, power estimates in @fig-finalpwr increase with both the number of subjects and the number of items.
-The CIs reported here indicate how precisely power was estimated by our simulation.
-Higher precision (which would be reflected in narrower CIs) could be obtained by increasing the number of repetitions (`rep`) in the simulation.
-In practice, data simulations are often run multiple times with adjusted combinations of sample sizes.
-When running for the first time, it might be revealed that power is way too low (or much higher than required) for some combinations of `n_subjects` and `n_items`.
-When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
-This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.
-
-Much has been written on the optimal amount of power to target in empirical research.
-The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints  [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
-When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
-For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation.
-This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice).
-If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items.
-However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule.
-The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
-Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
-
-#### Precision results
-
-When planning for precision, one could monitor the width of all four CIs at the same time.
-However, because the CIs of the four contrasts strongly differ in width, it is not trivial to decide which width one should target when deciding on the appropriate sample size.
-In contrast to planning for power, there are no common standards on how to specify the targeted precision.
-For our example, we use a simple heuristic but we strongly encourage readers to think about better alternatives that are appropriate in their own applications.
-
-Our simulations show that the smallest CI can be expected for the first contrast (expert with correct advice vs. expert without advice).
-The true contrast in probability for an average expert and an average item in this condition is `plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec) - plogis(b_0 + b_e) =` $`r round(plogis(b_0 + b_e + b_a + b_c + b_ea + b_ec) - plogis(b_0 + b_e), 3)`$.
-We want the width of this CI to be smaller than 0.1.
-This would mean that if the point estimate happens to be close to the true value, the plausible values inside of a 95% CI would all be positive.
-
-Thus in our example, precision is defined as the (long-run) average width of a 95% CI for the probability contrast between experts with correct advice and experts without advice.
-Of course, lower width implies better precision.
-Based on our simulation outcomes, we compute the precision estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.
-
-```{r}
-#| label: fig-finalprecision
-#| fig-cap: "Simulation-based precision estimates (expected width of confidence intervals) including 95% confidence interval of the case study for different numbers of subjects and items, based on a confidence level of 0.95."
-#| echo: true
-precision <- sim_result |>
-  pivot_wider(names_from = term, names_sep = "_", 
-    values_from = estimate:p.value.equiv) |>
-  group_by(n_subjects, n_items) |>
-  mutate(width = `conf.high_b8=b2` - `conf.low_b8=b2`) |>
-  summarise(precision = mean(width),
-    ci.lwr = t.test(width)$conf.int[1],
-    ci.upr = t.test(width)$conf.int[2], 
-    .groups = "drop")
-precision |>
-  mutate(across(c(n_subjects, n_items), factor)) |>
-  ggplot(aes(n_subjects, n_items, fill = precision)) +
-  geom_tile() +
-  geom_text(aes(label = sprintf("%.2f \n [%.2f; %.2f]", 
-                                precision, ci.lwr, ci.upr)), 
-    color = "white", size = 4) +
-  scale_fill_viridis_c(limits = c(0, 0.3), direction = -1) +
-  guides(fill = guide_legend(reverse=FALSE))
-```
-
-As should be the case, precision estimates in @fig-finalprecision increase  (i.e., average width of CI decreases) with both the number of subjects and the number of items.
-The CIs reported here indicate how precisely the expected width of the CI for our focal contrast was estimated by our simulation.
-Applying our simple heuristic of targeting an expected width smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.
-We could either choose 100 subjects and 30 items or 200 subjects and 10 items.
-
-Note that our simple heuristic for determining sample size in the planning for precision scenario was quite liberal.
-This is reflected by the result that we would need a smaller sample size than in the planning for power scenario.
-With a more conservative precision target, the result is generally the opposite: As a rule, precise parameter estimates usually require bigger samples than null hypothesis testing.
-
-<!-- ## Sensitivity analysis -->
-
-## Sensitivity analysis
-
-In our case study, we have performed simulation-based sample size planning from a single set of parameter values that reflect our assumptions of an expected effect size.
-Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials [e.g., @kumleEstimatingPowerGeneralized2021], we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
-Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
-When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present (or estimate the effect with satisfying precision).
-However, if the true effect is indeed smaller, power (or precision) will be lower, and the study might not be sufficiently informative.
-A common, more conservative strategy for sample size justification is to perform sample size planning for the smallest effect size of interest (SESOI).
-An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011; @lakensEquivalenceTestingPsychological2018].
-For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].
-When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power or precision for different effect sizes in what is called *sensitivity power analysis*.
-By simulating power (or precision) for different effect sizes (in addition to the different number of subjects and items), one can make sure that power (or precision) would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power (or precision) depends on the size of the true effect.
-For our case study that investigates combined hypotheses in a GLMM modeling framework, sensitivity analysis would require manually specifying additional sets of plausible parameter values that reflect scenarios with smaller or larger differences between groups with respect to our specific research question.
-Power (or precision) could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
-
-# Conclusion and outlook
-
-The goal of this tutorial was to teach researchers how to perform tailored simulation-based sample size planning for GLMMs.
-Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based sample size planning in experimental research:
-
-In light of the ongoing replication crisis and an existing literature full of underpowered studies, there is a growing need for simulation-based sample size planning in experimental research: In order to conduct informative research, GLMMs offer a flexible statistical framework to analyze complex experimental study designs.
-However, existing formula-based heuristics and user-friendly software tools for a priori power analysis are often not sufficient.
-Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
-
-Managing data simulations more easily with discrete predictor variables: Simulation-based sample size planning becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
-This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable estimates for power or precision without compromising realistic assumptions about the data-generating process implied by the study design.
-
-Teaching data simulation skills: The ability to conduct simulation-based sample size planning is a valuable skill that should be taught to experimental researchers.
-By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power or precision, and improve the quality of their experimental designs.
-Equipping researchers with the knowledge and tools to perform simulation-based sample size planning enables them to make informed decisions and enhance the rigor of their studies. 
-The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
-
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform tailored simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
-Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-It is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.
-
-Recognizing the value of simulation-based design analysis: Tailored data simulations and power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
-They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power or precision.
-Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
-
-Integration with Open Science and preregistration practices: Simulation-based sample size planning aligns well with the principles of Open Science and preregistration in experimental research.
-When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.
-By preregistering their simulation-based sample size plan, researchers enhance the transparency and accountability of their experimental procedures, contributing to the credibility and reproducibility of research.
-
-# References
-
-::: {#refs}
-:::

### 2024-06-11 manuscript.qmd
not ideal fix of probtable...

@@
   - name: Florian Pargent
-    corresponding: true
     equal-contributor: true
@@
 
-In Table XXX, we give a short review of existing R packages that can be used for power analysis for GLMMs.
+INSERT TABLE 1 HERE!
+
+In Table 1, we give a short review of existing R packages that can be used for power analysis for GLMMs.
 
@@
 ```{r}
-#| label: tab-probtable
+#| label: tbl-probtable
+#| tbl-cap: Assumed probabilities that an average subject solves an average item in each experimental condition.
+#| apa-note: Implied equations are derived based on the model equations and setting all random intercept terms to 0.
 #| echo: false
@@
       "$logit(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
-apa_table(
-  probtable
-  , col.names = c("Experimental condition",
-    "$P(Y=1|\\mathbf{X=x}, \\mathbf{U} = \\mathbf{0})$",
-    "Implied equation")
-  , caption = "Assumed probabilities that an average subject solves an average item in each experimental condition."
-  , note = "Implied equations are derived based on the model equations and setting all random intercept terms to 0."
-  , escape = FALSE
-)
-```
-
-@tab-probtable shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
+knitr::kable(probtable, col.names = c("Experimental condition",
+    "$P(Y=1 \\mid \\mathbf{X=x}, \\mathbf{U} = \\mathbf{0})$",
+    "Implied equation"), escape = FALSE)
+```
+
+@tbl-probtable shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
 The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations (`b_0` stands for the intercept $\beta_0$, `b_e` for the slope $\beta_e$, and so on):
@@
   ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45",
+    breaks = "Sturges") +
   scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
@@
 set.seed(1)
-dat <- simulate(n_subjects = 2000, n_items = 2000, sd_u0s = 0.01)
+dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.01)
 dat |> 
@@
 ```{r}
-#| message: false
 #| eval: false

### 2024-06-12 full_simulation.qmd
update rcode.qmd and full_simulation.qmd

@@
 ---
-title: "Full simulation"
+title: "Run the full simulation for the case study example"
+execute: 
+  eval: false
 ---
@@
 ::: {.callout-note}
-This Notebook is used to run the full simulation for the case study discussed in the manuscript. The simulation takes several hours to run on a Mac Book Pro (13-inch, M1, 2020) with 16 GB memory.
+This Notebook was used to run the full simulation for the case study discussed in the manuscript. The simulation takes more than 1 hour to run on a Mac Book Pro (13-inch, M1, 2020) with 16 GB memory.
 :::

### 2024-06-12 rcode.qmd
update rcode.qmd and full_simulation.qmd

@@
 ---
-title: "R code from the practice sessions"
+title: "R code from PRACTICE sections"
+execute:
+  eval: false
 ---
@@
 ::: {.callout-note}
-This Notebook contains only the R code from the practice sessions of the manuscript.
+This Notebook contains only the R code from the PRACTICE sections of the manuscript.
 The number of samples and repetitions for all simulations have been greatly reduced to ensure that the code runs quickly and with limited computational resources.
@@
  "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
-  group_by(condition) %>%
+ "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
+  group_by(condition) |>
   summarize(relative_frequency = sum(y_bin) / n())
@@
 library(ggdist)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
  "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+ "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
   ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45",
+    breaks = "Sturges") +
   scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
@@
 dat <- simulate(n_subjects = 500, n_items = 500, sd_u0i = 0.01)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
     "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
   ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45",
+    breaks = "Sturges") +
   scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
@@
 dat <- simulate(n_subjects = 500, n_items = 500, sd_u0s = 0.01)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
 dat <- simulate(n_subjects = 500, n_items = 500, sd_u0s = 3, sd_u0i = 3)
-dat %>% 
+dat |> 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
+    factor(expert), factor(advice_present), factor(advice_correct))) |>
   mutate(condition = fct_recode(condition,
@@
 library(tidyverse)
+library(lme4)
 set.seed(1)
 dat <- simulate(n_subjects = 100, n_items = 50)
-```
-
-```{r}
-#| message: false
-library(lme4)
 f <- y_bin ~ 1 + expert + advice_present + advice_correct + 
@@
   (1|subject) + (1|item)
-```
-
-```{r}
 fit <- glmer(f, data = dat, family = "binomial")
-```
-
-```{r}
 summary(fit)
@@
 ```{r}
-pred_df <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), 
+grid1 <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), 
   expert = c(1, 1))
-pred_df
-preds <- predict(fit, newdata = pred_df, type = "response", re.form = NA)
-preds
-preds[1] - preds[2]
+grid1
+pred <- predict(fit, newdata = grid1, type = "response", re.form = NA)
+pred
+pred[1] - pred[2]
 ```
@@
 ```{r}
-library(tidyverse)
-library(marginaleffects)
-pred_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
-pred_df
-probs <- predictions(fit, newdata = pred_df, type = "response", re.form = NA)
-probs
+grid2 <- expand_grid(advice_present = 0:1, 
+  advice_correct = 0:1, expert = 0:1)
+grid2
+preds <- predictions(fit, newdata = grid2, 
+  type = "response", re.form = NA)
+as.data.frame(preds)[,c(2,7:11)]
 ```
@@
 ```{r}
-probs %>% 
+contrasts <- preds |> 
   hypotheses(hypothesis = c(
-    "b8 = b2",
-    "b2 = b6",
-    "b7 = b1",
-    "b1 = b5"),
+    "b8 = b2",  # (correct advice, expert) - (no advice, expert)
+    "b2 = b6",  # (no advice, expert) - (wrong advice, expert) 
+    "b7 = b1",  # (correct advice, no expert) - (no advice, no expert)
+    "b1 = b5"), # (no advice, no expert) - (wrong advice, no expert)
     equivalence = c(0, 0))
+contrasts
 ```
@@
 alpha <- 0.05
-power <- sim_result %>%
+power <- sim_result |>
   pivot_wider(names_from = term, names_sep = "_", 
-    values_from = estimate:p.value.equiv) %>%
-  group_by(n_subjects, n_items) %>% 
+    values_from = estimate:p.value.equiv) |>
+  group_by(n_subjects, n_items) |> 
   summarise(
@@
 #| fig-cap: "Simulation-based precision estimates (expected width of confidence intervals) including 95% confidence interval of the case study for different numbers of subjects and items, based on a confidence level of 0.95."
-precision <- sim_result %>%
+precision <- sim_result |>
   pivot_wider(names_from = term, names_sep = "_", 
-    values_from = estimate:p.value.equiv) %>%
-  group_by(n_subjects, n_items) %>%
-  mutate(width = `conf.high_b8=b2` - `conf.low_b8=b2`) %>%
+    values_from = estimate:p.value.equiv) |>
+  group_by(n_subjects, n_items) |>
+  mutate(width = `conf.high_b8=b2` - `conf.low_b8=b2`) |>
   summarise(precision = mean(width),

### 2024-06-12 manuscript.qmd
various formatting improvements to manuscript.qmd

@@
   disclosures:
-    data-sharing: "All materials (reproducible manuscript incl. R code, simulation results) are available in the project's repository on the Open Science Framework (OSF) at https://osf.io/dhwf4/"
-    conflict-of-interest: "The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article."
-    financial-support: "This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525)."
+    data-sharing: This is version 2.5 of our preprint published at <https://doi.org/10.31234/osf.io/rpjem>.
+All materials (reproducible manuscript, including R code and all simulation results) are available in the project's repository on the Open Science Framework (OSF) at <https://osf.io/dhwf4/>.
+A Quarto Manuscripts website is hosted at <https://Timo-Ko.github.io/glmm_simulation_tutorial/>.
+    conflict-of-interest: The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
+    financial-support: This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525).
     gratitude: ~
-    authorship-agreements: "Florian Pargent and Timo K. Koch contributed equally to this work."
-abstract: "When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
+    authorship-agreements: Florian Pargent and Timo K. Koch contributed equally to this work.
+abstract: When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
 The recent replication crisis underlines the need for more rigorous statistical methodology and adequately powered designs.
@@
 We not only focus on power analysis but also explain how to use the precision of parameter estimates to determine appropriate sample sizes.
-We conclude with an outlook on the increasing importance of simulation-based sample size planning."
+We conclude with an outlook on the increasing importance of simulation-based sample size planning.
 keywords: [power analysis, precision, data simulation, sample size planning, generalized linear mixed model]
@@
 #| include: false
-library("papaja")
 library("tidyverse")
@@
 library("marginaleffects")
+library("tinytable")
 library("future")
@@
 library("binom")
+library("papaja")
 r_refs("r-references.bib")
@@
 
-Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in TABLE XXX.
+Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in TABLE 1.
 The most important circumstances under which tailored simulation-based sample size planning is necessary are 1) complex study designs, 2) complex statistical hypotheses, 3) planning for precision, 4) no available prior studies or pilot data.
@@
 Although planning for precision [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008; @cummingNewStatisticsWhy2014] has been increasingly discussed as a useful strategy for empirical research, all available software packages for sample size planning with GLMMs are based on power analysis.
-Therefore, researchers that want to want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE XXX.
+Therefore, researchers that want to want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE 1.
 <!-- As long as one is interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
@@
 We will introduce each step in a theoretical section, followed by the practical application based on a hypothetical case study.
-All code in this manuscript and simulation results are available in the project’s repository on the Open Science Framework (https://osf.io/dhwf4/).
+All code in this manuscript and simulation results are available in the project’s repository on the Open Science Framework (<https://osf.io/dhwf4/>).
 
@@
 However, this is not possible for all estimands, which is why the literature discusses many estimation strategies beyond regression [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
-
-<!-- When the empirical estimand has been specified, the last step is to define how a concrete estimate of the empirical estimand shall be computed based on the observed study data.
-When a regression model is used for estimation, this includes decisions like which statistical framework estimation technique and software package are used to fit the model and how to compute HTs or CIs.
--->
 
@@
 This contrast is quantified for a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.^[Note that a different estimand would be the so-called average treatment effect (ATE).
-For the ATE, the probability contrast is defined for each combination of expert and scan, and then these contrasts are averaged across all experts and scans from the target population [@lundbergWhatYourEstimand2021]].
+For the ATE, the probability contrast is defined for each combination of expert and scan, and then these contrasts are averaged across all experts and scans from the target population [@lundbergWhatYourEstimand2021].].
 To complete our definition of the estimand, we have to define our target population that consists of persons, stimuli, and AI-advice: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
@@
 Then, we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
-
 In the last section, we simulated a large dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
@@
 ```
-
 @fig-margdist1 shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
@@
 This difference increases with the variability of the random effects.
-
 We can use plots like the one above as a useful tool to decide whether the specified standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
-
-In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$). This gives us a better way to see the variability between persons.
 
@@
 ```
-
+In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$).
+This gives us a better way to see the variability between persons.
 As an example, @fig-margdist2 reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
@@
 We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
-
 ```{r}
@@
 For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
-
 ```{r}
@@
 
-<!-- The lme4 package includes a function called `simulate` that allows researchers to simulate the dependent variable based on the same model formula used for model fitting, enabling simulation-based power analyses and other related analyses.
--->
-
-<!-- [^1]: For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017]. -->
-
-<!-- However, the model parameterization used by the lme4 package is quite technical, making it difficult for applied researchers to determine whether their specified population model (i.e., the theoretical model that describes the underlying data generation process for a specific population of interest) implies plausible associations in their simulated data.
-Therefore, for our case study, we have written our own function to simulate data for GLMMs from first principles (i.e., creating synthetic data step by step instead of using black box functions).
--->
-<!-- We think this choice assists applied researchers in better understanding all model assumptions before using lme4 to analyze the simulated datasets.[^2] -->
-
-<!-- [^2]: A less flexible alternative would be to use the simr package [@greenSIMRPackagePower2016], which can be used to both simulate data and perform power analysis for models supported by the lme4 package.
--->
-
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]
+In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.^[For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].]
 For the current example, we simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
@@
 
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- & \text{omv}[\beta_0 + \beta_a \cdot 1 + \beta_c \cdot 1 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 1 + \beta_{ec} \cdot 1 \cdot 1]\\ -->
-<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_a \cdot 0 + \beta_c \cdot 0 + \beta_e \cdot 1 + \beta_{ea} \cdot 1 \cdot 0 + \beta_{ec} \cdot 1 \cdot 0] = \\ -->
-<!-- & = \text{inverse_logit}[\beta_0 + \beta_a  + \beta_c + \beta_e + \beta_{ea} + \beta_{ec} ]\\ -->
-<!-- & \quad - \text{inverse_logit}[\beta_0 + \beta_e ] \\ -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-
 We could use our knowledge of the structure of the GLMM to determine the exact formula needed to compute the contrasts of interest and then plug in the parameter estimates manually from the `summary(fit)` output.
 However, this would be tedious and we can use R to compute this contrast without doing the math.
-
-Using the first contrast as our example, we could apply the `predict` function of the lme4 package to compute the predicted probability for a correct diagnosis based on our fitted model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
-
-```{r}
-#| echo: true
-pred_df <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), 
+Using the first contrast *(correct advice, expert) - (no advice, expert)* as our example, we could apply the `predict` function of the lme4 package to compute the predicted probability for a correct diagnosis based on our fitted model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
+
+```{r}
+#| echo: true
+grid1 <- data.frame(advice_present = c(1, 0), advice_correct = c(1, 0), 
   expert = c(1, 1))
-pred_df
-preds <- predict(fit, newdata = pred_df, type = "response", re.form = NA)
-preds
-preds[1] - preds[2]
+grid1
+pred <- predict(fit, newdata = grid1, type = "response", re.form = NA)
+pred
+pred[1] - pred[2]
 ```
@@
 However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute HTs or CI.
-
 The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute HTs and CIs for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
@@
 library(marginaleffects)
-pred_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
-pred_df
-probs <- predictions(fit, newdata = pred_df, type = "response", re.form = NA)
-probs
-```
+library(tinytable)
+grid2 <- expand_grid(advice_present = 0:1, 
+  advice_correct = 0:1, expert = 0:1)
+grid2
+preds <- predictions(fit, newdata = grid2, 
+  type = "response", re.form = NA)
+print(preds, style = "tinytable") |> theme_tt(theme = "resize")
+```
+
+<!-- print(probs, style = "tinytable") -->
 
@@
 #| echo: true
-probs |> 
+#| size: footnotesize
+contrasts <- preds |> 
   hypotheses(hypothesis = c(
-    "b8 = b2",
-    "b2 = b6",
-    "b7 = b1",
-    "b1 = b5"),
+    "b8 = b2",  # (correct advice, expert) - (no advice, expert)
+    "b2 = b6",  # (no advice, expert) - (wrong advice, expert) 
+    "b7 = b1",  # (correct advice, no expert) - (no advice, no expert)
+    "b1 = b5"), # (no advice, no expert) - (wrong advice, no expert)
     equivalence = c(0, 0))
+print(contrasts, style = "tinytable") |> theme_tt(theme = "resize")
 ```
@@
 If we chose hypothesis testing for our case study, we would test a combined null hypothesis $H_0$ that consists of four separate null hypotheses:
-
 $$
@@
 The largest width is observed for the fourth contrast (non-expert with wrong advice vs. non-expert without advice), where both underlying probabilities are closer to 0.5.
-
-<!-- Due to this complicated relationship, we argue not to focus too much on interpreting single model parameters when working with GLMMs.
-Instead, it can be more intuitive to consider model predictions and the model-implied distribution of the dependent variable for each experimental condition across all subjects and items.
--->
-
-<!-- With the marginaleffects package [@R-marginaleffects], we can easily compute predictions for all observations in the dataset based on the fitted GLMM (including all fixed **and** random effects), and plot the average probability with confidence intervals for each experimental condition in Figure \@ref(fig:marginaleffects): -->
-
-<!-- (ref:marginaleffects) Marginal distributions including means and 95% confidence intervals for all experimental conditions computed with the marginaleffects package. -->
-
-<!-- ```{r marginaleffects, message=FALSE, fig.cap="(ref:marginaleffects)"} -->
-<!-- library(marginaleffects) -->
-<!-- plot_predictions(fit, by = c("advice_present", "advice_correct", "expert"), -->
-<!--   type = "response") + ylim(c(0.3, 1)) -->
-<!-- ``` -->
 
@@
 ```{r}
-#| echo: false
-#| eval: false
-library(future)
-library(furrr)
-plan("multisession", workers = 6)
-set.seed(2)
-sim_result <- crossing(
-  rep = 1:300,
-  n_subjects = c(100, 150, 200, 250),
-  n_items = c(10, 30, 50, 70)
-) |>
-  mutate(res = future_pmap(., sim_and_analyse, 
-    .options = furrr_options(seed = TRUE))) |>
-  unnest(col = res)
-saveRDS(sim_result, file = "results.rds")
-```
-
-```{r}
 #| eval: false
@@
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.[^3]
-
-[^3]: This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.^[This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.]
 
@@
 #| echo: false
+# The simulation results were computed in the full_simulation.qmd notebook
 sim_result <- readRDS(file = "results.rds")
@@
 
-<!-- How many participants/stimuli are needed for concrete example -->
-
 As should be the case, power estimates in @fig-finalpwr increase with both the number of subjects and the number of items.
@@
 For our example, we use a simple heuristic but we strongly encourage readers to think about better alternatives that are appropriate in their own applications.
-
 Our simulations show that the smallest CI can be expected for the first contrast (expert with correct advice vs. expert without advice).
@@
 Based on our simulation outcomes, we compute the precision estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.
-
 ```{r}
@@
 ```
-
 As should be the case, precision estimates in @fig-finalprecision increase  (i.e., average width of CI decreases) with both the number of subjects and the number of items.
@@
 We could either choose 100 subjects and 30 items or 200 subjects and 10 items.
-
 Note that our simple heuristic for determining sample size in the planning for precision scenario was quite liberal.
@@
 With a more conservative precision target, the result is generally the opposite: As a rule, precise parameter estimates usually require bigger samples than null hypothesis testing.
-
-<!-- ## Sensitivity analysis -->
 

### 2024-06-12 manuscript.qmd
fix small mistake, add date, add citation to website, add r-references.bib

@@
 shorttitle: "Simulation-Based Sample Size Planning for GLMMs"
+date: 2024-06-12
+date-format: iso
 author:
@@
     gratitude: ~
-    authorship-agreements: Florian Pargent and Timo K. Koch contributed equally to this work.
+    authorship-agreements: ~
 abstract: When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.
@@
 We conclude with an outlook on the increasing importance of simulation-based sample size planning.
-keywords: [power analysis, precision, data simulation, sample size planning, generalized linear mixed model]
+keywords: [tutorial, sample size planning, generalized linear mixed model, power analysis, precision, data simulation]
 bibliography: [Simulation Tutorial Paper.bib, r-references.bib]
@@
 library("papaja")
-r_refs("r-references.bib")
+r_refs("r-references.bib", append = FALSE)
 ```
@@
 
-With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value in the `p (NonInf)` column for the third hypothesis is not significant) and therefore also not reject the combined null hypothesis for this particular (simulated) dataset.
+With a standard significance level of $\alpha = 0.05$, we would not reject all four null hypotheses (the p-value in the `p (NonInf)` column for the second hypothesis is not significant) and therefore also not reject the combined null hypothesis for this particular (simulated) dataset.
 Note that this decision would be wrong because we have simulated the data such that the combined alternative hypothesis $H_1$ is actually true in the population.

### 2024-06-12 _tex/manuscript.tex
Built site for gh-pages

\documentclass[
  man,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{A Tutorial on Tailored Simulation-Based Sample Size Planning for
Experimental Designs with Generalized Linear Mixed Models}

\shorttitle{Simulation-Based Sample Size Planning for GLMMs}

\usepackage{etoolbox}

\authorsnames[{1},{1,2},{1},{1,3},{1,4}]{Florian Pargent,Timo K.
Koch,Anne-Kathrin Kleine,Eva Lermer,Susanne Gaube}

\authorsaffiliations{
{Department of Psychology, LMU Munich},{Institute of Behavioral Science
\& Technology, University of St.~Gallen},{Department of Business
Psychology, Technical University of Applied Sciences Augsburg},{Global
Business School for Health, University College London}}

\leftheader{Pargent, Koch, Kleine, Lermer and Gaube}

\date{2024-06-12}

\abstract{When planning experimental research, determining an
appropriate sample size and using suitable statistical models are
crucial for robust and informative results. The recent replication
crisis underlines the need for more rigorous statistical methodology and
adequately powered designs. Generalized linear mixed models (GLMMs)
offer a flexible statistical framework to analyze experimental data with
complex (e.g., dependent and hierarchical) data structures. However,
available methods and software for a priori sample size planning for
GLMMs are often limited to specific designs. Tailored data simulation
approaches are a more flexible alternative. Based on a practical case
study, the current tutorial equips researchers with a step-by-step guide
and corresponding code for conducting tailored a priori sample size
planning with GLMMs. We not only focus on power analysis but also
explain how to use the precision of parameter estimates to determine
appropriate sample sizes. We conclude with an outlook on the increasing
importance of simulation-based sample size planning.}

\keywords{tutorial, sample size planning, generalized linear mixed
model, power analysis, precision, data simulation}

\authornote{\par{\addORCIDlink{Florian
Pargent}{0000-0002-2388-553X}}\par{\addORCIDlink{Timo K.
Koch}{0000-0001-6728-2063}}

\par{ This is version 2.5 of our preprint published at
\url{https://doi.org/10.31234/osf.io/rpjem}. All materials (reproducible
manuscript, including R code and all simulation results) are available
in the project's repository on the Open Science Framework (OSF) at
\url{https://osf.io/dhwf4/}. A Quarto Manuscripts website is hosted at
\url{https://Timo-Ko.github.io/glmm_simulation_tutorial/}.  The authors
declare that there were no conflicts of interest with respect to the
authorship or the publication of this article. This research was funded
by a grant from the Volkswagen Foundation (Grant
No.~98525).   Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Florian
Pargent:   formal
analysis, methodology, visualization, writing, editing; Timo K.
Koch:   conceptualization, formal
analysis, writing, editing; Anne-Kathrin Kleine:   formal
analysis, editing; Eva Lermer:   conceptualization, funding
aquisition, editing; Susanne
Gaube:   conceptualization, supervision, editing}
\par{Correspondence concerning this article should be addressed to Timo
K. Koch, Institute of Behavioral Science \& Technology, University of
St.~Gallen, Torstrasse
25, St.~Gallen 9000, Switzerland, Email: timo.koch@unisg.ch}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

\usepackage{float}
\usepackage{tabularray}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\UseTblrLibrary{booktabs}
\UseTblrLibrary{siunitx}
\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{Introduction}\label{introduction}

When planning experimental research, it is essential to determine an
appropriate sample size and to use appropriate statistical models to
analyze the data to ensure that the results obtained are both robust and
informative (\citeproc{ref-lakensSampleSizeJustification2022}{Lakens,
2022a}). The recent replication crisis has illustrated many challenges
surrounding the reproducibility and reliability of study findings
(\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}). As a
result, there is a growing need for more rigorous statistical
methodology and the adoption of adequately powered experimental designs.
Multiple easy-to-use software solutions exist for simple statistical
models and experimental designs. However, many researchers lack the
skills and tools to conduct ``a priori'' (i.e., before data collection)
sample size planning for more complex research designs such as flexible
generalized linear mixed models (GLMM) framework. In the present work,
we provide a tutorial on how to determine adequate sample sizes by
performing tailored simulation-based sample size planning for GLMMs.
After introducing some theoretical background on sample size planning,
we review existing software solutions in R and discuss under which
circumstances tailored data simulations are necessary. Then we describe
general steps and decisions involved in tailored data simulation. To
illustrate the details of these steps, we finish with a hypothetical
case study from the field of human-AI (artificial intelligence)
interaction research.

To benefit most of this tutorial paper, we recommend readers to
familiarise themselves with basic statistical concepts like hypothesis
tests (HTs) and their statistical power as well as confidence intervals
(CIs) and their precision
(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021};
\citeproc{ref-lakensImprovingYourStatistical2022}{Lakens, 2022b};
\citeproc{ref-riesthuisSimulationBasedPowerAnalyses2024}{Riesthuis,
2024}). Some knowledge of causal inference is useful but not necessary
(\citeproc{ref-deffnerCausalFrameworkCrossCultural2022}{Deffner et al.,
2022}; \citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al.,
2021}). In addition, readers should have a rough understanding of R
(\citeproc{ref-wickhamDataScienceImport2023}{Wickham et al., 2023}) and
how to simulate data. For data simulation, we use functions from the
tidyverse (\citeproc{ref-wickhamWelcomeTidyverse2019}{Wickham et al.,
2019}) and the faux package (\citeproc{ref-R-faux}{DeBruine, 2023}).
Finally, readers should be familiar with regression modeling in general
and GLMMs in particular. In this tutorial, we simulate data by manually
specifying the model equation of a GLMM that represents our assumed data
generating process
(\citeproc{ref-debruineUnderstandingMixedEffectsModels2021}{DeBruine \&
Barr, 2021}). It is not necessary to understand the technical details of
how GLMMs are estimated. However it is crucial to understand the
structure of a basic GLMM (e.g., logistic regression with random
intercepts) and how the model assumes that the values in the dependent
variable are determined by the predictor variables and the random
effects.

\section{Theoretical background}\label{theoretical-background}

\subsection{Planning for statistical power or
precision}\label{planning-for-statistical-power-or-precision}

Conducting research with insufficiently large sample sizes can have many
negative consequences. First, experiments may yield inconclusive or
misleading results, hindering the accumulation of knowledge. Second,
studies that are doomed to never finding a postulated effect waste
resources by consuming time, effort, and funding without delivering
meaningful results. For these reasons, many journals and funding bodies
now require that a justification of sample size is included in study
protocols and grant proposals, recognizing its significance in ensuring
robust and meaningful findings. Although scientists often do not justify
sample size or use general heuristics from the literature, resource
constraints often play an important role
(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}). But
ideally, a suitable sample size should be determined a priori (i.e.,
before the study is conducted) based on some meaningful computation to
ensure that the study will be able to fulfill its purpose.

The majority of empirical studies in psychology and other social
sciences apply hypothesis testing. As a consequence, the dominant
approach for determining an adequate sample size is based on power
analysis (i.e., planning for power)
(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}).
Statistical power is defined as the probability that a HT has a
significant p-value when analyzing repeated samples from a population
with a true effect of some pre-specified size
(\citeproc{ref-cohenPowerPrimer1992}{Cohen, 1992}). Less formally, power
is described as the probability that a HT correctly rejects the null
hypothesis when the alternative hypothesis is true. If the sample size
(i.e., the number of participants and/or stimuli) used for data
collection is insufficient to detect the effects or relationships being
investigated with high probability, the study is considered
``underpowered''. When planning for power, a target is set for the
statistical power of a HT of interest. Assuming some effect size of
interest and a desired significance level, a minimum sample size can be
determined that, on average, would guarantee reaching this target.
Although it is recommended to justify the desired significance level and
power (\citeproc{ref-lakensJustifyYourAlpha2018}{Lakens, Adolfi, et al.,
2018}), most empirical studies adopt the heuristic of \(\alpha = 0.05\)
and \(1 - \beta = 0.80\).

In contrast to power analysis, sample size planning can also be based on
the precision of parameter estimates (i.e., planning for precision or
planning for accuracy)
(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}). Not
all research questions are best answered using hypothesis testing. It
has been argued that basic research rarely requires making discrete
decisions on whether some effect has been ``discovered'' and should thus
shift from a hypothesis testing towards an estimation framework
(\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014};
\citeproc{ref-kruschkeBayesianNewStatistics2018}{Kruschke \& Liddell,
2018};
\citeproc{ref-mcelreathStatisticalRethinkingBayesian2020}{McElreath,
2020}). Although this view is not without critique
(\citeproc{ref-uyguntuncEpistemicPragmaticFunction2023}{Uygun Tunç et
al., 2023}), at least for exploratory or pilot studies where little
previous research has been conducted, more scientists seem to agree that
simply estimating the effects of interest and making the estimation
uncertainty transparent by reporting CIs is more useful. Assuming that
no HTs are conducted for a planned study, power analysis is not relevant
for sample size planning. Nonetheless, the sample size still has a
crucial effect on how informative the planned study will be, because an
effect of interest is estimated more precisely with bigger samples. In
the precision framework, the target quantity commonly used for sample
size planning is the expected width of a CI
(\citeproc{ref-kelleySampleSizePlanning2006}{Kelley \& Rausch, 2006};
\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}). The
values inside a CI are often interpreted as plausible values for the
quantity of interest it is supposed to estimate. More formally, a CI
with confidence level 0.95 provides the smallest interval with the
property that upon repeated sampling, 95\% of individual CIs would
include the true quantity of interest. Thus, a narrow CI is more
informative about the size of the true effect than a wide interval.
Apart from the desired confidence level, the width of a CI depends
strongly on the sample size. Because bigger samples carry more
information, they lead to smaller CIs. When planning for precision, a
target can be set for the expected width of a CI of interest. Assuming
some effect size of interest and a desired confidence level, a minimum
sample size can be determined that, on average, would guarantee reaching
this target. Because planning for precision is still rare, there are no
common heuristics on how to choose the desired width of the CI
(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).

\subsection{Generalized linear mixed models
(GLMMs)}\label{generalized-linear-mixed-models-glmms}

As study designs become more complex, psychological researchers require
more sophisticated statistical models to capture the nuanced
relationships and grouping structures introduced by them
(\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}). GLMMs
(also called multilevel models) are gaining increasing popularity
because they offer great flexibility
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}). GLMMs are an extension of LMMs (Linear Mixed Models), which are,
in turn, extensions of linear regression models that account for
correlated data, including hierarchical structures
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}). In this context, correlated data means that the value in the
outcome variable for one observation may be related to the value of
another observation in a systematic way that is not already accounted
for by the usual (fixed) predictor variables (e.g., age of
participants). This correlation can arise for various reasons: For
instance, responses to some stimuli from some participants might be more
similar because the same person was measured twice (repeated
measurements), participants come from the same neighborhood
(clustering), or participants responded to the same stimulus (stimulus
effects). Thus, modeling such correlations is important whenever the
data has a clear structure, while the grouping variables can be
hierarchically nested (e.g., grouping variables students and schools:
each student belongs to exactly one school) or cross-classified (e.g.,
grouping variables students and math exercises: each student is
presented with several math exercises). LMMs are used when the outcome
variable is continuous and follows a normal distribution. They allow for
the modeling of fixed effects, which capture the relationships between
the usual predictors and the outcome, as well as random effects, which
account for the different types of correlation structure and grouping
effects. Random effects are typically assumed to follow a normal
distribution with a mean of zero and a variance that quantifies the
heterogeneity across groups. GLMMs extend the LMM framework to
accommodate non-normally distributed continuous and categorical outcome
variables. GLMMs involve a link function that connects the linear
combination of predictor variables to the expected value of the outcome
variable. The link function allows for modeling the relationship between
predictors and the outcome in a non-linear way that is appropriate for
the specific distribution family of the outcome variable. For example,
think of an experiment with different design factors (e.g., picture
positions, headline aesthetics) impacting the likelihood of users
clicking on an online advertisement. The participants' behavior is
measured repeatedly over several sessions. The click patterns of
participants in one session are likely to be correlated with their
previous sessions and the outcome variable is binary (click/no click)
for each session, which follows a binomial distribution.

\section{Simulation-based sample size planning with
GLMMs}\label{simulation-based-sample-size-planning-with-glmms}

To our knowledge, existing approaches for sample size planning for GLMMs
have exclusively focused on planning for power. Power analysis methods
for multilevel models can be categorized into formula-based methods and
simulation-based methods
(\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
et al., 2022}). Formula-based methods rely on formulas to calculate
power directly while simulation-based methods rely on repeatedly
simulating data with a known true effect size and estimating power
empirically (i.e., what percentage of simulated datasets produces a
significant p-value). Available formula-based software packages for
multilevel models often do not include GLMMs or are limited to very
simple designs
(\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
et al., 2022};
\citeproc{ref-westfallStatisticalPowerOptimal2014}{Westfall et al.,
2014}), making it necessary to build data simulations tailored
specifically to the study design. A number of tutorials have been
published describing how to perform such simulation-based power analysis
for multilevel models
(\citeproc{ref-arendStatisticalPowerTwolevel2019}{Arend \& Schäfer,
2019}; \citeproc{ref-brysbaertPowerAnalysisEffect2018}{Brysbaert \&
Stevens, 2018};
\citeproc{ref-debruineUnderstandingMixedEffectsModels2021}{DeBruine \&
Barr, 2021}; \citeproc{ref-greenSIMRPackagePower2016a}{Green \& MacLeod,
2016}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al., 2015};
\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021};
\citeproc{ref-lafitSelectionNumberParticipants2021}{Lafit et al., 2021};
\citeproc{ref-zimmerSampleSizePlanning2022}{Zimmer et al., 2022}).
However, many of these tutorials focus on linear mixed models (LMMs) and
the common study designs (but see
\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021},
for a tutorial that also covers more advanced settings). This narrow
focus provides limited guidance for researchers faced with more complex
study designs, especially when little prior knowledge about plausible
effect sizes is available (see the discussion in
\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}).
Simulation-based power analysis with GLMMs requires making a range of
assumptions: The (conditional) distribution assumption specifies the
distributional family for the outcome variable. Assumptions about the
random effects include the assumption of normality (i.e., that the
random effects follow a normal distribution) and the covariance
structure among the random effects (i.e., if and how they are
correlated). Making these decisions requires understanding the
underlying assumptions of the model and ensuring they align with the
characteristics of the data being analyzed. Existing tutorials often
rely on heuristics for specifying variance components (e.g., the
standard deviation of random intercepts) or assume that results from
meta-analyses or data from pilot studies are available to determine
plausible values for all model parameters. However, in practice,
knowledge about those parameters from prior studies is often limited,
which makes specifying assumptions a practical challenge (see the
discussion in (\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et
al., 2008}) and
(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al.,
2021})).

INSERT TABLE 1 HERE!

In Table 1, we give a short review of existing R packages that can be
used for power analysis for GLMMs.

\section{When to use tailored data
simulation?}\label{when-to-use-tailored-data-simulation}

Performing tailored simulation-based sample size planning is more
complicated and time-consuming than using the existing software tools
outlined in TABLE 1. The most important circumstances under which
tailored simulation-based sample size planning is necessary are 1)
complex study designs, 2) complex statistical hypotheses, 3) planning
for precision, 4) no available prior studies or pilot data.

Requirements of real-world studies are often more complex than the
simplified designs assumed by many user-friendly software packages for
sample size planning. One frequent issue in applied data analysis is
missing data, and there can be various reasons for this
(\citeproc{ref-littleStatisticalAnalysisMissing2014}{Little \& Rubin,
2014}). For example, data can be missing completely at random (e.g.,
because an electronic measurement device randomly failed for some
technical reasons). Alternatively, subjects might systematically drop
out or produce missing data, but this dropout can be explained by some
attributes also measured in the dataset (e.g., older subjects have a
higher probability to refuse answering a question on income). In a more
complicated scenario, missing data in some variable is caused by the
measured attribute itself (e.g., wealthy people are more likely to
refuse reporting their income). Moreover, many experimental designs
contain conditions in which values of the predictor variables are
missing by design. This can make data analysis more complicated because
predictors have to be coded in specific ways that prevent the estimated
GLMM from becoming unidentified. Whether missing data has an effect on
the sample size planning depends on our theoretical assumptions on how
the missingness is caused. However, it is often challenging to decide
whether missing data can be safely ignored in the data analysis and
sample size planning process based on a merely theoretical approach
(\citeproc{ref-gomilaMissingDataExperiments2022}{Gomila \& Clark,
2022}). Tailored simulation-based approaches offer the possibility to
include the assumed process of how data become missing in the data
simulation, thereby determining the required sample based on simulated
datasets that contain missing values (for an example, see
\citeproc{ref-lanePowerStrugglesEstimating2018}{Lane \& Hennes, 2018}).
As a byproduct, the simulated datasets can also be used to test whether
the intended data analysis provides the expected (unbiased) results,
despite the missing data. Although GLMMs can handle a large variety of
outcome variables, researchers are becoming increasingly aware that many
datasets might profit from even more sophisticated models. Common
examples are zero-inflated outcomes, censoring, and nonlinear predictor
effects that can be modeled with the R packages glmmTMB
(\citeproc{ref-brooksGlmmTMBBalancesSpeed2017}{Brooks et al., 2017}) or
brms (\citeproc{ref-burknerAdvancedBayesianMultilevel2018}{Bürkner,
2018}). Tailored simulation-based approaches do not share the same
limitation than the existing software solutions for power analysis that
focus exclusively on GLMMs. As long as there is a software package
available to estimate the model of interest, it is always possible to
perform tailored simulation-based sample size planning.

The most common hypotheses tested in psychological research are of the
type \(H_0: \beta = 0\), where \(\beta\) is a slope or intercept of a
regression model. However, many research questions in psychology
actually require testing more complex statistical hypotheses. In the new
era of preregistration and registered reports
(\citeproc{ref-chambersPresentFutureRegistered2022}{Chambers \&
Tzavella, 2022}), most research questions should be tested with directed
hypotheses because good theories at least postulate whether some
psychological effect of interest is positive or negative. Even better
theories should be able to specify the smallest effect sizes of interest
(SESOI) that must be exceeded if the effect has any practical relevance
(\citeproc{ref-lakensEquivalenceTestingPsychological2018}{Lakens,
Scheel, et al., 2018}). In combination, this might require a test such
as \(H_0: \beta \leq 0.1\). More elaborate research questions often
require testing hypotheses that consist of a combination of model
parameters, for example testing simple slopes
(\citeproc{ref-preacherComputationalToolsProbing2006}{Preacher et al.,
2006}) with a hypothesis such as \(H_0: \beta_0 + \beta_1 \leq 0\). If
the research question consists only of a single hypothesis of this sort,
it might be possible to reduce the hypothesis to a single regression
coefficient by clever coding and/or centering of predictor variables.
However, interesting research questions often consist of combined
hypotheses that consist of more than one separate statistical hypothesis
(for a tutorial on contrast analysis in GLMMs, see
\citeproc{ref-schadHowCapitalizePriori2020}{Schad et al., 2020}). For
example, a combined null hypothesis \(H_0\) might consist of two single
null hypotheses \(H_{01}: \beta_1 \leq 0\) and
\(H_{02}: \beta_0 + \beta_1 \leq 0\). For some research questions, the
combined null hypothesis \(H_0\) would be rejected if both \(H_{01}\)
\emph{AND} \(H_{02}\) are rejected. For other research questions, the
combined null hypothesis \(H_0\) would be rejected if \(H_{01}\)
\emph{OR} \(H_{02}\) \emph{OR} both are rejected. If the global
hypothesis \(H_0\) is combined with \emph{OR}, the p-values of the
single hypotheses must be corrected for multiple testing to avoid
\(\alpha\)-inflation for the global hypothesis
(\citeproc{ref-dmitrienkoTraditionalMultiplicityAdjustment2013}{Dmitrienko
\& D'Agostino, 2013}). However, if the global hypothesis \(H_0\) is
combined with \emph{AND}, a correction for multiple testing is not
necessary but rather a mistake that unnecessarily reduces the power of
the global HT. None of the software packages for sample size planning in
table X can handle combined hypotheses as discussed here and only some
can handle directed hypotheses. In contrast, our case study will
demonstrate how we can test directed combined hypotheses with tailored
simulation-based sample size planning.

Although planning for precision
(\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014};
\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}) has
been increasingly discussed as a useful strategy for empirical research,
all available software packages for sample size planning with GLMMs are
based on power analysis. Therefore, researchers that want to want to
apply an estimation strategy in their studies, without testing any
statistical hypotheses, currently cannot use the software packages
outlined in TABLE 1. However, tailored simulation-based sample size
planning can easily handle the planning for precision approach
(\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}).
The only change in procedure is that instead of computing HTs for each
simulated dataset and estimating statistical power across repetitions,
CIs are computed for each simulated dataset, and the expected width is
estimated.

All frameworks for sample size planning require the user to make
assumptions about the expected effect size. Assuming the true effect is
of this size (or greater), one can compute the (minimum) power of a HT
or the (maximum) expected width of a CI. Existing software packages for
sample size planning for GLMMs usually require to provide the assumed
effect in the unit of some standardized measure of effect size. When the
researcher has access to similar studies or pilot data, providing such
standardized effect sizes is feasible. However, note that choosing
effect sizes based on small pilot studies is generally not recommended,
as those estimates can be heavily biased
(\citeproc{ref-albersWhenPowerAnalyses2018a}{Albers \& Lakens, 2018};
\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).
Providing an informed standardized effect size can be an almost
impossible challenge when no prior studies of pilot data are available.
This problem is further exacerbated by the fact that GLMMs are so
flexible that general heuristics of what should be considered a small
effect do not exist or are difficult to defend. In our experience, using
domain knowledge to construct a tailored data simulation is the only
solution to determine plausible effect sizes in the absence of prior
evidence. It would be possible to use these tailored simulations to
extract plausible values for standardized effect sizes that could then
be inserted in existing software packages for sample size planning.
However, we would argue that when tailored data simulations necessary to
determine effect sizes anyway, performing the whole sample size planning
in a customized way is preferred over using the existing software
packages.

\section{General steps in tailored simulation-based sample size
planning}\label{general-steps-in-tailored-simulation-based-sample-size-planning}

Although the details differ depending on the specific study
characteristics, each tailored simulation-based sample size planning
requires a series of steps and decisions. We will introduce each step in
a theoretical section, followed by the practical application based on a
hypothetical case study. All code in this manuscript and simulation
results are available in the project's repository on the Open Science
Framework (\url{https://osf.io/dhwf4/}).

\subsection{Step 1: Define the
estimand}\label{step-1-define-the-estimand}

\subsubsection{THEORY}\label{theory}

The first step in every research process should be a clear definition of
the theoretical \emph{estimand}
(\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}),
i.e.~the theoretical quantity which is necessary to answer a specific
research question. The estimand consists of a quantity, that can be
described for each unit under investigation and a clear definition of
the target population, for which the quantity is of interest. For
example, an estimand might be the probability that a clinical
psychologists makes the correct diagnosis for a psychiatric patient with
major depression, averaged across all clinical psychologists and
depressed patients in psychiatric institutions in Germany.

The estimand should always be defined outside of any statistical model,
because there are usually a range of statistical methods that can be
used to estimate the same estimand, depending on the study design (e.g.,
a randomized experiment) that will produce the observed data in the
planned study. For many common research questions in psychology, the
estimand can be expressed as a statistical quantity that can be
estimated with a regression model, for example a single \(\beta\)
coefficient. However, this is not possible for all estimands, which is
why the literature discusses many estimation strategies beyond
regression
(\citeproc{ref-deffnerCausalFrameworkCrossCultural2022}{Deffner et al.,
2022}; \citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al.,
2021}).

\subsubsection{PRACTICE}\label{practice}

In the present hypothetical case study, we consider the effectiveness of
feedback provided by an artificial intelligence (AI) embedded in a
diagnostic decision support system. The context is a clinical setting,
where expert radiologists and students under training must detect
bleeding based on head scans from computer tomography (CT). In the
investigated AI-enabled diagnostic decision support system, an AI model
can provide initial diagnostic advice, which can be used as guidance by
the humans who are required to make the final diagnostic decision. The
research goal is to validate the effectiveness of the AI-enabled advice.
We consider the AI-enabled advice as effective, if the following pattern
holds, which we will first describe verbally:

\emph{We expect that for BOTH expert radiologists and medical students,
correct AI-advice leads to a higher probability of accurately diagnosing
a CT scan compared to no AI-advice presented, AND, we expect that for
BOTH experts and non-experts, incorrect advice leads to a lower
probability of accurately diagnosing a CT scan compared to no advice
presented.}

It becomes clear that our estimand consists of four comparisons between
experimental conditions
(\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}).
However, the verbal description is still somewhat vague, which is why we
try to give a more precise expression for each comparison:

\[
\begin{aligned}
& P(\text{correct diagnosis} | \text{correct advice, average expert, average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{no advice, average expert, average scan})
\end{aligned}
\]

\[
\begin{aligned}
& P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan})
\end{aligned}
\] \[
\begin{aligned}
& P(\text{correct diagnosis} | \text{correct advice, average student, average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{no advice, average student, average scan})
\end{aligned}
\]

\[
\begin{aligned}
& P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\
& \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan})
\end{aligned}
\] For example, the first expression is concerned with the difference
between the probability that a correct diagnosis is made if correct
AI-advice is presented and the probability that a correct diagnosis is
made if no AI-advice is presented. This contrast is quantified for a
hypothetical \emph{typical} expert and a \emph{typical} scan, where
\emph{typical} is usually defined as an average score on all attributes
of the expert or scan.\footnote{Note that a different estimand would be
  the so-called average treatment effect (ATE). For the ATE, the
  probability contrast is defined for each combination of expert and
  scan, and then these contrasts are averaged across all experts and
  scans from the target population
  (\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}).}.
To complete our definition of the estimand, we have to define our target
population that consists of persons, stimuli, and AI-advice: With
respect to persons, we are only interested in expert radiologists and
medical students at German universities. With respect to stimuli, we are
only interested in the head CT scans made from subjects that do or do
not suffer from intracerebral hemorrhage. Lastly, we are only interested
in AI-advice given by a specific AI-enabled diagnostic decision support
system.

Although the estimand is initially defined outside of any statistical
model, it is only useful if we find a way to estimate it based on
observed data. For our exemplary research question, it is possible to
construct an experimental study, where all participants are confronted
with the same set of head CT scans, but the kind of AI-advice given for
each scan is randomly assigned within participants. This random
intervention allows us to produce an empirical estimate of our estimand,
although, in reality, each person receives only one kind of AI-advice
(correct advice, wrong advice, no advice) for each scan. We will see
later how each of the probability expressions in our estimand can be
modeled with the same GLMM. Estimating this GLMM based on the data
observed in our planned study will produce an estimate for each
probability, and these estimates can then be combined to compute an
estimate for each of the four probability contrasts. For pedagogical
reasons, we will skip the concrete definition of our estimate until we
have discussed how to simulate data based on a concrete GLMM in the next
section.

\subsection{Step 2: Simulate the data generating
process}\label{step-2-simulate-the-data-generating-process}

\paragraph{THEORY.}\label{theory-1}

When the estimand has been defined, the next step in the research
process is to write code that simulates the (assumed) data generating
process of the planned study. This requires specifying a generative
process for all predictor variables used in the final data analysis.
While such assumptions can be quite challenging for observational
studies or continuous predictor variables, this is less of a problem for
experimental studies with only categorical predictor variables. When all
predictor variables have been simulated, one can use the structure of a
suitable GLMM to simulate the dependent variable. To simulate the GLMM,
one requires plausible values for all model parameters. We will discuss
strategies on how these values can be obtained later. Because we have
full control over the data generating process in a tailored simulation,
it is possible to model specific aspects of the planned study, like data
missing by design or assuming that some subjects might drop out. The
quality of the results of the sample size planning crucially depends on
the plausibility of the simulated data generating process. However, we
would argue that even a strongly simplified data generating process
(e.g.~only a small number of interaction effects; only random intercepts
and no random slopes; assuming that data is missing completely at
random) can yield informative results.

\paragraph{PRACTICE.}\label{practice-1}

In our case study, we simulate data for an experiment where the
diagnostic performance of users of an AI-enabled diagnostic decision
support system will be evaluated. Radiologists (task experts) and
students/interns (non experts) review a series of head computer
tomography (CT) scans to assess the presence of a bleeding. An AI model
provides initial diagnostic advice to assist their decision-making. In
the control condition, no AI advice is presented. When AI advice is
given, this advice can be either correct or incorrect. The type of
advice (no advice, wrong advice, correct advice) is randomized within
subjects across brain scans. After reviewing a CT scan, participants
deliver a medical diagnosis (bleeding or no bleeding), which may be
either accurate or inaccurate. This experimental design introduces some
missing values by design since the advice is neither correct nor
incorrect when no advice is present, which must be taken into account
when simulating and analyzing the data. In this example, recruiting task
experts (i.e., radiologists) is more challenging due to their limited
availability, while non-experts (i.e., students/interns) are more
readily accessible. The goal of simulation-based sample size planning is
to determine how many task experts and non-experts must be recruited to
achieve sufficient statistical power or precision in the planned
experiment.

\subparagraph{Our specific GLMM.}\label{our-specific-glmm}

In a GLMM, the expected value of the dependent variable \(Y\)
conditioned on the vector of predictor variables \(\mathbf{X}\) and
random effects \(\mathbf{U}\), transformed by a link function \(g()\) is
modeled as a linear combination \(\eta\) of the predictor variables
\(\mathbf{X}\), the random effects \(\mathbf{U}\) and the model
parameters \(\mathbf{\beta}\)
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}): \[
g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta
\] Equivalently, the conditional expected value is modeled as the linear
combination \(\eta\), transformed by the inverse link function
\(g^{-1}()\): \[
E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u}) = g^{-1}(\eta)
\] If the dependent variable (i.e., diagnostic decision) \(Y\) is a
binary variable with values \(0\) (i.e., inaccurate), or \(1\) (i.e.,
accurate), the conditional expected value is equivalent to the
probability: \[
P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
\] In our case study, \(P_{si}\) is the conditional probability that
subject \(s\) gives the correct response to item (i.e., CT scan) \(i\).

In such a setting, we model this probability as \[
P_{si} = \text{inverse\_logit}(\eta_{si})
\] with the inverse-logit link
\(g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}\)
or equivalently \[
\text{logit}(P_{si}) = \eta_{si}
\] with the logit link
\(g(P_{si}) = \text{logit}(P_{si}) = \text{ln} (\frac{P_{si}}{1 - P_{si}})\).

In our case study, the probability of making an accurate diagnostic
decision is assumed to depend on the predictors:

\begin{itemize}
\tightlist
\item
  \(advice\_present_{si}\): whether subject \(s\) was presented with AI
  advice (1) or not (0) when asked to assess item \(i\)
\item
  \(advice\_correct_{si}\): whether this advice was correct (1) or not
  (0)
\item
  \(expert_s\): whether subject \(s\) was a task expert (1) or not (0)
\end{itemize}

and the random effects:

\begin{itemize}
\tightlist
\item
  \(u_{0s}\): the deviation of subject \(s\) from the average ability to
  solve an item (i.e., CT scan) with average difficulty; assumed to be
  distributed as \(u_{0s} \sim N(0, \sigma_S^2)\)
\item
  \(u_{0i}\): the deviation of item (i.e., CT scan) \(i\) from the
  average difficulty to be solved by a person with average ability;
  assumed to be distributed as \(u_{0i} \sim N(0, \sigma_I^2)\)
\end{itemize}

In total, we assume the model \[
\begin{aligned}
\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
\end{aligned}
\] or equivalently \[
\begin{aligned}
P_{si} = \text{inverse\_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
\] with model parameters \(\beta_0\), \(\beta_e\), \(\beta_a\),
\(\beta_c\), \(\beta_{ea}\), \(\beta_{ec}\), \(\sigma_S\), and
\(\sigma_I\).

In the GLMM literature, this would be called a binomial GLMM with two
random intercepts (for subjects and items), two level-1 predictors
(\(advice\_present\), \(advice\_correct\)), one level-2 predictor
(\(expert\)) and two cross-level interactions
(\(expert \cdot advice\_present\), \(expert \cdot advice\_correct\)). To
limit complexity, we do not consider random slopes, additional
predictors or higher-level interactions.

\subparagraph{Simulation function in R.}\label{simulation-function-in-r}

The following R function simulates a full dataset structured according
to the design of our case study.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{simulate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{100}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{50}\NormalTok{,}
  \AttributeTok{b\_0 =} \FloatTok{0.847}\NormalTok{, }\AttributeTok{b\_e =} \FloatTok{1.350}\NormalTok{, }\AttributeTok{b\_a =} \SpecialCharTok{{-}}\FloatTok{1.253}\NormalTok{, }\AttributeTok{b\_c =} \FloatTok{2.603}\NormalTok{,}
  \AttributeTok{b\_ea =} \FloatTok{0.790}\NormalTok{, }\AttributeTok{b\_ec =} \SpecialCharTok{{-}}\FloatTok{1.393}\NormalTok{,}
  \AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{, ...)\{}
  \FunctionTok{require}\NormalTok{(dplyr)}
  \FunctionTok{require}\NormalTok{(faux)}
  \CommentTok{\# simulate design}
\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{add\_random}\NormalTok{(}\AttributeTok{subject =}\NormalTok{ n\_subjects, }\AttributeTok{item =}\NormalTok{ n\_items) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{add\_between}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{expert =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{.prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.25}\NormalTok{, }\FloatTok{0.75}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_present =} \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{, }\AttributeTok{prob =} \DecValTok{2}\SpecialCharTok{/}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_correct =} \FunctionTok{if\_else}\NormalTok{(advice\_present }\SpecialCharTok{==} \DecValTok{1}\NormalTok{L, }
                                    \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{L, }\AttributeTok{prob =} \FloatTok{0.8}\NormalTok{), }\DecValTok{0}\NormalTok{L)) }\SpecialCharTok{|\textgreater{}}
    \CommentTok{\# add random effects}
    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{u0s =}\NormalTok{ sd\_u0s) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"item"}\NormalTok{, }\AttributeTok{u0i =}\NormalTok{ sd\_u0i) }\SpecialCharTok{|\textgreater{}}
    \CommentTok{\# compute dependent variable}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{linpred =}\NormalTok{ b\_0 }\SpecialCharTok{+}\NormalTok{ u0i }\SpecialCharTok{+}\NormalTok{ u0s }\SpecialCharTok{+}
\NormalTok{        b\_e }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{*}\NormalTok{ advice\_correct }\SpecialCharTok{+}
\NormalTok{        b\_ea }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_ec }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_correct) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_prob =} \FunctionTok{plogis}\NormalTok{(linpred)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_bin =} \FunctionTok{rbinom}\NormalTok{(}\AttributeTok{n =} \FunctionTok{n}\NormalTok{(), }\AttributeTok{size =} \DecValTok{1}\NormalTok{, }\AttributeTok{prob =}\NormalTok{ y\_prob))}
\NormalTok{  dat}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

In the first four lines of the function definition, we set some default
parameter values (which we will explain in the next section) and load
the packages we use to manipulate and simulate data.\footnote{The faux
  package (\citeproc{ref-R-faux}{DeBruine, 2023}) contains useful
  functions when simulating factorial designs, including random effects.}
In our case study, each subject (\texttt{n\_subjects} in total) is
assumed to respond to each item (i.e., CT scan; \texttt{n\_items} in
total). Thus, the \texttt{add\_random} command creates a fully-crossed
\texttt{data.frame} with \texttt{n\_subjects} \(\times\)
\texttt{n\_items} rows. We add a between-subject effect with the
\texttt{add\_between} command, simulating that about \(25\%\) of
subjects are experts. The next two lines simulate that in
\(\frac{2}{3}\) of trials, subjects will be presented with AI advice,
and if advice is presented, the advice will be correct in about \(80\%\)
of cases (the variable \texttt{advice\_correct} is always 0 when no
advice is presented). Next, we simulate one random effect for each
subject (\texttt{u0s}) and for each item (\texttt{u0i}). As assumed by
standard GLMMs, the \texttt{add\_ranef} function draws the random
effects from a normal distribution with a mean 0 and a standard
deviation specified by the user. With all design variables done, we are
ready to simulate our model equation outlined in the last section. The
linear predictor variable \texttt{linpred} (\(\eta\) in the GLMM model
equations) combines the predictor variables, random effects, and model
parameters as assumed by our model. We then transform the linear
predictor with the inverse-link function to compute \texttt{y\_prob},
the probability that the subject correctly solved the item (in R, the
inverse-logit link is computed with \texttt{plogis} and the logit link
with \texttt{qlogis}). In the final step, we simulate the binary
dependent variable \texttt{y\_bin} (i.e., whether the subject makes an
accurate diagnostic decision for the CT scan) by -- for each trial --
drawing from a Bernoulli distribution with success probability
\texttt{y\_prob}.

\subsection{Step 3: Specify the population
parameters}\label{step-3-specify-the-population-parameters}

\subsubsection{THEORY}\label{theory-2}

In the absence of previous studies with the same design or pilot data,
researchers require strategies on how to specify the population
parameters used in their data simulation. Population parameters are all
model parameters estimated in a GLMM, in particular the regression
coefficients of the fixed effects and the standard deviation of the
random effects (and the correlation between random effects in more
complicated models). In contrast to non-hierarchical linear regression,
common heuristics based on standardized effect sizes are less useful or
not even available for GLMMs. Our strategies to specify population
parameters will require access to domain knowledge from domain experts.
Because most domain knowledge can only be expressed in unstandardized
measurement units of a specific application, we argue that
unstandardized effect sizes are usually preferable over standardized
effect sizes for tailored simulation-based sample size planning. The
basic idea of all strategies is how the data generating process implied
by certain values of population parameters can be quantified or
visualized in an intuitive way that enables a calibration of population
parameters based on the available knowledge of domain experts.

Although we use frequentist model estimation in our tutorial, many
strategies described in this chapter are inspired by research on how to
monitor the plausibility of model assumptions in applied Bayesian
statistics (\citeproc{ref-gelmanBayesianWorkflow2020}{Gelman et al.,
2020}).

\subsubsection{PRACTICE}\label{practice-2}

When introducing the simulation function for our case study, we have
used theoretically plausible values as defaults for all model parameters
(\(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\),
\(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\)), but have not talked
about where these numbers came from.

The starting point for all parameter values in our present case study
were based on results from distantly related study designs in the
literature. Additionally, we had repeated discussions with our
affiliated domain experts in radiology to check whether our assumptions
seem plausible.

We now outline our main strategy to determine plausible parameter values
for the fixed effects (\(\beta\) parameters): Unfortunately, the model
parameters in a binomial GLMM are hard to interpret in isolation because
1) the \(\beta\) parameters are connected to the modeled probability via
the non-linear inverse-logit link, and 2) we also have to consider the
random effects. The most simple interpretation, that allows us to ignore
the random effects for now, works by imagining a subject with average
ability (\(u_{0s} = 0\)) responding to an item (i.e., CT scan) with
average difficulty (\(u_{0i} = 0\)). Then the model implied probability
that such a person solves such an item accurately is given by:

\[
\begin{aligned}
P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
\end{aligned}
\] In fact, we would only need the full equation if the subject is an
expert and correct advice is presented. In all other experimental
conditions, some terms drop from the equation because they are
multiplied by \(0\). For example, the probability that a non-expert with
average ability solves an item with average difficulty when no advice is
presented, only requires the intercept: \[
\begin{aligned}
P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
= \text{inverse\_logit}[\beta_0]
\end{aligned}
\] We can revert this perspective by choosing plausible probability
values based on domain knowledge and deriving the parameter values
implied by these probabilities for each experimental condition.

\begin{table}

{\caption{{Assumed probabilities that an average subject solves an
average item in each experimental condition.}{\label{tbl-probtable}}}
\vspace{-20pt}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1646}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3291}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5063}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Experimental condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(P(Y=1 \mid \mathbf{X=x}, \mathbf{U} = \mathbf{0})\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Implied equation
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
no advice, no expert & 0.70 & \(logit(0.70) = \beta_0\) \\
no advice, expert & 0.90 & \(logit(0.90) = \beta_0 + \beta_e\) \\
false advice, no expert & 0.40 & \(logit(0.40) = \beta_0 + \beta_a\) \\
false advice, expert & 0.85 &
\(logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \\
correct advice, no expert & 0.90 &
\(logit(0.90) = \beta_0 + \beta_a + \beta_c\) \\
correct advice, expert & 0.95 &
\(logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\) \\
\end{longtable}

{\vspace{-20pt}
\noindent \emph{Note.} Implied equations are derived based on the model equations and setting all random intercept terms to 0.}

\end{table}

Table~\ref{tbl-probtable} shows our set of assumptions concerning the
probability that an average subject solves an average item for each
experimental condition, as well as the corresponding equations implied
by the model. The table can be used to compute the implied values for
the \(\beta\) parameters, starting with the first equation and
reinserting the computed \(\beta\) values in all following equations
(\texttt{b\_0} stands for the intercept \(\beta_0\), \texttt{b\_e} for
the slope \(\beta_e\), and so on):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{b\_0 }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.7}\NormalTok{)}
\NormalTok{b\_e }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
\NormalTok{b\_a }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.4}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0}
\NormalTok{b\_ea }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.85}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a}
\NormalTok{b\_c }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.9}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_a}
\NormalTok{b\_ec }\OtherTok{\textless{}{-}} \FunctionTok{qlogis}\NormalTok{(}\FloatTok{0.95}\NormalTok{) }\SpecialCharTok{{-}}\NormalTok{ b\_0 }\SpecialCharTok{{-}}\NormalTok{ b\_e }\SpecialCharTok{{-}}\NormalTok{ b\_a }\SpecialCharTok{{-}}\NormalTok{ b\_c }\SpecialCharTok{{-}}\NormalTok{ b\_ea}
\FunctionTok{c}\NormalTok{(}\AttributeTok{b\_0 =}\NormalTok{ b\_0, }\AttributeTok{b\_e =}\NormalTok{ b\_e, }\AttributeTok{b\_a =}\NormalTok{ b\_a, }\AttributeTok{b\_c =}\NormalTok{ b\_c, }\AttributeTok{b\_ea =}\NormalTok{ b\_ea, }\AttributeTok{b\_ec =}\NormalTok{ b\_ec)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
       b_0        b_e        b_a        b_c       b_ea       b_ec 
 0.8472979  1.3499267 -1.2527630  2.6026897  0.7901394 -1.3928518 
\end{verbatim}

It is always possible to double-check these computations by transforming
the parameter values back to probabilities, e.g.~ \[
\begin{aligned}
P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
\end{aligned}
\]

which we compute in R as:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plogis}\NormalTok{(b\_0 }\SpecialCharTok{+}\NormalTok{ b\_e }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{+}\NormalTok{ b\_ea }\SpecialCharTok{+}\NormalTok{ b\_ec)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.95
\end{verbatim}

This leaves us with the question on how to determine plausible values
for the two remaining model parameters (\(\sigma_S\), and \(\sigma_I\))
that are the standard deviations for the random intercepts. For this, we
introduce two more strategies.

\subsubsection{Insightful descriptive
statistics}\label{insightful-descriptive-statistics}

\paragraph{THEORY.}\label{theory-3}

The mathematical structure of GLMMs determines which patterns in data
would be produced by the model, if a specific set of values for the
population parameters would be specified. The knowledge of how to
simulate from a GLMM enables us to compute insightful descriptive
statistics that can be compared to available domain knowledge much more
easily than the opaque values of model parameters. For example, domain
experts might not be able to directly choose plausible values for the
coefficients in a logistic regression model (which are measured on the
log-odds scale). However, they should be able to reason about the
expected ratio of the binary dependent variable in different
experimental conditions, i.e.~which relative frequency they expect to
observe (on average). The job of the analyst who is familiar with the
mathematical structure of the GLMM is to produce the model implied value
of the insightful descriptive statistic expected by the domain expert.
Although insightful descriptive statistics usually depend on the model
parameters in a non-linear way, it is not necessary to solve the exact
relationship mathematically. Instead, one can simply adjust the
population parameters by trial and error until the model implied
quantities produce the desired result.

\paragraph{PRACTICE.}\label{practice-3}

In the last section, we showed how we can derive the model implied
probability that a subject with average ability solves an item with
average difficulty for each experimental condition. Although these
derivations are straightforward, it is important not to misinterpret
their implications: In binomial GLMMs, the average probability to solve
an item (averaged across persons of varying ability and items of varying
difficulty) is \textbf{not} equal to the probability that a person with
average ability solves an item with average difficulty
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}). The first perspective implies a so-called marginal
interpretation, while the second one implies a conditional
interpretation.

For example, we determined the \(\beta\) parameters in a way that
corresponds to a desired conditional probability of \(0.95\), that an
expert with average ability solves an item with average difficulty when
presented with correct advice (the conditional perspective). However,
even if the model were true, we would not observe that 95\% of experts
responding to items presented with correct advice from a big sample of
subjects drawn from their natural distribution of ability, and items
drawn from their natural distribution of difficulty (the marginal
perspective). How much the two probabilities differ depends on the
standard deviations of the random intercepts (the two probabilities are
only equal if both standard deviations would be zero). We want to use
the model implied observed proportion of correct diagnoses in each
experimental condition as an insightful descriptive statistics to
determine plausible values for the random effect standard deviations.

We will simulate a large dataset (for which the observed values of the
descriptive statistic will be close to their model implied true values)
and simply compute the relative frequency of correct diagnoses for each
experimental condition.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{3000}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{3000}\NormalTok{,}
  \AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{)}
\NormalTok{dat }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
 \StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
 \StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
 \StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{group\_by}\NormalTok{(condition) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{summarize}\NormalTok{(}\AttributeTok{relative\_frequency =} \FunctionTok{sum}\NormalTok{(y\_bin) }\SpecialCharTok{/} \FunctionTok{n}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 6 x 2
  condition                 relative_frequency
  <fct>                                  <dbl>
1 no expert, no advice                   0.683
2 expert, no advice                      0.881
3 no expert, wrong advice                0.409
4 expert, wrong advice                   0.828
5 no expert, correct advice              0.883
6 expert, correct advice                 0.938
\end{verbatim}

We tried using these descriptive statistics to judge together with
domain experts whether our chosen values for the random effect standard
deviations would produce data that align with out domain expertise.
However, although the result was deemed plausible, these statistics were
not informative enough to determine a final set of plausible parameter
values. For this reason, we will additionally look at insightful model
based quantities.

\subsubsection{Insightful model based
quantities}\label{insightful-model-based-quantities}

\paragraph{THEORY.}\label{theory-4}

Because GLMMs are complicated models, descriptive statistics alone are
usually not enough to specify plausible values for all model parameters.
This is especially true for the standard deviation of random effects
that have non-linear (and often unexpected) effects on the model-implied
results. An important advantage of data simulation (where one has full
control over parameter values and sample sizes) is that one can produce
insightful model based quantities that can never be directly observed in
an actual empirical dataset. For example, in a logistic model with
random intercepts for participants, one can produce a visualization of
the implied distribution of the probability that a participant on
average solves a cognitive task. Although domain knowledge will probably
not suffice to specify this distribution completely, it should be
possible to rule out implausible boundary conditions. For example, the
domain expert might deem it implausible that the 5\% most able
participants have a probability of more than 0.99 to solve the difficult
cognitive task.

\paragraph{PRACTICE.}\label{practice-4}

The discussed inequality of conditional and marginal effects in GLMMs
(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
2021}) makes their interpretation more difficult. One must be careful
when specifying parameter values based on previous studies or pilot data
that use the marginal interpretation (e.g., a pilot study providing an
estimate of how often radiologists make an accurate diagnosis based on
brain scans). However, this does not mean that we cannot use the
marginal interpretation (average probability across persons and items)
to inform plausible parameter values: When parameter values have been
selected, we can compute the implied marginal distributions and compare
this information to our domain knowledge. Then, we can iteratively
adjust the parameter values until we are satisfied with the implied
distributions. In the last section, we simulated a large dataset and
computed descriptive statistics, the relative frequencies of correct
diagnoses, for each experimental condition. We will now use the model
implied probability of each simulated data point (stored in the variable
\texttt{y\_prob}) to visualize the whole model implied marginal
distribution of correct diagnoses for each experimental condition.

\phantomsection\label{cell-fig-margdist1}
\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggdist)}
\NormalTok{dat }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
\StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
\StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
\StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ y\_prob, }\AttributeTok{y =}\NormalTok{ condition)) }\SpecialCharTok{+}
  \FunctionTok{stat\_histinterval}\NormalTok{(}\AttributeTok{point\_interval =} \StringTok{"mean\_qi"}\NormalTok{, }\AttributeTok{slab\_color =} \StringTok{"gray45"}\NormalTok{,}
    \AttributeTok{breaks =} \StringTok{"Sturges"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{breaks =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FloatTok{0.1}\NormalTok{), }\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{\label{fig-margdist1}Marginal distributions including means,
66\% and 95\% confidence intervals for all experimental conditions.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-margdist1-1.pdf}

}

\end{figure}%

Figure~\ref{fig-margdist1} shows the model implied marginal
distributions, including the mean, 66\% and 95\% intervals. We can see
that, indeed, the average probabilities (black dots) slightly differ
from the probabilities of average subjects and items considered in the
previous section. This difference increases with the variability of the
random effects. We can use plots like the one above as a useful tool to
decide whether the specified standard deviations of the subject and item
random intercepts (\(\sigma_S\) and \(\sigma_I\)) are reasonable by
comparing the ranges and overlap between conditions to domain knowledge.

\phantomsection\label{cell-fig-margdist2}
\begin{figure}[H]

\caption{\label{fig-margdist2}Marginal distributions including means,
66\% and 95\% confidence intervals for all experimental conditions while
setting the standard deviation of item random intercepts to 0.01.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-margdist2-1.pdf}

}

\end{figure}%

In the next plot, we have set the item standard deviation to almost zero
(\(\sigma_I = 0.01\)). This gives us a better way to see the variability
between persons. As an example, Figure~\ref{fig-margdist2} reveals a
number of implicit assumptions about the comparison between experts and
non-experts: With wrong advice, virtually all experts have a higher
probability of making a correct diagnosis compared to non-experts when
considering only items with average difficulty. In contrast, there is
considerable overlap in probability between experts and non-experts with
no advice and even higher overlap with correct advice. Patterns like
these should be considered carefully and discussed with the domain
experts. Parameter values (\(\beta\) parameters, and \(\sigma_S\))
should be adjusted if the implications do not seem reasonable. We could
also have a closer look at variability between items by setting the
subject standard deviation to almost zero (\(\sigma_S = 0.01\)).

The final plot demonstrates that these plots are also useful for
spotting standard deviations that are specified too high. For
Figure~\ref{fig-margdist3} we have set \(\sigma_S = 3\) and
\(\sigma_I = 3\). This implies that in each experimental condition, the
probabilities that a subject solves an item are usually close to either
0 or 1, which is not a plausible assumption. These high standard
deviations do not account for the inherent variability and complexity of
human performance. For example, we would expect that a participant with
low ability compared to other task experts to solve a difficult item
with a probability substantially larger than zero even when presented
with wrong advice.

\phantomsection\label{cell-fig-margdist3}
\begin{figure}[H]

\caption{\label{fig-margdist3}Marginal distributions including means,
66\% and 95\% confidence intervals for all experimental conditions while
setting the standard deviation of subject and item random intercepts to
3.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-margdist3-1.pdf}

}

\end{figure}%

\subsubsection{Iterative process with domain
experts}\label{iterative-process-with-domain-experts}

\paragraph{THEORY.}\label{theory-5}

In our experience, the gathering of domain knowledge by domain experts
and the consecutive specification of population parameter values used in
data simulation is not a one-time event but rather an iterative process.
In a first step, domain experts can be interviewed to ``elicit'' their
domain knowledge about how the future data of the planned study is
expected to look like. As most domain experts are no experts in
statistical modeling and GLMMs, they often struggle without further
guidance to communicate their knowledge in a way that is useful when
specifying the parameters for data simulation. For this reason, we
suggest that after an initial unstructured interview of domain experts,
the analyst who is familiar with the structure of the GLMM under study
selects an initial set of insightful descriptive statistics and model
based quantities. Then they reenter into an iterative discussion where
some set of population values are selected and the plausibility of
resulting implied quantities are discussed with the domain experts.
Then, the population parameters are updated based on this discussion
until the domain experts are satisfied with the final result. During
this process, the monitored model based quantities and descriptive
statistics can be updated or extended to capture as much available
domain knowledge as possible.

\subparagraph{PRACTICE.}\label{practice-5}

All parameter values in our present case study have been determined
based on repeated discussions with our affiliated domain experts in
Radiology to validate our assumptions. Initially, we reviewed the
literature to establish a reasonable baseline performance rate for
examining head CT scans for intracranial hemorrhage. Existing studies
indicate that radiologists typically demonstrate high accuracies, often
exceeding or hovering around 90\%, while interns have been shown to
perform below 80\%, and medical students fall even shorter. For
simplicity, we assumed plausible probability values of .90 for experts
and .70 for non-experts, respectively. Our experts confirmed that these
values are realistic baselines for reviewing diverse head CT images
without AI assistance. Subsequently, we consulted several published
papers investigating the effect of correct and incorrect advice on
decision-making performance in other settings. From their findings, we
inferred that both experts and non-experts should benefit from correct
and suffer losses from incorrect advice. However, the magnitude of these
effects should be substantially greater for non-experts, given their
demonstrated reliance on advice compared to experts. We further
validated the plausibility of our estimated gains and losses with the
collaborating radiologists. For our simulation, we used the
probabilities of average participants to solve an average case, as shown
in Table 1.

\subsection{Step 4: Estimate the statistical
model}\label{step-4-estimate-the-statistical-model}

\subsubsection{THEORY}\label{theory-6}

At this point, the researcher is capable of producing a simulated
dataset similar to the actual dataset that will later be collected in
the planned study. The next step is to specify how the statistical model
shall be estimated in the actual study collected later. This usually
includes the selection of 1) a statistical framework (e.g., frequentist
statistics), 2) a software package that is capable of estimating the
model class of interest (e.g., the lme4 R package), 3) an estimation
algorithm (e.g., the default optimizer ``bobyqa''), and 4) the specific
model structure including all fixed effects, random effects, and the
model family of the dependent variable.\\
Note that this does not always mean that one will specify the same GLMM
that was used when specifying the data generating process. On the one
hand, using a simpler model for data simulation than for model
estimation can be a useful strategy in scenarios where making plausible
assumptions for complicated random effect structures and interactions is
not feasible. On the other hand, using a more complex model for data
simulation than for model estimation can be a useful strategy in
scenarios where one has specific domain knowledge about aspects of the
data generating process that are still difficult to estimate with the
current state-of-the-art in multilevel modeling.

\subsubsection{PRACTICE}\label{practice-6}

In our case study, we use the lme4 R package
(\citeproc{ref-batesFittingLinearMixedEffects2015}{Bates et al., 2015}),
which is a state-of-the-art tool for fitting frequentist
GLMMs.\footnote{For Bayesian GLMMs, the brms R package is currently the
  most prominent option
  (\citeproc{ref-burknerBrmsPackageBayesian2017}{Bürkner, 2017}).} For
the current example, we simulate data according to our model, in which
100 subjects respond to 50 items (we use \texttt{set.seed} to make the
simulation reproducible). However, for the sake of the exercise, we can
imagine that this would be real data resulting from our future
experiment and think about how we would analyze this data.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(}\AttributeTok{n\_subjects =} \DecValTok{100}\NormalTok{, }\AttributeTok{n\_items =} \DecValTok{50}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The lme4 package uses a special syntax for model specification. Our
specific GLMM is represented by the formula:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(lme4)}
\NormalTok{f }\OtherTok{\textless{}{-}}\NormalTok{ y\_bin }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ expert }\SpecialCharTok{+}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ advice\_correct }\SpecialCharTok{+} 
\NormalTok{  expert}\SpecialCharTok{:}\NormalTok{advice\_present }\SpecialCharTok{+}\NormalTok{ expert}\SpecialCharTok{:}\NormalTok{advice\_correct }\SpecialCharTok{+}
\NormalTok{  (}\DecValTok{1}\SpecialCharTok{|}\NormalTok{subject) }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{|}\NormalTok{item)}
\end{Highlighting}
\end{Shaded}

The first two lines look similar to any linear model in R (general
intercept indicated by \texttt{1}; main effects indicated by variable
names in the dataset; interactions indicated by
\texttt{variable1:variable2}). The third line specifies a random
intercept for each subject \texttt{(1\textbar{}subject)} and for each
item \texttt{(1\textbar{}item)}. The complete set of rules for the
syntax is outlined in Bates et al.
(\citeproc{ref-batesFittingLinearMixedEffects2015}{2015}) and in the
documentation of the lme4 package.

In lme4, a GLMM is fitted with the \texttt{glmer} function. By setting
\texttt{family\ =\ \ "binomial"}, we request a binomial GLMM appropriate
for our binary dependent variable \texttt{y\_bin} (the binomial GLMM
uses the canonical logit link by default), which is defined as an
accurate (1) vs.~inaccurate (0) diagnosis.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(f, }\AttributeTok{data =}\NormalTok{ dat, }\AttributeTok{family =} \StringTok{"binomial"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We can inspect the estimates for all model parameters with the
\texttt{summary} command:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{summary}\NormalTok{(fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: 
y_bin ~ 1 + expert + advice_present + advice_correct + expert:advice_present +  
    expert:advice_correct + (1 | subject) + (1 | item)
   Data: dat

     AIC      BIC   logLik deviance df.resid 
  4149.4   4201.6  -2066.7   4133.4     4992

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-5.7669  0.2125  0.3046  0.4317  2.1056

Random effects:
 Groups  Name        Variance Std.Dev.
 subject (Intercept) 0.3148   0.5611  
 item    (Intercept) 0.1624   0.4029  
Number of obs: 5000, groups:  subject, 100; item, 50

Fixed effects:
                      Estimate Std. Error z value Pr(>|z|)    
(Intercept)             1.0339     0.1103   9.374  < 2e-16 ***
expert                  1.1849     0.2096   5.654 1.56e-08 ***
advice_present         -1.3436     0.1206 -11.143  < 2e-16 ***
advice_correct          2.6154     0.1273  20.540  < 2e-16 ***
expert:advice_present   1.0589     0.2940   3.601 0.000317 ***
expert:advice_correct  -1.8104     0.2915  -6.210 5.29e-10 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Correlation of Fixed Effects:
            (Intr) expert advc_p advc_c exprt:dvc_p
expert      -0.377                                 
advic_prsnt -0.349  0.176                          
advic_crrct  0.023  0.001 -0.668                   
exprt:dvc_p  0.143 -0.448 -0.412  0.276            
exprt:dvc_c -0.008  0.004  0.292 -0.435 -0.686     
\end{verbatim}

In the lme4 output, the \texttt{Estimate} column in the
\texttt{Fixed\ effects} table contains the estimates for the \(\beta\)
parameters, while the \texttt{Std.Dev.} column in the
\texttt{Random\ effects} table contains the estimates for \(\sigma_S\)
and \(\sigma_I\).

\subsection{Step 5: Compute the
estimate}\label{step-5-compute-the-estimate}

\subsubsection{THEORY}\label{theory-7}

In previous steps, we have defined the theoretical estimand, written a
data simulation function and specified how to estimate a GLMM using
simulated data. The next step is to specify how to compute a concrete
point estimate of the theoretical estimand within the framework of the
fitted GLMM. For some research question, the estimate corresponds with a
single regression coefficient. In more complicated scenarios, the
estimate is computed from a combination of coefficients. Beyond
computing the point estimate, we have already discussed that both
hypothesis testing and interval estimation can be used to answer the
research question. The decision on testing or estimating is then
followed by selecting the specific statistical method that shall be
applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with
the marginaleffects R package using the delta method).

\subsubsection{PRACTICE}\label{practice-7}

In the estimand section, we have translated a verbal description of our
research question into four probability statements that are specified
outside of any specific statistical model. For a concrete estimate
within the context of our specified GLMM, we must compute the following
probability contrasts:

\[
\begin{aligned}
& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\] \[
\begin{aligned}
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\]

\[
\begin{aligned}
& P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\]

\[
\begin{aligned}
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
& \quad - P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\] We have already discussed how to compute the involved probabilities
in the section on specifying population parameters. Plugging in the
model equation of the GLMM produces an equation on how to compute each
contrast if all model parameters were known. When we want to estimate
the above contrasts based on \emph{observed} data, the only difference
is that model parameters are not known and we instead use the
corresponding parameter \emph{estimates}.

We could use our knowledge of the structure of the GLMM to determine the
exact formula needed to compute the contrasts of interest and then plug
in the parameter estimates manually from the \texttt{summary(fit)}
output. However, this would be tedious and we can use R to compute this
contrast without doing the math. Using the first contrast \emph{(correct
advice, expert) - (no advice, expert)} as our example, we could apply
the \texttt{predict} function of the lme4 package to compute the
predicted probability for a correct diagnosis based on our fitted model,
plug in the two sets of predictor values, and compute the difference
between the two estimated probabilities.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{grid1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{advice\_present =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{advice\_correct =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }
  \AttributeTok{expert =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\NormalTok{grid1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
  advice_present advice_correct expert
1              1              1      1
2              0              0      1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pred }\OtherTok{\textless{}{-}} \FunctionTok{predict}\NormalTok{(fit, }\AttributeTok{newdata =}\NormalTok{ grid1, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{)}
\NormalTok{pred}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
       1        2 
0.939292 0.901923 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pred[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ pred[}\DecValTok{2}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
         1 
0.03736901 
\end{verbatim}

The argument \texttt{type\ =\ "response"} specifies that predictions are
made on the probability scale (instead of the log-odds scale of the
\(\beta\) parameters), while \texttt{re.form\ =\ NA} sets all random
effects to 0. We could use this method to compute point estimates for
all four contrasts that are part of our estimand. However, depending on
whether we are interested in hypothesis testing or parameter estimation,
we also need a method to compute HTs or CI. The marginaleffects package
(\citeproc{ref-R-marginaleffects}{Arel-Bundock, 2024}) is a very
flexible, increasingly popular package to compute HTs and CIs for
contrasts with a variety of statistical models, including GLMMs
estimated with lme4. First, we specify a grid of all combinations of
predictor variable and then compute estimated probabilities for all
experimental conditions in our experiment with the \texttt{predictions}
function:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(marginaleffects)}
\FunctionTok{library}\NormalTok{(tinytable)}
\NormalTok{grid2 }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}\AttributeTok{advice\_present =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }
  \AttributeTok{advice\_correct =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }\AttributeTok{expert =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{)}
\NormalTok{grid2}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 8 x 3
  advice_present advice_correct expert
           <int>          <int>  <int>
1              0              0      0
2              0              0      1
3              0              1      0
4              0              1      1
5              1              0      0
6              1              0      1
7              1              1      0
8              1              1      1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{preds }\OtherTok{\textless{}{-}} \FunctionTok{predictions}\NormalTok{(fit, }\AttributeTok{newdata =}\NormalTok{ grid2, }
  \AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{)}
\FunctionTok{print}\NormalTok{(preds, }\AttributeTok{style =} \StringTok{"tinytable"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{theme\_tt}\NormalTok{(}\AttributeTok{theme =} \StringTok{"resize"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\resizebox{\ifdim\width>\linewidth 1\linewidth\else\width\fi}{!}{
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={Type:  response 
},
note{ }={Columns: rowid, estimate, std.error, statistic, p.value, s.value, conf.low, conf.high, advice\_present, advice\_correct, expert, y\_bin 
},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]},
}                     %% tabularray inner close
\toprule
Estimate & Std. Error & z & Pr(>|z|) & S & 2.5 \% & 97.5 \% & advice\_present & advice\_correct & expert \\ \midrule %% TinyTableHeader
0.738 & 0.02134 &  34.6 & <0.001 & 867.2 & 0.696 & 0.779 & 0 & 0 & 0 \\
0.902 & 0.01739 &  51.9 & <0.001 & Inf & 0.868 & 0.936 & 0 & 0 & 1 \\
0.975 & 0.00421 & 231.6 & <0.001 & Inf & 0.966 & 0.983 & 0 & 1 & 0 \\
0.954 & 0.01454 &  65.6 & <0.001 & Inf & 0.925 & 0.982 & 0 & 1 & 1 \\
0.423 & 0.03221 &  13.1 & <0.001 & 128.6 & 0.360 & 0.486 & 1 & 0 & 0 \\
0.874 & 0.02793 &  31.3 & <0.001 & 711.4 & 0.819 & 0.928 & 1 & 0 & 1 \\
0.909 & 0.00967 &  94.1 & <0.001 & Inf & 0.890 & 0.928 & 1 & 1 & 0 \\
0.939 & 0.01091 &  86.1 & <0.001 & Inf & 0.918 & 0.961 & 1 & 1 & 1 \\
\bottomrule
\end{talltblr}
}
\end{table}

The point estimates for all experimental conditions are reported in the
\texttt{Estimate} column. Note that the output also contains the two
missing by design conditions that will never be observed in the actual
study (\(advice\_present = 0, advice\_correct = 1, expert = 1\) and
\(advice\_present = 0, advice\_correct = 1, expert = 0\)). This is no
problem as long as we never interpret those estimates. Next, we use the
estimated probabilities to compute the four specific contrasts that are
part of our estimand. For this we must specify which rows in
\texttt{probs} have to be subtracted from each other. We will use the
\texttt{hypotheses} function to compute our four contrasts of interest
together with HTs and CIs. We use the default inference options of the
marginaleffects package that compute HTs and CIs based on the
approximate delta method.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{contrasts }\OtherTok{\textless{}{-}}\NormalTok{ preds }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =} \FunctionTok{c}\NormalTok{(}
    \StringTok{"b8 = b2"}\NormalTok{,  }\CommentTok{\# (correct advice, expert) {-} (no advice, expert)}
    \StringTok{"b2 = b6"}\NormalTok{,  }\CommentTok{\# (no advice, expert) {-} (wrong advice, expert) }
    \StringTok{"b7 = b1"}\NormalTok{,  }\CommentTok{\# (correct advice, no expert) {-} (no advice, no expert)}
    \StringTok{"b1 = b5"}\NormalTok{), }\CommentTok{\# (no advice, no expert) {-} (wrong advice, no expert)}
    \AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{))}
\FunctionTok{print}\NormalTok{(contrasts, }\AttributeTok{style =} \StringTok{"tinytable"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{theme\_tt}\NormalTok{(}\AttributeTok{theme =} \StringTok{"resize"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\resizebox{\ifdim\width>\linewidth 1\linewidth\else\width\fi}{!}{
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={Type:  response 
},
note{ }={Columns: term, estimate, std.error, statistic, p.value, s.value, conf.low, conf.high, statistic.noninf, statistic.nonsup, p.value.noninf, p.value.nonsup, p.value.equiv 
},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]},
}                     %% tabularray inner close
\toprule
Term & Estimate & Std. Error & z & Pr(>|z|) & S & 2.5 \% & 97.5 \% & p (NonSup) & p (NonInf) & p (Equiv) \\ \midrule %% TinyTableHeader
b8=b2 & 0.0374 & 0.0162 &  2.31 & 0.021 & 5.6 &  0.00563 & 0.0691 & 0.989 & 0.0105 & 0.989 \\
b2=b6 & 0.0282 & 0.0279 &  1.01 & 0.312 & 1.7 & -0.02653 & 0.0830 & 0.844 & 0.1562 & 0.844 \\
b7=b1 & 0.1717 & 0.0173 &  9.93 & <0.001 & 74.8 &  0.13780 & 0.2056 & 1.000 & <0.001 & 1.000 \\
b1=b5 & 0.3145 & 0.0280 & 11.24 & <0.001 & 95.0 &  0.25965 & 0.3693 & 1.000 & <0.001 & 1.000 \\
\bottomrule
\end{talltblr}
}
\end{table}

The expression \texttt{"b8\ =\ b2"} is special syntax to subtract the
estimate in row number 8 from the estimate in row number 2 in the
\texttt{probs}-output. The argument \texttt{equivalence\ =\ c(0,\ 0)}
can be used to compute one-sided p-values, testing whether the contrast
in the population is smaller than 0 (\texttt{p\ (NonSub)} column) or
greater than 0 (\texttt{p\ (NonInf)} column). The point estimates for
four contrasts are reported in the \texttt{Estimate} column. Note that
to facilitate interpretation, we arranged the contrasts in a way that we
theoretically expect positive values for all four of them.

\paragraph{Hypothesis testing.}\label{hypothesis-testing}

If we chose hypothesis testing for our case study, we would test a
combined null hypothesis \(H_0\) that consists of four separate null
hypotheses: \[
\begin{aligned}
H_{01}:\ & P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
H_{02}:\ &P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 1, u_{0s} = 0, u_{0i} = 0) \\
H_{03}:\ &P(Y=1|advice\_present = 1, advice\_correct = 1, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \\
H_{04}:\ & P(Y=1|advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) \leq \\
& P(Y=1|advice\_present = 1, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)
\end{aligned}
\] The combined null hypothesis \(H_0\) should only be rejected if
\textbf{all} individual null hypotheses are rejected (i.e.,
intersection-union setting;
\citeproc{ref-dmitrienkoTraditionalMultiplicityAdjustment2013}{Dmitrienko
\& D'Agostino, 2013}). In such cases, the error probabilities do not
accumulate, and we would waste power when correcting for multiple tests.

With a standard significance level of \(\alpha = 0.05\), we would not
reject all four null hypotheses (the p-value in the \texttt{p\ (NonInf)}
column for the second hypothesis is not significant) and therefore also
not reject the combined null hypothesis for this particular (simulated)
dataset. Note that this decision would be wrong because we have
simulated the data such that the combined alternative hypothesis \(H_1\)
is actually true in the population.

\paragraph{Interval estimation.}\label{interval-estimation}

If we chose parameter estimation for our case study, we would focus on
the two-sided CIs of the four contrasts of interest. With a standard
confidence level of \(1 - \alpha = 0.95\), plausible values are clearly
in the positive range for the first, third and fourth contrast, while
both negative and positive values seem plausible for the second
contrast. Note that due to the constrained range of the probability
scale, the width of the CI differs between the four contrasts (which is
the expected behavior for binomial GLMMs). The smallest width is
observed for the first contrast (expert with correct advice vs.~expert
without advice) where both underlying probabilities are close to 1. The
largest width is observed for the fourth contrast (non-expert with wrong
advice vs.~non-expert without advice), where both underlying
probabilities are closer to 0.5.

\subsection{Step 6: Perform repeated
simulations}\label{step-6-perform-repeated-simulations}

\subsubsection{THEORY}\label{theory-8}

Conducting all previous steps enables the analyst to 1) simulate a
dataset, 2) estimate a GLMM, and 3) compute HTs or CIs for estimands of
interest, mirroring the analysis that will later be performed for the
actual dataset of the planned study. The last missing piece is to write
code to perform the above steps repeatedly and allow for a setting using
different sample sizes. On a conceptual level, we first require a
function that takes as input the sample size and the full set of
population parameter values. When planning for power, the function
should return the p-value(s) of the HT(s) of interest when conducted on
the simulated dataset. When planning for precision, the function should
return the width of the CI(s) of interest. Secondly, we must run this
function repeatedly with the same sample size and population parameters.
Because even fitting GLMMs with frequentist methods can quickly become
time-consuming, it is recommended to use parallel computing, that is
running simulations on multiple cores of the computer at the same time
to reduce total run time. Thirdly, the results of the repeated
simulation must be collected and aggregated. When planning for power, we
compute the relative frequency of (a) significant p-value(s) across
repeated simulations. When planning for precision, we compute the
average width of the CI(s). Lastly, we have to repeat the complete
simulation for different sample sizes, to determine how big the sample
must be in order to achieve the targeted power or precision.

\subsubsection{PRACTICE}\label{practice-8}

We are finally ready to run our simulation-based sample size planning
analyses to plan for power and for precision. Wrapping the
\texttt{simulate} function already constructed earlier, the helper
function \texttt{sim\_and\_analyse} performs all previous steps
(simulate a dataset, fit a GLMM, compute p-values and CIs) in a single
command.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sim\_and\_analyse }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}
  \AttributeTok{formula\_chr =} \StringTok{"y\_bin \textasciitilde{} 1 + expert + advice\_present + advice\_correct + }
\StringTok{    expert:advice\_present + expert:advice\_correct + (1|subject) + (1|item)"}\NormalTok{,}
  \AttributeTok{contrasts =} \FunctionTok{c}\NormalTok{(}\StringTok{"b8 = b2"}\NormalTok{, }\StringTok{"b2 = b6"}\NormalTok{, }\StringTok{"b7 = b1"}\NormalTok{, }\StringTok{"b1 = b5"}\NormalTok{), ...)\{}
  \FunctionTok{require}\NormalTok{(lme4)}
  \FunctionTok{require}\NormalTok{(marginaleffects)}
  \FunctionTok{require}\NormalTok{(tidyr)}
  \CommentTok{\# simulate data}
\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{simulate}\NormalTok{(...)}
  \CommentTok{\# fit model}
\NormalTok{  model }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(}\FunctionTok{as.formula}\NormalTok{(formula\_chr), }\AttributeTok{data =}\NormalTok{ dat, }\AttributeTok{family =} \StringTok{"binomial"}\NormalTok{)}
  \CommentTok{\# compute contrasts}
\NormalTok{  contr\_df }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}\AttributeTok{advice\_present =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }\AttributeTok{advice\_correct =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{,}
    \AttributeTok{expert =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{)}
  \FunctionTok{predictions}\NormalTok{(model, }\AttributeTok{newdata =}\NormalTok{ contr\_df, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =}\NormalTok{ contrasts, }\AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{data.frame}\NormalTok{()}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Simulation-based sample size planning can quickly become computationally
intensive when we repeatedly simulate data and fit models for different
parameter combinations or sample sizes. Thus, we use the future
(\citeproc{ref-R-RJ-2021-048}{Bengtsson, 2021}) and furrr
(\citeproc{ref-R-furrr}{Vaughan \& Dancho, 2022}) packages to perform
computations in parallel. First, we enable parallelization with the
\texttt{plan} function and specify how many parallel cores (``workers'')
of our computer to use (users can find out the maximum number of cores
on their computer with the command \texttt{parallel::detectCores()}),
and set a seed to make the simulation reproducible.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(future)}
\FunctionTok{plan}\NormalTok{(}\StringTok{"multisession"}\NormalTok{, }\AttributeTok{workers =} \DecValTok{6}\NormalTok{)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The next code chunk specifies a simulation grid with different settings
for both the number of subjects (\texttt{n\_subjects}) and the number of
items (\texttt{n\_items}), each combination being repeated \texttt{rep}
times. We chose 300 repetitions for the data simulation at hand as it
strikes a balance between achieving a robust statistical estimate and
remaining computationally feasible. With the current settings, this
simulation takes about one hour on a MacBook Pro from 2020 with M1 chip
and 16 GB working memory. If you want to quickly experiment with the
code yourself, a setting with \texttt{workers\ =\ 4} and
\texttt{rep\ =\ 5} should finish in less than 5 minutes, even on smaller
machines.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(furrr)}
\NormalTok{sim\_result }\OtherTok{\textless{}{-}} \FunctionTok{crossing}\NormalTok{(}
  \AttributeTok{rep =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{300}\NormalTok{,}
  \AttributeTok{n\_subjects =} \FunctionTok{c}\NormalTok{(}\DecValTok{100}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{250}\NormalTok{),}
  \AttributeTok{n\_items =} \FunctionTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{50}\NormalTok{, }\DecValTok{70}\NormalTok{)}
\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{res =} \FunctionTok{future\_pmap}\NormalTok{(., sim\_and\_analyse, }
    \AttributeTok{.options =} \FunctionTok{furrr\_options}\NormalTok{(}\AttributeTok{seed =} \ConstantTok{TRUE}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{unnest}\NormalTok{(}\AttributeTok{col =}\NormalTok{ res)}
\end{Highlighting}
\end{Shaded}

The result of this computation is a data frame that contains the
p-values and CIs of all specified contrasts for each simulated dataset.
In some iterations (predominantly in conditions with small sample
sizes), model estimation did not converge with the lme4 package. When
the model fails to converge, it means that the statistical model being
fitted to the data failed to reach a stable or valid solution during the
estimation process. We do not remove these results because
non-convergence can also happen when analyzing the real data we plan to
collect, thus, we want to factor in this possibility to keep our
simulation more realistic.

\paragraph{Power results.}\label{power-results}

For our exemplary combined hypothesis, power is defined as the
(long-run) percentage of simulations in which all four p-values of our
individual hypotheses are significant at the \(\alpha = 0.05\) level.
Based on our simulation outcomes, we compute a power estimate for each
combination of \texttt{n\_subjects} \(\times\) \texttt{n\_items}
(including 95\% CIs) and visualize the results with the following
code.\footnote{This code was inspired by the ``Mixed Design Simulation''
  vignette of the faux package at
  \url{https://debruine.github.io/faux/articles/sim_mixed.html}.}

\phantomsection\label{cell-fig-finalpwr}
\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(binom)}
\NormalTok{alpha }\OtherTok{\textless{}{-}} \FloatTok{0.05}
\NormalTok{power }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ term, }\AttributeTok{names\_sep =} \StringTok{"\_"}\NormalTok{, }
    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{power =} \FunctionTok{mean}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b1=b5}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b2=b6}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b7=b1}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha), }
    \AttributeTok{n\_sig =} \FunctionTok{sum}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b1=b5}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b2=b6}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha }\SpecialCharTok{\&} 
        \StringTok{\textasciigrave{}}\AttributeTok{p.value.noninf\_b7=b1}\StringTok{\textasciigrave{}} \SpecialCharTok{\textless{}}\NormalTok{ alpha),}
    \AttributeTok{n =} \FunctionTok{n}\NormalTok{(),}
    \AttributeTok{ci.lwr =} \FunctionTok{binom.confint}\NormalTok{(n\_sig, n, }\AttributeTok{method =} \StringTok{"wilson"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{lower,}
    \AttributeTok{ci.upr =} \FunctionTok{binom.confint}\NormalTok{(n\_sig, n, }\AttributeTok{method =} \StringTok{"wilson"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{upper, }
    \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{)}
\NormalTok{power }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(n\_subjects, n\_items, }\AttributeTok{fill =}\NormalTok{ power)) }\SpecialCharTok{+}
  \FunctionTok{geom\_tile}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{label =} \FunctionTok{sprintf}\NormalTok{(}\StringTok{"\%.2f }\SpecialCharTok{\textbackslash{}n}\StringTok{ [\%.2f; \%.2f]"}\NormalTok{, }
\NormalTok{                                power, ci.lwr, ci.upr)), }
    \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \DecValTok{4}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{xlab}\NormalTok{(}\StringTok{"number of subjects"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{ylab}\NormalTok{(}\StringTok{"number of items"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{\label{fig-finalpwr}Simulation-based power estimates including
95\% confidence interval of the case study for different numbers of
subjects and items, based on a significance level of 0.05.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-finalpwr-1.pdf}

}

\end{figure}%

As should be the case, power estimates in Figure~\ref{fig-finalpwr}
increase with both the number of subjects and the number of items. The
CIs reported here indicate how precisely power was estimated by our
simulation. Higher precision (which would be reflected in narrower CIs)
could be obtained by increasing the number of repetitions (\texttt{rep})
in the simulation. In practice, data simulations are often run multiple
times with adjusted combinations of sample sizes. When running for the
first time, it might be revealed that power is way too low (or much
higher than required) for some combinations of \texttt{n\_subjects} and
\texttt{n\_items}. When narrowing down the best combination that
achieves sufficient power while at the same time striking a good balance
of how many subjects and items are practically feasible, later rounds of
data simulation will typically include a smaller grid of sample sizes
combined with a higher number of repetitions. This will assure high
precision for the final power estimates, which are then used for the
sample size justification of the future study.

Much has been written on the optimal amount of power to target in
empirical research. The most prominent heuristic is to target a power of
0.8 (when combined with a type I error rate of \(\alpha = 0.05\)), but
depending on the research goals of the study, there are often good
reasons to move away from this standard depending on the research goals
and resource constraints
(\citeproc{ref-lakensJustifyYourAlpha2018}{Lakens, Adolfi, et al.,
2018}; \citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).
When target power has been specified, the number of subjects and the
number of items in our study design can be traded against each other
based on practical considerations. For the sake of the example, let the
targeted power be indeed about 0.8, using an \(\alpha\) of 0.05 to
detect an effect of the expected size implied by our data simulation.
This could be achieved by collecting data from 200 subjects (about 25\%
of which will be experts), each completing the same 50 items (with
advice present in about 67\% of cases, which is correct in about 80\% of
cases with present advice). If collecting data from 200 subjects is not
feasible, an alternative would be to recruit 150 subjects but increase
the length of the experiment to over 70 items. However, 70 items might
take too long to complete for the radiologists participating in the
study, who have a busy schedule. The simulation suggests that it might
also be possible to plan a shorter experiment with only 30 items if it
is feasible to recruit an even higher number of subjects (\textgreater{}
250, to be determined by additional rounds of power analysis). Design
parameters that also affect power, and which could be investigated in
the simulation to find a more optimal trade-off, are the ratio of
experts, the frequency of whether advice is presented and whether it is
correct.

\paragraph{Precision results.}\label{precision-results}

When planning for precision, one could monitor the width of all four CIs
at the same time. However, because the CIs of the four contrasts
strongly differ in width, it is not trivial to decide which width one
should target when deciding on the appropriate sample size. In contrast
to planning for power, there are no common standards on how to specify
the targeted precision. For our example, we use a simple heuristic but
we strongly encourage readers to think about better alternatives that
are appropriate in their own applications. Our simulations show that the
smallest CI can be expected for the first contrast (expert with correct
advice vs.~expert without advice). The true contrast in probability for
an average expert and an average item in this condition is
\texttt{plogis(b\_0\ +\ b\_e\ +\ b\_a\ +\ b\_c\ +\ b\_ea\ +\ b\_ec)\ -\ plogis(b\_0\ +\ b\_e)\ =}
\(0.05\). We want the width of this CI to be smaller than 0.1. This
would mean that if the point estimate happens to be close to the true
value, the plausible values inside of a 95\% CI would all be positive.

Thus in our example, precision is defined as the (long-run) average
width of a 95\% CI for the probability contrast between experts with
correct advice and experts without advice. Of course, lower width
implies better precision. Based on our simulation outcomes, we compute
the precision estimate for each combination of \texttt{n\_subjects}
\(\times\) \texttt{n\_items} (including 95\% CIs) and visualize the
results with the following code.

\phantomsection\label{cell-fig-finalprecision}
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{precision }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ term, }\AttributeTok{names\_sep =} \StringTok{"\_"}\NormalTok{, }
    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{width =} \StringTok{\textasciigrave{}}\AttributeTok{conf.high\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{{-}} \StringTok{\textasciigrave{}}\AttributeTok{conf.low\_b8=b2}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{precision =} \FunctionTok{mean}\NormalTok{(width),}
    \AttributeTok{ci.lwr =} \FunctionTok{t.test}\NormalTok{(width)}\SpecialCharTok{$}\NormalTok{conf.int[}\DecValTok{1}\NormalTok{],}
    \AttributeTok{ci.upr =} \FunctionTok{t.test}\NormalTok{(width)}\SpecialCharTok{$}\NormalTok{conf.int[}\DecValTok{2}\NormalTok{], }
    \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{)}
\NormalTok{precision }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(n\_subjects, n\_items, }\AttributeTok{fill =}\NormalTok{ precision)) }\SpecialCharTok{+}
  \FunctionTok{geom\_tile}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{label =} \FunctionTok{sprintf}\NormalTok{(}\StringTok{"\%.2f }\SpecialCharTok{\textbackslash{}n}\StringTok{ [\%.2f; \%.2f]"}\NormalTok{, }
\NormalTok{                                precision, ci.lwr, ci.upr)), }
    \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \DecValTok{4}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\AttributeTok{direction =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{guides}\NormalTok{(}\AttributeTok{fill =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{reverse=}\ConstantTok{FALSE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{\label{fig-finalprecision}Simulation-based precision estimates
(expected width of confidence intervals) including 95\% confidence
interval of the case study for different numbers of subjects and items,
based on a confidence level of 0.95.}

\centering{

\includegraphics{manuscript_files/figure-pdf/fig-finalprecision-1.pdf}

}

\end{figure}%

As should be the case, precision estimates in
Figure~\ref{fig-finalprecision} increase (i.e., average width of CI
decreases) with both the number of subjects and the number of items. The
CIs reported here indicate how precisely the expected width of the CI
for our focal contrast was estimated by our simulation. Applying our
simple heuristic of targeting an expected width smaller than 0.1, we see
the same trade-off between the number of subjects and the number of
items as with planning for power. We could either choose 100 subjects
and 30 items or 200 subjects and 10 items. Note that our simple
heuristic for determining sample size in the planning for precision
scenario was quite liberal. This is reflected by the result that we
would need a smaller sample size than in the planning for power
scenario. With a more conservative precision target, the result is
generally the opposite: As a rule, precise parameter estimates usually
require bigger samples than null hypothesis testing.

\subsection{Sensitivity analysis}\label{sensitivity-analysis}

In our case study, we have performed simulation-based sample size
planning from a single set of parameter values that reflect our
assumptions of an expected effect size. Instead of extracting this
expected effect size from meta-analyses or pilot data, which has been
the main focus of previous tutorials (e.g.,
\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}),
we have demonstrated some strategies to determine plausible parameter
values in GLMMs based on domain knowledge. Domain knowledge can be
considered a vague theoretical model about the data-generating process
that is less formal and can only be accessed by a back-and-forth
exchange in which domain experts assess the plausibility of simulated
data. When sample sizes are chosen based on the results of our
simulation-based power analysis, a future study will be informative to
reject the null hypothesis if an effect of our \emph{expected size} is
present (or estimate the effect with satisfying precision). However, if
the true effect is indeed smaller, power (or precision) will be lower,
and the study might not be sufficiently informative. A common, more
conservative strategy for sample size justification is to perform sample
size planning for the smallest effect size of interest (SESOI). An
effect smaller than the SESOI would be considered too small to be
interesting or practically meaningful, even if the effect is not
actually zero (\citeproc{ref-kingPointMinimalImportant2011}{King, 2011};
\citeproc{ref-lakensEquivalenceTestingPsychological2018}{Lakens, Scheel,
et al., 2018}). For strategies on the even more difficult task of
specifying a plausible SESOI, as well as a thorough discussion of
various topics concerning power analysis, see
(\citeproc{ref-lakensImprovingYourStatistical2022}{Lakens, 2022b}). When
domain knowledge or formal theories about the research topic of interest
are too vague to specify a meaningful SESOI, it is still recommended to
demonstrate power or precision for different effect sizes in what is
called \emph{sensitivity power analysis}. By simulating power (or
precision) for different effect sizes (in addition to the different
number of subjects and items), one can make sure that power (or
precision) would still be sufficient to detect smaller effect sizes than
our expected effect or at least get an impression of how strongly power
(or precision) depends on the size of the true effect. For our case
study that investigates combined hypotheses in a GLMM modeling
framework, sensitivity analysis would require manually specifying
additional sets of plausible parameter values that reflect scenarios
with smaller or larger differences between groups with respect to our
specific research question. Power (or precision) could then be simulated
for several of these scenarios (across different numbers of subjects and
items, as considered earlier).

\section{Conclusion and outlook}\label{conclusion-and-outlook}

The goal of this tutorial was to teach researchers how to perform
tailored simulation-based sample size planning for GLMMs. Beyond the
specifics of our concrete case study, we want to outline six
developments regarding the future role of simulation-based sample size
planning in experimental research:

In light of the ongoing replication crisis and an existing literature
full of underpowered studies, there is a growing need for
simulation-based sample size planning in experimental research: In order
to conduct informative research, GLMMs offer a flexible statistical
framework to analyze complex experimental study designs. However,
existing formula-based heuristics and user-friendly software tools for a
priori power analysis are often not sufficient. Therefore,
simulation-based power analysis is becoming increasingly needed since it
provides experimental researchers with a tailored approach to estimating
required sample sizes before data collection.

Managing data simulations more easily with discrete predictor variables:
Simulation-based sample size planning becomes more manageable when all
predictor variables are discrete (like in the presented case study) and
fixed by the study design. This allows researchers to focus on
simulating outcome variables while avoiding the need for complex
simulations of predictor values, which would introduce additional
assumptions. By simplifying the simulation process, researchers can
obtain reliable estimates for power or precision without compromising
realistic assumptions about the data-generating process implied by the
study design.

Teaching data simulation skills: The ability to conduct simulation-based
sample size planning is a valuable skill that should be taught to
experimental researchers. By incorporating such training into research
methods courses and workshops, researchers can gain a deeper
understanding of statistical power or precision, and improve the quality
of their experimental designs. Equipping researchers with the knowledge
and tools to perform simulation-based sample size planning enables them
to make informed decisions and enhance the rigor of their studies. The
need to reason about how to simulate plausible data that is in line with
the research hypothesis, while not violating domain expertise on how
plausible data should look, might also contribute to planning more
insightful studies that can answer more precise research questions
(\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}).

Addressing the mismatch in effort perception: There is often a
significant disconnect between the amount of effort required to perform
tailored simulation-based sample size planning and the perceived effort
estimated by researchers and collaborators in experimental research.
Many researchers request simulation-based power analyses from
statisticians or methodological experts without fully comprehending the
complexity and time-consuming nature of these tailored simulations. It
is crucial to raise awareness about the effort involved to ensure
realistic expectations and effective collaboration between researchers
and methodological experts.

Recognizing the value of simulation-based design analysis: Tailored data
simulations and power analyses are not mere technicalities; they are
valuable research contributions that deserve recognition in experimental
research. They offer insights into the robustness and sensitivity of
experimental designs, helping researchers make informed decisions about
sample sizes, effect sizes, and statistical power or precision. Their
importance can be reflected by allocating them a separate publication or
incorporating them as a significant component of stage 1 preregistered
reports (\citeproc{ref-chambersPresentFutureRegistered2022}{Chambers \&
Tzavella, 2022}).

Integration with Open Science and preregistration practices:
Simulation-based sample size planning aligns well with the principles of
Open Science and preregistration in experimental research. When
researchers have access to simulated data based on their pre-specified
model, analyzing the collected dataset becomes straightforward and
unambiguous. By preregistering their simulation-based sample size plan,
researchers enhance the transparency and accountability of their
experimental procedures, contributing to the credibility and
reproducibility of research.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-albersWhenPowerAnalyses2018a}
Albers, C., \& Lakens, D. (2018). When power analyses based on pilot
data are biased: {Inaccurate} effect size estimators and follow-up bias.
\emph{Journal of Experimental Social Psychology}, \emph{74}, 187--195.
\url{https://doi.org/10.1016/j.jesp.2017.09.004}

\bibitem[\citeproctext]{ref-R-marginaleffects}
Arel-Bundock, V. (2024). \emph{Marginaleffects: Predictions,
comparisons, slopes, marginal means, and hypothesis tests}.
\url{https://marginaleffects.com/}

\bibitem[\citeproctext]{ref-arendStatisticalPowerTwolevel2019}
Arend, M. G., \& Schäfer, T. (2019). Statistical power in two-level
models: {A} tutorial based on {Monte Carlo} simulation.
\emph{Psychological Methods}, \emph{24}(1), 1--19.
\url{https://doi.org/10.1037/met0000195}

\bibitem[\citeproctext]{ref-batesFittingLinearMixedEffects2015}
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting
{Linear Mixed-Effects Models Using} {\textbf{Lme4}}. \emph{Journal of
Statistical Software}, \emph{67}(1).
\url{https://doi.org/10.18637/jss.v067.i01}

\bibitem[\citeproctext]{ref-R-RJ-2021-048}
Bengtsson, H. (2021). A unifying framework for parallel and distributed
processing in r using futures. \emph{The R Journal}, \emph{13}(2),
208--227. \url{https://doi.org/10.32614/RJ-2021-048}

\bibitem[\citeproctext]{ref-brooksGlmmTMBBalancesSpeed2017}
Brooks, M., E., Kristensen, K., Benthem, van, J., Magnusson, A., Berg,
C., W., Nielsen, A., Skaug, H., J., Mächler, M., \& Bolker, B., M.
(2017). {glmmTMB Balances Speed} and {Flexibility Among Packages} for
{Zero-inflated Generalized Linear Mixed Modeling}. \emph{The R Journal},
\emph{9}(2), 378. \url{https://doi.org/10.32614/RJ-2017-066}

\bibitem[\citeproctext]{ref-brysbaertPowerAnalysisEffect2018}
Brysbaert, M., \& Stevens, M. (2018). Power {Analysis} and {Effect Size}
in {Mixed Effects Models}: {A Tutorial}. \emph{Journal of Cognition},
\emph{1}(1), 9. \url{https://doi.org/10.5334/joc.10}

\bibitem[\citeproctext]{ref-burknerBrmsPackageBayesian2017}
Bürkner, P.-C. (2017). Brms: {An R Package} for {Bayesian Multilevel
Models Using Stan}. \emph{Journal of Statistical Software}, \emph{80},
1--28. \url{https://doi.org/10.18637/jss.v080.i01}

\bibitem[\citeproctext]{ref-burknerAdvancedBayesianMultilevel2018}
Bürkner, P.-C. (2018). Advanced {Bayesian Multilevel Modeling} with the
{R Package} brms. \emph{The R Journal}, \emph{10}(1), 395.
\url{https://doi.org/10.32614/RJ-2018-017}

\bibitem[\citeproctext]{ref-chambersPresentFutureRegistered2022}
Chambers, C. D., \& Tzavella, L. (2022). The past, present and future of
{Registered Reports}. \emph{Nature Human Behaviour}, \emph{6}(1),
29--42. \url{https://doi.org/10.1038/s41562-021-01193-7}

\bibitem[\citeproctext]{ref-cohenPowerPrimer1992}
Cohen, J. (1992). A power primer. \emph{Psychological Bulletin},
\emph{112}(1), 155--159.
\url{https://doi.org/10.1037/0033-2909.112.1.155}

\bibitem[\citeproctext]{ref-cummingNewStatisticsWhy2014}
Cumming, G. (2014). The {New Statistics}: {Why} and {How}.
\emph{Psychological Science}, \emph{25}(1), 7--29.
\url{https://doi.org/10.1177/0956797613504966}

\bibitem[\citeproctext]{ref-R-faux}
DeBruine, L. (2023). \emph{Faux: Simulation for factorial designs}.
Zenodo. \url{https://doi.org/10.5281/zenodo.2669586}

\bibitem[\citeproctext]{ref-debruineUnderstandingMixedEffectsModels2021}
DeBruine, L., \& Barr, D. J. (2021). Understanding {Mixed-Effects Models
Through Data Simulation}. \emph{Advances in Methods and Practices in
Psychological Science}, \emph{4}(1), 2515245920965119.
\url{https://doi.org/10.1177/2515245920965119}

\bibitem[\citeproctext]{ref-deffnerCausalFrameworkCrossCultural2022}
Deffner, D., Rohrer, J. M., \& McElreath, R. (2022). A {Causal
Framework} for {Cross-Cultural Generalizability}. \emph{Advances in
Methods and Practices in Psychological Science}, \emph{5}(3),
251524592211063. \url{https://doi.org/10.1177/25152459221106366}

\bibitem[\citeproctext]{ref-dmitrienkoTraditionalMultiplicityAdjustment2013}
Dmitrienko, A., \& D'Agostino, R. (2013). Traditional multiplicity
adjustment methods in clinical trials. \emph{Statistics in Medicine},
\emph{32}(29), 5172--5218. \url{https://doi.org/10.1002/sim.5990}

\bibitem[\citeproctext]{ref-fahrmeirRegressionModelsMethods2021}
Fahrmeir, L., Kneib, T., Lang, S., \& Marx, B. D. (2021).
\emph{Regression: {Models}, {Methods} and {Applications}}. Springer
Berlin Heidelberg. \url{https://doi.org/10.1007/978-3-662-63882-8}

\bibitem[\citeproctext]{ref-gelmanBayesianWorkflow2020}
Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B.,
Yao, Y., Kennedy, L., Gabry, J., Bürkner, P.-C., \& Modrák, M. (2020).
\emph{Bayesian {Workflow}} (arXiv:2011.01808). arXiv.
\url{https://arxiv.org/abs/2011.01808}

\bibitem[\citeproctext]{ref-gomilaMissingDataExperiments2022}
Gomila, R., \& Clark, C. S. (2022). Missing data in experiments:
{Challenges} and solutions. \emph{Psychological Methods}, \emph{27}(2),
143--155. \url{https://doi.org/10.1037/met0000361}

\bibitem[\citeproctext]{ref-greenSIMRPackagePower2016a}
Green, P., \& MacLeod, C. J. (2016). {\textsc{SIMR}} : An {R} package
for power analysis of generalized linear mixed models by simulation.
\emph{Methods in Ecology and Evolution}, \emph{7}(4), 493--498.
\url{https://doi.org/10.1111/2041-210X.12504}

\bibitem[\citeproctext]{ref-kainPracticalGuidePower2015}
Kain, M. P., Bolker, B. M., \& McCoy, M. W. (2015). A practical guide
and power analysis for {GLMMs}: Detecting among treatment variation in
random effects. \emph{PeerJ}, \emph{3}, e1226.
\url{https://doi.org/10.7717/peerj.1226}

\bibitem[\citeproctext]{ref-kelleySampleSizePlanning2006}
Kelley, K., \& Rausch, J. R. (2006). Sample size planning for the
standardized mean difference: {Accuracy} in parameter estimation via
narrow confidence intervals. \emph{Psychological Methods}, \emph{11}(4),
363--385. \url{https://doi.org/10.1037/1082-989X.11.4.363}

\bibitem[\citeproctext]{ref-kingPointMinimalImportant2011}
King, M. T. (2011). A point of minimal important difference ({MID}): A
critique of terminology and methods. \emph{Expert Review of
Pharmacoeconomics \& Outcomes Research}, \emph{11}(2), 171--184.
\url{https://doi.org/10.1586/erp.11.9}

\bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
Kruschke, J. K., \& Liddell, T. M. (2018). The {Bayesian New
Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power
analysis from a {Bayesian} perspective. \emph{Psychonomic Bulletin \&
Review}, \emph{25}(1), 178--206.
\url{https://doi.org/10.3758/s13423-016-1221-4}

\bibitem[\citeproctext]{ref-kumleEstimatingPowerGeneralized2021}
Kumle, L., Võ, M. L.-H., \& Draschkow, D. (2021). Estimating power in
(generalized) linear mixed models: {An} open introduction and tutorial
in {R}. \emph{Behavior Research Methods}, \emph{53}(6), 2528--2543.
\url{https://doi.org/10.3758/s13428-021-01546-0}

\bibitem[\citeproctext]{ref-lafitSelectionNumberParticipants2021}
Lafit, G., Adolf, J. K., Dejonckheere, E., Myin-Germeys, I.,
Viechtbauer, W., \& Ceulemans, E. (2021). Selection of the {Number} of
{Participants} in {Intensive Longitudinal Studies}: {A User-Friendly
Shiny App} and {Tutorial} for {Performing Power Analysis} in {Multilevel
Regression Models That Account} for {Temporal Dependencies}.
\emph{Advances in Methods and Practices in Psychological Science},
\emph{4}(1), 251524592097873.
\url{https://doi.org/10.1177/2515245920978738}

\bibitem[\citeproctext]{ref-lakensSampleSizeJustification2022}
Lakens, D. (2022a). Sample {Size Justification}. \emph{Collabra:
Psychology}, \emph{8}(1), 33267.
\url{https://doi.org/10.1525/collabra.33267}

\bibitem[\citeproctext]{ref-lakensImprovingYourStatistical2022}
Lakens, D. (2022b). \emph{Improving {Your Statistical Inferences}}.
Zenodo. \url{https://doi.org/10.5281/ZENODO.6409077}

\bibitem[\citeproctext]{ref-lakensJustifyYourAlpha2018}
Lakens, D., Adolfi, F. G., Albers, C. J., Anvari, F., Apps, M. A. J.,
Argamon, S. E., Baguley, T., Becker, R. B., Benning, S. D., Bradford, D.
E., Buchanan, E. M., Caldwell, A. R., Van Calster, B., Carlsson, R.,
Chen, S.-C., Chung, B., Colling, L. J., Collins, G. S., Crook, Z.,
\ldots{} Zwaan, R. A. (2018). Justify your alpha. \emph{Nature Human
Behaviour}, \emph{2}(3), 168--171.
\url{https://doi.org/10.1038/s41562-018-0311-x}

\bibitem[\citeproctext]{ref-lakensEquivalenceTestingPsychological2018}
Lakens, D., Scheel, A. M., \& Isager, P. M. (2018). Equivalence
{Testing} for {Psychological Research}: {A Tutorial}. \emph{Advances in
Methods and Practices in Psychological Science}, \emph{1}(2), 259--269.
\url{https://doi.org/10.1177/2515245918770963}

\bibitem[\citeproctext]{ref-lanePowerStrugglesEstimating2018}
Lane, S. P., \& Hennes, E. P. (2018). Power struggles: {Estimating}
sample size for multilevel relationships research. \emph{Journal of
Social and Personal Relationships}, \emph{35}(1), 7--31.
\url{https://doi.org/10.1177/0265407517710342}

\bibitem[\citeproctext]{ref-littleStatisticalAnalysisMissing2014}
Little, R. J. A., \& Rubin, D. B. (2014). \emph{Statistical {Analysis}
with {Missing Data}} (2nd ed). Wiley.

\bibitem[\citeproctext]{ref-lundbergWhatYourEstimand2021}
Lundberg, I., Johnson, R., \& Stewart, B. M. (2021). What {Is Your
Estimand}? {Defining} the {Target Quantity Connects Statistical
Evidence} to {Theory}. \emph{American Sociological Review},
\emph{86}(3), 532--565. \url{https://doi.org/10.1177/00031224211004187}

\bibitem[\citeproctext]{ref-maxwellSampleSizePlanning2008}
Maxwell, S. E., Kelley, K., \& Rausch, J. R. (2008). Sample {Size
Planning} for {Statistical Power} and {Accuracy} in {Parameter
Estimation}. \emph{Annual Review of Psychology}, \emph{59}(1), 537--563.
\url{https://doi.org/10.1146/annurev.psych.59.103006.093735}

\bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2020}
McElreath, R. (2020). \emph{Statistical {Rethinking}: {A Bayesian
Course} with {Examples} in {R} and {Stan}} (2nd ed.). {Chapman and
Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}

\bibitem[\citeproctext]{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}
Murayama, K., Usami, S., \& Sakaki, M. (2022). Summary-statistics-based
power analysis: {A} new and practical method to determine sample size
for mixed-effects modeling. \emph{Psychological Methods}.
\url{https://doi.org/10.1037/met0000330}

\bibitem[\citeproctext]{ref-preacherComputationalToolsProbing2006}
Preacher, K. J., Curran, P. J., \& Bauer, D. J. (2006). Computational
{Tools} for {Probing Interactions} in {Multiple Linear Regression},
{Multilevel Modeling}, and {Latent Curve Analysis}. \emph{Journal of
Educational and Behavioral Statistics}, \emph{31}(4), 437--448.
\url{https://doi.org/10.3102/10769986031004437}

\bibitem[\citeproctext]{ref-riesthuisSimulationBasedPowerAnalyses2024}
Riesthuis, P. (2024). Simulation-{Based Power Analyses} for the
{Smallest Effect Size} of {Interest}: {A Confidence-Interval Approach}
for {Minimum-Effect} and {Equivalence Testing}. \emph{Advances in
Methods and Practices in Psychological Science}, \emph{7}(2),
25152459241240722. \url{https://doi.org/10.1177/25152459241240722}

\bibitem[\citeproctext]{ref-schadHowCapitalizePriori2020}
Schad, D. J., Vasishth, S., Hohenstein, S., \& Kliegl, R. (2020). How to
capitalize on a priori contrasts in linear (mixed) models: {A} tutorial.
\emph{Journal of Memory and Language}, \emph{110}, 104038.
\url{https://doi.org/10.1016/j.jml.2019.104038}

\bibitem[\citeproctext]{ref-uyguntuncEpistemicPragmaticFunction2023}
Uygun Tunç, D., Tunç, M. N., \& Lakens, D. (2023). The epistemic and
pragmatic function of dichotomous claims based on statistical hypothesis
tests. \emph{Theory \& Psychology}, \emph{33}(3), 403--423.
\url{https://doi.org/10.1177/09593543231160112}

\bibitem[\citeproctext]{ref-R-furrr}
Vaughan, D., \& Dancho, M. (2022). \emph{Furrr: Apply mapping functions
in parallel using futures}. \url{https://github.com/DavisVaughan/furrr}

\bibitem[\citeproctext]{ref-westfallStatisticalPowerOptimal2014}
Westfall, J., Kenny, D. A., \& Judd, C. M. (2014). Statistical power and
optimal design in experiments in which samples of participants respond
to samples of stimuli. \emph{Journal of Experimental Psychology:
General}, \emph{143}, 2020--2045.
\url{https://doi.org/10.1037/xge0000014}

\bibitem[\citeproctext]{ref-wickhamWelcomeTidyverse2019}
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L., François,
R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen,
T., Miller, E., Bache, S., Müller, K., Ooms, J., Robinson, D., Seidel,
D., Spinu, V., \ldots{} Yutani, H. (2019). Welcome to the {Tidyverse}.
\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
\url{https://doi.org/10.21105/joss.01686}

\bibitem[\citeproctext]{ref-wickhamDataScienceImport2023}
Wickham, H., Çetinkaya-Rundel, M., \& Grolemund, G. (2023). \emph{R for
data science: Import, tidy, transform, visualize, and model data} (2nd
edition). O'Reilly.

\bibitem[\citeproctext]{ref-yarkoniGeneralizabilityCrisis2022}
Yarkoni, T. (2022). The generalizability crisis. \emph{Behavioral and
Brain Sciences}, \emph{45}, e1.
\url{https://doi.org/10.1017/S0140525X20001685}

\bibitem[\citeproctext]{ref-zimmerSampleSizePlanning2022}
Zimmer, F., Henninger, M., \& Debelak, R. (2022). \emph{Sample {Size
Planning} for {Complex Study Designs}: {A Tutorial} for the mlpwr
{Package}}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/r9w6t}

\end{CSLReferences}

\end{document}

### 2024-06-19 manuscript.qmd
update title page

@@
   - name: Anne-Kathrin Kleine
+    orcid: 0000-0003-1919-2834
     roles:
@@
   - name: Eva Lermer
+    orcid: 0000-0002-6600-9580
     roles:
       - conceptualization
-      - funding aquisition
+      - funding acquisition
       - editing
@@
   - name: Susanne Gaube
+    orcid: 0000-0002-1633-4772
     roles:
@@
       - name: University College London
-        department: Global Business School for Health
+        department: UCL Global Business School for Health
 author-note:
@@
 A Quarto Manuscripts website is hosted at <https://Timo-Ko.github.io/glmm_simulation_tutorial/>.
-    conflict-of-interest: The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
+    conflict-of-interest: The authors declare that there were no conflicts of interest concerning the authorship or the publication of this article.
     financial-support: This research was funded by a grant from the Volkswagen Foundation (Grant No. 98525).
@@
     authorship-agreements: ~
+title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; <https://credit.niso.org/>):"
 abstract: When planning experimental research, determining an appropriate sample size and using suitable statistical models are crucial for robust and informative results.

### 2024-06-19 manuscript.qmd
work on comments

@@
 
-When planning experimental research, it is essential to determine an appropriate sample size and to use appropriate statistical models to analyze the data to ensure that the results obtained are both robust and informative [@lakensSampleSizeJustification2022].
+When planning experimental research, it is essential to determine an appropriate sample size and use appropriate statistical models to analyze the data to ensure that the results are robust and informative [@lakensSampleSizeJustification2022].
 The recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of study findings [@yarkoniGeneralizabilityCrisis2022].
@@
 After introducing some theoretical background on sample size planning, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
-Then we describe general steps and decisions involved in tailored data simulation.
-To illustrate the details of these steps, we finish with a hypothetical case study from the field of human-AI (artificial intelligence) interaction research.
+Then, we describe the general steps and decisions involved in tailored data simulation, illustrating them using a case study.
 
@@
 
-To benefit most of this tutorial paper, we recommend readers to familiarise themselves with basic statistical concepts like hypothesis tests (HTs) and their statistical power as well as confidence intervals (CIs) and their precision [@lakensImprovingYourStatistical2022; @riesthuisSimulationBasedPowerAnalyses2024; @kumleEstimatingPowerGeneralized2021].
-Some knowledge of causal inference is useful but not necessary [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
+To benefit most of this tutorial paper, we recommend readers to familiarize themselves with basic statistical concepts like hypothesis tests (HTs) and their statistical power as well as confidence intervals (CIs) and their precision [@lakensImprovingYourStatistical2022; @riesthuisSimulationBasedPowerAnalyses2024; @kumleEstimatingPowerGeneralized2021].
+Some knowledge of causal inference is beneficial but not necessary [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
 In addition, readers should have a rough understanding of R [@wickhamDataScienceImport2023] and how to simulate data.
 For data simulation, we use functions from the tidyverse [@wickhamWelcomeTidyverse2019] and the faux package [@R-faux].
-Finally, readers should be familiar with regression modeling in general and GLMMs in particular.
-In this tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data generating process [@debruineUnderstandingMixedEffectsModels2021].
+Finally, readers should be familiar with general regression modeling and GLMMs.
+In this tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data-generating process [@debruineUnderstandingMixedEffectsModels2021].
 It is not necessary to understand the technical details of how GLMMs are estimated.
-However it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the values in the dependent variable are determined by the predictor variables and the random effects.
+However, it is crucial to understand the structure of a basic GLMM (e.g., logistic regression with random intercepts) and how the model assumes that the dependent variable’s values are determined by the predictor variables and the random effects.
 
@@
 First, experiments may yield inconclusive or misleading results, hindering the accumulation of knowledge.
-Second, studies that are doomed to never finding a postulated effect waste resources by consuming time, effort, and funding without delivering meaningful results.
-For these reasons, many journals and funding bodies now require that a justification of sample size is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
+Second, studies that are doomed never to find a postulated effect waste resources by consuming time, effort, and funding.
+For these reasons, many journals and funding bodies now require a sample size justification in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
 Although scientists often do not justify sample size or use general heuristics from the literature, resource constraints often play an important role [@lakensSampleSizeJustification2022].
@@
 
-The majority of empirical studies in psychology and other social sciences apply hypothesis testing.
-As a consequence, the dominant approach for determining an adequate sample size is based on power analysis (i.e., planning for power) [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
+Most empirical studies in psychology and other social sciences apply hypothesis testing.
+Consequently, the dominant approach for determining an adequate sample size is based on power analysis (i.e., planning for power) [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
 Statistical power is defined as the probability that a HT has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size [@cohenPowerPrimer1992].
 Less formally, power is described as the probability that a HT correctly rejects the null hypothesis when the alternative hypothesis is true.
-If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study is considered "underpowered".
+If the sample size (i.e., the number of participants and/or stimuli) is insufficient to detect the effects or relationships being investigated with high probability, the study is considered "underpowered".
 When planning for power, a target is set for the statistical power of a HT of interest.
-Assuming some effect size of interest and a desired significance level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
+Assuming an effect size of interest and a desired significance level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
 Although it is recommended to justify the desired significance level and power [@lakensJustifyYourAlpha2018], most empirical studies adopt the heuristic of $\alpha = 0.05$ and $1 - \beta = 0.80$.
@@
 In contrast to power analysis, sample size planning can also be based on the precision of parameter estimates (i.e., planning for precision or planning for accuracy) [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
-Not all research questions are best answered using hypothesis testing.
-It has been argued that basic research rarely requires making discrete decisions on whether some effect has been "discovered" and should thus shift from a hypothesis testing towards an estimation framework [@mcelreathStatisticalRethinkingBayesian2020; @cummingNewStatisticsWhy2014; @kruschkeBayesianNewStatistics2018].
-Although this view is not without critique [@uyguntuncEpistemicPragmaticFunction2023], at least for exploratory or pilot studies where little previous research has been conducted, more scientists seem to agree that simply estimating the effects of interest and making the estimation uncertainty transparent by reporting CIs is more useful.
-Assuming that no HTs are conducted for a planned study, power analysis is not relevant for sample size planning.
-Nonetheless, the sample size still has a crucial effect on how informative the planned study will be, because an effect of interest is estimated more precisely with bigger samples.
+Not all research questions are best answered by hypothesis testing.
+It has been argued that basic research rarely requires discrete decisions on whether some effect has been "discovered" and should thus shift from hypothesis testing towards an estimation framework [@mcelreathStatisticalRethinkingBayesian2020; @cummingNewStatisticsWhy2014; @kruschkeBayesianNewStatistics2018].
+Although this view is not without critique [@uyguntuncEpistemicPragmaticFunction2023], at least for exploratory or pilot studies with little prior research,  estimating the effects of interest and making the estimation uncertainty transparent by reporting CIs seems more useful.
+Assuming that no HTs are conducted, power analysis is not relevant for sample size planning.
+Nonetheless, the sample size still has a crucial effect on how informative the study will be because an effect of interest is estimated more precisely with bigger samples.
 In the precision framework, the target quantity commonly used for sample size planning is the expected width of a CI [@kelleySampleSizePlanning2006; @lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
 The values inside a CI are often interpreted as plausible values for the quantity of interest it is supposed to estimate.
-More formally, a CI with confidence level 0.95 provides the smallest interval with the property that upon repeated sampling, 95% of individual CIs would include the true quantity of interest.
-Thus, a narrow CI is more informative about the size of the true effect than a wide interval.
-Apart from the desired confidence level, the width of a CI depends strongly on the sample size.
+More formally, CI with a confidence level of 0.95 provides the smallest interval with the property that 95% of individual CIs would include the true quantity of interest upon repeated sampling.
+Thus, a narrow CI is more informative about the size of the true effect than a wide CI.
+Apart from the confidence level, the width of a CI depends strongly on the sample size.
 Because bigger samples carry more information, they lead to smaller CIs.
@@
 Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
-Because planning for precision is still rare, there are no common heuristics on how to choose the desired width of the CI [@lakensSampleSizeJustification2022].
+Because planning for precision is still rare, there are no common heuristics for choosing the desired width of the CI [@lakensSampleSizeJustification2022].
 
@@
 
-As study designs become more complex, psychological researchers require more sophisticated statistical models to capture the nuanced relationships and grouping structures introduced by them [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining increasing popularity because they offer great flexibility [@fahrmeirRegressionModelsMethods2021].
+As study designs become more complex, psychological researchers require more sophisticated statistical models to capture their nuanced relationships and grouping structures [@yarkoniGeneralizabilityCrisis2022].
+GLMMs (also called multilevel models) are gaining popularity because they offer great flexibility [@fahrmeirRegressionModelsMethods2021].
 GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data, including hierarchical structures [@fahrmeirRegressionModelsMethods2021].
-In this context, correlated data means that the value in the outcome variable for one observation may be related to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., age of participants).
-This correlation can arise for various reasons: For instance, responses to some stimuli from some participants might be more similar because the same person was measured twice (repeated measurements), participants come from the same neighborhood (clustering), or participants responded to the same stimulus (stimulus effects).
+In this context, correlated data means that the value in the outcome variable for one observation may be related to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., the age of participants).
+This correlation can arise for various reasons: For instance, responses to some stimuli from some participants might be more similar because the same person was measured multiple times (repeated measurements), participants belong to the same group (clustering), or participants responded to the same stimulus (stimulus effects).
 Thus, modeling such correlations is important whenever the data has a clear structure, while the grouping variables can be hierarchically nested (e.g., grouping variables students and schools: each student belongs to exactly one school) or cross-classified (e.g., grouping variables students and math exercises: each student is presented with several math exercises).
@@
 The participants’ behavior is measured repeatedly over several sessions.
-The click patterns of participants in one session are likely to be correlated with their previous sessions and the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
+The click patterns of participants in one session are likely to be correlated with their previous sessions, and the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
 
@@
 Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on formulas to calculate power directly while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., what percentage of simulated datasets produces a significant p-value).
-Available formula-based software packages for multilevel models often do not include GLMMs or are limited to very simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
+Formula-based methods rely on formulas to calculate power directly, while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., what percentage of simulated datasets produces a significant p-value).
+Available formula-based software packages for multilevel models often do not include GLMMs or are limited to simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
 A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a].
-However, many of these tutorials focus on linear mixed models (LMMs) and the common study designs [but see @kumleEstimatingPowerGeneralized2021, for a tutorial that also covers more advanced settings].
+However, many of these tutorials focus on linear mixed models (LMMs) and the most common study designs [but see @kumleEstimatingPowerGeneralized2021, for a tutorial that also covers more advanced settings].
 This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
-Simulation-based power analysis with GLMMs requires making a range of assumptions: The (conditional) distribution assumption specifies the distributional family for the outcome variable.
+Simulation-based power analysis with GLMMs requires making a range of assumptions: The distribution assumption specifies the distributional family of the outcome variable, after conditioning on all predictor variables.
 Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e., if and how they are correlated).
@@
 Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
-However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge (see the discussion in [@maxwellSampleSizePlanning2008] and [@kumleEstimatingPowerGeneralized2021]).
+However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge (see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021).
 
@@
 
-In Table 1, we give a short review of existing R packages that can be used for power analysis for GLMMs.
+In Table 1, we review existing R packages that can be used for power analysis for GLMMs.
 
@@
 
-## Step 2: Simulate the data generating process
+## Step 2: Simulate the data-generating process
 
@@
 
-When the estimand has been defined, the next step in the research process is to write code that simulates the (assumed) data generating process of the planned study.
+When the estimand has been defined, the next step in the research process is to write code that simulates the (assumed) data-generating process of the planned study.
 This requires specifying a generative process for all predictor variables used in the final data analysis.
@@
 We will discuss strategies on how these values can be obtained later.
-Because we have full control over the data generating process in a tailored simulation, it is possible to model specific aspects of the planned study, like data missing by design or assuming that some subjects might drop out.
-The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data generating process.
-However, we would argue that even a strongly simplified data generating process (e.g.
+Because we have full control over the data-generating process in a tailored simulation, it is possible to model specific aspects of the planned study, like data missing by design or assuming that some subjects might drop out.
+The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data-generating process.
+However, we would argue that even a strongly simplified data-generating process (e.g.
 only a small number of interaction effects; only random intercepts and no random slopes; assuming that data is missing completely at random) can yield informative results.
@@
 Because most domain knowledge can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
-The basic idea of all strategies is how the data generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
+The basic idea of all strategies is how the data-generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
 
@@
 This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
-Note that this does not always mean that one will specify the same GLMM that was used when specifying the data generating process.
+Note that this does not always mean that one will specify the same GLMM that was used when specifying the data-generating process.
 On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
-On the other hand, using a more complex model for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
+On the other hand, using a more complex model for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data-generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
 

### 2024-06-19 manuscript.qmd
work on comments continued

@@
     authorship-agreements: ~
+language:
   title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; <https://credit.niso.org/>):"
@@
 However, available methods and software for a priori sample size planning for GLMMs are often limited to specific designs.
-Tailored data simulation approaches are a more flexible alternative.
+Tailored data simulation approaches offer a more flexible alternative.
 Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori sample size planning with GLMMs.
@@
 Some knowledge of causal inference is beneficial but not necessary [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
-In addition, readers should have a rough understanding of R [@wickhamDataScienceImport2023] and how to simulate data.
+In addition, readers should have an understanding of how to conduct statistical analyses with R [@wickhamDataScienceImport2023] and how to simulate data.
 For data simulation, we use functions from the tidyverse [@wickhamWelcomeTidyverse2019] and the faux package [@R-faux].
@@
 Second, studies that are doomed never to find a postulated effect waste resources by consuming time, effort, and funding.
-For these reasons, many journals and funding bodies now require a sample size justification in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings.
-Although scientists often do not justify sample size or use general heuristics from the literature, resource constraints often play an important role [@lakensSampleSizeJustification2022].
+For these reasons, many journals and funding bodies now require a sample size justification in study protocols and grant proposals, recognizing its relevance in ensuring robust and meaningful findings.
+Although scientists often do not justify sample sizes or use general heuristics from the literature, resource constraints often play an important role [@lakensSampleSizeJustification2022].
 But ideally, a suitable sample size should be determined a priori (i.e., before the study is conducted) based on some meaningful computation to ensure that the study will be able to fulfill its purpose.
@@
 A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a].
-However, many of these tutorials focus on linear mixed models (LMMs) and the most common study designs [but see @kumleEstimatingPowerGeneralized2021, for a tutorial that also covers more advanced settings].
-This narrow focus provides limited guidance for researchers faced with more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
+However, many of these tutorials focus on linear mixed models (LMMs) and the most common study designs [see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
+This narrow focus provides limited guidance for researchers using more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
 Simulation-based power analysis with GLMMs requires making a range of assumptions: The distribution assumption specifies the distributional family of the outcome variable, after conditioning on all predictor variables.
@@
 Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in TABLE 1.
-The most important circumstances under which tailored simulation-based sample size planning is necessary are 1) complex study designs, 2) complex statistical hypotheses, 3) planning for precision, 4) no available prior studies or pilot data.
+Circumstances under which tailored simulation-based sample size planning is necessary include 1) complex study designs, 2) complex statistical hypotheses, 3) planning for precision, and 4) lack of prior studies or pilot data.
 
@@
 Alternatively, subjects might systematically drop out or produce missing data, but this dropout can be explained by some attributes also measured in the dataset (e.g., older subjects have a higher probability to refuse answering a question on income).
-In a more complicated scenario, missing data in some variable is caused by the measured attribute itself (e.g., wealthy people are more likely to refuse reporting their income).
+In a more complicated scenario, missing data in some variables is caused by the measured attribute itself (e.g., wealthy people are more likely to refuse to report their income).
 Moreover, many experimental designs contain conditions in which values of the predictor variables are missing by design.
@@
 However, it is often challenging to decide whether missing data can be safely ignored in the data analysis and sample size planning process based on a merely theoretical approach [@gomilaMissingDataExperiments2022].
-Tailored simulation-based approaches offer the possibility to include the assumed process of how data become missing in the data simulation, thereby determining the required sample based on simulated datasets that contain missing values [for an example, see @lanePowerStrugglesEstimating2018].
+Tailored simulation-based approaches offer the possibility to include the assumed process of how data become missing in the data simulation, thereby determining the required sample based on simulated datasets that contain missing values [for example, see @lanePowerStrugglesEstimating2018].
 As a byproduct, the simulated datasets can also be used to test whether the intended data analysis provides the expected (unbiased) results, despite the missing data.
@@
 Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects that can be modeled with the R packages glmmTMB [@brooksGlmmTMBBalancesSpeed2017] or brms [@burknerAdvancedBayesianMultilevel2018].
-Tailored simulation-based approaches do not share the same limitation than the existing software solutions for power analysis that focus exclusively on GLMMs.
+Tailored simulation-based approaches do not share the same limitations as the existing software solutions for power analysis that focus exclusively on GLMMs.
 As long as there is a software package available to estimate the model of interest, it is always possible to perform tailored simulation-based sample size planning.
@@
 Even better theories should be able to specify the smallest effect sizes of interest (SESOI) that must be exceeded if the effect has any practical relevance [@lakensEquivalenceTestingPsychological2018].
-In combination, this might require a test such as $H_0: \beta \leq 0.1$.
-More elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example testing simple slopes [@preacherComputationalToolsProbing2006] with a hypothesis such as $H_0: \beta_0 + \beta_1 \leq 0$.
+This might require a test such as $H_0: \beta \leq 0.1$.
+More elaborate research questions often require testing hypotheses that consist of a combination of model parameters, for example, testing simple slopes [@preacherComputationalToolsProbing2006] with a hypothesis such as $H_0: \beta_0 + \beta_1 \leq 0$.
 If the research question consists only of a single hypothesis of this sort, it might be possible to reduce the hypothesis to a single regression coefficient by clever coding and/or centering of predictor variables.
-However, interesting research questions often consist of combined hypotheses that consist of more than one separate statistical hypothesis [for a tutorial on contrast analysis in GLMMs, see @schadHowCapitalizePriori2020].
+However, research questions often consist of combined hypotheses that consist of more than one separate statistical hypothesis [for a tutorial on contrast analysis in GLMMs, see @schadHowCapitalizePriori2020].
 For example, a combined null hypothesis $H_0$ might consist of two single null hypotheses $H_{01}: \beta_1 \leq 0$ and $H_{02}: \beta_0 + \beta_1 \leq 0$.
@@
 However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
-None of the software packages for sample size planning in table X can handle combined hypotheses as discussed here and only some can handle directed hypotheses.
+None of the software packages for sample size planning in Table 1 can handle combined hypotheses as discussed here and only some can handle directed hypotheses.
 In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation-based sample size planning.
@@
 It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
-However, we would argue that when tailored data simulations necessary to determine effect sizes anyway, performing the whole sample size planning in a customized way is preferred over using the existing software packages.
+However, when tailored data simulations are necessary to determine effect sizes anyway, performing the whole sample size planning in a customized way is preferred over using the existing software packages.
 
@@
 the theoretical quantity which is necessary to answer a specific research question.
-The estimand consists of a quantity, that can be described for each unit under investigation and a clear definition of the target population, for which the quantity is of interest.
-For example, an estimand might be the probability that a clinical psychologists makes the correct diagnosis for a psychiatric patient with major depression, averaged across all clinical psychologists and depressed patients in psychiatric institutions in Germany.
+The estimand consists of a quantity that can be described for each unit under investigation and a clear definition of the target population, for which the quantity is of interest.
+For example, an estimand might be the probability that a clinical psychologist makes the correct diagnosis for a psychiatric patient with major depression, averaged across all clinical psychologists and depressed patients in psychiatric institutions in a given country.
 
@@
 
-In the present hypothetical case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) embedded in a diagnostic decision support system.
-The context is a clinical setting, where expert radiologists and students under training must detect bleeding based on head scans from computer tomography (CT).
-In the investigated AI-enabled diagnostic decision support system, an AI model can provide initial diagnostic advice, which can be used as guidance by the humans who are required to make the final diagnostic decision.
+In the present case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) embedded in a clinical decision support system.
+The context is a clinical setting, where task experts, i.e., radiologists, and non-expert, i.e., medical students, must detect bleedings based on head scans from computer tomography (CT).
+The  AI model can provide initial diagnostic advice, which can be used as guidance by the humans who are required to make the final diagnostic decision.
 The research goal is to validate the effectiveness of the AI-enabled advice.
-We consider the AI-enabled advice as effective, if the following pattern holds, which we will first describe verbally:
-
-*We expect that for BOTH expert radiologists and medical students, correct AI-advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI-advice presented, AND, we expect that for BOTH experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
+We consider the AI-enabled advice effective, if the following pattern holds:
+
+*We expect that for BOTH radiologists and medical students, correct AI advice leads to a higher probability of accurately diagnosing a CT scan compared to no AI advice presented, AND, we expect that for BOTH task experts and non-experts, incorrect advice leads to a lower probability of accurately diagnosing a CT scan compared to no advice presented.*
 
@@
 $$
-For example, the first expression is concerned with the difference between the probability that a correct diagnosis is made if correct AI-advice is presented and the probability that a correct diagnosis is made if no AI-advice is presented.
+For example, the first expression is concerned with the difference between the probability that a correct diagnosis is made if correct AI advice is presented and the probability that a correct diagnosis is made if no AI advice is presented.
 This contrast is quantified for a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.^[Note that a different estimand would be the so-called average treatment effect (ATE).
 For the ATE, the probability contrast is defined for each combination of expert and scan, and then these contrasts are averaged across all experts and scans from the target population [@lundbergWhatYourEstimand2021].].
-To complete our definition of the estimand, we have to define our target population that consists of persons, stimuli, and AI-advice: With respect to persons, we are only interested in expert radiologists and medical students at German universities.
+To complete our definition of the estimand, we have to define our target population that consists of persons, stimuli, and AI advice: With respect to persons, we are only interested in experts (i.e., radiologists) and non-experts (i.e., medical students).
 With respect to stimuli, we are only interested in the head CT scans made from subjects that do or do not suffer from intracerebral hemorrhage.
-Lastly, we are only interested in AI-advice given by a specific AI-enabled diagnostic decision support system.
-
-Although the estimand is initially defined outside of any statistical model, it is only useful if we find a way to estimate it based on observed data.
-For our exemplary research question, it is possible to construct an experimental study, where all participants are confronted with the same set of head CT scans, but the kind of AI-advice given for each scan is randomly assigned within participants.
-This random intervention allows us to produce an empirical estimate of our estimand, although, in reality, each person receives only one kind of AI-advice (correct advice, wrong advice, no advice) for each scan.
+Lastly, we are only interested in AI advice given by a specific AI-enabled system.
+
+Although the estimand is initially defined outside of a statistical model, it is only useful if we find a way to estimate it based on observed data.
+For our exemplary research question, it is possible to construct an experimental study where all participants are confronted with the same set of head CT scans, but the kind of AI advice given for each scan is randomly assigned within participants.
+This random intervention allows us to produce an empirical estimate of our estimand, although, in reality, each person receives only one kind of AI advice (correct advice, incorrect advice, no advice at all) for each scan.
 We will see later how each of the probability expressions in our estimand can be modeled with the same GLMM.
@@
 
-When the estimand has been defined, the next step in the research process is to write code that simulates the (assumed) data-generating process of the planned study.
+When the estimand has been defined, the next step in the research process is to write code that simulates the data-generating process of the planned study.
 This requires specifying a generative process for all predictor variables used in the final data analysis.
@@
 The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data-generating process.
-However, we would argue that even a strongly simplified data-generating process (e.g.
-only a small number of interaction effects; only random intercepts and no random slopes; assuming that data is missing completely at random) can yield informative results.
+However, we argue that even a strongly simplified data-generating process (e.g., only a small number of interaction effects; only random intercepts, and no random slopes; assuming that data is missing completely at random) can yield informative results.
 
@@
 
-In our case study, we simulate data for an experiment where the diagnostic performance of users of an AI-enabled diagnostic decision support system will be evaluated.
-Radiologists (task experts) and students/interns (non experts) review a series of head computer tomography (CT) scans to assess the presence of a bleeding.
+In our case study, radiologists and students review a series of head CT scans to assess the presence of a bleeding.
 An AI model provides initial diagnostic advice to assist their decision-making.
@@
 When AI advice is given, this advice can be either correct or incorrect.
-The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across brain scans.
-After reviewing a CT scan, participants deliver a medical diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
+The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across CT scans.
+After reviewing a CT scan, participants deliver a diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
 This experimental design introduces some missing values by design since the advice is neither correct nor incorrect when no advice is present, which must be taken into account when simulating and analyzing the data.
-In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students/interns) are more readily accessible.
+In this example, recruiting task experts (i.e., radiologists) is more challenging due to their limited availability, while non-experts (i.e., students) are more readily accessible.
 The goal of simulation-based sample size planning is to determine how many task experts and non-experts must be recruited to achieve sufficient statistical power or precision in the planned experiment.
@@
 
-In a GLMM, the expected value of the dependent variable $Y$ conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
+In a GLMM, the expected value of the dependent variable $Y$ conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$, and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
 $$
@@
 In the GLMM literature, this would be called a binomial GLMM with two random intercepts (for subjects and items), two level-1 predictors ($advice\_present$, $advice\_correct$), one level-2 predictor ($expert$) and two cross-level interactions ($expert \cdot advice\_present$, $expert \cdot advice\_correct$).
-To limit complexity, we do not consider random slopes, additional predictors or higher-level interactions.
+To limit complexity, we do not consider random slopes, additional predictors, or higher-level interactions.
 
@@
 Because most domain knowledge can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
-The basic idea of all strategies is how the data-generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables a calibration of population parameters based on the available knowledge of domain experts.
-
-Although we use frequentist model estimation in our tutorial, many strategies described in this chapter are inspired by research on how to monitor the plausibility of model assumptions in applied Bayesian statistics [@gelmanBayesianWorkflow2020].
+The basic idea of all strategies is how the data-generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables calibration of population parameters based on the available knowledge of domain experts.
+
+Although we use frequentist model estimation in our tutorial, many of the strategies described in this chapter are inspired by research on monitoring the plausibility of model assumptions in applied Bayesian statistics [@gelmanBayesianWorkflow2020].
 
@@
 
-When introducing the simulation function for our case study, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$), but have not talked about where these numbers came from.
-
-The starting point for all parameter values in our present case study were based on results from distantly related study designs in the literature.
-Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions seem plausible.
+When introducing the simulation function for our case study, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$) but have not talked about where these numbers came from.
+
+The starting point for all parameter values in our present case study was based on results from distantly related study designs in the literature.
+Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions regarding participants’ diagnostic performance seemed plausible.
 
 We now outline our main strategy to determine plausible parameter values for the fixed effects ($\beta$ parameters):
-Unfortunately, the model parameters in a binomial GLMM are hard to interpret in isolation because 1) the $\beta$ parameters are connected to the modeled probability via the non-linear inverse-logit link, and 2) we also have to consider the random effects.
-The most simple interpretation, that allows us to ignore the random effects for now, works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
-Then the model implied probability that such a person solves such an item accurately is given by:
+Unfortunately, the model parameters in a binomial GLMM are hard to interpret in isolation because first, the  parameters are connected to the modeled probability via the non-linear inverse-logit link, and second, we also have to consider the random effects.
+The most simple interpretation, which allows us to ignore the random effects for now, works by imagining a subject with average ability ($u_{0s} = 0$) responding to an item (i.e., CT scan) with average difficulty ($u_{0i} = 0$).
+Then the model-implied probability that such a person solves such an item accurately is given by:
 
@@
 In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-For example, the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented, only requires the intercept:
+For example, the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented only requires the intercept:
 $$
@@
 For example, domain experts might not be able to directly choose plausible values for the  coefficients in a logistic regression model (which are measured on the log-odds scale).
-However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental conditions, i.e.
-which relative frequency they expect to observe (on average).
-The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model implied value of the insightful descriptive statistic expected by the domain expert.
+However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental conditions, i.e., which relative frequency they expect to observe (on average).
+The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model-implied value of the insightful descriptive statistic expected by the domain expert.
 Although insightful descriptive statistics usually depend on the model parameters in a non-linear way, it is not necessary to solve the exact relationship mathematically.
-Instead, one can simply adjust the population parameters by trial and error until the model implied quantities produce the desired result.
+Instead, one can simply adjust the population parameters by trial and error until the model-implied quantities produce the desired result.
 
@@
 
-In the last section, we showed how we can derive the model implied probability that a subject with average ability solves an item with average difficulty for each experimental condition.
+In the last section, we showed how we can derive the model-implied probability that a subject with average ability solves an item with average difficulty for each experimental condition.
 Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty [@fahrmeirRegressionModelsMethods2021].
@@
 For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice (the conditional perspective).
-However, even if the model were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability, and items drawn from their natural distribution of difficulty (the marginal perspective).
-How much the two probabilities differ depends on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations would be zero).
-We want to use the model implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistics to determine plausible values for the random effect standard deviations.
-
-We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model implied true values) and simply compute the relative frequency of correct diagnoses for each experimental condition.
+However, even if the model assumptions were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty (the marginal perspective).
+How much the two probabilities differ depends on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations are be zero).
+We want to use the model-implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistic to determine plausible values for the random effect standard deviations.
+
+We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model-implied true values) and simply compute the relative frequency of correct diagnoses for each experimental condition.
 
@@
 
-We tried using these descriptive statistics to judge together with domain experts whether our chosen values for the random effect standard deviations would produce data that align with out domain expertise.
-However, although the result was deemed plausible, these statistics were not informative enough to determine a final set of plausible parameter values.
-For this reason, we will additionally look at insightful model based quantities.
+We tried using these descriptive statistics to judge together with domain experts whether our chosen values for the random effect standard deviations would produce data that aligned with their domain expertise.
+Although the result was deemed plausible, these statistics were not informative enough to determine a final set of plausible parameter values.
+For this reason, we will additionally look at insightful model- based quantities.
 
@@
 An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
-For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant on average solves a cognitive task.
+For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant, on average, solves a cognitive task.
 Although domain knowledge will probably not suffice to specify this distribution completely, it should be possible to rule out implausible boundary conditions.
@@
 In the last section, we simulated a large dataset and computed descriptive statistics, the relative frequencies of correct diagnoses, for each experimental condition.
-We will now use the model implied probability of each simulated data point (stored in the variable `y_prob`) to visualize the whole model implied marginal distribution of correct diagnoses for each experimental condition.
+We will now use the model-implied probability of each simulated data point (stored in the variable `y_prob`) to visualize the whole model-implied marginal distribution of correct diagnoses for each experimental condition.
 
@@
 ```
-@fig-margdist1 shows the model implied marginal distributions, including the mean, 66% and 95% intervals.
+@fig-margdist1 shows the model-implied marginal distributions, including the mean, 66% and 95% intervals.
 We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
 This difference increases with the variability of the random effects.
-We can use plots like the one above as a useful tool to decide whether the specified standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
+We can use plots like @fig-margdist1 as a useful tool to decide whether the specified standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
 
@@
 ```
-In the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$).
+For the next plot, we have set the item standard deviation to almost zero ($\sigma_I = 0.01$).
 This gives us a better way to see the variability between persons.
@@
 Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
-We could also have a closer look at variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
+We could also have a closer look at the variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
 ```{r}
@@
 These high standard deviations do not account for the inherent variability and complexity of human performance.
-For example, we would expect that a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero even when presented with wrong advice.
+For example, we would expect a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero, even when presented with wrong advice.
 ```{r}
@@
 
-In our experience, the gathering of domain knowledge by domain experts and the consecutive specification of population parameter values used in data simulation is not a one-time event but rather an iterative process.
-In a first step, domain experts can be interviewed to "elicit" their domain knowledge about how the future data of the planned study is expected to look like.
+In our experience, gathering domain knowledge by domain experts and the specification of population parameter values is not a one-time event but rather an iterative process.
+As a first step, domain experts can be interviewed to “elicit” their domain knowledge about how the data of the planned study is expected to look.
 As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
-For this reason, we suggest that after an initial unstructured interview of domain experts, the analyst who is familiar with the structure of the GLMM under study selects an initial set of insightful descriptive statistics and model based quantities.
-Then they reenter into an iterative discussion where some set of population values are selected and the plausibility of resulting implied quantities are discussed with the domain experts.
-Then, the population parameters are updated based on this discussion until the domain experts are satisfied with the final result.
-During this process, the monitored model based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
+For this reason, we suggest that after an initial interview, the analyst who is familiar with the structure of the GLMM selects an initial set of insightful descriptive statistics and model-based quantities.
+Then they reenter into an iterative discussion with the domain experts where some set of population values are selected, and the plausibility of resulting implied quantities is evaluated.
+Then, the population parameters are updated until the domain experts are satisfied with the final result.
+During this process, the monitored model-based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
 
@@
 
-All parameter values in our present case study have been determined based on repeated discussions with our affiliated domain experts in Radiology to validate our assumptions.
+All parameter values in our present case study have been determined based on repeated discussions with our affiliated domain experts in radiology to validate our assumptions.
 Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for intracranial hemorrhage.
@@
 From their findings, we inferred that both experts and non-experts should benefit from correct and suffer losses from incorrect advice.
-However, the magnitude of these effects should be substantially greater for non-experts, given their demonstrated reliance on advice compared to experts.
+However, the magnitude of these effects should be substantially greater for non-experts, given their stronger reliance on advice compared to experts.
 We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
@@
 
-At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset that will later be collected in the planned study.
+At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset to be collected in the planned study.
 The next step is to specify how the statistical model shall be estimated in the actual study collected later.
@@
 The next step is to specify how to compute a concrete point estimate of the theoretical estimand within the framework of the fitted GLMM.
-For some research question, the estimate corresponds with a single regression coefficient. In more complicated scenarios, the estimate is computed from a combination of coefficients.
+For some research questions, the estimate corresponds with a single regression coefficient. In more complicated scenarios, the estimate is computed from a combination of coefficients.
 Beyond computing the point estimate, we have already discussed that both hypothesis testing and interval estimation can be used to answer the research question.

### 2024-06-19 manuscript.qmd
more work on comments

@@
 When planning experimental research, it is essential to determine an appropriate sample size and use appropriate statistical models to analyze the data to ensure that the results are robust and informative [@lakensSampleSizeJustification2022].
-The recent replication crisis has illustrated many challenges surrounding the reproducibility and reliability of study findings [@yarkoniGeneralizabilityCrisis2022].
+The recent replication crisis in Psychology and other disciplines has illustrated many challenges surrounding the reproducibility and reliability of study findings [@yarkoniGeneralizabilityCrisis2022].
 As a result, there is a growing need for more rigorous statistical methodology and the adoption of adequately powered experimental designs.
@@
 After introducing some theoretical background on sample size planning, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
-Then, we describe the general steps and decisions involved in tailored data simulation, illustrating them using a case study.
+We proceed by describing the relevant steps and decisions involved in tailored data simulation, illustrated in a case study.
 
@@
 For data simulation, we use functions from the tidyverse [@wickhamWelcomeTidyverse2019] and the faux package [@R-faux].
-Finally, readers should be familiar with general regression modeling and GLMMs.
+Finally, readers should have a basic understanding of regression modeling and GLMMs [@bolkerLinearGeneralizedLinear2015].
 In this tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data-generating process [@debruineUnderstandingMixedEffectsModels2021].
@@
 
-Conducting research with insufficiently large sample sizes can have many negative consequences.
+Conducting research with insufficiently large sample sizes can have many negative consequences [@buttonPowerFailureWhy2013].
 First, experiments may yield inconclusive or misleading results, hindering the accumulation of knowledge.
@@
 In the precision framework, the target quantity commonly used for sample size planning is the expected width of a CI [@kelleySampleSizePlanning2006; @lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
-The values inside a CI are often interpreted as plausible values for the quantity of interest it is supposed to estimate.
-More formally, CI with a confidence level of 0.95 provides the smallest interval with the property that 95% of individual CIs would include the true quantity of interest upon repeated sampling.
-Thus, a narrow CI is more informative about the size of the true effect than a wide CI.
+A CI with a confidence level of 0.95 provides the smallest interval with the property that 95% of individual CIs would include the true quantity of interest upon repeated sampling.
+Thus, a narrow CI with fewer plausible values for the quantity of interest is more informative about the size of the true effect than a wide CI.
 Apart from the confidence level, the width of a CI depends strongly on the sample size.
@@
 Requirements of real-world studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
-One frequent issue in applied data analysis is missing data, and there can be various reasons for this [@littleStatisticalAnalysisMissing2014].
-For example, data can be missing completely at random (e.g., because an electronic measurement device randomly failed for some technical reasons).
+One frequent issue in applied data analysis is missing data [@littleStatisticalAnalysisMissing2014].
+Data can be missing completely at random (e.g., because an electronic measurement device randomly failed for some technical reasons).
 Alternatively, subjects might systematically drop out or produce missing data, but this dropout can be explained by some attributes also measured in the dataset (e.g., older subjects have a higher probability to refuse answering a question on income).
@@
 However, it is often challenging to decide whether missing data can be safely ignored in the data analysis and sample size planning process based on a merely theoretical approach [@gomilaMissingDataExperiments2022].
-Tailored simulation-based approaches offer the possibility to include the assumed process of how data become missing in the data simulation, thereby determining the required sample based on simulated datasets that contain missing values [for example, see @lanePowerStrugglesEstimating2018].
+Tailored simulation-based approaches offer the possibility to include the assumed process of how data become missing in the data simulation, thereby determining the required sample size based on simulated datasets that contain missing values [for example, see @lanePowerStrugglesEstimating2018].
 As a byproduct, the simulated datasets can also be used to test whether the intended data analysis provides the expected (unbiased) results, despite the missing data.
@@
 However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
-None of the software packages for sample size planning in Table 1 can handle combined hypotheses as discussed here and only some can handle directed hypotheses.
+None of the software packages for sample size planning in Table 1 can handle combined hypotheses and only some can handle directed hypotheses.
 In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation-based sample size planning.
@@
 Although planning for precision [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008; @cummingNewStatisticsWhy2014] has been increasingly discussed as a useful strategy for empirical research, all available software packages for sample size planning with GLMMs are based on power analysis.
-Therefore, researchers that want to want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE 1.
+Therefore, researchers that want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE 1.
 <!-- As long as one is interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
@@
 However, tailored simulation-based sample size planning can easily handle the planning for precision approach [@maxwellSampleSizePlanning2008].
-The only change in procedure is that instead of computing HTs for each simulated dataset and estimating statistical power across repetitions, CIs are computed for each simulated dataset, and the expected width is estimated.
+The only change in procedure is that instead of computing HTs for each simulated dataset and estimating statistical power across repetitions, CIs are computed for each simulated dataset, and their expected width is estimated.
 
@@
 When the researcher has access to similar studies or pilot data, providing such standardized effect sizes is feasible.
-However, note that choosing effect sizes based on small pilot studies is generally not recommended, as those estimates can be heavily biased [@albersWhenPowerAnalyses2018a; @lakensSampleSizeJustification2022].
+However, choosing effect sizes based on small pilot studies is generally not recommended, as those estimates can be heavily biased [@albersWhenPowerAnalyses2018a; @lakensSampleSizeJustification2022].
 Providing an informed standardized effect size can be an almost impossible challenge when no prior studies of pilot data are available.
@@
 Although the details differ depending on the specific study characteristics, each tailored simulation-based sample size planning requires a series of steps and decisions.
-We will introduce each step in a theoretical section, followed by the practical application based on a hypothetical case study.
+We will introduce each step in a theoretical section, followed by the practical application based on a case study.
 All code in this manuscript and simulation results are available in the project’s repository on the Open Science Framework (<https://osf.io/dhwf4/>).
@@
 
-The first step in every research process should be a clear definition of the theoretical *estimand* [@lundbergWhatYourEstimand2021], i.e.
+The first step in every research process is a clear definition of the theoretical *estimand* [@lundbergWhatYourEstimand2021], i.e.
 the theoretical quantity which is necessary to answer a specific research question.
@@
 The estimand should always be defined outside of any statistical model, because there are usually a range of statistical methods that can be used to estimate the same estimand, depending on the study design (e.g., a randomized experiment) that will produce the observed data in the planned study.
-For many common research questions in psychology, the estimand can be expressed as a statistical quantity that can be estimated with a regression model, for example a single $\beta$ coefficient.
+For many common research questions in psychology, it is possible to express the estimand as a statistical quantity that can be estimated with a regression model, for example a single $\beta$ coefficient.
 However, this is not possible for all estimands, which is why the literature discusses many estimation strategies beyond regression [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
@@
 This leaves us with the question on how to determine plausible values for the two remaining model parameters ($\sigma_S$, and $\sigma_I$) that are the standard deviations for the random intercepts.
-For this, we introduce two more strategies.
+For this, we introduce two more strategies in the following sections.
 
@@
 
-The final plot demonstrates that these plots are also useful for spotting standard deviations that are specified too high.
+The final plot demonstrates that these plots are also useful for spotting standard deviations that were specified too high.
 For @fig-margdist3 we have set $\sigma_S = 3$ and $\sigma_I = 3$.
-This implies that in each experimental condition, the probabilities that a subject solves an item are usually close to either 0 or 1, which is not a plausible assumption.
+This implies that in each experimental condition, the probabilities that a subject solves an item are overwhelmingly close to either 0 or 1 and nothing in between, which is not a plausible assumption.
 These high standard deviations do not account for the inherent variability and complexity of human performance.
-For example, we would expect a participant with low ability compared to other task experts to solve a difficult item with a probability substantially larger than zero, even when presented with wrong advice.
+For example, the expectation that an expert with low ability and incorrect advice would solve a difficult item with a probability close to zero is not convincing.
 ```{r}
@@
 
-In our experience, gathering domain knowledge by domain experts and the specification of population parameter values is not a one-time event but rather an iterative process.
+Gathering domain knowledge by domain experts and the specification of population parameter values is an iterative process.
 As a first step, domain experts can be interviewed to “elicit” their domain knowledge about how the data of the planned study is expected to look.
@@
 Then they reenter into an iterative discussion with the domain experts where some set of population values are selected, and the plausibility of resulting implied quantities is evaluated.
-Then, the population parameters are updated until the domain experts are satisfied with the final result.
+Then, the population parameters are updated until the domain experts are satisfied with the result.
 During this process, the monitored model-based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
@@
 
-All parameter values in our present case study have been determined based on repeated discussions with our affiliated domain experts in radiology to validate our assumptions.
+All parameter values in our present case study have been determined based on repeated discussions with domain experts in radiology to validate our assumptions.
 Initially, we reviewed the literature to establish a reasonable baseline performance rate for examining head CT scans for intracranial hemorrhage.
-Existing studies indicate that radiologists typically demonstrate high accuracies, often exceeding or hovering around 90%, while interns have been shown to perform below 80%, and medical students fall even shorter.
+Existing studies indicate that radiologists typically demonstrate high accuracies around 90%, while interns have been shown to perform below 80%, and medical students fall even shorter.
 For simplicity, we assumed plausible probability values of .90 for experts and .70 for non-experts, respectively.
@@
 This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
+
 Note that this does not always mean that one will specify the same GLMM that was used when specifying the data-generating process.
@@
 We have already discussed how to compute the involved probabilities in the section on specifying population parameters.
-Plugging in the model equation of the GLMM produces an equation on how to compute each contrast if all model parameters were known.
-When we want to estimate the above contrasts based on *observed* data, the only difference is that model parameters are not known and we instead use the corresponding parameter *estimates*.
+Plugging in the GLMM model equation produces an equation for computing each contrast if all model parameters were known.
+When we want to estimate the above contrasts based on observed data, the only difference is that model parameters are not known and we instead use the corresponding parameter estimates.
 
@@
 The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute HTs and CIs for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
-First, we specify a grid of all combinations of predictor variable and then compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
+First, we specify a grid of all combinations of predictor variables and then compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
 
@@
 Next, we use the estimated probabilities to compute the four specific contrasts that are part of our estimand.
-For this we must specify which rows in `probs` have to be subtracted from each other.
+For this, we must specify which rows in `probs` have to be subtracted from each other.
 We will use the `hypotheses` function to compute our four contrasts of interest together with HTs and CIs.
@@
 If we chose parameter estimation for our case study, we would focus on the two-sided CIs of the four contrasts of interest.
-With a standard confidence level of $1 - \alpha = 0.95$, plausible values are clearly in the positive range for the first, third and fourth contrast, while both negative and positive values seem plausible for the second contrast.
+With a standard confidence level of $1 - \alpha = 0.95$ plausible values are clearly in the positive range for the first, third, and fourth contrast, while both negative and positive values seem plausible for the second contrast.
 Note that due to the constrained range of the probability scale, the width of the CI differs between the four contrasts (which is the expected behavior for binomial GLMMs).
-The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice) where both underlying probabilities are close to 1.
+The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice), where both underlying probabilities are close to 1.
 The largest width is observed for the fourth contrast (non-expert with wrong advice vs. non-expert without advice), where both underlying probabilities are closer to 0.5.
@@
 Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute HTs or CIs for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
-The last missing piece is to write code to perform the above steps repeatedly and allow for a setting using different sample sizes.
-On a conceptual level, we first require a function that takes as input the sample size and the full set of population parameter values.
+The last missing piece is to write code that performs the previous steps repeatedly and allows for setting different sample sizes.
+On a conceptual level, we first require a function that takes the sample size and the full set of population parameter values as input.
 When planning for power, the function should return the p-value(s) of the HT(s) of interest when conducted on the simulated dataset.
 When planning for precision, the function should return the width of the CI(s) of interest.
-Secondly, we must run this function repeatedly with the same sample size and population parameters.
-Because even fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computing, that is running simulations on multiple cores of the computer at the same time to reduce total run time.
-Thirdly, the results of the repeated simulation must be collected and aggregated.
+Second, we must run this function repeatedly with the same sample size and population parameters.
+Because fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computing, that is running simulations on multiple cores of the computer at the same time to reduce total run time.
+Third, the results of the repeated simulation must be collected and aggregated.
 When planning for power, we compute the relative frequency of (a) significant p-value(s) across repeated simulations.
 When planning for precision, we compute the average width of the CI(s).
-Lastly, we have to repeat the complete simulation for different sample sizes, to determine how big the sample must be in order to achieve the targeted power or precision.
+Finally, we have to repeat the complete simulation for different sample sizes to determine how big the sample must be in order to achieve the targeted power or precision.
 
@@
 
-We are finally ready to run our simulation-based sample size planning analyses to plan for power and for precision.
+We are finally ready to run our simulation-based sample size planning analyses to plan for power and precision.
 Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and CIs) in a single command.
@@
 
-Simulation-based sample size planning can quickly become computationally intensive when we repeatedly simulate data and fit models for different parameter combinations or sample sizes.
-Thus, we use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
+We use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
 First, we enable parallelization with the `plan` function and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
@@
 The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
-Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts, the frequency of whether advice is presented and whether it is correct.
+Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts to non-experts, the frequency of whether advice is presented at all, and whether it is correct or incorrect.
 
@@
 Thus in our example, precision is defined as the (long-run) average width of a 95% CI for the probability contrast between experts with correct advice and experts without advice.
-Of course, lower width implies better precision.
+Of course, a lower width implies better precision.
 Based on our simulation outcomes, we compute the precision estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.
@@
 ```
-As should be the case, precision estimates in @fig-finalprecision increase  (i.e., average width of CI decreases) with both the number of subjects and the number of items.
+As should be the case, precision estimates in @fig-finalprecision increase  (i.e., average width of CI decreases) with the number of included subjects and items.
 The CIs reported here indicate how precisely the expected width of the CI for our focal contrast was estimated by our simulation.
-Applying our simple heuristic of targeting an expected width smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.
+Applying our simple heuristic of targeting an expected CI width that is smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.
 We could either choose 100 subjects and 30 items or 200 subjects and 10 items.
 Note that our simple heuristic for determining sample size in the planning for precision scenario was quite liberal.
-This is reflected by the result that we would need a smaller sample size than in the planning for power scenario.
+This is reflected in the result that we would need a smaller sample size than in the planning for power scenario.
 With a more conservative precision target, the result is generally the opposite: As a rule, precise parameter estimates usually require bigger samples than null hypothesis testing.
@@
 
-The goal of this tutorial was to teach researchers how to perform tailored simulation-based sample size planning for GLMMs.
-Beyond the specifics of our concrete case study, we want to outline six developments regarding the future role of simulation-based sample size planning in experimental research:
-
-In light of the ongoing replication crisis and an existing literature full of underpowered studies, there is a growing need for simulation-based sample size planning in experimental research: In order to conduct informative research, GLMMs offer a flexible statistical framework to analyze complex experimental study designs.
-However, existing formula-based heuristics and user-friendly software tools for a priori power analysis are often not sufficient.
-Therefore, simulation-based power analysis is becoming increasingly needed since it provides experimental researchers with a tailored approach to estimating required sample sizes before data collection.
-
-Managing data simulations more easily with discrete predictor variables: Simulation-based sample size planning becomes more manageable when all predictor variables are discrete (like in the presented case study) and fixed by the study design.
+In this tutorial, we provided a step-by-step guide on  how to perform tailored simulation-based sample size planning for GLMMs.
+To conclude, we want to give an outlook on six developments regarding the future role of simulation-based sample size planning in experimental research:
+
+1.
+As experimental designs become more complex and the appropriate flexible statistical frameworks, like GLMMs, become more popular, there is an ever growing need for simulation-based sample size planning in experimental research.
+
+2.
+Managing data simulations more easily with discrete predictor variables.
+Simulation-based sample size planning becomes more manageable when all predictor variables are discrete and fixed by the study design.
 This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
@@
 
-Teaching data simulation skills: The ability to conduct simulation-based sample size planning is a valuable skill that should be taught to experimental researchers.
-By incorporating such training into research methods courses and workshops, researchers can gain a deeper understanding of statistical power or precision, and improve the quality of their experimental designs.
-Equipping researchers with the knowledge and tools to perform simulation-based sample size planning enables them to make informed decisions and enhance the rigor of their studies. 
-The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
-
-Addressing the mismatch in effort perception: There is often a significant disconnect between the amount of effort required to perform tailored simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research.
+3.
+The ability to conduct simulation-based sample size planning becomes an increasingly valuable skill that should be taught to experimental researchers.
+By incorporating such training into research methods courses and workshops, researchers can improve the quality of their experimental designs and enhance the rigor of their studies.
+The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
+
+4.
+Given the significant disconnect between the amount of effort required to perform tailored simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research, there is a need to address the mismatch in effort perception.
 Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-It is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.
-
-Recognizing the value of simulation-based design analysis: Tailored data simulations and power analyses are not mere technicalities; they are valuable research contributions that deserve recognition in experimental research.
-They offer insights into the robustness and sensitivity of experimental designs, helping researchers make informed decisions about sample sizes, effect sizes, and statistical power or precision.
-Their importance can be reflected by allocating them a separate publication or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
-
-Integration with Open Science and preregistration practices: Simulation-based sample size planning aligns well with the principles of Open Science and preregistration in experimental research.
+Therefore, it is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.
+
+5.
+Tailored data simulations and power analyses are not mere technicalities; they are valuable research contributions that deserve adequate recognition in experimental research.
+Their importance can be reflected by highlighting the simulation work in a publication or even allocating them a separate publication, or incorporating them as a significant component of stage 1 preregistered reports [@chambersPresentFutureRegistered2022].
+
+6.
+Simulation-based sample size planning aligns well with the principles of Open Science and preregistration and could be further integrated.
 When researchers have access to simulated data based on their pre-specified model, analyzing the collected dataset becomes straightforward and unambiguous.

### 2024-06-19 manuscript.qmd
more comments

@@
 Multiple easy-to-use software solutions exist for simple statistical models and experimental designs.
-However, many researchers lack the skills and tools to conduct "a priori" (i.e., before data collection) sample size planning for more complex research designs such as flexible generalized linear mixed models (GLMM) framework.
+A priori sample size planning for more complex research designs such as flexible generalized linear mixed models (GLMM) is not covered by standard software solutions.
+Researchers willing to use this framework will need to use data simulation.
 In the present work, we provide a tutorial on how to determine adequate sample sizes by performing tailored simulation-based sample size planning for GLMMs.
@@
 Some knowledge of causal inference is beneficial but not necessary [@lundbergWhatYourEstimand2021; @deffnerCausalFrameworkCrossCultural2022].
-In addition, readers should have an understanding of how to conduct statistical analyses with R [@wickhamDataScienceImport2023] and how to simulate data.
+In addition, readers should have an understanding of how to conduct statistical analyses with R [@wickhamDataScienceImport2023] and how to simulate data [@debruineUnderstandingMixedEffectsModels2021; @hallgrenConductingSimulationStudies2013; @leeUsingTidyversePackage2020].
 For data simulation, we use functions from the tidyverse [@wickhamWelcomeTidyverse2019] and the faux package [@R-faux].
@@
 For these reasons, many journals and funding bodies now require a sample size justification in study protocols and grant proposals, recognizing its relevance in ensuring robust and meaningful findings.
-Although scientists often do not justify sample sizes or use general heuristics from the literature, resource constraints often play an important role [@lakensSampleSizeJustification2022].
-But ideally, a suitable sample size should be determined a priori (i.e., before the study is conducted) based on some meaningful computation to ensure that the study will be able to fulfill its purpose.
+While sample sizes can be justified with resource constraints or general heuristics, statistical arguments based on power or precision are the gold standard [@lakensSampleSizeJustification2022].
 

### 2024-06-20 manuscript.qmd
rewording: change non-expert/no expert to student, false advice/ wrong advice to incorrect advice

@@
 In the present case study, we consider the effectiveness of feedback provided by an artificial intelligence (AI) embedded in a clinical decision support system.
-The context is a clinical setting, where task experts, i.e., radiologists, and non-expert, i.e., medical students, must detect bleedings based on head scans from computer tomography (CT).
+The context is a clinical setting, where task experts (i.e., radiologists), and non-expert (i.e., medical students) must detect bleedings based on head scans from computer tomography (CT).
 The AI model can provide initial diagnostic advice, which can be used as guidance by the humans who are required to make the final diagnostic decision.
@@
 & P(\text{correct diagnosis} | \text{no advice, average expert, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, average expert, average scan})
+& \quad - P(\text{correct diagnosis} | \text{incorrect advice, average expert, average scan})
 \end{aligned}
@@
 & P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan})
+& \quad - P(\text{correct diagnosis} | \text{incorrect advice, average student, average scan})
 \end{aligned}
@@
 When AI advice is given, this advice can be either correct or incorrect.
-The type of advice (no advice, wrong advice, correct advice) is randomized within subjects across CT scans.
+The type of advice (no advice, incorrect advice, correct advice) is randomized within subjects across CT scans.
 After reviewing a CT scan, participants deliver a diagnosis (bleeding or no bleeding), which may be either accurate or inaccurate.
@@
 In all other experimental conditions, some terms drop from the equation because they are multiplied by $0$.
-For example, the probability that a non-expert with average ability solves an item with average difficulty when no advice is presented only requires the intercept:
+For example, the probability that a student with average ability solves an item with average difficulty when no advice is presented only requires the intercept:
 $$
@@
   col1 = 
-    c("no advice, no expert", "no advice, expert", "false advice, no expert", 
-      "false advice, expert", "correct advice, no expert", "correct advice, expert"),
+    c("no advice, student", "no advice, expert", "incorrect advice, student", 
+      "incorrect advice, expert", "correct advice, student", "correct advice, expert"),
   col2 =
@@
   mutate(condition = fct_recode(condition,
- "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
- "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
- "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
+ "student, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+ "student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
+ "student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
   group_by(condition) |>
@@
   mutate(condition = fct_recode(condition,
-"no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-"no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-"no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
+"student, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+"student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
+"student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
   ggplot(aes(x = y_prob, y = condition)) +
@@
 This gives us a better way to see the variability between persons.
-As an example, @fig-margdist2 reveals a number of implicit assumptions about the comparison between experts and non-experts: With wrong advice, virtually all experts have a higher probability of making a correct diagnosis compared to non-experts when considering only items with average difficulty.
-In contrast, there is considerable overlap in probability between experts and non-experts with no advice and even higher overlap with correct advice.
+As an example, @fig-margdist2 reveals a number of implicit assumptions about the comparison between experts and students: With incorrect advice, virtually all experts have a higher probability of making a correct diagnosis compared to students when considering only items with average difficulty.
+In contrast, there is considerable overlap in probability between experts and students with no advice and even higher overlap with correct advice.
 Patterns like these should be considered carefully and discussed with the domain experts.
@@
   mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
+    "student, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+    "student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
+    "student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
   ggplot(aes(x = y_prob, y = condition)) +
@@
 Existing studies indicate that radiologists typically demonstrate high accuracies around 90%, while interns have been shown to perform below 80%, and medical students fall even shorter.
-For simplicity, we assumed plausible probability values of .90 for experts and .70 for non-experts, respectively.
+For simplicity, we assumed plausible probability values of .90 for experts and .70 for students, respectively.
 Our experts confirmed that these values are realistic baselines for reviewing diverse head CT images without AI assistance.
 Subsequently, we consulted several published papers investigating the effect of correct and incorrect advice on decision-making performance in other settings.
-From their findings, we inferred that both experts and non-experts should benefit from correct and suffer losses from incorrect advice.
-However, the magnitude of these effects should be substantially greater for non-experts, given their stronger reliance on advice compared to experts.
+From their findings, we inferred that both experts and students should benefit from correct and suffer losses from incorrect advice.
+However, the magnitude of these effects should be substantially greater for students, given their stronger reliance on advice compared to experts.
 We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
@@
     "b8 = b2",  # (correct advice, expert) - (no advice, expert)
-    "b2 = b6",  # (no advice, expert) - (wrong advice, expert) 
-    "b7 = b1",  # (correct advice, no expert) - (no advice, no expert)
-    "b1 = b5"), # (no advice, no expert) - (wrong advice, no expert)
+    "b2 = b6",  # (no advice, expert) - (incorrect advice, expert) 
+    "b7 = b1",  # (correct advice, student) - (no advice, student)
+    "b1 = b5"), # (no advice, student) - (incorrect advice, student)
     equivalence = c(0, 0))
@@
 The smallest width is observed for the first contrast (expert with correct advice vs. expert without advice), where both underlying probabilities are close to 1.
-The largest width is observed for the fourth contrast (non-expert with wrong advice vs. non-expert without advice), where both underlying probabilities are closer to 0.5.
+The largest width is observed for the fourth contrast (student with incorrect advice vs. student without advice), where both underlying probabilities are closer to 0.5.
 
@@
 The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
-Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts to non-experts, the frequency of whether advice is presented at all, and whether it is correct or incorrect.
+Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts to students, the frequency of whether advice is presented at all, and whether it is correct or incorrect.
 

### 2024-06-20 manuscript.qmd
more comments; R packages italic

@@
 In addition, readers should have an understanding of how to conduct statistical analyses with R [@wickhamDataScienceImport2023] and how to simulate data [@debruineUnderstandingMixedEffectsModels2021; @hallgrenConductingSimulationStudies2013; @leeUsingTidyversePackage2020].
-For data simulation, we use functions from the tidyverse [@wickhamWelcomeTidyverse2019] and the faux package [@R-faux].
+For data simulation, we use functions from the *tidyverse* [@wickhamWelcomeTidyverse2019] and the *faux* package [@R-faux].
 Finally, readers should have a basic understanding of regression modeling and GLMMs [@bolkerLinearGeneralizedLinear2015].
@@
 
-Requirements of real-world studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
+First, requirements of real-world studies are often more complex than the simplified designs assumed by many user-friendly software packages for sample size planning.
 One frequent issue in applied data analysis is missing data [@littleStatisticalAnalysisMissing2014].
@@
 Although GLMMs can handle a large variety of outcome variables, researchers are becoming increasingly aware that many datasets might profit from even more sophisticated models.
-Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects that can be modeled with the R packages glmmTMB [@brooksGlmmTMBBalancesSpeed2017] or brms [@burknerAdvancedBayesianMultilevel2018].
+Common examples are zero-inflated outcomes, censoring, and nonlinear predictor effects that can be modeled with the R packages *glmmTMB* [@brooksGlmmTMBBalancesSpeed2017] or *brms* [@burknerAdvancedBayesianMultilevel2018].
 Tailored simulation-based approaches do not share the same limitations as the existing software solutions for power analysis that focus exclusively on GLMMs.
@@
 
-The most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is a slope or intercept of a regression model.
+Second, the most common hypotheses tested in psychological research are of the type $H_0: \beta = 0$, where $\beta$ is a slope or intercept of a regression model.
 However, many research questions in psychology actually require testing more complex statistical hypotheses.
@@
 
-Although planning for precision [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008; @cummingNewStatisticsWhy2014] has been increasingly discussed as a useful strategy for empirical research, all available software packages for sample size planning with GLMMs are based on power analysis.
+Third, although planning for precision [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008; @cummingNewStatisticsWhy2014] has been increasingly discussed as a useful strategy for empirical research, all available software packages for sample size planning with GLMMs are based on power analysis.
 Therefore, researchers that want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE 1.
@@
 
-All frameworks for sample size planning require the user to make assumptions about the expected effect size.
+Fourth, all frameworks for sample size planning require the user to make assumptions about the expected effect size.
 Assuming the true effect is of this size (or greater), one can compute the (minimum) power of a HT or the (maximum) expected width of a CI.
@@
 
-Although the details differ depending on the specific study characteristics, each tailored simulation-based sample size planning requires a series of steps and decisions.
+Although the details differ for each study, every tailored simulation-based sample size planning requires a series of steps and decisions.
 We will introduce each step in a theoretical section, followed by the practical application based on a case study.
-All code in this manuscript and simulation results are available in the project’s repository on the Open Science Framework (<https://osf.io/dhwf4/>).
+All code in this manuscript and simulation results are available on the Open Science Framework (<https://osf.io/dhwf4/>) and on our tutorial website (<https://timo-ko.github.io/glmm_simulation_tutorial/>).
 
@@
 
-In the first four lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data.^[The faux package [@R-faux] contains useful functions when simulating factorial designs, including random effects.]
+In the first four lines of the function definition, we set some default parameter values (which we will explain in the next section) and load the packages we use to manipulate and simulate data.^[The *faux* package [@R-faux] contains useful functions when simulating factorial designs, including random effects.]
 In our case study, each subject (`n_subjects` in total) is assumed to respond to each item (i.e., CT scan; `n_items` in total).
@@
 The next step is to specify how the statistical model shall be estimated in the actual study collected later.
-This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the lme4 R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
+This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the *lme4* R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
 
@@
 
-In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.^[For Bayesian GLMMs, the brms R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].]
+In our case study, we use the *lme4* R package [@batesFittingLinearMixedEffects2015], which is a state-of-the-art tool for fitting frequentist GLMMs.^[For Bayesian GLMMs, the *brms* R package is currently the most prominent option [@burknerBrmsPackageBayesian2017].]
 For the current example, we simulate data according to our model, in which 100 subjects respond to 50 items (we use `set.seed` to make the simulation reproducible).
@@
 
-The lme4 package uses a special syntax for model specification. Our specific GLMM is represented by the formula:
+The *lme4* package uses a special syntax for model specification. Our specific GLMM is represented by the formula:
 
@@
 The third line specifies a random intercept for each subject `(1|subject)` and for each item `(1|item)`.
-The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the lme4 package.
-
-In lme4, a GLMM is fitted with the `glmer` function.
+The complete set of rules for the syntax is outlined in @batesFittingLinearMixedEffects2015 and in the documentation of the *lme4* package.
+
+In *lme4*, a GLMM is fitted with the `glmer` function.
 By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs.
@@
 
-In the lme4 output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
+In the *lme4* output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
 
@@
 Beyond computing the point estimate, we have already discussed that both hypothesis testing and interval estimation can be used to answer the research question.
-The decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with the marginaleffects R package using the delta method).
+The decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with the *marginaleffects* R package using the delta method).
 
@@
 However, this would be tedious and we can use R to compute this contrast without doing the math.
-Using the first contrast *(correct advice, expert) - (no advice, expert)* as our example, we could apply the `predict` function of the lme4 package to compute the predicted probability for a correct diagnosis based on our fitted model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
+Using the first contrast *(correct advice, expert) - (no advice, expert)* as our example, we could apply the `predict` function of the *lme4* package to compute the predicted probability for a correct diagnosis based on our fitted model, plug in the two sets of predictor values, and compute the difference between the two estimated probabilities.
 
@@
 However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute HTs or CI.
-The marginaleffects package [@R-marginaleffects] is a very flexible, increasingly popular package to compute HTs and CIs for contrasts with a variety of statistical models, including GLMMs estimated with lme4.
+The *marginaleffects* package [@R-marginaleffects] is a very flexible, increasingly popular package to compute HTs and CIs for contrasts with a variety of statistical models, including GLMMs estimated with *lme4*.
 First, we specify a grid of all combinations of predictor variables and then compute estimated probabilities for all experimental conditions in our experiment with the `predictions` function:
@@
 We will use the `hypotheses` function to compute our four contrasts of interest together with HTs and CIs.
-We use the default inference options of the marginaleffects package that compute HTs and CIs based on the approximate delta method.
+We use the default inference options of the *marginaleffects* package that compute HTs and CIs based on the approximate delta method.
 
@@
 
-We use the future [@R-RJ-2021-048] and furrr [@R-furrr] packages to perform computations in parallel.
+We use the *future* [@R-RJ-2021-048] and *furrr* [@R-furrr] packages to perform computations in parallel.
 First, we enable parallelization with the `plan` function and specify how many parallel cores ("workers") of our computer to use (users can find out the maximum number of cores on their computer with the command `parallel::detectCores()`), and set a seed to make the simulation reproducible.
@@
 The result of this computation is a data frame that contains the p-values and CIs of all specified contrasts for each simulated dataset.
-In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the lme4 package.
+In some iterations (predominantly in conditions with small sample sizes), model estimation did not converge with the *lme4* package.
 When the model fails to converge, it means that the statistical model being fitted to the data failed to reach a stable or valid solution during the estimation process.
@@
 For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the $\alpha = 0.05$ level.
-Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.^[This code was inspired by the "Mixed Design Simulation" vignette of the faux package at <https://debruine.github.io/faux/articles/sim_mixed.html>.]
+Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.^[This code was inspired by the "Mixed Design Simulation" vignette of the *faux* package at <https://debruine.github.io/faux/articles/sim_mixed.html>.]
 
@@
 
-## Sensitivity analysis
+## Step 7 (optional): Sensitivity analysis
 

### 2024-06-20 manuscript.qmd
more comments; add pwr package for citation; update lock file; update package versions with renv

@@
 library("papaja")
+library("pwr")
 r_refs("r-references.bib", append = FALSE)
@@
 As a result, there is a growing need for more rigorous statistical methodology and the adoption of adequately powered experimental designs.
-Multiple easy-to-use software solutions exist for simple statistical models and experimental designs.
+Multiple easy-to-use software solutions exist for simple statistical models and experimental designs [@R-pwr; @lakensSimulationBasedPowerAnalysis2021].
 A priori sample size planning for more complex research designs such as flexible generalized linear mixed models (GLMM) is not covered by standard software solutions.
@@
 Not all research questions are best answered by hypothesis testing.
-It has been argued that basic research rarely requires discrete decisions on whether some effect has been "discovered" and should thus shift from hypothesis testing towards an estimation framework [@mcelreathStatisticalRethinkingBayesian2020; @cummingNewStatisticsWhy2014; @kruschkeBayesianNewStatistics2018].
-Although this view is not without critique [@uyguntuncEpistemicPragmaticFunction2023], at least for exploratory or pilot studies with little prior research,  estimating the effects of interest and making the estimation uncertainty transparent by reporting CIs seems more useful.
+It has been argued that basic research rarely requires discrete decisions on whether some effect has been "discovered" and should thus shift from hypothesis testing towards an estimation framework [@mcelreathStatisticalRethinkingBayesian2020; @cummingNewStatisticsWhy2014; @kruschkeBayesianNewStatistics2018], although this view is not without critique [@uyguntuncEpistemicPragmaticFunction2023].
 Assuming that no HTs are conducted, power analysis is not relevant for sample size planning.
-Nonetheless, the sample size still has a crucial effect on how informative the study will be because an effect of interest is estimated more precisely with bigger samples.
 In the precision framework, the target quantity commonly used for sample size planning is the expected width of a CI [@kelleySampleSizePlanning2006; @lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
@@
 Thus, a narrow CI with fewer plausible values for the quantity of interest is more informative about the size of the true effect than a wide CI.
-Apart from the confidence level, the width of a CI depends strongly on the sample size.
+Apart from the confidence level, the width of a CI depends on the sample size.
 Because bigger samples carry more information, they lead to smaller CIs.
 When planning for precision, a target can be set for the expected width of a CI of interest.
-Assuming some effect size of interest and a desired confidence level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
+Assuming some effect size of interest and a certain confidence level, a minimum sample size can be determined that would guarantee reaching the targeted expected width.
 Because planning for precision is still rare, there are no common heuristics for choosing the desired width of the CI [@lakensSampleSizeJustification2022].
@@
 
-### Insightful descriptive statistics
+### Examine insightful descriptive statistics
 
@@
 We tried using these descriptive statistics to judge together with domain experts whether our chosen values for the random effect standard deviations would produce data that aligned with their domain expertise.
-Although the result was deemed plausible, these statistics were not informative enough to determine a final set of plausible parameter values.
-For this reason, we will additionally look at insightful model- based quantities.
-
-### Insightful model based quantities
+Although the result was deemed plausible, these statistics were not informative enough to determine a final set of plausible parameter values (e.g., doubling the standard deviations from 0.5 to 1 produces only minor changes in relative frequency).
+For this reason, we will additionally look at insightful model-based quantities.
+
+### Examine insightful model-based quantities
 
@@
 This is especially true for the standard deviation of random effects that have non-linear (and often unexpected) effects on the model-implied results.
-An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model based quantities that can never be directly observed in an actual empirical dataset.
+An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model-based quantities that can never be directly observed in an actual empirical dataset.
 For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant, on average, solves a cognitive task.
@@
 
-### Iterative process with domain experts
+### Iterate with domain experts
 
@@
 
-Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute HTs or CIs for estimands of interest, mirroring the analysis that will later be performed for the actual dataset of the planned study.
-The last missing piece is to write code that performs the previous steps repeatedly and allows for setting different sample sizes.
-On a conceptual level, we first require a function that takes the sample size and the full set of population parameter values as input.
+Conducting all previous steps enables the analyst to 1) simulate a dataset, 2) estimate a GLMM, and 3) compute HTs or CIs for estimands of interest, mirroring the analysis that will later be conducted on the actual dataset of the planned study.
+To produce an estimate of power or precision, the last step is to perform these previous steps repeatedly.
+On a conceptual level, we first require a function that takes a sample size and a full set of population parameter values as input.
 When planning for power, the function should return the p-value(s) of the HT(s) of interest when conducted on the simulated dataset.
 When planning for precision, the function should return the width of the CI(s) of interest.
-Second, we must run this function repeatedly with the same sample size and population parameters.
-Because fitting GLMMs with frequentist methods can quickly become time-consuming, it is recommended to use parallel computing, that is running simulations on multiple cores of the computer at the same time to reduce total run time.
-Third, the results of the repeated simulation must be collected and aggregated.
+Second, we run this function repeatedly with the same sample size and population parameters.
+Because fitting GLMMs can quickly become time-consuming, it is recommended to use parallel computing, that is running simulations on multiple cores of the computer at the same time to reduce total run time.
+Third, the results of the repeated simulation are collected and aggregated.
 When planning for power, we compute the relative frequency of (a) significant p-value(s) across repeated simulations.
 When planning for precision, we compute the average width of the CI(s).
-Finally, we have to repeat the complete simulation for different sample sizes to determine how big the sample must be in order to achieve the targeted power or precision.
+Finally, we repeat the complete simulation for different sample sizes to determine how big the sample must be in order to achieve the targeted power or precision.
 
@@
 
-We are finally ready to run our simulation-based sample size planning analyses to plan for power and precision.
 Wrapping the `simulate` function already constructed earlier, the helper function `sim_and_analyse` performs all previous steps (simulate a dataset, fit a GLMM, compute p-values and CIs) in a single command.
@@
 The next code chunk specifies a simulation grid with different settings for both the number of subjects (`n_subjects`) and the number of items (`n_items`), each combination being repeated `rep` times.
-We chose 300 repetitions for the data simulation at hand as it strikes a balance between achieving a robust statistical estimate and remaining computationally feasible.
+The more repetitions, the more accurately power and precision can be estimated.
+We chose 300 repetitions for the data simulation at hand to strike a balance between achieving a robust estimate and remaining computationally feasible.
 With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.

### 2024-06-20 manuscript.qmd
cut example from glmm section

@@
 The link function allows for modeling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the outcome variable.
-For example, think of an experiment with different design factors (e.g., picture positions, headline aesthetics) impacting the likelihood of users clicking on an online advertisement.
+
+<!-- For example, think of an experiment with different design factors (e.g., picture positions, headline aesthetics) impacting the likelihood of users clicking on an online advertisement.
 The participants’ behavior is measured repeatedly over several sessions.
 The click patterns of participants in one session are likely to be correlated with their previous sessions, and the outcome variable is binary (click/no click) for each session, which follows a binomial distribution.
+-->
 

### 2024-06-20 manuscript.qmd
more changes

@@
 They allow for the modeling of fixed effects, which capture the relationships between the usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.^[Correlated random effects can be assumed in models that contain both random intercepts and random slopes.]
 GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.
@@
 Simulation-based power analysis with GLMMs requires making a range of assumptions: The distribution assumption specifies the distributional family of the outcome variable, after conditioning on all predictor variables.
-Assumptions about the random effects include the assumption of normality (i.e., that the random effects follow a normal distribution) and the covariance structure among the random effects (i.e., if and how they are correlated).
+Assumptions about the random effects include the assumption of normality and the covariance structure among the random effects.
 Making these decisions requires understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
@@
 
-# When to use tailored data simulation?
+# Reasons to use tailored data simulation
 

### 2024-06-21 manuscript.qmd
more comments

@@
 When introducing the simulation function for our case study, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$) but have not talked about where these numbers came from.
-
 The starting point for all parameter values in our present case study was based on results from distantly related study designs in the literature.
@@
 
-Gathering domain knowledge by domain experts and the specification of population parameter values is an iterative process.
+Gathering domain knowledge and the specification of population parameter values is an iterative process, in which domain experts repeatedly assess the plausibility of simulated data.
 As a first step, domain experts can be interviewed to “elicit” their domain knowledge about how the data of the planned study is expected to look.
@@
 Instead of extracting this expected effect size from meta-analyses or pilot data, which has been the main focus of previous tutorials [e.g., @kumleEstimatingPowerGeneralized2021], we have demonstrated some strategies to determine plausible parameter values in GLMMs based on domain knowledge.
-Domain knowledge can be considered a vague theoretical model about the data-generating process that is less formal and can only be accessed by a back-and-forth exchange in which domain experts assess the plausibility of simulated data.
 When sample sizes are chosen based on the results of our simulation-based power analysis, a future study will be informative to reject the null hypothesis if an effect of our *expected size* is present (or estimate the effect with satisfying precision).

### 2024-06-21 manuscript.qmd
work on step 4

@@
 The next step is to specify how the statistical model shall be estimated in the actual study collected later.
-This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the *lme4* R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
+This requires the selection of a statistical framework, a software package that is capable of estimating the model class of interest, an estimation algorithm, and the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.
 
@@
 inaccurate (0) diagnosis.
+We use the default estimation algorithm (see `?glmerControl` for a list of alternative options).
 
@@
 
-In the *lme4* output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
+In the model output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
 

### 2024-06-21 manuscript.qmd
incorporate discrete predictor outlook into step 2

@@
 This requires specifying a generative process for all predictor variables used in the final data analysis.
-While such assumptions can be quite challenging for observational studies or continuous predictor variables, this is less of a problem for experimental studies with only categorical predictor variables.
+Realistic assumptions can be quite challenging for observational studies or continuous predictor variables, which is beyond the scope of this tutorial.
+However, this step is much easier for experimental studies with only categorical predictor variables because the distribution of predictors is fixed by the study design.
 When all predictor variables have been simulated, one can use the structure of a suitable GLMM to simulate the dependent variable.
@@
 
-In this tutorial, we provided a step-by-step guide on  how to perform tailored simulation-based sample size planning for GLMMs.
-To conclude, we want to give an outlook on six developments regarding the future role of simulation-based sample size planning in experimental research:
+In this tutorial, we provided a step-by-step guide on how to perform tailored simulation-based sample size planning for GLMMs based on a concrete case study.
+To conclude, we want to give an outlook on five developments regarding the future role of simulation-based sample size planning in experimental research:
 
@@
 2.
-Managing data simulations more easily with discrete predictor variables.
-Simulation-based sample size planning becomes more manageable when all predictor variables are discrete and fixed by the study design.
-This allows researchers to focus on simulating outcome variables while avoiding the need for complex simulations of predictor values, which would introduce additional assumptions.
-By simplifying the simulation process, researchers can obtain reliable estimates for power or precision without compromising realistic assumptions about the data-generating process implied by the study design.
-
-3.
 The ability to conduct simulation-based sample size planning becomes an increasingly valuable skill that should be taught to experimental researchers.
@@
 
-4.
+3.
 Given the significant disconnect between the amount of effort required to perform tailored simulation-based sample size planning and the perceived effort estimated by researchers and collaborators in experimental research, there is a need to address the mismatch in effort perception.
 Many researchers request simulation-based power analyses from statisticians or methodological experts without fully comprehending the complexity and time-consuming nature of these tailored simulations.
-Therefore, it is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration between researchers and methodological experts.
-
-5.
+Therefore, it is crucial to raise awareness about the effort involved to ensure realistic expectations and effective collaboration with methodological experts.
+
+4.
 Tailored data simulations and power analyses are not mere technicalities; they are valuable research contributions that deserve adequate recognition in experimental research.
@@
 
-6.
+5.
 Simulation-based sample size planning aligns well with the principles of Open Science and preregistration and could be further integrated.

### 2024-06-21 manuscript.qmd
add more references to glmm section

@@
 For data simulation, we use functions from the *tidyverse* [@wickhamWelcomeTidyverse2019] and the *faux* package [@R-faux].
-Finally, readers should have a basic understanding of regression modeling and GLMMs [@bolkerLinearGeneralizedLinear2015].
+Finally, readers should have a basic understanding of regression modeling and GLMMs [@brownIntroductionLinearMixedEffects2021].
 In this tutorial, we simulate data by manually specifying the model equation of a GLMM that represents our assumed data-generating process [@debruineUnderstandingMixedEffectsModels2021].
@@
 As study designs become more complex, psychological researchers require more sophisticated statistical models to capture their nuanced relationships and grouping structures [@yarkoniGeneralizabilityCrisis2022].
-GLMMs (also called multilevel models) are gaining popularity because they offer great flexibility [@fahrmeirRegressionModelsMethods2021].
-GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data, including hierarchical structures [@fahrmeirRegressionModelsMethods2021].
+GLMMs (also called multilevel models) are gaining popularity because they offer great flexibility when applied carefully [@brownIntroductionLinearMixedEffects2021; @matuschekBalancingTypeError2017; @meteyardBestPracticeGuidance2020a].
+GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of linear regression models that account for correlated data, including hierarchical structures [@fahrmeirRegressionModelsMethods2021; @bolkerLinearGeneralizedLinear2015].
 In this context, correlated data means that the value in the outcome variable for one observation may be related to the value of another observation in a systematic way that is not already accounted for by the usual (fixed) predictor variables (e.g., the age of participants).
@@
 They allow for the modeling of fixed effects, which capture the relationships between the usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects.
-Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.^[Correlated random effects can be assumed in models that contain both random intercepts and random slopes.]
+Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across groups.
+Correlated random effects can be assumed in models that contain both random intercepts and random slopes.
 GLMMs extend the LMM framework to accommodate non-normally distributed continuous and categorical outcome variables.

### 2024-06-21 manuscript.qmd
iterate on glmm and simulation-based sections

@@
 Thus, modeling such correlations is important whenever the data has a clear structure, while the grouping variables can be hierarchically nested (e.g., grouping variables students and schools: each student belongs to exactly one school) or cross-classified (e.g., grouping variables students and math exercises: each student is presented with several math exercises).
-LMMs are used when the outcome variable is continuous and follows a normal distribution.
+LMMs are used when the outcome variable is continuous and follows a normal distribution (after conditioning on all predictor variables).
 They allow for the modeling of fixed effects, which capture the relationships between the usual predictors and the outcome, as well as random effects, which account for the different types of correlation structure and grouping effects.
@@
 To our knowledge, existing approaches for sample size planning for GLMMs have exclusively focused on planning for power.
-Power analysis methods for multilevel models can be categorized into formula-based methods and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
-Formula-based methods rely on formulas to calculate power directly, while simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically (i.e., what percentage of simulated datasets produces a significant p-value).
-Available formula-based software packages for multilevel models often do not include GLMMs or are limited to simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
+In Table 1, we review available software packages that can be used to perform power analysis for multilevel models in `r cite_r()`.
+Power analysis methods can be categorized into formula-based, summary-statistics-based and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
+Formula-based methods rely on exact formulas to calculate power directly.
+Summary-statistics-based methods use statistical theory to approximate power based on formula-based methods developed for simple t-tests.
+Simulation-based methods rely on repeatedly simulating data with a known true effect size and estimating power empirically, that is what percentage of simulated datasets produces a significant p-value.
+Available formula-based and summary-statistics-based software packages for multilevel models often do not include GLMMs or are limited to simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
 A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a].
-However, many of these tutorials focus on linear mixed models (LMMs) and the most common study designs [see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
+However, many of these tutorials focus on LMMs and the most common study designs [see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
 This narrow focus provides limited guidance for researchers using more complex study designs, especially when little prior knowledge about plausible effect sizes is available [see the discussion in @kumleEstimatingPowerGeneralized2021].
-Simulation-based power analysis with GLMMs requires making a range of assumptions: The distribution assumption specifies the distributional family of the outcome variable, after conditioning on all predictor variables.
-Assumptions about the random effects include the assumption of normality and the covariance structure among the random effects.
-Making these decisions requires understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
+Simulation-based power analysis with GLMMs requires making a range of assumptions about the model structure that should align with the characteristics of the data being analyzed.
 Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
 However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge (see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021).
+We will discuss a number of strategies on how to specify model parameters for application-specific, tailored data simulations in a later chapter.
 
 INSERT TABLE 1 HERE!
-
-In Table 1, we review existing R packages that can be used for power analysis for GLMMs.
 

### 2024-06-25 manuscript.qmd
justify target power

@@
 When narrowing down the best combination that achieves sufficient power while at the same time striking a good balance of how many subjects and items are practically feasible, later rounds of data simulation will typically include a smaller grid of sample sizes combined with a higher number of repetitions.
-This will assure high precision for the final power estimates, which are then used for the sample size justification of the future study.
-
-Much has been written on the optimal amount of power to target in empirical research.
-The most prominent heuristic is to target a power of 0.8 (when combined with a type I error rate of $\alpha = 0.05$), but depending on the research goals of the study, there are often good reasons to move away from this standard depending on the research goals and resource constraints  [@lakensJustifyYourAlpha2018; @lakensSampleSizeJustification2022].
+This will assure high precision for the final power estimates, which are then used for the sample size justification of the planned study.
+
+The most prominent heuristic in sample size planning is to target a power of 0.8 in combination with a type I error rate of $\alpha = 0.05$ [@lakensSampleSizeJustification2022].
+However, depending on the research goals or resource constraints, there are often good reasons to move away from this standard [@benjaminRedefineStatisticalSignificance2017; @lakensJustifyYourAlpha2018].^[For example, the *Social Sciences Replication Project* targeted a power of 0.9 to safeguard against biased effect sizes in the original studies [@camererEvaluatingReplicabilitySocial2018]].
 When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
@@
 However, because the CIs of the four contrasts strongly differ in width, it is not trivial to decide which width one should target when deciding on the appropriate sample size.
-In contrast to planning for power, there are no common standards on how to specify the targeted precision.
+In contrast to planning for power, there are no common standards on how to specify the targeted precision [@lakensSampleSizeJustification2022].
 For our example, we use a simple heuristic but we strongly encourage readers to think about better alternatives that are appropriate in their own applications.

### 2024-06-25 manuscript.qmd
update sensitivity analysis section

@@
 An effect smaller than the SESOI would be considered too small to be interesting or practically meaningful, even if the effect is not actually zero [@kingPointMinimalImportant2011; @lakensEquivalenceTestingPsychological2018].
-For strategies on the even more difficult task of specifying a plausible SESOI, as well as a thorough discussion of various topics concerning power analysis, see [@lakensImprovingYourStatistical2022].
-When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power or precision for different effect sizes in what is called *sensitivity power analysis*.
+Unfortunately, specifying a plausible SESOI is a challenging task (for strategies on how to specify a SESOI, see @lakensSampleSizeJustification2022, @riesthuisSimulationBasedPowerAnalyses2024, or @lakensImprovingYourStatistical2022).
+When domain knowledge or formal theories about the research topic of interest are too vague to specify a meaningful SESOI, it is still recommended to demonstrate power or precision for different effect sizes in a *sensitivity power analysis* [@lakensSampleSizeJustification2022].
 By simulating power (or precision) for different effect sizes (in addition to the different number of subjects and items), one can make sure that power (or precision) would still be sufficient to detect smaller effect sizes than our expected effect or at least get an impression of how strongly power (or precision) depends on the size of the true effect.
@@
 Power (or precision) could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
+Although recommended, we do not present a sensitivity analysis for our case study to keep the length of the tutorial manageable.
 

### 2024-06-25 manuscript.qmd
shorten missing data sentence

@@
 One frequent issue in applied data analysis is missing data [@littleStatisticalAnalysisMissing2014].
-Data can be missing completely at random (e.g., because an electronic measurement device randomly failed for some technical reasons).
-Alternatively, subjects might systematically drop out or produce missing data, but this dropout can be explained by some attributes also measured in the dataset (e.g., older subjects have a higher probability to refuse answering a question on income).
-In a more complicated scenario, missing data in some variables is caused by the measured attribute itself (e.g., wealthy people are more likely to refuse to report their income).
+Missings can be completely random (e.g., an electronic measurement device randomly failed for technical reasons), depend on some attributes also measured in the dataset (e.g., older subjects refuse answering a question on income), or be caused by the measured attribute itself (e.g., wealthy people refuse to report their income).
 Moreover, many experimental designs contain conditions in which values of the predictor variables are missing by design.

### 2024-06-25 manuscript.qmd
adjust planning for precision reason section

@@
 It has been argued that basic research rarely requires discrete decisions on whether some effect has been "discovered" and should thus shift from hypothesis testing towards an estimation framework [@mcelreathStatisticalRethinkingBayesian2020; @cummingNewStatisticsWhy2014; @kruschkeBayesianNewStatistics2018], although this view is not without critique [@uyguntuncEpistemicPragmaticFunction2023].
-Assuming that no HTs are conducted, power analysis is not relevant for sample size planning.
+When no HTs are conducted, power analysis is not relevant for sample size planning.
 In the precision framework, the target quantity commonly used for sample size planning is the expected width of a CI [@kelleySampleSizePlanning2006; @lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008].
@@
 
-Third, although planning for precision [@lakensSampleSizeJustification2022; @maxwellSampleSizePlanning2008; @cummingNewStatisticsWhy2014] has been increasingly discussed as a useful strategy for empirical research, all available software packages for sample size planning with GLMMs are based on power analysis.
-Therefore, researchers that want to apply an estimation strategy in their studies, without testing any statistical hypotheses, currently cannot use the software packages outlined in TABLE 1.
+Third, all available software packages for sample size planning with GLMMs are based on power analysis and do not support planning for precision.
+Thus, researchers that want to apply an estimation strategy instead of testing statistical hypotheses [@cummingNewStatisticsWhy2014], currently cannot use the software packages outlined in TABLE 1.
+However, tailored simulation-based sample size planning can easily handle the planning for precision approach [@maxwellSampleSizePlanning2008].
+The only change in procedure is that instead of computing HTs for each simulated dataset and estimating statistical power across repetitions, CIs are computed for each simulated dataset, and their expected width is estimated.
 <!-- As long as one is interested only in confidence intervals of a single model parameter, it might be possible to adapt existing software packages to plan for precision.
@@
 -->
-However, tailored simulation-based sample size planning can easily handle the planning for precision approach [@maxwellSampleSizePlanning2008].
-The only change in procedure is that instead of computing HTs for each simulated dataset and estimating statistical power across repetitions, CIs are computed for each simulated dataset, and their expected width is estimated.
 

### 2024-06-25 manuscript.qmd
rephrase standardized effect size reason

@@
 This problem is further exacerbated by the fact that GLMMs are so flexible that general heuristics of what should be considered a small effect do not exist or are difficult to defend.
-In our experience, using domain knowledge to construct a tailored data simulation is the only solution to determine plausible effect sizes in the absence of prior evidence.
-It would be possible to use these tailored simulations to extract plausible values for standardized effect sizes that could then be inserted in existing software packages for sample size planning.
-However, when tailored data simulations are necessary to determine effect sizes anyway, performing the whole sample size planning in a customized way is preferred over using the existing software packages.
+In the absence of prior evidence, using domain knowledge to construct a tailored data simulation is considered the best solution to determine plausible effect sizes [see scenario 3 in @kumleEstimatingPowerGeneralized2021].
+It would be possible to use these tailored simulations to compute standardized effect sizes that could then be inserted in existing software packages for sample size planning.
+However, we would argue that when tailored data simulations are necessary to determine effect sizes anyway, performing the whole sample size planning in a customized way is preferred over using the existing software packages.
 

### 2024-06-25 manuscript.qmd
update population parameter setting

@@
 
-In the absence of previous studies with the same design or pilot data, researchers require strategies on how to specify the population parameters used in their data simulation.
 Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
-In contrast to non-hierarchical linear regression, common heuristics based on standardized effect sizes are less useful or not even available for GLMMs.
-Our strategies to specify population parameters will require access to domain knowledge from domain experts.
+In the absence of previous studies with the same design or pilot data, strategies to specify population parameters require access to domain knowledge from domain experts.
 Because most domain knowledge can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
 The basic idea of all strategies is how the data-generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables calibration of population parameters based on the available knowledge of domain experts.
-
-Although we use frequentist model estimation in our tutorial, many of the strategies described in this chapter are inspired by research on monitoring the plausibility of model assumptions in applied Bayesian statistics [@gelmanBayesianWorkflow2020].
+Although we use frequentist model estimation in our tutorial, many strategies demonstrated in this chapter are inspired by research on monitoring the plausibility of model assumptions in applied Bayesian statistics [@gelmanBayesianWorkflow2020].
 

### 2024-06-25 manuscript.qmd
update power results section

@@
 Assuming an effect size of interest and a desired significance level, a minimum sample size can be determined that, on average, would guarantee reaching this target.
-Although it is recommended to justify the desired significance level and power [@lakensJustifyYourAlpha2018], most empirical studies adopt the heuristic of $\alpha = 0.05$ and $1 - \beta = 0.80$.
+The most prominent heuristic is to target a power of $1- \beta = 0.8$ in combination with a type I error rate of $\alpha = 0.05$ [@lakensSampleSizeJustification2022].
+However, depending on the research goals or resource constraints, there are often good reasons to move away from this standard [@benjaminRedefineStatisticalSignificance2017; @lakensJustifyYourAlpha2018].
+For example, the *Social Sciences Replication Project* targeted a power of 0.9 to safeguard against biased effect sizes in the original studies [@camererEvaluatingReplicabilitySocial2018].
 
@@
 
-The most prominent heuristic in sample size planning is to target a power of 0.8 in combination with a type I error rate of $\alpha = 0.05$ [@lakensSampleSizeJustification2022].
-However, depending on the research goals or resource constraints, there are often good reasons to move away from this standard [@benjaminRedefineStatisticalSignificance2017; @lakensJustifyYourAlpha2018].^[For example, the *Social Sciences Replication Project* targeted a power of 0.9 to safeguard against biased effect sizes in the original studies [@camererEvaluatingReplicabilitySocial2018]].
 When target power has been specified, the number of subjects and the number of items in our study design can be traded against each other based on practical considerations.
-For the sake of the example, let the targeted power be indeed about 0.8, using an $\alpha$ of 0.05 to detect an effect of the expected size implied by our data simulation.
+Although it is recommended to justify the power [@lakensSampleSizeJustification2022], we adopt the common heuristic to target a power of 0.8 to detect an effect of the expected size implied by our data simulation.
 This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice).

### 2024-06-25 manuscript.qmd
work on last comment

@@
 Because most domain knowledge can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
-The basic idea of all strategies is how the data-generating process implied by certain values of population parameters can be quantified or visualized in an intuitive way that enables calibration of population parameters based on the available knowledge of domain experts.
+The basic idea of all strategies is to quantify or visualize the data-generating process implied by certain values of population parameters in an intuitive way that enables calibration of population parameters based on the available knowledge of domain experts.
 Although we use frequentist model estimation in our tutorial, many strategies demonstrated in this chapter are inspired by research on monitoring the plausibility of model assumptions in applied Bayesian statistics [@gelmanBayesianWorkflow2020].
@@
 When introducing the simulation function for our case study, we have used theoretically plausible values as defaults for all model parameters ($\beta_0$, $\beta_e$, $\beta_a$, $\beta_c$, $\beta_{ea}$, $\beta_{ec}$, $\sigma_S$, and $\sigma_I$) but have not talked about where these numbers came from.
-The starting point for all parameter values in our present case study was based on results from distantly related study designs in the literature.
+The starting point for all parameter values was based on results from distantly related study designs in the literature.
 Additionally, we had repeated discussions with our affiliated domain experts in radiology to check whether our assumptions regarding participants’ diagnostic performance seemed plausible.
@@
 
-The mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters would be specified.
-The knowledge of how to simulate from a GLMM enables us to compute insightful descriptive statistics that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
+The mathematical structure of GLMMs determines which patterns in data would be produced by the model, if a specific set of values for the population parameters is specified.
+The knowledge of how to simulate from a GLMM enables us to compute *insightful descriptive statistics* that can be compared to available domain knowledge much more easily than the opaque values of model parameters.
 For example, domain experts might not be able to directly choose plausible values for the  coefficients in a logistic regression model (which are measured on the log-odds scale).
-However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental conditions, i.e., which relative frequency they expect to observe (on average).
-The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model-implied value of the insightful descriptive statistic expected by the domain expert.
+However, they should be able to reason about the expected ratio of the binary dependent variable in different experimental conditions, i.e., which relative frequency they expect to observe.
+The job of the analyst who is familiar with the mathematical structure of the GLMM is to produce the model-implied value of the insightful descriptive statistic that is expected by the domain expert.
 Although insightful descriptive statistics usually depend on the model parameters in a non-linear way, it is not necessary to solve the exact relationship mathematically.
@@
 Although these derivations are straightforward, it is important not to misinterpret their implications: In binomial GLMMs, the average probability to solve an item (averaged across persons of varying ability and items of varying difficulty) is **not** equal to the probability that a person with average ability solves an item with average difficulty [@fahrmeirRegressionModelsMethods2021].
-The first perspective implies a so-called marginal interpretation, while the second one implies a conditional interpretation.
+The first perspective implies a so-called *marginal interpretation*, while the second one implies a *conditional interpretation*.
 
@@
 We want to use the model-implied observed proportion of correct diagnoses in each experimental condition as an insightful descriptive statistic to determine plausible values for the random effect standard deviations.
-
 We will simulate a large dataset (for which the observed values of the descriptive statistic will be close to their model-implied true values) and simply compute the relative frequency of correct diagnoses for each experimental condition.
@@
 
-We tried using these descriptive statistics to judge together with domain experts whether our chosen values for the random effect standard deviations would produce data that aligned with their domain expertise.
+We tried using these descriptive statistics to judge together with domain experts whether our chosen values for the random effect standard deviations would produce data that aligned with their expectations.
 Although the result was deemed plausible, these statistics were not informative enough to determine a final set of plausible parameter values (e.g., doubling the standard deviations from 0.5 to 1 produces only minor changes in relative frequency).
@@
 This is especially true for the standard deviation of random effects that have non-linear (and often unexpected) effects on the model-implied results.
-An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce insightful model-based quantities that can never be directly observed in an actual empirical dataset.
+An important advantage of data simulation (where one has full control over parameter values and sample sizes) is that one can produce *insightful model-based quantities* that can never be directly observed in an actual empirical dataset.
 For example, in a logistic model with random intercepts for participants, one can produce a visualization of the implied distribution of the probability that a participant, on average, solves a cognitive task.
@@
 @fig-margdist1 shows the model-implied marginal distributions, including the mean, 66% and 95% intervals.
-We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered in the previous section.
+We can see that, indeed, the average probabilities (black dots) slightly differ from the probabilities of average subjects and items considered previously.
 This difference increases with the variability of the random effects.
@@
 Then, the population parameters are updated until the domain experts are satisfied with the result.
-During this process, the monitored model-based quantities and descriptive statistics can be updated or extended to capture as much available domain knowledge as possible.
+During this process, the monitored descriptive statistics and model-based quantities can be updated or extended to capture as much available domain knowledge as possible.
 
@@
 However, the magnitude of these effects should be substantially greater for students, given their stronger reliance on advice compared to experts.
-We further validated the plausibility of our estimated gains and losses with the collaborating radiologists.
-For our simulation, we used the probabilities of average participants to solve an average case, as shown in Table 1.
+We further validated the plausibility of our estimated gains and losses with the collaborating radiologists until we settled on the probabilities shown in @tbl-probtable.
 
@@
 
-At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset to be collected in the planned study.
+At this point, the researcher is capable of producing a simulated dataset that is comparable to the actual dataset to be collected in the planned study.
 The next step is to specify how the statistical model shall be estimated in the actual study collected later.
@@
 For some research questions, the estimate corresponds with a single regression coefficient. In more complicated scenarios, the estimate is computed from a combination of coefficients.
-Beyond computing the point estimate, we have already discussed that both hypothesis testing and interval estimation can be used to answer the research question.
-The decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with the *marginaleffects* R package using the delta method).
+Beyond computing the point estimate, we have already discussed that both hypothesis testing and interval estimation could be used to answer the research question.
+The decision on testing or estimating is then followed by selecting the specific statistical method that shall be applied to compute the HTs or CIs (e.g., compute HTs and CIs with the *marginaleffects* R package using the delta method).
 
@@
 We could use this method to compute point estimates for all four contrasts that are part of our estimand.
-However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute HTs or CI.
+However, depending on whether we are interested in hypothesis testing or parameter estimation, we also need a method to compute HTs or CIs.
 The *marginaleffects* package [@R-marginaleffects] is a very flexible, increasingly popular package to compute HTs and CIs for contrasts with a variety of statistical models, including GLMMs estimated with *lme4*.
@@
 
-<!-- print(probs, style = "tinytable") -->
-
 The point estimates for all experimental conditions are reported in the `Estimate` column.
@@
 Next, we use the estimated probabilities to compute the four specific contrasts that are part of our estimand.
-For this, we must specify which rows in `probs` have to be subtracted from each other.
+For this, we must specify which rows in `preds` have to be subtracted from each other.
 We will use the `hypotheses` function to compute our four contrasts of interest together with HTs and CIs.
@@
 
-The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2 in the `probs`-output.
+The expression `"b8 = b2"` is special syntax to subtract the estimate in row number 8 from the estimate in row number 2 in the `preds`-output.
 The argument `equivalence = c(0, 0)` can be used to compute one-sided p-values, testing whether the contrast in the population is smaller than 0 (`p (NonSub)` column) or greater than 0 (`p (NonInf)` column).

### 2024-06-25 _manuscript/_tex/manuscript.tex
update _freeze and _manuscript

--- 2024-06-12 _tex/manuscript.tex
@@
 \& Technology, University of St.~Gallen},{Department of Business
-Psychology, Technical University of Applied Sciences Augsburg},{Global
-Business School for Health, University College London}}
+Psychology, Technical University of Applied Sciences Augsburg},{UCL
+Global Business School for Health, University College London}}
 
@@
 GLMMs are often limited to specific designs. Tailored data simulation
-approaches are a more flexible alternative. Based on a practical case
+approaches offer a more flexible alternative. Based on a practical case
 study, the current tutorial equips researchers with a step-by-step guide
@@
 Pargent}{0000-0002-2388-553X}}\par{\addORCIDlink{Timo K.
-Koch}{0000-0001-6728-2063}}
+Koch}{0000-0001-6728-2063}}\par{\addORCIDlink{Anne-Kathrin
+Kleine}{0000-0003-1919-2834}}\par{\addORCIDlink{Eva
+Lermer}{0000-0002-6600-9580}}\par{\addORCIDlink{Susanne
+Gaube}{0000-0002-1633-4772}}
 
@@
 \url{https://Timo-Ko.github.io/glmm_simulation_tutorial/}.  The authors
-declare that there were no conflicts of interest with respect to the
+declare that there were no conflicts of interest concerning the
 authorship or the publication of this article. This research was funded
@@
 analysis, editing; Eva Lermer:   conceptualization, funding
-aquisition, editing; Susanne
+acquisition, editing; Susanne
 Gaube:   conceptualization, supervision, editing}
@@
 When planning experimental research, it is essential to determine an
-appropriate sample size and to use appropriate statistical models to
-analyze the data to ensure that the results obtained are both robust and
-informative (\citeproc{ref-lakensSampleSizeJustification2022}{Lakens,
-2022a}). The recent replication crisis has illustrated many challenges
-surrounding the reproducibility and reliability of study findings
+appropriate sample size and use appropriate statistical models to
+analyze the data to ensure that the results are robust and informative
+(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}). The
+recent replication crisis in Psychology and other disciplines has
+illustrated many challenges surrounding the reproducibility and
+reliability of study findings
 (\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}). As a
@@
 Multiple easy-to-use software solutions exist for simple statistical
-models and experimental designs. However, many researchers lack the
-skills and tools to conduct ``a priori'' (i.e., before data collection)
-sample size planning for more complex research designs such as flexible
-generalized linear mixed models (GLMM) framework. In the present work,
-we provide a tutorial on how to determine adequate sample sizes by
+models and experimental designs (\citeproc{ref-R-pwr}{Champely, 2020};
+\citeproc{ref-lakensSimulationBasedPowerAnalysis2021}{Lakens \&
+Caldwell, 2021}). A priori sample size planning for more complex
+research designs such as flexible generalized linear mixed models (GLMM)
+is not covered by standard software solutions. Researchers willing to
+use this framework will need to use data simulation. In the present
+work, we provide a tutorial on how to determine adequate sample sizes by
 performing tailored simulation-based sample size planning for GLMMs.
@@
 we review existing software solutions in R and discuss under which
-circumstances tailored data simulations are necessary. Then we describe
-general steps and decisions involved in tailored data simulation. To
-illustrate the details of these steps, we finish with a hypothetical
-case study from the field of human-AI (artificial intelligence)
-interaction research.
+circumstances tailored data simulations are necessary. We proceed by
+describing the relevant steps and decisions involved in tailored data
+simulation, illustrated in a case study.
 
 To benefit most of this tutorial paper, we recommend readers to
-familiarise themselves with basic statistical concepts like hypothesis
+familiarize themselves with basic statistical concepts like hypothesis
 tests (HTs) and their statistical power as well as confidence intervals
@@
 \citeproc{ref-riesthuisSimulationBasedPowerAnalyses2024}{Riesthuis,
-2024}). Some knowledge of causal inference is useful but not necessary
+2024}). Some knowledge of causal inference is beneficial but not
+necessary
 (\citeproc{ref-deffnerCausalFrameworkCrossCultural2022}{Deffner et al.,
 2022}; \citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al.,
-2021}). In addition, readers should have a rough understanding of R
+2021}). In addition, readers should have an understanding of how to
+conduct statistical analyses with R
 (\citeproc{ref-wickhamDataScienceImport2023}{Wickham et al., 2023}) and
-how to simulate data. For data simulation, we use functions from the
-tidyverse (\citeproc{ref-wickhamWelcomeTidyverse2019}{Wickham et al.,
-2019}) and the faux package (\citeproc{ref-R-faux}{DeBruine, 2023}).
-Finally, readers should be familiar with regression modeling in general
-and GLMMs in particular. In this tutorial, we simulate data by manually
-specifying the model equation of a GLMM that represents our assumed data
-generating process
+how to simulate data
+(\citeproc{ref-debruineUnderstandingMixedEffectsModels2021}{DeBruine \&
+Barr, 2021};
+\citeproc{ref-hallgrenConductingSimulationStudies2013}{Hallgren, 2013};
+\citeproc{ref-leeUsingTidyversePackage2020}{Lee et al., 2020}). For data
+simulation, we use functions from the \emph{tidyverse}
+(\citeproc{ref-wickhamWelcomeTidyverse2019}{Wickham et al., 2019}) and
+the \emph{faux} package (\citeproc{ref-R-faux}{DeBruine, 2023}).
+Finally, readers should have a basic understanding of regression
+modeling and GLMMs
+(\citeproc{ref-brownIntroductionLinearMixedEffects2021}{Brown, 2021}).
+In this tutorial, we simulate data by manually specifying the model
+equation of a GLMM that represents our assumed data-generating process
 (\citeproc{ref-debruineUnderstandingMixedEffectsModels2021}{DeBruine \&
 Barr, 2021}). It is not necessary to understand the technical details of
-how GLMMs are estimated. However it is crucial to understand the
+how GLMMs are estimated. However, it is crucial to understand the
 structure of a basic GLMM (e.g., logistic regression with random
-intercepts) and how the model assumes that the values in the dependent
-variable are determined by the predictor variables and the random
-effects.
+intercepts) and how the model assumes that the dependent variable's
+values are determined by the predictor variables and the random effects.
 
@@
 Conducting research with insufficiently large sample sizes can have many
-negative consequences. First, experiments may yield inconclusive or
-misleading results, hindering the accumulation of knowledge. Second,
-studies that are doomed to never finding a postulated effect waste
-resources by consuming time, effort, and funding without delivering
-meaningful results. For these reasons, many journals and funding bodies
-now require that a justification of sample size is included in study
-protocols and grant proposals, recognizing its significance in ensuring
-robust and meaningful findings. Although scientists often do not justify
-sample size or use general heuristics from the literature, resource
-constraints often play an important role
-(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}). But
-ideally, a suitable sample size should be determined a priori (i.e.,
-before the study is conducted) based on some meaningful computation to
-ensure that the study will be able to fulfill its purpose.
-
-The majority of empirical studies in psychology and other social
-sciences apply hypothesis testing. As a consequence, the dominant
-approach for determining an adequate sample size is based on power
-analysis (i.e., planning for power)
-(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
+negative consequences (\citeproc{ref-buttonPowerFailureWhy2013}{Button
+et al., 2013}). First, experiments may yield inconclusive or misleading
+results, hindering the accumulation of knowledge. Second, studies that
+are doomed never to find a postulated effect waste resources by
+consuming time, effort, and funding. For these reasons, many journals
+and funding bodies now require a sample size justification in study
+protocols and grant proposals, recognizing its relevance in ensuring
+robust and meaningful findings. While sample sizes can be justified with
+resource constraints or general heuristics, statistical arguments based
+on power or precision are the gold standard
+(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).
+
+Most empirical studies in psychology and other social sciences apply
+hypothesis testing. Consequently, the dominant approach for determining
+an adequate sample size is based on power analysis (i.e., planning for
+power) (\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
 \citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}).
@@
 hypothesis when the alternative hypothesis is true. If the sample size
-(i.e., the number of participants and/or stimuli) used for data
-collection is insufficient to detect the effects or relationships being
-investigated with high probability, the study is considered
-``underpowered''. When planning for power, a target is set for the
-statistical power of a HT of interest. Assuming some effect size of
-interest and a desired significance level, a minimum sample size can be
-determined that, on average, would guarantee reaching this target.
-Although it is recommended to justify the desired significance level and
-power (\citeproc{ref-lakensJustifyYourAlpha2018}{Lakens, Adolfi, et al.,
-2018}), most empirical studies adopt the heuristic of \(\alpha = 0.05\)
-and \(1 - \beta = 0.80\).
+(i.e., the number of participants and/or stimuli) is insufficient to
+detect the effects or relationships being investigated with high
+probability, the study is considered ``underpowered''. When planning for
+power, a target is set for the statistical power of a HT of interest.
+Assuming an effect size of interest and a desired significance level, a
+minimum sample size can be determined that, on average, would guarantee
+reaching this target. The most prominent heuristic is to target a power
+of \(1- \beta = 0.8\) in combination with a type I error rate of
+\(\alpha = 0.05\)
+(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).
+However, depending on the research goals or resource constraints, there
+are often good reasons to move away from this standard
+(\citeproc{ref-benjaminRedefineStatisticalSignificance2017}{Benjamin et
+al., 2017}; \citeproc{ref-lakensJustifyYourAlpha2018}{Lakens, Adolfi, et
+al., 2018}). For example, the \emph{Social Sciences Replication Project}
+targeted a power of 0.9 to safeguard against biased effect sizes in the
+original studies
+(\citeproc{ref-camererEvaluatingReplicabilitySocial2018}{Camerer et al.,
+2018}).
 
@@
 \citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}). Not
-all research questions are best answered using hypothesis testing. It
-has been argued that basic research rarely requires making discrete
-decisions on whether some effect has been ``discovered'' and should thus
-shift from a hypothesis testing towards an estimation framework
+all research questions are best answered by hypothesis testing. It has
+been argued that basic research rarely requires discrete decisions on
+whether some effect has been ``discovered'' and should thus shift from
+hypothesis testing towards an estimation framework
 (\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014};
@@
 \citeproc{ref-mcelreathStatisticalRethinkingBayesian2020}{McElreath,
-2020}). Although this view is not without critique
+2020}), although this view is not without critique
 (\citeproc{ref-uyguntuncEpistemicPragmaticFunction2023}{Uygun Tunç et
-al., 2023}), at least for exploratory or pilot studies where little
-previous research has been conducted, more scientists seem to agree that
-simply estimating the effects of interest and making the estimation
-uncertainty transparent by reporting CIs is more useful. Assuming that
-no HTs are conducted for a planned study, power analysis is not relevant
-for sample size planning. Nonetheless, the sample size still has a
-crucial effect on how informative the planned study will be, because an
-effect of interest is estimated more precisely with bigger samples. In
-the precision framework, the target quantity commonly used for sample
-size planning is the expected width of a CI
-(\citeproc{ref-kelleySampleSizePlanning2006}{Kelley \& Rausch, 2006};
-\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
-\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}). The
-values inside a CI are often interpreted as plausible values for the
-quantity of interest it is supposed to estimate. More formally, a CI
-with confidence level 0.95 provides the smallest interval with the
-property that upon repeated sampling, 95\% of individual CIs would
-include the true quantity of interest. Thus, a narrow CI is more
-informative about the size of the true effect than a wide interval.
-Apart from the desired confidence level, the width of a CI depends
-strongly on the sample size. Because bigger samples carry more
-information, they lead to smaller CIs. When planning for precision, a
-target can be set for the expected width of a CI of interest. Assuming
-some effect size of interest and a desired confidence level, a minimum
-sample size can be determined that, on average, would guarantee reaching
-this target. Because planning for precision is still rare, there are no
-common heuristics on how to choose the desired width of the CI
+al., 2023}). When no HTs are conducted, power analysis is not relevant
+for sample size planning. In the precision framework, the target
+quantity commonly used for sample size planning is the expected width of
+a CI (\citeproc{ref-kelleySampleSizePlanning2006}{Kelley \& Rausch,
+2006}; \citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
+\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}). A
+CI with a confidence level of 0.95 provides the smallest interval with
+the property that 95\% of individual CIs would include the true quantity
+of interest upon repeated sampling. Thus, a narrow CI with fewer
+plausible values for the quantity of interest is more informative about
+the size of the true effect than a wide CI. Apart from the confidence
+level, the width of a CI depends on the sample size. Because bigger
+samples carry more information, they lead to smaller CIs. When planning
+for precision, a target can be set for the expected width of a CI of
+interest. Assuming some effect size of interest and a certain confidence
+level, a minimum sample size can be determined that would guarantee
+reaching the targeted expected width. Because planning for precision is
+still rare, there are no common heuristics for choosing the desired
+width of the CI
 (\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).
@@
 As study designs become more complex, psychological researchers require
-more sophisticated statistical models to capture the nuanced
-relationships and grouping structures introduced by them
+more sophisticated statistical models to capture their nuanced
+relationships and grouping structures
 (\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}). GLMMs
-(also called multilevel models) are gaining increasing popularity
-because they offer great flexibility
-(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
-2021}). GLMMs are an extension of LMMs (Linear Mixed Models), which are,
+(also called multilevel models) are gaining popularity because they
+offer great flexibility when applied carefully
+(\citeproc{ref-brownIntroductionLinearMixedEffects2021}{Brown, 2021};
+\citeproc{ref-matuschekBalancingTypeError2017}{Matuschek et al., 2017};
+\citeproc{ref-meteyardBestPracticeGuidance2020a}{Meteyard \& Davies,
+2020}). GLMMs are an extension of LMMs (Linear Mixed Models), which are,
 in turn, extensions of linear regression models that account for
 correlated data, including hierarchical structures
-(\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
+(\citeproc{ref-bolkerLinearGeneralizedLinear2015}{Bolker, 2015};
+\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
 2021}). In this context, correlated data means that the value in the
@@
 another observation in a systematic way that is not already accounted
-for by the usual (fixed) predictor variables (e.g., age of
+for by the usual (fixed) predictor variables (e.g., the age of
 participants). This correlation can arise for various reasons: For
 instance, responses to some stimuli from some participants might be more
-similar because the same person was measured twice (repeated
-measurements), participants come from the same neighborhood
-(clustering), or participants responded to the same stimulus (stimulus
-effects). Thus, modeling such correlations is important whenever the
-data has a clear structure, while the grouping variables can be
-hierarchically nested (e.g., grouping variables students and schools:
-each student belongs to exactly one school) or cross-classified (e.g.,
-grouping variables students and math exercises: each student is
-presented with several math exercises). LMMs are used when the outcome
-variable is continuous and follows a normal distribution. They allow for
-the modeling of fixed effects, which capture the relationships between
-the usual predictors and the outcome, as well as random effects, which
-account for the different types of correlation structure and grouping
-effects. Random effects are typically assumed to follow a normal
-distribution with a mean of zero and a variance that quantifies the
-heterogeneity across groups. GLMMs extend the LMM framework to
-accommodate non-normally distributed continuous and categorical outcome
-variables. GLMMs involve a link function that connects the linear
-combination of predictor variables to the expected value of the outcome
-variable. The link function allows for modeling the relationship between
-predictors and the outcome in a non-linear way that is appropriate for
-the specific distribution family of the outcome variable. For example,
-think of an experiment with different design factors (e.g., picture
-positions, headline aesthetics) impacting the likelihood of users
-clicking on an online advertisement. The participants' behavior is
-measured repeatedly over several sessions. The click patterns of
-participants in one session are likely to be correlated with their
-previous sessions and the outcome variable is binary (click/no click)
-for each session, which follows a binomial distribution.
+similar because the same person was measured multiple times (repeated
+measurements), participants belong to the same group (clustering), or
+participants responded to the same stimulus (stimulus effects). Thus,
+modeling such correlations is important whenever the data has a clear
+structure, while the grouping variables can be hierarchically nested
+(e.g., grouping variables students and schools: each student belongs to
+exactly one school) or cross-classified (e.g., grouping variables
+students and math exercises: each student is presented with several math
+exercises). LMMs are used when the outcome variable is continuous and
+follows a normal distribution (after conditioning on all predictor
+variables). They allow for the modeling of fixed effects, which capture
+the relationships between the usual predictors and the outcome, as well
+as random effects, which account for the different types of correlation
+structure and grouping effects. Random effects are typically assumed to
+follow a normal distribution with a mean of zero and a variance that
+quantifies the heterogeneity across groups. Correlated random effects
+can be assumed in models that contain both random intercepts and random
+slopes. GLMMs extend the LMM framework to accommodate non-normally
+distributed continuous and categorical outcome variables. GLMMs involve
+a link function that connects the linear combination of predictor
+variables to the expected value of the outcome variable. The link
+function allows for modeling the relationship between predictors and the
+outcome in a non-linear way that is appropriate for the specific
+distribution family of the outcome variable.
 
@@
 To our knowledge, existing approaches for sample size planning for GLMMs
-have exclusively focused on planning for power. Power analysis methods
-for multilevel models can be categorized into formula-based methods and
+have exclusively focused on planning for power. In Table 1, we review
+available software packages that can be used to perform power analysis
+for multilevel models in R {[}Version 4.4.0; R Core Team
+(\citeproc{ref-R-base}{2024}){]}. Power analysis methods can be
+categorized into formula-based, summary-statistics-based and
 simulation-based methods
 (\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
-et al., 2022}). Formula-based methods rely on formulas to calculate
-power directly while simulation-based methods rely on repeatedly
-simulating data with a known true effect size and estimating power
-empirically (i.e., what percentage of simulated datasets produces a
-significant p-value). Available formula-based software packages for
-multilevel models often do not include GLMMs or are limited to very
-simple designs
+et al., 2022}). Formula-based methods rely on exact formulas to
+calculate power directly. Summary-statistics-based methods use
+statistical theory to approximate power based on formula-based methods
+developed for simple t-tests. Simulation-based methods rely on
+repeatedly simulating data with a known true effect size and estimating
+power empirically, that is what percentage of simulated datasets
+produces a significant p-value. Available formula-based and
+summary-statistics-based software packages for multilevel models often
+do not include GLMMs or are limited to simple designs
 (\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
@@
 \citeproc{ref-zimmerSampleSizePlanning2022}{Zimmer et al., 2022}).
-However, many of these tutorials focus on linear mixed models (LMMs) and
-the common study designs (but see
-\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021},
-for a tutorial that also covers more advanced settings). This narrow
-focus provides limited guidance for researchers faced with more complex
-study designs, especially when little prior knowledge about plausible
-effect sizes is available (see the discussion in
+However, many of these tutorials focus on LMMs and the most common study
+designs (see \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et
+al., 2021} for a tutorial that also covers more advanced settings). This
+narrow focus provides limited guidance for researchers using more
+complex study designs, especially when little prior knowledge about
+plausible effect sizes is available (see the discussion in
 \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}).
 Simulation-based power analysis with GLMMs requires making a range of
-assumptions: The (conditional) distribution assumption specifies the
-distributional family for the outcome variable. Assumptions about the
-random effects include the assumption of normality (i.e., that the
-random effects follow a normal distribution) and the covariance
-structure among the random effects (i.e., if and how they are
-correlated). Making these decisions requires understanding the
-underlying assumptions of the model and ensuring they align with the
+assumptions about the model structure that should align with the
 characteristics of the data being analyzed. Existing tutorials often
@@
 which makes specifying assumptions a practical challenge (see the
-discussion in (\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et
-al., 2008}) and
-(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al.,
-2021})).
+discussion in Maxwell et al.
+(\citeproc{ref-maxwellSampleSizePlanning2008}{2008}) and Kumle et al.
+(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{2021})). We will
+discuss a number of strategies on how to specify model parameters for
+application-specific, tailored data simulations in a later chapter.
 
@@
 
-In Table 1, we give a short review of existing R packages that can be
-used for power analysis for GLMMs.
-
-\section{When to use tailored data
-simulation?}\label{when-to-use-tailored-data-simulation}
+\section{Reasons to use tailored data
+simulation}\label{reasons-to-use-tailored-data-simulation}
 
@@
 complicated and time-consuming than using the existing software tools
-outlined in TABLE 1. The most important circumstances under which
-tailored simulation-based sample size planning is necessary are 1)
-complex study designs, 2) complex statistical hypotheses, 3) planning
-for precision, 4) no available prior studies or pilot data.
-
-Requirements of real-world studies are often more complex than the
-simplified designs assumed by many user-friendly software packages for
-sample size planning. One frequent issue in applied data analysis is
-missing data, and there can be various reasons for this
-(\citeproc{ref-littleStatisticalAnalysisMissing2014}{Little \& Rubin,
-2014}). For example, data can be missing completely at random (e.g.,
-because an electronic measurement device randomly failed for some
-technical reasons). Alternatively, subjects might systematically drop
-out or produce missing data, but this dropout can be explained by some
-attributes also measured in the dataset (e.g., older subjects have a
-higher probability to refuse answering a question on income). In a more
-complicated scenario, missing data in some variable is caused by the
-measured attribute itself (e.g., wealthy people are more likely to
-refuse reporting their income). Moreover, many experimental designs
-contain conditions in which values of the predictor variables are
-missing by design. This can make data analysis more complicated because
-predictors have to be coded in specific ways that prevent the estimated
-GLMM from becoming unidentified. Whether missing data has an effect on
-the sample size planning depends on our theoretical assumptions on how
-the missingness is caused. However, it is often challenging to decide
-whether missing data can be safely ignored in the data analysis and
-sample size planning process based on a merely theoretical approach
+outlined in TABLE 1. Circumstances under which tailored simulation-based
+sample size planning is necessary include 1) complex study designs, 2)
+complex statistical hypotheses, 3) planning for precision, and 4) lack
+of prior studies or pilot data.
+
+First, requirements of real-world studies are often more complex than
+the simplified designs assumed by many user-friendly software packages
+for sample size planning. One frequent issue in applied data analysis is
+missing data (\citeproc{ref-littleStatisticalAnalysisMissing2014}{Little
+\& Rubin, 2014}). Missings can be completely random (e.g., an electronic
+measurement device randomly failed for technical reasons), depend on
+some attributes also measured in the dataset (e.g., older subjects
+refuse answering a question on income), or be caused by the measured
+attribute itself (e.g., wealthy people refuse to report their income).
+Moreover, many experimental designs contain conditions in which values
+of the predictor variables are missing by design. This can make data
+analysis more complicated because predictors have to be coded in
+specific ways that prevent the estimated GLMM from becoming
+unidentified. Whether missing data has an effect on the sample size
+planning depends on our theoretical assumptions on how the missingness
+is caused. However, it is often challenging to decide whether missing
+data can be safely ignored in the data analysis and sample size planning
+process based on a merely theoretical approach
 (\citeproc{ref-gomilaMissingDataExperiments2022}{Gomila \& Clark,
@@
 include the assumed process of how data become missing in the data
-simulation, thereby determining the required sample based on simulated
-datasets that contain missing values (for an example, see
+simulation, thereby determining the required sample size based on
+simulated datasets that contain missing values (for example, see
 \citeproc{ref-lanePowerStrugglesEstimating2018}{Lane \& Hennes, 2018}).
@@
 examples are zero-inflated outcomes, censoring, and nonlinear predictor
-effects that can be modeled with the R packages glmmTMB
+effects that can be modeled with the R packages \emph{glmmTMB}
 (\citeproc{ref-brooksGlmmTMBBalancesSpeed2017}{Brooks et al., 2017}) or
-brms (\citeproc{ref-burknerAdvancedBayesianMultilevel2018}{Bürkner,
-2018}). Tailored simulation-based approaches do not share the same
-limitation than the existing software solutions for power analysis that
-focus exclusively on GLMMs. As long as there is a software package
-available to estimate the model of interest, it is always possible to
-perform tailored simulation-based sample size planning.
-
-The most common hypotheses tested in psychological research are of the
-type \(H_0: \beta = 0\), where \(\beta\) is a slope or intercept of a
-regression model. However, many research questions in psychology
+\emph{brms}
+(\citeproc{ref-burknerAdvancedBayesianMultilevel2018}{Bürkner, 2018}).
+Tailored simulation-based approaches do not share the same limitations
+as the existing software solutions for power analysis that focus
+exclusively on GLMMs. As long as there is a software package available
+to estimate the model of interest, it is always possible to perform
+tailored simulation-based sample size planning.
+
+Second, the most common hypotheses tested in psychological research are
+of the type \(H_0: \beta = 0\), where \(\beta\) is a slope or intercept
+of a regression model. However, many research questions in psychology
 actually require testing more complex statistical hypotheses. In the new
@@
 (\citeproc{ref-lakensEquivalenceTestingPsychological2018}{Lakens,
-Scheel, et al., 2018}). In combination, this might require a test such
-as \(H_0: \beta \leq 0.1\). More elaborate research questions often
-require testing hypotheses that consist of a combination of model
-parameters, for example testing simple slopes
+Scheel, et al., 2018}). This might require a test such as
+\(H_0: \beta \leq 0.1\). More elaborate research questions often require
+testing hypotheses that consist of a combination of model parameters,
+for example, testing simple slopes
 (\citeproc{ref-preacherComputationalToolsProbing2006}{Preacher et al.,
@@
 coefficient by clever coding and/or centering of predictor variables.
-However, interesting research questions often consist of combined
-hypotheses that consist of more than one separate statistical hypothesis
-(for a tutorial on contrast analysis in GLMMs, see
+However, research questions often consist of combined hypotheses that
+consist of more than one separate statistical hypothesis (for a tutorial
+on contrast analysis in GLMMs, see
 \citeproc{ref-schadHowCapitalizePriori2020}{Schad et al., 2020}). For
@@
 the global HT. None of the software packages for sample size planning in
-table X can handle combined hypotheses as discussed here and only some
-can handle directed hypotheses. In contrast, our case study will
-demonstrate how we can test directed combined hypotheses with tailored
-simulation-based sample size planning.
-
-Although planning for precision
-(\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014};
-\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
-\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}) has
-been increasingly discussed as a useful strategy for empirical research,
-all available software packages for sample size planning with GLMMs are
-based on power analysis. Therefore, researchers that want to want to
-apply an estimation strategy in their studies, without testing any
-statistical hypotheses, currently cannot use the software packages
-outlined in TABLE 1. However, tailored simulation-based sample size
-planning can easily handle the planning for precision approach
-(\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}).
-The only change in procedure is that instead of computing HTs for each
-simulated dataset and estimating statistical power across repetitions,
-CIs are computed for each simulated dataset, and the expected width is
-estimated.
-
-All frameworks for sample size planning require the user to make
+Table 1 can handle combined hypotheses and only some can handle directed
+hypotheses. In contrast, our case study will demonstrate how we can test
+directed combined hypotheses with tailored simulation-based sample size
+planning.
+
+Third, all available software packages for sample size planning with
+GLMMs are based on power analysis and do not support planning for
+precision. Thus, researchers that want to apply an estimation strategy
+instead of testing statistical hypotheses
+(\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014}), currently
+cannot use the software packages outlined in TABLE 1. However, tailored
+simulation-based sample size planning can easily handle the planning for
+precision approach (\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell
+et al., 2008}). The only change in procedure is that instead of
+computing HTs for each simulated dataset and estimating statistical
+power across repetitions, CIs are computed for each simulated dataset,
+and their expected width is estimated.
+
+Fourth, all frameworks for sample size planning require the user to make
 assumptions about the expected effect size. Assuming the true effect is
@@
 researcher has access to similar studies or pilot data, providing such
-standardized effect sizes is feasible. However, note that choosing
-effect sizes based on small pilot studies is generally not recommended,
-as those estimates can be heavily biased
+standardized effect sizes is feasible. However, choosing effect sizes
+based on small pilot studies is generally not recommended, as those
+estimates can be heavily biased
 (\citeproc{ref-albersWhenPowerAnalyses2018a}{Albers \& Lakens, 2018};
@@
 flexible that general heuristics of what should be considered a small
-effect do not exist or are difficult to defend. In our experience, using
-domain knowledge to construct a tailored data simulation is the only
-solution to determine plausible effect sizes in the absence of prior
-evidence. It would be possible to use these tailored simulations to
-extract plausible values for standardized effect sizes that could then
-be inserted in existing software packages for sample size planning.
-However, we would argue that when tailored data simulations necessary to
-determine effect sizes anyway, performing the whole sample size planning
-in a customized way is preferred over using the existing software
-packages.
+effect do not exist or are difficult to defend. In the absence of prior
+evidence, using domain knowledge to construct a tailored data simulation
+is considered the best solution to determine plausible effect sizes (see
+scenario 3 in \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle
+et al., 2021}). It would be possible to use these tailored simulations
+to compute standardized effect sizes that could then be inserted in
+existing software packages for sample size planning. However, we would
+argue that when tailored data simulations are necessary to determine
+effect sizes anyway, performing the whole sample size planning in a
+customized way is preferred over using the existing software packages.
 
@@
 
-Although the details differ depending on the specific study
-characteristics, each tailored simulation-based sample size planning
-requires a series of steps and decisions. We will introduce each step in
-a theoretical section, followed by the practical application based on a
-hypothetical case study. All code in this manuscript and simulation
-results are available in the project's repository on the Open Science
-Framework (\url{https://osf.io/dhwf4/}).
+Although the details differ for each study, every tailored
+simulation-based sample size planning requires a series of steps and
+decisions. We will introduce each step in a theoretical section,
+followed by the practical application based on a case study. All code in
+this manuscript and simulation results are available on the Open Science
+Framework (\url{https://osf.io/dhwf4/}) and on our tutorial website
+(\url{https://timo-ko.github.io/glmm_simulation_tutorial/}).
 
@@
 
-The first step in every research process should be a clear definition of
-the theoretical \emph{estimand}
+The first step in every research process is a clear definition of the
+theoretical \emph{estimand}
 (\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}),
 i.e.~the theoretical quantity which is necessary to answer a specific
-research question. The estimand consists of a quantity, that can be
+research question. The estimand consists of a quantity that can be
 described for each unit under investigation and a clear definition of
@@
 example, an estimand might be the probability that a clinical
-psychologists makes the correct diagnosis for a psychiatric patient with
+psychologist makes the correct diagnosis for a psychiatric patient with
 major depression, averaged across all clinical psychologists and
-depressed patients in psychiatric institutions in Germany.
+depressed patients in psychiatric institutions in a given country.
 
@@
 a randomized experiment) that will produce the observed data in the
-planned study. For many common research questions in psychology, the
-estimand can be expressed as a statistical quantity that can be
+planned study. For many common research questions in psychology, it is
+possible to express the estimand as a statistical quantity that can be
 estimated with a regression model, for example a single \(\beta\)
@@
 
-In the present hypothetical case study, we consider the effectiveness of
-feedback provided by an artificial intelligence (AI) embedded in a
-diagnostic decision support system. The context is a clinical setting,
-where expert radiologists and students under training must detect
-bleeding based on head scans from computer tomography (CT). In the
-investigated AI-enabled diagnostic decision support system, an AI model
-can provide initial diagnostic advice, which can be used as guidance by
-the humans who are required to make the final diagnostic decision. The
-research goal is to validate the effectiveness of the AI-enabled advice.
-We consider the AI-enabled advice as effective, if the following pattern
-holds, which we will first describe verbally:
-
-\emph{We expect that for BOTH expert radiologists and medical students,
-correct AI-advice leads to a higher probability of accurately diagnosing
-a CT scan compared to no AI-advice presented, AND, we expect that for
-BOTH experts and non-experts, incorrect advice leads to a lower
+In the present case study, we consider the effectiveness of feedback
+provided by an artificial intelligence (AI) embedded in a clinical
+decision support system. The context is a clinical setting, where task
+experts (i.e., radiologists), and non-expert (i.e., medical students)
+must detect bleedings based on head scans from computer tomography (CT).
+The AI model can provide initial diagnostic advice, which can be used as
+guidance by the humans who are required to make the final diagnostic
+decision. The research goal is to validate the effectiveness of the
+AI-enabled advice. We consider the AI-enabled advice effective, if the
+following pattern holds:
+
+\emph{We expect that for BOTH radiologists and medical students, correct
+AI advice leads to a higher probability of accurately diagnosing a CT
+scan compared to no AI advice presented, AND, we expect that for BOTH
+task experts and non-experts, incorrect advice leads to a lower
 probability of accurately diagnosing a CT scan compared to no advice
@@
 & P(\text{correct diagnosis} | \text{no advice, average student, average scan}) \\
-& \quad - P(\text{correct diagnosis} | \text{wrong advice, average student, average scan})
+& \quad - P(\text{correct diagnosis} | \text{incorrect advice, average student, average scan})
 \end{aligned}
 \] For example, the first expression is concerned with the difference
-between the probability that a correct diagnosis is made if correct
-AI-advice is presented and the probability that a correct diagnosis is
-made if no AI-advice is presented. This contrast is quantified for a
+between the probability that a correct diagnosis is made if correct AI
+advice is presented and the probability that a correct diagnosis is made
+if no AI advice is presented. This contrast is quantified for a
 hypothetical \emph{typical} expert and a \emph{typical} scan, where
@@
 To complete our definition of the estimand, we have to define our target
-population that consists of persons, stimuli, and AI-advice: With
-respect to persons, we are only interested in expert radiologists and
-medical students at German universities. With respect to stimuli, we are
-only interested in the head CT scans made from subjects that do or do
-not suffer from intracerebral hemorrhage. Lastly, we are only interested
-in AI-advice given by a specific AI-enabled diagnostic decision support
-system.
-
-Although the estimand is initially defined outside of any statistical
+population that consists of persons, stimuli, and AI advice: With
+respect to persons, we are only interested in experts (i.e.,
+radiologists) and non-experts (i.e., medical students). With respect to
+stimuli, we are only interested in the head CT scans made from subjects
+that do or do not suffer from intracerebral hemorrhage. Lastly, we are
+only interested in AI advice given by a specific AI-enabled system.
+
+Although the estimand is initially defined outside of a statistical
 model, it is only useful if we find a way to estimate it based on
 observed data. For our exemplary research question, it is possible to
-construct an experimental study, where all participants are confronted
-with the same set of head CT scans, but the kind of AI-advice given for
+construct an experimental study where all participants are confronted
+with the same set of head CT scans, but the kind of AI advice given for
 each scan is randomly assigned within participants. This random
 intervention allows us to produce an empirical estimate of our estimand,
-although, in reality, each person receives only one kind of AI-advice
-(correct advice, wrong advice, no advice) for each scan. We will see
-later how each of the probability expressions in our estimand can be
-modeled with the same GLMM. Estimating this GLMM based on the data
-observed in our planned study will produce an estimate for each
+although, in reality, each person receives only one kind of AI advice
+(correct advice, incorrect advice, no advice at all) for each scan. We
+will see later how each of the probability expressions in our estimand
+can be modeled with the same GLMM. Estimating this GLMM based on the
+data observed in our planned study will produce an estimate for each
 probability, and these estimates can then be combined to compute an
@@
 
-\subsection{Step 2: Simulate the data generating
+\subsection{Step 2: Simulate the data-generating
 process}\label{step-2-simulate-the-data-generating-process}
@@
 When the estimand has been defined, the next step in the research
-process is to write code that simulates the (assumed) data generating
-process of the planned study. This requires specifying a generative
-process for all predictor variables used in the final data analysis.
-While such assumptions can be quite challenging for observational
-studies or continuous predictor variables, this is less of a problem for
-experimental studies with only categorical predictor variables. When all
-predictor variables have been simulated, one can use the structure of a
-suitable GLMM to simulate the dependent variable. To simulate the GLMM,
-one requires plausible values for all model parameters. We will discuss
-strategies on how these values can be obtained later. Because we have
-full control over the data generating process in a tailored simulation,
-it is possible to model specific aspects of the planned study, like data
+process is to write code that simulates the data-generating process of
+the planned study. This requires specifying a generative process for all
+predictor variables used in the final data analysis. Realistic
+assumptions can be quite challenging for observational studies or
+continuous predictor variables, which is beyond the scope of this
+tutorial. However, this step is much easier for experimental studies
+with only categorical predictor variables because the distribution of
+predictors is fixed by the study design. When all predictor variables
+have been simulated, one can use the structure of a suitable GLMM to
+simulate the dependent variable. To simulate the GLMM, one requires
+plausible values for all model parameters. We will discuss strategies on
+how these values can be obtained later. Because we have full control
+over the data-generating process in a tailored simulation, it is
+possible to model specific aspects of the planned study, like data
 missing by design or assuming that some subjects might drop out. The
 quality of the results of the sample size planning crucially depends on
-the plausibility of the simulated data generating process. However, we
-would argue that even a strongly simplified data generating process
-(e.g.~only a small number of interaction effects; only random intercepts
-and no random slopes; assuming that data is missing completely at
-random) can yield informative results.
+the plausibility of the simulated data-generating process. However, we
+argue that even a strongly simplified data-generating process (e.g.,
+only a small number of interaction effects; only random intercepts, and
+no random slopes; assuming that data is missing completely at random)
+can yield informative results.
 
@@
 
-In our case study, we simulate data for an experiment where the
-diagnostic performance of users of an AI-enabled diagnostic decision
-support system will be evaluated. Radiologists (task experts) and
-students/interns (non experts) review a series of head computer
-tomography (CT) scans to assess the presence of a bleeding. An AI model
-provides initial diagnostic advice to assist their decision-making. In
-the control condition, no AI advice is presented. When AI advice is
-given, this advice can be either correct or incorrect. The type of
-advice (no advice, wrong advice, correct advice) is randomized within
-subjects across brain scans. After reviewing a CT scan, participants
-deliver a medical diagnosis (bleeding or no bleeding), which may be
-either accurate or inaccurate. This experimental design introduces some
-missing values by design since the advice is neither correct nor
-incorrect when no advice is present, which must be taken into account
-when simulating and analyzing the data. In this example, recruiting task
-experts (i.e., radiologists) is more challenging due to their limited
-availability, while non-experts (i.e., students/interns) are more
-readily accessible. The goal of simulation-based sample size planning is
-to determine how many task experts and non-experts must be recruited to
-achieve sufficient statistical power or precision in the planned
-experiment.
+In our case study, radiologists and students review a series of head CT
+scans to assess the presence of a bleeding. An AI model provides initial
+diagnostic advice to assist their decision-making. In the control
+condition, no AI advice is presented. When AI advice is given, this
+advice can be either correct or incorrect. The type of advice (no
+advice, incorrect advice, correct advice) is randomized within subjects
+across CT scans. After reviewing a CT scan, participants deliver a
+diagnosis (bleeding or no bleeding), which may be either accurate or
+inaccurate. This experimental design introduces some missing values by
+design since the advice is neither correct nor incorrect when no advice
+is present, which must be taken into account when simulating and
+analyzing the data. In this example, recruiting task experts (i.e.,
+radiologists) is more challenging due to their limited availability,
+while non-experts (i.e., students) are more readily accessible. The goal
+of simulation-based sample size planning is to determine how many task
+experts and non-experts must be recruited to achieve sufficient
+statistical power or precision in the planned experiment.
 
@@
 modeled as a linear combination \(\eta\) of the predictor variables
-\(\mathbf{X}\), the random effects \(\mathbf{U}\) and the model
+\(\mathbf{X}\), the random effects \(\mathbf{U}\), and the model
 parameters \(\mathbf{\beta}\)
@@
 limit complexity, we do not consider random slopes, additional
-predictors or higher-level interactions.
+predictors, or higher-level interactions.
 
@@
 parameter values (which we will explain in the next section) and load
-the packages we use to manipulate and simulate data.\footnote{The faux
-  package (\citeproc{ref-R-faux}{DeBruine, 2023}) contains useful
-  functions when simulating factorial designs, including random effects.}
-In our case study, each subject (\texttt{n\_subjects} in total) is
-assumed to respond to each item (i.e., CT scan; \texttt{n\_items} in
-total). Thus, the \texttt{add\_random} command creates a fully-crossed
-\texttt{data.frame} with \texttt{n\_subjects} \(\times\)
-\texttt{n\_items} rows. We add a between-subject effect with the
-\texttt{add\_between} command, simulating that about \(25\%\) of
+the packages we use to manipulate and simulate data.\footnote{The
+  \emph{faux} package (\citeproc{ref-R-faux}{DeBruine, 2023}) contains
+  useful functions when simulating factorial designs, including random
+  effects.} In our case study, each subject (\texttt{n\_subjects} in
+total) is assumed to respond to each item (i.e., CT scan;
+\texttt{n\_items} in total). Thus, the \texttt{add\_random} command
+creates a fully-crossed \texttt{data.frame} with \texttt{n\_subjects}
+\(\times\) \texttt{n\_items} rows. We add a between-subject effect with
+the \texttt{add\_between} command, simulating that about \(25\%\) of
 subjects are experts. The next two lines simulate that in
@@
 
-In the absence of previous studies with the same design or pilot data,
-researchers require strategies on how to specify the population
-parameters used in their data simulation. Population parameters are all
-model parameters estimated in a GLMM, in particular the regression
-coefficients of the fixed effects and the standard deviation of the
-random effects (and the correlation between random effects in more
-complicated models). In contrast to non-hierarchical linear regression,
-common heuristics based on standardized effect sizes are less useful or
-not even available for GLMMs. Our strategies to specify population
-parameters will require access to domain knowledge from domain experts.
-Because most domain knowledge can only be expressed in unstandardized
-measurement units of a specific application, we argue that
-unstandardized effect sizes are usually preferable over standardized
-effect sizes for tailored simulation-based sample size planning. The
-basic idea of all strategies is how the data generating process implied
-by certain values of population parameters can be quantified or
-visualized in an intuitive way that enables a calibration of population
-parameters based on the available knowledge of domain experts.
-
-Although we use frequentist model estimation in our tutorial, many
-strategies described in this chapter are inspired by research on how to
-monitor the plausibility of model assumptions in applied Bayesian
-statistics (\citeproc{ref-gelmanBayesianWorkflow2020}{Gelman et al.,
-2020}).
+Population parameters are all model parameters estimated in a GLMM, in
+particular the regression coefficients of the fixed effects and the
+standard deviation of the random effects (and the correlation between
+random effects in more complicated models). In the absence of previous
+studies with the same design or pilot data, strategies to specify
+population parameters require access to domain knowledge from domain
+experts. Because most domain knowledge can only be expressed in
+unstandardized measurement units of a specific application, we argue
+that unstandardized effect sizes are usually preferable over
+standardized effect sizes for tailored simulation-based sample size
+planning. The basic idea of all strategies is to quantify or visualize
+the data-generating process implied by certain values of population
+parameters in an intuitive way that enables calibration of population
+parameters based on the available knowledge of domain experts. Although
+we use frequentist model estimation in our tutorial, many strategies
+demonstrated in this chapter are inspired by research on monitoring the
+plausibility of model assumptions in applied Bayesian statistics
+(\citeproc{ref-gelmanBayesianWorkflow2020}{Gelman et al., 2020}).
 
@@
 (\(\beta_0\), \(\beta_e\), \(\beta_a\), \(\beta_c\), \(\beta_{ea}\),
-\(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\)), but have not talked
-about where these numbers came from.
-
-The starting point for all parameter values in our present case study
-were based on results from distantly related study designs in the
-literature. Additionally, we had repeated discussions with our
-affiliated domain experts in radiology to check whether our assumptions
-seem plausible.
+\(\beta_{ec}\), \(\sigma_S\), and \(\sigma_I\)) but have not talked
+about where these numbers came from. The starting point for all
+parameter values was based on results from distantly related study
+designs in the literature. Additionally, we had repeated discussions
+with our affiliated domain experts in radiology to check whether our
+assumptions regarding participants' diagnostic performance seemed
+plausible.
 
@@
 parameters in a binomial GLMM are hard to interpret in isolation because
-1) the \(\beta\) parameters are connected to the modeled probability via
-the non-linear inverse-logit link, and 2) we also have to consider the
-random effects. The most simple interpretation, that allows us to ignore
-the random effects for now, works by imagining a subject with average
-ability (\(u_{0s} = 0\)) responding to an item (i.e., CT scan) with
-average difficulty (\(u_{0i} = 0\)). Then the model implied probability
-that such a person solves such an item accurately is given by:
+first, the parameters are connected to the modeled probability via the
+non-linear inverse-logit link, and second, we also have to consider the
+random effects. The most simple interpretation, which allows us to
+ignore the random effects for now, works by imagining a subject with
+average ability (\(u_{0s} = 0\)) responding to an item (i.e., CT scan)
+with average difficulty (\(u_{0i} = 0\)). Then the model-implied
+probability that such a person solves such an item accurately is given
+by:
 
@@
 conditions, some terms drop from the equation because they are
-multiplied by \(0\). For example, the probability that a non-expert with
+multiplied by \(0\). For example, the probability that a student with
 average ability solves an item with average difficulty when no advice is
-presented, only requires the intercept: \[
+presented only requires the intercept: \[
 \begin{aligned}
@@
 \endlastfoot
-no advice, no expert & 0.70 & \(logit(0.70) = \beta_0\) \\
+no advice, student & 0.70 & \(logit(0.70) = \beta_0\) \\
 no advice, expert & 0.90 & \(logit(0.90) = \beta_0 + \beta_e\) \\
-false advice, no expert & 0.40 & \(logit(0.40) = \beta_0 + \beta_a\) \\
-false advice, expert & 0.85 &
+incorrect advice, student & 0.40 &
+\(logit(0.40) = \beta_0 + \beta_a\) \\
+incorrect advice, expert & 0.85 &
 \(logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \\
-correct advice, no expert & 0.90 &
+correct advice, student & 0.90 &
 \(logit(0.90) = \beta_0 + \beta_a + \beta_c\) \\
@@
 that are the standard deviations for the random intercepts. For this, we
-introduce two more strategies.
-
-\subsubsection{Insightful descriptive
-statistics}\label{insightful-descriptive-statistics}
+introduce two more strategies in the following sections.
+
+\subsubsection{Examine insightful descriptive
+statistics}\label{examine-insightful-descriptive-statistics}
 
@@
 would be produced by the model, if a specific set of values for the
-population parameters would be specified. The knowledge of how to
-simulate from a GLMM enables us to compute insightful descriptive
-statistics that can be compared to available domain knowledge much more
+population parameters is specified. The knowledge of how to simulate
+from a GLMM enables us to compute \emph{insightful descriptive
+statistics} that can be compared to available domain knowledge much more
 easily than the opaque values of model parameters. For example, domain
@@
 expected ratio of the binary dependent variable in different
-experimental conditions, i.e.~which relative frequency they expect to
-observe (on average). The job of the analyst who is familiar with the
-mathematical structure of the GLMM is to produce the model implied value
-of the insightful descriptive statistic expected by the domain expert.
+experimental conditions, i.e., which relative frequency they expect to
+observe. The job of the analyst who is familiar with the mathematical
+structure of the GLMM is to produce the model-implied value of the
+insightful descriptive statistic that is expected by the domain expert.
 Although insightful descriptive statistics usually depend on the model
@@
 relationship mathematically. Instead, one can simply adjust the
-population parameters by trial and error until the model implied
+population parameters by trial and error until the model-implied
 quantities produce the desired result.
@@
 
-In the last section, we showed how we can derive the model implied
+In the last section, we showed how we can derive the model-implied
 probability that a subject with average ability solves an item with
@@
 (\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
-2021}). The first perspective implies a so-called marginal
-interpretation, while the second one implies a conditional
-interpretation.
+2021}). The first perspective implies a so-called \emph{marginal
+interpretation}, while the second one implies a \emph{conditional
+interpretation}.
 
@@
 presented with correct advice (the conditional perspective). However,
-even if the model were true, we would not observe that 95\% of experts
-responding to items presented with correct advice from a big sample of
-subjects drawn from their natural distribution of ability, and items
-drawn from their natural distribution of difficulty (the marginal
+even if the model assumptions were true, we would not observe that 95\%
+of experts responding to items presented with correct advice from a big
+sample of subjects drawn from their natural distribution of ability and
+items drawn from their natural distribution of difficulty (the marginal
 perspective). How much the two probabilities differ depends on the
 standard deviations of the random intercepts (the two probabilities are
-only equal if both standard deviations would be zero). We want to use
-the model implied observed proportion of correct diagnoses in each
-experimental condition as an insightful descriptive statistics to
-determine plausible values for the random effect standard deviations.
-
-We will simulate a large dataset (for which the observed values of the
-descriptive statistic will be close to their model implied true values)
+only equal if both standard deviations are be zero). We want to use the
+model-implied observed proportion of correct diagnoses in each
+experimental condition as an insightful descriptive statistic to
+determine plausible values for the random effect standard deviations. We
+will simulate a large dataset (for which the observed values of the
+descriptive statistic will be close to their model-implied true values)
 and simply compute the relative frequency of correct diagnoses for each
@@
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
- \StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
- \StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
- \StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
+ \StringTok{"student, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
+ \StringTok{"student, incorrect advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, incorrect advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
+ \StringTok{"student, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{group\_by}\NormalTok{(condition) }\SpecialCharTok{|\textgreater{}}
@@
   <fct>                                  <dbl>
-1 no expert, no advice                   0.683
+1 student, no advice                     0.683
 2 expert, no advice                      0.881
-3 no expert, wrong advice                0.409
-4 expert, wrong advice                   0.828
-5 no expert, correct advice              0.883
+3 student, incorrect advice              0.409
+4 expert, incorrect advice               0.828
+5 student, correct advice                0.883
 6 expert, correct advice                 0.938
@@
 domain experts whether our chosen values for the random effect standard
-deviations would produce data that align with out domain expertise.
-However, although the result was deemed plausible, these statistics were
-not informative enough to determine a final set of plausible parameter
-values. For this reason, we will additionally look at insightful model
-based quantities.
-
-\subsubsection{Insightful model based
-quantities}\label{insightful-model-based-quantities}
+deviations would produce data that aligned with their expectations.
+Although the result was deemed plausible, these statistics were not
+informative enough to determine a final set of plausible parameter
+values (e.g., doubling the standard deviations from 0.5 to 1 produces
+only minor changes in relative frequency). For this reason, we will
+additionally look at insightful model-based quantities.
+
+\subsubsection{Examine insightful model-based
+quantities}\label{examine-insightful-model-based-quantities}
 
@@
 control over parameter values and sample sizes) is that one can produce
-insightful model based quantities that can never be directly observed in
-an actual empirical dataset. For example, in a logistic model with
-random intercepts for participants, one can produce a visualization of
-the implied distribution of the probability that a participant on
-average solves a cognitive task. Although domain knowledge will probably
-not suffice to specify this distribution completely, it should be
-possible to rule out implausible boundary conditions. For example, the
-domain expert might deem it implausible that the 5\% most able
-participants have a probability of more than 0.99 to solve the difficult
-cognitive task.
+\emph{insightful model-based quantities} that can never be directly
+observed in an actual empirical dataset. For example, in a logistic
+model with random intercepts for participants, one can produce a
+visualization of the implied distribution of the probability that a
+participant, on average, solves a cognitive task. Although domain
+knowledge will probably not suffice to specify this distribution
+completely, it should be possible to rule out implausible boundary
+conditions. For example, the domain expert might deem it implausible
+that the 5\% most able participants have a probability of more than 0.99
+to solve the difficult cognitive task.
 
@@
 computed descriptive statistics, the relative frequencies of correct
-diagnoses, for each experimental condition. We will now use the model
-implied probability of each simulated data point (stored in the variable
-\texttt{y\_prob}) to visualize the whole model implied marginal
+diagnoses, for each experimental condition. We will now use the
+model-implied probability of each simulated data point (stored in the
+variable \texttt{y\_prob}) to visualize the whole model-implied marginal
 distribution of correct diagnoses for each experimental condition.
@@
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
-\StringTok{"no expert, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
-\StringTok{"no expert, wrong advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, wrong advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
-\StringTok{"no expert, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
+\StringTok{"student, no advice"} \OtherTok{=} \StringTok{"0:0:0"}\NormalTok{, }\StringTok{"expert, no advice"} \OtherTok{=} \StringTok{"1:0:0"}\NormalTok{, }
+\StringTok{"student, incorrect advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, incorrect advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
+\StringTok{"student, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ y\_prob, }\AttributeTok{y =}\NormalTok{ condition)) }\SpecialCharTok{+}
@@
 
-Figure~\ref{fig-margdist1} shows the model implied marginal
+Figure~\ref{fig-margdist1} shows the model-implied marginal
 distributions, including the mean, 66\% and 95\% intervals. We can see
 that, indeed, the average probabilities (black dots) slightly differ
-from the probabilities of average subjects and items considered in the
-previous section. This difference increases with the variability of the
-random effects. We can use plots like the one above as a useful tool to
-decide whether the specified standard deviations of the subject and item
-random intercepts (\(\sigma_S\) and \(\sigma_I\)) are reasonable by
-comparing the ranges and overlap between conditions to domain knowledge.
+from the probabilities of average subjects and items considered
+previously. This difference increases with the variability of the random
+effects. We can use plots like Figure~\ref{fig-margdist1} as a useful
+tool to decide whether the specified standard deviations of the subject
+and item random intercepts (\(\sigma_S\) and \(\sigma_I\)) are
+reasonable by comparing the ranges and overlap between conditions to
+domain knowledge.
 
@@
 
-In the next plot, we have set the item standard deviation to almost zero
-(\(\sigma_I = 0.01\)). This gives us a better way to see the variability
-between persons. As an example, Figure~\ref{fig-margdist2} reveals a
-number of implicit assumptions about the comparison between experts and
-non-experts: With wrong advice, virtually all experts have a higher
-probability of making a correct diagnosis compared to non-experts when
-considering only items with average difficulty. In contrast, there is
-considerable overlap in probability between experts and non-experts with
+For the next plot, we have set the item standard deviation to almost
+zero (\(\sigma_I = 0.01\)). This gives us a better way to see the
+variability between persons. As an example, Figure~\ref{fig-margdist2}
+reveals a number of implicit assumptions about the comparison between
+experts and students: With incorrect advice, virtually all experts have
+a higher probability of making a correct diagnosis compared to students
+when considering only items with average difficulty. In contrast, there
+is considerable overlap in probability between experts and students with
 no advice and even higher overlap with correct advice. Patterns like
@@
 should be adjusted if the implications do not seem reasonable. We could
-also have a closer look at variability between items by setting the
+also have a closer look at the variability between items by setting the
 subject standard deviation to almost zero (\(\sigma_S = 0.01\)).
@@
 The final plot demonstrates that these plots are also useful for
-spotting standard deviations that are specified too high. For
+spotting standard deviations that were specified too high. For
 Figure~\ref{fig-margdist3} we have set \(\sigma_S = 3\) and
 \(\sigma_I = 3\). This implies that in each experimental condition, the
-probabilities that a subject solves an item are usually close to either
-0 or 1, which is not a plausible assumption. These high standard
-deviations do not account for the inherent variability and complexity of
-human performance. For example, we would expect that a participant with
-low ability compared to other task experts to solve a difficult item
-with a probability substantially larger than zero even when presented
-with wrong advice.
+probabilities that a subject solves an item are overwhelmingly close to
+either 0 or 1 and nothing in between, which is not a plausible
+assumption. These high standard deviations do not account for the
+inherent variability and complexity of human performance. For example,
+the expectation that an expert with low ability and incorrect advice
+would solve a difficult item with a probability close to zero is not
+convincing.
 
@@
 
-\subsubsection{Iterative process with domain
-experts}\label{iterative-process-with-domain-experts}
+\subsubsection{Iterate with domain
+experts}\label{iterate-with-domain-experts}
 
@@
 
-In our experience, the gathering of domain knowledge by domain experts
-and the consecutive specification of population parameter values used in
-data simulation is not a one-time event but rather an iterative process.
-In a first step, domain experts can be interviewed to ``elicit'' their
-domain knowledge about how the future data of the planned study is
-expected to look like. As most domain experts are no experts in
-statistical modeling and GLMMs, they often struggle without further
-guidance to communicate their knowledge in a way that is useful when
-specifying the parameters for data simulation. For this reason, we
-suggest that after an initial unstructured interview of domain experts,
-the analyst who is familiar with the structure of the GLMM under study
-selects an initial set of insightful descriptive statistics and model
-based quantities. Then they reenter into an iterative discussion where
-some set of population values are selected and the plausibility of
-resulting implied quantities are discussed with the domain experts.
-Then, the population parameters are updated based on this discussion
-until the domain experts are satisfied with the final result. During
-this process, the monitored model based quantities and descriptive
-statistics can be updated or extended to capture as much available
+Gathering domain knowledge and the specification of population parameter
+values is an iterative process, in which domain experts repeatedly
+assess the plausibility of simulated data. As a first step, domain
+experts can be interviewed to ``elicit'' their domain knowledge about
+how the data of the planned study is expected to look. As most domain
+experts are no experts in statistical modeling and GLMMs, they often
+struggle without further guidance to communicate their knowledge in a
+way that is useful when specifying the parameters for data simulation.
+For this reason, we suggest that after an initial interview, the analyst
+who is familiar with the structure of the GLMM selects an initial set of
+insightful descriptive statistics and model-based quantities. Then they
+reenter into an iterative discussion with the domain experts where some
+set of population values are selected, and the plausibility of resulting
+implied quantities is evaluated. Then, the population parameters are
+updated until the domain experts are satisfied with the result. During
+this process, the monitored descriptive statistics and model-based
+quantities can be updated or extended to capture as much available
 domain knowledge as possible.
@@
 All parameter values in our present case study have been determined
-based on repeated discussions with our affiliated domain experts in
-Radiology to validate our assumptions. Initially, we reviewed the
-literature to establish a reasonable baseline performance rate for
-examining head CT scans for intracranial hemorrhage. Existing studies
-indicate that radiologists typically demonstrate high accuracies, often
-exceeding or hovering around 90\%, while interns have been shown to
-perform below 80\%, and medical students fall even shorter. For
-simplicity, we assumed plausible probability values of .90 for experts
-and .70 for non-experts, respectively. Our experts confirmed that these
-values are realistic baselines for reviewing diverse head CT images
-without AI assistance. Subsequently, we consulted several published
-papers investigating the effect of correct and incorrect advice on
-decision-making performance in other settings. From their findings, we
-inferred that both experts and non-experts should benefit from correct
-and suffer losses from incorrect advice. However, the magnitude of these
-effects should be substantially greater for non-experts, given their
-demonstrated reliance on advice compared to experts. We further
-validated the plausibility of our estimated gains and losses with the
-collaborating radiologists. For our simulation, we used the
-probabilities of average participants to solve an average case, as shown
-in Table 1.
+based on repeated discussions with domain experts in radiology to
+validate our assumptions. Initially, we reviewed the literature to
+establish a reasonable baseline performance rate for examining head CT
+scans for intracranial hemorrhage. Existing studies indicate that
+radiologists typically demonstrate high accuracies around 90\%, while
+interns have been shown to perform below 80\%, and medical students fall
+even shorter. For simplicity, we assumed plausible probability values of
+.90 for experts and .70 for students, respectively. Our experts
+confirmed that these values are realistic baselines for reviewing
+diverse head CT images without AI assistance. Subsequently, we consulted
+several published papers investigating the effect of correct and
+incorrect advice on decision-making performance in other settings. From
+their findings, we inferred that both experts and students should
+benefit from correct and suffer losses from incorrect advice. However,
+the magnitude of these effects should be substantially greater for
+students, given their stronger reliance on advice compared to experts.
+We further validated the plausibility of our estimated gains and losses
+with the collaborating radiologists until we settled on the
+probabilities shown in Table~\ref{tbl-probtable}.
 
@@
 At this point, the researcher is capable of producing a simulated
-dataset similar to the actual dataset that will later be collected in
-the planned study. The next step is to specify how the statistical model
-shall be estimated in the actual study collected later. This usually
-includes the selection of 1) a statistical framework (e.g., frequentist
-statistics), 2) a software package that is capable of estimating the
-model class of interest (e.g., the lme4 R package), 3) an estimation
-algorithm (e.g., the default optimizer ``bobyqa''), and 4) the specific
-model structure including all fixed effects, random effects, and the
-model family of the dependent variable.\\
+dataset that is comparable to the actual dataset to be collected in the
+planned study. The next step is to specify how the statistical model
+shall be estimated in the actual study collected later. This requires
+the selection of a statistical framework, a software package that is
+capable of estimating the model class of interest, an estimation
+algorithm, and the specific model structure including all fixed effects,
+random effects, and the model family of the dependent variable.
+
 Note that this does not always mean that one will specify the same GLMM
-that was used when specifying the data generating process. On the one
+that was used when specifying the data-generating process. On the one
 hand, using a simpler model for data simulation than for model
@@
 scenarios where one has specific domain knowledge about aspects of the
-data generating process that are still difficult to estimate with the
+data-generating process that are still difficult to estimate with the
 current state-of-the-art in multilevel modeling.
@@
 
-In our case study, we use the lme4 R package
+In our case study, we use the \emph{lme4} R package
 (\citeproc{ref-batesFittingLinearMixedEffects2015}{Bates et al., 2015}),
 which is a state-of-the-art tool for fitting frequentist
-GLMMs.\footnote{For Bayesian GLMMs, the brms R package is currently the
-  most prominent option
+GLMMs.\footnote{For Bayesian GLMMs, the \emph{brms} R package is
+  currently the most prominent option
   (\citeproc{ref-burknerBrmsPackageBayesian2017}{Bürkner, 2017}).} For
@@
 
-The lme4 package uses a special syntax for model specification. Our
-specific GLMM is represented by the formula:
+The \emph{lme4} package uses a special syntax for model specification.
+Our specific GLMM is represented by the formula:
 
@@
 (\citeproc{ref-batesFittingLinearMixedEffects2015}{2015}) and in the
-documentation of the lme4 package.
-
-In lme4, a GLMM is fitted with the \texttt{glmer} function. By setting
-\texttt{family\ =\ \ "binomial"}, we request a binomial GLMM appropriate
-for our binary dependent variable \texttt{y\_bin} (the binomial GLMM
-uses the canonical logit link by default), which is defined as an
-accurate (1) vs.~inaccurate (0) diagnosis.
+documentation of the \emph{lme4} package.
+
+In \emph{lme4}, a GLMM is fitted with the \texttt{glmer} function. By
+setting \texttt{family\ =\ \ "binomial"}, we request a binomial GLMM
+appropriate for our binary dependent variable \texttt{y\_bin} (the
+binomial GLMM uses the canonical logit link by default), which is
+defined as an accurate (1) vs.~inaccurate (0) diagnosis. We use the
+default estimation algorithm (see \texttt{?glmerControl} for a list of
+alternative options).
 
@@
 
-In the lme4 output, the \texttt{Estimate} column in the
+In the model output, the \texttt{Estimate} column in the
 \texttt{Fixed\ effects} table contains the estimates for the \(\beta\)
@@
 point estimate of the theoretical estimand within the framework of the
-fitted GLMM. For some research question, the estimate corresponds with a
-single regression coefficient. In more complicated scenarios, the
+fitted GLMM. For some research questions, the estimate corresponds with
+a single regression coefficient. In more complicated scenarios, the
 estimate is computed from a combination of coefficients. Beyond
 computing the point estimate, we have already discussed that both
-hypothesis testing and interval estimation can be used to answer the
+hypothesis testing and interval estimation could be used to answer the
 research question. The decision on testing or estimating is then
 followed by selecting the specific statistical method that shall be
-applied to compute the HT(s) or CI(s) (e.g., compute HTs and CIs with
-the marginaleffects R package using the delta method).
+applied to compute the HTs or CIs (e.g., compute HTs and CIs with the
+\emph{marginaleffects} R package using the delta method).
 
@@
 \] We have already discussed how to compute the involved probabilities
-in the section on specifying population parameters. Plugging in the
-model equation of the GLMM produces an equation on how to compute each
-contrast if all model parameters were known. When we want to estimate
-the above contrasts based on \emph{observed} data, the only difference
-is that model parameters are not known and we instead use the
-corresponding parameter \emph{estimates}.
+in the section on specifying population parameters. Plugging in the GLMM
+model equation produces an equation for computing each contrast if all
+model parameters were known. When we want to estimate the above
+contrasts based on observed data, the only difference is that model
+parameters are not known and we instead use the corresponding parameter
+estimates.
 
@@
 advice, expert) - (no advice, expert)} as our example, we could apply
-the \texttt{predict} function of the lme4 package to compute the
+the \texttt{predict} function of the \emph{lme4} package to compute the
 predicted probability for a correct diagnosis based on our fitted model,
@@
 whether we are interested in hypothesis testing or parameter estimation,
-we also need a method to compute HTs or CI. The marginaleffects package
-(\citeproc{ref-R-marginaleffects}{Arel-Bundock, 2024}) is a very
-flexible, increasingly popular package to compute HTs and CIs for
-contrasts with a variety of statistical models, including GLMMs
-estimated with lme4. First, we specify a grid of all combinations of
-predictor variable and then compute estimated probabilities for all
-experimental conditions in our experiment with the \texttt{predictions}
-function:
+we also need a method to compute HTs or CIs. The \emph{marginaleffects}
+package (\citeproc{ref-R-marginaleffects}{Arel-Bundock et al.,
+Forthcoming}) is a very flexible, increasingly popular package to
+compute HTs and CIs for contrasts with a variety of statistical models,
+including GLMMs estimated with \emph{lme4}. First, we specify a grid of
+all combinations of predictor variables and then compute estimated
+probabilities for all experimental conditions in our experiment with the
+\texttt{predictions} function:
 
@@
 estimated probabilities to compute the four specific contrasts that are
-part of our estimand. For this we must specify which rows in
-\texttt{probs} have to be subtracted from each other. We will use the
+part of our estimand. For this, we must specify which rows in
+\texttt{preds} have to be subtracted from each other. We will use the
 \texttt{hypotheses} function to compute our four contrasts of interest
 together with HTs and CIs. We use the default inference options of the
-marginaleffects package that compute HTs and CIs based on the
+\emph{marginaleffects} package that compute HTs and CIs based on the
 approximate delta method.
@@
     \StringTok{"b8 = b2"}\NormalTok{,  }\CommentTok{\# (correct advice, expert) {-} (no advice, expert)}
-    \StringTok{"b2 = b6"}\NormalTok{,  }\CommentTok{\# (no advice, expert) {-} (wrong advice, expert) }
-    \StringTok{"b7 = b1"}\NormalTok{,  }\CommentTok{\# (correct advice, no expert) {-} (no advice, no expert)}
-    \StringTok{"b1 = b5"}\NormalTok{), }\CommentTok{\# (no advice, no expert) {-} (wrong advice, no expert)}
+    \StringTok{"b2 = b6"}\NormalTok{,  }\CommentTok{\# (no advice, expert) {-} (incorrect advice, expert) }
+    \StringTok{"b7 = b1"}\NormalTok{,  }\CommentTok{\# (correct advice, student) {-} (no advice, student)}
+    \StringTok{"b1 = b5"}\NormalTok{), }\CommentTok{\# (no advice, student) {-} (incorrect advice, student)}
     \AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{))}
@@
 estimate in row number 8 from the estimate in row number 2 in the
-\texttt{probs}-output. The argument \texttt{equivalence\ =\ c(0,\ 0)}
+\texttt{preds}-output. The argument \texttt{equivalence\ =\ c(0,\ 0)}
 can be used to compute one-sided p-values, testing whether the contrast
@@
 the two-sided CIs of the four contrasts of interest. With a standard
-confidence level of \(1 - \alpha = 0.95\), plausible values are clearly
-in the positive range for the first, third and fourth contrast, while
+confidence level of \(1 - \alpha = 0.95\) plausible values are clearly
+in the positive range for the first, third, and fourth contrast, while
 both negative and positive values seem plausible for the second
@@
 observed for the first contrast (expert with correct advice vs.~expert
-without advice) where both underlying probabilities are close to 1. The
-largest width is observed for the fourth contrast (non-expert with wrong
-advice vs.~non-expert without advice), where both underlying
+without advice), where both underlying probabilities are close to 1. The
+largest width is observed for the fourth contrast (student with
+incorrect advice vs.~student without advice), where both underlying
 probabilities are closer to 0.5.
@@
 dataset, 2) estimate a GLMM, and 3) compute HTs or CIs for estimands of
-interest, mirroring the analysis that will later be performed for the
-actual dataset of the planned study. The last missing piece is to write
-code to perform the above steps repeatedly and allow for a setting using
-different sample sizes. On a conceptual level, we first require a
-function that takes as input the sample size and the full set of
-population parameter values. When planning for power, the function
-should return the p-value(s) of the HT(s) of interest when conducted on
-the simulated dataset. When planning for precision, the function should
-return the width of the CI(s) of interest. Secondly, we must run this
-function repeatedly with the same sample size and population parameters.
-Because even fitting GLMMs with frequentist methods can quickly become
+interest, mirroring the analysis that will later be conducted on the
+actual dataset of the planned study. To produce an estimate of power or
+precision, the last step is to perform these previous steps repeatedly.
+On a conceptual level, we first require a function that takes a sample
+size and a full set of population parameter values as input. When
+planning for power, the function should return the p-value(s) of the
+HT(s) of interest when conducted on the simulated dataset. When planning
+for precision, the function should return the width of the CI(s) of
+interest. Second, we run this function repeatedly with the same sample
+size and population parameters. Because fitting GLMMs can quickly become
 time-consuming, it is recommended to use parallel computing, that is
 running simulations on multiple cores of the computer at the same time
-to reduce total run time. Thirdly, the results of the repeated
-simulation must be collected and aggregated. When planning for power, we
-compute the relative frequency of (a) significant p-value(s) across
-repeated simulations. When planning for precision, we compute the
-average width of the CI(s). Lastly, we have to repeat the complete
-simulation for different sample sizes, to determine how big the sample
-must be in order to achieve the targeted power or precision.
+to reduce total run time. Third, the results of the repeated simulation
+are collected and aggregated. When planning for power, we compute the
+relative frequency of (a) significant p-value(s) across repeated
+simulations. When planning for precision, we compute the average width
+of the CI(s). Finally, we repeat the complete simulation for different
+sample sizes to determine how big the sample must be in order to achieve
+the targeted power or precision.
 
@@
 
-We are finally ready to run our simulation-based sample size planning
-analyses to plan for power and for precision. Wrapping the
-\texttt{simulate} function already constructed earlier, the helper
-function \texttt{sim\_and\_analyse} performs all previous steps
+Wrapping the \texttt{simulate} function already constructed earlier, the
+helper function \texttt{sim\_and\_analyse} performs all previous steps
 (simulate a dataset, fit a GLMM, compute p-values and CIs) in a single
@@
 
-Simulation-based sample size planning can quickly become computationally
-intensive when we repeatedly simulate data and fit models for different
-parameter combinations or sample sizes. Thus, we use the future
-(\citeproc{ref-R-RJ-2021-048}{Bengtsson, 2021}) and furrr
-(\citeproc{ref-R-furrr}{Vaughan \& Dancho, 2022}) packages to perform
-computations in parallel. First, we enable parallelization with the
-\texttt{plan} function and specify how many parallel cores (``workers'')
-of our computer to use (users can find out the maximum number of cores
-on their computer with the command \texttt{parallel::detectCores()}),
-and set a seed to make the simulation reproducible.
+We use the \emph{future} (\citeproc{ref-R-RJ-2021-048}{Bengtsson, 2021})
+and \emph{furrr} (\citeproc{ref-R-furrr}{Vaughan \& Dancho, 2022})
+packages to perform computations in parallel. First, we enable
+parallelization with the \texttt{plan} function and specify how many
+parallel cores (``workers'') of our computer to use (users can find out
+the maximum number of cores on their computer with the command
+\texttt{parallel::detectCores()}), and set a seed to make the simulation
+reproducible.
 
@@
 items (\texttt{n\_items}), each combination being repeated \texttt{rep}
-times. We chose 300 repetitions for the data simulation at hand as it
-strikes a balance between achieving a robust statistical estimate and
-remaining computationally feasible. With the current settings, this
-simulation takes about one hour on a MacBook Pro from 2020 with M1 chip
-and 16 GB working memory. If you want to quickly experiment with the
-code yourself, a setting with \texttt{workers\ =\ 4} and
-\texttt{rep\ =\ 5} should finish in less than 5 minutes, even on smaller
-machines.
+times. The more repetitions, the more accurately power and precision can
+be estimated. We chose 300 repetitions for the data simulation at hand
+to strike a balance between achieving a robust estimate and remaining
+computationally feasible. With the current settings, this simulation
+takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB
+working memory. If you want to quickly experiment with the code
+yourself, a setting with \texttt{workers\ =\ 4} and \texttt{rep\ =\ 5}
+should finish in less than 5 minutes, even on smaller machines.
 
@@
 In some iterations (predominantly in conditions with small sample
-sizes), model estimation did not converge with the lme4 package. When
-the model fails to converge, it means that the statistical model being
-fitted to the data failed to reach a stable or valid solution during the
-estimation process. We do not remove these results because
+sizes), model estimation did not converge with the \emph{lme4} package.
+When the model fails to converge, it means that the statistical model
+being fitted to the data failed to reach a stable or valid solution
+during the estimation process. We do not remove these results because
 non-convergence can also happen when analyzing the real data we plan to
@@
 code.\footnote{This code was inspired by the ``Mixed Design Simulation''
-  vignette of the faux package at
+  vignette of the \emph{faux} package at
   \url{https://debruine.github.io/faux/articles/sim_mixed.html}.}
@@
 precision for the final power estimates, which are then used for the
-sample size justification of the future study.
-
-Much has been written on the optimal amount of power to target in
-empirical research. The most prominent heuristic is to target a power of
-0.8 (when combined with a type I error rate of \(\alpha = 0.05\)), but
-depending on the research goals of the study, there are often good
-reasons to move away from this standard depending on the research goals
-and resource constraints
-(\citeproc{ref-lakensJustifyYourAlpha2018}{Lakens, Adolfi, et al.,
-2018}; \citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}).
+sample size justification of the planned study.
+
 When target power has been specified, the number of subjects and the
 number of items in our study design can be traded against each other
-based on practical considerations. For the sake of the example, let the
-targeted power be indeed about 0.8, using an \(\alpha\) of 0.05 to
+based on practical considerations. Although it is recommended to justify
+the power (\citeproc{ref-lakensSampleSizeJustification2022}{Lakens,
+2022a}), we adopt the common heuristic to target a power of 0.8 to
 detect an effect of the expected size implied by our data simulation.
@@
 the simulation to find a more optimal trade-off, are the ratio of
-experts, the frequency of whether advice is presented and whether it is
-correct.
+experts to students, the frequency of whether advice is presented at
+all, and whether it is correct or incorrect.
 
@@
 to planning for power, there are no common standards on how to specify
-the targeted precision. For our example, we use a simple heuristic but
-we strongly encourage readers to think about better alternatives that
-are appropriate in their own applications. Our simulations show that the
-smallest CI can be expected for the first contrast (expert with correct
-advice vs.~expert without advice). The true contrast in probability for
-an average expert and an average item in this condition is
+the targeted precision
+(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}). For
+our example, we use a simple heuristic but we strongly encourage readers
+to think about better alternatives that are appropriate in their own
+applications. Our simulations show that the smallest CI can be expected
+for the first contrast (expert with correct advice vs.~expert without
+advice). The true contrast in probability for an average expert and an
+average item in this condition is
 \texttt{plogis(b\_0\ +\ b\_e\ +\ b\_a\ +\ b\_c\ +\ b\_ea\ +\ b\_ec)\ -\ plogis(b\_0\ +\ b\_e)\ =}
@@
 width of a 95\% CI for the probability contrast between experts with
-correct advice and experts without advice. Of course, lower width
+correct advice and experts without advice. Of course, a lower width
 implies better precision. Based on our simulation outcomes, we compute
@@
 Figure~\ref{fig-finalprecision} increase (i.e., average width of CI
-decreases) with both the number of subjects and the number of items. The
-CIs reported here indicate how precisely the expected width of the CI
-for our focal contrast was estimated by our simulation. Applying our
-simple heuristic of targeting an expected width smaller than 0.1, we see
-the same trade-off between the number of subjects and the number of
+decreases) with the number of included subjects and items. The CIs
+reported here indicate how precisely the expected width of the CI for
+our focal contrast was estimated by our simulation. Applying our simple
+heuristic of targeting an expected CI width that is smaller than 0.1, we
+see the same trade-off between the number of subjects and the number of
 items as with planning for power. We could either choose 100 subjects
@@
 heuristic for determining sample size in the planning for precision
-scenario was quite liberal. This is reflected by the result that we
+scenario was quite liberal. This is reflected in the result that we
 would need a smaller sample size than in the planning for power
@@
 
-\subsection{Sensitivity analysis}\label{sensitivity-analysis}
+\subsection{Step 7 (optional): Sensitivity
+analysis}\label{step-7-optional-sensitivity-analysis}
 
@@
 we have demonstrated some strategies to determine plausible parameter
-values in GLMMs based on domain knowledge. Domain knowledge can be
-considered a vague theoretical model about the data-generating process
-that is less formal and can only be accessed by a back-and-forth
-exchange in which domain experts assess the plausibility of simulated
-data. When sample sizes are chosen based on the results of our
-simulation-based power analysis, a future study will be informative to
-reject the null hypothesis if an effect of our \emph{expected size} is
-present (or estimate the effect with satisfying precision). However, if
-the true effect is indeed smaller, power (or precision) will be lower,
-and the study might not be sufficiently informative. A common, more
-conservative strategy for sample size justification is to perform sample
-size planning for the smallest effect size of interest (SESOI). An
-effect smaller than the SESOI would be considered too small to be
-interesting or practically meaningful, even if the effect is not
-actually zero (\citeproc{ref-kingPointMinimalImportant2011}{King, 2011};
+values in GLMMs based on domain knowledge. When sample sizes are chosen
+based on the results of our simulation-based power analysis, a future
+study will be informative to reject the null hypothesis if an effect of
+our \emph{expected size} is present (or estimate the effect with
+satisfying precision). However, if the true effect is indeed smaller,
+power (or precision) will be lower, and the study might not be
+sufficiently informative. A common, more conservative strategy for
+sample size justification is to perform sample size planning for the
+smallest effect size of interest (SESOI). An effect smaller than the
+SESOI would be considered too small to be interesting or practically
+meaningful, even if the effect is not actually zero
+(\citeproc{ref-kingPointMinimalImportant2011}{King, 2011};
 \citeproc{ref-lakensEquivalenceTestingPsychological2018}{Lakens, Scheel,
-et al., 2018}). For strategies on the even more difficult task of
-specifying a plausible SESOI, as well as a thorough discussion of
-various topics concerning power analysis, see
-(\citeproc{ref-lakensImprovingYourStatistical2022}{Lakens, 2022b}). When
+et al., 2018}). Unfortunately, specifying a plausible SESOI is a
+challenging task (for strategies on how to specify a SESOI, see Lakens
+(\citeproc{ref-lakensSampleSizeJustification2022}{2022a}), Riesthuis
+(\citeproc{ref-riesthuisSimulationBasedPowerAnalyses2024}{2024}), or
+Lakens (\citeproc{ref-lakensImprovingYourStatistical2022}{2022b})). When
 domain knowledge or formal theories about the research topic of interest
 are too vague to specify a meaningful SESOI, it is still recommended to
-demonstrate power or precision for different effect sizes in what is
-called \emph{sensitivity power analysis}. By simulating power (or
-precision) for different effect sizes (in addition to the different
-number of subjects and items), one can make sure that power (or
-precision) would still be sufficient to detect smaller effect sizes than
-our expected effect or at least get an impression of how strongly power
-(or precision) depends on the size of the true effect. For our case
-study that investigates combined hypotheses in a GLMM modeling
-framework, sensitivity analysis would require manually specifying
-additional sets of plausible parameter values that reflect scenarios
-with smaller or larger differences between groups with respect to our
-specific research question. Power (or precision) could then be simulated
-for several of these scenarios (across different numbers of subjects and
-items, as considered earlier).
+demonstrate power or precision for different effect sizes in a
+\emph{sensitivity power analysis}
+(\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}). By
+simulating power (or precision) for different effect sizes (in addition
+to the different number of subjects and items), one can make sure that
+power (or precision) would still be sufficient to detect smaller effect
+sizes than our expected effect or at least get an impression of how
+strongly power (or precision) depends on the size of the true effect.
+For our case study that investigates combined hypotheses in a GLMM
+modeling framework, sensitivity analysis would require manually
+specifying additional sets of plausible parameter values that reflect
+scenarios with smaller or larger differences between groups with respect
+to our specific research question. Power (or precision) could then be
+simulated for several of these scenarios (across different numbers of
+subjects and items, as considered earlier). Although recommended, we do
+not present a sensitivity analysis for our case study to keep the length
+of the tutorial manageable.
 
@@
 
-The goal of this tutorial was to teach researchers how to perform
-tailored simulation-based sample size planning for GLMMs. Beyond the
-specifics of our concrete case study, we want to outline six
+In this tutorial, we provided a step-by-step guide on how to perform
+tailored simulation-based sample size planning for GLMMs based on a
+concrete case study. To conclude, we want to give an outlook on five
 developments regarding the future role of simulation-based sample size
@@
 
-In light of the ongoing replication crisis and an existing literature
-full of underpowered studies, there is a growing need for
-simulation-based sample size planning in experimental research: In order
-to conduct informative research, GLMMs offer a flexible statistical
-framework to analyze complex experimental study designs. However,
-existing formula-based heuristics and user-friendly software tools for a
-priori power analysis are often not sufficient. Therefore,
-simulation-based power analysis is becoming increasingly needed since it
-provides experimental researchers with a tailored approach to estimating
-required sample sizes before data collection.
-
-Managing data simulations more easily with discrete predictor variables:
-Simulation-based sample size planning becomes more manageable when all
-predictor variables are discrete (like in the presented case study) and
-fixed by the study design. This allows researchers to focus on
-simulating outcome variables while avoiding the need for complex
-simulations of predictor values, which would introduce additional
-assumptions. By simplifying the simulation process, researchers can
-obtain reliable estimates for power or precision without compromising
-realistic assumptions about the data-generating process implied by the
-study design.
-
-Teaching data simulation skills: The ability to conduct simulation-based
-sample size planning is a valuable skill that should be taught to
-experimental researchers. By incorporating such training into research
-methods courses and workshops, researchers can gain a deeper
-understanding of statistical power or precision, and improve the quality
-of their experimental designs. Equipping researchers with the knowledge
-and tools to perform simulation-based sample size planning enables them
-to make informed decisions and enhance the rigor of their studies. The
-need to reason about how to simulate plausible data that is in line with
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  As experimental designs become more complex and the appropriate
+  flexible statistical frameworks, like GLMMs, become more popular,
+  there is an ever growing need for simulation-based sample size
+  planning in experimental research.
+\item
+  The ability to conduct simulation-based sample size planning becomes
+  an increasingly valuable skill that should be taught to experimental
+  researchers. By incorporating such training into research methods
+  courses and workshops, researchers can improve the quality of their
+  experimental designs and enhance the rigor of their studies. The need
+  to reason about how to simulate plausible data that is in line with
   the research hypothesis, while not violating domain expertise on how
-plausible data should look, might also contribute to planning more
+  plausible data should look might also contribute to planning more
   insightful studies that can answer more precise research questions
   (\citeproc{ref-yarkoniGeneralizabilityCrisis2022}{Yarkoni, 2022}).
-
-Addressing the mismatch in effort perception: There is often a
-significant disconnect between the amount of effort required to perform
-tailored simulation-based sample size planning and the perceived effort
-estimated by researchers and collaborators in experimental research.
-Many researchers request simulation-based power analyses from
-statisticians or methodological experts without fully comprehending the
-complexity and time-consuming nature of these tailored simulations. It
-is crucial to raise awareness about the effort involved to ensure
-realistic expectations and effective collaboration between researchers
-and methodological experts.
-
-Recognizing the value of simulation-based design analysis: Tailored data
-simulations and power analyses are not mere technicalities; they are
-valuable research contributions that deserve recognition in experimental
-research. They offer insights into the robustness and sensitivity of
-experimental designs, helping researchers make informed decisions about
-sample sizes, effect sizes, and statistical power or precision. Their
-importance can be reflected by allocating them a separate publication or
-incorporating them as a significant component of stage 1 preregistered
-reports (\citeproc{ref-chambersPresentFutureRegistered2022}{Chambers \&
+\item
+  Given the significant disconnect between the amount of effort required
+  to perform tailored simulation-based sample size planning and the
+  perceived effort estimated by researchers and collaborators in
+  experimental research, there is a need to address the mismatch in
+  effort perception. Many researchers request simulation-based power
+  analyses from statisticians or methodological experts without fully
+  comprehending the complexity and time-consuming nature of these
+  tailored simulations. Therefore, it is crucial to raise awareness
+  about the effort involved to ensure realistic expectations and
+  effective collaboration with methodological experts.
+\item
+  Tailored data simulations and power analyses are not mere
+  technicalities; they are valuable research contributions that deserve
+  adequate recognition in experimental research. Their importance can be
+  reflected by highlighting the simulation work in a publication or even
+  allocating them a separate publication, or incorporating them as a
+  significant component of stage 1 preregistered reports
+  (\citeproc{ref-chambersPresentFutureRegistered2022}{Chambers \&
   Tzavella, 2022}).
-
-Integration with Open Science and preregistration practices:
-Simulation-based sample size planning aligns well with the principles of
-Open Science and preregistration in experimental research. When
-researchers have access to simulated data based on their pre-specified
-model, analyzing the collected dataset becomes straightforward and
-unambiguous. By preregistering their simulation-based sample size plan,
-researchers enhance the transparency and accountability of their
-experimental procedures, contributing to the credibility and
-reproducibility of research.
+\item
+  Simulation-based sample size planning aligns well with the principles
+  of Open Science and preregistration and could be further integrated.
+  When researchers have access to simulated data based on their
+  pre-specified model, analyzing the collected dataset becomes
+  straightforward and unambiguous. By preregistering their
+  simulation-based sample size plan, researchers enhance the
+  transparency and accountability of their experimental procedures,
+  contributing to the credibility and reproducibility of research.
+\end{enumerate}
 
@@
 \bibitem[\citeproctext]{ref-R-marginaleffects}
-Arel-Bundock, V. (2024). \emph{Marginaleffects: Predictions,
-comparisons, slopes, marginal means, and hypothesis tests}.
-\url{https://marginaleffects.com/}
+Arel-Bundock, V., Greifer, N., \& Heiss, A. (Forthcoming). How to
+interpret statistical models using {marginaleffects} in {R} and
+{Python}. \emph{Journal of Statistical Software}.
 
@@
 208--227. \url{https://doi.org/10.32614/RJ-2021-048}
+
+\bibitem[\citeproctext]{ref-benjaminRedefineStatisticalSignificance2017}
+Benjamin, D. J., Berger, J. O., Johannesson, M., Nosek, B. A.,
+Wagenmakers, E.-J., Berk, R., Bollen, K. A., Brembs, B., Brown, L.,
+Camerer, C., Cesarini, D., Chambers, C. D., Clyde, M., Cook, T. D., De
+Boeck, P., Dienes, Z., Dreber, A., Easwaran, K., Efferson, C., \ldots{}
+Johnson, V. E. (2017). Redefine statistical significance. \emph{Nature
+Human Behaviour}, \emph{2}(1), 6--10.
+\url{https://doi.org/10.1038/s41562-017-0189-z}
+
+\bibitem[\citeproctext]{ref-bolkerLinearGeneralizedLinear2015}
+Bolker, B. M. (2015). Linear and generalized linear mixed models. In G.
+A. Fox, S. Negrete-Yankelevich, \& V. J. Sosa (Eds.), \emph{Ecological
+{Statistics}} (1st ed., pp. 309--333). Oxford University PressOxford.
+\url{https://doi.org/10.1093/acprof:oso/9780199672547.003.0014}
 
@@
 
+\bibitem[\citeproctext]{ref-brownIntroductionLinearMixedEffects2021}
+Brown, V. A. (2021). An {Introduction} to {Linear Mixed-Effects
+Modeling} in {R}. \emph{Advances in Methods and Practices in
+Psychological Science}, \emph{4}(1), 2515245920960351.
+\url{https://doi.org/10.1177/2515245920960351}
+
 \bibitem[\citeproctext]{ref-brysbaertPowerAnalysisEffect2018}
@@
 
+\bibitem[\citeproctext]{ref-buttonPowerFailureWhy2013}
+Button, K. S., Ioannidis, J. P. A., Mokrysz, C., Nosek, B. A., Flint,
+J., Robinson, E. S. J., \& Munafò, M. R. (2013). Power failure: Why
+small sample size undermines the reliability of neuroscience.
+\emph{Nature Reviews Neuroscience}, \emph{14}(5), 365--376.
+\url{https://doi.org/10.1038/nrn3475}
+
+\bibitem[\citeproctext]{ref-camererEvaluatingReplicabilitySocial2018}
+Camerer, C. F., Dreber, A., Holzmeister, F., Ho, T.-H., Huber, J.,
+Johannesson, M., Kirchler, M., Nave, G., Nosek, B. A., Pfeiffer, T.,
+Altmejd, A., Buttrick, N., Chan, T., Chen, Y., Forsell, E., Gampa, A.,
+Heikensten, E., Hummer, L., Imai, T., \ldots{} Wu, H. (2018). Evaluating
+the replicability of social science experiments in {Nature} and
+{Science} between 2010 and 2015. \emph{Nature Human Behaviour},
+\emph{2}(9), 637--644. \url{https://doi.org/10.1038/s41562-018-0399-z}
+
 \bibitem[\citeproctext]{ref-chambersPresentFutureRegistered2022}
@@
 29--42. \url{https://doi.org/10.1038/s41562-021-01193-7}
+
+\bibitem[\citeproctext]{ref-R-pwr}
+Champely, S. (2020). \emph{Pwr: Basic functions for power analysis}.
+\url{https://github.com/heliosdrm/pwr}
 
@@
 \url{https://doi.org/10.1111/2041-210X.12504}
+
+\bibitem[\citeproctext]{ref-hallgrenConductingSimulationStudies2013}
+Hallgren, K. A. (2013). Conducting {Simulation Studies} in the {R
+Programming Environment}. \emph{Tutorials in Quantitative Methods for
+Psychology}, \emph{9}(2), 43--60.
+\url{https://doi.org/10.20982/tqmp.09.2.p043}
 
@@
 
+\bibitem[\citeproctext]{ref-lakensSimulationBasedPowerAnalysis2021}
+Lakens, D., \& Caldwell, A. R. (2021). Simulation-{Based Power Analysis}
+for {Factorial Analysis} of {Variance Designs}. \emph{Advances in
+Methods and Practices in Psychological Science}, \emph{4}(1),
+251524592095150. \url{https://doi.org/10.1177/2515245920951503}
+
 \bibitem[\citeproctext]{ref-lakensEquivalenceTestingPsychological2018}
@@
 
+\bibitem[\citeproctext]{ref-leeUsingTidyversePackage2020}
+Lee, S., Sriutaisuk, S., \& Kim, H. (2020). Using the {Tidyverse
+Package} in {R} for {Simulation Studies} in {SEM}. \emph{Structural
+Equation Modeling: A Multidisciplinary Journal}, \emph{27}(3), 468--482.
+\url{https://doi.org/10.1080/10705511.2019.1644515}
+
 \bibitem[\citeproctext]{ref-littleStatisticalAnalysisMissing2014}
@@
 
+\bibitem[\citeproctext]{ref-matuschekBalancingTypeError2017}
+Matuschek, H., Kliegl, R., Vasishth, S., Baayen, H., \& Bates, D.
+(2017). Balancing {Type I} error and power in linear mixed models.
+\emph{Journal of Memory and Language}, \emph{94}, 305--315.
+\url{https://doi.org/10.1016/j.jml.2017.01.001}
+
 \bibitem[\citeproctext]{ref-maxwellSampleSizePlanning2008}
@@
 Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}
+
+\bibitem[\citeproctext]{ref-meteyardBestPracticeGuidance2020a}
+Meteyard, L., \& Davies, R. A. I. (2020). Best practice guidance for
+linear mixed-effects models in psychological science. \emph{Journal of
+Memory and Language}, \emph{112}, 104092.
+\url{https://doi.org/10.1016/j.jml.2020.104092}
 
@@
 \url{https://doi.org/10.3102/10769986031004437}
+
+\bibitem[\citeproctext]{ref-R-base}
+R Core Team. (2024). \emph{R: A language and environment for statistical
+computing}. R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}
 

### 2024-06-26 manuscript.qmd
added table 1 (r package overview) to manuscript

@@
 To our knowledge, existing approaches for sample size planning for GLMMs have exclusively focused on planning for power.
-In Table 1, we review available software packages that can be used to perform power analysis for multilevel models in `r cite_r()`.
+In @tbl-rpackages, we review available software packages that can be used to perform power analysis for multilevel models in `r cite_r()`.
 Power analysis methods can be categorized into formula-based, summary-statistics-based and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
@@
 
-INSERT TABLE 1 HERE!
+```{r}
+#| label: tbl-rpackages
+#| tbl-cap: Overview of available software packages for power analysis for multilevel models in R.
+#| echo: false
+
+packagetable <- data.frame(
+  col1 = c("glmmrBase", "longpower", "mixedpower", "mlmpower", "pamm", "pass.lme", "PowerAnalysisIL", "powerlmm", "simglm", "simr", "summary_statistics_based_power", "WebPower"),
+  col2 = c("Formula- and simulation-based power analysis; non-linear fixed effects and flexible covariance functions",
+              "Formula-based power analysis; focus on longitudinal data",
+              "Simulation-based power analysis; based on 'simr' package",
+              "Simulation-based power analysis; missing data mechanisms",
+              "Simulation-based power analysis; based on 'lme4' package",
+              "Formula-based power analysis; limited documentation",
+              "Simulation-based power analysis; focus on longitudinal data",
+              "Simulation-based power analysis; missing data mechanisms; focus on longitudinal data",
+              "Simulation-based power analysis; missing data mechanisms",
+              "Simulation-based power analysis; based on 'lme4' package",
+              "Summary-statistics-based power analysis; based on 'pwr' package",
+              "Formula-based power analysis; based on 'longpower' package; focus on longitudinal data"),
+  col3 = c("no", "yes", "no", "no", "no", "no", "yes", "no", "no", "no", "yes", "yes"),
+  col4 = c("yes", "no", "yes", "no", "no", "no", "no", "no", "yes", "yes", "yes", "no"),
+  col5 = c("none", "limited set of study designs with two levels", "none", "limited to two levels", "none", "none", "limited set of study designs with two levels", "limited set of study designs with two or three levels", "limited to two or three levels", "none", "limited set of study designs with two levels", "limited set of study designs with two levels"),
+  col6 = c("@watsonGeneralisedLinearMixed2023", "@iddiPowerSampleSize2022", "@kumleEstimatingPowerGeneralized2021", "@endersSimpleMonteCarlo2023", "@martinMeasuringIndividualDifferences2011", "@marcochakyanyuPassLmePower2019", "@lafitSelectionNumberParticipants2021", "@magnussonConsequencesIgnoringTherapist2018", "@lebeauSimglmSimulateModels2017", "@greenSIMRPackagePower2016", "@murayamaSummarystatisticsbasedPowerAnalysis2022", "@zhangPracticalStatisticalPower2018")
+)
+
+knitr::kable(packagetable, col.names = c("Package name", "Summary", "GUI available", "Supports GLMMs", "Design limitations", "Reference"), escape = FALSE)
+
+```
 
@@
 
-Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in TABLE 1.
+Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in @tbl-rpackage.
 Circumstances under which tailored simulation-based sample size planning is necessary include 1) complex study designs, 2) complex statistical hypotheses, 3) planning for precision, and 4) lack of prior studies or pilot data.
@@
 However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
-None of the software packages for sample size planning in Table 1 can handle combined hypotheses and only some can handle directed hypotheses.
+None of the software packages for sample size planning in @tbl-rpackages can handle combined hypotheses and only some can handle directed hypotheses.
 In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation-based sample size planning.
@@
 Third, all available software packages for sample size planning with GLMMs are based on power analysis and do not support planning for precision.
-Thus, researchers that want to apply an estimation strategy instead of testing statistical hypotheses [@cummingNewStatisticsWhy2014], currently cannot use the software packages outlined in TABLE 1.
+Thus, researchers that want to apply an estimation strategy instead of testing statistical hypotheses [@cummingNewStatisticsWhy2014], currently cannot use the software packages outlined in @tbl-rpackages.
 However, tailored simulation-based sample size planning can easily handle the planning for precision approach [@maxwellSampleSizePlanning2008].

### 2024-06-26 manuscript.qmd
update tbl-rpackages

--- 2024-06-25 manuscript.qmd
@@
 
-INSERT TABLE 1 HERE!
+```{r}
+#| label: tbl-rpackages
+#| tbl-cap: Overview of software packages for power analysis for (generalized) linear mixed models in R.
+#| apa-note: GUI available = whether a graphical user interface is provided (usually a Shiny app that is available online or can be downloaded to run locally), GLMM support = whether power analysis can be performed for generalized linear mixed model (otherwise only linear mixed models are supported), Design limitations = whether only a limited selection of study designs can be specified (otherwise arbitrary designs within the model class are supported).
+#| echo: false
+
+packagetable <- data.frame(
+  col1 = c("glmmrBase", "longpower", "mixedpower", "mlmpower", "pamm", "pass.lme", "PowerAnalysisIL", "powerlmm", "simglm", "simr", "summary_statistics based_power", "WebPower"),
+  col2 = c("Formula- and simulation-based power analysis; non-linear fixed effects and flexible covariance functions",
+              "Formula-based power analysis; focus on longitudinal data",
+              "Simulation-based power analysis; based on 'simr' package",
+              "Simulation-based power analysis; missing data mechanisms",
+              "Simulation-based power analysis; based on 'lme4' package",
+              "Formula-based power analysis; limited documentation",
+              "Simulation-based power analysis; focus on longitudinal data",
+              "Simulation-based power analysis; missing data mechanisms; focus on longitudinal data",
+              "Simulation-based power analysis; missing data mechanisms",
+              "Simulation-based power analysis; based on 'lme4' package",
+              "Summary-statistics-based power analysis; based on 'pwr' package",
+              "Formula-based power analysis; based on 'longpower' package; focus on longitudinal data"),
+  col3 = c("no", "[yes](https://atrihub.shinyapps.io/power/)", "no", "no", "no", "no", "[yes](https://github.com/ginettelafit/PowerAnalysisIL?tab=readme-ov-file#shiny-app-and-r-package-to-perform-power-analysis-to-select-the-number-of-participants-in-intensive-longitudinal-studies)", "[yes](https://github.com/rpsychologist/powerlmm?tab=readme-ov-file#launch-interactive-web-application)", "[yes](https://simglm.brandonlebeau.org/reference/run_shiny.html)", "no", "[yes](https://koumurayama.shinyapps.io/summary_statistics_based_power/)", "[yes](https://webpower.psychstat.org/wiki/start)"),
+  col4 = c("yes", "no", "yes", "no", "no", "no", "no", "no", "yes", "yes", "yes", "no"),
+  col5 = c("none", "limited set of study designs with two levels", "none", "limited to two levels", "none", "none", "limited set of study designs with two levels", "limited set of study designs with two or three levels", "limited to two or three levels", "none", "limited set of study designs with two levels", "limited set of study designs with two levels"),
+  col6 = c("@watsonGeneralisedLinearMixed2023", "@iddiPowerSampleSize2022", "@kumleEstimatingPowerGeneralized2021", "@endersSimpleMonteCarlo2023", "@martinMeasuringIndividualDifferences2011", "@marcochakyanyuPassLmePower2019", "@lafitSelectionNumberParticipants2021", "@magnussonConsequencesIgnoringTherapist2018", "@lebeauSimglmSimulateModels2017", "@greenSIMRPackagePower2016", "@murayamaSummarystatisticsbasedPowerAnalysis2022", "@zhangPracticalStatisticalPower2018")
+)
+
+knitr::kable(packagetable, col.names = c("Package name", "Summary", "GUI available", "GLMM support", "Design limitations", "Reference"), escape = FALSE)
+```
 
@@
 
-Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in TABLE 1.
+Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in @tbl-rpackages.
 Circumstances under which tailored simulation-based sample size planning is necessary include 1) complex study designs, 2) complex statistical hypotheses, 3) planning for precision, and 4) lack of prior studies or pilot data.

### 2024-06-26 _tex/manuscript.tex
Built site for gh-pages

@@
 To our knowledge, existing approaches for sample size planning for GLMMs
-have exclusively focused on planning for power. Power analysis methods
-for multilevel models can be categorized into formula-based methods and
-simulation-based methods
+have exclusively focused on planning for power. In
+Table~\ref{tbl-rpackages}, we review available software packages that
+can be used to perform power analysis for multilevel models in R
+{[}Version 4.4.0; R Core Team (\citeproc{ref-R-base}{2024}){]}. Power
+analysis methods can be categorized into formula-based,
+summary-statistics-based and simulation-based methods
 (\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
-et al., 2022}). Formula-based methods rely on formulas to calculate
-power directly while simulation-based methods rely on repeatedly
-simulating data with a known true effect size and estimating power
-empirically (i.e., what percentage of simulated datasets produces a
-significant p-value). Available formula-based software packages for
-multilevel models often do not include GLMMs or are limited to very
-simple designs
+et al., 2022}). Formula-based methods rely on exact formulas to
+calculate power directly. Summary-statistics-based methods use
+statistical theory to approximate power based on formula-based methods
+developed for simple t-tests. Simulation-based methods rely on
+repeatedly simulating data with a known true effect size and estimating
+power empirically, that is what percentage of simulated datasets
+produces a significant p-value. Available formula-based and
+summary-statistics-based software packages for multilevel models often
+do not include GLMMs or are limited to simple designs
 (\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
@@
 Barr, 2021}; \citeproc{ref-greenSIMRPackagePower2016a}{Green \& MacLeod,
-2016}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al., 2015};
+2016a}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al., 2015};
 \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021};
@@
 which makes specifying assumptions a practical challenge (see the
-discussion in (\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et
-al., 2008}) and
-(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al.,
-2021})).
-
-INSERT TABLE 1 HERE!
-
-In Table 1, we give a short review of existing R packages that can be
-used for power analysis for GLMMs.
-
-\section{When to use tailored data
-simulation?}\label{when-to-use-tailored-data-simulation}
+discussion in Maxwell et al.
+(\citeproc{ref-maxwellSampleSizePlanning2008}{2008}) and Kumle et al.
+(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{2021})). We will
+discuss a number of strategies on how to specify model parameters for
+application-specific, tailored data simulations in a later chapter.
+
+\begin{table}
+
+{\caption{{Overview of software packages for power analysis for
+(generalized) linear mixed models in R.}{\label{tbl-rpackages}}}
+\vspace{-20pt}}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1165}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.3947}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0526}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0489}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2030}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1842}}@{}}
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Package name
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Summary
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+GUI available
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+GLMM support
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Design limitations
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Reference
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+glmmrBase & Formula- and simulation-based power analysis; non-linear
+fixed effects and flexible covariance functions & no & yes & none &
+Watson (\citeproc{ref-watsonGeneralisedLinearMixed2023}{2023}) \\
+longpower & Formula-based power analysis; focus on longitudinal data &
+\href{https://atrihub.shinyapps.io/power/}{yes} & no & limited set of
+study designs with two levels & Iddi and C Donohue
+(\citeproc{ref-iddiPowerSampleSize2022}{2022}) \\
+mixedpower & Simulation-based power analysis; based on `simr' package &
+no & yes & none & Kumle et al.
+(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{2021}) \\
+mlmpower & Simulation-based power analysis; missing data mechanisms & no
+& no & limited to two levels & Enders et al.
+(\citeproc{ref-endersSimpleMonteCarlo2023}{2023}) \\
+pamm & Simulation-based power analysis; based on `lme4' package & no &
+no & none & Martin et al.
+(\citeproc{ref-martinMeasuringIndividualDifferences2011}{2011}) \\
+pass.lme & Formula-based power analysis; limited documentation & no & no
+& none & Marco Chak Yan YU
+(\citeproc{ref-marcochakyanyuPassLmePower2019}{2019}) \\
+PowerAnalysisIL & Simulation-based power analysis; focus on longitudinal
+data &
+\href{https://github.com/ginettelafit/PowerAnalysisIL?tab=readme-ov-file\#shiny-app-and-r-package-to-perform-power-analysis-to-select-the-number-of-participants-in-intensive-longitudinal-studies}{yes}
+& no & limited set of study designs with two levels & Lafit et al.
+(\citeproc{ref-lafitSelectionNumberParticipants2021}{2021}) \\
+powerlmm & Simulation-based power analysis; missing data mechanisms;
+focus on longitudinal data &
+\href{https://github.com/rpsychologist/powerlmm?tab=readme-ov-file\#launch-interactive-web-application}{yes}
+& no & limited set of study designs with two or three levels & Magnusson
+et al.
+(\citeproc{ref-magnussonConsequencesIgnoringTherapist2018}{2018}) \\
+simglm & Simulation-based power analysis; missing data mechanisms &
+\href{https://simglm.brandonlebeau.org/reference/run_shiny.html}{yes} &
+yes & limited to two or three levels & LeBeau
+(\citeproc{ref-lebeauSimglmSimulateModels2017}{2017}) \\
+simr & Simulation-based power analysis; based on `lme4' package & no &
+yes & none & Green and MacLeod
+(\citeproc{ref-greenSIMRPackagePower2016}{2016b}) \\
+summary\_statistics based\_power & Summary-statistics-based power
+analysis; based on `pwr' package &
+\href{https://koumurayama.shinyapps.io/summary_statistics_based_power/}{yes}
+& yes & limited set of study designs with two levels & Murayama et al.
+(\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{2022}) \\
+WebPower & Formula-based power analysis; based on `longpower' package;
+focus on longitudinal data &
+\href{https://webpower.psychstat.org/wiki/start}{yes} & no & limited set
+of study designs with two levels & Zhang and Yuan
+(\citeproc{ref-zhangPracticalStatisticalPower2018}{2018}) \\
+\end{longtable}
+
+{\vspace{-20pt}
+\noindent \emph{Note.} GUI available = whether a graphical user interface is provided (usually a Shiny app that is available online or can be downloaded to run locally), GLMM support = whether power analysis can be performed for generalized linear mixed model (otherwise only linear mixed models are supported).
+Design limitations = whether only a limited selection of study designs can be specified (otherwise arbitrary designs within the model class are supported).}
+
+\end{table}
+
+\section{Reasons to use tailored data
+simulation}\label{reasons-to-use-tailored-data-simulation}
 
@@
 complicated and time-consuming than using the existing software tools
-outlined in TABLE 1. The most important circumstances under which
-tailored simulation-based sample size planning is necessary are 1)
+outlined in Table~\ref{tbl-rpackages}. Circumstances under which
+tailored simulation-based sample size planning is necessary include 1)
 complex study designs, 2) complex statistical hypotheses, 3) planning
-for precision, 4) no available prior studies or pilot data.
-
-Requirements of real-world studies are often more complex than the
-simplified designs assumed by many user-friendly software packages for
-sample size planning. One frequent issue in applied data analysis is
-missing data, and there can be various reasons for this
-(\citeproc{ref-littleStatisticalAnalysisMissing2014}{Little \& Rubin,
-2014}). For example, data can be missing completely at random (e.g.,
-because an electronic measurement device randomly failed for some
-technical reasons). Alternatively, subjects might systematically drop
-out or produce missing data, but this dropout can be explained by some
-attributes also measured in the dataset (e.g., older subjects have a
-higher probability to refuse answering a question on income). In a more
-complicated scenario, missing data in some variable is caused by the
-measured attribute itself (e.g., wealthy people are more likely to
-refuse reporting their income). Moreover, many experimental designs
-contain conditions in which values of the predictor variables are
-missing by design. This can make data analysis more complicated because
-predictors have to be coded in specific ways that prevent the estimated
-GLMM from becoming unidentified. Whether missing data has an effect on
-the sample size planning depends on our theoretical assumptions on how
-the missingness is caused. However, it is often challenging to decide
-whether missing data can be safely ignored in the data analysis and
-sample size planning process based on a merely theoretical approach
+for precision, and 4) lack of prior studies or pilot data.
+
+First, requirements of real-world studies are often more complex than
+the simplified designs assumed by many user-friendly software packages
+for sample size planning. One frequent issue in applied data analysis is
+missing data (\citeproc{ref-littleStatisticalAnalysisMissing2014}{Little
+\& Rubin, 2014}). Missings can be completely random (e.g., an electronic
+measurement device randomly failed for technical reasons), depend on
+some attributes also measured in the dataset (e.g., older subjects
+refuse answering a question on income), or be caused by the measured
+attribute itself (e.g., wealthy people refuse to report their income).
+Moreover, many experimental designs contain conditions in which values
+of the predictor variables are missing by design. This can make data
+analysis more complicated because predictors have to be coded in
+specific ways that prevent the estimated GLMM from becoming
+unidentified. Whether missing data has an effect on the sample size
+planning depends on our theoretical assumptions on how the missingness
+is caused. However, it is often challenging to decide whether missing
+data can be safely ignored in the data analysis and sample size planning
+process based on a merely theoretical approach
 (\citeproc{ref-gomilaMissingDataExperiments2022}{Gomila \& Clark,
@@
 the global HT. None of the software packages for sample size planning in
-table X can handle combined hypotheses as discussed here and only some
+Table~\ref{tbl-rpackages} can handle combined hypotheses and only some
 can handle directed hypotheses. In contrast, our case study will
@@
 
-Although planning for precision
-(\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014};
-\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a};
-\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}) has
-been increasingly discussed as a useful strategy for empirical research,
-all available software packages for sample size planning with GLMMs are
-based on power analysis. Therefore, researchers that want to want to
-apply an estimation strategy in their studies, without testing any
-statistical hypotheses, currently cannot use the software packages
-outlined in TABLE 1. However, tailored simulation-based sample size
-planning can easily handle the planning for precision approach
+Third, all available software packages for sample size planning with
+GLMMs are based on power analysis and do not support planning for
+precision. Thus, researchers that want to apply an estimation strategy
+instead of testing statistical hypotheses
+(\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014}), currently
+cannot use the software packages outlined in Table~\ref{tbl-rpackages}.
+However, tailored simulation-based sample size planning can easily
+handle the planning for precision approach
 (\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}).
@@
 simulated dataset and estimating statistical power across repetitions,
-CIs are computed for each simulated dataset, and the expected width is
+CIs are computed for each simulated dataset, and their expected width is
 estimated.
 
-All frameworks for sample size planning require the user to make
+Fourth, all frameworks for sample size planning require the user to make
 assumptions about the expected effect size. Assuming the true effect is
@@
 
+\bibitem[\citeproctext]{ref-endersSimpleMonteCarlo2023}
+Enders, C. K., Keller, B. T., \& Woller, M. P. (2023). A simple {Monte
+Carlo} method for estimating power in multilevel designs.
+\emph{Psychological Methods}. \url{https://doi.org/10.1037/met0000614}
+
 \bibitem[\citeproctext]{ref-fahrmeirRegressionModelsMethods2021}
@@
 \bibitem[\citeproctext]{ref-greenSIMRPackagePower2016a}
-Green, P., \& MacLeod, C. J. (2016). {\textsc{SIMR}} : An {R} package
+Green, P., \& MacLeod, C. J. (2016a). {\textsc{SIMR}} : An {R} package
 for power analysis of generalized linear mixed models by simulation.
@@
 \url{https://doi.org/10.1111/2041-210X.12504}
+
+\bibitem[\citeproctext]{ref-greenSIMRPackagePower2016}
+Green, P., \& MacLeod, C. J. (2016b). {SIMR}: An {R} package for power
+analysis of generalized linear mixed models by simulation. \emph{Methods
+in Ecology and Evolution}, \emph{7}(4), 493--498.
+\url{https://doi.org/10.1111/2041-210X.12504}
+
+\bibitem[\citeproctext]{ref-hallgrenConductingSimulationStudies2013}
+Hallgren, K. A. (2013). Conducting {Simulation Studies} in the {R
+Programming Environment}. \emph{Tutorials in Quantitative Methods for
+Psychology}, \emph{9}(2), 43--60.
+\url{https://doi.org/10.20982/tqmp.09.2.p043}
+
+\bibitem[\citeproctext]{ref-iddiPowerSampleSize2022}
+Iddi, S., \& C Donohue, M. (2022). Power and {Sample Size} for
+{Longitudinal Models} in {R} -- {The} longpower {Package} and {Shiny
+App}. \emph{The R Journal}, \emph{14}(1), 264--282.
+\url{https://doi.org/10.32614/RJ-2022-022}
 
@@
 
+\bibitem[\citeproctext]{ref-lebeauSimglmSimulateModels2017}
+LeBeau, B. (2017). \emph{Simglm: {Simulate Models Based} on the
+{Generalized Linear Model}} (p. 0.8.9). Comprehensive R Archive Network.
+\url{https://doi.org/10.32614/CRAN.package.simglm}
+
+\bibitem[\citeproctext]{ref-leeUsingTidyversePackage2020}
+Lee, S., Sriutaisuk, S., \& Kim, H. (2020). Using the {Tidyverse
+Package} in {R} for {Simulation Studies} in {SEM}. \emph{Structural
+Equation Modeling: A Multidisciplinary Journal}, \emph{27}(3), 468--482.
+\url{https://doi.org/10.1080/10705511.2019.1644515}
+
 \bibitem[\citeproctext]{ref-littleStatisticalAnalysisMissing2014}
@@
 
+\bibitem[\citeproctext]{ref-magnussonConsequencesIgnoringTherapist2018}
+Magnusson, K., Andersson, G., \& Carlbring, P. (2018). The consequences
+of ignoring therapist effects in trials with longitudinal data: {A}
+simulation study. \emph{Journal of Consulting and Clinical Psychology},
+\emph{86}(9), 711--725. \url{https://doi.org/10.1037/ccp0000333}
+
+\bibitem[\citeproctext]{ref-marcochakyanyuPassLmePower2019}
+Marco Chak Yan YU. (2019). \emph{Pass.lme: {Power} and {Sample Size} for
+{Linear Mixed Effect Models}} (p. 0.9.0). Comprehensive R Archive
+Network. \url{https://doi.org/10.32614/CRAN.package.pass.lme}
+
+\bibitem[\citeproctext]{ref-martinMeasuringIndividualDifferences2011}
+Martin, J. G. A., Nussey, D. H., Wilson, A. J., \& Réale, D. (2011).
+Measuring individual differences in reaction norms in field and
+experimental studies: A power analysis of random regression models.
+\emph{Methods in Ecology and Evolution}, \emph{2}(4), 362--374.
+\url{https://doi.org/10.1111/j.2041-210X.2010.00084.x}
+
+\bibitem[\citeproctext]{ref-matuschekBalancingTypeError2017}
+Matuschek, H., Kliegl, R., Vasishth, S., Baayen, H., \& Bates, D.
+(2017). Balancing {Type I} error and power in linear mixed models.
+\emph{Journal of Memory and Language}, \emph{94}, 305--315.
+\url{https://doi.org/10.1016/j.jml.2017.01.001}
+
 \bibitem[\citeproctext]{ref-maxwellSampleSizePlanning2008}
@@
 
+\bibitem[\citeproctext]{ref-R-base}
+R Core Team. (2024). \emph{R: A language and environment for statistical
+computing}. R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}
+
 \bibitem[\citeproctext]{ref-riesthuisSimulationBasedPowerAnalyses2024}
@@
 in parallel using futures}. \url{https://github.com/DavisVaughan/furrr}
+
+\bibitem[\citeproctext]{ref-watsonGeneralisedLinearMixed2023}
+Watson, S. I. (2023). \emph{Generalised {Linear Mixed Model
+Specification}, {Analysis}, {Fitting}, and {Optimal Design} in {R} with
+the glmmr {Packages}}. arXiv.
+\url{https://doi.org/10.48550/ARXIV.2303.12657}
 
@@
 
+\bibitem[\citeproctext]{ref-zhangPracticalStatisticalPower2018}
+Zhang, Z., \& Yuan, K.-H. (2018). \emph{Practical statistical power
+analysis using {Webpower} and {R}}. ISDSA Press.
+\url{https://doi.org/10.35566/power}
+
 \bibitem[\citeproctext]{ref-zimmerSampleSizePlanning2022}

### 2024-06-28 manuscript.qmd
update mlpwr reference and add footnote on performing cost-efficient design analysis with mlpwr; fix some typos

@@
 Available formula-based and summary-statistics-based software packages for multilevel models often do not include GLMMs or are limited to simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
-A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a].
+A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2023; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a].
 However, many of these tutorials focus on LMMs and the most common study designs [see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
@@
 This difference increases with the variability of the random effects.
-We can use plots like @fig-margdist1 as a useful tool to decide whether the specified standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
+We can use plots like @fig-margdist1 as a useful tool to determine whether the specified standard deviations of the subject and item random intercepts ($\sigma_S$ and $\sigma_I$) are reasonable by comparing the ranges and overlap between conditions to domain knowledge.
 
@@
 The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
-Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts to students, the frequency of whether advice is presented at all, and whether it is correct or incorrect.
+Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts to students, the frequency of whether advice is presented at all, and whether it is correct or incorrect.^[An advanced alternative to our heuristic trade-off between subjects and items would be to explicitly optimize for cost-efficient study designs using the *mlpwr* package [@zimmerSampleSizePlanning2023].
+The package performs simulation based power analysis in a surrogate modeling framework that can evaluate power for a large grid of design combinations more efficiently.
+The framework also requires users to write their own `sim_and_analyse` function and can be considered a great extension on what we cover in our tutorial.]
 
@@
 By incorporating such training into research methods courses and workshops, researchers can improve the quality of their experimental designs and enhance the rigor of their studies.
-The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
+The need to reason about how to simulate plausible data that is in line with the research hypothesis, while not violating domain expertise on how plausible data should look, might also contribute to planning more insightful studies that can answer more precise research questions [@yarkoniGeneralizabilityCrisis2022].
 

### 2024-06-28 manuscript.qmd
fix tbl-rpackages to work on pdf with flextable to avoid bug cannot use table referencing and must hard code the table note; add comment on the design used in case study

@@
 To our knowledge, existing approaches for sample size planning for GLMMs have exclusively focused on planning for power.
-In @tbl-rpackages, we review available software packages that can be used to perform power analysis for multilevel models in `r cite_r()`.
+In Table 1, we review available software packages that can be used to perform power analysis for multilevel models in `r cite_r()`.
 Power analysis methods can be categorized into formula-based, summary-statistics-based and simulation-based methods [@murayamaSummarystatisticsbasedPowerAnalysis2022].
@@
 ```{r}
-#| label: tbl-rpackages
 #| tbl-cap: Overview of software packages for power analysis for (generalized) linear mixed models in R.
@@
 #| echo: false
+#| message: false
+# label: tbl-rpackages # There is currently a bug in apaquarto that prevents referencing for multipage tables (https://github.com/wjschne/apaquarto/issues/71)
 
@@
 
-knitr::kable(packagetable, col.names = c("Package name", "Summary", "GUI available", "GLMM support", "Design limitations", "Reference"), escape = FALSE)
-```
+flextable::flextable(packagetable) |>
+  flextable::width(j = "Package name", width = 3, unit = "cm") |>
+  flextable::width(j = "Summary", width = 4, unit = "cm") |>
+  flextable::width(j = "GUI available", width = 1.5, unit = "cm") |>
+  flextable::width(j = "GLMM support", width = 1.5, unit = "cm") |>
+  flextable::width(j = "Design limitations", width = 3, unit = "cm") |>
+  flextable::width(j = "Reference", width = 2, unit = "cm") |>
+  flextable::fontsize(size = 10, part = "all") |>
+  flextable::theme_apa() |>
+  ftExtra::colformat_md()
+```
+:::{.NoIndent}
+*Note.* GUI available = whether a graphical user interface is provided (usually a Shiny app that is available online or can be downloaded to run locally), GLMM support = whether power analysis can be performed for generalized linear mixed model (otherwise only linear mixed models are supported), Design limitations = whether only a limited selection of study designs can be specified (otherwise arbitrary designs within the model class are supported).
+:::
 
@@
 
-Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in @tbl-rpackages.
+Performing tailored simulation-based sample size planning is more complicated and time-consuming than using the existing software tools outlined in Table 1.
 Circumstances under which tailored simulation-based sample size planning is necessary include 1) complex study designs, 2) complex statistical hypotheses, 3) planning for precision, and 4) lack of prior studies or pilot data.
@@
 However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
-None of the software packages for sample size planning in @tbl-rpackages can handle combined hypotheses and only some can handle directed hypotheses.
+None of the software packages for sample size planning in Table 1 can handle combined hypotheses and only some can handle directed hypotheses.
 In contrast, our case study will demonstrate how we can test directed combined hypotheses with tailored simulation-based sample size planning.
@@
 Third, all available software packages for sample size planning with GLMMs are based on power analysis and do not support planning for precision.
-Thus, researchers that want to apply an estimation strategy instead of testing statistical hypotheses [@cummingNewStatisticsWhy2014], currently cannot use the software packages outlined in @tbl-rpackages.
+Thus, researchers that want to apply an estimation strategy instead of testing statistical hypotheses [@cummingNewStatisticsWhy2014], currently cannot use the software packages outlined in Table 1.
 However, tailored simulation-based sample size planning can easily handle the planning for precision approach [@maxwellSampleSizePlanning2008].
@@
 We will introduce each step in a theoretical section, followed by the practical application based on a case study.
+To keep the tutorial manageable we focus on a binomial GLMM with two random intercepts and discrete predictor variables.
 All code in this manuscript and simulation results are available on the Open Science Framework (<https://osf.io/dhwf4/>) and on our tutorial website (<https://timo-ko.github.io/glmm_simulation_tutorial/>).
@@
 
-@tbl-probtable shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
+Table 2 shows our set of assumptions concerning the probability that an average subject solves an average item for each experimental condition, as well as the corresponding equations implied by the model.
 The table can be used to compute the implied values for the $\beta$ parameters, starting with the first equation and reinserting the computed $\beta$ values in all following equations (`b_0` stands for the intercept $\beta_0$, `b_e` for the slope $\beta_e$, and so on):
@@
 However, the magnitude of these effects should be substantially greater for students, given their stronger reliance on advice compared to experts.
-We further validated the plausibility of our estimated gains and losses with the collaborating radiologists until we settled on the probabilities shown in @tbl-probtable.
+We further validated the plausibility of our estimated gains and losses with the collaborating radiologists until we settled on the probabilities shown in Table 2.
 

### 2024-06-29 manuscript.qmd
switch to R 4.3.3 to be able to install flextable in renv project; update packages and lock file

@@
 )
+names(packagetable) <- c("Package name", "Summary", "GUI available", "GLMM support", "Design limitations", "Reference")
 

### 2024-06-29 manuscript.qmd
add note on equal author contribution; add johnson citation; add many labs citation and move comment to footnote; update simglm citation; update bibfile

--- 2024-06-28 manuscript.qmd
@@
     gratitude: ~
-    authorship-agreements: ~
+    authorship-agreements: Florian Pargent and Timo Koch contributed equally to this manuscript.
 language:
@@
 The most prominent heuristic is to target a power of $1- \beta = 0.8$ in combination with a type I error rate of $\alpha = 0.05$ [@lakensSampleSizeJustification2022].
-However, depending on the research goals or resource constraints, there are often good reasons to move away from this standard [@benjaminRedefineStatisticalSignificance2017; @lakensJustifyYourAlpha2018].
-For example, the *Social Sciences Replication Project* targeted a power of 0.9 to safeguard against biased effect sizes in the original studies [@camererEvaluatingReplicabilitySocial2018].
+However, depending on the research goals or resource constraints, there are often good reasons to move away from this standard [@benjaminRedefineStatisticalSignificance2017; @lakensJustifyYourAlpha2018].^[For example, the *Social Sciences Replication Project* targeted a power of 0.90 to safeguard against biased effect sizes in the original studies [@camererEvaluatingReplicabilitySocial2018], and *Many Labs 5* even targeted a power of 0.95 [@ebersoleManyLabsTesting2020].]
 
@@
 Available formula-based and summary-statistics-based software packages for multilevel models often do not include GLMMs or are limited to simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
-A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2023; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a].
+A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2023; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a; @johnsonPowerAnalysisGeneralized2015].
 However, many of these tutorials focus on LMMs and the most common study designs [see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
@@
   col5 = c("none", "limited set of study designs with two levels", "none", "limited to two levels", "none", "none", "limited set of study designs with two levels", "limited set of study designs with two or three levels", "limited to two or three levels", "none", "limited set of study designs with two levels", "limited set of study designs with two levels"),
-  col6 = c("@watsonGeneralisedLinearMixed2023", "@iddiPowerSampleSize2022", "@kumleEstimatingPowerGeneralized2021", "@endersSimpleMonteCarlo2023", "@martinMeasuringIndividualDifferences2011", "@marcochakyanyuPassLmePower2019", "@lafitSelectionNumberParticipants2021", "@magnussonConsequencesIgnoringTherapist2018", "@lebeauSimglmSimulateModels2017", "@greenSIMRPackagePower2016", "@murayamaSummarystatisticsbasedPowerAnalysis2022", "@zhangPracticalStatisticalPower2018")
+  col6 = c("@watsonGeneralisedLinearMixed2023", "@iddiPowerSampleSize2022", "@kumleEstimatingPowerGeneralized2021", "@endersSimpleMonteCarlo2023", "@martinMeasuringIndividualDifferences2011", "@marcochakyanyuPassLmePower2019", "@lafitSelectionNumberParticipants2021", "@magnussonConsequencesIgnoringTherapist2018", "@lebeauPowerAnalysisSimulation2019", "@greenSIMRPackagePower2016", "@murayamaSummarystatisticsbasedPowerAnalysis2022", "@zhangPracticalStatisticalPower2018")
 )
+names(packagetable) <- c("Package name", "Summary", "GUI available", "GLMM support", "Design limitations", "Reference")
 

### 2024-06-29 manuscript.qmd
update date and preprint version

@@
 shorttitle: "Simulation-Based Sample Size Planning for GLMMs"
-date: 2024-06-12
+date: 2024-07-01
 date-format: iso
@@
   disclosures:
-    data-sharing: This is version 2.5 of our preprint published at <https://doi.org/10.31234/osf.io/rpjem>.
+    data-sharing: This is version 3 of our preprint published at <https://doi.org/10.31234/osf.io/rpjem>.
 All materials (reproducible manuscript, including R code and all simulation results) are available in the project's repository on the Open Science Framework (OSF) at <https://osf.io/dhwf4/>.

### 2024-07-01 manuscript.qmd
switch back to %>% pipe; update results text to new full_simulation; fix some typos

@@
 ```{r}
-#| tbl-cap: Overview of software packages for power analysis for (generalized) linear mixed models in R.
+#| tbl-cap: Table 1: Overview of software packages for power analysis for (generalized) linear mixed models in R.
 #| apa-note: GUI available = whether a graphical user interface is provided (usually a Shiny app that is available online or can be downloaded to run locally), GLMM support = whether power analysis can be performed for generalized linear mixed model (otherwise only linear mixed models are supported), Design limitations = whether only a limited selection of study designs can be specified (otherwise arbitrary designs within the model class are supported).
@@
 
-flextable::flextable(packagetable) |>
-  flextable::width(j = "Package name", width = 3, unit = "cm") |>
-  flextable::width(j = "Summary", width = 4, unit = "cm") |>
-  flextable::width(j = "GUI available", width = 1.5, unit = "cm") |>
-  flextable::width(j = "GLMM support", width = 1.5, unit = "cm") |>
-  flextable::width(j = "Design limitations", width = 3, unit = "cm") |>
-  flextable::width(j = "Reference", width = 2, unit = "cm") |>
-  flextable::fontsize(size = 10, part = "all") |>
-  flextable::theme_apa() |>
+flextable::flextable(packagetable) %>%
+  flextable::width(j = "Package name", width = 3, unit = "cm") %>%
+  flextable::width(j = "Summary", width = 4, unit = "cm") %>%
+  flextable::width(j = "GUI available", width = 1.5, unit = "cm") %>%
+  flextable::width(j = "GLMM support", width = 1.5, unit = "cm") %>%
+  flextable::width(j = "Design limitations", width = 3, unit = "cm") %>%
+  flextable::width(j = "Reference", width = 2, unit = "cm") %>%
+  flextable::fontsize(size = 10, part = "all") %>%
+  flextable::theme_apa() %>%
   ftExtra::colformat_md()
@@
 We will introduce each step in a theoretical section, followed by the practical application based on a case study.
-To keep the tutorial manageable we focus on a binomial GLMM with two random intercepts and discrete predictor variables.
+To keep the tutorial manageable, we focus on a binomial GLMM with two random intercepts and discrete predictor variables.
 All code in this manuscript and simulation results are available on the Open Science Framework (<https://osf.io/dhwf4/>) and on our tutorial website (<https://timo-ko.github.io/glmm_simulation_tutorial/>).
@@
 This contrast is quantified for a hypothetical *typical* expert and a *typical* scan, where *typical* is usually defined as an average score on all attributes of the expert or scan.^[Note that a different estimand would be the so-called average treatment effect (ATE).
-For the ATE, the probability contrast is defined for each combination of expert and scan, and then these contrasts are averaged across all experts and scans from the target population [@lundbergWhatYourEstimand2021].].
+For the ATE, the probability contrast is defined for each combination of expert and scan, and then these contrasts are averaged across all experts and scans from the target population [@lundbergWhatYourEstimand2021].]
 To complete our definition of the estimand, we have to define our target population that consists of persons, stimuli, and AI advice: With respect to persons, we are only interested in experts (i.e., radiologists) and non-experts (i.e., medical students).
@@
   # simulate design
-  dat <- add_random(subject = n_subjects, item = n_items) |>
-    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) |>
-    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) |>
+  dat <- add_random(subject = n_subjects, item = n_items) %>%
+    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
+    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
     mutate(advice_correct = if_else(advice_present == 1L, 
-                                    rbinom(n(), 1L, prob = 0.8), 0L)) |>
+                                    rbinom(n(), 1L, prob = 0.8), 0L)) %>%
     # add random effects
-    add_ranef("subject", u0s = sd_u0s) |>
-    add_ranef("item", u0i = sd_u0i) |>
+    add_ranef("subject", u0s = sd_u0s) %>%
+    add_ranef("item", u0i = sd_u0i) %>%
     # compute dependent variable
@@
       b_e * expert + b_a * advice_present + b_c * advice_correct +
-        b_ea * expert * advice_present + b_ec * expert * advice_correct) |>
-    mutate(y_prob = plogis(linpred)) |>
+      b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
+    mutate(y_prob = plogis(linpred)) %>%
     mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
@@
 ```{r}
-#| label: tbl-probtable
-#| tbl-cap: Assumed probabilities that an average subject solves an average item in each experimental condition.
+#| tbl-cap: Table 2: Assumed probabilities that an average subject solves an average item in each experimental condition.
 #| apa-note: Implied equations are derived based on the model equations and setting all random intercept terms to 0.
@@
   sd_u0s = 0.5, sd_u0i = 0.5)
-dat |> 
+dat %>% 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
   mutate(condition = fct_recode(condition,
@@
  "student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
- "student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
-  group_by(condition) |>
+ "student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+  group_by(condition) %>%
   summarize(relative_frequency = sum(y_bin) / n())
@@
 library(ggdist)
-dat |> 
+dat %>% 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
   mutate(condition = fct_recode(condition,
@@
 "student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
-"student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
+"student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
   ggplot(aes(x = y_prob, y = condition)) +
@@
 dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0i = 0.01)
-dat |> 
+dat %>% 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
   mutate(condition = fct_recode(condition,
@@
 dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.01)
-dat |> 
+dat %>% 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
   mutate(condition = fct_recode(condition,
@@
 dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 3, sd_u0i = 3)
-dat |> 
+dat %>% 
   mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) |>
+    factor(expert), factor(advice_present), factor(advice_correct))) %>%
   mutate(condition = fct_recode(condition,
@@
 "student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
-    "student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) |> 
+"student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
   ggplot(aes(x = y_prob, y = condition)) +
@@
   type = "response", re.form = NA)
-print(preds, style = "tinytable") |> theme_tt(theme = "resize")
+print(preds, style = "tinytable") %>% theme_tt(theme = "resize")
 ```
@@
 #| size: footnotesize
-contrasts <- preds |> 
+contrasts <- preds %>% 
   hypotheses(hypothesis = c(
@@
     equivalence = c(0, 0))
-print(contrasts, style = "tinytable") |> theme_tt(theme = "resize")
+print(contrasts, style = "tinytable") %>% theme_tt(theme = "resize")
 ```
@@
     expert = 0:1)
-  predictions(model, newdata = contr_df, type = "response", re.form = NA) |>
-    hypotheses(hypothesis = contrasts, equivalence = c(0, 0)) |>
+  predictions(model, newdata = contr_df, type = "response", re.form = NA) %>%
+    hypotheses(hypothesis = contrasts, equivalence = c(0, 0)) %>%
     data.frame()
@@
 We chose 300 repetitions for the data simulation at hand to strike a balance between achieving a robust estimate and remaining computationally feasible.
-With the current settings, this simulation takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
+With the current settings, this simulation takes several hours on a MacBook Pro from 2020 with M1 chip and 16 GB working memory.
 If you want to quickly experiment with the code yourself, a setting with `workers = 4` and `rep = 5` should finish in less than 5 minutes, even on smaller machines.
@@
 sim_result <- crossing(
-  rep = 1:300,
+  rep = 1:500,
   n_subjects = c(100, 150, 200, 250),
   n_items = c(10, 30, 50, 70)
-) |>
+) %>%
   mutate(res = future_pmap(., sim_and_analyse, 
-    .options = furrr_options(seed = TRUE))) |>
+    .options = furrr_options(seed = TRUE))) %>%
   unnest(col = res)
@@
 alpha <- 0.05
-power <- sim_result |>
+power <- sim_result %>%
   pivot_wider(names_from = term, names_sep = "_", 
-    values_from = estimate:p.value.equiv) |>
-  group_by(n_subjects, n_items) |> 
+    values_from = estimate:p.value.equiv) %>%
+  group_by(n_subjects, n_items) %>% 
   summarise(
@@
     .groups = "drop")
-power |>
-  mutate(across(c(n_subjects, n_items), factor)) |>
+power %>%
+  mutate(across(c(n_subjects, n_items), factor)) %>%
   ggplot(aes(n_subjects, n_items, fill = power)) +
@@
 Although it is recommended to justify the power [@lakensSampleSizeJustification2022], we adopt the common heuristic to target a power of 0.8 to detect an effect of the expected size implied by our data simulation.
-This could be achieved by collecting data from 200 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice).
-If collecting data from 200 subjects is not feasible, an alternative would be to recruit 150 subjects but increase the length of the experiment to over 70 items.
+This could be achieved by collecting data from 250 subjects (about 25% of which will be experts), each completing the same 50 items (with advice present in about 67% of cases, which is correct in about 80% of cases with present advice).
+If collecting data from 250 subjects is not feasible, an alternative would be to recruit 200 subjects but increase the length of the experiment to 70 items.
 However, 70 items might take too long to complete for the radiologists participating in the study, who have a busy schedule.
-The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (\> 250, to be determined by additional rounds of power analysis).
+The simulation suggests that it might also be possible to plan a shorter experiment with only 30 items if it is feasible to recruit an even higher number of subjects (to be determined by additional rounds of power analysis).
 Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts to students, the frequency of whether advice is presented at all, and whether it is correct or incorrect.^[An advanced alternative to our heuristic trade-off between subjects and items would be to explicitly optimize for cost-efficient study designs using the *mlpwr* package [@zimmerSampleSizePlanning2023].
@@
 #| echo: true
-precision <- sim_result |>
+precision <- sim_result %>%
   pivot_wider(names_from = term, names_sep = "_", 
-    values_from = estimate:p.value.equiv) |>
-  group_by(n_subjects, n_items) |>
-  mutate(width = `conf.high_b8=b2` - `conf.low_b8=b2`) |>
+    values_from = estimate:p.value.equiv) %>%
+  group_by(n_subjects, n_items) %>%
+  mutate(width = `conf.high_b8=b2` - `conf.low_b8=b2`) %>%
   summarise(precision = mean(width),
@@
     .groups = "drop")
-precision |>
-  mutate(across(c(n_subjects, n_items), factor)) |>
+precision %>%
+  mutate(across(c(n_subjects, n_items), factor)) %>%
   ggplot(aes(n_subjects, n_items, fill = precision)) +
@@
 Applying our simple heuristic of targeting an expected CI width that is smaller than 0.1, we see the same trade-off between the number of subjects and the number of items as with planning for power.
-We could either choose 100 subjects and 30 items or 200 subjects and 10 items.
+We could either choose 100 subjects and 30 items or 250 subjects and 10 items.
 Note that our simple heuristic for determining sample size in the planning for precision scenario was quite liberal.

### 2024-07-01 manuscript.qmd
fix table headers

@@
 ```{r}
-#| tbl-cap: Table 1: Overview of software packages for power analysis for (generalized) linear mixed models in R.
+#| tbl-cap: Table 1. Overview of software packages for power analysis for (generalized) linear mixed models in R.
 #| apa-note: GUI available = whether a graphical user interface is provided (usually a Shiny app that is available online or can be downloaded to run locally), GLMM support = whether power analysis can be performed for generalized linear mixed model (otherwise only linear mixed models are supported), Design limitations = whether only a limited selection of study designs can be specified (otherwise arbitrary designs within the model class are supported).
@@
 ```{r}
-#| tbl-cap: Table 2: Assumed probabilities that an average subject solves an average item in each experimental condition.
+#| tbl-cap: Table 2. Assumed probabilities that an average subject solves an average item in each experimental condition.
 #| apa-note: Implied equations are derived based on the model equations and setting all random intercept terms to 0.

### 2024-07-01 _manuscript/_tex/manuscript.tex
update _freeze and _manuscript

@@
   longtable,
+  a4paper,
   nolmodern,
@@
 
-\date{2024-06-12}
+\date{2024-07-01}
 
@@
 
-\par{ This is version 2.5 of our preprint published at
+\par{ This is version 3 of our preprint published at
 \url{https://doi.org/10.31234/osf.io/rpjem}. All materials (reproducible
@@
 authorship or the publication of this article. This research was funded
-by a grant from the Volkswagen Foundation (Grant
-No.~98525).   Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Florian
+by a grant from the Volkswagen Foundation (Grant No.~98525).  Florian
+Pargent and Timo Koch contributed equally to this
+manuscript. Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Florian
 Pargent:   formal
@@
 
+\usepackage{fontspec}
+\usepackage{multirow}
+\usepackage{multicol}
+\usepackage{colortbl}
+\usepackage{hhline}
+\newlength\Oldarrayrulewidth
+\newlength\Oldtabcolsep
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{hyperref}
 \usepackage{float}
+\usepackage{wrapfig}
 \usepackage{tabularray}
@@
 al., 2017}; \citeproc{ref-lakensJustifyYourAlpha2018}{Lakens, Adolfi, et
-al., 2018}). For example, the \emph{Social Sciences Replication Project}
-targeted a power of 0.9 to safeguard against biased effect sizes in the
-original studies
-(\citeproc{ref-camererEvaluatingReplicabilitySocial2018}{Camerer et al.,
-2018}).
+al., 2018}).\footnote{For example, the \emph{Social Sciences Replication
+  Project} targeted a power of 0.90 to safeguard against biased effect
+  sizes in the original studies
+  (\citeproc{ref-camererEvaluatingReplicabilitySocial2018}{Camerer et
+  al., 2018}), and \emph{Many Labs 5} even targeted a power of 0.95
+  (\citeproc{ref-ebersoleManyLabsTesting2020}{Ebersole et al., 2020}).}
 
@@
 available software packages that can be used to perform power analysis
-for multilevel models in R {[}Version 4.4.0; R Core Team
+for multilevel models in R {[}Version 4.3.3; R Core Team
 (\citeproc{ref-R-base}{2024}){]}. Power analysis methods can be
@@
 Barr, 2021}; \citeproc{ref-greenSIMRPackagePower2016a}{Green \& MacLeod,
-2016}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al., 2015};
-\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021};
-\citeproc{ref-lafitSelectionNumberParticipants2021}{Lafit et al., 2021};
-\citeproc{ref-zimmerSampleSizePlanning2022}{Zimmer et al., 2022}).
-However, many of these tutorials focus on LMMs and the most common study
-designs (see \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et
-al., 2021} for a tutorial that also covers more advanced settings). This
-narrow focus provides limited guidance for researchers using more
-complex study designs, especially when little prior knowledge about
-plausible effect sizes is available (see the discussion in
+2016a}; \citeproc{ref-johnsonPowerAnalysisGeneralized2015}{Johnson et
+al., 2015}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al.,
+2015}; \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al.,
+2021}; \citeproc{ref-lafitSelectionNumberParticipants2021}{Lafit et al.,
+2021}; \citeproc{ref-zimmerSampleSizePlanning2023}{Zimmer et al.,
+2023}). However, many of these tutorials focus on LMMs and the most
+common study designs (see
+\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}
+for a tutorial that also covers more advanced settings). This narrow
+focus provides limited guidance for researchers using more complex study
+designs, especially when little prior knowledge about plausible effect
+sizes is available (see the discussion in
 \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}).
@@
 
-INSERT TABLE 1 HERE!
+\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}
+
+\global\setlength{\Oldtabcolsep}{\tabcolsep}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\begin{longtable}[c]{|p{1.18in}|p{1.57in}|p{0.59in}|p{0.59in}|p{1.18in}|p{0.79in}}
+\caption{Table 1. Overview of software packages for power analysis for
+(generalized) linear mixed models in R.}\tabularnewline
+
+\ascline{0.75pt}{000000}{1-6}
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Package\ name}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Summary}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{GUI\ available}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{GLMM\ support}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Design\ limitations}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Reference}}}} \\
+
+\ascline{0.75pt}{000000}{1-6}\endfirsthead
+
+\ascline{0.75pt}{000000}{1-6}
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Package\ name}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Summary}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{GUI\ available}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{GLMM\ support}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Design\ limitations}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Reference}}}} \\
+
+\ascline{0.75pt}{000000}{1-6}\endhead
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{glmmrBase}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{and}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{non-linear}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{fixed}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{effects}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{and}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{flexible}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{covariance}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{functions}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Watson}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2023)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://atrihub.shinyapps.io/power/}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Iddi}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\&}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{C}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Donohue}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2022)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mixedpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{simr}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Kumle}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2021)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mlmpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{missing}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mechanisms}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{to}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Enders}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2023)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{pamm}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{lme4}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Martin}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2011)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{pass.lme}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{documentation}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Marco}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Chak}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Yan}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{YU}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2019)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{PowerAnalysisIL}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://github.com/ginettelafit/PowerAnalysisIL?tab=readme-ov-file\#shiny-app-and-r-package-to-perform-power-analysis-to-select-the-number-of-participants-in-intensive-longitudinal-studies}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Lafit}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2021)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{powerlmm}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{missing}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mechanisms;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://github.com/rpsychologist/powerlmm?tab=readme-ov-file\#launch-interactive-web-application}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{or}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{three}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Magnusson}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2018)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{simglm}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{missing}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mechanisms}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://simglm.brandonlebeau.org/reference/run\_shiny.html}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{to}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{or}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{three}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{LeBeau}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2019)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{simr}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{lme4}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Green}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\&}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{MacLeod}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2016)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{summary\_statistics}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based\_power}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Summary-statistics-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{pwr}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://koumurayama.shinyapps.io/summary\_statistics\_based\_power/}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Murayama}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2022)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{WebPower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longpower}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://webpower.psychstat.org/wiki/start}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Zhang}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\&}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Yuan}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2018)}}}} \\
+
+\ascline{0.75pt}{000000}{1-6}
+
+\end{longtable}
+
+\arrayrulecolor[HTML]{000000}
+
+\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}
+
+\global\setlength{\tabcolsep}{\Oldtabcolsep}
+
+\renewcommand*{\arraystretch}{1}
+
+\noindent \emph{Note.} GUI available = whether a graphical user
+interface is provided (usually a Shiny app that is available online or
+can be downloaded to run locally), GLMM support = whether power analysis
+can be performed for generalized linear mixed model (otherwise only
+linear mixed models are supported), Design limitations = whether only a
+limited selection of study designs can be specified (otherwise arbitrary
+designs within the model class are supported).
 
@@
 complicated and time-consuming than using the existing software tools
-outlined in TABLE 1. Circumstances under which tailored simulation-based
+outlined in Table 1. Circumstances under which tailored simulation-based
 sample size planning is necessary include 1) complex study designs, 2)
@@
 (\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014}), currently
-cannot use the software packages outlined in TABLE 1. However, tailored
+cannot use the software packages outlined in Table 1. However, tailored
 simulation-based sample size planning can easily handle the planning for
@@
 decisions. We will introduce each step in a theoretical section,
-followed by the practical application based on a case study. All code in
-this manuscript and simulation results are available on the Open Science
-Framework (\url{https://osf.io/dhwf4/}) and on our tutorial website
+followed by the practical application based on a case study. To keep the
+tutorial manageable, we focus on a binomial GLMM with two random
+intercepts and discrete predictor variables. All code in this manuscript
+and simulation results are available on the Open Science Framework
+(\url{https://osf.io/dhwf4/}) and on our tutorial website
 (\url{https://timo-ko.github.io/glmm_simulation_tutorial/}).
@@
   scans from the target population
-  (\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}).}.
+  (\citeproc{ref-lundbergWhatYourEstimand2021}{Lundberg et al., 2021}).}
 To complete our definition of the estimand, we have to define our target
@@
   \CommentTok{\# simulate design}
-\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{add\_random}\NormalTok{(}\AttributeTok{subject =}\NormalTok{ n\_subjects, }\AttributeTok{item =}\NormalTok{ n\_items) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{add\_between}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{expert =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{.prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.25}\NormalTok{, }\FloatTok{0.75}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_present =} \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{, }\AttributeTok{prob =} \DecValTok{2}\SpecialCharTok{/}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
+\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{add\_random}\NormalTok{(}\AttributeTok{subject =}\NormalTok{ n\_subjects, }\AttributeTok{item =}\NormalTok{ n\_items) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{add\_between}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{expert =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{.prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.25}\NormalTok{, }\FloatTok{0.75}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_present =} \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{, }\AttributeTok{prob =} \DecValTok{2}\SpecialCharTok{/}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
     \FunctionTok{mutate}\NormalTok{(}\AttributeTok{advice\_correct =} \FunctionTok{if\_else}\NormalTok{(advice\_present }\SpecialCharTok{==} \DecValTok{1}\NormalTok{L, }
-                                    \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{L, }\AttributeTok{prob =} \FloatTok{0.8}\NormalTok{), }\DecValTok{0}\NormalTok{L)) }\SpecialCharTok{|\textgreater{}}
+                                    \FunctionTok{rbinom}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{1}\NormalTok{L, }\AttributeTok{prob =} \FloatTok{0.8}\NormalTok{), }\DecValTok{0}\NormalTok{L)) }\SpecialCharTok{\%\textgreater{}\%}
     \CommentTok{\# add random effects}
-    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{u0s =}\NormalTok{ sd\_u0s) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"item"}\NormalTok{, }\AttributeTok{u0i =}\NormalTok{ sd\_u0i) }\SpecialCharTok{|\textgreater{}}
+    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"subject"}\NormalTok{, }\AttributeTok{u0s =}\NormalTok{ sd\_u0s) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{add\_ranef}\NormalTok{(}\StringTok{"item"}\NormalTok{, }\AttributeTok{u0i =}\NormalTok{ sd\_u0i) }\SpecialCharTok{\%\textgreater{}\%}
     \CommentTok{\# compute dependent variable}
@@
 \NormalTok{      b\_e }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{+}\NormalTok{ b\_a }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_c }\SpecialCharTok{*}\NormalTok{ advice\_correct }\SpecialCharTok{+}
-\NormalTok{        b\_ea }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_ec }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_correct) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_prob =} \FunctionTok{plogis}\NormalTok{(linpred)) }\SpecialCharTok{|\textgreater{}}
+\NormalTok{      b\_ea }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_present }\SpecialCharTok{+}\NormalTok{ b\_ec }\SpecialCharTok{*}\NormalTok{ expert }\SpecialCharTok{*}\NormalTok{ advice\_correct) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_prob =} \FunctionTok{plogis}\NormalTok{(linpred)) }\SpecialCharTok{\%\textgreater{}\%}
     \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_bin =} \FunctionTok{rbinom}\NormalTok{(}\AttributeTok{n =} \FunctionTok{n}\NormalTok{(), }\AttributeTok{size =} \DecValTok{1}\NormalTok{, }\AttributeTok{prob =}\NormalTok{ y\_prob))}
@@
 
-\begin{table}
-
-{\caption{{Assumed probabilities that an average subject solves an
-average item in each experimental condition.}{\label{tbl-probtable}}}
-\vspace{-20pt}}
-
 \begin{longtable}[]{@{}
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5063}}@{}}
+\caption{Table 2. Assumed probabilities that an average subject solves
+an average item in each experimental condition.}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Experimental condition
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(P(Y=1 \mid \mathbf{X=x}, \mathbf{U} = \mathbf{0})\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Implied equation
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
 \toprule\noalign{}
@@
 
-{\vspace{-20pt}
-\noindent \emph{Note.} Implied equations are derived based on the model equations and setting all random intercept terms to 0.}
-
-\end{table}
-
-Table~\ref{tbl-probtable} shows our set of assumptions concerning the
-probability that an average subject solves an average item for each
-experimental condition, as well as the corresponding equations implied
-by the model. The table can be used to compute the implied values for
-the \(\beta\) parameters, starting with the first equation and
-reinserting the computed \(\beta\) values in all following equations
-(\texttt{b\_0} stands for the intercept \(\beta_0\), \texttt{b\_e} for
-the slope \(\beta_e\), and so on):
+Table 2 shows our set of assumptions concerning the probability that an
+average subject solves an average item for each experimental condition,
+as well as the corresponding equations implied by the model. The table
+can be used to compute the implied values for the \(\beta\) parameters,
+starting with the first equation and reinserting the computed \(\beta\)
+values in all following equations (\texttt{b\_0} stands for the
+intercept \(\beta_0\), \texttt{b\_e} for the slope \(\beta_e\), and so
+on):
 
@@
   \AttributeTok{sd\_u0s =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{sd\_u0i =} \FloatTok{0.5}\NormalTok{)}
-\NormalTok{dat }\SpecialCharTok{|\textgreater{}} 
+\NormalTok{dat }\SpecialCharTok{\%\textgreater{}\%} 
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
-    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{|\textgreater{}}
+    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
@@
  \StringTok{"student, incorrect advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, incorrect advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
- \StringTok{"student, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{(condition) }\SpecialCharTok{|\textgreater{}}
+ \StringTok{"student, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(condition) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{summarize}\NormalTok{(}\AttributeTok{relative\_frequency =} \FunctionTok{sum}\NormalTok{(y\_bin) }\SpecialCharTok{/} \FunctionTok{n}\NormalTok{())}
@@
 \FunctionTok{library}\NormalTok{(ggdist)}
-\NormalTok{dat }\SpecialCharTok{|\textgreater{}} 
+\NormalTok{dat }\SpecialCharTok{\%\textgreater{}\%} 
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_cross}\NormalTok{(}
-    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{|\textgreater{}}
+    \FunctionTok{factor}\NormalTok{(expert), }\FunctionTok{factor}\NormalTok{(advice\_present), }\FunctionTok{factor}\NormalTok{(advice\_correct))) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{fct\_recode}\NormalTok{(condition,}
@@
 \StringTok{"student, incorrect advice"} \OtherTok{=} \StringTok{"0:1:0"}\NormalTok{, }\StringTok{"expert, incorrect advice"} \OtherTok{=} \StringTok{"1:1:0"}\NormalTok{,}
-\StringTok{"student, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
+\StringTok{"student, correct advice"} \OtherTok{=} \StringTok{"0:1:1"}\NormalTok{, }\StringTok{"expert, correct advice"} \OtherTok{=} \StringTok{"1:1:1"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ y\_prob, }\AttributeTok{y =}\NormalTok{ condition)) }\SpecialCharTok{+}
@@
 effects. We can use plots like Figure~\ref{fig-margdist1} as a useful
-tool to decide whether the specified standard deviations of the subject
-and item random intercepts (\(\sigma_S\) and \(\sigma_I\)) are
+tool to determine whether the specified standard deviations of the
+subject and item random intercepts (\(\sigma_S\) and \(\sigma_I\)) are
 reasonable by comparing the ranges and overlap between conditions to
@@
 with the collaborating radiologists until we settled on the
-probabilities shown in Table~\ref{tbl-probtable}.
+probabilities shown in Table 2.
 
@@
 (Intercept)             1.0339     0.1103   9.374  < 2e-16 ***
-expert                  1.1849     0.2096   5.654 1.56e-08 ***
+expert                  1.1849     0.2096   5.654 1.57e-08 ***
 advice_present         -1.3436     0.1206 -11.143  < 2e-16 ***
@@
 expert:advice_present   1.0589     0.2940   3.601 0.000317 ***
-expert:advice_correct  -1.8104     0.2915  -6.210 5.29e-10 ***
+expert:advice_correct  -1.8104     0.2915  -6.210 5.28e-10 ***
 ---
@@
   \AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{)}
-\FunctionTok{print}\NormalTok{(preds, }\AttributeTok{style =} \StringTok{"tinytable"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{theme\_tt}\NormalTok{(}\AttributeTok{theme =} \StringTok{"resize"}\NormalTok{)}
+\FunctionTok{print}\NormalTok{(preds, }\AttributeTok{style =} \StringTok{"tinytable"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{theme\_tt}\NormalTok{(}\AttributeTok{theme =} \StringTok{"resize"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{contrasts }\OtherTok{\textless{}{-}}\NormalTok{ preds }\SpecialCharTok{|\textgreater{}} 
+\NormalTok{contrasts }\OtherTok{\textless{}{-}}\NormalTok{ preds }\SpecialCharTok{\%\textgreater{}\%} 
   \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =} \FunctionTok{c}\NormalTok{(}
@@
     \AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{))}
-\FunctionTok{print}\NormalTok{(contrasts, }\AttributeTok{style =} \StringTok{"tinytable"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{theme\_tt}\NormalTok{(}\AttributeTok{theme =} \StringTok{"resize"}\NormalTok{)}
+\FunctionTok{print}\NormalTok{(contrasts, }\AttributeTok{style =} \StringTok{"tinytable"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{theme\_tt}\NormalTok{(}\AttributeTok{theme =} \StringTok{"resize"}\NormalTok{)}
 \end{Highlighting}
@@
     \AttributeTok{expert =} \DecValTok{0}\SpecialCharTok{:}\DecValTok{1}\NormalTok{)}
-  \FunctionTok{predictions}\NormalTok{(model, }\AttributeTok{newdata =}\NormalTok{ contr\_df, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =}\NormalTok{ contrasts, }\AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{predictions}\NormalTok{(model, }\AttributeTok{newdata =}\NormalTok{ contr\_df, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{, }\AttributeTok{re.form =} \ConstantTok{NA}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{hypotheses}\NormalTok{(}\AttributeTok{hypothesis =}\NormalTok{ contrasts, }\AttributeTok{equivalence =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
     \FunctionTok{data.frame}\NormalTok{()}
@@
 computationally feasible. With the current settings, this simulation
-takes about one hour on a MacBook Pro from 2020 with M1 chip and 16 GB
+takes several hours on a MacBook Pro from 2020 with M1 chip and 16 GB
 working memory. If you want to quickly experiment with the code
@@
 \NormalTok{sim\_result }\OtherTok{\textless{}{-}} \FunctionTok{crossing}\NormalTok{(}
-  \AttributeTok{rep =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{300}\NormalTok{,}
+  \AttributeTok{rep =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{500}\NormalTok{,}
   \AttributeTok{n\_subjects =} \FunctionTok{c}\NormalTok{(}\DecValTok{100}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{250}\NormalTok{),}
   \AttributeTok{n\_items =} \FunctionTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{50}\NormalTok{, }\DecValTok{70}\NormalTok{)}
-\NormalTok{) }\SpecialCharTok{|\textgreater{}}
+\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{res =} \FunctionTok{future\_pmap}\NormalTok{(., sim\_and\_analyse, }
-    \AttributeTok{.options =} \FunctionTok{furrr\_options}\NormalTok{(}\AttributeTok{seed =} \ConstantTok{TRUE}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
+    \AttributeTok{.options =} \FunctionTok{furrr\_options}\NormalTok{(}\AttributeTok{seed =} \ConstantTok{TRUE}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{unnest}\NormalTok{(}\AttributeTok{col =}\NormalTok{ res)}
@@
 \NormalTok{alpha }\OtherTok{\textless{}{-}} \FloatTok{0.05}
-\NormalTok{power }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{|\textgreater{}}
+\NormalTok{power }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ term, }\AttributeTok{names\_sep =} \StringTok{"\_"}\NormalTok{, }
-    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{|\textgreater{}} 
+    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{\%\textgreater{}\%} 
   \FunctionTok{summarise}\NormalTok{(}
@@
     \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{)}
-\NormalTok{power }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{|\textgreater{}}
+\NormalTok{power }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(n\_subjects, n\_items, }\AttributeTok{fill =}\NormalTok{ power)) }\SpecialCharTok{+}
@@
 detect an effect of the expected size implied by our data simulation.
-This could be achieved by collecting data from 200 subjects (about 25\%
+This could be achieved by collecting data from 250 subjects (about 25\%
 of which will be experts), each completing the same 50 items (with
 advice present in about 67\% of cases, which is correct in about 80\% of
-cases with present advice). If collecting data from 200 subjects is not
-feasible, an alternative would be to recruit 150 subjects but increase
-the length of the experiment to over 70 items. However, 70 items might
-take too long to complete for the radiologists participating in the
-study, who have a busy schedule. The simulation suggests that it might
-also be possible to plan a shorter experiment with only 30 items if it
-is feasible to recruit an even higher number of subjects (\textgreater{}
-250, to be determined by additional rounds of power analysis). Design
-parameters that also affect power, and which could be investigated in
-the simulation to find a more optimal trade-off, are the ratio of
-experts to students, the frequency of whether advice is presented at
-all, and whether it is correct or incorrect.
+cases with present advice). If collecting data from 250 subjects is not
+feasible, an alternative would be to recruit 200 subjects but increase
+the length of the experiment to 70 items. However, 70 items might take
+too long to complete for the radiologists participating in the study,
+who have a busy schedule. The simulation suggests that it might also be
+possible to plan a shorter experiment with only 30 items if it is
+feasible to recruit an even higher number of subjects (to be determined
+by additional rounds of power analysis). Design parameters that also
+affect power, and which could be investigated in the simulation to find
+a more optimal trade-off, are the ratio of experts to students, the
+frequency of whether advice is presented at all, and whether it is
+correct or incorrect.\footnote{An advanced alternative to our heuristic
+  trade-off between subjects and items would be to explicitly optimize
+  for cost-efficient study designs using the \emph{mlpwr} package
+  (\citeproc{ref-zimmerSampleSizePlanning2023}{Zimmer et al., 2023}).
+  The package performs simulation based power analysis in a surrogate
+  modeling framework that can evaluate power for a large grid of design
+  combinations more efficiently. The framework also requires users to
+  write their own \texttt{sim\_and\_analyse} function and can be
+  considered a great extension on what we cover in our tutorial.}
 
@@
 \begin{Highlighting}[]
-\NormalTok{precision }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{|\textgreater{}}
+\NormalTok{precision }\OtherTok{\textless{}{-}}\NormalTok{ sim\_result }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ term, }\AttributeTok{names\_sep =} \StringTok{"\_"}\NormalTok{, }
-    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{width =} \StringTok{\textasciigrave{}}\AttributeTok{conf.high\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{{-}} \StringTok{\textasciigrave{}}\AttributeTok{conf.low\_b8=b2}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
+    \AttributeTok{values\_from =}\NormalTok{ estimate}\SpecialCharTok{:}\NormalTok{p.value.equiv) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(n\_subjects, n\_items) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{width =} \StringTok{\textasciigrave{}}\AttributeTok{conf.high\_b8=b2}\StringTok{\textasciigrave{}} \SpecialCharTok{{-}} \StringTok{\textasciigrave{}}\AttributeTok{conf.low\_b8=b2}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{summarise}\NormalTok{(}\AttributeTok{precision =} \FunctionTok{mean}\NormalTok{(width),}
@@
     \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{)}
-\NormalTok{precision }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{|\textgreater{}}
+\NormalTok{precision }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(n\_subjects, n\_items), factor)) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(n\_subjects, n\_items, }\AttributeTok{fill =}\NormalTok{ precision)) }\SpecialCharTok{+}
@@
 items as with planning for power. We could either choose 100 subjects
-and 30 items or 200 subjects and 10 items. Note that our simple
+and 30 items or 250 subjects and 10 items. Note that our simple
 heuristic for determining sample size in the planning for precision
@@
   the research hypothesis, while not violating domain expertise on how
-  plausible data should look might also contribute to planning more
+  plausible data should look, might also contribute to planning more
   insightful studies that can answer more precise research questions
@@
 
+\bibitem[\citeproctext]{ref-ebersoleManyLabsTesting2020}
+Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J.,
+Buttrick, N. R., Chartier, C. R., Corker, K. S., Corley, M., Hartshorne,
+J. K., IJzerman, H., Lazarević, L. B., Rabagliati, H., Ropovik, I.,
+Aczel, B., Aeschbach, L. F., Andrighetto, L., Arnal, J. D., Arrow, H.,
+Babincak, P., \ldots{} Nosek, B. A. (2020). Many {Labs} 5: {Testing
+Pre-Data-Collection Peer Review} as an {Intervention} to {Increase
+Replicability}. \emph{Advances in Methods and Practices in Psychological
+Science}, \emph{3}(3), 309--331.
+\url{https://doi.org/10.1177/2515245920958687}
+
+\bibitem[\citeproctext]{ref-endersSimpleMonteCarlo2023}
+Enders, C. K., Keller, B. T., \& Woller, M. P. (2023). A simple {Monte
+Carlo} method for estimating power in multilevel designs.
+\emph{Psychological Methods}. \url{https://doi.org/10.1037/met0000614}
+
 \bibitem[\citeproctext]{ref-fahrmeirRegressionModelsMethods2021}
@@
 \bibitem[\citeproctext]{ref-greenSIMRPackagePower2016a}
-Green, P., \& MacLeod, C. J. (2016). {\textsc{SIMR}} : An {R} package
+Green, P., \& MacLeod, C. J. (2016a). {\textsc{SIMR}} : An {R} package
 for power analysis of generalized linear mixed models by simulation.
 \emph{Methods in Ecology and Evolution}, \emph{7}(4), 493--498.
+\url{https://doi.org/10.1111/2041-210X.12504}
+
+\bibitem[\citeproctext]{ref-greenSIMRPackagePower2016}
+Green, P., \& MacLeod, C. J. (2016b). {SIMR}: An {R} package for power
+analysis of generalized linear mixed models by simulation. \emph{Methods
+in Ecology and Evolution}, \emph{7}(4), 493--498.
 \url{https://doi.org/10.1111/2041-210X.12504}
@@
 \url{https://doi.org/10.20982/tqmp.09.2.p043}
+
+\bibitem[\citeproctext]{ref-iddiPowerSampleSize2022}
+Iddi, S., \& C Donohue, M. (2022). Power and {Sample Size} for
+{Longitudinal Models} in {R} -- {The} longpower {Package} and {Shiny
+App}. \emph{The R Journal}, \emph{14}(1), 264--282.
+\url{https://doi.org/10.32614/RJ-2022-022}
+
+\bibitem[\citeproctext]{ref-johnsonPowerAnalysisGeneralized2015}
+Johnson, P. C. D., Barry, S. J. E., Ferguson, H. M., \& Müller, P.
+(2015). Power analysis for generalized linear mixed models in ecology
+and evolution. \emph{Methods in Ecology and Evolution}, \emph{6}(2),
+133--142. \url{https://doi.org/10.1111/2041-210X.12306}
 
@@
 
+\bibitem[\citeproctext]{ref-lebeauPowerAnalysisSimulation2019}
+LeBeau, B. (2019). \emph{Power {Analysis} by {Simulation} using {R} and
+simglm}. \url{https://doi.org/10.17077/f7kk-6w7f}
+
 \bibitem[\citeproctext]{ref-leeUsingTidyversePackage2020}
@@
 \emph{86}(3), 532--565. \url{https://doi.org/10.1177/00031224211004187}
+
+\bibitem[\citeproctext]{ref-magnussonConsequencesIgnoringTherapist2018}
+Magnusson, K., Andersson, G., \& Carlbring, P. (2018). The consequences
+of ignoring therapist effects in trials with longitudinal data: {A}
+simulation study. \emph{Journal of Consulting and Clinical Psychology},
+\emph{86}(9), 711--725. \url{https://doi.org/10.1037/ccp0000333}
+
+\bibitem[\citeproctext]{ref-marcochakyanyuPassLmePower2019}
+Marco Chak Yan YU. (2019). \emph{Pass.lme: {Power} and {Sample Size} for
+{Linear Mixed Effect Models}} (p. 0.9.0). Comprehensive R Archive
+Network. \url{https://doi.org/10.32614/CRAN.package.pass.lme}
+
+\bibitem[\citeproctext]{ref-martinMeasuringIndividualDifferences2011}
+Martin, J. G. A., Nussey, D. H., Wilson, A. J., \& Réale, D. (2011).
+Measuring individual differences in reaction norms in field and
+experimental studies: A power analysis of random regression models.
+\emph{Methods in Ecology and Evolution}, \emph{2}(4), 362--374.
+\url{https://doi.org/10.1111/j.2041-210X.2010.00084.x}
 
@@
 Vaughan, D., \& Dancho, M. (2022). \emph{Furrr: Apply mapping functions
-in parallel using futures}. \url{https://github.com/DavisVaughan/furrr}
+in parallel using futures}.
+\url{https://CRAN.R-project.org/package=furrr}
+
+\bibitem[\citeproctext]{ref-watsonGeneralisedLinearMixed2023}
+Watson, S. I. (2023). \emph{Generalised {Linear Mixed Model
+Specification}, {Analysis}, {Fitting}, and {Optimal Design} in {R} with
+the glmmr {Packages}}. arXiv.
+\url{https://doi.org/10.48550/ARXIV.2303.12657}
 
@@
 
-\bibitem[\citeproctext]{ref-zimmerSampleSizePlanning2022}
-Zimmer, F., Henninger, M., \& Debelak, R. (2022). \emph{Sample {Size
-Planning} for {Complex Study Designs}: {A Tutorial} for the mlpwr
-{Package}}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/r9w6t}
+\bibitem[\citeproctext]{ref-zhangPracticalStatisticalPower2018}
+Zhang, Z., \& Yuan, K.-H. (2018). \emph{Practical statistical power
+analysis using {Webpower} and {R}}. ISDSA Press.
+\url{https://doi.org/10.35566/power}
+
+\bibitem[\citeproctext]{ref-zimmerSampleSizePlanning2023}
+Zimmer, F., Henninger, M., \& Debelak, R. (2023). Sample size planning
+for complex study designs: {A} tutorial for the mlpwr package.
+\emph{Behavior Research Methods}.
+\url{https://doi.org/10.3758/s13428-023-02269-0}
 

### 2024-07-01 rcode.qmd
Built site for gh-pages

--- 2024-04-14 rcode.qmd
@@
 
+## Specify the population parameters
+
 ```{r}
@@
   mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+ "student, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+ "student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
+ "student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
   group_by(condition) %>%
@@
   mutate(condition = fct_recode(condition,
-    "no expert, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-    "no expert, wrong advice" = "0:1:0", "expert, wrong advice" = "1:1:0",
-    "no expert, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
+"student, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
+"student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
+"student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
   ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45") +
+  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45",
+    breaks = "Sturges") +
   scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
@@
 ```{r}
-library(tidyverse)
 library(marginaleffects)
-pred_df <- expand_grid(advice_present = 0:1, advice_correct = 0:1, expert = 0:1)
-pred_df
-probs <- predictions(fit, newdata = pred_df, type = "response", re.form = NA)
-probs
+library(tinytable)
+grid2 <- expand_grid(advice_present = 0:1, 
+  advice_correct = 0:1, expert = 0:1)
+grid2
+preds <- predictions(fit, newdata = grid2, 
+  type = "response", re.form = NA)
+print(preds, style = "tinytable") %>% theme_tt(theme = "resize")
 ```
@@
 ```{r}
-probs %>% 
+contrasts <- preds %>% 
   hypotheses(hypothesis = c(
-    "b8 = b2",
-    "b2 = b6",
-    "b7 = b1",
-    "b1 = b5"),
+    "b8 = b2",  # (correct advice, expert) - (no advice, expert)
+    "b2 = b6",  # (no advice, expert) - (incorrect advice, expert) 
+    "b7 = b1",  # (correct advice, student) - (no advice, student)
+    "b1 = b5"), # (no advice, student) - (incorrect advice, student)
     equivalence = c(0, 0))
+print(contrasts, style = "tinytable") %>% theme_tt(theme = "resize")
 ```

### 2024-07-01 manuscript.qmd
fix subsection formatting; fix some names and double reference in bibfile; fix step 7 section name; add whitespace before table 2 in pdf

@@
 Available formula-based and summary-statistics-based software packages for multilevel models often do not include GLMMs or are limited to simple designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022], making it necessary to build data simulations tailored specifically to the study design.
-A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2023; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016a; @johnsonPowerAnalysisGeneralized2015].
+A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2023; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021; @kainPracticalGuidePower2015; @greenSIMRPackagePower2016; @johnsonPowerAnalysisGeneralized2015].
 However, many of these tutorials focus on LMMs and the most common study designs [see @kumleEstimatingPowerGeneralized2021 for a tutorial that also covers more advanced settings].
@@
   col5 = c("none", "limited set of study designs with two levels", "none", "limited to two levels", "none", "none", "limited set of study designs with two levels", "limited set of study designs with two or three levels", "limited to two or three levels", "none", "limited set of study designs with two levels", "limited set of study designs with two levels"),
-  col6 = c("@watsonGeneralisedLinearMixed2023", "@iddiPowerSampleSize2022", "@kumleEstimatingPowerGeneralized2021", "@endersSimpleMonteCarlo2023", "@martinMeasuringIndividualDifferences2011", "@marcochakyanyuPassLmePower2019", "@lafitSelectionNumberParticipants2021", "@magnussonConsequencesIgnoringTherapist2018", "@lebeauPowerAnalysisSimulation2019", "@greenSIMRPackagePower2016", "@murayamaSummarystatisticsbasedPowerAnalysis2022", "@zhangPracticalStatisticalPower2018")
+  col6 = c("@watsonGeneralisedLinearMixed2023", "@iddiPowerSampleSize2022", "@kumleEstimatingPowerGeneralized2021", "@endersSimpleMonteCarlo2023", "@martinMeasuringIndividualDifferences2011", "@yumarcoPassLmePower2019", "@lafitSelectionNumberParticipants2021", "@magnussonConsequencesIgnoringTherapist2018", "@lebeauPowerAnalysisSimulation2019", "@greenSIMRPackagePower2016", "@murayamaSummarystatisticsbasedPowerAnalysis2022", "@zhangPracticalStatisticalPower2018")
 )
@@
 
-#### THEORY
+### THEORY
 
@@
 
-##### Our specific GLMM
-
-In a GLMM, the expected value of the dependent variable $Y$ conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$, and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
+**Our specific GLMM.** In a GLMM, the expected value of the dependent variable $Y$ conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$, and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]:
 $$
@@
 
-##### Simulation function in R
-
-The following R function simulates a full dataset structured according to the design of our case study.
+**Simulation function in R.** The following R function simulates a full dataset structured according to the design of our case study.
 
@@
 
+\newpage
+
 ```{r}
@@
 Parameter values ($\beta$ parameters, and $\sigma_S$) should be adjusted if the implications do not seem reasonable.
-We could also have a closer look at the variability between items by setting the subject standard deviation to almost zero ($\sigma_S = 0.01$).
-```{r}
-#| include: false
-set.seed(1)
-dat <- simulate(n_subjects = 3000, n_items = 3000, sd_u0s = 0.01)
-dat %>% 
-  mutate(condition = fct_cross(
-    factor(expert), factor(advice_present), factor(advice_correct))) %>%
-  mutate(condition = fct_recode(condition,
-"student, no advice" = "0:0:0", "expert, no advice" = "1:0:0", 
-"student, incorrect advice" = "0:1:0", "expert, incorrect advice" = "1:1:0",
-"student, correct advice" = "0:1:1", "expert, correct advice" = "1:1:1")) %>% 
-  ggplot(aes(x = y_prob, y = condition)) +
-  stat_histinterval(point_interval = "mean_qi", slab_color = "gray45",
-    breaks = "Sturges") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
-```
-
 The final plot demonstrates that these plots are also useful for spotting standard deviations that were specified too high.
@@
 
-##### PRACTICE
+### PRACTICE
 
@@
 
-#### Hypothesis testing
-
-If we chose hypothesis testing for our case study, we would test a combined null hypothesis $H_0$ that consists of four separate null hypotheses:
+**Hypothesis testing.** If we chose hypothesis testing for our case study, we would test a combined null hypothesis $H_0$ that consists of four separate null hypotheses:
 $$
@@
 
-#### Interval estimation
-
-If we chose parameter estimation for our case study, we would focus on the two-sided CIs of the four contrasts of interest.
+**Interval estimation.** If we chose parameter estimation for our case study, we would focus on the two-sided CIs of the four contrasts of interest.
 With a standard confidence level of $1 - \alpha = 0.95$ plausible values are clearly in the positive range for the first, third, and fourth contrast, while both negative and positive values seem plausible for the second contrast.
@@
 
-#### Power results
-
-For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the $\alpha = 0.05$ level.
+**Power results.** For our exemplary combined hypothesis, power is defined as the (long-run) percentage of simulations in which all four p-values of our individual hypotheses are significant at the $\alpha = 0.05$ level.
 Based on our simulation outcomes, we compute a power estimate for each combination of `n_subjects` $\times$ `n_items` (including 95% CIs) and visualize the results with the following code.^[This code was inspired by the "Mixed Design Simulation" vignette of the *faux* package at <https://debruine.github.io/faux/articles/sim_mixed.html>.]
@@
 
-#### Precision results
-
-When planning for precision, one could monitor the width of all four CIs at the same time.
+**Precision results.** When planning for precision, one could monitor the width of all four CIs at the same time.
 However, because the CIs of the four contrasts strongly differ in width, it is not trivial to decide which width one should target when deciding on the appropriate sample size.
@@
 
-## Step 7 (optional): Sensitivity analysis
+## Step 7 (optional): Conduct sensitivity analysis
 

### 2024-07-01 manuscript.qmd
fix yu reference again...

@@
   col5 = c("none", "limited set of study designs with two levels", "none", "limited to two levels", "none", "none", "limited set of study designs with two levels", "limited set of study designs with two or three levels", "limited to two or three levels", "none", "limited set of study designs with two levels", "limited set of study designs with two levels"),
-  col6 = c("@watsonGeneralisedLinearMixed2023", "@iddiPowerSampleSize2022", "@kumleEstimatingPowerGeneralized2021", "@endersSimpleMonteCarlo2023", "@martinMeasuringIndividualDifferences2011", "@yumarcoPassLmePower2019", "@lafitSelectionNumberParticipants2021", "@magnussonConsequencesIgnoringTherapist2018", "@lebeauPowerAnalysisSimulation2019", "@greenSIMRPackagePower2016", "@murayamaSummarystatisticsbasedPowerAnalysis2022", "@zhangPracticalStatisticalPower2018")
+  col6 = c("@watsonGeneralisedLinearMixed2023", "@iddiPowerSampleSize2022", "@kumleEstimatingPowerGeneralized2021", "@endersSimpleMonteCarlo2023", "@martinMeasuringIndividualDifferences2011", "@yuPassLmePower2019", "@lafitSelectionNumberParticipants2021", "@magnussonConsequencesIgnoringTherapist2018", "@lebeauPowerAnalysisSimulation2019", "@greenSIMRPackagePower2016", "@murayamaSummarystatisticsbasedPowerAnalysis2022", "@zhangPracticalStatisticalPower2018")
 )

### 2024-07-01 _manuscript/_tex/manuscript.tex
update _freeze and _manuscript

@@
 \citeproc{ref-debruineUnderstandingMixedEffectsModels2021}{DeBruine \&
-Barr, 2021}; \citeproc{ref-greenSIMRPackagePower2016a}{Green \& MacLeod,
-2016a}; \citeproc{ref-johnsonPowerAnalysisGeneralized2015}{Johnson et
+Barr, 2021}; \citeproc{ref-greenSIMRPackagePower2016}{Green \& MacLeod,
+2016}; \citeproc{ref-johnsonPowerAnalysisGeneralized2015}{Johnson et
 al., 2015}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al.,
@@
 
-\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://atrihub.shinyapps.io/power/}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Iddi}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\&}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{C}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Donohue}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2022)}}}} \\
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://atrihub.shinyapps.io/power/}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Iddi}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\&}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Donohue}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2022)}}}} \\
 
@@
 
-\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{pass.lme}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{documentation}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Marco}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Chak}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Yan}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{YU}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2019)}}}} \\
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{pass.lme}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{documentation}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Yu}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2019)}}}} \\
 
@@
 
-\paragraph{THEORY.}\label{theory-1}
+\subsubsection{THEORY}\label{theory-1}
 
@@
 
-\paragraph{PRACTICE.}\label{practice-1}
+\subsubsection{PRACTICE}\label{practice-1}
 
@@
 
-\subparagraph{Our specific GLMM.}\label{our-specific-glmm}
-
-In a GLMM, the expected value of the dependent variable \(Y\)
-conditioned on the vector of predictor variables \(\mathbf{X}\) and
-random effects \(\mathbf{U}\), transformed by a link function \(g()\) is
-modeled as a linear combination \(\eta\) of the predictor variables
-\(\mathbf{X}\), the random effects \(\mathbf{U}\), and the model
-parameters \(\mathbf{\beta}\)
+\textbf{Our specific GLMM.} In a GLMM, the expected value of the
+dependent variable \(Y\) conditioned on the vector of predictor
+variables \(\mathbf{X}\) and random effects \(\mathbf{U}\), transformed
+by a link function \(g()\) is modeled as a linear combination \(\eta\)
+of the predictor variables \(\mathbf{X}\), the random effects
+\(\mathbf{U}\), and the model parameters \(\mathbf{\beta}\)
 (\citeproc{ref-fahrmeirRegressionModelsMethods2021}{Fahrmeir et al.,
@@
 
-\subparagraph{Simulation function in R.}\label{simulation-function-in-r}
-
-The following R function simulates a full dataset structured according
-to the design of our case study.
+\textbf{Simulation function in R.} The following R function simulates a
+full dataset structured according to the design of our case study.
 
@@
 implied by these probabilities for each experimental condition.
+
+\newpage
 
@@
 
-\paragraph{THEORY.}\label{theory-3}
+\subsubsection{THEORY}\label{theory-3}
 
@@
 
-\paragraph{PRACTICE.}\label{practice-3}
+\subsubsection{PRACTICE}\label{practice-3}
 
@@
 
-\paragraph{THEORY.}\label{theory-4}
+\subsubsection{THEORY}\label{theory-4}
 
@@
 
-\paragraph{PRACTICE.}\label{practice-4}
+\subsubsection{PRACTICE}\label{practice-4}
 
@@
 experts. Parameter values (\(\beta\) parameters, and \(\sigma_S\))
-should be adjusted if the implications do not seem reasonable. We could
-also have a closer look at the variability between items by setting the
-subject standard deviation to almost zero (\(\sigma_S = 0.01\)).
-
+should be adjusted if the implications do not seem reasonable.\\
 The final plot demonstrates that these plots are also useful for
@@
 
-\paragraph{THEORY.}\label{theory-5}
+\subsubsection{THEORY}\label{theory-5}
 
@@
 
-\subparagraph{PRACTICE.}\label{practice-5}
+\subsubsection{PRACTICE}\label{practice-5}
 
@@
 
-\paragraph{Hypothesis testing.}\label{hypothesis-testing}
-
-If we chose hypothesis testing for our case study, we would test a
-combined null hypothesis \(H_0\) that consists of four separate null
-hypotheses: \[
+\textbf{Hypothesis testing.} If we chose hypothesis testing for our case
+study, we would test a combined null hypothesis \(H_0\) that consists of
+four separate null hypotheses: \[
 \begin{aligned}
@@
 
-\paragraph{Interval estimation.}\label{interval-estimation}
-
-If we chose parameter estimation for our case study, we would focus on
-the two-sided CIs of the four contrasts of interest. With a standard
-confidence level of \(1 - \alpha = 0.95\) plausible values are clearly
-in the positive range for the first, third, and fourth contrast, while
-both negative and positive values seem plausible for the second
-contrast. Note that due to the constrained range of the probability
-scale, the width of the CI differs between the four contrasts (which is
-the expected behavior for binomial GLMMs). The smallest width is
-observed for the first contrast (expert with correct advice vs.~expert
-without advice), where both underlying probabilities are close to 1. The
-largest width is observed for the fourth contrast (student with
-incorrect advice vs.~student without advice), where both underlying
-probabilities are closer to 0.5.
+\textbf{Interval estimation.} If we chose parameter estimation for our
+case study, we would focus on the two-sided CIs of the four contrasts of
+interest. With a standard confidence level of \(1 - \alpha = 0.95\)
+plausible values are clearly in the positive range for the first, third,
+and fourth contrast, while both negative and positive values seem
+plausible for the second contrast. Note that due to the constrained
+range of the probability scale, the width of the CI differs between the
+four contrasts (which is the expected behavior for binomial GLMMs). The
+smallest width is observed for the first contrast (expert with correct
+advice vs.~expert without advice), where both underlying probabilities
+are close to 1. The largest width is observed for the fourth contrast
+(student with incorrect advice vs.~student without advice), where both
+underlying probabilities are closer to 0.5.
 
@@
 
-\paragraph{Power results.}\label{power-results}
-
-For our exemplary combined hypothesis, power is defined as the
-(long-run) percentage of simulations in which all four p-values of our
-individual hypotheses are significant at the \(\alpha = 0.05\) level.
-Based on our simulation outcomes, we compute a power estimate for each
-combination of \texttt{n\_subjects} \(\times\) \texttt{n\_items}
-(including 95\% CIs) and visualize the results with the following
-code.\footnote{This code was inspired by the ``Mixed Design Simulation''
-  vignette of the \emph{faux} package at
+\textbf{Power results.} For our exemplary combined hypothesis, power is
+defined as the (long-run) percentage of simulations in which all four
+p-values of our individual hypotheses are significant at the
+\(\alpha = 0.05\) level. Based on our simulation outcomes, we compute a
+power estimate for each combination of \texttt{n\_subjects} \(\times\)
+\texttt{n\_items} (including 95\% CIs) and visualize the results with
+the following code.\footnote{This code was inspired by the ``Mixed
+  Design Simulation'' vignette of the \emph{faux} package at
   \url{https://debruine.github.io/faux/articles/sim_mixed.html}.}
@@
 
-\paragraph{Precision results.}\label{precision-results}
-
-When planning for precision, one could monitor the width of all four CIs
-at the same time. However, because the CIs of the four contrasts
-strongly differ in width, it is not trivial to decide which width one
-should target when deciding on the appropriate sample size. In contrast
-to planning for power, there are no common standards on how to specify
-the targeted precision
+\textbf{Precision results.} When planning for precision, one could
+monitor the width of all four CIs at the same time. However, because the
+CIs of the four contrasts strongly differ in width, it is not trivial to
+decide which width one should target when deciding on the appropriate
+sample size. In contrast to planning for power, there are no common
+standards on how to specify the targeted precision
 (\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}). For
@@
 
-\subsection{Step 7 (optional): Sensitivity
-analysis}\label{step-7-optional-sensitivity-analysis}
+\subsection{Step 7 (optional): Conduct sensitivity
+analysis}\label{step-7-optional-conduct-sensitivity-analysis}
 
@@
 
-\bibitem[\citeproctext]{ref-greenSIMRPackagePower2016a}
-Green, P., \& MacLeod, C. J. (2016a). {\textsc{SIMR}} : An {R} package
-for power analysis of generalized linear mixed models by simulation.
-\emph{Methods in Ecology and Evolution}, \emph{7}(4), 493--498.
-\url{https://doi.org/10.1111/2041-210X.12504}
-
 \bibitem[\citeproctext]{ref-greenSIMRPackagePower2016}
-Green, P., \& MacLeod, C. J. (2016b). {SIMR}: An {R} package for power
+Green, P., \& MacLeod, C. J. (2016). {SIMR}: An {R} package for power
 analysis of generalized linear mixed models by simulation. \emph{Methods
@@
 \bibitem[\citeproctext]{ref-iddiPowerSampleSize2022}
-Iddi, S., \& C Donohue, M. (2022). Power and {Sample Size} for
+Iddi, S., \& Donohue, M. C. (2022). Power and {Sample Size} for
 {Longitudinal Models} in {R} -- {The} longpower {Package} and {Shiny
@@
 
-\bibitem[\citeproctext]{ref-marcochakyanyuPassLmePower2019}
-Marco Chak Yan YU. (2019). \emph{Pass.lme: {Power} and {Sample Size} for
-{Linear Mixed Effect Models}} (p. 0.9.0). Comprehensive R Archive
-Network. \url{https://doi.org/10.32614/CRAN.package.pass.lme}
-
 \bibitem[\citeproctext]{ref-martinMeasuringIndividualDifferences2011}
@@
 
+\bibitem[\citeproctext]{ref-yuPassLmePower2019}
+Yu, M. (2019). \emph{Pass.lme: {Power} and {Sample Size} for {Linear
+Mixed Effect Models}} (p. 0.9.0). Comprehensive R Archive Network.
+\url{https://doi.org/10.32614/CRAN.package.pass.lme}
+
 \bibitem[\citeproctext]{ref-zhangPracticalStatisticalPower2018}

### 2024-07-24 manuscript.qmd
correct mistake: mixedpower is based on lme4 not simr

@@
               "Formula-based power analysis; focus on longitudinal data",
-              "Simulation-based power analysis; based on 'simr' package",
+              "Simulation-based power analysis; based on 'lme4' package",
               "Simulation-based power analysis; missing data mechanisms",

### 2024-07-24 manuscript.qmd
add limitation of case study design to abstract and introduction

@@
 Tailored data simulation approaches offer a more flexible alternative.
-Based on a practical case study, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori sample size planning with GLMMs.
+Based on a practical case study where we focus on a binomial GLMM with two random intercepts and discrete predictor variables, the current tutorial equips researchers with a step-by-step guide and corresponding code for conducting tailored a priori sample size planning with GLMMs.
 We not only focus on power analysis but also explain how to use the precision of parameter estimates to determine appropriate sample sizes.
@@
 After introducing some theoretical background on sample size planning, we review existing software solutions in R and discuss under which circumstances tailored data simulations are necessary.
-We proceed by describing the relevant steps and decisions involved in tailored data simulation, illustrated in a case study.
+We proceed by describing the relevant steps and decisions involved in tailored data simulation, illustrated in a case study where we focus on a binomial GLMM with two random intercepts and discrete predictor variables.
 

### 2024-07-24 manuscript.qmd
add further discussion of simplified data generating process, maximal vs. parsimonious models based on matuschek et al. (2017)

@@
 The quality of the results of the sample size planning crucially depends on the plausibility of the simulated data-generating process.
-However, we argue that even a strongly simplified data-generating process (e.g., only a small number of interaction effects; only random intercepts, and no random slopes; assuming that data is missing completely at random) can yield informative results.
+If the simulated data generating process is less complex than the true process, one can expect that sample size planning underestimates the required sample size and the planned analysis has inflated type I error rates [@matuschekBalancingTypeError2017].
+However, @matuschekBalancingTypeError2017 have also shown in simulations with LMMs that while fitting *maximal models* (i.e.
+that include all possible random effects) safeguards against inflated type I error rates, this can lead to a great loss in power if the variance in the random effects is actually small.
+Thus, we argue that even a simplified data-generating process (e.g., only a small number of interaction effects; only random intercepts, and no random slopes; assuming that data is missing completely at random) that is only plausible under idealized circumstances, can yield informative results and is preferred over performing no systematic sample size planning.
 
@@
 Note that this does not always mean that one will specify the same GLMM that was used when specifying the data-generating process.
-On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible.
+On the one hand, using a simpler model for data simulation than for model estimation can be a useful strategy in scenarios where making plausible assumptions for complicated random effect structures and interactions is not feasible, or the omitted components are expected to have small effects [@matuschekBalancingTypeError2017].
 On the other hand, using a more complex model for data simulation than for model estimation can be a useful strategy in scenarios where one has specific domain knowledge about aspects of the data-generating process that are still difficult to estimate with the current state-of-the-art in multilevel modeling.
+In fact, applying models that are more parsimonious than the true data generating process can provide a better trade-off between type I error rates (or bias) and power (or precision) in many scenarios [@matuschekBalancingTypeError2017].
 

### 2024-07-29 manuscript.qmd
add literature and extend the sections specify population parameters and iterate with domain expertise

@@
 Population parameters are all model parameters estimated in a GLMM, in particular the regression coefficients of the fixed effects and the standard deviation of the random effects (and the correlation between random effects in more complicated models).
-In the absence of previous studies with the same design or pilot data, strategies to specify population parameters require access to domain knowledge from domain experts.
+In the absence of previous studies with the same design or pilot data, strategies to specify population parameters will always require access to some domain knowledge.
+In our experience, research teams for projects where tailored simulation-based sample size planning is necessary typically consist of *analysts* (who are primarily responsible for study design and statistical analysis), and *domain experts* (who often provide the original research question and are responsible for data collection).
+In smaller projects, analysts and domain experts might be the same persons. In these research settings, a broad review of the available literature typically plays a more important role.
 Because most domain knowledge can only be expressed in unstandardized measurement units of a specific application, we argue that unstandardized effect sizes are usually preferable over standardized effect sizes for tailored simulation-based sample size planning.
-The basic idea of all strategies is to quantify or visualize the data-generating process implied by certain values of population parameters in an intuitive way that enables calibration of population parameters based on the available knowledge of domain experts.
-Although we use frequentist model estimation in our tutorial, many strategies demonstrated in this chapter are inspired by research on monitoring the plausibility of model assumptions in applied Bayesian statistics [@gelmanBayesianWorkflow2020].
+The basic idea of all strategies is to quantify or visualize the data-generating process implied by certain values of population parameters in an intuitive way that enables calibration of population parameters based on the available domain knowledge.
+Although we use frequentist model estimation in our tutorial, many strategies demonstrated in this chapter are inspired by research on monitoring the plausibility of model assumptions [@gelmanBayesianWorkflow2020; @schadPrincipledBayesianWorkflow2021] and eliciting prior information [@bocktingSimulationBasedPriorKnowledge2023; @mikkolaPriorKnowledgeElicitation2023; @hartmannFlexiblePriorElicitation2020; @stefanPracticalChallengesMethodological2022] in applied Bayesian statistics.
 
@@
 
-### Iterate with domain experts
+### Iterate with domain expertise
 
@@
 
-Gathering domain knowledge and the specification of population parameter values is an iterative process, in which domain experts repeatedly assess the plausibility of simulated data.
-As a first step, domain experts can be interviewed to “elicit” their domain knowledge about how the data of the planned study is expected to look.
-As most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
+Tailored simulation-based sample size planning always requires access to domain knowledge.
+However, gathering domain knowledge and the specification of population parameter values is not always straightforward and can be better described as an iterative process, in which the plausibility of simulated data is repeatedly compared with the available domain expertise.
+As a first step, dedicated domain experts (if available) can be interviewed to *elicit* their domain knowledge about how the data of the planned study is expected to look.
+Because most domain experts are no experts in statistical modeling and GLMMs, they often struggle without further guidance to communicate their knowledge in a way that is useful when specifying the parameters for data simulation.
 For this reason, we suggest that after an initial interview, the analyst who is familiar with the structure of the GLMM selects an initial set of insightful descriptive statistics and model-based quantities.
-Then they reenter into an iterative discussion with the domain experts where some set of population values are selected, and the plausibility of resulting implied quantities is evaluated.
-Then, the population parameters are updated until the domain experts are satisfied with the result.
+Then, they reenter into an iterative discussion with the domain experts where some set of population values are selected, and the plausibility of resulting implied quantities is evaluated.
+The population parameters are updated until the domain experts are satisfied with the result.
 During this process, the monitored descriptive statistics and model-based quantities can be updated or extended to capture as much available domain knowledge as possible.
+Even if no dedicated domain experts are available, the basic principles stay the same except that initial parameter values are often derived from a more extensive literature review.
+Descriptive statistics and model-based quantities are still used to evaluate whether the implied assumptions about the data generating process are plausible.
+Often, this judgment evolves over several iterations as the analyst becomes more familiar with the model and the study design.^[Developing more concrete strategies on how to elicit and visualize domain knowledge is currently an active area of research, primarily in the context of Bayesian modeling [for reviews, see for example @mikkolaPriorKnowledgeElicitation2023 and @stefanPracticalChallengesMethodological2022].
+We expect that future advances in this field will also be highly relevant for tailored simulation-based sample size planning with frequentist models.]
 
@@
 Design parameters that also affect power, and which could be investigated in the simulation to find a more optimal trade-off, are the ratio of experts to students, the frequency of whether advice is presented at all, and whether it is correct or incorrect.^[An advanced alternative to our heuristic trade-off between subjects and items would be to explicitly optimize for cost-efficient study designs using the *mlpwr* package [@zimmerSampleSizePlanning2023].
-The package performs simulation based power analysis in a surrogate modeling framework that can evaluate power for a large grid of design combinations more efficiently.
+The package performs simulation-based power analysis in a surrogate modeling framework that can evaluate power for a large grid of design combinations more efficiently.
 The framework also requires users to write their own `sim_and_analyse` function and can be considered a great extension on what we cover in our tutorial.]

### 2024-07-31 manuscript.qmd
set preprint version to 3.5 (between major psyarxiv releases)
added a comment on how sensitivity analysis can be used to investigate the consequences of a mismatch between data-generating process and statistical model

@@
   disclosures:
-    data-sharing: This is version 3 of our preprint published at <https://doi.org/10.31234/osf.io/rpjem>.
+    data-sharing: This is version 3.5 of our preprint published at <https://doi.org/10.31234/osf.io/rpjem>.
 All materials (reproducible manuscript, including R code and all simulation results) are available in the project's repository on the Open Science Framework (OSF) at <https://osf.io/dhwf4/>.
@@
 Power (or precision) could then be simulated for several of these scenarios (across different numbers of subjects and items, as considered earlier).
+In steps 2 and 4 of the tutorial, we have briefly discussed scenarios where the applied statistical model is more (or less) parsimonious than the data-generating process.
+Sensitivity analysis can be used to assess the consequences of a mismatch by investigating different combinations of data-generating processes and statistical models.
+Such extended simulations can be a great resource to make an informed decision about the concrete model to estimate for the planned study where the true data-generating process is unknown.
 Although recommended, we do not present a sensitivity analysis for our case study to keep the length of the tutorial manageable.

### 2024-07-31 manuscript.qmd
Built site for gh-pages

@@
 The basic idea of all strategies is to quantify or visualize the data-generating process implied by certain values of population parameters in an intuitive way that enables calibration of population parameters based on the available domain knowledge.
-Although we use frequentist model estimation in our tutorial, many strategies demonstrated in this chapter are inspired by research on monitoring the plausibility of model assumptions [@gelmanBayesianWorkflow2020; @schadPrincipledBayesianWorkflow2021] and eliciting prior information [@bocktingSimulationBasedPriorKnowledge2023; @mikkolaPriorKnowledgeElicitation2023; @hartmannFlexiblePriorElicitation2020; @stefanPracticalChallengesMethodological2022] in applied Bayesian statistics.
+Although we use frequentist model estimation in our tutorial, many strategies demonstrated in this chapter are inspired by research on monitoring the plausibility of model assumptions [@gelmanBayesianWorkflow2020; @schadPrincipledBayesianWorkflow2021] and eliciting prior information [@bocktingSimulationbasedPriorKnowledge2024; @mikkolaPriorKnowledgeElicitation2023; @hartmannFlexiblePriorElicitation2020; @stefanPracticalChallengesMethodological2022] in applied Bayesian statistics.
 
@@
 Descriptive statistics and model-based quantities are still used to evaluate whether the implied assumptions about the data generating process are plausible.
-Often, this judgment evolves over several iterations as the analyst becomes more familiar with the model and the study design.^[Developing more concrete strategies on how to elicit and visualize domain knowledge is currently an active area of research, primarily in the context of Bayesian modeling [for reviews, see for example @mikkolaPriorKnowledgeElicitation2023 and @stefanPracticalChallengesMethodological2022].
+Often, this judgment evolves over several iterations as the analyst becomes more familiar with the model and the study design.^[Developing more effective strategies on how to elicit and visualize domain knowledge is currently an active area of research, primarily in the context of Bayesian modeling [@bocktingSimulationbasedPriorKnowledge2024; @mikkolaPriorKnowledgeElicitation2023; @stefanPracticalChallengesMethodological2022; @hartmannFlexiblePriorElicitation2020].
 We expect that future advances in this field will also be highly relevant for tailored simulation-based sample size planning with frequentist models.]

### 2024-07-31 _manuscript/_tex/manuscript.tex
update website and commit changes to _freeze and _manuscript

@@
 approaches offer a more flexible alternative. Based on a practical case
-study, the current tutorial equips researchers with a step-by-step guide
-and corresponding code for conducting tailored a priori sample size
-planning with GLMMs. We not only focus on power analysis but also
-explain how to use the precision of parameter estimates to determine
-appropriate sample sizes. We conclude with an outlook on the increasing
-importance of simulation-based sample size planning.}
+study where we focus on a binomial GLMM with two random intercepts and
+discrete predictor variables, the current tutorial equips researchers
+with a step-by-step guide and corresponding code for conducting tailored
+a priori sample size planning with GLMMs. We not only focus on power
+analysis but also explain how to use the precision of parameter
+estimates to determine appropriate sample sizes. We conclude with an
+outlook on the increasing importance of simulation-based sample size
+planning.}
 
@@
 
-\par{ This is version 3 of our preprint published at
+\par{ This is version 3.5 of our preprint published at
 \url{https://doi.org/10.31234/osf.io/rpjem}. All materials (reproducible
@@
 describing the relevant steps and decisions involved in tailored data
-simulation, illustrated in a case study.
+simulation, illustrated in a case study where we focus on a binomial
+GLMM with two random intercepts and discrete predictor variables.
 
@@
 
-\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mixedpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{simr}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Kumle}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2021)}}}} \\
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mixedpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{lme4}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Kumle}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2021)}}}} \\
 
@@
 quality of the results of the sample size planning crucially depends on
-the plausibility of the simulated data-generating process. However, we
-argue that even a strongly simplified data-generating process (e.g.,
-only a small number of interaction effects; only random intercepts, and
-no random slopes; assuming that data is missing completely at random)
-can yield informative results.
+the plausibility of the simulated data-generating process. If the
+simulated data generating process is less complex than the true process,
+one can expect that sample size planning underestimates the required
+sample size and the planned analysis has inflated type I error rates
+(\citeproc{ref-matuschekBalancingTypeError2017}{Matuschek et al.,
+2017}). However, Matuschek et al.
+(\citeproc{ref-matuschekBalancingTypeError2017}{2017}) have also shown
+in simulations with LMMs that while fitting \emph{maximal models}
+(i.e.~that include all possible random effects) safeguards against
+inflated type I error rates, this can lead to a great loss in power if
+the variance in the random effects is actually small. Thus, we argue
+that even a simplified data-generating process (e.g., only a small
+number of interaction effects; only random intercepts, and no random
+slopes; assuming that data is missing completely at random) that is only
+plausible under idealized circumstances, can yield informative results
+and is preferred over performing no systematic sample size planning.
 
@@
 studies with the same design or pilot data, strategies to specify
-population parameters require access to domain knowledge from domain
-experts. Because most domain knowledge can only be expressed in
-unstandardized measurement units of a specific application, we argue
-that unstandardized effect sizes are usually preferable over
-standardized effect sizes for tailored simulation-based sample size
-planning. The basic idea of all strategies is to quantify or visualize
-the data-generating process implied by certain values of population
-parameters in an intuitive way that enables calibration of population
-parameters based on the available knowledge of domain experts. Although
+population parameters will always require access to some domain
+knowledge. In our experience, research teams for projects where tailored
+simulation-based sample size planning is necessary typically consist of
+\emph{analysts} (who are primarily responsible for study design and
+statistical analysis), and \emph{domain experts} (who often provide the
+original research question and are responsible for data collection). In
+smaller projects, analysts and domain experts might be the same persons.
+In these research settings, a broad review of the available literature
+typically plays a more important role. Because most domain knowledge can
+only be expressed in unstandardized measurement units of a specific
+application, we argue that unstandardized effect sizes are usually
+preferable over standardized effect sizes for tailored simulation-based
+sample size planning. The basic idea of all strategies is to quantify or
+visualize the data-generating process implied by certain values of
+population parameters in an intuitive way that enables calibration of
+population parameters based on the available domain knowledge. Although
 we use frequentist model estimation in our tutorial, many strategies
 demonstrated in this chapter are inspired by research on monitoring the
-plausibility of model assumptions in applied Bayesian statistics
-(\citeproc{ref-gelmanBayesianWorkflow2020}{Gelman et al., 2020}).
+plausibility of model assumptions
+(\citeproc{ref-gelmanBayesianWorkflow2020}{Gelman et al., 2020};
+\citeproc{ref-schadPrincipledBayesianWorkflow2021}{Schad et al., 2021})
+and eliciting prior information
+(\citeproc{ref-bocktingSimulationbasedPriorKnowledge2024}{Bockting et
+al., 2024}; \citeproc{ref-hartmannFlexiblePriorElicitation2020}{Hartmann
+et al., 2020};
+\citeproc{ref-mikkolaPriorKnowledgeElicitation2023}{Mikkola et al.,
+2023}; \citeproc{ref-stefanPracticalChallengesMethodological2022}{Stefan
+et al., 2022}) in applied Bayesian statistics.
 
@@
 \subsubsection{Iterate with domain
-experts}\label{iterate-with-domain-experts}
+expertise}\label{iterate-with-domain-expertise}
 
@@
 
-Gathering domain knowledge and the specification of population parameter
-values is an iterative process, in which domain experts repeatedly
-assess the plausibility of simulated data. As a first step, domain
-experts can be interviewed to ``elicit'' their domain knowledge about
-how the data of the planned study is expected to look. As most domain
-experts are no experts in statistical modeling and GLMMs, they often
-struggle without further guidance to communicate their knowledge in a
-way that is useful when specifying the parameters for data simulation.
-For this reason, we suggest that after an initial interview, the analyst
-who is familiar with the structure of the GLMM selects an initial set of
-insightful descriptive statistics and model-based quantities. Then they
-reenter into an iterative discussion with the domain experts where some
-set of population values are selected, and the plausibility of resulting
-implied quantities is evaluated. Then, the population parameters are
-updated until the domain experts are satisfied with the result. During
-this process, the monitored descriptive statistics and model-based
-quantities can be updated or extended to capture as much available
-domain knowledge as possible.
+Tailored simulation-based sample size planning always requires access to
+domain knowledge. However, gathering domain knowledge and the
+specification of population parameter values is not always
+straightforward and can be better described as an iterative process, in
+which the plausibility of simulated data is repeatedly compared with the
+available domain expertise. As a first step, dedicated domain experts
+(if available) can be interviewed to \emph{elicit} their domain
+knowledge about how the data of the planned study is expected to look.
+Because most domain experts are no experts in statistical modeling and
+GLMMs, they often struggle without further guidance to communicate their
+knowledge in a way that is useful when specifying the parameters for
+data simulation. For this reason, we suggest that after an initial
+interview, the analyst who is familiar with the structure of the GLMM
+selects an initial set of insightful descriptive statistics and
+model-based quantities. Then, they reenter into an iterative discussion
+with the domain experts where some set of population values are
+selected, and the plausibility of resulting implied quantities is
+evaluated. The population parameters are updated until the domain
+experts are satisfied with the result. During this process, the
+monitored descriptive statistics and model-based quantities can be
+updated or extended to capture as much available domain knowledge as
+possible. Even if no dedicated domain experts are available, the basic
+principles stay the same except that initial parameter values are often
+derived from a more extensive literature review. Descriptive statistics
+and model-based quantities are still used to evaluate whether the
+implied assumptions about the data generating process are plausible.
+Often, this judgment evolves over several iterations as the analyst
+becomes more familiar with the model and the study design.\footnote{Developing
+  more effective strategies on how to elicit and visualize domain
+  knowledge is currently an active area of research, primarily in the
+  context of Bayesian modeling
+  (\citeproc{ref-bocktingSimulationbasedPriorKnowledge2024}{Bockting et
+  al., 2024};
+  \citeproc{ref-hartmannFlexiblePriorElicitation2020}{Hartmann et al.,
+  2020}; \citeproc{ref-mikkolaPriorKnowledgeElicitation2023}{Mikkola et
+  al., 2023};
+  \citeproc{ref-stefanPracticalChallengesMethodological2022}{Stefan et
+  al., 2022}). We expect that future advances in this field will also be
+  highly relevant for tailored simulation-based sample size planning
+  with frequentist models.}
 
@@
 assumptions for complicated random effect structures and interactions is
-not feasible. On the other hand, using a more complex model for data
+not feasible, or the omitted components are expected to have small
+effects (\citeproc{ref-matuschekBalancingTypeError2017}{Matuschek et
+al., 2017}). On the other hand, using a more complex model for data
 simulation than for model estimation can be a useful strategy in
@@
 data-generating process that are still difficult to estimate with the
-current state-of-the-art in multilevel modeling.
+current state-of-the-art in multilevel modeling. In fact, applying
+models that are more parsimonious than the true data generating process
+can provide a better trade-off between type I error rates (or bias) and
+power (or precision) in many scenarios
+(\citeproc{ref-matuschekBalancingTypeError2017}{Matuschek et al.,
+2017}).
 
@@
   (\citeproc{ref-zimmerSampleSizePlanning2023}{Zimmer et al., 2023}).
-  The package performs simulation based power analysis in a surrogate
+  The package performs simulation-based power analysis in a surrogate
   modeling framework that can evaluate power for a large grid of design
@@
 simulated for several of these scenarios (across different numbers of
-subjects and items, as considered earlier). Although recommended, we do
-not present a sensitivity analysis for our case study to keep the length
-of the tutorial manageable.
+subjects and items, as considered earlier). In steps 2 and 4 of the
+tutorial, we have briefly discussed scenarios where the applied
+statistical model is more (or less) parsimonious than the
+data-generating process. Sensitivity analysis can be used to assess the
+consequences of a mismatch by investigating different combinations of
+data-generating processes and statistical models. Such extended
+simulations can be a great resource to make an informed decision about
+the concrete model to estimate for the planned study where the true
+data-generating process is unknown. Although recommended, we do not
+present a sensitivity analysis for our case study to keep the length of
+the tutorial manageable.
 
@@
 
+\bibitem[\citeproctext]{ref-bocktingSimulationbasedPriorKnowledge2024}
+Bockting, F., Radev, S. T., \& Bürkner, P.-C. (2024). Simulation-based
+prior knowledge elicitation for parametric {Bayesian} models.
+\emph{Scientific Reports}, \emph{14}(1), 17330.
+\url{https://doi.org/10.1038/s41598-024-68090-7}
+
 \bibitem[\citeproctext]{ref-bolkerLinearGeneralizedLinear2015}
@@
 
+\bibitem[\citeproctext]{ref-hartmannFlexiblePriorElicitation2020}
+Hartmann, M., Agiashvili, G., Bürkner, P., \& Klami, A. (2020). Flexible
+prior elicitation via the prior predictive distribution. In J. Peters \&
+D. Sontag (Eds.), \emph{Proceedings of the 36th conference on
+uncertainty in artificial intelligence ({UAI})} (Vol. 124, pp.
+1129--1138). PMLR.
+
 \bibitem[\citeproctext]{ref-iddiPowerSampleSize2022}
@@
 
+\bibitem[\citeproctext]{ref-mikkolaPriorKnowledgeElicitation2023}
+Mikkola, P., Martin, O. A., Chandramouli, S., Hartmann, M., Abril Pla,
+O., Thomas, O., Pesonen, H., Corander, J., Vehtari, A., Kaski, S.,
+Bürkner, P.-C., \& Klami, A. (2023). Prior {Knowledge Elicitation}: {The
+Past}, {Present}, and {Future}. \emph{Bayesian Analysis}, \emph{-1}(-1).
+\url{https://doi.org/10.1214/23-BA1381}
+
 \bibitem[\citeproctext]{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}
@@
 
+\bibitem[\citeproctext]{ref-schadPrincipledBayesianWorkflow2021}
+Schad, D. J., Betancourt, M., \& Vasishth, S. (2021). Toward a
+principled {Bayesian} workflow in cognitive science. \emph{Psychological
+Methods}, \emph{26}(1), 103--126.
+\url{https://doi.org/10.1037/met0000275}
+
 \bibitem[\citeproctext]{ref-schadHowCapitalizePriori2020}
@@
 \url{https://doi.org/10.1016/j.jml.2019.104038}
+
+\bibitem[\citeproctext]{ref-stefanPracticalChallengesMethodological2022}
+Stefan, A. M., Evans, N. J., \& Wagenmakers, E.-J. (2022). Practical
+challenges and methodological flexibility in prior elicitation.
+\emph{Psychological Methods}, \emph{27}(2), 177--197.
+\url{https://doi.org/10.1037/met0000354}
 

### 2024-11-05 _manuscript/_tex/manuscript.tex
minor updates based on proof-reading; include citation notice; reformat precision results figure

@@
 
-\par{ This is version 3.5 of our preprint published at
+\par{ This is version 4 of our preprint published at
 \url{https://doi.org/10.31234/osf.io/rpjem}. All materials (reproducible
@@
 \newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
+\definecolor{quarto-callout-color}{HTML}{909090}
+\definecolor{quarto-callout-note-color}{HTML}{0758E5}
+\definecolor{quarto-callout-important-color}{HTML}{CC1914}
+\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
+\definecolor{quarto-callout-tip-color}{HTML}{00A047}
+\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
+\definecolor{quarto-callout-color-frame}{HTML}{acacac}
+\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
+\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
+\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
+\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
+\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
+\makeatother
 \makeatletter
@@
 \setlength\LTleft{0pt}
+
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, arc=.35mm, colframe=quarto-callout-important-color-frame, toprule=.15mm, colback=white, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, bottomtitle=1mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-important-color!10!white, left=2mm, opacityback=0, rightrule=.15mm, breakable, coltitle=black, toptitle=1mm, titlerule=0mm, bottomrule=.15mm]
+
+This manuscript has been accepted by the journal \emph{Advances in
+Methods and Practices in Psychological Science (AMPPS)}. Please cite the
+official version of record as:
+
+Pargent, F., K. Koch, T., Kleine, A.-K., Lermer, E., \& Gaube, S.
+(2024). A tutorial on tailored simulation-based sample-size planning for
+experimental designs with generalized linear mixed models.
+\emph{Advances in Methods and Practices in Psychological Science}, 7(4).
+\url{https://doi.org/10.1177/25152459241287132}
+
+\end{tcolorbox}
 
@@
 \citeproc{ref-lanePowerStrugglesEstimating2018}{Lane \& Hennes, 2018}).
-As a byproduct, the simulated datasets can also be used to test whether
+As a by-product, the simulated datasets can also be used to test whether
 the intended data analysis provides the expected (unbiased) results,
@@
 \emph{OR} \(H_{02}\) \emph{OR} both are rejected. If the global
-hypothesis \(H_0\) is combined with \emph{OR}, the p-values of the
-single hypotheses must be corrected for multiple testing to avoid
+hypothesis \(H_0\) is combined with a logical \emph{OR}, the p-values of
+the single hypotheses must be corrected for multiple testing to avoid
 \(\alpha\)-inflation for the global hypothesis
@@
 \& D'Agostino, 2013}). However, if the global hypothesis \(H_0\) is
-combined with \emph{AND}, a correction for multiple testing is not
-necessary but rather a mistake that unnecessarily reduces the power of
-the global HT. None of the software packages for sample size planning in
-Table 1 can handle combined hypotheses and only some can handle directed
-hypotheses. In contrast, our case study will demonstrate how we can test
-directed combined hypotheses with tailored simulation-based sample size
-planning.
+combined with a logical \emph{AND}, a correction for multiple testing is
+not necessary but rather a mistake that unnecessarily reduces the power
+of the global HT. None of the software packages for sample size planning
+in Table 1 can handle combined hypotheses and only some can handle
+directed hypotheses. In contrast, our case study will demonstrate how we
+can test directed combined hypotheses with tailored simulation-based
+sample size planning.
 
@@
 \] with the inverse-logit link
-\(g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}\)
+\(g^{-1}(\eta_{si}) = \text{inverse\_logit}(\eta_{si}) = \frac{\text{exp}(\eta_{si})}{1 + \text{exp}(\eta_{si})}\)
 or equivalently \[
@@
 \begin{aligned}
-\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i}) + \\
-&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
+\text{logit}[P_{si}] =\ (&\beta_0 + u_{0s} + u_{0i})\\
+&+ \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s\\
+&+\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}
 \end{aligned}
@@
 \begin{aligned}
-P_{si} = \text{inverse\_logit}[&(\beta_0 + u_{0s} + u_{0i}) + \\
-&\beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+P_{si} = \text{inverse\_logit}[&(\beta_0 + u_{0s} + u_{0i})\\
+&+ \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s\\
+&+ \beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
 \end{aligned}
@@
 \begin{aligned}
-P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0}) = \\
-= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} + \beta_e \cdot expert_s + \\
-&\beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
+P(Y=1|\mathbf{X=x}, \mathbf{U} = \mathbf{0})\\
+= \text{inverse\_logit}[&\beta_0 + \beta_a \cdot advice\_present_{si} + \beta_c \cdot advice\_correct_{si} +  \beta_e \cdot expert_s\\
+&+ \beta_{ea} \cdot expert_{s} \cdot advice\_present_{si} + \beta_{ec} \cdot expert_{s} \cdot advice\_correct_{si}]
 \end{aligned}
@@
 \begin{aligned}
-P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0) = \\
+P(Y=1| advice\_present = 0, advice\_correct = 0, expert = 0, u_{0s} = 0, u_{0i} = 0)\\
 = \text{inverse\_logit}[\beta_0]
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1646}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3291}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5063}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1576}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3152}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5273}}@{}}
 \caption{Table 2. Assumed probabilities that an average subject solves
@@
 \endlastfoot
-no advice, student & 0.70 & \(logit(0.70) = \beta_0\) \\
-no advice, expert & 0.90 & \(logit(0.90) = \beta_0 + \beta_e\) \\
+no advice, student & 0.70 & \(\text{logit}(0.70) = \beta_0\) \\
+no advice, expert & 0.90 & \(\text{logit}(0.90) = \beta_0 + \beta_e\) \\
 incorrect advice, student & 0.40 &
-\(logit(0.40) = \beta_0 + \beta_a\) \\
+\(\text{logit}(0.40) = \beta_0 + \beta_a\) \\
 incorrect advice, expert & 0.85 &
-\(logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \\
+\(\text{logit}(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \\
 correct advice, student & 0.90 &
-\(logit(0.90) = \beta_0 + \beta_a + \beta_c\) \\
+\(\text{logit}(0.90) = \beta_0 + \beta_a + \beta_c\) \\
 correct advice, expert & 0.95 &
-\(logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\) \\
+\(\text{logit}(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\) \\
 \end{longtable}
@@
 \begin{aligned}
-P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0) = \\
-= inverse\_logit[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
+P(Y=1|expert = 1, advice\_present = 1, advice\_correct = 1, u_{0s} = 0, u_{0i} = 0)\\
+= \text{inverse\_logit}[\beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}]
 \end{aligned}
@@
 presented with correct advice (the conditional perspective). However,
-even if the model assumptions were true, we would not observe that 95\%
-of experts responding to items presented with correct advice from a big
-sample of subjects drawn from their natural distribution of ability and
-items drawn from their natural distribution of difficulty (the marginal
-perspective). How much the two probabilities differ depends on the
-standard deviations of the random intercepts (the two probabilities are
-only equal if both standard deviations are be zero). We want to use the
-model-implied observed proportion of correct diagnoses in each
-experimental condition as an insightful descriptive statistic to
-determine plausible values for the random effect standard deviations. We
-will simulate a large dataset (for which the observed values of the
-descriptive statistic will be close to their model-implied true values)
-and simply compute the relative frequency of correct diagnoses for each
-experimental condition.
+even if the model assumptions were true, the relative frequency of
+experts that solve items presented with correct advice would not be
+\(0.95\) in a large sample of subjects drawn from their natural
+distribution of ability and items drawn from their natural distribution
+of difficulty (the marginal perspective). How much the two probabilities
+differ depends on the standard deviations of the random intercepts (the
+two probabilities are only equal if both standard deviations are be
+zero). We want to use the model-implied observed proportion of correct
+diagnoses in each experimental condition as an insightful descriptive
+statistic to determine plausible values for the random effect standard
+deviations. We will simulate a large dataset (for which the observed
+values of the descriptive statistic will be close to their model-implied
+true values) and simply compute the relative frequency of correct
+diagnoses for each experimental condition.
 
@@
     \AttributeTok{color =} \StringTok{"white"}\NormalTok{, }\AttributeTok{size =} \DecValTok{4}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\AttributeTok{direction =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{guides}\NormalTok{(}\AttributeTok{fill =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{reverse=}\ConstantTok{FALSE}\NormalTok{))}
+  \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.2}\NormalTok{), }\AttributeTok{direction =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{fill =} \FunctionTok{guide\_colourbar}\NormalTok{(}\AttributeTok{reverse =} \ConstantTok{TRUE}\NormalTok{))}
 \end{Highlighting}

### 2024-11-05 _manuscript/manuscript.qmd
minor updates based on proof-reading; include citation notice; reformat precision results figure

--- 2024-07-31 manuscript.qmd
@@
 date: 2024-07-01
+date-modified: 2024-11-05
 date-format: iso
@@
   disclosures:
-    data-sharing: This is version 3.5 of our preprint published at <https://doi.org/10.31234/osf.io/rpjem>.
+    data-sharing: This is version 4 of our preprint published at <https://doi.org/10.31234/osf.io/rpjem>.
 All materials (reproducible manuscript, including R code and all simulation results) are available in the project's repository on the Open Science Framework (OSF) at <https://osf.io/dhwf4/>.
@@
 
+:::{.callout-important}
+This manuscript has been accepted by the journal *Advances in Methods and Practices in Psychological Science (AMPPS)*.
+Please cite the official version of record as:
+
+Pargent, F., K.
+Koch, T., Kleine, A.-K., Lermer, E., & Gaube, S.
+(2024).
+A tutorial on tailored simulation-based sample-size planning for experimental designs with generalized linear mixed models.
+*Advances in Methods and Practices in Psychological Science*, 7(4).
+<https://doi.org/10.1177/25152459241287132>
+:::
+
 # Introduction
@@
 Tailored simulation-based approaches offer the possibility to include the assumed process of how data become missing in the data simulation, thereby determining the required sample size based on simulated datasets that contain missing values [for example, see @lanePowerStrugglesEstimating2018].
-As a byproduct, the simulated datasets can also be used to test whether the intended data analysis provides the expected (unbiased) results, despite the missing data.
+As a by-product, the simulated datasets can also be used to test whether the intended data analysis provides the expected (unbiased) results, despite the missing data.
 Although GLMMs can handle a large variety of outcome variables, researchers are becoming increasingly aware that many datasets might profit from even more sophisticated models.
@@
 For other research questions, the combined null hypothesis $H_0$ would be rejected if $H_{01}$ *OR* $H_{02}$ *OR* both are rejected.
-If the global hypothesis $H_0$ is combined with *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis [@dmitrienkoTraditionalMultiplicityAdjustment2013].
-However, if the global hypothesis $H_0$ is combined with *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
+If the global hypothesis $H_0$ is combined with a logical *OR*, the p-values of the single hypotheses must be corrected for multiple testing to avoid $\alpha$-inflation for the global hypothesis [@dmitrienkoTraditionalMultiplicityAdjustment2013].
+However, if the global hypothesis $H_0$ is combined with a logical *AND*, a correction for multiple testing is not necessary but rather a mistake that unnecessarily reduces the power of the global HT.
 None of the software packages for sample size planning in Table 1 can handle combined hypotheses and only some can handle directed hypotheses.
@@
 $$
-with the inverse-logit link $g^{-1}(\eta_{si}) = inverse\_logit(\eta_{si}) = \frac{exp(\eta_{si})}{1 + exp(\eta_{si})}$ or equivalently
+with the inverse-logit link $g^{-1}(\eta_{si}) = \text{inverse\_logit}(\eta_{si}) = \frac{\text{exp}(\eta_{si})}{1 + \text{exp}(\eta_{si})}$ or equivalently
 $$
@@
   col3 = 
-    c("$logit(0.70) = \\beta_0$",
-      "$logit(0.90) = \\beta_0 + \\beta_e$", 
-      "$logit(0.40) = \\beta_0 + \\beta_a$",
-      "$logit(0.85) = \\beta_0 + \\beta_e + \\beta_{a} + \\beta_{ea}$",
-      "$logit(0.90) = \\beta_0 + \\beta_a + \\beta_c$",
-      "$logit(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
+    c("$\\text{logit}(0.70) = \\beta_0$",
+      "$\\text{logit}(0.90) = \\beta_0 + \\beta_e$", 
+      "$\\text{logit}(0.40) = \\beta_0 + \\beta_a$",
+      "$\\text{logit}(0.85) = \\beta_0 + \\beta_e + \\beta_{a} + \\beta_{ea}$",
+      "$\\text{logit}(0.90) = \\beta_0 + \\beta_a + \\beta_c$",
+      "$\\text{logit}(0.95) = \\beta_0 + \\beta_e + \\beta_a + \\beta_c + \\beta_{ea} + \\beta_{ec}$"))
 knitr::kable(probtable, col.names = c("Experimental condition",
@@
 For example, we determined the $\beta$ parameters in a way that corresponds to a desired conditional probability of $0.95$, that an expert with average ability solves an item with average difficulty when presented with correct advice (the conditional perspective).
-However, even if the model assumptions were true, we would not observe that 95% of experts responding to items presented with correct advice from a big sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty (the marginal perspective).
+However, even if the model assumptions were true, the relative frequency of experts that solve items presented with correct advice would not be $0.95$ in a large sample of subjects drawn from their natural distribution of ability and items drawn from their natural distribution of difficulty (the marginal perspective).
 How much the two probabilities differ depends on the standard deviations of the random intercepts (the two probabilities are only equal if both standard deviations are be zero).
@@
     color = "white", size = 4) +
-  scale_fill_viridis_c(limits = c(0, 0.3), direction = -1) +
-  guides(fill = guide_legend(reverse=FALSE))
+  scale_fill_viridis_c(limits = c(0, 0.2), direction = -1) +
+  guides(fill = guide_colourbar(reverse = TRUE))
 ```

### 2024-11-05 _tex/manuscript.tex
Built site for gh-pages

@@
 
-\par{ This is version 2.5 of our preprint published at
+\par{ This is version 4 of our preprint published at
 \url{https://doi.org/10.31234/osf.io/rpjem}. All materials (reproducible
@@
 
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomrule=.15mm, bottomtitle=1mm, opacityback=0, colbacktitle=quarto-callout-important-color!10!white, coltitle=black, left=2mm, colback=white, breakable, toptitle=1mm, titlerule=0mm, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, arc=.35mm, rightrule=.15mm, toprule=.15mm, leftrule=.75mm, colframe=quarto-callout-important-color-frame]
+
+This manuscript has been accepted by the journal \emph{Advances in
+Methods and Practices in Psychological Science (AMPPS)}. Please cite the
+official version of record as:
+
+Pargent, F., K. Koch, T., Kleine, A.-K., Lermer, E., \& Gaube, S.
+(2024). A tutorial on tailored simulation-based sample-size planning for
+experimental designs with generalized linear mixed models.
+\emph{Advances in Methods and Practices in Psychological Science}, 7(4).
+\url{https://doi.org/10.1177/25152459241287132}
+
+\end{tcolorbox}
+
 \section{Introduction}\label{introduction}
@@
 To our knowledge, existing approaches for sample size planning for GLMMs
-have exclusively focused on planning for power. In
-Table~\ref{tbl-rpackages}, we review available software packages that
-can be used to perform power analysis for multilevel models in R
-{[}Version 4.4.0; R Core Team (\citeproc{ref-R-base}{2024}){]}. Power
-analysis methods can be categorized into formula-based,
-summary-statistics-based and simulation-based methods
+have exclusively focused on planning for power. In Table 1, we review
+available software packages that can be used to perform power analysis
+for multilevel models in R {[}Version 4.3.3; R Core Team
+(\citeproc{ref-R-base}{2024}){]}. Power analysis methods can be
+categorized into formula-based, summary-statistics-based and
+simulation-based methods
 (\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{Murayama
@@
 \citeproc{ref-debruineUnderstandingMixedEffectsModels2021}{DeBruine \&
-Barr, 2021}; \citeproc{ref-greenSIMRPackagePower2016a}{Green \& MacLeod,
-2016a}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al., 2015};
-\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021};
-\citeproc{ref-lafitSelectionNumberParticipants2021}{Lafit et al., 2021};
-\citeproc{ref-zimmerSampleSizePlanning2022}{Zimmer et al., 2022}).
-However, many of these tutorials focus on LMMs and the most common study
-designs (see \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et
-al., 2021} for a tutorial that also covers more advanced settings). This
-narrow focus provides limited guidance for researchers using more
-complex study designs, especially when little prior knowledge about
-plausible effect sizes is available (see the discussion in
+Barr, 2021}; \citeproc{ref-greenSIMRPackagePower2016}{Green \& MacLeod,
+2016}; \citeproc{ref-johnsonPowerAnalysisGeneralized2015}{Johnson et
+al., 2015}; \citeproc{ref-kainPracticalGuidePower2015}{Kain et al.,
+2015}; \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al.,
+2021}; \citeproc{ref-lafitSelectionNumberParticipants2021}{Lafit et al.,
+2021}; \citeproc{ref-zimmerSampleSizePlanning2023}{Zimmer et al.,
+2023}). However, many of these tutorials focus on LMMs and the most
+common study designs (see
+\citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}
+for a tutorial that also covers more advanced settings). This narrow
+focus provides limited guidance for researchers using more complex study
+designs, especially when little prior knowledge about plausible effect
+sizes is available (see the discussion in
 \citeproc{ref-kumleEstimatingPowerGeneralized2021}{Kumle et al., 2021}).
@@
 
-\begin{table}
-
-{\caption{{Overview of software packages for power analysis for
-(generalized) linear mixed models in R.}{\label{tbl-rpackages}}}
-\vspace{-20pt}}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1165}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.3947}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0526}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0489}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2030}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1842}}@{}}
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Package name
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Summary
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-GUI available
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-GLMM support
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Design limitations
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Reference
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-glmmrBase & Formula- and simulation-based power analysis; non-linear
-fixed effects and flexible covariance functions & no & yes & none &
-Watson (\citeproc{ref-watsonGeneralisedLinearMixed2023}{2023}) \\
-longpower & Formula-based power analysis; focus on longitudinal data &
-\href{https://atrihub.shinyapps.io/power/}{yes} & no & limited set of
-study designs with two levels & Iddi and C Donohue
-(\citeproc{ref-iddiPowerSampleSize2022}{2022}) \\
-mixedpower & Simulation-based power analysis; based on `simr' package &
-no & yes & none & Kumle et al.
-(\citeproc{ref-kumleEstimatingPowerGeneralized2021}{2021}) \\
-mlmpower & Simulation-based power analysis; missing data mechanisms & no
-& no & limited to two levels & Enders et al.
-(\citeproc{ref-endersSimpleMonteCarlo2023}{2023}) \\
-pamm & Simulation-based power analysis; based on `lme4' package & no &
-no & none & Martin et al.
-(\citeproc{ref-martinMeasuringIndividualDifferences2011}{2011}) \\
-pass.lme & Formula-based power analysis; limited documentation & no & no
-& none & Marco Chak Yan YU
-(\citeproc{ref-marcochakyanyuPassLmePower2019}{2019}) \\
-PowerAnalysisIL & Simulation-based power analysis; focus on longitudinal
-data &
-\href{https://github.com/ginettelafit/PowerAnalysisIL?tab=readme-ov-file\#shiny-app-and-r-package-to-perform-power-analysis-to-select-the-number-of-participants-in-intensive-longitudinal-studies}{yes}
-& no & limited set of study designs with two levels & Lafit et al.
-(\citeproc{ref-lafitSelectionNumberParticipants2021}{2021}) \\
-powerlmm & Simulation-based power analysis; missing data mechanisms;
-focus on longitudinal data &
-\href{https://github.com/rpsychologist/powerlmm?tab=readme-ov-file\#launch-interactive-web-application}{yes}
-& no & limited set of study designs with two or three levels & Magnusson
-et al.
-(\citeproc{ref-magnussonConsequencesIgnoringTherapist2018}{2018}) \\
-simglm & Simulation-based power analysis; missing data mechanisms &
-\href{https://simglm.brandonlebeau.org/reference/run_shiny.html}{yes} &
-yes & limited to two or three levels & LeBeau
-(\citeproc{ref-lebeauSimglmSimulateModels2017}{2017}) \\
-simr & Simulation-based power analysis; based on `lme4' package & no &
-yes & none & Green and MacLeod
-(\citeproc{ref-greenSIMRPackagePower2016}{2016b}) \\
-summary\_statistics based\_power & Summary-statistics-based power
-analysis; based on `pwr' package &
-\href{https://koumurayama.shinyapps.io/summary_statistics_based_power/}{yes}
-& yes & limited set of study designs with two levels & Murayama et al.
-(\citeproc{ref-murayamaSummarystatisticsbasedPowerAnalysis2022}{2022}) \\
-WebPower & Formula-based power analysis; based on `longpower' package;
-focus on longitudinal data &
-\href{https://webpower.psychstat.org/wiki/start}{yes} & no & limited set
-of study designs with two levels & Zhang and Yuan
-(\citeproc{ref-zhangPracticalStatisticalPower2018}{2018}) \\
+\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}
+
+\global\setlength{\Oldtabcolsep}{\tabcolsep}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\begin{longtable}[c]{|p{1.18in}|p{1.57in}|p{0.59in}|p{0.59in}|p{1.18in}|p{0.79in}}
+\caption{Table 1. Overview of software packages for power analysis for
+(generalized) linear mixed models in R.}\tabularnewline
+
+\ascline{0.75pt}{000000}{1-6}
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Package\ name}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Summary}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{GUI\ available}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{GLMM\ support}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Design\ limitations}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Reference}}}} \\
+
+\ascline{0.75pt}{000000}{1-6}\endfirsthead
+
+\ascline{0.75pt}{000000}{1-6}
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Package\ name}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Summary}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{GUI\ available}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{GLMM\ support}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Design\ limitations}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Reference}}}} \\
+
+\ascline{0.75pt}{000000}{1-6}\endhead
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{glmmrBase}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{and}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{non-linear}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{fixed}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{effects}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{and}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{flexible}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{covariance}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{functions}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Watson}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2023)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://atrihub.shinyapps.io/power/}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Iddi}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\&}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Donohue}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2022)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mixedpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{lme4}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Kumle}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2021)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mlmpower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{missing}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mechanisms}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{to}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Enders}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2023)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{pamm}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{lme4}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Martin}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2011)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{pass.lme}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{documentation}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Yu}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2019)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{PowerAnalysisIL}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://github.com/ginettelafit/PowerAnalysisIL?tab=readme-ov-file\#shiny-app-and-r-package-to-perform-power-analysis-to-select-the-number-of-participants-in-intensive-longitudinal-studies}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Lafit}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2021)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{powerlmm}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{missing}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mechanisms;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://github.com/rpsychologist/powerlmm?tab=readme-ov-file\#launch-interactive-web-application}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{or}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{three}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Magnusson}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2018)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{simglm}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{missing}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{mechanisms}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://simglm.brandonlebeau.org/reference/run\_shiny.html}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{to}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{or}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{three}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{LeBeau}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2019)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{simr}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Simulation-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{lme4}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{none}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Green}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\&}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{MacLeod}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2016)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{summary\_statistics}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based\_power}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Summary-statistics-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{pwr}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://koumurayama.shinyapps.io/summary\_statistics\_based\_power/}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{yes}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Murayama}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{et}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{al.}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2022)}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{WebPower}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.57in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Formula-based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{power}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{analysis;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{based}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{‘}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longpower}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{’}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{package;}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{focus}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{on}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{longitudinal}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{data}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{0000FF}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\href{https://webpower.psychstat.org/wiki/start}{yes}}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.59in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{no}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{limited}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{set}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{of}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{study}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{designs}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{with}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{two}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{levels}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.79in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Zhang}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\&}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{Yuan}}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{\ }}}\textcolor[HTML]{000000}{\fontsize{10}{20}\selectfont{\global\setmainfont{Times New Roman}{(2018)}}}} \\
+
+\ascline{0.75pt}{000000}{1-6}
+
 \end{longtable}
 
-{\vspace{-20pt}
-\noindent \emph{Note.} GUI available = whether a graphical user interface is provided (usually a Shiny app that is available online or can be downloaded to run locally), GLMM support = whether power analysis can be performed for generalized linear mixed model (otherwise only linear mixed models are supported).
-Design limitations = whether only a limited selection of study designs can be specified (otherwise arbitrary designs within the model class are supported).}
-
-\end{table}
+\arrayrulecolor[HTML]{000000}
+
+\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}
+
+\global\setlength{\tabcolsep}{\Oldtabcolsep}
+
+\renewcommand*{\arraystretch}{1}
+
+\noindent \emph{Note.} GUI available = whether a graphical user
+interface is provided (usually a Shiny app that is available online or
+can be downloaded to run locally), GLMM support = whether power analysis
+can be performed for generalized linear mixed model (otherwise only
+linear mixed models are supported), Design limitations = whether only a
+limited selection of study designs can be specified (otherwise arbitrary
+designs within the model class are supported).
 
@@
 complicated and time-consuming than using the existing software tools
-outlined in Table~\ref{tbl-rpackages}. Circumstances under which
-tailored simulation-based sample size planning is necessary include 1)
-complex study designs, 2) complex statistical hypotheses, 3) planning
-for precision, and 4) lack of prior studies or pilot data.
+outlined in Table 1. Circumstances under which tailored simulation-based
+sample size planning is necessary include 1) complex study designs, 2)
+complex statistical hypotheses, 3) planning for precision, and 4) lack
+of prior studies or pilot data.
 
@@
 \& D'Agostino, 2013}). However, if the global hypothesis \(H_0\) is
-combined with \emph{AND}, a correction for multiple testing is not
-necessary but rather a mistake that unnecessarily reduces the power of
-the global HT. None of the software packages for sample size planning in
-Table~\ref{tbl-rpackages} can handle combined hypotheses and only some
-can handle directed hypotheses. In contrast, our case study will
-demonstrate how we can test directed combined hypotheses with tailored
-simulation-based sample size planning.
+combined with a logical \emph{AND}, a correction for multiple testing is
+not necessary but rather a mistake that unnecessarily reduces the power
+of the global HT. None of the software packages for sample size planning
+in Table 1 can handle combined hypotheses and only some can handle
+directed hypotheses. In contrast, our case study will demonstrate how we
+can test directed combined hypotheses with tailored simulation-based
+sample size planning.
 
@@
 (\citeproc{ref-cummingNewStatisticsWhy2014}{Cumming, 2014}), currently
-cannot use the software packages outlined in Table~\ref{tbl-rpackages}.
-However, tailored simulation-based sample size planning can easily
-handle the planning for precision approach
-(\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell et al., 2008}).
-The only change in procedure is that instead of computing HTs for each
-simulated dataset and estimating statistical power across repetitions,
-CIs are computed for each simulated dataset, and their expected width is
-estimated.
+cannot use the software packages outlined in Table 1. However, tailored
+simulation-based sample size planning can easily handle the planning for
+precision approach (\citeproc{ref-maxwellSampleSizePlanning2008}{Maxwell
+et al., 2008}). The only change in procedure is that instead of
+computing HTs for each simulated dataset and estimating statistical
+power across repetitions, CIs are computed for each simulated dataset,
+and their expected width is estimated.
 
@@
 quality of the results of the sample size planning crucially depends on
-the plausibility of the simulated data-generating process. However, we
-argue that even a strongly simplified data-generating process (e.g.,
-only a small number of interaction effects; only random intercepts, and
-no random slopes; assuming that data is missing completely at random)
-can yield informative results.
-
-\paragraph{PRACTICE.}\label{practice-1}
+the plausibility of the simulated data-generating process. If the
+simulated data generating process is less complex than the true process,
+one can expect that sample size planning underestimates the required
+sample size and the planned analysis has inflated type I error rates
+(\citeproc{ref-matuschekBalancingTypeError2017}{Matuschek et al.,
+2017}). However, Matuschek et al.
+(\citeproc{ref-matuschekBalancingTypeError2017}{2017}) have also shown
+in simulations with LMMs that while fitting \emph{maximal models}
+(i.e.~that include all possible random effects) safeguards against
+inflated type I error rates, this can lead to a great loss in power if
+the variance in the random effects is actually small. Thus, we argue
+that even a simplified data-generating process (e.g., only a small
+number of interaction effects; only random intercepts, and no random
+slopes; assuming that data is missing completely at random) that is only
+plausible under idealized circumstances, can yield informative results
+and is preferred over performing no systematic sample size planning.
+
+\subsubsection{PRACTICE}\label{practice-1}
 
@@
 
-\begin{table}
-
-{\caption{{Assumed probabilities that an average subject solves an
-average item in each experimental condition.}{\label{tbl-probtable}}}
-\vspace{-20pt}}
+\newpage
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1646}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3291}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5063}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1576}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3152}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5273}}@{}}
+\caption{Table 2. Assumed probabilities that an average subject solves
+an average item in each experimental condition.}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Experimental condition
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(P(Y=1 \mid \mathbf{X=x}, \mathbf{U} = \mathbf{0})\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Implied equation
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
 \toprule\noalign{}
@@
 \endlastfoot
-no advice, student & 0.70 & \(logit(0.70) = \beta_0\) \\
-no advice, expert & 0.90 & \(logit(0.90) = \beta_0 + \beta_e\) \\
+no advice, student & 0.70 & \(\text{logit}(0.70) = \beta_0\) \\
+no advice, expert & 0.90 & \(\text{logit}(0.90) = \beta_0 + \beta_e\) \\
 incorrect advice, student & 0.40 &
-\(logit(0.40) = \beta_0 + \beta_a\) \\
+\(\text{logit}(0.40) = \beta_0 + \beta_a\) \\
 incorrect advice, expert & 0.85 &
-\(logit(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \\
+\(\text{logit}(0.85) = \beta_0 + \beta_e + \beta_{a} + \beta_{ea}\) \\
 correct advice, student & 0.90 &
-\(logit(0.90) = \beta_0 + \beta_a + \beta_c\) \\
+\(\text{logit}(0.90) = \beta_0 + \beta_a + \beta_c\) \\
 correct advice, expert & 0.95 &
-\(logit(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\) \\
+\(\text{logit}(0.95) = \beta_0 + \beta_e + \beta_a + \beta_c + \beta_{ea} + \beta_{ec}\) \\
 \end{longtable}
 
-{\vspace{-20pt}
-\noindent \emph{Note.} Implied equations are derived based on the model equations and setting all random intercept terms to 0.}
-
-\end{table}
-
-Table~\ref{tbl-probtable} shows our set of assumptions concerning the
-probability that an average subject solves an average item for each
-experimental condition, as well as the corresponding equations implied
-by the model. The table can be used to compute the implied values for
-the \(\beta\) parameters, starting with the first equation and
-reinserting the computed \(\beta\) values in all following equations
-(\texttt{b\_0} stands for the intercept \(\beta_0\), \texttt{b\_e} for
-the slope \(\beta_e\), and so on):
+Table 2 shows our set of assumptions concerning the probability that an
+average subject solves an average item for each experimental condition,
+as well as the corresponding equations implied by the model. The table
+can be used to compute the implied values for the \(\beta\) parameters,
+starting with the first equation and reinserting the computed \(\beta\)
+values in all following equations (\texttt{b\_0} stands for the
+intercept \(\beta_0\), \texttt{b\_e} for the slope \(\beta_e\), and so
+on):
 
@@
 \subsubsection{Iterate with domain
-experts}\label{iterate-with-domain-experts}
-
-\paragraph{THEORY.}\label{theory-5}
-
-Gathering domain knowledge and the specification of population parameter
-values is an iterative process, in which domain experts repeatedly
-assess the plausibility of simulated data. As a first step, domain
-experts can be interviewed to ``elicit'' their domain knowledge about
-how the data of the planned study is expected to look. As most domain
-experts are no experts in statistical modeling and GLMMs, they often
-struggle without further guidance to communicate their knowledge in a
-way that is useful when specifying the parameters for data simulation.
-For this reason, we suggest that after an initial interview, the analyst
-who is familiar with the structure of the GLMM selects an initial set of
-insightful descriptive statistics and model-based quantities. Then they
-reenter into an iterative discussion with the domain experts where some
-set of population values are selected, and the plausibility of resulting
-implied quantities is evaluated. Then, the population parameters are
-updated until the domain experts are satisfied with the result. During
-this process, the monitored descriptive statistics and model-based
-quantities can be updated or extended to capture as much available
-domain knowledge as possible.
-
-\subparagraph{PRACTICE.}\label{practice-5}
+expertise}\label{iterate-with-domain-expertise}
+
+\subsubsection{THEORY}\label{theory-5}
+
+Tailored simulation-based sample size planning always requires access to
+domain knowledge. However, gathering domain knowledge and the
+specification of population parameter values is not always
+straightforward and can be better described as an iterative process, in
+which the plausibility of simulated data is repeatedly compared with the
+available domain expertise. As a first step, dedicated domain experts
+(if available) can be interviewed to \emph{elicit} their domain
+knowledge about how the data of the planned study is expected to look.
+Because most domain experts are no experts in statistical modeling and
+GLMMs, they often struggle without further guidance to communicate their
+knowledge in a way that is useful when specifying the parameters for
+data simulation. For this reason, we suggest that after an initial
+interview, the analyst who is familiar with the structure of the GLMM
+selects an initial set of insightful descriptive statistics and
+model-based quantities. Then, they reenter into an iterative discussion
+with the domain experts where some set of population values are
+selected, and the plausibility of resulting implied quantities is
+evaluated. The population parameters are updated until the domain
+experts are satisfied with the result. During this process, the
+monitored descriptive statistics and model-based quantities can be
+updated or extended to capture as much available domain knowledge as
+possible. Even if no dedicated domain experts are available, the basic
+principles stay the same except that initial parameter values are often
+derived from a more extensive literature review. Descriptive statistics
+and model-based quantities are still used to evaluate whether the
+implied assumptions about the data generating process are plausible.
+Often, this judgment evolves over several iterations as the analyst
+becomes more familiar with the model and the study design.\footnote{Developing
+  more effective strategies on how to elicit and visualize domain
+  knowledge is currently an active area of research, primarily in the
+  context of Bayesian modeling
+  (\citeproc{ref-bocktingSimulationbasedPriorKnowledge2024}{Bockting et
+  al., 2024};
+  \citeproc{ref-hartmannFlexiblePriorElicitation2020}{Hartmann et al.,
+  2020}; \citeproc{ref-mikkolaPriorKnowledgeElicitation2023}{Mikkola et
+  al., 2023};
+  \citeproc{ref-stefanPracticalChallengesMethodological2022}{Stefan et
+  al., 2022}). We expect that future advances in this field will also be
+  highly relevant for tailored simulation-based sample size planning
+  with frequentist models.}
+
+\subsubsection{PRACTICE}\label{practice-5}
 
@@
 detect an effect of the expected size implied by our data simulation.
-This could be achieved by collecting data from 200 subjects (about 25\%
+This could be achieved by collecting data from 250 subjects (about 25\%
 of which will be experts), each completing the same 50 items (with
 advice present in about 67\% of cases, which is correct in about 80\% of
-cases with present advice). If collecting data from 200 subjects is not
-feasible, an alternative would be to recruit 150 subjects but increase
-the length of the experiment to over 70 items. However, 70 items might
-take too long to complete for the radiologists participating in the
-study, who have a busy schedule. The simulation suggests that it might
-also be possible to plan a shorter experiment with only 30 items if it
-is feasible to recruit an even higher number of subjects (\textgreater{}
-250, to be determined by additional rounds of power analysis). Design
-parameters that also affect power, and which could be investigated in
-the simulation to find a more optimal trade-off, are the ratio of
-experts to students, the frequency of whether advice is presented at
-all, and whether it is correct or incorrect.
-
-\paragraph{Precision results.}\label{precision-results}
-
-When planning for precision, one could monitor the width of all four CIs
-at the same time. However, because the CIs of the four contrasts
-strongly differ in width, it is not trivial to decide which width one
-should target when deciding on the appropriate sample size. In contrast
-to planning for power, there are no common standards on how to specify
-the targeted precision
+cases with present advice). If collecting data from 250 subjects is not
+feasible, an alternative would be to recruit 200 subjects but increase
+the length of the experiment to 70 items. However, 70 items might take
+too long to complete for the radiologists participating in the study,
+who have a busy schedule. The simulation suggests that it might also be
+possible to plan a shorter experiment with only 30 items if it is
+feasible to recruit an even higher number of subjects (to be determined
+by additional rounds of power analysis). Design parameters that also
+affect power, and which could be investigated in the simulation to find
+a more optimal trade-off, are the ratio of experts to students, the
+frequency of whether advice is presented at all, and whether it is
+correct or incorrect.\footnote{An advanced alternative to our heuristic
+  trade-off between subjects and items would be to explicitly optimize
+  for cost-efficient study designs using the \emph{mlpwr} package
+  (\citeproc{ref-zimmerSampleSizePlanning2023}{Zimmer et al., 2023}).
+  The package performs simulation-based power analysis in a surrogate
+  modeling framework that can evaluate power for a large grid of design
+  combinations more efficiently. The framework also requires users to
+  write their own \texttt{sim\_and\_analyse} function and can be
+  considered a great extension on what we cover in our tutorial.}
+
+\textbf{Precision results.} When planning for precision, one could
+monitor the width of all four CIs at the same time. However, because the
+CIs of the four contrasts strongly differ in width, it is not trivial to
+decide which width one should target when deciding on the appropriate
+sample size. In contrast to planning for power, there are no common
+standards on how to specify the targeted precision
 (\citeproc{ref-lakensSampleSizeJustification2022}{Lakens, 2022a}). For
@@
 
+\bibitem[\citeproctext]{ref-ebersoleManyLabsTesting2020}
+Ebersole, C. R., Mathur, M. B., Baranski, E., Bart-Plange, D.-J.,
+Buttrick, N. R., Chartier, C. R., Corker, K. S., Corley, M., Hartshorne,
+J. K., IJzerman, H., Lazarević, L. B., Rabagliati, H., Ropovik, I.,
+Aczel, B., Aeschbach, L. F., Andrighetto, L., Arnal, J. D., Arrow, H.,
+Babincak, P., \ldots{} Nosek, B. A. (2020). Many {Labs} 5: {Testing
+Pre-Data-Collection Peer Review} as an {Intervention} to {Increase
+Replicability}. \emph{Advances in Methods and Practices in Psychological
+Science}, \emph{3}(3), 309--331.
+\url{https://doi.org/10.1177/2515245920958687}
+
 \bibitem[\citeproctext]{ref-endersSimpleMonteCarlo2023}
@@
 
+\bibitem[\citeproctext]{ref-hartmannFlexiblePriorElicitation2020}
+Hartmann, M., Agiashvili, G., Bürkner, P., \& Klami, A. (2020). Flexible
+prior elicitation via the prior predictive distribution. In J. Peters \&
+D. Sontag (Eds.), \emph{Proceedings of the 36th conference on
+uncertainty in artificial intelligence ({UAI})} (Vol. 124, pp.
+1129--1138). PMLR.
+
 \bibitem[\citeproctext]{ref-iddiPowerSampleSize2022}
-Iddi, S., \& C Donohue, M. (2022). Power and {Sample Size} for
+Iddi, S., \& Donohue, M. C. (2022). Power and {Sample Size} for
 {Longitudinal Models} in {R} -- {The} longpower {Package} and {Shiny
@@
 \url{https://doi.org/10.32614/RJ-2022-022}
+
+\bibitem[\citeproctext]{ref-johnsonPowerAnalysisGeneralized2015}
+Johnson, P. C. D., Barry, S. J. E., Ferguson, H. M., \& Müller, P.
+(2015). Power analysis for generalized linear mixed models in ecology
+and evolution. \emph{Methods in Ecology and Evolution}, \emph{6}(2),
+133--142. \url{https://doi.org/10.1111/2041-210X.12306}
 
@@
 
-\bibitem[\citeproctext]{ref-lebeauSimglmSimulateModels2017}
-LeBeau, B. (2017). \emph{Simglm: {Simulate Models Based} on the
-{Generalized Linear Model}} (p. 0.8.9). Comprehensive R Archive Network.
-\url{https://doi.org/10.32614/CRAN.package.simglm}
+\bibitem[\citeproctext]{ref-lebeauPowerAnalysisSimulation2019}
+LeBeau, B. (2019). \emph{Power {Analysis} by {Simulation} using {R} and
+simglm}. \url{https://doi.org/10.17077/f7kk-6w7f}
 
@@
 \emph{86}(9), 711--725. \url{https://doi.org/10.1037/ccp0000333}
-
-\bibitem[\citeproctext]{ref-marcochakyanyuPassLmePower2019}
-Marco Chak Yan YU. (2019). \emph{Pass.lme: {Power} and {Sample Size} for
-{Linear Mixed Effect Models}} (p. 0.9.0). Comprehensive R Archive
-Network. \url{https://doi.org/10.32614/CRAN.package.pass.lme}
 
@@
 
+\bibitem[\citeproctext]{ref-stefanPracticalChallengesMethodological2022}
+Stefan, A. M., Evans, N. J., \& Wagenmakers, E.-J. (2022). Practical
+challenges and methodological flexibility in prior elicitation.
+\emph{Psychological Methods}, \emph{27}(2), 177--197.
+\url{https://doi.org/10.1037/met0000354}
+
 \bibitem[\citeproctext]{ref-uyguntuncEpistemicPragmaticFunction2023}
@@
 Vaughan, D., \& Dancho, M. (2022). \emph{Furrr: Apply mapping functions
-in parallel using futures}. \url{https://github.com/DavisVaughan/furrr}
+in parallel using futures}.
+\url{https://CRAN.R-project.org/package=furrr}
 
@@
 
-\bibitem[\citeproctext]{ref-zimmerSampleSizePlanning2022}
-Zimmer, F., Henninger, M., \& Debelak, R. (2022). \emph{Sample {Size
-Planning} for {Complex Study Designs}: {A Tutorial} for the mlpwr
-{Package}}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/r9w6t}
+\bibitem[\citeproctext]{ref-zimmerSampleSizePlanning2023}
+Zimmer, F., Henninger, M., \& Debelak, R. (2023). Sample size planning
+for complex study designs: {A} tutorial for the mlpwr package.
+\emph{Behavior Research Methods}.
+\url{https://doi.org/10.3758/s13428-023-02269-0}
 

### 2024-11-05 _manuscript/_tex/manuscript.tex
fix precision result figure labels

@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, arc=.35mm, colframe=quarto-callout-important-color-frame, toprule=.15mm, colback=white, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, bottomtitle=1mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-important-color!10!white, left=2mm, opacityback=0, rightrule=.15mm, breakable, coltitle=black, toptitle=1mm, titlerule=0mm, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, left=2mm, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, colframe=quarto-callout-important-color-frame, breakable, rightrule=.15mm, coltitle=black, titlerule=0mm, arc=.35mm, toprule=.15mm, bottomtitle=1mm, colbacktitle=quarto-callout-important-color!10!white, opacitybacktitle=0.6, bottomrule=.15mm, toptitle=1mm, opacityback=0, colback=white]
 
@@
 discuss a number of strategies on how to specify model parameters for
-application-specific, tailored data simulations in a later chapter.
+application-specific, tailored data simulations in a later section.
 
@@
   \FunctionTok{scale\_fill\_viridis\_c}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.2}\NormalTok{), }\AttributeTok{direction =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{guides}\NormalTok{(}\AttributeTok{fill =} \FunctionTok{guide\_colourbar}\NormalTok{(}\AttributeTok{reverse =} \ConstantTok{TRUE}\NormalTok{))}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{fill =} \FunctionTok{guide\_colourbar}\NormalTok{(}\AttributeTok{reverse =} \ConstantTok{TRUE}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"number of subjects"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{ylab}\NormalTok{(}\StringTok{"number of items"}\NormalTok{)}
 \end{Highlighting}

### 2024-11-05 manuscript.qmd
Built site for gh-pages
fix precision result figure labels

@@
 However, in practice, knowledge about those parameters from prior studies is often limited, which makes specifying assumptions a practical challenge (see the discussion in @maxwellSampleSizePlanning2008 and @kumleEstimatingPowerGeneralized2021).
-We will discuss a number of strategies on how to specify model parameters for application-specific, tailored data simulations in a later chapter.
+We will discuss a number of strategies on how to specify model parameters for application-specific, tailored data simulations in a later section.
 
@@
     color = "white", size = 4) +
-  scale_fill_viridis_c(limits = c(0, 0.3), direction = -1) +
-  guides(fill = guide_legend(reverse=FALSE))
+  scale_fill_viridis_c(limits = c(0, 0.2), direction = -1) +
+  guides(fill = guide_colourbar(reverse = TRUE)) +
+  xlab("number of subjects") + ylab("number of items")
 ```

### 2024-11-05 _tex/manuscript.tex
Built site for gh-pages

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomrule=.15mm, bottomtitle=1mm, opacityback=0, colbacktitle=quarto-callout-important-color!10!white, coltitle=black, left=2mm, colback=white, breakable, toptitle=1mm, titlerule=0mm, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, arc=.35mm, rightrule=.15mm, toprule=.15mm, leftrule=.75mm, colframe=quarto-callout-important-color-frame]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, coltitle=black, opacityback=0, breakable, toptitle=1mm, colbacktitle=quarto-callout-important-color!10!white, toprule=.15mm, arc=.35mm, colback=white, bottomrule=.15mm, left=2mm, colframe=quarto-callout-important-color-frame, rightrule=.15mm, opacitybacktitle=0.6, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, bottomtitle=1mm, titlerule=0mm]
 

### 2024-11-06 _manuscript/_tex/manuscript.tex
update citation metadata; small fixes

@@
 
-\title{A Tutorial on Tailored Simulation-Based Sample Size Planning for
+\title{A Tutorial on Tailored Simulation-Based Sample-Size Planning for
 Experimental Designs with Generalized Linear Mixed Models}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, left=2mm, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, colframe=quarto-callout-important-color-frame, breakable, rightrule=.15mm, coltitle=black, titlerule=0mm, arc=.35mm, toprule=.15mm, bottomtitle=1mm, colbacktitle=quarto-callout-important-color!10!white, opacitybacktitle=0.6, bottomrule=.15mm, toptitle=1mm, opacityback=0, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, bottomtitle=1mm, arc=.35mm, toptitle=1mm, left=2mm, opacityback=0, colbacktitle=quarto-callout-important-color!10!white, colback=white, opacitybacktitle=0.6, rightrule=.15mm, breakable, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Important}, colframe=quarto-callout-important-color-frame, titlerule=0mm, bottomrule=.15mm, coltitle=black, leftrule=.75mm]
 
@@
 Providing an informed standardized effect size can be an almost
-impossible challenge when no prior studies of pilot data are available.
+impossible challenge when no prior studies or pilot data are available.
 This problem is further exacerbated by the fact that GLMMs are so
